### 2023-11-24 co687/project/00-introduction.tex
Converted project into latex

\section{Introduction}

Trapdoor functions are the foundations of public-key cryptography. The most famous and popular trapdoor functions include the RSA trapdoor (it is easy to compute $m^e$ from $m$ and $e$ but difficult to recover $m$ from $m^e$ and $e$) and discrete log (it is easy to compute $g^x$ from $g, x$ but hard to recover $x$ from $g^x, g$). While these two trapdoors have been proved very successful over their decades of application, they are unfortunately both based on the hardness of integer factorization. Trapdoor functions based on other classes of hard problems have been of research interest since the early days of public-key cryptography, and they have gained increased attention as a result of the discovery of efficient quantum algorithm for integer factorization that threatens to break encryption and signature schemes based on RSA and Diffie-Hellman.

In this project is a survey a particular class of cryptographic trapdoor functions whose construction is based on lattices and whose security is based on the hardness of lattice problems such as the shortest vector problem (SVP) and the closest vector problem (CVP). Some preliminary mathematics surrounding lattice are introduced in chapter 2, followed by some additional discussion of hard lattice problems and known best algorithms to solve them in chapter 3. A lattice-based trapdoor, initially proposed by Goldreich-Goldwasser-Halevi in 1997, is described and its applications discussed in chapter 4. In chapter 5 we will discuss an improvement due to Micciancio that uses Hermite normal form to reduce key sizes while achieving optimal security.

### 2023-11-24 co687/project/01-preliminaries.tex
Converted project into latex

\section{Preliminaries}
Throughout this survey, we denote vectors by bolded letters $\mathbf{v} \in \mathbb{Z}^n$. While lattices are defined as discrete subgroups over the real, in practice it is easier to exclusively work with integer lattices $\mathcal{L} \subseteq \mathbb{Z}^n$. We also exclusively work with full-rank lattices (whose basis $B \in \mathbb{Z}^{n \times n}$ is non-singular full-rank matrix).

There are a few preliminaries in the mathematics of lattices that are tremendously helpful with understanding the difficulty of hard lattice problems and with the trapdoor constructions based on these hard problems. We will discuss them in this chapter.

\subsection{Unimodular relationship}
First, we describe how two different basis of the same lattice relate to each other. This relationship plays a crucial role in the construction of the GGH trapdoor because the trapdoor scheme first generates a secret basis, then uses this relationship to generate a public basis of the same lattice.

> Let $\mathcal{L}(B)$ the lattice generated by the column vectors of matrix $B \in \mathbb{Z}^{n \times n}$, then $\mathcal{L}(B_1) = \mathcal{L}(B_2)$ if and only if there exists a unimodular matrix $U \in \mathbb{Z}^{n \times n}$ such that $B_1 = B_2U$

Here a unimodular matrix is a square matrix whose entries are integers and whose determinant is $\pm 1$. In the forward direction, if $B_1, B_2$ generate the same lattice, then each column of $B_1, B_2$ is a lattice point and can be represented using the other basis, which means that there exists integer matrix $U, V$ such that $B_1 = B_2U$ and $B_2 = B_1V$. With simple substitution, we find that $B_1 = B_2U = (B_1V)U$, so $B_1(I - VU) = 1$. Because $B_1$ is non-singular, it necessarily follows that $VU = 1$. By a symmetric argument, we also have $UV = 1$, which means that $V = U^{-1}$. we know $U, V$ to both be integer matrix, so their determinants both necessarily have to be $1$, thus proving that $U, V$ to both be unimodular matrix.

In the backward direction, if $B_1 = B_2U$ for some unimodular matrix $U$, then columns of $B_1$ are already points in the lattice generated by $B_2$, meaning that $\mathcal{L}(B_1) \subseteq \mathcal{L}(B_2)$. Because $U$ is unimodular, we also hve $B_2 = B_1U^{-1}$, where $U^{-1}$ is an integer matrix (in fact $U^{-1}$ is also unimodular), so by a symmetric argument we also have $\mathcal{L}(B_2) \subseteq \mathcal{L}(B_1)$. Therefore, $\mathcal{L}(B_1) = \mathcal{L}(B_2)$. $\blacksquare$

Knowing that two basis generate the same lattice if and only if they are related by a unimodular matrix, it is not hard to deduce that swapping columns, negating a column, and adding to a column an integer linear combination of all other columns all preserve the lattice generated by a matrix because each of the such action can be represented as a right multiplication by a unimodular matrix.

\subsection{Orthogonality and determinant}
The determinant of a lattice $\mathcal{L}(B)$ generated by a basis $B$ is the determinant of the matrix $B$. Knowing that two basis generate the same lattice if and only if they are related by a unimodular matrix (whose determinant is 1), it is not hard to see that the determinant is the same regardless of choice of basis. We can thus define the "determinant of a lattice" $\det(\mathcal{L})$ as an invariant of the lattice regardless of the choice of basis.

The determinant of a lattice can be geometrically interpreted as the "n-dimensional volume" of the fundamental parallelpiped, which is defined as a subset of $\mathbb{R}^n$:

$$
P(B) = B\cdot [0, 1)^n = \{\sum_{i=1}^n x_i\mathbf{b_i} \mid 0\leq x < 1 \}
$$

Note that it is often more convenient to "center" the fundamental parallelpiped around $\mathbf{0}$ and define it as $P(B) = B\cdot [-\frac{1}{2}, \frac{1}{2})^n$.

On the other hand, we are also concerned with "how orthogonal" the basis is. For that we define the orthogonality defect:

$$
\mathop{\text{ortho-defect}} = \frac{\prod_{i=1}^n \Vert \mathbf{b}_i \Vert}{\vert\det(B)\vert}
$$

Observe that if $B$ is also an orthogonal matrix whose columns are pair-wise orthogonal, then $\text{ortho-defect}(B) = 1$. Otherwise, the orthogonality defect of $B$ will be greater than 1. Geometrically, a greater orthogonality defect corresponds to a basis that is more "skewed".

Intuitively it is not hard to imagine that, for a fundamental parallelpiped with a fixed n-dimensional volume, the more skewed the basis is, the longer each basis will have to be to preserve the n-dimensional volume, and the further away the basis is from the origin. This is the principal observation on the hardness of the shortset vector problems and their siblings (such as the closest vector problem used in GGH's construction): **the longer, more skewed a basis is, the harder it is to find a short vector in the lattice**. It turns out that the gap in difficult of finding short vector between good (short, orthogonal) basis and bad (long, skewed) basis is exponential, and thus hard lattice problems pose provide an excellent foundation to build trapdoor functions on.

### 2023-11-24 co687/project/02-basis-reduction.tex
Converted project into latex

\section{Lattice basis reduction}
In this section we discuss some known best bounds on the shortest vector (of a lattice), as well as the known best algorithm for approximating the shortest vector.

By convention, we define the minimum distance of a lattice to be the minimum of all distances between two distinct points on a lattice. Because lattice is a discrete sub-group of $\mathbb{Z}^n$, this minimum distance is always achieved by some pair of points. Also because a lattice is closed under addition (and thus subtraction), the minimum distance is exactly the length of the shortest vector on the lattice. Denote the mininum distance by $\lambda(\mathcal{L})$. For convenience, where there is no confusion of which lattice we are working with, we simply denote the minimum distance by $\lambda$

\subsection{A lower bound on the shortest vector}
We can first provide a lower bound on the length of the shortest vector of a lattice, given a basis $B$. First, recall the Gram-Schmidt orthognoalization process for converting a basis $B = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n]$ into an orthogonal basis $B^\ast = [\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_n^\ast]$ via the following algorithm:

$$
\mathbf{b}_i^\ast \leftarrow \mathbf{b}_i - \sum_{j<i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^*\rangle}{\langle \mathbf{b}_j^*, \mathbf{b}_j^*\rangle}\mathbf{b}_j^\ast
$$

For convenience, we denote the orthogonalization coefficient by:

$$
\mu_{i, j} = \frac{\langle \mathbf{b}_i, \mathbf{b}_j^*\rangle}{\langle \mathbf{b}_j^*, \mathbf{b}_j^*\rangle}
$$

Also, we define a function $\pi_i$ that projects a vector from $\mathbb{R}^n$ onto the orthogonal basis formed by $\mathbf{b}_i^\ast, \mathbf{b}_{i+1}^\ast, \ldots, \mathbf{b}_n^\ast$:

$$
\pi_i(\mathbf{x}) = \sum_{j\geq i}\frac{\langle \mathbf{x}, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast
$$

Observe that for $i = 1$, $\pi_i$ is the identity function, because the orthogonalized basis still span the same vector space as the original basis. Also, $\pi_i(\mathbf{b}_i) = \mathbf{b}_i^\ast$:

$$
\begin{aligned}
\pi_i(\mathbf{b}_i) &= \sum_{j\geq i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
&= \sum_{1 \leq j \leq n}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast  - \sum_{j < i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
&= \pi_1(\mathbf{b}_i) - \sum_{j < i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
&= \mathbf{b}_i - \sum_{j < i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
&= \mathbf{b}_i^\ast
\end{aligned}
$$

With these notations established, we can bound $\lambda(\mathcal{L(B)})$ from below using the orthogonalized basis. We will do that by proving the following recursive inequality:

$$
\lambda(\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n)) \geq \min (\Vert \mathbf{b}_n^\ast \Vert, \lambda(\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}))
$$

To prove this inequality, we represent an arbitrary non-zero point on the lattice by $B\mathbf{x} = \sum_{i=1}^n\mathbf{b}_i x_i$:

$$
\begin{aligned}
\Vert B\mathbf{x} \Vert^2 &= \Vert \sum_{i=1}^n\mathbf{b}_i x_i \Vert^2 \\
&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + \mathbf{b}_nx_n \Vert^2
\end{aligned}
$$

From here substitute the orthogonalization of $\mathbf{b}_n^\ast = \mathbf{b}_n - \sum_{j<n}\mu_{j, n} \mathbf{b}_j^\ast$:

$$
\begin{aligned}
\Vert B\mathbf{x} \Vert^2 &= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + \mathbf{b}_nx_n \Vert^2 \\
&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\mathbf{b}_n^* + \sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast)x_n \Vert^2 \\
&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast) x_n  + \mathbf{b}_n^\ast x_n\Vert^2
\end{aligned}
$$

Observe that by the definition of the orthogonalization process, $\mathbf{b}_n^\ast$ is orthogonal to all other orthogonal basis $\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_{n-1}^\ast$, as well as all except the identically indexed original basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}$. By Pythagoras theorem we can separate the RHS into orthogonal components:

$$
\begin{aligned}
\Vert B\mathbf{x} \Vert^2 
&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast) x_n  + \mathbf{b}_n^\ast x_n\Vert^2 \\
&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast) x_n \Vert^2 + \Vert \mathbf{b}_n^\ast x_n \Vert^2 \\
&\geq \Vert \mathbf{b}_n^\ast x_n \Vert^2
\end{aligned}
$$

where $x_n > 0$, the inequality trivially implies $\Vert B\mathbf{x} \Vert \geq \Vert\mathbf{b}_n^\ast\Vert$. where $x_n = 0$, $\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n) = \mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1})$, so by definition of minimum distance we have $\Vert B\mathbf{x} \Vert \geq \lambda(\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}))$. Combining the two cases gives us the overall inequality. $\blacksquare$

Having proved the recursive inequality, we can simply apply recursion $n$ times and arrive at a concrete lower bound for $\lambda$.

$$
\lambda \geq \min_{1 \leq i \leq n} \Vert \mathbf{b}_i^\ast \Vert
$$

\subsection{LLL lattice basis reduction algorithm}
At the time of this survey, the best algorithm for solving the (approximate) shortest vector problem is the LLL lattice basis reduction algorithm, attributed to Arjen Lenstra, Hendrik Lenstra, and László Lovász. The reduction algorithm transforms an input basis into an LLL-reduced form, where the first base vector of the reduced basis is an approximation of the shortest vector.

**If time permits, we will discuss the algorithm itself. However, I found it okay to first treat the algorithm as a black box, and instead focus on the properties of the reduced basis.**

The definition of a reduced basis is parameterized by a real number $\frac{1}{4} < \delta \leq 1$. A basis $B$ is $\delta$-LLL reduced if the following conditions are satisfied:

- For all $j > i$, $\vert\mu_{j, i}\vert \leq \frac{1}{2}$
- For all $1 \leq i \leq n-1$, $\delta\Vert \pi_i(\mathbf{b}_i)\Vert^2 \leq \Vert\pi_i(\mathbf{b}_{i+1})\Vert^2$

The second clause of the reduced form is of particular interest because we can use it to measure how well $\mathbf{b}_1$ approximates the shortest vector. First observe the RHS of the condition:

$$
\begin{aligned}
\Vert \pi_i(\mathbf{b}_{i+1})\Vert^2 &= \Vert \sum_{j\geq i}\mu_{i+1, j}\mathbf{b}_j^\ast\Vert^2 \\
&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast + \sum_{j\geq i+1}\mu_{i+1, j}\mathbf{b}_j^\ast\Vert^2 \\
&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast + \pi_{i+1}(\mathbf{b}_{i+1}) \Vert^2 \\
&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast + \mathbf{b}_{i+1}^\ast \Vert^2 \\
&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast \Vert^2 + \Vert \mathbf{b}_{i+1}^\ast \Vert^2 \\
\end{aligned}
$$

Substituting the equation back into the condition:

$$
\delta\Vert \pi_i(\mathbf{b}_i)\Vert^2 \leq \Vert \mu_{i+1, i}\mathbf{b}_i^\ast \Vert^2 + \Vert \mathbf{b}_{i+1}^\ast \Vert^2
$$

Which transforms into:

$$
\Vert \mathbf{b}_i^\ast \Vert^2 \leq \frac{1}{(\delta - \mu_{i+1, i}^2)} \Vert \mathbf{b}_{i+1}^\ast \Vert^2
$$

By the first clause of the reduced basis we know $\mu_{i+1, i}^2 \leq \frac{1}{4}$, which means that $\frac{1}{\delta - \mu_{i+1, i}^2} \leq \frac{1}{\delta - \frac{1}{4}}$. Denote $\alpha = \frac{1}{\delta - \frac{1}{4}}$, then $\alpha > \frac{4}{3}$, and we have the following inequality:

$$
\Vert \mathbf{b}_i^\ast \Vert^2 \leq \alpha \Vert \mathbf{b}_{i+1}^\ast \Vert^2
$$

Note that because $\mathbf{b}_1 = \mathbf{b}_1^\ast$, we can recusively evaluate the inequality above can obtain the following closed inequality:

$$
\Vert \mathbf{b}_1 \Vert^2 \leq \alpha^{i-1} \Vert \mathbf{b}_{i}^\ast \Vert^2 \leq \alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert^2
$$

The inequality above states that $\mathbf{b}_1$ is not longer than $\alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert$ for all $i$, so it must be not longer than the minimum among $\alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert$:

$$
\Vert \mathbf{b}_1 \Vert^2 \leq \min_{1 \leq i \leq n}\alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert^2 = \alpha^{n-1} \min_{1 \leq i \leq n}\Vert \mathbf{b}_{i}^\ast \Vert^2
$$

From the previou section we've derived that $\lambda \geq \min_{1 \leq i \leq n} \Vert \mathbf{b}_i^\ast \Vert$, which we can plug into the inequality above:

$$
\Vert \mathbf{b}_1 \Vert \leq \alpha^\frac{n-1}{2} \lambda
$$

This in equality bounds the length of the first base vector of the reduced basis by some multiples of the minimum distance, where the multiple is exponential with respect to the number of dimensions of the lattice.

In other words, the LLL reduction algorithm computes an approximation of the shortest vector within a factor of $\gamma \in O(\alpha^n)$. With specific choices of "bad" basis and sufficiently high dimension $n$, the basis reduction algorithm will not be able to provide meaningfully tight approximation of the shortest vector, making the shortest vector problem suitably hard for cryptographic applications.

### 2023-11-24 co687/project/03-nearest-plane-cvp.tex
Converted project into latex

\section{Nearest plane algorithm}
In this section we state the closest vector problem (CVP), which is a hard lattice problem that the GGH construction is based on. We will then discuss Babai's nearest plane algorithm, which can be used to approximate solutions to the CVP.

Let $\mathcal{L}(B) \subseteq \mathbb{Z}^n$ be a full-rank lattice, and let $\mathbf{t} \in \mathbb{Z}^n$ be a target vector. The closest vector problem asks to find $\mathbf{v} \in \mathcal{L}(B)$ that minimizes $\Vert \mathbf{t} - \mathbf{v} \Vert$, where the norm is the Euclidean norm (over real numbers).

\subsection{Fundamental regions}
Before attempting the closest vector problem, we first introduce a highly relevant concept called the **fundamental region** of a lattice. A fundamental region $S$ is a subset of the (real) linear span of the basis such that it tiles the linear span of the basis and each tile contains exactly one lattice point. If we have a fundamental region $S$ defined, then each point in the linear span of the basis can be uniquely decomposed into the sum of a lattice point and a point in the fundamental region:

$$
\mathcal{L}(B) = \{S + B\mathbf{x} \mid \mathbf{x} \in \mathbb{Z}^n\}
$$

A few notable examples include the fundamental parallelpiped

$$
\mathcal{P}(B) = \{B\mathbf{x} \mid \mathbf{x} \in [0, 1)^n\}
$$

and the centered fundamental parallelpiped

$$
\mathcal{C}(B) = \{B\mathbf{x} \mid \mathbf{x} \in [-\frac{1}{2}, \frac{1}{2})^n\}
$$

Computing the decomposition using the (centered) fundmental parallelpiped is easy. First compute $B^{-1}\mathbf{t}$ (over $\mathbb{R}$), then perform some kind of rounding (flooring or nearest integer) depending on whether the fundamental parallelpiped is centered or or not. $B\lfloor B^{-1}\mathbf{T} \rceil$ is the lattice point whose corresponding fundamental region contains the target. This is Babai's rounding algorithm. As we will see in later section, this algorithm works well if $B$ is short and orthogonal, but falls apart badly if $B$ is long and skewed.

The Voronoi region is defined by the set of points in the linear span that are closer to $\mathbf{0}$ than to any other lattice points. There is some additional details that need to be specified to ensure that the Voronoi region properly closes and forms a partition of $\text{span}(B)$, which we will not discuss in details here. If we have an algorithm that can efficiently decompose $\mathbf{t} \in \text{span}(B)$ using the Voronoi region, then we automatically have a way to solve the closest vector problem. Unfortunately, no such algorithm is known.

Of particular interest is the (centered) orthogonalized fundamental parallelpiped, which is defined using the orthogonalized basis $B^\ast = \text{Gram-Schmidt}(B)$:

$$
\mathcal{C}(B^\ast) = \{B^\ast\mathbf{x} \mid \mathbf{x} \in [-\frac{1}{2}, \frac{1}{2})\}
$$

(If time permits we should try to prove that $\mathcal{C}(B^\ast)$ is indeed a fundamental region. For now, we will take that for granted without proof).

Because $B^\ast$ is an orthogonal basis, it's easy to see that the sphere centered at some lattice point $\mathbf{v} \in \mathcal{L}$ and whose radius is $\frac{1}{2}\min \Vert \mathbf{b}_i^\ast\Vert$ is entirely contained in the (shifted) fundamental region $\text{sphere} \subset \{\mathcal{C}(B^\ast) + \mathbf{v}\}$.

\subsection{Babai's nearest plane algorithm}
Babai's nearest plane algorithm, attributed to László Babai, is a recursive algorithm that can approximate the closest vector under a given basis to a target vector. More specifically, it returns a vector point $\mathbf{v}$ such that, if target vector is projected onto the orthogonalized basis, the projection is contained in $\mathbf{v} + \mathcal{C}(B^\ast)$. If the target vector is in the linear span of the basis, then the target vector itself is contained in $\mathbf{v} + \mathcal{C}(B^\ast)$.

The algorithm $\text{NearestPlane}(B, \mathbf{t})$ takes a set of basis $B = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n}]$ and a target vector $\mathbf{t}$, then performs the following procedure:

1. If $B$ is an empty set of basis, return $\mathbf{0}$
2. $B^\ast \leftarrow \text{GramSchmidt}(B)$
3. $c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil$, where $\mathbf{b}_n^\ast$ is the last base vector in $B^\ast$
4. $B' = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}]$ is obtained by removing the last base vector from $B$
5. return $c\mathbf{b}_n + \text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$

Denote the output of this algorithm by $\mathbf{v}$, the key theorem about this algorithm is as follows:

$$
\forall 1 \leq i \leq n, 
\frac{
    \langle
        \mathbf{t} - \mathbf{v}, \mathbf{b}_i^\ast
    \rangle
}{
    \langle
        \mathbf{b}_i^\ast, \mathbf{b}_i^\ast
    \rangle
} 
\in [-\frac{1}{2}, \frac{1}{2})
$$

Intuitively, the inequality above states that the deviation of $\mathbf{t}$ from $\mathbf{v}$ is between $-\frac{1}{2}\mathbf{b}_i^\ast$ and $\frac{1}{2}\mathbf{b}_i^\ast$ in any of the chosen direction $\mathbf{b}_i^\ast$, which means that $\mathbf{t} - \mathbf{v}$ is indeed contained in the orthogonalized fundamental parallelpiped.

We can prove this result inductively. In the base case, if the input set of basis is empty, then this result is trivially correct. In the inductive case, we denote the output of $\text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$ by $\mathbf{v}^\prime$. Assuming that the algorithm produces the desired result for the sublattice $\mathcal{L}(B^\prime)$:

$$
\forall 1 \leq i \leq (n-1), 
\frac{
    \langle
        (\mathbf{t} - c\mathbf{b}_n) - \mathbf{v}^\prime, 
        \mathbf{b}_i^\ast
    \rangle
}{
    \langle
        \mathbf{b}_i^\ast, \mathbf{b}_i^\ast
    \rangle
} \in [-\frac{1}{2}, \frac{1}{2})
$$

However, notice in step 5 of the algorithm $c\mathbf{b}_n + \mathbf{v}^\prime$ is the output of the current iteration of the algorithm, so we have

$$
\begin{aligned}
(\mathbf{t} - c\mathbf{b}_n) - \mathbf{v}^\prime 
&= \mathbf{t} - (c\mathbf{b}_n + \mathbf{v}^\prime) \\
&= \mathbf{t} - \mathbf{v}
\end{aligned}
$$

which means that

$$
\forall 1 \leq i \leq (n-1), 
\frac{
    \langle
        \mathbf{t} - \mathbf{v}, 
        \mathbf{b}_i^\ast
    \rangle
}{
    \langle
        \mathbf{b}_i^\ast, \mathbf{b}_i^\ast
    \rangle
} \in [-\frac{1}{2}, \frac{1}{2})
$$

So all that remains is to show that this relationship also holds for $i = n$.

To prove that the relationship holds for $i = n$, first observe how $c$ is computed:

$$
c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil
$$

which is equivalent to saying that

$$
\frac{
    \langle\mathbf{t}, \mathbf{b}_n^\ast\rangle
}{
    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
}
- c \in [-\frac{1}{2}, \frac{1}{2})
$$

Also recall the Gram-Schmid orthogonalization process:

$$
\mathbf{b}_n^\ast +\sum_{i<n}\mu_{n, i}\mathbf{b}_i^\ast = \mathbf{b}_n
$$

Therefore

$$
\begin{aligned}
\langle\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
&= \langle 
    \mathbf{b}_n^\ast +\sum_{i<n}\mu_{n, i}\mathbf{b}_i^\ast, \mathbf{b}_n^\ast
\rangle \\
&= \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
\end{aligned}
$$

All non $\mathbf{b}_n^\ast$ terms can be cleared because they are orthogonal to $\mathbf{b}_n^\ast$, so their inner product is 0. This equality means that:

$$
\frac{
    \langle\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
}{
    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
} = 1
$$

Finally we can put everything together:

$$
\begin{aligned}
\frac{
    \langle\mathbf{t}, \mathbf{b}_n^\ast\rangle
}{
    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
}
- c 
&= \frac{
    \langle\mathbf{t}, \mathbf{b}_n^\ast\rangle
}{
    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
}
- c \frac{
    \langle\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
}{
    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
} \\
&= \frac{
    \langle\mathbf{t} - c\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
}{
    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
}
\in [-\frac{1}{2}, \frac{1}{2})
\end{aligned}
$$

Hence we've proved the relationship for $i = n$. $\blacksquare$

### 2023-11-24 co687/project/04-ggh-trapdoor.tex
Converted project into latex

\section{The Goldreich-Goldwasser-Halevi trapdoor}
One class of constructing lattice trapdoor uses a pair of public and secret basis for the same lattice. Since the two basis generate the same lattice, they are equally good at mapping integer coordinates into a lattice point. On the other hand, the secret basis is very "good" and can be used to efficiently find the closest lattice point, but the public basis is very "bad" at recovering the closest lattice point.

One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper *"Public-key cryptosystem from lattice reduction problem"*. At a high level, the trapdoor is parameterized by three items: a "bad" basis $B$, a "good" basis $R$, and an error bound $\sigma$. In the forward direction, the function maps a pair of lattice coordinate $\mathbf{v} \in \mathbb{Z}^n$ and a small error vector $\mathbf{e} \leftarrow \{ -\sigma, \sigma \}^n$ to $\mathbf{x} = B\mathbf{v} + \mathbf{e} \in \mathbb{R}^n$. If the parameters are generated correctly, then the closest lattice point in $\mathcal{L}(B)$ is exactly $B\mathbf{v}$.

Inverting the function involves recovering the integer coordinate $\mathbf{v}$ (or the error vector $\mathbf{e}$, since recovering one of them automatically gives you the other). However, the inversion is exactly the cloest vector problem (CVP), and should be hard if $B$ is a sufficiently bad basis. On the other hand, since $R$ is a good basis, finding the closest vector point should be "easy" if we have $R$.

From here, GGH '97 proposed a public-key cryptosystem as well as a digital signature scheme that uses such a trapdoor construction, and the security of the two schemes naturally rest on the hardness of the underlying hard lattice problem.

\subsection{The trapdoor scheme}
The GGH trapdoor scheme contains four components:

1. Parameter generation  
The main security parameter in this scheme is the number of dimensions $n$ of the lattice. In the original paper, the authors claimed $n \approx 150$ is sufficient for making inverting the function without the private basis hard and $n \approx 250$ should be a safe choice for the foreseeable future.
2. Key generation  
First generate the "good" basis $R \in \mathbb{R}^{n \times }$, then apply some (unimodular) transformation to $R$ to obtain the "bad" basis $B$. The error bound $\sigma > 0 \in \mathbb{R}$ is dependent on the choice of $R$ and the choice of "probability of inversion error" $\epsilon >= 0$, which will be discussed in a later section.
3. Forward evaluation  
$f_{B}(\mathbf{v}, \mathbf{e}) = B\mathbf{v} + \mathbf{e}$, where $\mathbf{v} \in \{-n, \ldots, n\}^n$ and $\mathbf{e} \in \{-\sigma, \sigma\}$. According to the authors, the choice of bounds for values of $\mathbf{v}$ is arbitrary and not a significant contributor to the overall security of the scheme.
4. Inversion  
Denote the output by $\mathbf{x} = f_{B}(\mathbf{v}, \mathbf{e})$, first attempt to recover the integer coordinate $\mathbf{v} \leftarrow B^{-1}R\lfloor R^{-1}\mathbf{x} \rceil$. From here it is trivial to recompute the lattice point $B\mathbf{v}$ and recover the error term $\mathbf{e} = \mathbf{x} - B\mathbf{v}$.

\subsection{Correctness of trapdoor inversion}
Without the error term, the function $f_B: \mathbf{v} \mapsto B\mathbf{v}$ is trivially invertible with either choice of the basis. However, with a non-zero the error term, the quality of the basis makes a substantial difference in how much error can be added before the points can no longer be recovered.

Observe the calculation used for recovering the integer coordinate $\mathbf{v}$:

$$
\begin{aligned}
\lfloor R^{-1}\mathbf{x} \rceil &= B^{-1}R\lfloor R^{-1}(B\mathbf{v} + \mathbf{e})\rceil \\
&= B^{-1}R\lfloor R^{-1}B\mathbf{v} + R^{-1}\mathbf{e}\rceil
\end{aligned}
$$

Since $R, B$ are related by a unimodular matrix and $\mathbf{v}$ is an integer vector, $R^{-1}B\mathbf{v}$ is an integer vector and can be moved out of the "rounding" operator:

$$
\begin{aligned}
\lfloor R^{-1}\mathbf{x} \rceil &= B^{-1}RR^{-1}B\mathbf{v} + B^{-1}R\lfloor R^{-1}\mathbf{e}\rceil \\
&= \mathbf{v} + B^{-1}R\lfloor R^{-1}\mathbf{e} \rceil
\end{aligned}
$$

Since $B^{-1}R$ is also a unimodular matrix, we can conclude that the equation above is successful at recovering the original coordinate if and only if $R^{-1}\mathbf{e} = \mathbf{0}$.

To guarantee that inversion error never happens, we can bound the error term $\sigma > 0$ by $\frac{1}{2\rho}$, where $\rho$ is maximal $L_1$ norm among the rows of $R^{-1}$. This bound is excessively conservative, however, and we might want to relax the bound to enhance the security of the trapdoor scheme (larger error terms makes it harder to invert the function using only the public basis). The authors provided one such relaxation based on the Hoeffding inequality. This relaxation is stated as follows

$$
P(\text{inversion error}) \leq 2n \cdot \exp(-\frac{1}{8\sigma^2\gamma^2})
$$

Where $\gamma = \sqrt{n} \cdot \max(L_\infty \text{ norm of rows of } R^{-1})$. Simple reorganization of inequality shows that to bound the inversion error by $\epsilon$, the error term will be bounded by $\sigma \leq (\gamma\sqrt{8\ln{2n/\epsilon}})^{-1}$.

\subsection{Generating the pair of basis}
The authors described two similar ways of generating the private basis. The first way is to sample each coordinate of $R \in \mathbb{R}^{n \times n}$ from a uniform distribution on $\{-l, -l + 1, \ldots, l-1, l\}$, where according to the authors, the value of the bound $l$ has negligible impact on the quality of the generated basis (the authors chose $\pm 4$ in their implementation). A second method is to first generate a square lattice $L(kI)$ for some positive integer $k$, then add a small amount of noise $R^\prime \in \{-l, \ldots, l\}^{n \times n}$. With this method of sampling the private basis, it is important to balance the choice of values between $k$ and $l$, where a larger $k$ value gives a more orthogonal basis, but also weakens the security of the trapdoor function by making it easier to reduce the public basis into a short, orthogonal basis using basis reduction algorithm.

The authors also described two methods for generating the public basis $B$ from the private basis. The first method is to directly generate random unimodular matrix $T$ and set $B = TR$, then repeat until satisfactory. A second method is to repeatedly apply column mixing, where at each mixing a column of $R$ is chosen, and a linear combination of all other columns is added to the chosen column.

Mathematically, the two methods are equivalent. However, in implemenetation, we would like the values of $B$ to be smaller for space efficiency while maintaining sufficient security so that the function cannot be easily inverted using the public basis alone and so that $B$ cannot be easily reduced using basis reduction algorithm. The authors preferred column mixing for requiring less computation and for producing public basis $B$ with smaller values.

Unfortunately, there is no known rigorous description of how skewed $B$ needs to be for the trapdoor function to be secure. The authors relied on experimental methods and determined that for $n \approx 100$, $2n$ steps of column mixing is enough to render LLL basis reduction ineffective at meaningfully improving the quality of the public basis $B$.

\subsection{Inverting the function without the trapdoor}
From a high level, inverting the function corresponds to finding the an approximate closest lattice point, so the best attacks on the trapdoor are also algorithms for solving the CVP problem.

The most straightforward attack on the trapdoor is to simply run the inversion algorithm with the public basis instead of the private basis:

$$
B^{-1}\mathbf{x} = B^{-1}(B\mathbf{v} + \mathbf{e}) = \mathbf{v} + B^{-1}\mathbf{e}
$$

Due to $B$ being very skewed, this procedure does not immediately yield the correct value for $\mathbf{v}$ since $B^{-1}\mathbf{e}$  is not $\mathbf{0}$. However, since the possible values of $\mathbf{e}$ is finite, we can perform an exhaustive search on all possible values of $\mathbf{d} = B^{-1}\mathbf{e}$, although the search space will grow exponentially with the number of dimensions $n$. In experimental settings, with $n \geq 100$ the search space reaches 168 bits of entropy.

Using a better approximation algorithm for CVP, such as the nearest plane algorithm, yields a more efficient inversion than the brute-force search described above, although the complexity of the algorithm still scales exponentially with $n$. In experimental settings, with $n=150$ the workload of the nearest plane algorithm reaches 104 bits.

\subsection{Applications}
The GGH paper presented two possible applications of a GGH-like trapdoor function: a public-key cryptosystem and a digital signature.

What is the cryptosystem???

What is the digital signature???

Unfortunately, in 1999 the public-key cryptosystem was cryptanalyzed by Nguyen, and in 2006 the digital signature scheme was also broken by Nguyen and Regev. Since then, newer lattice trapdoors based on gadget matrix have been proposed, so we will not cover details from the cryptanalysis in this survey.

### 2023-11-24 co687/project/main.tex
Converted project into latex

\documentclass[letterpaper,12pt]{article}

\usepackage{amsmath,amssymb,amstext}
\setlength{\marginparwidth}{0pt} % width of margin notes
% N.B. If margin notes are used, you must adjust \textwidth, \marginparwidth
% and \marginparsep so that the space left between the margin notes and page
% edge is less than 15 mm (0.6 in.)
\setlength{\marginparsep}{0pt} % width of space between body text and margin notes
\setlength{\evensidemargin}{0.125in} % Adds 1/8 in. to binding side of all 
% even-numbered pages when the "twoside" printing option is selected
\setlength{\oddsidemargin}{0.125in} % Adds 1/8 in. to the left of all pages when "oneside" printing is selected, and to the left of all odd-numbered pages when "twoside" printing is selected
\setlength{\textwidth}{6.375in} % assuming US letter paper (8.5 in. x 11 in.) and side margins as above

\begin{document}

\input{00-introduction.tex}
\input{01-preliminaries.tex}
\input{02-basis-reduction.tex}
\input{03-nearest-plane-cvp.tex}
\input{04-ggh-trapdoor.tex}

\end{document}

### 2023-11-24 co687/project/03-nearest-plane-cvp.tex
Some new formatting

@@
 
-1. If $B$ is an empty set of basis, return $\mathbf{0}$
-2. $B^\ast \leftarrow \text{GramSchmidt}(B)$
-3. $c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil$, where $\mathbf{b}_n^\ast$ is the last base vector in $B^\ast$
-4. $B' = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}]$ is obtained by removing the last base vector from $B$
-5. return $c\mathbf{b}_n + \text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$
+\begin{enumerate}
+    \item If $B$ is an empty set of basis, return $\mathbf{0}$
+    \item $B^\ast \leftarrow \text{GramSchmidt}(B)$
+    \item $c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil$, where $\mathbf{b}_n^\ast$ is the last base vector in $B^\ast$
+    \item $B' = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}]$ is obtained by removing the last base vector from $B$
+    \item return $c\mathbf{b}_n + \text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$
+\end{enumerate}
 

### 2023-11-24 co687/project/04-ggh-trapdoor.tex
Some new formatting

@@
 
-One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper *"Public-key cryptosystem from lattice reduction problem"*.
+One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper \textit{"Public-key cryptosystem from lattice reduction problem"}.
 At a high level, the trapdoor is parameterized by three items: a "bad" basis $B$, a "good" basis $R$, and an error bound $\sigma$.
@@
 
-1. Parameter generation  
-The main security parameter in this scheme is the number of dimensions $n$ of the lattice.
+\begin{enumerate}
+    \item \textbf{Parameter generation} The main security parameter in this scheme is the number of dimensions $n$ of the lattice.
 In the original paper, the authors claimed $n \approx 150$ is sufficient for making inverting the function without the private basis hard and $n \approx 250$ should be a safe choice for the foreseeable future.
-2. Key generation  
-First generate the "good" basis $R \in \mathbb{R}^{n \times }$, then apply some (unimodular) transformation to $R$ to obtain the "bad" basis $B$.
+    \item \textbf{Key generation} First generate the "good" basis $R \in \mathbb{R}^{n \times }$, then apply some (unimodular) transformation to $R$ to obtain the "bad" basis $B$.
 The error bound $\sigma > 0 \in \mathbb{R}$ is dependent on the choice of $R$ and the choice of "probability of inversion error" $\epsilon >= 0$, which will be discussed in a later section.
-3. Forward evaluation  
-$f_{B}(\mathbf{v}, \mathbf{e}) = B\mathbf{v} + \mathbf{e}$, where $\mathbf{v} \in \{-n, \ldots, n\}^n$ and $\mathbf{e} \in \{-\sigma, \sigma\}$.
+    \item \textbf{Forward evaluation} $f_{B}(\mathbf{v}, \mathbf{e}) = B\mathbf{v} + \mathbf{e}$, where $\mathbf{v} \in \{-n, \ldots, n\}^n$ and $\mathbf{e} \in \{-\sigma, \sigma\}$.
 According to the authors, the choice of bounds for values of $\mathbf{v}$ is arbitrary and not a significant contributor to the overall security of the scheme.
-4. Inversion  
-Denote the output by $\mathbf{x} = f_{B}(\mathbf{v}, \mathbf{e})$, first attempt to recover the integer coordinate $\mathbf{v} \leftarrow B^{-1}R\lfloor R^{-1}\mathbf{x} \rceil$.
+    \item \textbf{Inversion} Denote the output by $\mathbf{x} = f_{B}(\mathbf{v}, \mathbf{e})$, first attempt to recover the integer coordinate $\mathbf{v} \leftarrow B^{-1}R\lfloor R^{-1}\mathbf{x} \rceil$.
 From here it is trivial to recompute the lattice point $B\mathbf{v}$ and recover the error term $\mathbf{e} = \mathbf{x} - B\mathbf{v}$.
+\end{enumerate}
 

### 2023-11-24 co687/project/main.tex
Some new formatting

@@
 \documentclass[letterpaper,12pt]{article}
+\usepackage{amsmath,amssymb,amstext}
+\usepackage[margin=1in,letterpaper]{geometry}
+\usepackage{amsmath,amssymb}
 
-\usepackage{amsmath,amssymb,amstext}
-\setlength{\marginparwidth}{0pt} % width of margin notes
-% N.B. If margin notes are used, you must adjust \textwidth, \marginparwidth
-% and \marginparsep so that the space left between the margin notes and page
-% edge is less than 15 mm (0.6 in.)
-\setlength{\marginparsep}{0pt} % width of space between body text and margin notes
-\setlength{\evensidemargin}{0.125in} % Adds 1/8 in. to binding side of all 
-% even-numbered pages when the "twoside" printing option is selected
-\setlength{\oddsidemargin}{0.125in} % Adds 1/8 in. to the left of all pages when "oneside" printing is selected, and to the left of all odd-numbered pages when "twoside" printing is selected
-\setlength{\textwidth}{6.375in} % assuming US letter paper (8.5 in. x 11 in.) and side margins as above
+\title{A introduction to lattice trapdoor}
+\author{Ganyu Xu (g66xu)}
+\date{Fall, 2023}
 
 \begin{document}
+\maketitle
 

### 2023-11-24 co687/project/main.tex
A rough draft

@@
 
-\input{00-introduction.tex}
-\input{01-preliminaries.tex}
-\input{02-basis-reduction.tex}
-\input{03-nearest-plane-cvp.tex}
-\input{04-ggh-trapdoor.tex}
+\section{Introduction}
+
+Trapdoor functions are the foundations of public-key cryptography.
+The most famous and popular trapdoor functions include the RSA trapdoor (it is easy to compute $m^e$ from $m$ and $e$ but difficult to recover $m$ from $m^e$ and $e$) and discrete log (it is easy to compute $g^x$ from $g, x$ but hard to recover $x$ from $g^x, g$).
+While these two trapdoors have been proved very successful over their decades of application, they are unfortunately both based on the hardness of integer factorization.
+Trapdoor functions based on other classes of hard problems have been of research interest since the early days of public-key cryptography, and they have gained increased attention as a result of the discovery of efficient quantum algorithm for integer factorization that threatens to break encryption and signature schemes based on RSA and Diffie-Hellman.
+
+In this project is a survey a particular class of cryptographic trapdoor functions whose construction is based on lattices and whose security is based on the hardness of lattice problems such as the shortest vector problem (SVP) and the closest vector problem (CVP).
+Some preliminary mathematics surrounding lattice are introduced in chapter 2, followed by some additional discussion of hard lattice problems and known best algorithms to solve them in chapter 3.
+A lattice-based trapdoor, initially proposed by Goldreich-Goldwasser-Halevi in 1997, is described and its applications discussed in chapter 4.
+In chapter 5 we will discuss an improvement due to Micciancio that uses Hermite normal form to reduce key sizes while achieving optimal security.
+
+\section{Preliminaries}
+Throughout this survey, we denote vectors by bolded letters $\mathbf{v} \in \mathbb{Z}^n$.
+While lattices are defined as discrete subgroups over the real, in practice it is easier to exclusively work with integer lattices $\mathcal{L} \subseteq \mathbb{Z}^n$.
+We also exclusively work with full-rank lattices (whose basis $B \in \mathbb{Z}^{n \times n}$ is non-singular full-rank matrix).
+
+There are a few preliminaries in the mathematics of lattices that are tremendously helpful with understanding the difficulty of hard lattice problems and with the trapdoor constructions based on these hard problems.
+We will discuss them in this chapter.
+
+\subsection{Unimodular relationship}
+First, we describe how two different basis of the same lattice relate to each other.
+This relationship plays a crucial role in the construction of the GGH trapdoor because the trapdoor scheme first generates a secret basis, then uses this relationship to generate a public basis of the same lattice.
+
+> Let $\mathcal{L}(B)$ the lattice generated by the column vectors of matrix $B \in \mathbb{Z}^{n \times n}$, then $\mathcal{L}(B_1) = \mathcal{L}(B_2)$ if and only if there exists a unimodular matrix $U \in \mathbb{Z}^{n \times n}$ such that $B_1 = B_2U$
+
+Here a unimodular matrix is a square matrix whose entries are integers and whose determinant is $\pm 1$.
+In the forward direction, if $B_1, B_2$ generate the same lattice, then each column of $B_1, B_2$ is a lattice point and can be represented using the other basis, which means that there exists integer matrix $U, V$ such that $B_1 = B_2U$ and $B_2 = B_1V$.
+With simple substitution, we find that $B_1 = B_2U = (B_1V)U$, so $B_1(I - VU) = 1$.
+Because $B_1$ is non-singular, it necessarily follows that $VU = 1$.
+By a symmetric argument, we also have $UV = 1$, which means that $V = U^{-1}$.
+we know $U, V$ to both be integer matrix, so their determinants both necessarily have to be $1$, thus proving that $U, V$ to both be unimodular matrix.
+
+In the backward direction, if $B_1 = B_2U$ for some unimodular matrix $U$, then columns of $B_1$ are already points in the lattice generated by $B_2$, meaning that $\mathcal{L}(B_1) \subseteq \mathcal{L}(B_2)$.
+Because $U$ is unimodular, we also hve $B_2 = B_1U^{-1}$, where $U^{-1}$ is an integer matrix (in fact $U^{-1}$ is also unimodular), so by a symmetric argument we also have $\mathcal{L}(B_2) \subseteq \mathcal{L}(B_1)$.
+Therefore, $\mathcal{L}(B_1) = \mathcal{L}(B_2)$.
+$\blacksquare$
+
+Knowing that two basis generate the same lattice if and only if they are related by a unimodular matrix, it is not hard to deduce that swapping columns, negating a column, and adding to a column an integer linear combination of all other columns all preserve the lattice generated by a matrix because each of the such action can be represented as a right multiplication by a unimodular matrix.
+
+\subsection{Orthogonality and determinant}
+The determinant of a lattice $\mathcal{L}(B)$ generated by a basis $B$ is the determinant of the matrix $B$.
+Knowing that two basis generate the same lattice if and only if they are related by a unimodular matrix (whose determinant is 1), it is not hard to see that the determinant is the same regardless of choice of basis.
+We can thus define the "determinant of a lattice" $\det(\mathcal{L})$ as an invariant of the lattice regardless of the choice of basis.
+
+The determinant of a lattice can be geometrically interpreted as the "n-dimensional volume" of the fundamental parallelpiped, which is defined as a subset of $\mathbb{R}^n$:
+
+$$
+P(B) = B\cdot [0, 1)^n = \{\sum_{i=1}^n x_i\mathbf{b_i} \mid 0\leq x < 1 \}
+$$
+
+Note that it is often more convenient to "center" the fundamental parallelpiped around $\mathbf{0}$ and define it as $P(B) = B\cdot [-\frac{1}{2}, \frac{1}{2})^n$.
+
+On the other hand, we are also concerned with "how orthogonal" the basis is. For that we define the orthogonality defect:
+
+$$
+\mathop{\text{ortho-defect}} = \frac{\prod_{i=1}^n \Vert \mathbf{b}_i \Vert}{\vert\det(B)\vert}
+$$
+
+Observe that if $B$ is also an orthogonal matrix whose columns are pair-wise orthogonal, then $\text{ortho-defect}(B) = 1$.
+Otherwise, the orthogonality defect of $B$ will be greater than 1.
+Geometrically, a greater orthogonality defect corresponds to a basis that is more "skewed".
+
+Intuitively it is not hard to imagine that, for a fundamental parallelpiped with a fixed n-dimensional volume, the more skewed the basis is, the longer each basis will have to be to preserve the n-dimensional volume, and the further away the basis is from the origin.
+This is the principal observation on the hardness of the shortset vector problems and their siblings (such as the closest vector problem used in GGH's construction): \textbf{the longer, more skewed a basis is, the harder it is to find a short vector in the lattice}.
+It turns out that the gap in difficult of finding short vector between good (short, orthogonal) basis and bad (long, skewed) basis is exponential, and thus hard lattice problems pose provide an excellent foundation to build trapdoor functions on.
+
+\section{Lattice basis reduction}
+In this section we discuss some known best bounds on the shortest vector (of a lattice), as well as the known best algorithm for approximating the shortest vector.
+
+By convention, we define the minimum distance of a lattice to be the minimum of all distances between two distinct points on a lattice.
+Because lattice is a discrete sub-group of $\mathbb{Z}^n$, this minimum distance is always achieved by some pair of points.
+Also because a lattice is closed under addition (and thus subtraction), the minimum distance is exactly the length of the shortest vector on the lattice.
+Denote the mininum distance by $\lambda(\mathcal{L})$.
+For convenience, where there is no confusion of which lattice we are working with, we simply denote the minimum distance by $\lambda$
+
+\subsection{A lower bound on the shortest vector}
+We can first provide a lower bound on the length of the shortest vector of a lattice, given a basis $B$.
+First, recall the Gram-Schmidt orthognoalization process for converting a basis $B = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n]$ into an orthogonal basis $B^\ast = [\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_n^\ast]$ via the following algorithm:
+
+$$
+\mathbf{b}_i^\ast \leftarrow \mathbf{b}_i - \sum_{j<i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^*\rangle}{\langle \mathbf{b}_j^*, \mathbf{b}_j^*\rangle}\mathbf{b}_j^\ast
+$$
+
+For convenience, we denote the orthogonalization coefficient by:
+
+$$
+\mu_{i, j} = \frac{\langle \mathbf{b}_i, \mathbf{b}_j^*\rangle}{\langle \mathbf{b}_j^*, \mathbf{b}_j^*\rangle}
+$$
+
+Also, we define a function $\pi_i$ that projects a vector from $\mathbb{R}^n$ onto the orthogonal basis formed by $\mathbf{b}_i^\ast, \mathbf{b}_{i+1}^\ast, \ldots, \mathbf{b}_n^\ast$:
+
+$$
+\pi_i(\mathbf{x}) = \sum_{j\geq i}\frac{\langle \mathbf{x}, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast
+$$
+
+Observe that for $i = 1$, $\pi_i$ is the identity function, because the orthogonalized basis still span the same vector space as the original basis. Also, $\pi_i(\mathbf{b}_i) = \mathbf{b}_i^\ast$:
+
+$$
+\begin{aligned}
+\pi_i(\mathbf{b}_i) &= \sum_{j\geq i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
+&= \sum_{1 \leq j \leq n}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast  - \sum_{j < i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
+&= \pi_1(\mathbf{b}_i) - \sum_{j < i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
+&= \mathbf{b}_i - \sum_{j < i}\frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast\rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast \rangle}\mathbf{b}_j^\ast \\
+&= \mathbf{b}_i^\ast
+\end{aligned}
+$$
+
+With these notations established, we can bound $\lambda(\mathcal{L(B)})$ from below using the orthogonalized basis. We will do that by proving the following recursive inequality:
+
+$$
+\lambda(\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n)) \geq \min (\Vert \mathbf{b}_n^\ast \Vert, \lambda(\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}))
+$$
+
+To prove this inequality, we represent an arbitrary non-zero point on the lattice by $B\mathbf{x} = \sum_{i=1}^n\mathbf{b}_i x_i$:
+
+$$
+\begin{aligned}
+\Vert B\mathbf{x} \Vert^2 &= \Vert \sum_{i=1}^n\mathbf{b}_i x_i \Vert^2 \\
+&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + \mathbf{b}_nx_n \Vert^2
+\end{aligned}
+$$
+
+From here substitute the orthogonalization of $\mathbf{b}_n^\ast = \mathbf{b}_n - \sum_{j<n}\mu_{j, n} \mathbf{b}_j^\ast$:
+
+$$
+\begin{aligned}
+\Vert B\mathbf{x} \Vert^2 &= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + \mathbf{b}_nx_n \Vert^2 \\
+&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\mathbf{b}_n^* + \sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast)x_n \Vert^2 \\
+&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast) x_n  + \mathbf{b}_n^\ast x_n\Vert^2
+\end{aligned}
+$$
+
+Observe that by the definition of the orthogonalization process, $\mathbf{b}_n^\ast$ is orthogonal to all other orthogonal basis $\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_{n-1}^\ast$, as well as all except the identically indexed original basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}$.
+By Pythagoras theorem we can separate the RHS into orthogonal components:
+
+$$
+\begin{aligned}
+\Vert B\mathbf{x} \Vert^2 
+&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast) x_n  + \mathbf{b}_n^\ast x_n\Vert^2 \\
+&= \Vert \sum_{i=1}^{n-1}\mathbf{b}_ix_i + (\sum_{j<n}\mu_{j, n}\mathbf{b}_j^\ast) x_n \Vert^2 + \Vert \mathbf{b}_n^\ast x_n \Vert^2 \\
+&\geq \Vert \mathbf{b}_n^\ast x_n \Vert^2
+\end{aligned}
+$$
+
+where $x_n > 0$, the inequality trivially implies $\Vert B\mathbf{x} \Vert \geq \Vert\mathbf{b}_n^\ast\Vert$.
+where $x_n = 0$, $\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n) = \mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1})$, so by definition of minimum distance we have $\Vert B\mathbf{x} \Vert \geq \lambda(\mathcal{L}(\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}))$.
+Combining the two cases gives us the overall inequality.
+$\blacksquare$
+
+Having proved the recursive inequality, we can simply apply recursion $n$ times and arrive at a concrete lower bound for $\lambda$.
+
+$$
+\lambda \geq \min_{1 \leq i \leq n} \Vert \mathbf{b}_i^\ast \Vert
+$$
+
+\subsection{LLL lattice basis reduction algorithm}
+At the time of this survey, the best algorithm for solving the (approximate) shortest vector problem is the LLL lattice basis reduction algorithm, attributed to Arjen Lenstra, Hendrik Lenstra, and László Lovász.
+The reduction algorithm transforms an input basis into an LLL-reduced form, where the first base vector of the reduced basis is an approximation of the shortest vector.
+
+If time permits, we will discuss the algorithm itself. However, I found it okay to first treat the algorithm as a black box, and instead focus on the properties of the reduced basis.
+
+The definition of a reduced basis is parameterized by a real number $\frac{1}{4} < \delta \leq 1$. A basis $B$ is $\delta$-LLL reduced if the following conditions are satisfied:
+
+- For all $j > i$, $\vert\mu_{j, i}\vert \leq \frac{1}{2}$
+- For all $1 \leq i \leq n-1$, $\delta\Vert \pi_i(\mathbf{b}_i)\Vert^2 \leq \Vert\pi_i(\mathbf{b}_{i+1})\Vert^2$
+
+The second clause of the reduced form is of particular interest because we can use it to measure how well $\mathbf{b}_1$ approximates the shortest vector. First observe the RHS of the condition:
+
+$$
+\begin{aligned}
+\Vert \pi_i(\mathbf{b}_{i+1})\Vert^2 &= \Vert \sum_{j\geq i}\mu_{i+1, j}\mathbf{b}_j^\ast\Vert^2 \\
+&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast + \sum_{j\geq i+1}\mu_{i+1, j}\mathbf{b}_j^\ast\Vert^2 \\
+&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast + \pi_{i+1}(\mathbf{b}_{i+1}) \Vert^2 \\
+&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast + \mathbf{b}_{i+1}^\ast \Vert^2 \\
+&= \Vert \mu_{i+1, i}\mathbf{b}_i^\ast \Vert^2 + \Vert \mathbf{b}_{i+1}^\ast \Vert^2 \\
+\end{aligned}
+$$
+
+Substituting the equation back into the condition:
+
+$$
+\delta\Vert \pi_i(\mathbf{b}_i)\Vert^2 \leq \Vert \mu_{i+1, i}\mathbf{b}_i^\ast \Vert^2 + \Vert \mathbf{b}_{i+1}^\ast \Vert^2
+$$
+
+Which transforms into:
+
+$$
+\Vert \mathbf{b}_i^\ast \Vert^2 \leq \frac{1}{(\delta - \mu_{i+1, i}^2)} \Vert \mathbf{b}_{i+1}^\ast \Vert^2
+$$
+
+By the first clause of the reduced basis we know $\mu_{i+1, i}^2 \leq \frac{1}{4}$, which means that $\frac{1}{\delta - \mu_{i+1, i}^2} \leq \frac{1}{\delta - \frac{1}{4}}$.
+Denote $\alpha = \frac{1}{\delta - \frac{1}{4}}$, then $\alpha > \frac{4}{3}$, and we have the following inequality:
+
+$$
+\Vert \mathbf{b}_i^\ast \Vert^2 \leq \alpha \Vert \mathbf{b}_{i+1}^\ast \Vert^2
+$$
+
+Note that because $\mathbf{b}_1 = \mathbf{b}_1^\ast$, we can recusively evaluate the inequality above can obtain the following closed inequality:
+
+$$
+\Vert \mathbf{b}_1 \Vert^2 \leq \alpha^{i-1} \Vert \mathbf{b}_{i}^\ast \Vert^2 \leq \alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert^2
+$$
+
+The inequality above states that $\mathbf{b}_1$ is not longer than $\alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert$ for all $i$, so it must be not longer than the minimum among $\alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert$:
+
+$$
+\Vert \mathbf{b}_1 \Vert^2 \leq \min_{1 \leq i \leq n}\alpha^{n-1} \Vert \mathbf{b}_{i}^\ast \Vert^2 = \alpha^{n-1} \min_{1 \leq i \leq n}\Vert \mathbf{b}_{i}^\ast \Vert^2
+$$
+
+From the previou section we've derived that $\lambda \geq \min_{1 \leq i \leq n} \Vert \mathbf{b}_i^\ast \Vert$, which we can plug into the inequality above:
+
+$$
+\Vert \mathbf{b}_1 \Vert \leq \alpha^\frac{n-1}{2} \lambda
+$$
+
+This in equality bounds the length of the first base vector of the reduced basis by some multiples of the minimum distance, where the multiple is exponential with respect to the number of dimensions of the lattice.
+
+In other words, the LLL reduction algorithm computes an approximation of the shortest vector within a factor of $\gamma \in O(\alpha^n)$.
+With specific choices of "bad" basis and sufficiently high dimension $n$, the basis reduction algorithm will not be able to provide meaningfully tight approximation of the shortest vector, making the shortest vector problem suitably hard for cryptographic applications.
+
+\section{Nearest plane algorithm}
+In this section we state the closest vector problem (CVP), which is a hard lattice problem that the GGH construction is based on.
+We will then discuss Babai's nearest plane algorithm, which can be used to approximate solutions to the CVP.
+
+Let $\mathcal{L}(B) \subseteq \mathbb{Z}^n$ be a full-rank lattice, and let $\mathbf{t} \in \mathbb{Z}^n$ be a target vector.
+The closest vector problem asks to find $\mathbf{v} \in \mathcal{L}(B)$ that minimizes $\Vert \mathbf{t} - \mathbf{v} \Vert$, where the norm is the Euclidean norm (over real numbers).
+
+\subsection{Fundamental regions}
+Before attempting the closest vector problem, we first introduce a highly relevant concept called the \textbf{fundamental region} of a lattice.
+A fundamental region $S$ is a subset of the (real) linear span of the basis such that it tiles the linear span of the basis and each tile contains exactly one lattice point.
+If we have a fundamental region $S$ defined, then each point in the linear span of the basis can be uniquely decomposed into the sum of a lattice point and a point in the fundamental region:
+
+$$
+\mathcal{L}(B) = \{S + B\mathbf{x} \mid \mathbf{x} \in \mathbb{Z}^n\}
+$$
+
+A few notable examples include the fundamental parallelpiped
+
+$$
+\mathcal{P}(B) = \{B\mathbf{x} \mid \mathbf{x} \in [0, 1)^n\}
+$$
+
+and the centered fundamental parallelpiped
+
+$$
+\mathcal{C}(B) = \{B\mathbf{x} \mid \mathbf{x} \in [-\frac{1}{2}, \frac{1}{2})^n\}
+$$
+
+Computing the decomposition using the (centered) fundmental parallelpiped is easy.
+First compute $B^{-1}\mathbf{t}$ (over $\mathbb{R}$), then perform some kind of rounding (flooring or nearest integer) depending on whether the fundamental parallelpiped is centered or or not.
+$B\lfloor B^{-1}\mathbf{T} \rceil$ is the lattice point whose corresponding fundamental region contains the target.
+This is Babai's rounding algorithm.
+As we will see in later section, this algorithm works well if $B$ is short and orthogonal, but falls apart badly if $B$ is long and skewed.
+
+The Voronoi region is defined by the set of points in the linear span that are closer to $\mathbf{0}$ than to any other lattice points.
+There is some additional details that need to be specified to ensure that the Voronoi region properly closes and forms a partition of $\text{span}(B)$, which we will not discuss in details here.
+If we have an algorithm that can efficiently decompose $\mathbf{t} \in \text{span}(B)$ using the Voronoi region, then we automatically have a way to solve the closest vector problem.
+Unfortunately, no such algorithm is known.
+
+Of particular interest is the (centered) orthogonalized fundamental parallelpiped, which is defined using the orthogonalized basis $B^\ast = \text{Gram-Schmidt}(B)$:
+
+$$
+\mathcal{C}(B^\ast) = \{B^\ast\mathbf{x} \mid \mathbf{x} \in [-\frac{1}{2}, \frac{1}{2})\}
+$$
+
+(If time permits we should try to prove that $\mathcal{C}(B^\ast)$ is indeed a fundamental region. For now, we will take that for granted without proof).
+
+Because $B^\ast$ is an orthogonal basis, it's easy to see that the sphere centered at some lattice point $\mathbf{v} \in \mathcal{L}$ and whose radius is $\frac{1}{2}\min \Vert \mathbf{b}_i^\ast\Vert$ is entirely contained in the (shifted) fundamental region $\text{sphere} \subset \{\mathcal{C}(B^\ast) + \mathbf{v}\}$.
+
+\subsection{Babai's nearest plane algorithm}
+Babai's nearest plane algorithm, attributed to László Babai, is a recursive algorithm that can approximate the closest vector under a given basis to a target vector.
+More specifically, it returns a vector point $\mathbf{v}$ such that, if target vector is projected onto the orthogonalized basis, the projection is contained in $\mathbf{v} + \mathcal{C}(B^\ast)$.
+If the target vector is in the linear span of the basis, then the target vector itself is contained in $\mathbf{v} + \mathcal{C}(B^\ast)$.
+
+The algorithm $\text{NearestPlane}(B, \mathbf{t})$ takes a set of basis $B = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n}]$ and a target vector $\mathbf{t}$, then performs the following procedure:
+
+\begin{enumerate}
+    \item If $B$ is an empty set of basis, return $\mathbf{0}$
+    \item $B^\ast \leftarrow \text{GramSchmidt}(B)$
+    \item $c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil$, where $\mathbf{b}_n^\ast$ is the last base vector in $B^\ast$
+    \item $B' = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_{n-1}]$ is obtained by removing the last base vector from $B$
+    \item return $c\mathbf{b}_n + \text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$
+\end{enumerate}
+
+Denote the output of this algorithm by $\mathbf{v}$, the key theorem about this algorithm is as follows:
+
+$$
+\forall 1 \leq i \leq n, 
+\frac{
+    \langle
+        \mathbf{t} - \mathbf{v}, \mathbf{b}_i^\ast
+    \rangle
+}{
+    \langle
+        \mathbf{b}_i^\ast, \mathbf{b}_i^\ast
+    \rangle
+} 
+\in [-\frac{1}{2}, \frac{1}{2})
+$$
+
+Intuitively, the inequality above states that the deviation of $\mathbf{t}$ from $\mathbf{v}$ is between $-\frac{1}{2}\mathbf{b}_i^\ast$ and $\frac{1}{2}\mathbf{b}_i^\ast$ in any of the chosen direction $\mathbf{b}_i^\ast$, which means that $\mathbf{t} - \mathbf{v}$ is indeed contained in the orthogonalized fundamental parallelpiped.
+
+We can prove this result inductively.
+In the base case, if the input set of basis is empty, then this result is trivially correct.
+In the inductive case, we denote the output of $\text{NearestPlane}(B', \mathbf{t} - c\mathbf{b}_n)$ by $\mathbf{v}^\prime$.
+Assuming that the algorithm produces the desired result for the sublattice $\mathcal{L}(B^\prime)$:
+
+$$
+\forall 1 \leq i \leq (n-1), 
+\frac{
+    \langle
+        (\mathbf{t} - c\mathbf{b}_n) - \mathbf{v}^\prime, 
+        \mathbf{b}_i^\ast
+    \rangle
+}{
+    \langle
+        \mathbf{b}_i^\ast, \mathbf{b}_i^\ast
+    \rangle
+} \in [-\frac{1}{2}, \frac{1}{2})
+$$
+
+However, notice in step 5 of the algorithm $c\mathbf{b}_n + \mathbf{v}^\prime$ is the output of the current iteration of the algorithm, so we have
+
+$$
+\begin{aligned}
+(\mathbf{t} - c\mathbf{b}_n) - \mathbf{v}^\prime 
+&= \mathbf{t} - (c\mathbf{b}_n + \mathbf{v}^\prime) \\
+&= \mathbf{t} - \mathbf{v}
+\end{aligned}
+$$
+
+which means that
+
+$$
+\forall 1 \leq i \leq (n-1), 
+\frac{
+    \langle
+        \mathbf{t} - \mathbf{v}, 
+        \mathbf{b}_i^\ast
+    \rangle
+}{
+    \langle
+        \mathbf{b}_i^\ast, \mathbf{b}_i^\ast
+    \rangle
+} \in [-\frac{1}{2}, \frac{1}{2})
+$$
+
+So all that remains is to show that this relationship also holds for $i = n$.
+
+To prove that the relationship holds for $i = n$, first observe how $c$ is computed:
+
+$$
+c \leftarrow \lfloor \frac{\langle\mathbf{t}, \mathbf{b}_n^\ast\rangle}{\langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle} \rceil
+$$
+
+which is equivalent to saying that
+
+$$
+\frac{
+    \langle\mathbf{t}, \mathbf{b}_n^\ast\rangle
+}{
+    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+}
+- c \in [-\frac{1}{2}, \frac{1}{2})
+$$
+
+Also recall the Gram-Schmid orthogonalization process:
+
+$$
+\mathbf{b}_n^\ast +\sum_{i<n}\mu_{n, i}\mathbf{b}_i^\ast = \mathbf{b}_n
+$$
+
+Therefore
+
+$$
+\begin{aligned}
+\langle\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
+&= \langle 
+    \mathbf{b}_n^\ast +\sum_{i<n}\mu_{n, i}\mathbf{b}_i^\ast, \mathbf{b}_n^\ast
+\rangle \\
+&= \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+\end{aligned}
+$$
+
+All non $\mathbf{b}_n^\ast$ terms can be cleared because they are orthogonal to $\mathbf{b}_n^\ast$, so their inner product is 0. This equality means that:
+
+$$
+\frac{
+    \langle\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
+}{
+    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+} = 1
+$$
+
+Finally we can put everything together:
+
+$$
+\begin{aligned}
+\frac{
+    \langle\mathbf{t}, \mathbf{b}_n^\ast\rangle
+}{
+    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+}
+- c 
+&= \frac{
+    \langle\mathbf{t}, \mathbf{b}_n^\ast\rangle
+}{
+    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+}
+- c \frac{
+    \langle\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
+}{
+    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+} \\
+&= \frac{
+    \langle\mathbf{t} - c\mathbf{b}_n, \mathbf{b}_n^\ast\rangle
+}{
+    \langle\mathbf{b}_n^\ast, \mathbf{b}_n^\ast\rangle
+}
+\in [-\frac{1}{2}, \frac{1}{2})
+\end{aligned}
+$$
+
+Hence we've proved the relationship for $i = n$. $\blacksquare$
+
+\section{The Goldreich-Goldwasser-Halevi trapdoor}
+One class of constructing lattice trapdoor uses a pair of public and secret basis for the same lattice.
+Since the two basis generate the same lattice, they are equally good at mapping integer coordinates into a lattice point.
+On the other hand, the secret basis is very "good" and can be used to efficiently find the closest lattice point, but the public basis is very "bad" at recovering the closest lattice point.
+
+One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper \textit{"Public-key cryptosystem from lattice reduction problem"}.
+At a high level, the trapdoor is parameterized by three items: a "bad" basis $B$, a "good" basis $R$, and an error bound $\sigma$.
+In the forward direction, the function maps a pair of lattice coordinate $\mathbf{v} \in \mathbb{Z}^n$ and a small error vector $\mathbf{e} \leftarrow \{ -\sigma, \sigma \}^n$ to $\mathbf{x} = B\mathbf{v} + \mathbf{e} \in \mathbb{R}^n$.
+If the parameters are generated correctly, then the closest lattice point in $\mathcal{L}(B)$ is exactly $B\mathbf{v}$.
+
+Inverting the function involves recovering the integer coordinate $\mathbf{v}$ (or the error vector $\mathbf{e}$, since recovering one of them automatically gives you the other).
+However, the inversion is exactly the cloest vector problem (CVP), and should be hard if $B$ is a sufficiently bad basis.
+On the other hand, since $R$ is a good basis, finding the closest vector point should be "easy" if we have $R$.
+
+From here, GGH '97 proposed a public-key cryptosystem as well as a digital signature scheme that uses such a trapdoor construction, and the security of the two schemes naturally rest on the hardness of the underlying hard lattice problem.
+
+\subsection{The trapdoor scheme}
+The GGH trapdoor scheme contains four components:
+
+\begin{enumerate}
+    \item \textbf{Parameter generation} The main security parameter in this scheme is the number of dimensions $n$ of the lattice.
+In the original paper, the authors claimed $n \approx 150$ is sufficient for making inverting the function without the private basis hard and $n \approx 250$ should be a safe choice for the foreseeable future.
+    \item \textbf{Key generation} First generate the "good" basis $R \in \mathbb{R}^{n \times }$, then apply some (unimodular) transformation to $R$ to obtain the "bad" basis $B$.
+The error bound $\sigma > 0 \in \mathbb{R}$ is dependent on the choice of $R$ and the choice of "probability of inversion error" $\epsilon >= 0$, which will be discussed in a later section.
+    \item \textbf{Forward evaluation} $f_{B}(\mathbf{v}, \mathbf{e}) = B\mathbf{v} + \mathbf{e}$, where $\mathbf{v} \in \{-n, \ldots, n\}^n$ and $\mathbf{e} \in \{-\sigma, \sigma\}$.
+According to the authors, the choice of bounds for values of $\mathbf{v}$ is arbitrary and not a significant contributor to the overall security of the scheme.
+    \item \textbf{Inversion} Denote the output by $\mathbf{x} = f_{B}(\mathbf{v}, \mathbf{e})$, first attempt to recover the integer coordinate $\mathbf{v} \leftarrow B^{-1}R\lfloor R^{-1}\mathbf{x} \rceil$.
+From here it is trivial to recompute the lattice point $B\mathbf{v}$ and recover the error term $\mathbf{e} = \mathbf{x} - B\mathbf{v}$.
+\end{enumerate}
+
+\subsection{Correctness of trapdoor inversion}
+Without the error term, the function $f_B: \mathbf{v} \mapsto B\mathbf{v}$ is trivially invertible with either choice of the basis.
+However, with a non-zero the error term, the quality of the basis makes a substantial difference in how much error can be added before the points can no longer be recovered.
+
+Observe the calculation used for recovering the integer coordinate $\mathbf{v}$:
+
+$$
+\begin{aligned}
+\lfloor R^{-1}\mathbf{x} \rceil &= B^{-1}R\lfloor R^{-1}(B\mathbf{v} + \mathbf{e})\rceil \\
+&= B^{-1}R\lfloor R^{-1}B\mathbf{v} + R^{-1}\mathbf{e}\rceil
+\end{aligned}
+$$
+
+Since $R, B$ are related by a unimodular matrix and $\mathbf{v}$ is an integer vector, $R^{-1}B\mathbf{v}$ is an integer vector and can be moved out of the "rounding" operator:
+
+$$
+\begin{aligned}
+\lfloor R^{-1}\mathbf{x} \rceil &= B^{-1}RR^{-1}B\mathbf{v} + B^{-1}R\lfloor R^{-1}\mathbf{e}\rceil \\
+&= \mathbf{v} + B^{-1}R\lfloor R^{-1}\mathbf{e} \rceil
+\end{aligned}
+$$
+
+Since $B^{-1}R$ is also a unimodular matrix, we can conclude that the equation above is successful at recovering the original coordinate if and only if $R^{-1}\mathbf{e} = \mathbf{0}$.
+
+To guarantee that inversion error never happens, we can bound the error term $\sigma > 0$ by $\frac{1}{2\rho}$, where $\rho$ is maximal $L_1$ norm among the rows of $R^{-1}$.
+This bound is excessively conservative, however, and we might want to relax the bound to enhance the security of the trapdoor scheme (larger error terms makes it harder to invert the function using only the public basis).
+The authors provided one such relaxation based on the Hoeffding inequality.
+This relaxation is stated as follows
+
+$$
+P(\text{inversion error}) \leq 2n \cdot \exp(-\frac{1}{8\sigma^2\gamma^2})
+$$
+
+Where $\gamma = \sqrt{n} \cdot \max(L_\infty \text{ norm of rows of } R^{-1})$.
+Simple reorganization of inequality shows that to bound the inversion error by $\epsilon$, the error term will be bounded by $\sigma \leq (\gamma\sqrt{8\ln{2n/\epsilon}})^{-1}$.
+
+\subsection{Generating the pair of basis}
+The authors described two similar ways of generating the private basis.
+The first way is to sample each coordinate of $R \in \mathbb{R}^{n \times n}$ from a uniform distribution on $\{-l, -l + 1, \ldots, l-1, l\}$, where according to the authors, the value of the bound $l$ has negligible impact on the quality of the generated basis (the authors chose $\pm 4$ in their implementation).
+A second method is to first generate a square lattice $L(kI)$ for some positive integer $k$, then add a small amount of noise $R^\prime \in \{-l, \ldots, l\}^{n \times n}$.
+With this method of sampling the private basis, it is important to balance the choice of values between $k$ and $l$, where a larger $k$ value gives a more orthogonal basis, but also weakens the security of the trapdoor function by making it easier to reduce the public basis into a short, orthogonal basis using basis reduction algorithm.
+
+The authors also described two methods for generating the public basis $B$ from the private basis.
+The first method is to directly generate random unimodular matrix $T$ and set $B = TR$, then repeat until satisfactory.
+A second method is to repeatedly apply column mixing, where at each mixing a column of $R$ is chosen, and a linear combination of all other columns is added to the chosen column.
+
+Mathematically, the two methods are equivalent.
+However, in implemenetation, we would like the values of $B$ to be smaller for space efficiency while maintaining sufficient security so that the function cannot be easily inverted using the public basis alone and so that $B$ cannot be easily reduced using basis reduction algorithm.
+The authors preferred column mixing for requiring less computation and for producing public basis $B$ with smaller values.
+
+Unfortunately, there is no known rigorous description of how skewed $B$ needs to be for the trapdoor function to be secure.
+The authors relied on experimental methods and determined that for $n \approx 100$, $2n$ steps of column mixing is enough to render LLL basis reduction ineffective at meaningfully improving the quality of the public basis $B$.
+
+\subsection{Inverting the function without the trapdoor}
+From a high level, inverting the function corresponds to finding the an approximate closest lattice point, so the best attacks on the trapdoor are also algorithms for solving the CVP problem.
+
+The most straightforward attack on the trapdoor is to simply run the inversion algorithm with the public basis instead of the private basis:
+
+$$
+B^{-1}\mathbf{x} = B^{-1}(B\mathbf{v} + \mathbf{e}) = \mathbf{v} + B^{-1}\mathbf{e}
+$$
+
+Due to $B$ being very skewed, this procedure does not immediately yield the correct value for $\mathbf{v}$ since $B^{-1}\mathbf{e}$  is not $\mathbf{0}$.
+However, since the possible values of $\mathbf{e}$ is finite, we can perform an exhaustive search on all possible values of $\mathbf{d} = B^{-1}\mathbf{e}$, although the search space will grow exponentially with the number of dimensions $n$.
+In experimental settings, with $n \geq 100$ the search space reaches 168 bits of entropy.
+
+Using a better approximation algorithm for CVP, such as the nearest plane algorithm, yields a more efficient inversion than the brute-force search described above, although the complexity of the algorithm still scales exponentially with $n$.
+In experimental settings, with $n=150$ the workload of the nearest plane algorithm reaches 104 bits.
+
+\subsection{Public key cryptosystem}
+The authors of the GGH paper presented a public key cryptosystem that makes direct usage of the trapdoor.
+The main security parameter is $n$ the number of dimension of the lattice.
+The secret key is the short, orthogonal basis $R$ generated by applying a small amount of noise to a truly orthogonal basis.
+The public key is the long basis $B$ generated by applying column mixing to $R$.
+The encryption function is exactly the evaluation of the trapdoor, and the decryption function is the inversion of the trapdoor.
+
+What remains is the question of "where to encode the message", for which the authors discussed a few options:
+
+\begin{enumerate}
+    \item Use a generic encoding that takes advantage of the hard-core bits of the one-way function, although this encoding is inefficient (it only encodes $\log{n}$ bits at a time), and it does not take advantage of any specific features of the trapdoor construction itself
+    \item Directly encode the message into a lattice point $\mathbf{v}$.
+However, this scheme is insecure because an adversary can compute $B^{-1}\mathbf{c} = \mathbf{v} + B^{-1}\mathbf{e}$, where $B^{-1}\mathbf{e}$ might not be big enough to obscure all information about $\mathbf{v}$, which means this adversary can obtain partial information about the message
+    \item Encode the message into the least significant bits of each entry of $\mathbf{v}$ and choose all other bits randomly.
+The authors argued that with appropriate choice of probability distribution for the error term $\mathbf{e}$, no polynomial-time adversaries will be able to distinguish the parity of the entries of the ciphertext from truly random, thus achieving IND-CPA security.
+\end{enumerate}
+
+Later Micciancio also proposed to encode the message in the error term $\mathbf{e}$, and instead chooses the lattice point $\mathbf{v}$ at random.
+However, this scheme is not trivial because special care is needed for security requirement, and we also have the problem of a lattice being a countably infinite set, making defining probability distribution a non-trivial problem.
+
+\subsection{Digital signature}
+The GGH trapdoor is also suitable for a digital signature: the message is an arbitrary point in the linear span of the good basis, and the signature is the (approximate) closest lattice point, which can be efficiently computed using the good basis $R$.
+The verifier checks that the signature is valid by first using the public basis to verify that the signature is a lattice point, then computing the norm between the signature and the message to check that the distance is sufficiently small.
+
+\section{Conclusion and final comments}
+In this project, we reviewed some fundamental results of lattices, analyzed the difficulty of hard lattice problems and the efficiency of their best known solutions.
+We then discussed a lattice trapdoor construction that takes advantage of an exponential gap between short basis and long basis in their ability to solve the closest vector problem, and applied the trapdoor to build a public key cryptosystem and a digital signature scheme.
+
+Both the public-key cryptosystem and the signature scheme have been cryptanalyzed (by Oded Regev and Phong Q.
+Nguyen respectively) in their original proposed form.
+However, the idea of using a pair of short and long basis to build trapdoors has been extended in later studies.
+Gentry, Peikert, and Vaikuntanathan developed this idea into the GPV framework that later led to FALCON, a digital signature scheme that was submitted to NIST's post-quantum cryptography competition.
+This idea was also behind NTRUSign, whose post-quantum variant was also submitted to NIST's post-quantum cryptography competition.
 

### 2023-11-26 co687/project/main.tex
Use NearestPlane to implement SizeReduce

@@
 
-\subsection{LLL lattice basis reduction algorithm}
+\subsection{Reduced lattice basis}
 At the time of this survey, the best algorithm for solving the (approximate) shortest vector problem is the LLL lattice basis reduction algorithm, attributed to Arjen Lenstra, Hendrik Lenstra, and László Lovász.
 The reduction algorithm transforms an input basis into an LLL-reduced form, where the first base vector of the reduced basis is an approximation of the shortest vector.
-
-If time permits, we will discuss the algorithm itself. However, I found it okay to first treat the algorithm as a black box, and instead focus on the properties of the reduced basis.
+Details of the actual algorithm for obtaining a reduced basis will be discussed in section 4.
+Meanwhile, it is helpful to first discuss the properties of the reduced form itself.
 
@@
 
-- For all $j > i$, $\vert\mu_{j, i}\vert \leq \frac{1}{2}$
-- For all $1 \leq i \leq n-1$, $\delta\Vert \pi_i(\mathbf{b}_i)\Vert^2 \leq \Vert\pi_i(\mathbf{b}_{i+1})\Vert^2$
+\begin{enumerate}
+    \item For all $j > i$, $\vert\mu_{j, i}\vert \leq \frac{1}{2}$
+    \item For all $1 \leq i \leq n-1$, $\delta\Vert \pi_i(\mathbf{b}_i)\Vert^2 \leq \Vert\pi_i(\mathbf{b}_{i+1})\Vert^2$
+\end{enumerate}
 
@@
 
-Also recall the Gram-Schmid orthogonalization process:
+Also recall the Gram-Schmidt orthogonalization process:
 
@@
 
+\subsection{Basis size reduction}
+In addition, to being the best known algorithm for solving the CVP problem, Babai's nearest plane algorithm is also an essential component to the LLL lattice basis reduction algorithm.
+Specifically, the nearest plane algorithm is used to reduce the size of the basis vector so the first condition of the reduced basis (shown below) can be satisfied:
+
+$$
+\forall j<i, \vert\mu_{i, j}\vert \leq \frac{1}{2}
+$$
+
+The size reduction algorithm, which we will denote by $\operatorname{SizeReduce}$, takes a basis $B$ and returns a size-reduced basis $B^\prime$ such that the condition above is true.
+The procedure for $\operatorname{SizeReduce}$ is as follows:
+
+For each of $i \in \{2, 3, \ldots, n\}$
+\begin{enumerate}
+    \item $\mathbf{v} \in \mathcal{L}([\mathbf{b}_1, \ldots, \mathbf{b}_{i-1}]) \leftarrow \operatorname{NearestPlane}([\mathbf{b}_1, \ldots, \mathbf{b}_{i-1}], \mathbf{b}_i - \mathbf{b}_i^\ast)$
+    \item substitute $\mathbf{b}_i$ with $\mathbf{b}_i - \mathbf{v}$, then update $B$ and $B^\ast$
+\end{enumerate}
+
+Recall the result of the nearest plane algorithm we know that:
+
+$$
+\forall j < i,
+\frac{
+    \langle
+        \mathbf{b}_i - \mathbf{b}_i^\ast - \mathbf{v}, \mathbf{b}_j^\ast
+    \rangle
+}{
+    \langle 
+        \mathbf{b}_j^\ast, \mathbf{b}_j^\ast
+    \rangle
+} \in [-\frac{1}{2}, \frac{1}{2})
+$$
+
+Notice in the equation above, we have $\mathbf{b}_i^\ast \perp \mathbf{b}_j^\ast$ because $j < i$.
+We also have $\mathbf{b}_i - \mathbf{v}$ being the new value for $\mathbf{v}_i$ after the substitution.
+This means that after the substitution:
+
+$$
+\forall j < i,
+\frac{
+    \langle
+        \mathbf{b}_i, \mathbf{b}_j^\ast
+    \rangle
+}{
+    \langle
+        \mathbf{b}_j^\ast, \mathbf{b}_j^\ast
+    \rangle
+} \in [-\frac{1}{2}, \frac{1}{2})
+$$
+
+The LHS of the equation above is exactly $\mu_{i,j}$.
+Thus we have satisfied the first condition of $\delta$-LLL reduced basis.
+In addition, because $\mathbf{v} \in \mathcal{L}([\mathbf{b}_1, \ldots, \mathbf{b}_{i-1}])$, the substitution $\mathbf{b}_i \leftarrow \mathbf{b}_i - \mathbf{v}$ is equivalent to adding onto the substituted basis a linear combination of other basis vectors, which corresponds to a unimodular matrix multiplication and thus does not change the lattice itself before or after the size reduction.
+
+\subsection{The LLL basis reduction algorithm}
+By applying $\operatorname{SizeReduce}$ to the basis, we can obtain a size-reduced basis that satisfy the first condition of $\delta$-LLL reduced form, but there is no guarantee that the second condition will hold.
+Pairs of adjacent basis that violate the second condition can be swapped so that the swapped pair does not violate the second condition.
+Because swapping columns is equivalent to right multiply by unimodular matrix, the lattice it preserved, but the first condition will not hold, so we will need to apply $\operatorname{SizeReduce}$ again.
+
+The LLL basis reduction algorithm is exactly a repetition of alternating between size reduction and column swapping.
+For $\frac{1}{4} < \delta < 1$, it can be proved that the number of iteration is bounded by some polynomial factors of the dimension $n$ of the lattice.
+For $\delta = 1$ we will obtain the best possible reduction, but there is no guarantee that the basis reduction will run in polynomial time.
+
 \section{The Goldreich-Goldwasser-Halevi trapdoor}
@@
 
-One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, GoldWasser, and Halevi in their 1997 paper \textit{"Public-key cryptosystem from lattice reduction problem"}.
+One of the earliest instances of such class of lattice trapdoor is proposed by Goldreich, Goldwasser, and Halevi in their 1997 paper \textit{"Public-key cryptosystem from lattice reduction problem"}.
 At a high level, the trapdoor is parameterized by three items: a "bad" basis $B$, a "good" basis $R$, and an error bound $\sigma$.

### 2024-01-27 co789/homework-01/main.tex
Fixed Q3(1)

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% For code
\usepackage{listings}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Q1}
\subsection*{(1)}
The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.

For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.

There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from. On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors. Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span. Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.

In other words:

$$
P(
    \mathbf{a}_t 
    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
) = 1 - q^{t-1-n}
$$

From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:

$$
\begin{aligned}
&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
&= \prod_{j=0}^{m-1} P(
    \mathbf{a}_{j+1} 
    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
) \\
&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
&= \prod_{j=0}^{m-1}(1 - q^{j - n})
\end{aligned}
$$

\subsection*{(2)}
% TODO: this isn't quite right
We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:

$$
\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
$$

And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$

Thus:

$$
\begin{aligned}
&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
\end{aligned}
$$

\subsection*{(3)}
For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:

\lstinputlisting[language=Python]{q1_3.py}

The result is $0.999699519257883$

\newpage

\section*{Q2}
Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$. Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$. As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise. Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.

\newpage

\section*{Q3}
Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean. Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:

$$
P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
$$

\subsection*{(1)}
The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:

$$
P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
$$

On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:

$$
P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
$$

I used some Python code to approximate the statistical distance:

\lstinputlisting[language=Python]{q3_1.py}

The result is $0.017725703977230414$.

\subsection*{(2)}
I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in. Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$. Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$. The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:

$$
P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
$$

\subsection*{(3)}
Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:

$$
P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
$$

Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$. The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.

In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.

\subsection*{(4)}
\subsubsection*{(a)}
A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.

\subsubsection*{(b)}
The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:

$$
n = \frac{512!}{160!120!120!48!48!8!8!}
$$

Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$. The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$. The expectation of such a distribution is $\frac{n-1}{n}$

\newpage

\section*{Q4}
Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.

Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:

$$
\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
$$

Therefore:

$$
E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
$$

Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$. On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$. This is true because shifting a random variable by a constant does not change its variability.

Putting everything together:

$$
E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
$$

On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:

$$
\begin{aligned}
\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
&= \sum_{i=1}^n \var{S_i^2} \\
&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
&= \sum_{i=1}^n \bigg(
    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
    - (\frac{3}{2})^2 
\bigg) \\
&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
&= \frac{15}{4}n
\end{aligned}
$$

Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:

$$
\begin{aligned}
E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
\var{\norm{s}^2} &= \frac{15}{4}n \\
E[\norm{\mathbf{e}}^2] &= n \\
\var{\norm{e}^2} &= \frac{3}{2}n
\end{aligned}
$$

\newpage

\section*{Q5}
\subsection*{(1)}
Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$. Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.

It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that

$$
A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
$$

On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that

$$
\begin{aligned}
\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
\end{aligned}
$$

Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.

\subsection*{(2)}
Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$. Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$. Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:

$$
\begin{aligned}
\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
\end{aligned}
$$

Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random. Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$. We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$. Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$

% TODO: how to recover s_1, s_2 from s_1 + Bs_2?

\newpage

\section*{Q6}
\subsection*{(1)}
Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution. Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:

$$
\begin{cases}
P(I_i=1) = p \\
P(I_i=0) = 1-p
\end{cases}
$$

Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:

$$
\begin{aligned}
X_1 &= \sum_{i=1}^{n_1} I_i \\
X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
\end{aligned}
$$

Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.

Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$. Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:

$$
\begin{aligned}
C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
&= (X_1 + X_2) - E[X_1 + X_2]
\end{aligned}
$$

From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.

\subsection*{(2)}
From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.

Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$. Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$. Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d. centered binomial with parameters $(kn, p)$

\newpage

\section*{Q7}
We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.

Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.

Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$. Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.

It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis. The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$. The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$

\newpage

\section*{Q8}
Recall that the decryption is computed with the following routine:

$$
\begin{aligned}
D(\text{sk}, (\mathbf{c}_1, c_2))
&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
\end{aligned}
$$

Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.

According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$. So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.

With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$. Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions. For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$. The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.

\newpage

\section*{Q9}
Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$. In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.

I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection. This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).

Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.

Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$. Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$. If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.

The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well. Therefore, the two problems are equivalent.

\newpage

\section*{Q10}
Recall from textbook LWE the decryption routine:

$$
\begin{aligned}
D(\text{sk}, (\mathbf{c}_1, c_2))
&= c_2 - \mathbf{c}_1\mathbf{s} \\
&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
\end{aligned}
$$

Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.

\subsection*{(1)}
Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.

If $s_i = 2k$ is even, then

$$
\begin{aligned}
D(\text{sk}, (\mathbf{c}_1, c_2)) 
&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
\end{aligned}
$$

Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.

If $s_i = 2k + 1$ is odd, then

$$
\begin{aligned}
D(\text{sk}, (\mathbf{c}_1, c_2)) 
&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
\end{aligned}
$$

Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability. Thus, with high probability, decryption will be correct if and only if $s_i$ is even.

\subsection*{(2)}
Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$. For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response. The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.

\subsection*{(3)}
We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$. From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key. If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.

Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$. Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:

$$
-\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
$$

Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.

In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$. In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range. Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.

\subsection*{(4)}
Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial. In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem. Due to time constraint, the details will not be covered in this write-up.

\end{document}

### 2024-01-27 co789/homework-01/q1_3.py

@@ -1,0 +1,9 @@
+from sympy import Rational
+
+if __name__ == "__main__":
+    q = 3329
+    n = 256
+    prod = 1
+    for i in range(n):
+        prod *= 1 - Rational(q) ** (i - n)
+    print(f"Prob is {prod.evalf()}")

### 2024-01-27 co789/homework-01/q3_1.py

@@ -1,0 +1,33 @@
+import math
+
+KYBER_Q = 3329
+
+def centered_bin_pmf(val, n, p):
+    if not (0 <= val + n * p <= n):
+        return 0
+    return (
+        math.comb(n, int(val + n * p))
+        * (p ** (val + n * p))
+        * ((1-p) ** (n - val - n * p))
+    )
+
+def rho(val, mu, var):
+    return math.exp(-(val - mu) ** 2 / (2 * var))
+
+def dgaus(val, mu , var):
+    return rho(val, mu, var) / sum(
+        [rho(y, mu, var) for y in range(-KYBER_Q // 2, KYBER_Q // 2 + 1)]
+    )
+
+if __name__ == "__main__":
+    n, p = 6, 0.5
+    mu, var = 0, n * p * (1-p)
+    dist = 0
+    for val in range(
+        math.ceil(mu - KYBER_Q / 2),
+        math.ceil(mu + KYBER_Q / 2),
+    ):
+        lhs = centered_bin_pmf(val, n, p)
+        rhs = dgaus(val, mu, var)
+        dist += 0.5 * abs(lhs - rhs)
+    print(dist)

### 2024-01-28 co789/homework-01/q03/main.tex
Homework 1 reorganization

--- 2024-01-27 co789/homework-01/main.tex
@@
 \usepackage{amsmath,amsfonts,amssymb,amsthm}
-
-\newcommand{\round}[1]{\lfloor {#1} \rceil}
-\newcommand{\norm}[1]{\Vert {#1} \Vert}
-\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
 
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
 
@@
 $$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
+P(X=x) = \binom{n}{x + np}p^{x + np}(1-p)^{n - x - np} = \binom{6}{x + 3}2^{-6}
 $$
@@
 
-The result is $0.017725703977230414$.
+The result is {0.017725703977230414}.
 
@@
 Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
+https://www.overleaf.com/project/65a54df4dabb2a4ef8fbbc0c
 The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
@@
 
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-01-28 co789/homework-01/q04/main.tex
Homework 1 reorganization

--- 2024-01-27 co789/homework-01/main.tex
@@
 
+% For code
+\usepackage{listings}
+
 \newcommand{\round}[1]{\lfloor {#1} \rceil}
@@
 \newcommand{\var}[1]{\operatorname{Var}[{#1}]}
-
-% For code
-\usepackage{listings}
 
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
 
@@
 &= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
+    (\sum_{j=-3}^3(j^4 \cdot \binom{6}{j} \cdot 2^{-6}) 
     - (\frac{3}{2})^2 
@@
 
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-01-28 co789/homework-01/q05/main.tex
Homework 1 reorganization

--- 2024-01-27 co789/homework-01/main.tex
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
 
@@
 
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-01-28 co789/homework-01/q06/main.tex
Homework 1 reorganization

--- 2024-01-27 co789/homework-01/main.tex
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
 
@@
 
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
+Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = \binom{m}{k}$.
 Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
@@
 
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-01-28 co789/homework-01/q08/main.tex
Homework 1 reorganization

--- 2024-01-27 co789/homework-01/main.tex
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
 
@@
 
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-01-28 co789/homework-01/q10/main.tex
Homework 1 reorganization

--- 2024-01-27 co789/homework-01/main.tex
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
 

### 2024-01-29 co789/homework-01/q05/main.tex
Flawed argument about Q5(2)

@@
 \subsection*{(2)}
+For the remainder of this problem, we assume that full-rank LWE with parameters $(m, n, q, U_s, \chi_e)$ exist, which means that $n \leq m$.
+
 Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
@@
 
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
+we can discard the linearly dependent columns of $A$ and feed the truncated sample $(A_1, \mathbf{b})$ into the full-rank LWE oracle.
+If the corresponding full-rank Search-LWE has unique solution denoted by $\mathbf{s}^\prime$, then it must be that $\mathbf{b} - A_1\mathbf{s}^\prime = \mathbf{e}$, where $\mathbf{e}$ is exactly the error term from the original Search-LWE instance $(A, \mathbf{b})$.
 
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
+With the error term recovered, the Search-LWE instance becomes solving noiseless linear equations.
+Because $A$ is not necessarily full-rank, there may be more than one solutions, but since the secret is uniformly randomly sampled, each solution is equally likely the true secret.
+I claim that this is the best we can do in terms of solving Search-LWE for non-full-rank $A$.
 

### 2024-01-30 co789/homework-01/q03/main.tex
Office hour notes

@@
 Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-https://www.overleaf.com/project/65a54df4dabb2a4ef8fbbc0c
+
 The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
@@
 \subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
+A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow (\mathbf{b} - A\hat{\mathbf{s}}) \mod q$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
 

### 2024-01-30 co789/homework-01/q05/main.tex
Office hour notes

@@
 For the remainder of this problem, we assume that full-rank LWE with parameters $(m, n, q, U_s, \chi_e)$ exist, which means that $n \leq m$.
+To differentiate full-rank LWE with possibly-not-full-rank LWE, we denote full-rank LWE with $\operatorname{LWE}^\ast$ and possibly-not-full-rank LWE.
 
@@
 
-we can discard the linearly dependent columns of $A$ and feed the truncated sample $(A_1, \mathbf{b})$ into the full-rank LWE oracle.
-If the corresponding full-rank Search-LWE has unique solution denoted by $\mathbf{s}^\prime$, then it must be that $\mathbf{b} - A_1\mathbf{s}^\prime = \mathbf{e}$, where $\mathbf{e}$ is exactly the error term from the original Search-LWE instance $(A, \mathbf{b})$.
+Since the secret is sampled from a uniformly random distribution, $\mathbf{s}_1$ and $\mathbf{s}_2$ are both uniformly random.
+Therefore, $\mathbf{s}_1 + B\mathbf{s}_2$ is also uniformly random.
+This means that $(A_1, \mathbf{b})$ is a valid sample from $\operatorname{LWE}^\ast(m, n_1, q, U_s, \chi_e)$ where $n_1 < n$.
+Assuming that there is a unique solution to $(A_1, \mathbf{b})$, if we can solve this instance of full-rank LWE, then the solution must be $\mathbf{s}^\prime$ such that $\mathbf{b} - A_1\mathbf{s}^\prime = \mathbf{e}$ is exactly the error term of the original $(A, \mathbf{b})$.
+With the error term recovered, the LWE problem $(A, \mathbf{b})$ becomes solving linear equations, which can be efficiently computed using Gaussian eliminations.
 
-With the error term recovered, the Search-LWE instance becomes solving noiseless linear equations.
-Because $A$ is not necessarily full-rank, there may be more than one solutions, but since the secret is uniformly randomly sampled, each solution is equally likely the true secret.
-I claim that this is the best we can do in terms of solving Search-LWE for non-full-rank $A$.
+It remains to show that given an oracle that solves full-rank LWE of a higher dimension $n$, we can solve full-rank LWE of a lower dimension $n_1 < n$.
+
+Given $(A_1, \mathbf{b})$ where $A \in \mathbb{Z}_q^{m \times n_1}$ is full-rank, I claim without proof that there exists an efficient algorithm that can output a vector $\mathbf{a}^\ast \in \mathbb{Z}_q^m$ such that $[A \;\vert\; \mathbf{a}]$ is still full-rank (see problem 1).
+Using this algorithm, we can augment $A_1$ into a full-rank matrix with $n$ columns $A^\prime = [ A_1 \;\vert\; A_2^\prime ]$.
+We can sample $\mathbf{s}_2^\prime$ from $U_s$ and compute $\mathbf{b}^\prime = \mathbf{b} + A_2^\prime\mathbf{s}_2^\prime$, then $(A^\prime, \mathbf{b}^\prime)$ is a valid instance of $\operatorname{LWE}^\ast(m, n, q, U_s, \chi_e)$.
+Assuming that $(A^\prime, \mathbf{b}^\prime)$ has unique solution, then the solution must be $\mathbf{s}_1 \;\vert\; \mathbf{s}_2^\prime$, and we can obtain $\mathbf{s}_1$, which is the solution to $(A_1, \mathbf{b})$.
+Thus we have solved full-rank LWE of a lower dimension.
+
+P.S.
+the second reduction is very similar to the reduction in which an oracle for higher dimension LWE (not necessarily full-rank) can be used to solve lower dimension LWE by augmenting the lower dimension LWE instance.
+The only difference is that augmenting the lower dimension instance requires sampling linearly independent columns instead of sampling random columns.
 

### 2024-02-02 co789/homework-02/q1/main.tex
Attempt at homework 2

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 1}
Recall the Gram-Schmidt orthogonalization algorithm:

$$
\mathbf{b}_i^\ast = \mathbf{b}_i - \sum_{j<i} \mu_{i,j} \mathbf{b}_j^\ast
$$

where $\mu_{i, j} = \frac{\langle \mathbf{b}_i, \mathbf{b}_j^\ast \rangle}{\langle \mathbf{b}_j^\ast, \mathbf{b}_j^\ast\rangle}$. Re-arranging the procedure gives us a decomposition of the original base vector by the orthogonalized vector:

$$
\mathbf{b}_i = \mathbf{b}_i^\ast + \sum_{j<i}\mu_{i,j}\mathbf{b}_j^\ast
$$

Because $\mathbf{b}_i \bot \mathbf{b}_j$ when $i \neq j$, it is easy to see that for any $j > i$, $\mathbf{b}_j^\ast \bot \mathbf{b}_i$. This is true because $\mathbf{b}_i$ is a linear combination of orthogonalized base vector with index less than or equal to $i$, all such base vectors are orthogonal to $\mathbf{b}_j^\ast$ as is its linear combination.

Let $B = [\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n]$ be the basis of the lattice $\mathcal{L}$ and $B^\ast = [\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_n^\ast]$ be its orthogonalization. For each $\mathbf{v} \in \mathcal{L}$ lattice point, there exists a unique $\mathbf{x} \in \mathbb{Z}^n$ such that $\mathbf{v} = B\mathbf{x}$.

Because $\mathbf{x}$ has finite number of entries, there exists a maximal index $k \in \{1, 2, \ldots, n\}$ such that $x_k$ is non-zero (in other words, $x_l = 0$ for all $l > k$). Observe the inner product between $\mathbf{v}$ and $\mathbf{b}_k^\ast$:

$$
\begin{aligned}
\langle \mathbf{v}, \mathbf{b}_k^\ast \rangle
&= \langle B\mathbf{x}, \mathbf{b}_k^\ast\rangle \\
&= \langle x_k\mathbf{b}_k, \mathbf{b}_k^\ast \rangle \\
&= x_k \norm{\mathbf{b}_k^\ast}^2
\end{aligned}
$$

By Cauchy-Schwarz inequality we know that:

$$
\norm{\mathbf{v}}^2 \cdot \norm{\mathbf{b}_k^\ast}^2 
\geq \langle \mathbf{v}, \mathbf{b}_k^\ast \rangle^2 
= x_k^2 \norm{\mathbf{b}_k^\ast}^4
$$

Because $x_k$ is non-zero and an integer, $x_k^2 \geq 1$. Re-arranging the inequality gives us:

$$
\norm{\mathbf{v}} \geq \norm{\mathbf{b}_k^\ast}
$$

In other words, for each lattice point $\mathbf{v} \in \mathcal{L}$, there exists some orthogonalized base vector $\mathbf{b}_k^\ast$ that is at most as long as $\mathbf{v}$. Therefore, the shortest lattice point is at least as long as the shortest orthogonalized base vector. $\blacksquare$

\end{document}

### 2024-02-02 co789/homework-02/q2/main.tex
Attempt at homework 2

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 2}
\subsection*{(1)}
First, notice that for the given shortest vector $\mathbf{v} \in \mathcal{L}$ and base vector $\mathbf{b}_i$, $\mathbf{v} + \mathbf{b}_i \in \mathcal{L}(B^\prime)$. This is true because:

$$
\begin{aligned}
\mathbf{v} + \mathbf{b}_i
&= \mathbf{b}_1a_1 + \mathbf{b}_2a_2 + \ldots + \mathbf{b}_i(a_i + 1) + \ldots + \mathbf{b}_na_n \\
&= \mathbf{b}_1a_1 + \mathbf{b}_2a_2 + \ldots + \mathbf{b}_i(2k + 1 + 1) + \ldots + \mathbf{b}_na_n \\
&= \mathbf{b}_1a_1 + \mathbf{b}_2a_2 + \ldots + 2\mathbf{b}_i(k + 1) + \ldots + \mathbf{b}_na_n
\end{aligned}
$$

Denote the output of $\operatorname{CVP}_\gamma(B^\prime, \mathbf{b}_i)$ by $\mathbf{u}$, then by the definition of $\gamma$-CVP:

$$
\begin{aligned}
\norm{\mathbf{u} - \mathbf{b}_i} 
&\leq \gamma \min_{\mathbf{x} \in \mathcal{L}(B^\prime)}\norm{\mathbf{b}_i - \mathbf{x}} \\
& \leq \gamma \norm{\mathbf{b}_i - (\mathbf{v} + \mathbf{b}_i)} \\
&= \gamma \norm{\mathbf{v}} = \gamma \lambda_1(\mathcal{L}(B)) 
\end{aligned}
$$

In other words, $\mathbf{u} - \mathbf{b}_i$ is a solution to $\operatorname{SVP}_\gamma(B)$

\subsection*{(2)}
Let $B$ be the basis of a lattice for which we want to solve $\operatorname{SVP}_\gamma(B)$.

We can modify $B$ by replacing one of its base vector $\mathbf{b}_i$ with $2\mathbf{b}_i$. For a chosen $i$, denote the modified basis by $B_i$. In other words:

$$
B_i = \{
    \mathbf{b}_1, \mathbf{b}_2, \ldots, 2 \mathbf{b}_i, \ldots, \mathbf{b}_n
\}
$$

With a $\operatorname{CVP}_\gamma$ oracle, we can solve $\operatorname{CVP}_\gamma(B_i, \mathbf{b}_i)$. Denote the output by $\mathbf{w}_i$. It's easy to see that $\mathbf{w}_i - \mathbf{b}_i \in \mathcal{L}(B)$ because $B_i$ generates a sub-lattice of $\mathcal{L}(B)$.

Notice that if $\mathbf{v} = \sum_{i=1}^n a_i\mathbf{b}_i \in \mathcal{L}(B)$ is a shortest lattice point, then at least one of the coefficient $a_i$ must be odd. This is true because if all of coefficients are even, then $\frac{1}{2}\mathbf{v}$ is necessarily a shorter vector than $\mathbf{v}$, creating a contradiction.

Therefore, for at least one such $i \in \{1, 2, \ldots, n\}$, $\mathbf{u}_i - \mathbf{b}_i$ falls into the scenario described in part (1), and is thus a solution to $\operatorname{SVP}_\gamma(B)$. Any shorter $\mathbf{u}_j - \mathbf{b}_j$ will also suffice.

\begin{algorithm}
\caption{Solve $\gamma$-SVP with $\gamma$-CVP oracle}\label{alg:cap}
\begin{algorithmic}
\State $\mathbf{v} \leftarrow \mathbf{b}_1$ \Comment{Start with some arbitrary lattice point}

\For{$i \in \{1, 2, \ldots, n\}$}
    \State $B_i \leftarrow $ replacing $\mathbf{b}_i$ with $2\mathbf{b}_i$
    \State $\mathbf{u}_i \leftarrow \operatorname{CVP}_\gamma(B_i, \mathbf{b}_i)$
    \If{$\mathbf{u}_i - \mathbf{b}_i$ is shorter than $\mathbf{v}$}
        \State $\mathbf{v} \leftarrow \mathbf{u}_i - \mathbf{b}_i$
    \EndIf
\EndFor

\Return $\mathbf{v}$

\end{algorithmic}
\end{algorithm}

\end{document}

### 2024-02-02 co789/homework-02/q3/main.tex
Attempt at homework 2

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 3}
First, the result of the multiplication in the quotient ring is as follows:

\begin{equation}
    (a_1 + a_2x)(b_1 + b_2x) \equiv (a_1b_1 + a_2b_2\zeta) + (a_1b_2 + a_2b_1)x \mod x^2 - \zeta
\end{equation}

Using schoolbook multiplication, the R.H.S. from above requires 5 multiplication. Using Karatsuba we can compute $a_1b_2 + a_2b_1$ using only one multiplication (but at the expense at more addition/subtraction):

\begin{equation}
    a_1b_2 + a_2b_1 = (a_1 + a_2)(b_1 + b_2) - a_1b_1 - a_2b_2
\end{equation}

The R.H.S. of equation (2) only takes one multiplication because $a_1b_1$ and $a_2b_2$ have already been computed from previous steps of schoolbook multiplication.

Putting everything together:

\begin{algorithm}
\caption{Karatsuba-ish monomial multiplication}\label{alg:cap}
\begin{algorithmic}

    \State Start with $a_1 + a_2x$ and $b_1 + b_2x$
    \State $c_1 \leftarrow a_1 b_1$ \Comment first multiplication
    \State $c_3 \leftarrow a_2b_2$ \Comment second multiplication
    \State $c_2 \leftarrow (a_1 + a_2)(b_1 + b_2) - c_1 - c_3$ \Comment third multiplication
    \State $c_3 \leftarrow c_3 \zeta$ \Comment fourth multiplication

\Return $(c_1 + c_3) + c_2x$

\end{algorithmic}
\end{algorithm}

\end{document}

### 2024-02-02 co789/homework-02/q4/main.tex
Attempt at homework 2

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 4}

\subsection*{(a)}
By the definition of the (quotient) ring $R_q = \mathbb{Z}_q[x] / \langle p(x) \rangle$ we know that:

\begin{equation}
    a(x)s(x) + e(x) = p(x)g(x) + b(x)
\end{equation}

Where $g(x)$ is some polynomial in $\mathbb{Z}_q[x]$.

Where $\omega$ is a root of $p(x)$, evaluating equation (1) at $x = \omega$ is as follows:

$$
a(\omega)s(\omega) + e(\omega) = 0 \cdot g(\omega) + b(\omega) = b(\omega)
$$

$\blacksquare$

\subsection*{(b)}
From equation (1) we know that $a(\omega)s(\omega) + e(\omega) = b(\omega)$ if and only if $p(\omega)g(\omega) = 0$. Where $\omega$ is not a root of $p(x)$, the equality holds if and only if $g(\omega) = 0$.

While $g(\omega) = 0$ does not hold in general, in specific cases it can still happen. For example, if the sum of degrees of $a(x)$ and $s(x)$ is less than the degree of $p(x)$, then $g(x) = 0$ (aka $b(x)$ does not need to be reduced modulus $p(x)$).

\end{document}

### 2024-02-02 co789/homework-02/q5/main.tex
Attempt at homework 2

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 5}
By the Chinese Remainder Theorem we know that

\begin{equation}
    \frac{\mathbb{Z}_q[x]}{\langle x^{256} + 1\rangle} \cong
    \frac{\mathbb{Z}_q[x]}{\langle x - \zeta\rangle}
    \times \frac{\mathbb{Z}_q[x]}{\langle x - \zeta^3\rangle}
    \ldots
    \times \frac{\mathbb{Z}_q[x]}{\langle x - \zeta^{511}\rangle}
\end{equation}

Because of the isomorphism between the large quotient ring $R_q$ and the product space of the smaller rings, an element in $R_q$ is invertible if and only if its NTT representation is invertible, which happens if and only if each entry in the NTT representation is invertible.

Notice that because each of the smaller ring is quotient on a degree-1 polynomial, $\mathbb{Z}_q[x] / \langle x - \zeta^{2k + 1} \rangle$ is also isomorphic to $\mathbb{Z}_q$. There are $\phi(q) = q-1$ invertible elements in $\mathbb{Z}_q$ for a prime $q$, so the probability of drawing an invertible element from $\mathbb{Z}_q$ is $\frac{q-1}{q} = (1 - \frac{1}{q})$, and the probability of independently drawing 256 invertible elements is $(1 - \frac{1}{q})^{256}$
\end{document}

### 2024-02-02 co789/homework-02/q6/main.tex
Attempt at homework 2

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 6}
\subsection*{(a)}
We denote a polynomial with degree $\phi(n) - 1$ by $f(x) = f_0 + f_1x + \ldots + f_{\phi(n)-1}x^{\phi(n)-1}$. It is trivially true that if $f(x) = g(x)$, then $f(\zeta_i) = g(\zeta_i)$ for all $1 \leq i \leq \phi(n)$.

On the other hand, if for two polynomials $f, g$ with degree $\phi(n) - 1$, their NTT representations are exactly identical, then for all $1 \leq i \leq \phi(n)$:

$$
f(\zeta_i) = g(\zeta_i)
$$

Define $h(x) = f(x) - g(x)$, then $\zeta_1, \zeta_2, \ldots, \zeta_{\phi(n)}$ are distinct roots of $h(x)$. This means that $h(x)$ must have form:

$$
h(x) = h^\prime(x)(x - \zeta_1)(x - \zeta_2) \ldots (x - \zeta_{\phi(n)})
$$

Because $f(x), g(x)$ both have degree $\phi(n) - 1$, the degree of $f(x) - g(x)$ cannot be more than $\phi(n) - 1$. Therefore, $h(x)$ must be $0$, because otherwise it must have degree of at least $\phi(n)$. $\blacksquare$

\subsection*{(b)}

\end{document}

### 2024-02-07 co789/homework-02/q4/main.tex
Attribution to collaborators; Q4.2's more specific instructions

@@
 
-While $g(\omega) = 0$ does not hold in general, in specific cases it can still happen.
-For example, if the sum of degrees of $a(x)$ and $s(x)$ is less than the degree of $p(x)$, then $g(x) = 0$ (aka $b(x)$ does not need to be reduced modulus $p(x)$).
+Because $a(x), s(x), e(x) \in R_q$ are all polynomials of degree (up to) $d-1$, and $p(x)$ is a polynomial of degree $d$, the degree of $g(x)$ cannot be more than $2 \cdot (d - 1) - d = d - 2$.
+By the fundamental theorem of algebra we know that $g(x)$ cannot have more than $d-2$ roots in $\mathbb{Z}_q$.
+Therefore, if $q > d-2$, then there exists $\omega \in \mathbb{Z}_q$ such that $\omega$ is not a root of $g(x)$, which means that $b(\omega) \neq a(\omega)s(\omega) + e(\omega)$.
+
+On the other hand, if the sum of degrees of $a(x)$ and $s(x)$ is less than the degree of $p(x)$, then $g(x) = 0$ (aka $b(x)$ does not need to be reduced modulus $p(x)$).
 

### 2024-02-17 co789/homework-03/q5/main.tex
Skeleton for homework 03

--- 2024-02-02 co789/homework-02/q6/main.tex
@@
 
-\section*{Question 6}
-\subsection*{(a)}
-We denote a polynomial with degree $\phi(n) - 1$ by $f(x) = f_0 + f_1x + \ldots + f_{\phi(n)-1}x^{\phi(n)-1}$.
-It is trivially true that if $f(x) = g(x)$, then $f(\zeta_i) = g(\zeta_i)$ for all $1 \leq i \leq \phi(n)$.
-
-On the other hand, if for two polynomials $f, g$ with degree $\phi(n) - 1$, their NTT representations are exactly identical, then for all $1 \leq i \leq \phi(n)$:
-
-$$
-f(\zeta_i) = g(\zeta_i)
-$$
-
-Define $h(x) = f(x) - g(x)$, then $\zeta_1, \zeta_2, \ldots, \zeta_{\phi(n)}$ are distinct roots of $h(x)$. This means that $h(x)$ must have form:
-
-$$
-h(x) = h^\prime(x)(x - \zeta_1)(x - \zeta_2) \ldots (x - \zeta_{\phi(n)})
-$$
-
-Because $f(x), g(x)$ both have degree $\phi(n) - 1$, the degree of $f(x) - g(x)$ cannot be more than $\phi(n) - 1$.
-Therefore, $h(x)$ must be $0$, because otherwise it must have degree of at least $\phi(n)$.
-$\blacksquare$
-
-\subsection*{(b)}
+\section*{Question 1}
 

### 2024-02-21 co789/homework-03/q1/main.tex
homework 03 draft

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\indcpa}{\mathcal{A}_\text{IND-CPA}}
\newcommand{\indcca}{\mathcal{A}_\text{IND-CCA}}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 1}

We show that IND-CCA security implies IND-CPA security by showing that if an IND-CPA adversary can win with non-negligible advantage, then it can be used to build an IND-CCA adversary who can win with non-negligible advantage.

The key generation routines are identical between the IND-CPA and the IND-CCA game. When $\indcca$ receives the keypair, it directly passes it to $\indcpa$. $\indcpa$ can use the public key to perform encryption queries but does not submit any decryption queries.

When $\indcpa$ generates the challenge plaintext $m_0, m_1$, $\indcca$ passes them to the IND-CCA challenger and receives the challenge ciphertext $c_b$. $\indcca$ passes $c_b$ to $\indcpa$ and receives the guess $b^\ast$ from $\indcpa$.

Because both the IND-CPA and the IND-CCA games are played with identical keypairs, $\indcpa$'s guess is correct if and only if $b^\ast$ is equal to $b$. Therefore, $\indcca$'s advantage is equal to $\indcpa$'s advantage, meaning that if $\indcpa$ has non-negligible advantage, then $\indcca$ has non-negligble advantage.

\end{document}

### 2024-02-21 co789/homework-03/q2/main.tex
homework 03 draft

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 2}

\subsection*{(1)}
Recall in the IND-CPA security proof, we defined three games:

\begin{enumerate}
    \item Game 0 is the standard IND-CPA game for $\operatorname{Module-LWE}$
    \item Game 1 is identical to game 0, except in key generation, $\mathbf{b} = A\mathbf{s} + \mathbf{e}$ is replaced with a uniformly random sample $\mathbf{b} \leftarrow R_q^k$
    \item Game 2 is identical to game 1, except the challenge ciphertexts are replaced with uniformly random samples $\mathbf{c}_1^\ast \leftarrow R_q^k, c_2^\ast \leftarrow R_q$
\end{enumerate}

Further more, we defined two solvers for Module-decisional-LWE: solver 1 solves dLWE with $A \in R_q^{k \times k}$ and solver 2 solves dLWE with $A \in R_q^{(k+1) \times k}$.

Solver 2 decomposes $A, \mathbf{b}$ into the first $k$ rows and the last row: $A = [A_1 \in R_q^{k \times k},  A_2 \in R_q^{1 \times k}]$, $\mathbf{b} = [\mathbf{b}_1\in R_q^{k}, b_2 \in R_q]$. $A_1, A_2$ is given to the IND-CPA adversary as the public key, and $\mathbf{c}_1^\ast = \mathbf{b}_1, c_2^\ast = b_2 + m \round{\frac{q}{2}}$ as the challenge ciphertext. If Solver 2 receives LWE sample, then the IND-CPA adversary is playing game 1; if solver 2 receives truly random sample, then IND-CPA adversary is playing game 2. Therefore, the advantage of solver 2 is $\frac{1}{2}(\operatorname{adv}_1 - \operatorname{adv}_2)$

If in the encryption routine, the second error term $e^{\prime\prime}$ is removed, then $c_2^\ast = b_2 + m\round{\frac{q}{2}} = A_2\mathbf{s} + e_2 + m\round{\frac{q}{2}}$ is no longer a valid encryption of $m$. This means that when $A, \mathbf{b}$ is a LWE sample, the IND-CPA adversary is not playing game 1, but a new game that is identical to game 1 but with the second error term in the encryption routine. Denote the IND-CPA adversary's advantage in this game by $\operatorname{adv}_3$.

Following the same procedure as in the IND-CPA security proof, we can show that:

$$
\text{Adv in solving dLWE($R_q^{k \times k}$)} + 
\text{Adv in solving dLWE($R_q^{k+1 \times k}$)}
= \frac{1}{2}(\operatorname{adv}_0 - \operatorname{adv}_1) 
+ \frac{1}{2}(\operatorname{adv}_3 - \operatorname{adv}_2)
$$

Knowing that game 2 is unwinnable and that solving dLWE with higher dimension is harder, we can rearrange the equation above:

$$
\operatorname{adv}_0 
\geq 4 \cdot \operatorname{adv}_\text{dLWE(k)} + (\operatorname{adv}_1 - \operatorname{adv}_3)
$$

It's possible that $\operatorname{adv}_1 - \operatorname{adv}_3$ is non-negligible, so $\operatorname{adv}_0$ might be non-negligible, thus breaking IND-CPA security of the modified encryption routine.

\subsection*{(2)}
Observe that when $m = \mathbf{0} \in R_{\{0,1\}}$, $c_2 = \mathbf{r}^\intercal \mathbf{b}$ is divisible by $\mathbf{b}$. On the other hand, if $m$ is non-zero, then with very high probability, $c_2$ will not be divisible by $\mathbf{b}$.

Thus, an IND-CPA can challenge plaintexts as follows: $m_0 = \mathbf{0}$, $m_1$ is some arbitrary non-zero polynomial. Upon receiving the challenge ciphertext $c^\ast$, check if $c_2$ is divisible by $\mathbf{b}$ (which is part of the public key and hence accessible). If yes, then $c^\ast$ is the encryption of $m_0$; otherwise $c^\ast$ is the encryption of $m_1$.

\end{document}

### 2024-02-21 co789/homework-03/q3/main.tex
homework 03 draft

--- 2024-02-02 co789/homework-02/q6/main.tex
@@
 
-\section*{Question 6}
-\subsection*{(a)}
-We denote a polynomial with degree $\phi(n) - 1$ by $f(x) = f_0 + f_1x + \ldots + f_{\phi(n)-1}x^{\phi(n)-1}$.
-It is trivially true that if $f(x) = g(x)$, then $f(\zeta_i) = g(\zeta_i)$ for all $1 \leq i \leq \phi(n)$.
+\section*{Question 3}
 
-On the other hand, if for two polynomials $f, g$ with degree $\phi(n) - 1$, their NTT representations are exactly identical, then for all $1 \leq i \leq \phi(n)$:
+Observe the following:
 
 $$
-f(\zeta_i) = g(\zeta_i)
+\begin{aligned}
+A\mathbf{z} &= A(\mathbf{y} + c\mathbf{s}) \\
+&= \mathbf{w} + cA\mathbf{s} \\
+&= \mathbf{w} + c(\mathbf{t} - \mathbf{e})
+\end{aligned}
 $$
 
-Define $h(x) = f(x) - g(x)$, then $\zeta_1, \zeta_2, \ldots, \zeta_{\phi(n)}$ are distinct roots of $h(x)$. This means that $h(x)$ must have form:
+Assuming that $c$ is invertible, re-arranging the equation above gives us:
 
 $$
-h(x) = h^\prime(x)(x - \zeta_1)(x - \zeta_2) \ldots (x - \zeta_{\phi(n)})
+\mathbf{e} = \mathbf{t} - c^{-1}(A\mathbf{z} - \mathbf{w})
 $$
 
-Because $f(x), g(x)$ both have degree $\phi(n) - 1$, the degree of $f(x) - g(x)$ cannot be more than $\phi(n) - 1$.
-Therefore, $h(x)$ must be $0$, because otherwise it must have degree of at least $\phi(n)$.
-$\blacksquare$
-
-\subsection*{(b)}
+From here we can attempt to recover the secret key $\mathbf{s}$ by solving $A\mathbf{s} = \mathbf{t} - \mathbf{e}$.
+While this is an instance of inhomogeneous SIS, with proto-Dilithium $A$ is a wide matrix, which makes SIS easier to solve.
+In the lecture notes, we simply assume that wide SIS is "too easy".
 

### 2024-02-21 co789/homework-03/q5/main.tex
homework 03 draft

@@
 
-\section*{Question 1}
+\section*{Question 5}
+
+\subsection*{1}
+For a given message $m$, we can forge a signature under the public key $(A, \mathbf{t})$ using the following procedure:
+
+\begin{enumerate}
+    \item Sample $\mathbf{y} \leftarrow \chi_y$ 
+    \item Compute $\mathbf{w} \leftarrow A\mathbf{y}$
+    \item Compute $c \leftarrow H(\mathbf{w}, \mathbf{t})$
+    \item Give $(A, \mathbf{w} + c\mathbf{t})$ to the $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1 - np\tau)$ solver, which returns some $\mathbf{z}$
+    \item Output $\sigma = (\mathbf{w}, c, \mathbf{z})$
+\end{enumerate}
+
+$\sigma$ is a valid forgery because $\mathbf{z}$ as a solution to the $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1 - np\tau)$ problem satisfies the verification conditions:
+
+\begin{enumerate}
+    \item $\norm{\mathbf{z}}_\infty \leq \gamma_1 - np\tau$
+    \item $A\mathbf{z} = \mathbf{w} + c\mathbf{t}$
+\end{enumerate}
+
+\subsection*{2}
+The key recovery attack is as follows:
+
+\begin{enumerate}
+    \item Sample some random message $m$
+    \item Query the signature of $m$, which is $(\mathbf{w}, c, \mathbf{z})$
+    \item Give $(A, \mathbf{w})$ to $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1)$ solver, which returns some $\mathbf{y}$
+    \item Compute $\mathbf{s} = c^{-1}(\mathbf{z} - \mathbf{y})$. $\mathbf{s}$ is the secret key
+\end{enumerate}
+
+This procesure works because for the queried signature to be valid, it must satisfy $\mathbf{z} = \mathbf{y} + c\mathbf{s}$, meaning that if we can recover $\mathbf{y}$, then we can recover $\mathbf{s}$.
+In this instance, a valid $\mathbf{y}$ is recovered using the solver on $A\mathbf{y} = \mathbf{w}$.
 

### 2024-03-08 co789/homework-04/q1/main.tex
template for homework 4

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage{algorithm}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{First section}
\begin{definition}[Lattice]\label{lattice-def}
A lattice is a discrete subgroup of $\mathbb{R}^n$
\end{definition}

\begin{theorem}[Minkowski's bound]
    let $\mathcal{L}(B)$ be a full-rank lattice with basis $B \in \mathbb{R}^{n \times n}$, and $B^\ast = [\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_n^\ast]$ be the Gram-Schmidt orthogonalization of $B$, then
    \begin{equation}
        \lambda_1(\mathcal{L}(B)) \geq \min_{1 \leq i \leq n}\norm{\mathbf{b}_i^\ast}
    \end{equation}
\end{theorem}

\begin{algorithm}
\caption{Caption for the algorithm}\label{alg:cap}
\begin{algorithmic}[1]  % [1] is for displaying line number
    \Require This input, and that input
    \State Make some statement with the "State" command

    \For{Wrap for loop condition within the "For" command}
    \EndFor

    \While{Wrap while loop condition within "While" command}
    \EndWhile

    \If{Wrap if else in the "If" command}
    \ElsIf{Another condition}
    \Else
    \EndIf

    \Return Return the output using the "Return" command

\end{algorithmic}
\end{algorithm}

\end{document}

### 2024-03-10 co789/homework-04/q1/main.tex
Homework 4 draft

@@
 %%%% TITLE %%%%%
-\maketitle
+% \maketitle
 
-\section{First section}
-\begin{definition}[Lattice]\label{lattice-def}
-A lattice is a discrete subgroup of $\mathbb{R}^n$
-\end{definition}
+\section*{Question 1}
+P.S. SPHINCS is surprisingly annoying to type. I will abbreviate it with SPX.
 
-\begin{theorem}[Minkowski's bound]
-    let $\mathcal{L}(B)$ be a full-rank lattice with basis $B \in \mathbb{R}^{n \times n}$, and $B^\ast = [\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_n^\ast]$ be the Gram-Schmidt orthogonalization of $B$, then
-    \begin{equation}
-        \lambda_1(\mathcal{L}(B)) \geq \min_{1 \leq i \leq n}\norm{\mathbf{b}_i^\ast}
-    \end{equation}
-\end{theorem}
+We first briefly recall the key generation and signing procedure of SPX.
+In SPX key generation, a secret seed is randomly sampled, and a root WOTS keypair $(\text{PK}^\text{WOTS}_\text{root}, \text{SK}^\text{WOTS}_\text{root})$ is pseudorandomly generated from the secret seed.
+The SPX public key is the hash of the root WOTS public key, and the SPX secret key is the secret seed.
 
-\begin{algorithm}
-\caption{Caption for the algorithm}\label{alg:cap}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \Require This input, and that input
-    \State Make some statement with the "State" command
+When signing a message, a FORS keypair $(\text{PK}^\text{FORS}, \text{SK}^\text{FORS})$ is pseudorandomly generated based on the message and the seed.
+The message is signed by $\text{FORS}$.
+The $\text{FORS}$ public key is authenticated by a SPX hypertree (where each node is a XMDSS tree with $2^t$ WOTS keypairs).
+The SPX hypertree is finally authenticated using the root WOTS keypair by signing the root of the root XMSS tree in the SPX hypertree.
+The hyper tree is pseudorandomly generated using the secret seed alone so that under the same SPX keypair, the root WOTS keypair always signs the same value.
 
-    \For{Wrap for loop condition within the "For" command}
-    \EndFor
+If the hyper tree is pseudorandomly generated using BOTH the secret seed and the message, then for distinct messages, each XMSS tree in the hyper tree will be distinct, and the root WOTS keypair will sign the root of a distinct XMSS tree, meaning that the root WOTS keypair signs distinct messages.
+\textbf{Therefore, with each distinct query signature query, the security of the root WOTS keypair degrades}.
 
-    \While{Wrap while loop condition within "While" command}
-    \EndWhile
+\subsection*{A EF-CMA adversary}
 
-    \If{Wrap if else in the "If" command}
-    \ElsIf{Another condition}
-    \Else
-    \EndIf
+Let $\mathcal{A}$ denote an EF-CMA adversary against the modified SPX.
+It first makes a large number of distinct queries.
+For each query $(m_i, \sigma_i)$, the returned signature $\sigma_i$ contains a WOTS signature $\sigma_i^\text{WOTS}$ signed using the root WOTS keypair.
+$\mathcal{A}$ chooses a message $m^\ast$, then executes the $\text{SPX.KeyGen}()$ and $\text{SPX.Sign()}$ routine up to the point where the root of the root XMSS tree of the hyper tree has been computed, which we denote by $\text{root}(T_0^\ast)$.
 
-    \Return Return the output using the "Return" command
-
-\end{algorithmic}
-\end{algorithm}
+Because WOTS is a one-time signature, we know that with a large number of known signatures on distinct messages, the probability that $\mathcal{A}$ can forge a signature on $\text{root}(T_0^\ast)$ under the same keypair is non-negligible.
+Since $\mathcal{A}$ self-generated the FORS and the hypertree, all verification involving FORS and the hypertree will pass; if $\mathcal{A}$ can forge a signature under the true root WOTS keypair, then the entire verification will pass (we claim without proof that if $(\hat{m}, \hat{\sigma})$ is a valid forgery, then $\text{WOTS.PubKeyRegen}(\hat{m}, \hat{\sigma})$ will correctly recover the true $\text{PK}^\text{WOTS}$).
+Thus $\mathcal{A}$ has a non-negligible chance of successful forgery (and can quickly improve the probability of forgery by repeatedly sample different seeds).
 

### 2024-03-10 co789/homework-04/q2/main.tex
Homework 4 draft

--- 2024-03-08 co789/homework-04/q1/main.tex
@@
 %%%% TITLE %%%%%
-\maketitle
+% \maketitle
 
-\section{First section}
-\begin{definition}[Lattice]\label{lattice-def}
-A lattice is a discrete subgroup of $\mathbb{R}^n$
-\end{definition}
+\section*{Question 2}
+In SPX, each WOTS keypair $\text{WOTS}$ signs the root of child XMSS tree(s).
+If WOTS is replaced by some $k$-time signature scheme, then each leaf node in each XMSS tree can sign the roots of $k$ child XMSS tress.
+Fixing each XMSS tree to still have $2^t$ leaf notes, then each XMSS tree can have $k \cdot 2^t$ children.
+If the entire SPX hypertree has $d$ layers, then it can sign a total of $(k \cdot 2^t)^d$ messages.
 
-\begin{theorem}[Minkowski's bound]
-    let $\mathcal{L}(B)$ be a full-rank lattice with basis $B \in \mathbb{R}^{n \times n}$, and $B^\ast = [\mathbf{b}_1^\ast, \mathbf{b}_2^\ast, \ldots, \mathbf{b}_n^\ast]$ be the Gram-Schmidt orthogonalization of $B$, then
-    \begin{equation}
-        \lambda_1(\mathcal{L}(B)) \geq \min_{1 \leq i \leq n}\norm{\mathbf{b}_i^\ast}
-    \end{equation}
-\end{theorem}
+From the lecture notes we know that for a target security level $\lambda$, we want:
 
-\begin{algorithm}
-\caption{Caption for the algorithm}\label{alg:cap}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \Require This input, and that input
-    \State Make some statement with the "State" command
+$$
+(k \cdot 2^t)^d = 2^{2\lambda}
+$$
 
-    \For{Wrap for loop condition within the "For" command}
-    \EndFor
-
-    \While{Wrap while loop condition within "While" command}
-    \EndWhile
-
-    \If{Wrap if else in the "If" command}
-    \ElsIf{Another condition}
-    \Else
-    \EndIf
-
-    \Return Return the output using the "Return" command
-
-\end{algorithmic}
-\end{algorithm}
+Which solves to $d = \frac{2\lambda}{\log_2{k} + t}$.
+With larger $k$ (aka the signature scheme can sign more messages without losing security), \textbf{we need fewer layers in the hypertree to accomplish the same security level}.
+Since the signature, signing routine, and verification routine all iterate through all $d$ layers of the hypertree, \textbf{signature size, signing time, and verification time all decrease linearly as $d$ decreases}
 

### 2024-03-10 co789/homework-04/q6/main.tex
Homework 4 draft

--- 2024-03-10 co789/homework-04/q2/main.tex
@@
 
-\section*{Question 2}
-In SPX, each WOTS keypair $\text{WOTS}$ signs the root of child XMSS tree(s).
-If WOTS is replaced by some $k$-time signature scheme, then each leaf node in each XMSS tree can sign the roots of $k$ child XMSS tress.
-Fixing each XMSS tree to still have $2^t$ leaf notes, then each XMSS tree can have $k \cdot 2^t$ children.
-If the entire SPX hypertree has $d$ layers, then it can sign a total of $(k \cdot 2^t)^d$ messages.
+\section*{Question 6}
+We will prove equivalence by showing that each statement implies the other.
 
-From the lecture notes we know that for a target security level $\lambda$, we want:
+If (1) is true for all $\mathbf{c}_1 \neq \mathbf{c}_2$, then we can fix $\mathbf{c}_2 = \mathbf{0}$, and statement (1) becomes
 
 $$
-(k \cdot 2^t)^d = 2^{2\lambda}
+\norm{\mathbf{c}_1 - \mathbf{0}}_H = \norm{\mathbf{c}_1}_H \geq d
 $$
 
-Which solves to $d = \frac{2\lambda}{\log_2{k} + t}$.
-With larger $k$ (aka the signature scheme can sign more messages without losing security), \textbf{we need fewer layers in the hypertree to accomplish the same security level}.
-Since the signature, signing routine, and verification routine all iterate through all $d$ layers of the hypertree, \textbf{signature size, signing time, and verification time all decrease linearly as $d$ decreases}
+for all non-zero $\mathbf{c}_1$, thus statement (2) is true.
+
+Now suppose (2) is true. For all $\mathbf{c}_2 \neq \mathbf{c}_1$, $\mathbf{c}_2 - \mathbf{c}_1$ is a non-zero code in $C$ because $C$ is linear. Therefore:
+
+$$
+\norm{\mathbf{c}_2 - \mathbf{c}_1}_H \geq d
+$$
+
+thus statement (1) is true.
 

### 2024-03-13 co789/project/manuscript/main.tex
I can read more about CCA transformation

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage{algorithm}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{A survey of IND-CCA constructions}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{CO 789, Winter 2024}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{The Fujisaki-Okamoto Transformation}
The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} takes a IND-CPA public-key encryption scheme (PKE) and a few other cryptographic primitives of weaker security as inputs, and constructs a hybrid encryption scheme that achieves IND-CCA2 security.

In the conference paper, the security reduction is performed under the random oracle model, and the security of the hybrid scheme degrades linearly with the number of hash queries and the number of decryption queries. On the other hand, this scheme is desirable for its simplicity, and the low requirements for the input primitives: the PKE only needs to be one-time one-way secure, and the symetric cipher only needs to be indistinguishable under one-time attack.

% TODO: Need more introduction?

\subsection{The hybrid scheme and security result}
The hybrid encryption scheme contains three routines: key generation, encryption, and decryption. The input for the hybrid scheme includes a public-key encryption scheme $\text{KeyGen}^\text{asym}, E^\text{asym}, D^\text{asym}$, a symmetric encryption scheme $E^\text{sym}, D^\text{sym}$, and two hash functions $G: \mathcal{M}^\text{asym} \mapsto \mathcal{K}^\text{sym}, H: \mathcal{M}^\text{asym} \times \mathcal{C}^\text{sym} \mapsto \text{Coin}^\text{asym}$.

\begin{algorithm}
\caption{FO Key generation}\label{fo-key-gen}
\begin{algorithmic}[1]  % [1] is for displaying line number
    \State $
        (\text{PK}^\text{asym}, \text{SK}^\text{asym}) 
        \leftarrow \text{PKE.KeyGen}()
    $
    \State $
        \text{PK}^\text{hy} \leftarrow \text{PK}^\text{asym},
        \text{SK}^\text{hy} \leftarrow \text{SK}^\text{asym},
    $

    \Return $(\text{PK}^\text{hy}, \text{SK}^\text{hy})$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{FO Key encryption}\label{fo-key-enc}
\begin{algorithmic}[1]  % [1] is for displaying line number
    \Require $m \in \mathcal{M}^\text{sym}$
    \State Sample from the PKE's message space $\sigma \leftsample \mathcal{M}^\text{asym}$
    \State $a \leftarrow G(\sigma)$, $c \leftarrow E^\text{sym}_a(m)$
    \State $h \leftarrow H(\sigma, c)$
    \State $e \leftarrow E^\text{asym}(\text{PK}^\text{hy}, \sigma, h)$

    \Return $(e, c)$
\end{algorithmic}
\end{algorithm}

\pagebreak

The decryption routine:

\begin{algorithm}
\caption{FO Key decryption}\label{fo-key-dec}
\begin{algorithmic}[1]  % [1] is for displaying line number
    \Require The ciphertext $(e, c)$

    \State $
        \hat{\sigma} \leftarrow D^\text{asym}(\text{SK}^\text{hy}, e)
    $

    \State $
        \hat{h} \leftarrow H(\hat{\sigma}, c)
    $

    \State $
        \hat{e} \leftarrow E^\text{asym}(\text{PK}^\text{hy})
    $

    \If{$\hat{e} \neq e$}
        \State \Return $\bot$
    \EndIf

    \State $\hat{a} \leftarrow G(\hat{\sigma})$
    \State $\hat{m} \leftarrow D^\text{sym}_{\hat{a}}(c)$
    \State \Return $\hat{m}$
\end{algorithmic}
\end{algorithm}

\begin{theorem}\label{fo-security-theorem}
    For all IND-CCA adversary $\mathcal{A}^\text{hy}$ against the hybrid encryption scheme with advantage $\epsilon^\text{hy}$, there exists a one-way one-time-encryption adversary against the public-key encryption scheme with advantage $\epsilon^\text{asym}$ and an one-time indistinguishability adversary against the symmetric encryption scheme with advantage $\epsilon^\text{sym}$ such that

    \begin{equation*}
        \epsilon^\text{hy} \leq q_\text{H}\epsilon^\text{asym} + \epsilon^\text{sym} + q_\text{D}2^{-\gamma}
    \end{equation*}

    where $\gamma$ is the spread of the public-key encryption scheme, $q_\text{H}$ is the total number of hash queries, and $q_\text{D}$ is the total number of decryption queries
\end{theorem}

\subsection{Proof of security result}
Theorem \ref{fo-security-theorem} is proved using a sequence of games that involves $\mathcal{A}^\text{hy}$ as the main routine, and two games that involves $\mathcal{A}^\text{hy}$ as a sub-routine. The sequence of games is as follows:

\begin{enumerate}
    \item Game 0 is the standard IND-CCA2 game

    \item Game 1 is identical to game 0, except that the decryption oracle $\mathcal{O}_D$ is modified. Instead of using the true secret key $\text{SK}^\text{hy}$ to decrypt the query $(e_q, c_q)$, the decryption oracle checks the tape of hash function $H$ for the existence of hash query $(\sigma_H, c_H, h_H)$ such that $c_q = c_H$ and $e_q = E^\text{asym}(\text{PK}^\text{asym}, \sigma_H, h_H)$. If such a query exists, then $\mathcal{O}_D$ uses $\sigma_H$ to derive the symmetric key $a_q \leftarrow G(\sigma_H)$ and decrypt the queried ciphertext $c_q$. If no such query exists, then $\mathcal{O}_D$ will reject the queried ciphertext and output a decryption error. It is worth noting that this modified decryption oracle does not require the hybrid secret key $\text{SK}^\text{hy}$ to process decryption queries.

    \item Game 2 is identical to game 1, except the routine of encrypting the challenge ciphertext is modified: $a^\ast \leftsample \mathcal{K}^\text{sym}$ is randomly sampled from the symmetric key space instead of being queried from $G$, and $h^\ast \leftsample \text{COIN}^\text{asym}$ is randomly sampled from the asymmetric coin space instead of being queried from $H$.
\end{enumerate}

Let $S_0, S_1, S_2$ denote the event that $\mathcal{A}^\text{hy}$ wins game 0, game 1, and game 2, respectively.

\begin{lemma}\label{fo-win0-win1}
    Let $q_D$ denote the number of decryption queries, and $\gamma$ denote the spread of the PKE, then
    \begin{equation*}
        P[S_0] - P[S_1] \leq q_D 2^{-\gamma}
    \end{equation*}
\end{lemma}

\begin{proof}
    For each decryption query $(e_q, c_q)$, there are three mutually exclusive possibilities:

    \begin{enumerate}
        \item The queried ciphertext is \textbf{honest}, meaning that there is a matching record on the tape of the hash function $H$
        \item There is no matching record on the tape of the hash function $H$, and the check in step 4 in algorithm \ref{fo-key-dec} will fail, outputing decryption error. Such a query is called \textbf{invalid}
        \item There is no matching record on the tape of the hash function $H$, but the check in step 4 of algorithm \ref{fo-key-dec} will succeed. Such a query is called \textbf{almost valid}
    \end{enumerate}

    Observe that for both $S_0, S_1$:

    $$
    \begin{aligned}
        P[S] &= P[S \cap \text{ all decryption queries are honest }] \\
        &+ P[S \cap \text{ some decryption queries are dishonest, but none is almost valid }] \\
        &+ P[S \cap \text{ at least one almost valid decryption queries }]
    \end{aligned}
    $$

    When all decryption queries are honest, the decryption oracles will correctly decrypt the query in both game 0 and game 1. When all dishonest decryption queries are invalid, the decryption oracles will reject the query in both games. The only difference between the two games lies in how the decryption oracle processes almost valid decryption queries. Therefore:

    $$
    \begin{aligned}
        &P[S_0] - P[S_1] \\
        &= P[S_0 \cap \text{ at least one almost valid decryption queries }] \\
        &- P[S_1 \cap \text{ at least one almost valid decryption queries }] \\
        &\leq P[\text{ at least one almost valid decryption queries }]
    \end{aligned}
    $$

    Let $(e, c)$ be some decryption query made without querying $H$, then in the true decryption routine, $\hat{h} \leftarrow H(\hat{\sigma}, c)$ will be a truly random coin, and $\hat{e} \leftarrow E^\text{asym}(\hat{\sigma}, \hat{h})$ will be a truly random ciphertext for the given public key and $\hat{\sigma}$. Since the the PKE has $\gamma$ spread, we know that $P[e = \hat{e}] = P[(e, c) \text{ is almost valid}] \leq 2^{-\gamma}$. Among $q_D$ decryption query, the probability of having at least one almost valid query is bounded by sum of probability of each decryption query being almost valid: $P[\text{ at least one almost valid query }] \leq q_D 2^{-\gamma}$.
\end{proof}

% TODO: bound the difference between S_1 and S_2
% TODO: bound ask sigma
% TODO: bound S_2

\bibliographystyle{plain}
\bibliography{./references.bib}

\end{document}

### 2024-03-18 co789/project/manuscript/main.tex
Finished drafting FO transformation; started Hofheinz

@@
 
+\section{Introduction}
+
+\section{Preliminaries}
+
 \section{The Fujisaki-Okamoto Transformation}
@@
 
-    \Return $(\text{PK}^\text{hy}, \text{SK}^\text{hy})$
+    \State \Return $(\text{PK}^\text{hy}, \text{SK}^\text{hy})$
 \end{algorithmic}
@@
 
-    \Return $(e, c)$
+    \State \Return $(e, c)$
 \end{algorithmic}
@@
 
-\begin{enumerate}
+\begin{itemize}
     \item Game 0 is the standard IND-CCA2 game
@@
     \item Game 2 is identical to game 1, except the routine of encrypting the challenge ciphertext is modified: $a^\ast \leftsample \mathcal{K}^\text{sym}$ is randomly sampled from the symmetric key space instead of being queried from $G$, and $h^\ast \leftsample \text{COIN}^\text{asym}$ is randomly sampled from the asymmetric coin space instead of being queried from $H$.
-\end{enumerate}
+\end{itemize}
 
@@
 
-% TODO: bound the difference between S_1 and S_2
-% TODO: bound ask sigma
-% TODO: bound S_2
+If during the IND-CCA game, $\mathcal{A}^\text{hy}_\text{IND-CCA}$ never makes hash query that involves $\sigma^\ast$, then under the random oracle model, there is no difference between sampling $a^\ast, h^\ast$ randomly or pseudorandomly.
+On the other hand, if the adversary does make such a query, then there is an inconsistency between the challenge ciphertext and the results of such hash query.
+In other words, the difference between game 1 and game 2 is the even that the IND-CCA adversary makes such a hash query:
+
+\begin{lemma}
+    $$
+    P[S_1] - P[S_2] = P[\text{query}]
+    $$
+\end{lemma}
+
+Denote the challenge ciphertext in game 2 by $(e^\ast, c^\ast)$.
+Notice that $e^\ast$ is the encryption of a truly random $\sigma^\ast$ using a truly random coin $h^\ast$, while $c^\ast$ is the encryption of $m_b^\ast$ under a truly random key $a^\ast$.
+This allows a OW-CPA game for the PKE, and an IND-CPA game for the symmetric cipher to be perfectly simulated by some $\mathcal{A}^\text{asym}_\text{OW-CPA}$ and/or $\mathcal{A}^\text{sym}_\text{IND-CPA}$.
+Using this strategy, we can bound $P[S_2]$ and $P[\text{query}]$ using the advantage of some OW-CPA and/or IND-CPA adversaries.
+
+\begin{lemma}\label{fo-bound-s2}
+    For every IND-CCA adversary $A^\text{hy}_\text{IND-CCA}$, there exists an IND-OTE adversary against the underlying symmetric cipher $\mathcal{A}^\text{sym}_\text{IND-OTE}$ with advantage $\epsilon^\text{sym}_\text{IND-OTE}$ such that
+
+    $$
+    P[S_2] = \frac{1}{2} + \epsilon^\text{sym}_\text{IND-OTE}
+    $$
+\end{lemma}
+
+\begin{proof}
+    $\mathcal{A}^\text{sym}_\text{IND-OTE}$ can perfectly simulate the hybrid key generation $\text{PK}^\text{hy} = \text{PK}^\text{asym}, \text{SK}^\text{hy} = \text{SK}^\text{asym}$, the random oracles $G, H$, as well as the modified decryption oracle.
+
+    When $\mathcal{A}^\text{hy}_\text{IND-CCA}$ submits the challenge plaintexts $(m_0, m_1)$, $\mathcal{A}^\text{sym}_\text{IND-OTE}$ passes them to the symmetric cipher challenger and receives the symmetric challenge ciphertext $c^\ast$.
+$\mathcal{A}^\text{sym}_\text{IND-OTE}$ then randomly samples $\sigma^\ast \leftsample \mathcal{M}^\text{asym}$ and $h^\ast \leftsample \text{COIN}^\text{asym}$ and computes $e^\ast \leftarrow E^\text{asym}(\text{PK}^\text{asym}, \sigma^\ast, h^\ast)$.
+$(e^\ast, c^\ast)$ is given to $\mathcal{A}^\text{hy}_\text{IND-CCA}$ as the challenge ciphertext.
+It is easy to verify that from $\mathcal{A}^\text{hy}_\text{IND-CCA}$'s perspective, this game is exactly game 2; in addition $\mathcal{A}^\text{hy}_\text{IND-CCA}$ wins the game if and only if $\mathcal{A}^\text{sym}_\text{IND-OTE}$ wins the game.
+\end{proof}
+
+\begin{lemma}
+    For every IND-CCA adversray $\mathcal{A}^\text{hy}_\text{IND-CCA}$, there exists an OW-CPA adversary $\mathcal{A}^\text{asym}_\text{OW-CPA}$ with advantage $\epsilon^\text{asym}_\text{OW-CPA}$ such that
+
+    $$
+    \epsilon^\text{asym}_\text{OW-CPA} = P[\text{query}] \cdot \frac{1}{q_H}
+    $$
+
+    where $q_H$ is the number of hash queries $\mathcal{A}^\text{hy}_\text{IND-CCA}$ makes to either $H$ or $G$.
+\end{lemma}
+
+\begin{proof}
+    Similar to the proof of lemma \ref{fo-bound-s2}, $\mathcal{A}^\text{asym}_\text{OW-CPA}$ can perfectly simulate key generation, hash oracles, and decryption oracle.
+
+    Let $e^\ast$ denote the asymmetric challenge ciphertext $\mathcal{A}^\text{asym}_\text{OW-CPA}$ receives from the OW-CPA challenger, let $\sigma^\ast$ denote the plaintext that corresponds with $e^\ast$, and let $(m_0, m_1)$ denote the challenge plaintexts submitted by $\mathcal{A}^\text{hy}_\text{IND-CCA}$.
+$\mathcal{A}^\text{asym}_\text{OW-CPA}$ samples a random symmetric key $a^\ast \leftsample \mathcal{K}^\text{sym}$ and a coin flip $b \leftsample \{0, 1\}$, then computes $c^\ast \leftarrow E^\text{sym}_{a^\ast}(m_b)$.
+$(e^\ast, c^\ast)$ is the IND-CCA challenge ciphertext.
+
+    After the $\mathcal{A}^\text{hy}_\text{IND-CCA}$ halts, $\mathcal{A}^\text{asym}_\text{OW-CPA}$ looks through the tape of the hash oracles $\mathcal{O}^H = \{(\tilde{\sigma}, \tilde{c})\}$ and $\mathcal{O}^G = \{\tilde{\sigma}\}$, then picks a random value among all possible $\tilde{\sigma}$'s to output.
+$\mathcal{A}^\text{asym}_\text{OW-CPA}$ wins if $\mathcal{A}^\text{hy}_\text{OW-CPA}$ makes a query $\tilde{\sigma} = \sigma^\ast$ and the correct query is chosen.
+\end{proof}
+
+\section{Tweaking FO Transform for IND-CCA KEM}
+While the FO transformation can construct IND-CCA public-key encryption scheme from OW-CPA PKE and IND-OTE symmetric cipher, there are a few drawbacks to the hybrid scheme as it is:
+
+\begin{itemize}
+    \item The security proof is not tight. The security of the hybrid scheme degrades linearly with the number of hash queries.
+    \item The hybrid scheme assumes the underlying PKE to be always correct.
+This is not the case in PKEs such as schemes based on "Learning with Errors" (LWE), which can have a small but non-zero probability of decryption error
+\end{itemize}
+
+In 2017, Hofheinz et al \cite{hofheinz2017modular} improved the FO transformation by addressing the non-tight security and accounting for possible decryption errors.
+In addition, the paper proposed various IND-CCA KEM constructions from weak PKE's that does not require a symmetric cipher.
+In fact, both Kyber \cite{avanzi2019crystals} and Classic McEliece \cite{albrecht2022classic} make direct use of transformations proposed in this paper to achieve IND-CCA security for their respective KEMs.
+
+The strategy for achieving IND-CCA KEM consists of two parts
+
+\begin{itemize}
+    \item Transform a OW-CPA and/or IND-CPA PKE into an OW-PCVA PKE
+    \item Transform the OW-PCVA PKE into a IND-CCA KEM
+\end{itemize}
+
+\subsection{Preliminaries}
+
+\subsection{From OW-CPA to OW-PCVA}
+
+\subsection{IND-CCA KEM}
+
+\section{OAEP and RSA-OAEP}
 

### 2024-03-20 co789/project/manuscript/main.tex
Progress

@@
 \section{Preliminaries}
+% TODO: from IND-CPA to OW-CPA
 
@@
 
-\subsection{Preliminaries}
+The intermediary security definition PCVA is defined by two oracles: the \textbf{plaintext checking oracle (PCO)} and the \textbf{ciphertext validation oracle (CVO)}.
+The details of these two oracles will be discussed in the following section
 
-\subsection{From OW-CPA to OW-PCVA}
+\subsection{From OW/IND-CPA to OW-PCVA}
+Let $(\text{KeyGen}, E, D)$ denote the input public-key encryption scheme, let $(\text{KeyGen}^T, E^T, D^T)$ denote the transformed public-key encryption scheme, let $G$ be a hash function.
+The transformed routines are as follows:
+
+\begin{algorithm}
+    \caption[keygen-t]{$\text{KeyGen}^T$}
+    \begin{algorithmic}[1]
+        \State $\text{pk}, \text{sk} \leftsample \text{KeyGen}()$
+        \State \Return $\text{pk}^T \leftarrow \text{pk}, \text{sk}^T \leftarrow \text{sk}$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption[encrypt-t]{Encryption routine: $E^T$}
+    \begin{algorithmic}[1]
+        \Require The message $m \in \mathcal{M}$
+        \State $r \leftarrow G(m)$ \Comment $r$ serves as the coin
+        \State $c \leftarrow E(\text{pk}^T, m, r)$
+        \State \Return $c$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption[decrypt-t]{Decryption routine: $D^T$}
+    \begin{algorithmic}[1]
+        \Require ciphertext $c \in \mathcal{C}$
+        \State $\hat{m} \leftarrow D(\text{sk}, c)$
+        \State $\hat{r} \leftarrow G(\hat{m})$
+        \If{$E(\text{pk}, \hat{m}, \hat{r}) \neq c$}
+            \State \Return $\bot$
+        \EndIf
+        \State \Return $\hat{m}$
+    \end{algorithmic}
+\end{algorithm}
+
+To discuss the security of the transformed scheme, we first need to lay out the implementations of the PCO and the CVO.
+In the vanilla PCVA game, the two oracles have access to the challenge secret key and perform decryption routines:
+
+\begin{algorithm}
+    \caption[pco-0]{Plaintext checking oracle $\text{PCO}_0$}
+    \begin{algorithmic}[1]
+        \Require A query plaintext-ciphertext pair $(\tilde{m}, \tilde{c})$
+        \State $\hat{m} \leftarrow D(\text{sk}, \tilde{c})$
+        \If{$\hat{m} \neq \tilde{m}$}
+            \State \Return 0 \Comment query is not valid
+        \EndIf
+        \State $\tilde{r} \leftarrow G(\tilde{m})$
+        \If{$E(\text{pk}, \tilde{m}, \tilde{r}) \neq \tilde{c}$}
+            \State \Return 0  \Comment query is also not valid
+        \EndIf
+        \State \Return 1  \Comment query is valid
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption[cvo-0]{Ciphertext validation oracle $\text{CVO}_0$}
+    \begin{algorithmic}[1]
+        \Require A query ciphertext $\tilde{c}$
+        \State $\hat{m} \leftarrow D(\text{sk}, \tilde{c})$
+        \If{$\hat{m} = \bot$}
+            \State \Return 0 \Comment ciphertext is not valid
+        \ElsIf{$E(\text{pk}, \hat{m}, G(\hat{m})) \neq \tilde{c}$}
+            \State \Return 0 \Comment ciphertext is not valid
+        \EndIf
+        \State \Return 1  \Comment ciphertext is valid
+    \end{algorithmic}
+\end{algorithm}
+\pagebreak
+
+To link the advantage of an OW-PCVA adversary, a sequence of games is devised in which PCO and CVO are incrementally modified to not require the secret key and the challenge encryption to be performed with a truly random coin.
+These modifications will allow the OW-PCVA game to be perfectly simulated by an OW/IND-CPA adversary, thus establishing the security of the transformed encryption scheme upon the security of the input encryption scheme.
+
+\begin{theorem}[IND-CPA to OW-PCVA]\label{cpa-to-pcva}
+    Under the random-oracle model, for every OW-PCVA adversary $\mathcal{A}^T_\text{OW-PCVA}$ against the transformed PKE with advantage $\epsilon^T_\text{OW-PCVA}$, there exists an IND-CPA adversary $\mathcal{A}_\text{IND-CPA}$ against the input PKE with advantage $\epsilon_\text{IND-CPA}$ such that
+
+    $$
+    \epsilon^T_\text{OW-PCVA} 
+        \leq q_V \cdot 2^{-\gamma} 
+            + q_G \cdot \delta 
+            + 3\cdot\epsilon_\text{IND-CPA}
+    $$
+
+    Where $\gamma$ is the spread of ciphertext of the input PKE, $\delta$ is the probability of decryption error of the input PKE, $q_V$ is the number of ciphertext validation queries, and $q_G$ is the number of hash queries to $G$.
+\end{theorem}
+
+\begin{proof}\label{cpa-to-pcva-proof}
+    \textit{Admittedly this is only a sketch. Some nuance is omitted for the sake of clarity, but the overall result will be off by no more than some negligible constants.}
+
+    We device a sequence of games that $\mathcal{A}^T_\text{OW-PCVA}$ plays. Game 0 is the standard OW-PCVA game.
+
+    Game 1 is identical to game 0, except that CVO has been modified to check the tape of the hash function $G$ for plaintext $\tilde{m}$ that encrypts into the queried ciphertext $\tilde{c}$ under $E^T$:
+
+    \begin{algorithm}\label{cvo-1}
+        \caption{$\text{CVO}_1$}
+        \begin{algorithmic}[1]
+            \Require queried ciphertext $\tilde{c}$
+            \For{$(\tilde{m}, \tilde{r})$ on the tape of $G$}
+                \If{$E(\text{pk}, \tilde{m}, \tilde{r}) = \tilde{c}$} 
+                    \State \Return 1  \Comment queried ciphertext is valid
+                \EndIf
+            \EndFor
+            \State \Return 0  \Comment queried ciphertext is invalid
+        \end{algorithmic}
+    \end{algorithm}
+
+    Following similar logic in the FO transformation's security reduction, it's easy to see that game 0 and game 1 diverge when $\mathcal{A}^T_\text{OW-PCVA}$ makes at least one CVO query $\tilde{c}$ whose decryption encrypts correctly (aka passes CVO in game 0) but without querying $G$ (aka fails CVO in game 1)
+
+    \begin{equation}
+        P[\text{wins game 0}] - P[\text{wins game 1}] \leq q_V \cdot 2^{-\gamma}
+    \end{equation}
+
+    Game 2 is identical to game 1, except that PCO is modified to check only the encryption instead of checking both encryption and decryption
+
+    \begin{algorithm}
+        \caption{$\text{PCO}_2$}
+        \begin{algorithmic}[1]
+            \Require query $(\tilde{m}, \tilde{c})$
+            \For{$(m, r)$ from the tape of $G$}
+                \If{$m = \tilde{m}$ and $E(\text{pk}, m, r) = \tilde{c}$}
+                    \State \Return 1
+                \EndIf
+            \EndFor
+            \State \Return 0
+        \end{algorithmic}
+    \end{algorithm}\pagebreak
+
+    Game 2 diverges from game 1 if $\mathcal{A}^T_\text{OW-PCVA}$ makes at least one hash query $(\tilde{m}, \tilde{r})$ such that
+
+    $$D(\text{sk}, E(\text{pk}, \tilde{m}, \tilde{r})) \neq \tilde{m}$$
+
+    Therefore
+
+    $$P[\text{wins game 1}] - P[\text{wins game 2}] \leq q_G \cdot \delta$$
+
+    Game 3 is identical to game 2, except that when encrypting the challenge ciphertext $c^\ast$, a truly random coin $(r^\ast \leftsample \text{Coin})$ is used instead of a pseudorandom coin $r^\ast \leftarrow G(m^\ast)$.
+Game 3 diverges from game 2 if $\mathcal{A}^T_\text{OW-PCVA}$ ever queries the hash of the challenge plaintext $m^\ast$:
+
+    $$
+    P[\text{wins game 2}] - P[\text{wins game 3}] \leq P[\text{query $m^\ast$}]
+    $$
+
+    In game 3, PCO and CVO can be simulated without the secret key, and the challenge ciphertext is computed using a truly random coin instead of the hash of the challenge plaintext.
+This allows game 3 to be perfectly simulated by some OW-CPA adversary $\mathcal{A}_\text{OW-CPA}$ against the input PKE:
+
+    \begin{itemize}
+        \item When $\mathcal{A}_\text{OW-CPA}$ receives $\text{pk}$, the public key is passed to $\mathcal{A}^T_\text{OW-PCVA}$
+        \item $\mathcal{A}_\text{OW-CPA}$ simulates $G$, $\text{PCO}_2$, $\text{CVO}_1$ for $\mathcal{A}^T_\text{OW-PCVA}$
+        \item When $\mathcal{A}_\text{OW-CPA}$ receives the challenge ciphertext $c^\ast$, it is passed to $\mathcal{A}^T_\text{OW-PCVA}$ as the OW-PCVA challenge ciphertext
+        \item When $\mathcal{A}^T_\text{OW-PCVA}$ outputs $m$, $\mathcal{A}_\text{OW-CPA}$ outputs the same value
+    \end{itemize}
+
+    It's easy to see that from $\mathcal{A}^T_\text{OW-PCVA}$'s perspective, it is playing game 3, and $\mathcal{A}^T_\text{OW-PCVA}$ wins if and only if $\mathcal{A}_\text{OW-CPA}$ wins:
+
+    $$
+    P[\text{$\mathcal{A}^T_\text{OW-PCVA}$ wins game 3}] = \epsilon_\text{OW-CPA} = \epsilon_\text{IND-CPA} + \frac{1}{\vert\mathcal{M}\vert}
+    $$
+
+    % TODO: bound P[query] by P[IND-CPA wins]
+
+\end{proof}
 

### 2024-03-22 co789/project/manuscript/main.tex
A very rough draft

@@
             + q_G \cdot \delta 
+            + \frac{1}{\vert\mathcal{M}\vert}
             + 3\cdot\epsilon_\text{IND-CPA}
@@
 
-    \begin{equation}
+    \begin{equation}\label{ow-pcva-g0-g1}
         P[\text{wins game 0}] - P[\text{wins game 1}] \leq q_V \cdot 2^{-\gamma}
@@
 
-    $$P[\text{wins game 1}] - P[\text{wins game 2}] \leq q_G \cdot \delta$$
+    \begin{equation}\label{ow-pcva-g1-g2}
+        P[\text{wins game 1}] - P[\text{wins game 2}] \leq q_G \cdot \delta  
+    \end{equation}
 
@@
 
-    $$
+    \begin{equation}\label{ow-pcva-g2-g3}
         P[\text{wins game 2}] - P[\text{wins game 3}] \leq P[\text{query $m^\ast$}]
-    $$
+    \end{equation}
 
@@
 
+    \begin{equation}\label{ow-pcva-g3}
+        P[\text{$\mathcal{A}^T_\text{OW-PCVA}$ wins game 3}] 
+        = \epsilon_\text{OW-CPA} 
+        = \epsilon_\text{IND-CPA} + \frac{1}{\vert\mathcal{M}\vert}
+    \end{equation}
+
+    % TODO: this "well-known results" probably needs some explanation
+    Note that $\epsilon_\text{OW-CPA} = \epsilon_\text{IND-CPA} + \frac{1}{\vert\mathcal{M}\vert}$ is a well-known results
+
+    The final piece of this puzzle is to bound the probability of the event that the $\mathcal{A}^T_\text{OW-PCVA}$ queries the hash of the challenge plaintext $m^\ast$.
+Similar to the argument above, game 3 can be perfectly simulated by an IND-CPA adversary $\mathcal{A}_\text{IND-CPA}$ against the underlying PKE:
+
+    \begin{itemize}
+        \item The IND-CPA $\text{pk}$ is directly passed to $\mathcal{A}^T_\text{OW-PCVA}$
+        \item $\mathcal{A}_\text{IND-CPA}$ simulates $G$, $\text{PCO}_2$, $\text{CVO}_1$ for $\mathcal{A}^T_\text{OW-PCVA}$
+        \item $\mathcal{A}_\text{IND-CPA}$ submits randomly sampled messages $(m_0, m_1)$ as the challenge plaintexts.
+Upon receiving the IND-CPA challenge ciphertext $c^\ast$, $\mathcal{A}_\text{IND-CPA}$ passes $c^\ast$ to $\mathcal{A}^T_\text{OW-PCVA}$ as the OW-PCVA challenge ciphertext
+        \item After $\mathcal{A}^T_\text{OW-PCVA}$ halts, $\mathcal{A}_\text{IND-CPA}$ checks the tape of the hash function $G$.
+If either $m_0$ or $m_1$ is on the tape, then it is returned; if neither is on the tape, then $\mathcal{A}_\text{IND-CPA}$ randomly returns 0 or 1
+    \end{itemize}
+
+    Intuitively we can argue the following: if $\mathcal{A}^T_\text{OW-PCVA}$ indeed queries $m^\ast$, then $\mathcal{A}_\text{IND-CPA}$ wins with probability 1; otherwise, $\mathcal{A}_\text{IND-CPA}$ wins with probability $\frac{1}{2}$:
+
     $$
-    P[\text{$\mathcal{A}^T_\text{OW-PCVA}$ wins game 3}] = \epsilon_\text{OW-CPA} = \epsilon_\text{IND-CPA} + \frac{1}{\vert\mathcal{M}\vert}
+    P[\text{IND-CPA adversary wins}] 
+        = P[\text{query $m^\ast$}] 
+            + (1 - P[\text{query $m^\ast$}]) \cdot \frac{1}{2}
     $$
 
-    % TODO: bound P[query] by P[IND-CPA wins]
+    Which transforms into
+
+    \begin{equation}\label{ow-pcva-query}
+        P[\text{query $m^\ast$}] = 2\epsilon_\text{IND-CPA}
+    \end{equation}
+
+    Putting equations \ref{ow-pcva-g0-g1}, \ref{ow-pcva-g1-g2}, \ref{ow-pcva-g2-g3}, \ref{ow-pcva-g3}, \ref{ow-pcva-query} together gives us the result
 
@@
 \subsection{IND-CCA KEM}
+Let $H$ be a hash function, $(\text{KeyGen}, E, D)$ be an IND-CPA public-key encryption scheme, and $(E^T, D^T)$ be the OW-PCVA public-key encryption obtained by running the IND-CPA PKE through the transformation described above.
+We first present the KEM, then discuss its security properties
+
+The KEM's key generation routine is identical to the key generation routine of the IND-CPA PKE.
+
+\begin{algorithm}
+    \caption{Encapsulate}
+    \begin{algorithmic}[1]
+        \Require $\text{pk}$
+        \State Sample random $m \leftsample \mathcal{M}$
+        \State $c \leftarrow E^T(\text{pk}, m)$ \Comment the ciphertext
+        \State $K \leftarrow H(m, c)$ \Comment the shared secret
+        \State \Return $(c, K)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{Decapsulate}
+    \begin{algorithmic}[1]
+        \Require $\text{sk}$, the ciphertext $c$
+        \State $m \leftarrow D^T(\text{sk}, c)$
+        \If{$m = \bot$}
+            \State \Return $\bot$
+        \EndIf
+        \State \Return $H(m, c)$
+    \end{algorithmic}
+\end{algorithm}
+
+We denote the KEM by its routines $(\text{KeyGen}, E^{U^\bot}, D^{U^\bot})$
+
+\begin{theorem}\label{ind-cca-kem-adv}
+    For every IND-CCA adversary $\mathcal{A}^{U^\bot}_\text{IND-CCA}$ against the KEM $(\text{KeyGen}, E^{U^\bot}, D^{U^\bot})$ with advantage $\epsilon^{U^\bot}_\text{IND-CCA}$, there exists an OW-PCVA adversary $\mathcal{A}^T_\text{OW-PCVA}$ against the transformed PKE $(E^T, D^T)$ with advantage $\epsilon^{T}_\text{OW-PCVA}$ such that
+
+    \begin{equation}
+        \epsilon^{U^\bot}_\text{IND-CCA} \leq \epsilon^{T}_\text{OW-PCVA}
+    \end{equation}
+\end{theorem}
+
+This security result is obtained through a sequence of games.
+When we start at the IND-CCA game, there is a decapsulation oracle that uses the true secret key, but through the sequence of games, the decapsulation oracle is modified to not require the secret key, so it can be simulated by an OW-PCVA adversary.
+
+\begin{proof}\label{ind-cca-kem-proof}
+    Game 0 is the KEM IND-CCA game
+
+    Game 1 is identical to game 0, but we modify both the hash oracle $H$ and the decapsulation oracle $\mathcal{O}^D$ to remove the use of the secret key.
+
+    \begin{algorithm}
+        \caption{$H_1$, the hash oracle in game 1}
+        \begin{algorithmic}[1]
+            \Require input $\tilde{m}, \tilde{c}$
+            \Require plaintext checking oracle PCO
+            \If{$\exists (m, c, K) \in \text{tape}_H$ such that $m=\tilde{m}, c=\tilde{c}$}
+                \State \Return $K$
+            \EndIf
+            \State $\tilde{K} \leftsample \{0, 1\}^n$
+            \If{PCO validates $\tilde{m}, \tilde{c}$}
+                \State Add $(\tilde{c}, \tilde{K})$ to the tape of $\mathcal{O}^D$
+            \EndIf
+            \State Add $(\tilde{m}, \tilde{c}, \tilde{K})$ to $\text{tape}_H$
+            \State \Return $\tilde{K}$
+        \end{algorithmic}
+    \end{algorithm}
+
+    \begin{algorithm}
+        \caption{$\mathcal{O}^D_1$, the decapsulation oracle in game 1}
+        \begin{algorithmic}[1]
+            \Require the queried ciphertext $\tilde{c}$
+            \Require a ciphertext validity oracle CVO
+            \If{CVO fails to validate $\tilde{c}$}
+                \State \Return $\bot$
+            \EndIf
+            \If{$\exists (c, K)$ on $\mathcal{O}^D$'s tape such that $c = \tilde{c}$}
+                \State \Return $K$
+            \EndIf
+            \State Sample $\tilde{K} \leftsample \{0, 1\}^n$ 
+            \State Add $(\tilde{c}, \tilde{K})$ to the tape of $\mathcal{O}^D$
+            \State \Return $\tilde{K}$
+        \end{algorithmic}
+    \end{algorithm}
+    \pagebreak
+
+    Intuitively, we can argue that the PCO and the CVO ensures the integrity of the ciphertext, and that the hash tape and the decapsulation tape ensure the consistency of the queries.
+Therefore, from the adversary's point of view, game 1 and game 0 are perfectly identical.
+
+    \begin{equation}\label{kem-g0-g1}
+        P[S_0] - P[S_1] = 0 
+    \end{equation}
+
+    Game 2 is identical to game 1, except that when queried on the challenge plaintext/ciphertext pair $(m^\ast, c^\ast)$ by the adversary, $H$ consistently returns a separately randomly sampled value $K$.
+Game 2 diverges from game 1 when $\mathcal{A}^{U^\bot}_\text{IND-CCA}$ makes such a query
+
+    \begin{equation}\label{kem-g1-g2}
+        P[S_1] - P[S_2] \leq P[\text{query}]
+    \end{equation}
+
+    In game 2, since $H(m^\ast, c^\ast)$ returns truly random value instead of pseudorandom value, the KEM adversary can obtain no information that correlates with $K^\ast$.
+In other words, no adversary can have any advantage in game 2:
+
+    \begin{equation}\label{kem-g2}
+        P[S_2] = \frac{1}{2}
+    \end{equation}
+
+    On the other hand, $\mathcal{A}^{U^\bot}_\text{IND-CCA}$ making query to $m^\ast$ can be taken advantage of by the OW-PCVA adversary to recover $m^\ast$ in the OW-PCVA game:
+
+    \begin{itemize}
+        \item The OW-PCVA public key is passed directly to $\mathcal{A}^{U^\bot}_\text{IND-CCA}$
+        \item The hash function $H$ and the decapsulation oracle (implemented using PCO and CVO) can all be perfectly simulated by $\mathcal{A}^T_\text{OW-PCVA}$
+        \item When $\mathcal{A}^T_\text{OW-PCVA}$ receives the OW-PCVA challenge ciphertext $c^\ast$, it samples a random key $K^\ast \leftsample \{0, 1\}^n$ and passes $c^\ast, K^\ast$ to $\mathcal{A}^{U^\bot}_\text{IND-CCA}$
+        \item When $\mathcal{A}^{U^\bot}_\text{IND-CCA}$ halts, $\mathcal{A}^T_\text{OW-PCVA}$ checks the tape of $H$.
+If there exists $(\tilde{m}, \tilde{c})$ such that $\tilde{c} = c^\ast$, then $\mathcal{A}^T_\text{OW-PCVA}$ returns $\tilde{m}$.
+If no such query exists, a random value is returned
+    \end{itemize}
+
+    Using the simulation above, we know that if $\mathcal{A}^{U^\bot}_\text{IND-CCA}$ makes the query $(m^\ast, c^\ast)$, then $\mathcal{A}^T_\text{OW-PCVA}$ will win its game.
+Therefore, the probability of "query" cannot be more than the advantage of the OW-PCVA adversary:
+
+    \begin{equation}\label{kem-query}
+        P[\text{query}] \leq \epsilon^T_\text{OW-PCVA}
+    \end{equation}
+
+    Combining equations \ref{kem-g0-g1}, \ref{kem-g1-g2}, \ref{kem-g2}, \ref{kem-query} provides the desired result.
+\end{proof}
 

### 2024-03-30 co789/homework-05/q1/main.tex
Drafted solution to problem 1, 2, 4

--- 2024-03-18 co789/project/manuscript/main.tex
@@
 % Algorithms and pseudocode
-\usepackage{algorithm}
+\usepackage[ruled,vlined]{algorithm2e}
 \usepackage{algpseudocode}
@@
 \newcommand{\norm}[1]{\Vert {#1} \Vert}
+\newcommand{\metric}[1]{\vert {#1} \vert}
 \newcommand{\var}[1]{\operatorname{Var}[{#1}]}
-\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
 
@@
 %    Corollaries are numbered within the theorem, though they are rarely used
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
+\newtheorem{definition}{Definition}
+\newtheorem{theorem}{Theorem}
 \newtheorem*{remark}{Remark}
@@
 
-\title{A survey of IND-CCA constructions}
+\title{CO 789, Homework 1}
 \author{Ganyu (Bruce) Xu (g66xu)}
-\date{CO 789, Winter 2024}
+\date{Fall 2023}
 
@@
 %%%% TITLE %%%%%
-\maketitle
+% \maketitle
 
-\section{Introduction}
+\section*{Question 1}
 
-\section{Preliminaries}
+We will prove $\metric{c_1 - (c_2 + e)} \geq \metric{c_1 - c_2} - \metric{e}$ by proving a more general result:
 
-\section{The Fujisaki-Okamoto Transformation}
-The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} takes a IND-CPA public-key encryption scheme (PKE) and a few other cryptographic primitives of weaker security as inputs, and constructs a hybrid encryption scheme that achieves IND-CCA2 security.
-
-In the conference paper, the security reduction is performed under the random oracle model, and the security of the hybrid scheme degrades linearly with the number of hash queries and the number of decryption queries.
-On the other hand, this scheme is desirable for its simplicity, and the low requirements for the input primitives: the PKE only needs to be one-time one-way secure, and the symetric cipher only needs to be indistinguishable under one-time attack.
-
-% TODO: Need more introduction?
-
-\subsection{The hybrid scheme and security result}
-The hybrid encryption scheme contains three routines: key generation, encryption, and decryption.
-The input for the hybrid scheme includes a public-key encryption scheme $\text{KeyGen}^\text{asym}, E^\text{asym}, D^\text{asym}$, a symmetric encryption scheme $E^\text{sym}, D^\text{sym}$, and two hash functions $G: \mathcal{M}^\text{asym} \mapsto \mathcal{K}^\text{sym}, H: \mathcal{M}^\text{asym} \times \mathcal{C}^\text{sym} \mapsto \text{Coin}^\text{asym}$.
-
-\begin{algorithm}
-\caption{FO Key generation}\label{fo-key-gen}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \State $
-        (\text{PK}^\text{asym}, \text{SK}^\text{asym}) 
-        \leftarrow \text{PKE.KeyGen}()
-    $
-    \State $
-        \text{PK}^\text{hy} \leftarrow \text{PK}^\text{asym},
-        \text{SK}^\text{hy} \leftarrow \text{SK}^\text{asym},
-    $
-
-    \State \Return $(\text{PK}^\text{hy}, \text{SK}^\text{hy})$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{FO Key encryption}\label{fo-key-enc}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \Require $m \in \mathcal{M}^\text{sym}$
-    \State Sample from the PKE's message space $\sigma \leftsample \mathcal{M}^\text{asym}$
-    \State $a \leftarrow G(\sigma)$, $c \leftarrow E^\text{sym}_a(m)$
-    \State $h \leftarrow H(\sigma, c)$
-    \State $e \leftarrow E^\text{asym}(\text{PK}^\text{hy}, \sigma, h)$
-
-    \State \Return $(e, c)$
-\end{algorithmic}
-\end{algorithm}
-
-\pagebreak
-
-The decryption routine:
-
-\begin{algorithm}
-\caption{FO Key decryption}\label{fo-key-dec}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \Require The ciphertext $(e, c)$
-
-    \State $
-        \hat{\sigma} \leftarrow D^\text{asym}(\text{SK}^\text{hy}, e)
-    $
-
-    \State $
-        \hat{h} \leftarrow H(\hat{\sigma}, c)
-    $
-
-    \State $
-        \hat{e} \leftarrow E^\text{asym}(\text{PK}^\text{hy})
-    $
-
-    \If{$\hat{e} \neq e$}
-        \State \Return $\bot$
-    \EndIf
-
-    \State $\hat{a} \leftarrow G(\hat{\sigma})$
-    \State $\hat{m} \leftarrow D^\text{sym}_{\hat{a}}(c)$
-    \State \Return $\hat{m}$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{theorem}\label{fo-security-theorem}
-    For all IND-CCA adversary $\mathcal{A}^\text{hy}$ against the hybrid encryption scheme with advantage $\epsilon^\text{hy}$, there exists a one-way one-time-encryption adversary against the public-key encryption scheme with advantage $\epsilon^\text{asym}$ and an one-time indistinguishability adversary against the symmetric encryption scheme with advantage $\epsilon^\text{sym}$ such that
+\begin{theorem}[variation of triangle inequality]
+    let $\vert \cdot \vert$ be a metric, then for $a, b$ in the vector space:
 
     \begin{equation*}
-        \epsilon^\text{hy} \leq q_\text{H}\epsilon^\text{asym} + \epsilon^\text{sym} + q_\text{D}2^{-\gamma}
+        \metric{a - b} \geq \metric{a} - \metric{b}
+    \end{equation*}
+\end{theorem}
+
+The desired result is a special case of the theorem above where $a = c_1 - c_2$ and $b = e$
+
+\begin{proof}
+    By the triangle inequality we know that $\metric{a + b} \leq \metric{a} + \metric{b}$. Let $c = a + b$, then $a = c - b$, and the triangle inequality becomes
+
+    \begin{equation*}
+        \metric{c} \leq \metric{c - b} + \metric{b}
     \end{equation*}
 
-    where $\gamma$ is the spread of the public-key encryption scheme, $q_\text{H}$ is the total number of hash queries, and $q_\text{D}$ is the total number of decryption queries
-\end{theorem}
+    Re-arranging the equation above gives us the following:
 
-\subsection{Proof of security result}
-Theorem \ref{fo-security-theorem} is proved using a sequence of games that involves $\mathcal{A}^\text{hy}$ as the main routine, and two games that involves $\mathcal{A}^\text{hy}$ as a sub-routine.
-The sequence of games is as follows:
+    \begin{equation*}
+        \metric{c - b} \geq \metric{c} - \metric{b}
+    \end{equation*}
 
-\begin{itemize}
-    \item Game 0 is the standard IND-CCA2 game
-
-    \item Game 1 is identical to game 0, except that the decryption oracle $\mathcal{O}_D$ is modified.
-Instead of using the true secret key $\text{SK}^\text{hy}$ to decrypt the query $(e_q, c_q)$, the decryption oracle checks the tape of hash function $H$ for the existence of hash query $(\sigma_H, c_H, h_H)$ such that $c_q = c_H$ and $e_q = E^\text{asym}(\text{PK}^\text{asym}, \sigma_H, h_H)$.
-If such a query exists, then $\mathcal{O}_D$ uses $\sigma_H$ to derive the symmetric key $a_q \leftarrow G(\sigma_H)$ and decrypt the queried ciphertext $c_q$.
-If no such query exists, then $\mathcal{O}_D$ will reject the queried ciphertext and output a decryption error.
-It is worth noting that this modified decryption oracle does not require the hybrid secret key $\text{SK}^\text{hy}$ to process decryption queries.
-
-    \item Game 2 is identical to game 1, except the routine of encrypting the challenge ciphertext is modified: $a^\ast \leftsample \mathcal{K}^\text{sym}$ is randomly sampled from the symmetric key space instead of being queried from $G$, and $h^\ast \leftsample \text{COIN}^\text{asym}$ is randomly sampled from the asymmetric coin space instead of being queried from $H$.
-\end{itemize}
-
-Let $S_0, S_1, S_2$ denote the event that $\mathcal{A}^\text{hy}$ wins game 0, game 1, and game 2, respectively.
-
-\begin{lemma}\label{fo-win0-win1}
-    Let $q_D$ denote the number of decryption queries, and $\gamma$ denote the spread of the PKE, then
-    \begin{equation*}
-        P[S_0] - P[S_1] \leq q_D 2^{-\gamma}
-    \end{equation*}
-\end{lemma}
-
-\begin{proof}
-    For each decryption query $(e_q, c_q)$, there are three mutually exclusive possibilities:
-
-    \begin{enumerate}
-        \item The queried ciphertext is \textbf{honest}, meaning that there is a matching record on the tape of the hash function $H$
-        \item There is no matching record on the tape of the hash function $H$, and the check in step 4 in algorithm \ref{fo-key-dec} will fail, outputing decryption error.
-Such a query is called \textbf{invalid}
-        \item There is no matching record on the tape of the hash function $H$, but the check in step 4 of algorithm \ref{fo-key-dec} will succeed. Such a query is called \textbf{almost valid}
-    \end{enumerate}
-
-    Observe that for both $S_0, S_1$:
-
-    $$
-    \begin{aligned}
-        P[S] &= P[S \cap \text{ all decryption queries are honest }] \\
-        &+ P[S \cap \text{ some decryption queries are dishonest, but none is almost valid }] \\
-        &+ P[S \cap \text{ at least one almost valid decryption queries }]
-    \end{aligned}
-    $$
-
-    When all decryption queries are honest, the decryption oracles will correctly decrypt the query in both game 0 and game 1.
-When all dishonest decryption queries are invalid, the decryption oracles will reject the query in both games.
-The only difference between the two games lies in how the decryption oracle processes almost valid decryption queries.
-Therefore:
-
-    $$
-    \begin{aligned}
-        &P[S_0] - P[S_1] \\
-        &= P[S_0 \cap \text{ at least one almost valid decryption queries }] \\
-        &- P[S_1 \cap \text{ at least one almost valid decryption queries }] \\
-        &\leq P[\text{ at least one almost valid decryption queries }]
-    \end{aligned}
-    $$
-
-    Let $(e, c)$ be some decryption query made without querying $H$, then in the true decryption routine, $\hat{h} \leftarrow H(\hat{\sigma}, c)$ will be a truly random coin, and $\hat{e} \leftarrow E^\text{asym}(\hat{\sigma}, \hat{h})$ will be a truly random ciphertext for the given public key and $\hat{\sigma}$.
-Since the the PKE has $\gamma$ spread, we know that $P[e = \hat{e}] = P[(e, c) \text{ is almost valid}] \leq 2^{-\gamma}$.
-Among $q_D$ decryption query, the probability of having at least one almost valid query is bounded by sum of probability of each decryption query being almost valid: $P[\text{ at least one almost valid query }] \leq q_D 2^{-\gamma}$.
+    We made no assumption of $a$, $b$, or $c$, so the inequality above holds for all vectors $b, c$ in the vector space
 \end{proof}
 
-If during the IND-CCA game, $\mathcal{A}^\text{hy}_\text{IND-CCA}$ never makes hash query that involves $\sigma^\ast$, then under the random oracle model, there is no difference between sampling $a^\ast, h^\ast$ randomly or pseudorandomly.
-On the other hand, if the adversary does make such a query, then there is an inconsistency between the challenge ciphertext and the results of such hash query.
-In other words, the difference between game 1 and game 2 is the even that the IND-CCA adversary makes such a hash query:
-
-\begin{lemma}
-    $$
-    P[S_1] - P[S_2] = P[\text{query}]
-    $$
-\end{lemma}
-
-Denote the challenge ciphertext in game 2 by $(e^\ast, c^\ast)$.
-Notice that $e^\ast$ is the encryption of a truly random $\sigma^\ast$ using a truly random coin $h^\ast$, while $c^\ast$ is the encryption of $m_b^\ast$ under a truly random key $a^\ast$.
-This allows a OW-CPA game for the PKE, and an IND-CPA game for the symmetric cipher to be perfectly simulated by some $\mathcal{A}^\text{asym}_\text{OW-CPA}$ and/or $\mathcal{A}^\text{sym}_\text{IND-CPA}$.
-Using this strategy, we can bound $P[S_2]$ and $P[\text{query}]$ using the advantage of some OW-CPA and/or IND-CPA adversaries.
-
-\begin{lemma}\label{fo-bound-s2}
-    For every IND-CCA adversary $A^\text{hy}_\text{IND-CCA}$, there exists an IND-OTE adversary against the underlying symmetric cipher $\mathcal{A}^\text{sym}_\text{IND-OTE}$ with advantage $\epsilon^\text{sym}_\text{IND-OTE}$ such that
-
-    $$
-    P[S_2] = \frac{1}{2} + \epsilon^\text{sym}_\text{IND-OTE}
-    $$
-\end{lemma}
-
-\begin{proof}
-    $\mathcal{A}^\text{sym}_\text{IND-OTE}$ can perfectly simulate the hybrid key generation $\text{PK}^\text{hy} = \text{PK}^\text{asym}, \text{SK}^\text{hy} = \text{SK}^\text{asym}$, the random oracles $G, H$, as well as the modified decryption oracle.
-
-    When $\mathcal{A}^\text{hy}_\text{IND-CCA}$ submits the challenge plaintexts $(m_0, m_1)$, $\mathcal{A}^\text{sym}_\text{IND-OTE}$ passes them to the symmetric cipher challenger and receives the symmetric challenge ciphertext $c^\ast$.
-$\mathcal{A}^\text{sym}_\text{IND-OTE}$ then randomly samples $\sigma^\ast \leftsample \mathcal{M}^\text{asym}$ and $h^\ast \leftsample \text{COIN}^\text{asym}$ and computes $e^\ast \leftarrow E^\text{asym}(\text{PK}^\text{asym}, \sigma^\ast, h^\ast)$.
-$(e^\ast, c^\ast)$ is given to $\mathcal{A}^\text{hy}_\text{IND-CCA}$ as the challenge ciphertext.
-It is easy to verify that from $\mathcal{A}^\text{hy}_\text{IND-CCA}$'s perspective, this game is exactly game 2; in addition $\mathcal{A}^\text{hy}_\text{IND-CCA}$ wins the game if and only if $\mathcal{A}^\text{sym}_\text{IND-OTE}$ wins the game.
-\end{proof}
-
-\begin{lemma}
-    For every IND-CCA adversray $\mathcal{A}^\text{hy}_\text{IND-CCA}$, there exists an OW-CPA adversary $\mathcal{A}^\text{asym}_\text{OW-CPA}$ with advantage $\epsilon^\text{asym}_\text{OW-CPA}$ such that
-
-    $$
-    \epsilon^\text{asym}_\text{OW-CPA} = P[\text{query}] \cdot \frac{1}{q_H}
-    $$
-
-    where $q_H$ is the number of hash queries $\mathcal{A}^\text{hy}_\text{IND-CCA}$ makes to either $H$ or $G$.
-\end{lemma}
-
-\begin{proof}
-    Similar to the proof of lemma \ref{fo-bound-s2}, $\mathcal{A}^\text{asym}_\text{OW-CPA}$ can perfectly simulate key generation, hash oracles, and decryption oracle.
-
-    Let $e^\ast$ denote the asymmetric challenge ciphertext $\mathcal{A}^\text{asym}_\text{OW-CPA}$ receives from the OW-CPA challenger, let $\sigma^\ast$ denote the plaintext that corresponds with $e^\ast$, and let $(m_0, m_1)$ denote the challenge plaintexts submitted by $\mathcal{A}^\text{hy}_\text{IND-CCA}$.
-$\mathcal{A}^\text{asym}_\text{OW-CPA}$ samples a random symmetric key $a^\ast \leftsample \mathcal{K}^\text{sym}$ and a coin flip $b \leftsample \{0, 1\}$, then computes $c^\ast \leftarrow E^\text{sym}_{a^\ast}(m_b)$.
-$(e^\ast, c^\ast)$ is the IND-CCA challenge ciphertext.
-
-    After the $\mathcal{A}^\text{hy}_\text{IND-CCA}$ halts, $\mathcal{A}^\text{asym}_\text{OW-CPA}$ looks through the tape of the hash oracles $\mathcal{O}^H = \{(\tilde{\sigma}, \tilde{c})\}$ and $\mathcal{O}^G = \{\tilde{\sigma}\}$, then picks a random value among all possible $\tilde{\sigma}$'s to output.
-$\mathcal{A}^\text{asym}_\text{OW-CPA}$ wins if $\mathcal{A}^\text{hy}_\text{OW-CPA}$ makes a query $\tilde{\sigma} = \sigma^\ast$ and the correct query is chosen.
-\end{proof}
-
-\section{Tweaking FO Transform for IND-CCA KEM}
-While the FO transformation can construct IND-CCA public-key encryption scheme from OW-CPA PKE and IND-OTE symmetric cipher, there are a few drawbacks to the hybrid scheme as it is:
-
-\begin{itemize}
-    \item The security proof is not tight. The security of the hybrid scheme degrades linearly with the number of hash queries.
-    \item The hybrid scheme assumes the underlying PKE to be always correct.
-This is not the case in PKEs such as schemes based on "Learning with Errors" (LWE), which can have a small but non-zero probability of decryption error
-\end{itemize}
-
-In 2017, Hofheinz et al \cite{hofheinz2017modular} improved the FO transformation by addressing the non-tight security and accounting for possible decryption errors.
-In addition, the paper proposed various IND-CCA KEM constructions from weak PKE's that does not require a symmetric cipher.
-In fact, both Kyber \cite{avanzi2019crystals} and Classic McEliece \cite{albrecht2022classic} make direct use of transformations proposed in this paper to achieve IND-CCA security for their respective KEMs.
-
-The strategy for achieving IND-CCA KEM consists of two parts
-
-\begin{itemize}
-    \item Transform a OW-CPA and/or IND-CPA PKE into an OW-PCVA PKE
-    \item Transform the OW-PCVA PKE into a IND-CCA KEM
-\end{itemize}
-
-\subsection{Preliminaries}
-
-\subsection{From OW-CPA to OW-PCVA}
-
-\subsection{IND-CCA KEM}
-
-\section{OAEP and RSA-OAEP}
-
-\bibliographystyle{plain}
-\bibliography{./references.bib}
-
 \end{document}

### 2024-03-30 co789/homework-05/q2/main.tex
Drafted solution to problem 1, 2, 4

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 5}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Winter 2024}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 2}

\subsection*{(1)}
Here is the decryption algorithm

\begin{algorithm}
    \caption{Compressed McEliece decryption}
    \KwIn{$c \in \mathbb{F}_2^{n - l}$, $\text{sk} = (P, \mathcal{C}, S)$}

    Pad the low-order bits of $c$ with 0's such that $
        c^\prime = [c, 0, 0, \ldots, 0] \in \mathbb{F}^n
    $\;
    $c^{\prime\prime} \leftarrow P^{-1}c^\prime$\;
    $\hat{m} \leftarrow \mathcal{C}.\text{decode}(c^{\prime\prime})$\;
    $\hat{m} \leftarrow S^{-1}\hat{m}$\;
    \Return{$\hat{m}$}\;
\end{algorithm}

Among the high-order bits of $c^{\prime\prime}$, there are exactly $t - l$ bits of error, introduced by the sampled error term $e$. Among the low-order $l$ bits of $c^{\prime\prime}$, there are up to $l$ bits of errors since the padded 0's are blind guesses. Therefore, there are up to $t$ bits of errors in $c^{\prime\prime}$, and since $t \leq \frac{d-1}{2}$, the $(n, k, d)$-code is guaranteed to correct the error and recover the true $m$.

\subsection*{(2)}
The vanilla McEliece encryption scheme outputs the entire (partially corrupted codeword) $c$, which takes $n$ bits. This compressed scheme discarded the low-order $l$ bits, so the ciphertext takes $n - l$ bits, which saves space by a factor of $\frac{l}{n}$

\subsection*{(3)}
We already know that the McEliece encryption scheme is not IND-CPA. Instead, we will estimate the difficulty of breaking the encryption scheme by computing the probability of correctly decrypting some ciphertext without using the secret key. Specifically, since $c = Am + e$ where $A \in \mathbb{F}^{n \times k}$ is an overdetermined linear system, if the adversary can recover $e$, then it can recover $m$.

In the un-compressed McEliece scheme, the encryption routine corrupts exactly $t$ out of $n$ bits. There are a total of $\binom{n}{t}$ possible error terms to choose from. Without knowing additional information, the adversary can do no better than a blind guess:

\begin{equation*}
    \epsilon_0 = \frac{1}{\binom{n}{t}}
\end{equation*}

In the compressed scheme, the encryption routine corrupts $t-l$ out of the high-order $n-l$ bits. The low-order $l$ bits are blind guesses. There are a total of $\binom{n-l}{t-l} \cdot 2^l$ possible error values to choose from:

\begin{equation*}
    \epsilon_1 = \frac{1}{\binom{n-l}{t-l} \cdot 2^l}
\end{equation*}

Observe that

$$
\begin{aligned}
    \frac{\epsilon_0}{\epsilon_1} 
    &= \frac{\binom{n-l}{t-l} \cdot 2^l}{\binom{n}{t}} \\
    &= \frac{
        t \cdot (t-1) \cdot \ldots \cdot (t - l + 1)
    }{
        n \cdot (n-1) \cdot \ldots \cdot (n - l + 1)
    } \cdot 2^l
\end{aligned}
$$

Since $t = \frac{d-1}{2} \leq \frac{d}{2} < \frac{n}{2}$, we know $\frac{t}{n} \leq \frac{1}{2}$. Furthermore, we claim without proof that if $0 < a < b$ then $\frac{a}{b} \geq \frac{a-1}{b-1}$, which means

$$
\frac{1}{2} > \frac{t}{n} > \frac{t-1}{n-1} > \ldots > \frac{t-l+1}{n-l+1}
$$

Therefore:

$$
\frac{\epsilon_0}{\epsilon_1} = \frac{
    t \cdot (t-1) \cdot \ldots \cdot (t - l + 1)
}{
    n \cdot (n-1) \cdot \ldots \cdot (n - l + 1)
} \cdot 2^l \leq 1
$$

This means that a blindly-guessing adversary against the compressed scheme has higher advantage than a blindly-guessing adversary against the vanilla scheme. In other words, the compressed scheme is less secure.

\subsection*{(4)}
For a linear amount of reduction in ciphertext size, we lose an exponential amount of security. This tradeoff is not worth it.
\end{document}

### 2024-03-31 co789/homework-05/q5/main.tex
Draft for q5 to q7

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 5}
To clarify the question, we rephrase the erroneous implementation of the McEliece KEM:

\begin{algorithm}
    \caption{Key generation}
    Sample a random Goppa code $\mathcal{C}(g(x), \{\alpha_j\}_{j=1}^n)$ with parity check matrix $H$\;
    \Return{$\text{pk} = H, \text{sk} = \mathcal{C}$}
\end{algorithm}

\begin{algorithm}
    \caption{Encapsulation}
    Sample a random $e$ such that $\norm{e} = t$\;
    $c \leftarrow He$\;
    $K \leftarrow \operatorname{KDF}(m, c, 0)$
    \Return{$(c, K)$}\;
\end{algorithm}

Recall that the parity check matrix is defined by its individual entries:

\begin{equation*}
    H_{i,j} = \frac{\alpha_j^i}{g(\alpha_j)}
\end{equation*}

We can first recover individual values of $\alpha_j$ using adjacent values of the same column. For each of $1 \leq j \leq n$:

\begin{equation*}
    \frac{H_{1,j}}{H_{0,j}} 
    = \frac{\alpha_j}{g(\alpha_j)} \frac{g(\alpha_j)}{1}
    = \alpha_j
\end{equation*}

In addition, we can recover $g(\alpha_j)$ for $1 \leq j \leq n$ by inverting $H_{0,j}$:

\begin{equation*}
    H_{0,j} = g(\alpha_j)
\end{equation*}

From the construction of the Goppa code we know that the degree $t$ of the polynomial $g(x)$ is less than the number of $\alpha_j$'s, which means that with $n$ points on this polynomial we can uniquely determined the polynomial $g(x)$. Thus, we have fully recovered all parameters of the Goppa code from its parity check matrix.

Having recovered the Goppa code $\mathcal{C}$ is equivalent to posessing the secret key since the random permutation and row reduction are both omitted (which is equivalent to having $S = I_{n-k}$ and $P = I_n$). In other words, we have recovered the secret key from the public key, which trivially breaks the KEM.
\end{document}

### 2024-03-31 co789/homework-05/q6/main.tex
Draft for q5 to q7

--- 2024-03-10 co789/homework-04/q6/main.tex
@@
 \section*{Question 6}
-We will prove equivalence by showing that each statement implies the other.
 
-If (1) is true for all $\mathbf{c}_1 \neq \mathbf{c}_2$, then we can fix $\mathbf{c}_2 = \mathbf{0}$, and statement (1) becomes
+% TODO: develop the IND-CCA attack
 
-$$
-\norm{\mathbf{c}_1 - \mathbf{0}}_H = \norm{\mathbf{c}_1}_H \geq d
-$$
-
-for all non-zero $\mathbf{c}_1$, thus statement (2) is true.
-
-Now suppose (2) is true. For all $\mathbf{c}_2 \neq \mathbf{c}_1$, $\mathbf{c}_2 - \mathbf{c}_1$ is a non-zero code in $C$ because $C$ is linear. Therefore:
-
-$$
-\norm{\mathbf{c}_2 - \mathbf{c}_1}_H \geq d
-$$
-
-thus statement (1) is true.
+\subsection*{Comparing to McEliece}
+\begin{enumerate}
+    \item Classic McEliece checks that the norm (Hamming norm in this case) of the error term is exactly $t$ as defined by the security parameters.
+Any codeword with strictly fewer than $t$ bits of errors can still be correctly decoded, but the decryption routine will detect that the ciphertext has been tempered with and will appropriately reject the ciphertext as invalid.
+    \item Kyber will accept any ciphertexts whose error term's norm is no more than the threshold for decryption failure, which allows substantially more ciphertext malleability.
+This increased tolerance for ciphertext malleability is what makes the IND-CCA attack above work for Kyber, but not for McEliece.
+\end{enumerate}
 

### 2024-04-02 information-theory/main.tex
Getting started with ECE612 cheatsheet

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{ECE 612, Information Theory}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Winter, 2024}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section*{Preliminares}
\begin{definition}
    The normal distribution $N(\mu, \sigma^2)$ has the probability density function

    $$
    f(x) = \frac{1}{\sigma\sqrt{2\pi}}\exp(-\frac{1}{2}(\frac{x-\mu}{\sigma})^2)
    $$
\end{definition}

\begin{definition}
    The joint normal distribution $N(\mathbf{\mu}, K)$ is defined by probability density function:

    $$
    f(\mathbf{x}) = \frac{1}{\sqrt{(2\pi)^n\det(K)}}
    \exp(-\frac{1}{2}
        (\mathbf{x} - \mathbf{\mu})^\intercal
        K^{-1}
        (\mathbf{x} - \mathbf{\mu})
    )
    $$
\end{definition}

\section{Entropy, mutual information, divergence}

\section{Entropy rate}

\section{Asymptotic equipartition property}

\section{Data compressions}

\section{Channel capacity}

\section{Differential entropy}
\begin{theorem}[Differential entropy of Gaussian distribution]
    Let $X$ be Gaussian $N(0, \sigma^2)$, then

    $$
    h(X) = \frac{1}{2}\log{(2\pi e \sigma^2)}
    $$
\end{theorem}

\begin{theorem}
    Let $\mathbf{X}$ follow joint Gaussian distribution $N(\mathbf{0}, K)$, then:

    \begin{equation*}
        h(\mathbf{X}) = \frac{1}{2}\log ((2\pi e)^n \det{K})
    \end{equation*}
\end{theorem}

\section{Gaussian channel}
\begin{definition}[Gaussian channel with power constraint]

\end{definition}

\begin{definition}[Information channel capacity]

\end{definition}

\begin{theorem}
    The information channel capacity of a Gaussian channel is

    \begin{equation}
        \max_{f_X: E[X^2] \leq P} I(X; Y) = \frac{1}{2}\log{(1 + \frac{P}{\sigma^2})}
    \end{equation}

    Where $P$ is the power constraint, and $\sigma^2$ is the variance of the Gaussian noise. The maximum is achieved when $X$ follows Gaussian distribution $X \leftsample N(0, P)$
\end{theorem}

\section{Rate distortion theory}
\end{document}

### 2024-04-03 information-theory/main.tex
Checkpoint on 612 cheatsheet

@@
 
+\begin{theorem}[Joint normality implies marginal normality]
+    If $\mathbf{X} = (X_1, X_2, \ldots, X_n)$ follows a joint normal distribution, then any linear combination of $\mathbf{X}$ follows normal distribution.
+\end{theorem}
+
 \section{Entropy, mutual information, divergence}
+
+\begin{theorem}[Fano's inequality]
+    Let $X \rightarrow Y \rightarrow \hat{X}$ represent an encode-decode process, where $X, \hat{X} \in \mathcal{X}$ have the same support. Let $e$ denote decoding error $\hat{X} \neq X$, then:
+
+    $$
+    H(X \mid Y) \leq H(P_e) + P_e \log(\norm{\mathcal{X}})
+    $$
+\end{theorem}
 
@@
 \section{Gaussian channel}
-\begin{definition}[Gaussian channel with power constraint]
+\begin{definition}[Information channel capacity]
+    Let $Y = X + Z$, where $Z \leftsample N(0, \sigma^2)$ and $E[X^2] \leq P$ for some power level constraint $P$. The \textbf{information channel capacity} is defined by
 
-\end{definition}
-
-\begin{definition}[Information channel capacity]
-
+    $$
+    C^I = \max_{f_X : E[X^2] \leq P} I(X; Y)
+    $$
 \end{definition}

### 2024-04-03 co789/homework-05/q6/main.tex
Draft of homework 5

@@
 \section*{Question 6}
+Recall the parameters of Kyber:
 
-% TODO: develop the IND-CCA attack
+\begin{itemize}
+    \item $R_q = \mathbb{Z}_q[x] / p(x)$ for $q = 3329$ and $p(x) = x^{256} + 1$.
+    \item $A \in R_q^{k \times k}$ for $k = 2, 3, 4$ depending on the security level. $\mathbf{s}, \mathbf{e}, \mathbf{b} \in R_q^k$
+    \item $m \in R_q$ but each coefficient is either 0 or 1
+\end{itemize}
+
+Also recall that for $p(x) = x^{256} + 1$, the NTT representation of each $f(x) \in R_q$ consists of 128 degree-1 polynomials in $\mathbb{Z}_q[x]$:
+
+\begin{equation*}
+    \operatorname{NTT}(f) = (
+        \hat{f}_0 + \hat{f}_1x, \hat{f}_2 + \hat{f}_3x, \ldots, \hat{f}_{254} + \hat{f}_{255}x
+    )
+\end{equation*}
+
+Finally, we claim without proof that:
+
+\begin{enumerate}
+    \item Polynomial addition and multiplication in $R_q$ are all performed coordinate-wise
+    \item $\operatorname{NTT}(f) = \mathbf{0}$ if and only if $f = 0$
+\end{enumerate}
+
+For some $\mathbf{s} \in R_q^k$, let $\mathbf{s}[i] = R_q$ denote the i-th polynomial in the vector, and $\mathbf{s}[i][j] = s_0 + s_1x$ denote the j-th entry of the NTT representation of $\mathbf{s}[i]$
+
+\textbf{
+    For some $l \in \{1, 2, \ldots, k\}$ and some $j \in \{1, 2, \ldots, 128\}$ we can recover the j-th entry in the NTT representation of the $l$-th polynomial in $\mathbf{s} \in R_q^k$ in $q^2$ steps.
+}
+
+\begin{enumerate}
+    \item Set $c_2 \in R_q$ to have NTT representation $\operatorname{NTT}(c_2) = (0, 0, \ldots, q(x), \ldots, 0)$ where all entries are 0 except for the j-th entry, which is some random guess $g(x) = g_0 + g_1x \in \mathbb{Z}_q$.
+    \item Set $\mathbf{c}_1 \in R_q^k$ such that all except for the $l$-th polynomial is 0, and all but the j-th entry to the NTT representation of the $l$-th polynomial is exactly 1
+    \item Feed $c = (\mathbf{c}_1, c_2, H(0))$ (aka I expect $m = 0$) to the decapsulation oracle and get $K = \mathcal{O}^D(c)$
+    \item If $K = H(0, (\mathbf{c}_1, c_2), 0)$, then $g(x)$ is the correct guess for the j-th entry of the NTT representation of the $l$-th polynomial in $\mathbf{s} \in R_q^k$.
+This works because  $\operatorname{NTT}(\mathbf{c}_1)$ is all $0$ except for the j-th entry in the l-th polynomial in NTT domain, which means $\operatorname{NTT}(\mathbf{c}_1 \cdot \mathbf{s}) = \operatorname{NTT}(\mathbf{c}_1) \circ \operatorname{NTT}(\mathbf{s}) = (0, 0, \ldots, \mathbf{s}[l][j], \ldots, 0)$, so $\operatorname{NTT}(c_2 - \mathbf{c}_1 \cdot \mathbf{s}) = (0, 0, \ldots, g(x) - \mathbf{s}[l][j], \ldots, 0)$.
+I claim without proof that if $g(x) - \mathbf{s}[l][j] \neq 0$, then $c_2 - \mathbf{c}_1 \cdot \mathbf{s}$ will not round to $0 \in R_q$, which means that the decryption $\hat{m} \neq 0$, which will cause the decapsulation oracle to output the implicit rejection $H(\mathbf{s}, c, 1)$ instead of $H(0, (\mathbf{c}_1, c_2), 0)$
+\end{enumerate}
+
+Since $g(x)$ is a degree-1 polynomial, it takes at most $q^2$ guesses to recover a single entry.
+We can then repeat the attack for all $j \in \{1, 2, \ldots, 128\}$ and $l \in \{1, 2, \ldots, k\}$.
+In other words, we can recover the secret key $\mathbf{s}$ (by recovering its NTT representation) in $q^2 \cdot \frac{n}{2} \cdot k$ steps.
 

### 2024-04-09 information-theory/main.tex
Draft of homework 5
differential entropy

@@
 
+\begin{theorem}
+    Let $X, Y$ be random variables. $I(X; Y)$ is concave with respect to the probability distribution of X. For a fixed marignal distribution of $X$, $I(X; Y)$ is convex with respect to $f_{Y \mid X}$
+\end{theorem}
+
 \begin{theorem}[Fano's inequality]
@@
 \section{Differential entropy}
-\begin{theorem}[Differential entropy of Gaussian distribution]
-    Let $X$ be Gaussian $N(0, \sigma^2)$, then
+\begin{definition}
+    Let $X \in \mathbb{R}$ be a random variable with probability density function $f_X$. The \textbf{differential entropy} of $X$ is defined by
 
-    $$
-    h(X) = \frac{1}{2}\log{(2\pi e \sigma^2)}
-    $$
-\end{theorem}
+    \begin{equation*}
+        h(X) = -\int_{x \in \mathbb{R}} f_X(x)\ln(f_X(x))dx = -E[\ln(f_X(X))]
+    \end{equation*}
+\end{definition}
+
+Some notable results:
+
+\begin{itemize}
+    \item For uniform distribution over $[0, a]$, $f_X(x) = \frac{1}{a}$, $h(X) = \ln(a)$
+    \item For $X \sim N(0, \sigma^2)$, $h(X) = \frac{1}{2}\ln(2\pi e \sigma^2)$
+    \item For multivariate normal $\mathbf{X} \sim N(\mathbf{0}, K)$, $h(\mathbf{X}) = \frac{1}{2}\log ((2\pi e)^n \det{K})$
+    \item For some constant $a$, $h(X + a) = h(X)$
+    \item For some constant $a$, $h(aX) = \ln(\norm{a}) + h(X)$
+\end{itemize}
 
 \begin{theorem}
-    Let $\mathbf{X}$ follow joint Gaussian distribution $N(\mathbf{0}, K)$, then:
+    Let $X$ be continuous random variable with $0$ mean and $\sigma^2$ variance, then
 
     \begin{equation*}
-        h(\mathbf{X}) = \frac{1}{2}\log ((2\pi e)^n \det{K})
+        h(X) \leq \frac{1}{2}\ln(2\pi e \sigma^2)
     \end{equation*}
+
+    Equality is reached if and only if $X$ is Gaussian
 \end{theorem}
@@
 \section{Rate distortion theory}
+\begin{theorem}
+    Let $X$ follow $\operatorname{Bernoulli}(p)$, then
+
+    $$
+    R(D) = \begin{cases}
+        h(p) - h(D)  & \text{When $D < \min(p, 1-p)$} \\
+        0   &  \text{otherwise}
+    \end{cases}
+$$
+\end{theorem}
+
+\begin{theorem}
+    For $X \leftsample N(0, \sigma^2)$:
+
+    $$
+    R(D) = \begin{cases}
+        \frac{1}{2}\log(\frac{\sigma^2}{D}) & (D \leq \sigma^2) \\
+        0 & \text{otherwise}
+    \end{cases}
+    $$
+\end{theorem}
 \end{document}

### 2024-04-09 information-theory/main.tex
Checkpoint

@@
 
-    Where $P$ is the power constraint, and $\sigma^2$ is the variance of the Gaussian noise. The maximum is achieved when $X$ follows Gaussian distribution $X \leftsample N(0, P)$
+    Where $P$ is the power constraint, and $\sigma^2$ is the variance of the Gaussian noise. The maximum is achieved when $X$ follows Gaussian distribution $X \sim N(0, P)$
 \end{theorem}
+
+\subsection{Parallel Gaussian Channel}
+Suppose for $1 \leq l \leq n$, $Y_l = X_l + Z_l$, where $Z_l \sim N(0, N_l)$ is independent Gaussian noise, then $\mathbf{Y} = (Y_l)_{l=1}^n$ can be seen as the output of $n$ parallel Gaussian channels combined into a single channel.
+The capacity of the combined channel is
+
+\begin{equation*}
+    C = \max_{f_\mathbf{X}}I(\mathbf{X}; \mathbf{Y})
+\end{equation*}
+
+The optimal strategy for maximizing the capacity is to make $X_l$ independent and individually Gaussian.
+If there is a combined power constraint $\sum_{l=1}^nE[X_l^2] \leq P$, then \textit{power should be first given to the channel with the lowest amount of noise until the combined noise + power exceeds the next lowest noise} (\textbf{waterfilling}).
 

### 2024-04-10 information-theory/main.tex
information theory cheatsheet

@@
 \usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{graphicx}
 
@@
 \section{Entropy, mutual information, divergence}
+    \begin{definition}[Entropy]
+    The entropy of a random variable $X \in \mathcal{X}$ is defined by
+
+        \begin{equation*}
+            H(X) = -\sum_{x \in \mathcal{X}}p_X(x)\log{p_X(x)}
+        \end{equation*}
+    \end{definition}
+
+    \begin{definition}[Mutual informaiton]
+    The mutual information between two random variables $X, Y$ is defined by
+
+    \begin{equation*}
+        I(X; Y) = H(X) - H(X \mid Y)
+    \end{equation*}
+    \end{definition}
+
+    \begin{proposition}
+    Conditioning does not increase entropy
+        \begin{equation*}
+            H(X \mid Y) \leq H(X)
+        \end{equation*}
+    \end{proposition}
+
+    A nice trick for reasoning about the entropy of "sum of random variables":
+
+    $$
+    \begin{aligned}
+        H(X + Y \mid X) &= \sum_{x \in \mathcal{X}}p_X(x)H(X + Y \mid X=x) \\
+        &= \sum_{x \in \mathcal{X}}p_X(x)H(x + Y \mid X=x) \\
+        &= \sum_{x \in \mathcal{X}}p_X(x)H(Y \mid X=x) \\
+        &= H(Y \mid X)
+    \end{aligned}
+    $$
+
+    \subsection{Convexity}
+
+    \begin{definition}
+        A function $f$ is convex over the interval $x_1 \leq x \leq x_2$ if for all $0 \leq t \leq 1$:
+
+        \begin{equation*}
+            (1-t)f(x_1) + tf(x_2) \geq f((1-t)x_1 + tx_2)
+        \end{equation*}
+    \end{definition}
+
+    Some notable functions' convexity:
+    \begin{enumerate}
+        \item Entropy is a concave function with respect to the probability distribution
+        \item Mutual information $I(X; Y)$ is concave with respect to $p_X$ and is convex with respect to $p_{Y \mid X}$ for some fixed $p_X$
+        \item $\log$ is a concave function
+        \item The negative of a convex function is concave
+        \item The sum or product of a linear function and a concave function is concave
+        \item The composition of a linear function and a concave function is concave
+    \end{enumerate}
+
+    \begin{theorem}[Jensen's Inequality]
+        If $f$ is a convex function and $X$ is a random variable, then
+        \begin{equation*}
+            E[f(X)] \geq f(E[X])
+        \end{equation*}
+    \end{theorem}
+
+    \subsection{Markov chain}
+    \begin{definition}
+        Random variables $X, Y, Z$ form a Markove chain (denoted by $X \rightarrow Y \rightarrow Z$) if $p(z \mid x, y) = p(z \mid y)$
+    \end{definition}
+
+    \begin{theorem}[Data processing inequality]
+        If $X \rightarrow Y \rightarrow Z$, then
+        \begin{equation*}
+            I(X; Z) \leq I(Y; Z)
+        \end{equation*}
+    \end{theorem}
+
+    \begin{proposition}
+        \begin{equation*}
+            I(X; Y \mid Z) = 0 \Leftrightarrow X \rightarrow Z \rightarrow Y
+        \end{equation*}
+    \end{proposition}
+
+    \begin{proposition}
+        If $X \rightarrow Y \rightarrow Z$, then $Z \rightarrow Y \rightarrow X$
+    \end{proposition}
 
     \begin{theorem}
-    Let $X, Y$ be random variables. $I(X; Y)$ is concave with respect to the probability distribution of X. For a fixed marignal distribution of $X$, $I(X; Y)$ is convex with respect to $f_{Y \mid X}$
+        Let $X, Y$ be random variables.
+$I(X; Y)$ is concave with respect to the probability distribution of X.
+For a fixed marignal distribution of $X$, $I(X; Y)$ is convex with respect to $f_{Y \mid X}$
     \end{theorem}
 
-\begin{theorem}[Fano's inequality]
+    \begin{theorem}[Fano's inequality]\label{fanos-inequality}
         Let $X \rightarrow Y \rightarrow \hat{X}$ represent an encode-decode process, where $X, \hat{X} \in \mathcal{X}$ have the same support. Let $e$ denote decoding error $\hat{X} \neq X$, then:
@@
         $$
-    H(X \mid Y) \leq H(P_e) + P_e \log(\norm{\mathcal{X}})
+        H(X \mid Y) \leq h(P_e) + P_e \log(\norm{\mathcal{X}})
         $$
@@
 \section{Entropy rate}
+    \begin{definition}
+        A stochastic process is stationary if for all integer $n \geq 1$ and integer time shift (possibly negative) $l$:
+        \begin{equation*}
+            P[X_1^n] = P[X_{1+l}^{n+l}]
+        \end{equation*}
+    \end{definition}
+
+    \begin{definition}
+        A Markov chain is time invariant if for all $n \geq 1$:
+        \begin{equation*}
+            P[X_{n+1} \mid X_n] = P[X_2 \mid X_1]
+        \end{equation*}
+    \end{definition}
+
+    \begin{definition}
+        There are two definitions of entropy rate:
+        \begin{enumerate}
+            \item $H(\mathcal{X}) = \lim_{n\rightarrow\infty}\frac{1}{n}H(X_1^n)$
+            \item $H^\prime(\mathcal{X}) = \lim_{n\rightarrow\infty}H(X_n\mid X_1^{n-1})$
+        \end{enumerate}
+    \end{definition}
+
+    \begin{theorem}
+        If $X_i$ is a stationary process, then the two entropy rates both exist and are equal
+    \end{theorem}
+    \begin{lemma}
+        If $X_i$ is a stationary process, then $H(X_n \mid X_1^{n-1})$ is non-increasing with respect to $n$ and has a limit
+    \end{lemma}
 
 \section{Asymptotic equipartition property}
+    \begin{theorem}[Weak law of large numbers (WLLN)]
+        Let $X \sim p_X$ be a random variable, and let $X_l \sim p_X$ be i.i.d. random variables following the same distributions, then:
+
+        \begin{equation*}
+            \lim_{n \rightarrow \infty}\frac{1}{n}\sum_{i=1}^n X_i = E[X]
+        \end{equation*}
+
+        In other words, for any $\delta, \epsilon > 0$, there exists $N > 0$ such that for $n > N$:
+
+        \begin{equation*}
+            P[\norm{\frac{1}{n}\sum_{i=1}^n X_i - E[X]} < \delta] < \epsilon
+        \end{equation*}
+    \end{theorem}
+
+    Recall that the entropy of a random variable can be defined as $H(X) = -E[\log(P_X)]$, so the weak law of large numbers also applies: the joint distribution of i.i.d.
+$X_i$ converges to the probability $2^{-nH(X)}$.
+
+    \begin{definition}[Typical sequence]
+        For random variable $X \sim p_X$, the set of \textbf{typical sequence} $A_\epsilon^n$ denotes the set of sequences whose probability is close to the entropy of the random variable
+
+        \begin{equation*}
+            A_\epsilon^{(n)} = \{
+                \mathbf{x} \in \mathcal{X}^n 
+                \mid \norm{
+                    -\frac{1}{n}\log p_{\mathbf{X}}(\mathbf{x})
+                    - H(X)
+                } < \epsilon
+            \}
+        \end{equation*}
+    \end{definition}
+
+    \begin{theorem}[Asymptotic equipartition property]
+        Let $X \sim p_X$ and let $\mathbf{X} \in \mathcal{X}^n$ denote i.i.d. sequence following the same distribution, then:
+
+        \begin{enumerate}
+            \item $\lim_{n \rightarrow \infty}P[\mathbf{X} \in A_\epsilon^{(n)}] = 1$
+            \item $\norm{A_\epsilon^{(n)}} \leq 2^{n(H - \epsilon)}$
+            \item For sufficiently large $n$, $\norm{A_\epsilon^{(n)}} \geq (1-\epsilon)2^{n(H+\epsilon)}$
+        \end{enumerate}
+    \end{theorem}
+
+    \begin{proof}
+        \textit{A sketch of the proof}
+        (1) is the direct consequence of the weak law of large numbers
+
+        (2): begin with \textbf{the sum of probability of all possible sequence is 1}
+
+        (3): begin with \textbf{for sufficiently large $n$, $P[\mathbf{x} \in A_\epsilon^{(n)}] \geq 1 - \epsilon$}
+    \end{proof}
 
 \section{Data compressions}
+    \begin{theorem}[Kraft inequality]
+    A $D$-ary prefix code for a finite set of $m$ symbols $\mathcal{X}$ exists if and only if the lengths of the code words $l_1, l_2, \ldots, l_m$ satisfy:
+    \begin{equation*}
+        \sum_{x \in \mathcal{X}} D^{-l(x)} \leq 1
+    \end{equation*}
+    \end{theorem}
+
+    Using the method of Lagrange multiplier, we can optimize the expected codeword length $E[L]$ under the constraint fo Kraft inequality:
+
+    \begin{equation*}
+        l_i = \lceil \log_D\frac{1}{p_i} \rceil
+    \end{equation*}
+
+    \begin{theorem}[Optimal code length]
+    $H_D(X) \leq E[L] \leq H_D(X) + 1$
+    \end{theorem}
+
+    \begin{theorem}[McMillan inequality]
+        A D-ary uniquely decodable code exists if and only if
+
+        \begin{equation*}
+            \sum_{i}D^{-l_i} \leq 1
+        \end{equation*}
+    \end{theorem}
+
+    In practice, \textbf{Huffman code} offers the optimal noiseless compression:
+
+    \includegraphics[width=0.8\textwidth]{./img/huffman-example.png}
 
 \section{Channel capacity}
+    In this section we discuss \textbf{how much information can be transmitted through a noisy channel}
+
+    \begin{definition}
+        A discrete channel consists of a discrete set of words $\mathcal{W}$, two sets of symbols $\mathcal{X}, \mathcal{Y}$, an encoder $\mathcal{W} \rightarrow \mathcal{X}$, a decoder $\mathcal{Y} \rightarrow \mathcal{W}$, and a channel described by the conditional distribution $p_{Y \mid X}$:
+
+        \begin{equation*}
+            W \rightarrow X \rightarrow Y \rightarrow \hat{W}
+        \end{equation*}
+
+        For this section we are only concerned with \textbf{memoryless} channel: given $X_n$, $Y_n$ is independent of all other variables:
+
+        \begin{equation*}
+            p_{\mathbf{Y} \mid \mathbf{X}} = \prod_{l=1}^n p_{Y_l \mid X_l}
+        \end{equation*}
+
+        For a fixed $p_{Y \mid X}$, \textbf{information channel capacity} is defined by
+
+        \begin{equation*}
+            C^I = \max_{p_X} I(p_X; p_{Y \mid X})
+        \end{equation*}
+    \end{definition}
+
+    \subsection{Channel rate}
+    \begin{definition}[(M, n) code]
+        An $(M, n)$-code consists of a collection of $M$ words $\mathcal{W}: \norm{\mathcal{W}} = M$, two sets of symbols $\mathcal{X}, \mathcal{Y}$, an encoder $f: \mathcal{W} \rightarrow \mathcal{X}^n$, and a decoder $g: \mathcal{Y}^n \rightarrow \mathcal{W}$
+    \end{definition}
+
+    \begin{definition}[Decoding error]
+        For a single word $w \in \mathcal{W}$, the \textbf{conditional probability of error} is defined by
+
+        \begin{equation*}
+            \lambda_l = P[g(\mathbf{Y}) \neq l \mid \mathbf{X} = f(w)]
+        \end{equation*}
+
+        Denote the \textbf{maximal probability of error} across all words by
+
+        \begin{equation*}
+            \lambda^(n) = \max_{w \in \mathcal{W}} \lambda_w
+        \end{equation*}
+
+        Denote the \textbf{average probability of error} across all words by
+
+        \begin{equation*}
+            P_e^{(n)} = \frac{1}{M}\sum_{w \in \mathcal{W}}\lambda_w
+        \end{equation*}
+
+        The \textbf{rate of an $(M, n)$-code} denotes the number of bits of information transmitted per channel use:
+
+        \begin{equation*}
+            R = \frac{\log(M)}{n}
+        \end{equation*}
+    \end{definition}
+
+    \begin{definition}[Achievable rate]
+        A rate $R > 0$ is \textbf{achievable} if there exists a sequence of $(2^{nR}, n)$-code such that:
+
+        \begin{equation*}
+            \lim_{n \rightarrow \infty}\lambda^{(n)} = 0
+        \end{equation*}
+
+        The \textbf{capacity} of a channel is the supremum of all achievable rates
+    \end{definition}
+
+    \subsection{Jointly typical sequence}
+    Let $(\mathbf{X}, \mathbf{Y}) \in \mathcal{X}^n \times \mathcal{Y}^n$ be i.i.d.
+sequences drawn according to the joint distribution $P_{X, Y}$, then by the weak law of large numbers, $\lim_{n \rightarrow \infty}-\frac{1}{n}\log(p_{X, Y}(\mathbf{x}, \mathbf{y})) = H(X, Y)$.
+We can define notion of typicality on the joint sequence similar to the notion of typicality on individual sequences
+
+    \begin{definition}[Jointly typical sequence]
+        The set of jointly typical sequence $A_\epsilon^{(n)}$ is the set of $(\mathbf{x}, \mathbf{y})$ such that:
+
+        \begin{enumerate}
+            \item $\norm{-\frac{1}{n}\log(p_X(\mathbf{x})) - H(X)} \leq \epsilon$
+            \item $\norm{-\frac{1}{n}\log(p_Y(\mathbf{y})) - H(Y)} \leq \epsilon$
+            \item $\norm{
+                -\frac{1}{n}\log(p_{X, Y}(\mathbf{x}, \mathbf{y})) - H(X, Y)
+            } \leq \epsilon$
+        \end{enumerate}
+
+        In other words, $(\mathbf{x}, \mathbf{y})$ is jointly typical if $\mathbf{x}$ and $\mathbf{y}$ are each individually typical according to their marginal distribution, and $(\mathbf{x}, \mathbf{y})$ is typical according to the joint distribution
+    \end{definition}
+
+    \begin{theorem}[Joint asymptotic equipartition property]
+        Let $(\mathbf{X}, \mathbf{Y}) \in \mathcal{X}^n \times \mathcal{Y}^n$ be i.i.d. sequences drawn according to $p_{X, Y}$, then
+
+        \begin{enumerate}
+            \item $
+                \lim_{n \rightarrow \infty}
+                P[(\mathbf{X}, \mathbf{Y}) \in A_\epsilon^{(n)}] = 1
+            $
+            \item $\norm{A_\epsilon^{(n)}} \leq 2^{
+                n(H(X, Y) + \epsilon)
+            }$
+            \item For sufficiently large $n$:
+                $$
+                \norm{A_\epsilon^{(n)}} \geq (1-\epsilon)2^{n(H(X, Y) - \epsilon)}
+                $$
+            \item Let $\tilde{\mathbf{X}}$ and $\tilde{\mathbf{Y}}$ be drawn independently from $p_X$ and $p_Y$, then
+                $$
+                    P[(\tilde{\mathbf{X}}, \tilde{\mathbf{Y}}) \in A_\epsilon^n] 
+                    \leq 2^{-n(I(X; Y) - 3\epsilon)}
+                $$
+            \item Let $\tilde{\mathbf{X}}$ and $\tilde{\mathbf{Y}}$ be drawn independently from $p_X$ and $p_Y$, then for sufficiently large $n$:
+                $$
+                    P[(\tilde{\mathbf{X}}, \tilde{\mathbf{Y}}) \in A_\epsilon^n] 
+                    \geq (1-\epsilon)2^{-n(I(X; Y) + 3\epsilon)}
+                $$
+        \end{enumerate}
+    \end{theorem}
+
+    \begin{proof}
+        (1) is the direct consequence of the weak law of large numbers
+
+        (2):
+        \begin{equation*}
+            \begin{aligned}
+                1 &= \sum_{
+                    (\mathbf{x}, \mathbf{y}) \in \mathcal{X}^n \times \mathcal{Y}^n
+                    } p(\mathbf{x}, \mathbf{y}) \\
+                &\geq \sum_{
+                    (\mathbf{x}, \mathbf{y}) \in A_\epsilon^{(n)}
+                    } p(\mathbf{x}, \mathbf{y}) \\
+                &\geq \sum_{
+                    (\mathbf{x}, \mathbf{y}) \in A_\epsilon^{(n)}
+                    } 2^{-n(H(X, Y) + \epsilon)} \\
+                &=  \norm{A_\epsilon^{(n)}} 2^{-n(H(X, Y) + \epsilon)}
+            \end{aligned}
+        \end{equation*}
+
+        (3): for sufficiently large $n$, $P[(\mathbf{X}, \mathbf{Y}) \in A_\epsilon^{(n)}] \geq 1 - \epsilon$. On the other hand:
+
+        \begin{equation*}
+            \begin{aligned}
+                P[(\mathbf{X}, \mathbf{Y}) \in A_\epsilon^{(n)}]
+                &= \sum_{(\mathbf{x}, \mathbf{y}) \in A_\epsilon}
+                    P[(\mathbf{X}, \mathbf{Y}) = (\mathbf{x}, \mathbf{y})] \\
+                &\leq \sum_{(\mathbf{x}, \mathbf{y}) \in A_\epsilon}
+                    2^{-n(H_{X,Y} - \epsilon)} \\
+                &= \norm{A_\epsilon} \cdot
+                    2^{-n(H_{X,Y} - \epsilon)} \\
+            \end{aligned}
+        \end{equation*}
+
+        Putting the two inequalities above together:
+
+        \begin{equation*}
+            \norm{A_\epsilon} \cdot 2^{-n(H_{X,Y} - \epsilon)} \geq 1 - \epsilon
+        \end{equation*}
+
+        (4): consider the probability of $(\tilde{\mathbf{X}}, \tilde{\mathbf{Y}})$ falling into the jointly typical set:
+        \begin{equation*}
+            \begin{aligned}
+                P[(\tilde{\mathbf{X}}, \tilde{\mathbf{Y}}) \in A_\epsilon^{(n)}]
+                &= \sum_{(\tilde{\mathbf{x}}, \tilde{\mathbf{y}}) \in A_\epsilon^{(n)}}
+                    P[
+                        \tilde{\mathbf{X}} = \tilde{\mathbf{x}},
+                        \tilde{\mathbf{Y}} = \tilde{\mathbf{y}}
+                    ] \\
+                &= \sum_{(\tilde{\mathbf{x}}, \tilde{\mathbf{y}}) \in A_\epsilon^{(n)}}
+                    P[\tilde{\mathbf{X}} = \tilde{\mathbf{x}}] \cdot
+                    P[\tilde{\mathbf{Y}} = \tilde{\mathbf{y}}] \\
+                &\leq \sum_{(\tilde{\mathbf{x}}, \tilde{\mathbf{y}}) \in A_\epsilon^{(n)}}
+                    2^{-n(H_X - \epsilon)} \cdot 2^{-n(H_Y - \epsilon)} \\
+                &= \norm{A} \cdot 2^{-n(H_X - \epsilon)} \cdot 2^{-n(H_Y - \epsilon)} \\
+                &\leq 2^{n(H(X, Y) + \epsilon)} 
+                    \cdot 2^{-n(H_X - \epsilon)} \cdot 2^{-n(H_Y - \epsilon)}
+            \end{aligned}
+        \end{equation*}
+
+        (5): Using similar logic as shown above:
+        \begin{equation*}
+            \begin{aligned}
+                P[(\tilde{\mathbf{X}}, \tilde{\mathbf{Y}}) \in A_\epsilon^{(n)}]
+                &= \sum_{(\tilde{\mathbf{x}}, \tilde{\mathbf{y}}) \in A_\epsilon^{(n)}}
+                    P[
+                        \tilde{\mathbf{X}} = \tilde{\mathbf{x}},
+                        \tilde{\mathbf{Y}} = \tilde{\mathbf{y}}
+                    ] \\
+                &= \sum_{(\tilde{\mathbf{x}}, \tilde{\mathbf{y}}) \in A_\epsilon^{(n)}}
+                    P[\tilde{\mathbf{X}} = \tilde{\mathbf{x}}] \cdot
+                    P[\tilde{\mathbf{Y}} = \tilde{\mathbf{y}}] \\
+                &\geq \sum_{(\tilde{\mathbf{x}}, \tilde{\mathbf{y}}) \in A_\epsilon^{(n)}}
+                    2^{-n(H_X + \epsilon)} \cdot 2^{-n(H_Y + \epsilon)} \\
+                &= \norm{A_e}
+                    2^{-n(H_X + \epsilon)} \cdot 2^{-n(H_Y + \epsilon)} \\
+                &\geq (1-\epsilon)2^{n(H_{X,Y} - \epsilon)} \cdot
+                    2^{-n(H_X + \epsilon)} \cdot 2^{-n(H_Y + \epsilon)}
+            \end{aligned}
+        \end{equation*}
+    \end{proof}
+
+    \subsection{Channel coding theorem}
+    \begin{theorem}
+        A rate $R$ is achievable if and only if it is below the information channel capacity
+    \end{theorem}
+    \begin{proof}
+        \textit{This is a sketch of proof}
+
+        \textbf{First we show that rate below information capacity is achievable}:
+        Let $R$ be a rate such that $R < C$, then we can construct a $(2^{nR}, n)$-code through the following procedure:
+        \begin{enumerate}
+            \item Find some distribution of $p_X$ and $\epsilon > 0$ such that $R = I(p_X;p_{Y \mid X}) - 4\epsilon$
+            \item For each $w \in \{1, \ldots, 2^{nR}\}$, the encoding is $f(w) = (x_1(w), x_2(w), \ldots, x_n(w))$ where $x_i(w)$ is i.i.d. sampled from $p_X$
+            \item Transmit each symbol $x_i(w)$ according to $p_{Y \mid X}$
+            \item Upon receiving $\mathbf{y}$, find some $\hat{w} \in \{1, \ldots, 2^{nR}\}$ such that $(\mathbf{x}, \mathbf{y})$ is jointly typical according to $p_X$ and $p_{Y \mid X}$ (which can be used to compute $p_{X, Y}$ and $p_Y$)
+        \end{enumerate}
+
+        For each word $w$, decoding error occurs if and only if one of two scenarios occurs:
+
+        \begin{enumerate}
+            \item $(\mathbf{x}(w), \mathbf{y})$ is not jointly typical, in which case no appropriate $\hat{w}$ can be found in the decoding step
+            \item there is another $w^\prime \neq w$ such that $(\mathbf{x}(w^\prime), \mathbf{y})$ is also jointly typical
+        \end{enumerate}
+
+        The probability of (1) converges to 0 because $\lim_{n \rightarrow \infty}P[(\mathbf{x}, \mathbf{y}) \not \in A_\epsilon^{(n)}] = 0$.
+
+        The probability of (2) converges to 0 because $\mathbf{y}$ is independent of $\mathbf{x}(w^\prime)$.
+By joint AEP, the probability of independently sampled sequences being jointly typical is at most $2^{-n(I(X; Y) - 3\epsilon)}$, hence:
+
+        \begin{equation*}
+            \begin{aligned}
+                \sum_{w^\prime \neq w} P[(\mathbf{x}(w^\prime), \mathbf{y}) \in A_\epsilon^{(n)}]
+                &= (2^{nR}-1)P[(\mathbf{x}(w^\prime), \mathbf{y}) \in A_\epsilon^{(n)}] \\
+                &\leq (2^{nR}-1)2^{-n(I(X; Y) - 3\epsilon)} \\
+                &\leq 2^{nR - n(I_{X;Y} - 3\epsilon)} \\
+                &= 2^{n(I_{X;Y} - 4\epsilon) - n(I_{X;Y} - 3\epsilon)} \\
+                &= 2^{-n\epsilon}
+            \end{aligned}
+        \end{equation*}
+
+        Which also converges to 0 as $n \rightarrow \infty$
+
+        \textbf{Converse: achievable rates are below information capacity}
+
+        Recall that the channel represents a Markov chain $W \rightarrow \mathbf{X} \rightarrow \mathbf{Y} \rightarrow \hat{W}$. By Fano's inequality, we know that:
+
+        \begin{equation*}
+            H(W \mid \hat{W}) \leq h(p_e) + p_e \cdot \log(\norm{\mathcal{W}})
+        \end{equation*}
+
+        Consider the mutual information between $W$ and $\hat{W}$:
+
+        \begin{equation*}
+            I(W; \hat{W}) = H(W) - H(W \mid \hat{W})
+        \end{equation*}
+
+        Assuming that $W$ is uniformly distributed, $H(W) = \log(\norm{\mathcal{W}}) = nR$.
+
+        By the data-processing inequality $I(W; \hat{W}) \leq I(\mathbf{X}; \mathbf{Y})$.
+By the chain rule of mutual information: $I(\mathbf{X}; \mathbf{Y}) \leq nI(X; Y)$.
+By the definition of information channel capacity $I(X; Y) \leq C$.
+Putting everything together, we have
+
+        \begin{equation*}
+            nR - nC \leq p_e
+        \end{equation*}
+
+        Assuming the rate is achievable, then $\lim_{n \rightarrow \infty}p_e = 0$, which implies $R \leq C$.
+    \end{proof}
 
@@
 \section{Rate distortion theory}
+    \begin{definition}
+        For some fixed $p_X$, the \textbf{information rate distortion} is defined by
+
+        $$
+        R^I(D) = \min_{p_{\hat{X} \mid X}: E[d(X, \hat{X})] \leq D} I(X; \hat{X})
+        $$
+    \end{definition}
+
     \begin{theorem}

### 2024-04-11 information-theory/main.tex
Added divergence stuff; fixed typos

@@
 
-    \begin{definition}[Mutual informaiton]
+    \begin{definition}[KL Divergence]
+        The \textbf{KL Divergence} between two probability distributions $p, q$ on the same support $\mathcal{X}$ is defined by
+
+        \begin{equation*}
+            D(p \Vert q) = \sum_{x \in \mathcal{X}} p(x) \log{
+                \frac{p(x)}{q(x)}
+            }
+        \end{equation*}
+    \end{definition}
+
+    \begin{definition}[Mutual information]
         The mutual information between two random variables $X, Y$ is defined by
@@
             I(X; Y) = H(X) - H(X \mid Y)
+        \end{equation*}
+
+        It can also be expressed as the KL divergence between the joint distribution $p_{X, Y}$ and the product of marginal distribution $p_Xp_Y$:
+
+        \begin{equation*}
+            I(X; Y) = D(p_{X,Y} \Vert p_X \cdot p_Y)
         \end{equation*}

### 2024-04-16 co789/project/manuscript/main.tex
Checkpoint

@@
 \newcommand{\round}[1]{\lfloor {#1} \rceil}
-\newcommand{\norm}[1]{\Vert {#1} \Vert}
+\newcommand{\norm}[1]{\vert {#1} \vert}
+\newcommand{\Norm}[1]{\Vert {#1} \Vert}
 \newcommand{\var}[1]{\operatorname{Var}[{#1}]}
@@
 \newtheorem{proposition}{Proposition}[section]
-\newtheorem{lemma}{Lemma}[theorem]
+\newtheorem{lemma}{Lemma}
 
@@
 \section{Introduction}
+Indistinguishability under (adaptive) chosen-ciphertext attack (IND-CCA2) is widely recognized as the desirable security notion for public-key cryptography.
+However, directly achieving IND-CCA2 security is difficult.
+Previous attempts at deploying public-key cryptography in production, such as the usage of RSA PKCS1 v1.5 in early versions of SSl/TLS, were found to be vulnerable to adaptive chosen-ciphertext attacks.
+
+Instead of directly constructing IND-CCA2 secure cryptosystem from NP-hard problems, recent works approached this problem by proposing generic transformation that take cryptographic primitives of lesser strengths (e.g.
+OW-CPA, IND-CPA) and produce encryption schemes that lose only a negligible amount of security.
+One such transformation was proposed by Fujisaki and Okamoto in 1999 and later improved by Hofheinz, Hovelmann, and Kiltz in 2017.
+With simple construction and robust security reduction, the FO transformation is adopted by submissions to NIST's post-quantum cryptography competition (notably by Kyber, which has been standardized in FIPS 203 in 2024).
+
+In this paper, we will review the constructions of the Fujisaki-Okamoto transformation and its variations.
+We will also review their security results, including the techniques used in the security reduction.
+Finally, we will discuss open problems and propose some optimization.
 
@@
 % TODO: from IND-CPA to OW-CPA
+
+\subsection{Spread and correctness}
+The spread of measures the randomness of ciphertext in a probabilistic encryption scheme, where the randomness of the encryption routine is determined by a random coin $r \leftsample \text{Coin}$ from the coin space.
+
+\begin{definition}[$\gamma$-spread]
+    The spread $\gamma$ of an asymmetric encryption scheme $(\operatorname{KeyGen}, E, D)$ for some fixed public key $\text{pk}$ and plaintext $m$ is defined by
+
+    \begin{equation*}
+        \gamma(\text{pk}, m) = -\log{\Norm{E_\text{pk}(m)}_\infty}
+    \end{equation*}
+
+    Where $\Norm{E_\text{pk}(m)}_\infty$ is the min-entropy of the ciphertext with respect to the coin:
+
+    \begin{equation*}
+        \Norm{E_\text{pk}(m)}_\infty 
+        = \max_{c \in \mathcal{C}} P[E_\text{pk}(m) = c]
+    \end{equation*}
+\end{definition}
+
+\begin{definition}[correctness]
+    A public-key encrpytion scheme $(\operatorname{KeyGen}, E, D)$ is $\delta$-correct if across all possible keypairs and messages, the probability of decryption error is at most $\delta$:
+
+    \begin{equation*}
+        \max_{
+            m \in \mathcal{M},
+            (\text{pk}, \text{sk}) \leftsample \operatorname{KeyGen}()
+        } P[D_\text{sk}(E_\text{pk}(m)) \neq m]
+        \leq \delta
+    \end{equation*}
+\end{definition}
+
+\subsection{Difference lemma}
+The difference lemma is used extensively in the sequence of games to estimate the loss of security when one game is transformed into another
+
+\begin{lemma}[Difference lemma]
+    Let $A, B, F$ be events defined on the same probability space. If $P[A \cap \neg F] = P[B \cap \neg F]$, then $P[A] - P[B] \leq P[F]$
+\end{lemma}
+
+\begin{proof}
+
+\end{proof}
+
+\subsection{IND-CPA to OW-CPA}
 

### 2024-04-17 co789/project/manuscript/main.tex
project checkpoint

--- 2024-03-20 co789/project/manuscript/main.tex
@@
 % Algorithms and pseudocode
-\usepackage{algorithm}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
 \usepackage{algpseudocode}
+\usepackage{multicol}
 
@@
 
-\section{The Fujisaki-Okamoto Transformation}
-The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} takes a IND-CPA public-key encryption scheme (PKE) and a few other cryptographic primitives of weaker security as inputs, and constructs a hybrid encryption scheme that achieves IND-CCA2 security.
+\subsection{Spread and correctness}
+The spread of measures the randomness of ciphertext in a probabilistic encryption scheme, where the randomness of the encryption routine is determined by a random coin $r \leftsample \text{Coin}$ from the coin space.
 
-In the conference paper, the security reduction is performed under the random oracle model, and the security of the hybrid scheme degrades linearly with the number of hash queries and the number of decryption queries.
-On the other hand, this scheme is desirable for its simplicity, and the low requirements for the input primitives: the PKE only needs to be one-time one-way secure, and the symetric cipher only needs to be indistinguishable under one-time attack.
-
-% TODO: Need more introduction?
-
-\subsection{The hybrid scheme and security result}
-The hybrid encryption scheme contains three routines: key generation, encryption, and decryption.
-The input for the hybrid scheme includes a public-key encryption scheme $\text{KeyGen}^\text{asym}, E^\text{asym}, D^\text{asym}$, a symmetric encryption scheme $E^\text{sym}, D^\text{sym}$, and two hash functions $G: \mathcal{M}^\text{asym} \mapsto \mathcal{K}^\text{sym}, H: \mathcal{M}^\text{asym} \times \mathcal{C}^\text{sym} \mapsto \text{Coin}^\text{asym}$.
-
-\begin{algorithm}
-\caption{FO Key generation}\label{fo-key-gen}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \State $
-        (\text{PK}^\text{asym}, \text{SK}^\text{asym}) 
-        \leftarrow \text{PKE.KeyGen}()
-    $
-    \State $
-        \text{PK}^\text{hy} \leftarrow \text{PK}^\text{asym},
-        \text{SK}^\text{hy} \leftarrow \text{SK}^\text{asym},
-    $
-
-    \State \Return $(\text{PK}^\text{hy}, \text{SK}^\text{hy})$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-\caption{FO Key encryption}\label{fo-key-enc}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \Require $m \in \mathcal{M}^\text{sym}$
-    \State Sample from the PKE's message space $\sigma \leftsample \mathcal{M}^\text{asym}$
-    \State $a \leftarrow G(\sigma)$, $c \leftarrow E^\text{sym}_a(m)$
-    \State $h \leftarrow H(\sigma, c)$
-    \State $e \leftarrow E^\text{asym}(\text{PK}^\text{hy}, \sigma, h)$
-
-    \State \Return $(e, c)$
-\end{algorithmic}
-\end{algorithm}
-
-\pagebreak
-
-The decryption routine:
-
-\begin{algorithm}
-\caption{FO Key decryption}\label{fo-key-dec}
-\begin{algorithmic}[1]  % [1] is for displaying line number
-    \Require The ciphertext $(e, c)$
-
-    \State $
-        \hat{\sigma} \leftarrow D^\text{asym}(\text{SK}^\text{hy}, e)
-    $
-
-    \State $
-        \hat{h} \leftarrow H(\hat{\sigma}, c)
-    $
-
-    \State $
-        \hat{e} \leftarrow E^\text{asym}(\text{PK}^\text{hy})
-    $
-
-    \If{$\hat{e} \neq e$}
-        \State \Return $\bot$
-    \EndIf
-
-    \State $\hat{a} \leftarrow G(\hat{\sigma})$
-    \State $\hat{m} \leftarrow D^\text{sym}_{\hat{a}}(c)$
-    \State \Return $\hat{m}$
-\end{algorithmic}
-\end{algorithm}
-
-\begin{theorem}\label{fo-security-theorem}
-    For all IND-CCA adversary $\mathcal{A}^\text{hy}$ against the hybrid encryption scheme with advantage $\epsilon^\text{hy}$, there exists a one-way one-time-encryption adversary against the public-key encryption scheme with advantage $\epsilon^\text{asym}$ and an one-time indistinguishability adversary against the symmetric encryption scheme with advantage $\epsilon^\text{sym}$ such that
+\begin{definition}[$\gamma$-spread]
+    The spread $\gamma$ of an asymmetric encryption scheme $(\operatorname{KeyGen}, E, D)$ for some fixed public key $\text{pk}$ and plaintext $m$ is defined by
 
     \begin{equation*}
-        \epsilon^\text{hy} \leq q_\text{H}\epsilon^\text{asym} + \epsilon^\text{sym} + q_\text{D}2^{-\gamma}
+        \gamma(\text{pk}, m) = -\log{\Norm{E_\text{pk}(m)}_\infty}
     \end{equation*}
 
-    where $\gamma$ is the spread of the public-key encryption scheme, $q_\text{H}$ is the total number of hash queries, and $q_\text{D}$ is the total number of decryption queries
-\end{theorem}
+    Where $\Norm{E_\text{pk}(m)}_\infty$ is the min-entropy of the ciphertext with respect to the coin:
 
-\subsection{Proof of security result}
-Theorem \ref{fo-security-theorem} is proved using a sequence of games that involves $\mathcal{A}^\text{hy}$ as the main routine, and two games that involves $\mathcal{A}^\text{hy}$ as a sub-routine.
-The sequence of games is as follows:
+    \begin{equation*}
+        \Norm{E_\text{pk}(m)}_\infty 
+        = \max_{c \in \mathcal{C}} P[E_\text{pk}(m) = c]
+    \end{equation*}
+\end{definition}
 
-\begin{itemize}
-    \item Game 0 is the standard IND-CCA2 game
+\begin{definition}[correctness]
+    A public-key encrpytion scheme $(\operatorname{KeyGen}, E, D)$ is $\delta$-correct if across all possible keypairs and messages, the probability of decryption error is at most $\delta$:
 
-    \item Game 1 is identical to game 0, except that the decryption oracle $\mathcal{O}_D$ is modified.
-Instead of using the true secret key $\text{SK}^\text{hy}$ to decrypt the query $(e_q, c_q)$, the decryption oracle checks the tape of hash function $H$ for the existence of hash query $(\sigma_H, c_H, h_H)$ such that $c_q = c_H$ and $e_q = E^\text{asym}(\text{PK}^\text{asym}, \sigma_H, h_H)$.
-If such a query exists, then $\mathcal{O}_D$ uses $\sigma_H$ to derive the symmetric key $a_q \leftarrow G(\sigma_H)$ and decrypt the queried ciphertext $c_q$.
-If no such query exists, then $\mathcal{O}_D$ will reject the queried ciphertext and output a decryption error.
-It is worth noting that this modified decryption oracle does not require the hybrid secret key $\text{SK}^\text{hy}$ to process decryption queries.
+    \begin{equation*}
+        \max_{
+            m \in \mathcal{M},
+            (\text{pk}, \text{sk}) \leftsample \operatorname{KeyGen}()
+        } P[D_\text{sk}(E_\text{pk}(m)) \neq m]
+        \leq \delta
+    \end{equation*}
+\end{definition}
 
-    \item Game 2 is identical to game 1, except the routine of encrypting the challenge ciphertext is modified: $a^\ast \leftsample \mathcal{K}^\text{sym}$ is randomly sampled from the symmetric key space instead of being queried from $G$, and $h^\ast \leftsample \text{COIN}^\text{asym}$ is randomly sampled from the asymmetric coin space instead of being queried from $H$.
-\end{itemize}
+\subsection{Difference lemma}
+The difference lemma is used extensively in the sequence of games to estimate the loss of security when one game is transformed into another
 
-Let $S_0, S_1, S_2$ denote the event that $\mathcal{A}^\text{hy}$ wins game 0, game 1, and game 2, respectively.
+\begin{lemma}[Difference lemma]
+    Let $A, B, F$ be events defined on the same probability space. If $P[A \cap \neg F] = P[B \cap \neg F]$, then $P[A] - P[B] \leq P[F]$
+\end{lemma}
 
-\begin{lemma}\label{fo-win0-win1}
-    Let $q_D$ denote the number of decryption queries, and $\gamma$ denote the spread of the PKE, then
+\begin{proof}
     \begin{equation*}
-        P[S_0] - P[S_1] \leq q_D 2^{-\gamma}
+        \begin{aligned}
+            P[A] - P[B] &= (P[A \cap F] + P[A \cap \neg F]) 
+                - (P[B \cap F] + P[B \cap \neg F]) \\
+            &= (P[A \cap F] - P[B \cap F]) 
+                + (P[A \cap \neg F] - P[B \cap \neg F]) \\
+            &= P[A \cap F] - P[B \cap F] \\
+            &\leq P[A \cap F] \\
+            &\leq P[F]
+        \end{aligned}
+    \end{equation*}
+\end{proof}
+
+In the context of security reduction using sequence of game, the difference lemma is invoked with event $A$ and $B$ set to "adversary winning one game" and "adversary winning another game", and event $F$ set to the event that would cause the two games two diverge.
+If the diverging event $F$ does not happen, then the two games are identical to each other, so the advantage of the adversary remains the same across the two games; if the diverging event $F$ happens, then the advantage of the adversary does not change by more than the probability of the diverging event $F$.
+
+\subsection{IND-CPA to OW-CPA}
+It is trivially true that if a public-key encryption scheme is not one-way secure, then it is not indistinguishable.
+The converse if also almost true, with the caveat that the message space must be sufficiently large:
+
+\begin{lemma}
+    Let $(\operatorname{KeyGen}, E, D)$ be a public-key encryption scheme with a finite message space $\mathcal{M}$. For every IND-CPA adversary $A$, there exists an OW-CPA adversary $B$ such that
+
+    \begin{equation*}
+        \epsilon_\text{OW-CPA} = \epsilon_\text{IND-CPA} + \frac{1}{\norm{\mathcal{M}}}
     \end{equation*}
@@
 
-\begin{proof}
-    For each decryption query $(e_q, c_q)$, there are three mutually exclusive possibilities:
+\section{Related works}
+    Key encapsulation mechanisms (KEM) usually have weaker security requirements than public-key encryption schemes (PKE) thanks to the fact that properly designed KEMs obfuscate the plaintext through a key-derivation function (which is usually a hash function and thus a random oracle under the RO assumption) in the decryption routine.
+As a result, we can implement a simple IND-CCA2 KEM using an IND-CCA2 PKE:
+
+    \begin{multicols}{3}
+        \begin{algorithm}[H]
+            \SetAlgoLined
+            \caption{KeyGen}
+            $(\text{pk}, \text{sk}) \leftsample \operatorname{KeyGen}()$ \;
+            \Return{$(\text{pk}, \text{sk})$}
+        \end{algorithm}
+
+        \begin{algorithm}[H]
+            \SetAlgoLined
+            \caption{Encap}
+            \KwIn{$\text{pk}$}
+            $m \leftsample \mathcal{M}$\;
+            $c \leftsample E(\text{pk}, m)$\;
+            $K \leftarrow H(m, c)$\;
+            \Return{$(c, K)$}\;
+        \end{algorithm}
+
+        \begin{algorithm}[H]
+            \SetAlgoLined
+            \caption{Decap}
+            \KwIn{$\text{sk}, c \in \mathcal{C}$}
+            $\hat{m} \leftarrow D(\text{sk}, c)$\;
+            $\hat{K} \leftarrow H(\hat{m}, c)$\;
+            \Return{$\hat{K}$}\;
+        \end{algorithm}
+    \end{multicols}
+
+    It is straightforward to simulate the KEM game using a PKE adversary.
+The PKE adversary samples two random messages $(m_0, m_1) \leftsample \mathcal{M}$ and obtains the challenge encryption for one of them $c^\ast \leftarrow E(\text{pk}, m_b)$.
+The PKE adversary then computes $K^\ast \leftarrow H(c^\ast, m_0)$ and passes $(c^\ast, K^\ast)$ as the challenge encapsulation to the KEM adversary.
+If $c^\ast$ is an encryption of $m_0$, then $K^\ast$ is pseudorandom, otherwise under the random oralce model, $K^\ast$ is truly random from the perspective of the KEM adversary.
+Therefore, $\epsilon_\text{PKE} = \epsilon_\text{KEM}$.
+
+    Using IND-CCA2 PKE as an IND-CCA2 KEM has been in production for a long time.
+For example, TLS 1.2 \cite{dierks2008rfc} offers this implementation for key exchange using RSA-OAEP\cite{bellare1995optimal} as the underlying IND-CCA2 PKE.
+The key exchange protocol proceeds as follows:
 
     \begin{enumerate}
-        \item The queried ciphertext is \textbf{honest}, meaning that there is a matching record on the tape of the hash function $H$
-        \item There is no matching record on the tape of the hash function $H$, and the check in step 4 in algorithm \ref{fo-key-dec} will fail, outputing decryption error.
-Such a query is called \textbf{invalid}
-        \item There is no matching record on the tape of the hash function $H$, but the check in step 4 of algorithm \ref{fo-key-dec} will succeed. Such a query is called \textbf{almost valid}
+        \item Server sends its public key through the X.509 certificate
+        \item Client samples a master secret, encrypts the secret using server's public key, then sends the ciphertext to the server
+        \item Server decrypts the ciphertext to obtain the master secret
     \end{enumerate}
 
-    Observe that for both $S_0, S_1$:
+    This approach to key exchange has since been deprecated, and RSA is no longer supported for key exchange starting with TLS 1.3.
+In addition to the difficulty of implementing RSA correctly and securely (RSA is notoriously hard to implement without side-channel vulnerabilities), using the server's long-term keypair for establishing common secrets fails to provide forward secrecy, meaning that if the long-term key is compromised, all prior communications can be trivially decrypted.
+
+    While the modern approach is to use ephemeral Diffie-Hellman (with either prime field or elliptic curve), it is difficult to adapt this approach to the post-quantum setting.
+Instead, post-quantum key exchange usually builds an IND-CCA2 KEM, then perform the key exchange using ephemeral keypairs.
+At each session's handshake, the client generates the a new keypair and sends the public key to the server.
+The server uses the public key to run the encapsulation routine and sends back the ciphertext.
+Finally, the client runs the decapsulation routine on the ciphertext to recover the shared secret.
+
+\section{Fujisaki-Okamoto transformation and variations}
+In 1999, Fujisaki and Okamoto \cite{fujisaki1999secure} proposed a generic transformation that outputs an IND-CCA2-secure public-key encryption scheme using an IND-CPA symmetric cipher, an OW-CPA asymmetric cipher, and two hash functions.
+Later on in 2017, Hofheinz, Hovelmann, and Kiltz \cite{hofheinz2017modular} improved on the security of the Fujisaki-Transformation and adapted the original approach to constructing IND-CCA2 KEMs.
+In this section, we will give a quick overview of the individual transformations and their security results.
+We will then discuss the techniques used in the security reduction.
+Finally, we will provide a sketch of the actual proof.
+
+\subsection{An overview of transformed routines and security results}
+The original 1999 Fujisaki-Okamoto transformation takes as input a symmetric cipher $(E^\text{sym}, D^\text{sym})$, an asymmetric cipher $(\operatorname{KeyGen}, E^\text{asym}, D^\text{asym})$, a key derivation function (aka a hash function) $G: \mathcal{M}^\text{asym} \rightarrow \mathcal{K}^\text{sym}$, and a hash functino $H: \{0, 1\}^\ast \rightarrow \text{Coin}^\text{asym}$.
+
+Specifically, the asymmetric cipher is assumed to be probabilistic, but with pseudorandomness seeded on some coin $r \in \text{Coin}^\text{asym}$.
+When $E^\text{asym}$ is called without specifying a value for coin, it is assumed that a coin is uniformly sampled from the coin space and passed in as the pseudorandom seed:
 
 $$
-    \begin{aligned}
-        P[S] &= P[S \cap \text{ all decryption queries are honest }] \\
-        &+ P[S \cap \text{ some decryption queries are dishonest, but none is almost valid }] \\
-        &+ P[S \cap \text{ at least one almost valid decryption queries }]
-    \end{aligned}
+    c \leftsample E^\text{asym}(\text{pk}, m) 
+    \Leftrightarrow 
+    c \leftarrow E^\text{asym}(\text{pk}, m, r \leftsample \text{Coin})
 $$
 
-    When all decryption queries are honest, the decryption oracles will correctly decrypt the query in both game 0 and game 1.
-When all dishonest decryption queries are invalid, the decryption oracles will reject the query in both games.
-The only difference between the two games lies in how the decryption oracle processes almost valid decryption queries.
-Therefore:
+The output of the transformation is a public-key encryption scheme $(\operatorname{KeyGen}, E^\text{hy}, D^\text{hy})$.
+Note that the key generation routine is exactly the same as the input asymmetric cipher's routine, so no distinction of notation is needed.
+The encryption and decryption routines are as follows:
 
-    $$
-    \begin{aligned}
-        &P[S_0] - P[S_1] \\
-        &= P[S_0 \cap \text{ at least one almost valid decryption queries }] \\
-        &- P[S_1 \cap \text{ at least one almost valid decryption queries }] \\
-        &\leq P[\text{ at least one almost valid decryption queries }]
-    \end{aligned}
-    $$
-
-    Let $(e, c)$ be some decryption query made without querying $H$, then in the true decryption routine, $\hat{h} \leftarrow H(\hat{\sigma}, c)$ will be a truly random coin, and $\hat{e} \leftarrow E^\text{asym}(\hat{\sigma}, \hat{h})$ will be a truly random ciphertext for the given public key and $\hat{\sigma}$.
-Since the the PKE has $\gamma$ spread, we know that $P[e = \hat{e}] = P[(e, c) \text{ is almost valid}] \leq 2^{-\gamma}$.
-Among $q_D$ decryption query, the probability of having at least one almost valid query is bounded by sum of probability of each decryption query being almost valid: $P[\text{ at least one almost valid query }] \leq q_D 2^{-\gamma}$.
-\end{proof}
-
-If during the IND-CCA game, $\mathcal{A}^\text{hy}_\text{IND-CCA}$ never makes hash query that involves $\sigma^\ast$, then under the random oracle model, there is no difference between sampling $a^\ast, h^\ast$ randomly or pseudorandomly.
-On the other hand, if the adversary does make such a query, then there is an inconsistency between the challenge ciphertext and the results of such hash query.
-In other words, the difference between game 1 and game 2 is the even that the IND-CCA adversary makes such a hash query:
-
-\begin{lemma}
-    $$
-    P[S_1] - P[S_2] = P[\text{query}]
-    $$
-\end{lemma}
-
-Denote the challenge ciphertext in game 2 by $(e^\ast, c^\ast)$.
-Notice that $e^\ast$ is the encryption of a truly random $\sigma^\ast$ using a truly random coin $h^\ast$, while $c^\ast$ is the encryption of $m_b^\ast$ under a truly random key $a^\ast$.
-This allows a OW-CPA game for the PKE, and an IND-CPA game for the symmetric cipher to be perfectly simulated by some $\mathcal{A}^\text{asym}_\text{OW-CPA}$ and/or $\mathcal{A}^\text{sym}_\text{IND-CPA}$.
-Using this strategy, we can bound $P[S_2]$ and $P[\text{query}]$ using the advantage of some OW-CPA and/or IND-CPA adversaries.
-
-\begin{lemma}\label{fo-bound-s2}
-    For every IND-CCA adversary $A^\text{hy}_\text{IND-CCA}$, there exists an IND-OTE adversary against the underlying symmetric cipher $\mathcal{A}^\text{sym}_\text{IND-OTE}$ with advantage $\epsilon^\text{sym}_\text{IND-OTE}$ such that
-
-    $$
-    P[S_2] = \frac{1}{2} + \epsilon^\text{sym}_\text{IND-OTE}
-    $$
-\end{lemma}
-
-\begin{proof}
-    $\mathcal{A}^\text{sym}_\text{IND-OTE}$ can perfectly simulate the hybrid key generation $\text{PK}^\text{hy} = \text{PK}^\text{asym}, \text{SK}^\text{hy} = \text{SK}^\text{asym}$, the random oracles $G, H$, as well as the modified decryption oracle.
-
-    When $\mathcal{A}^\text{hy}_\text{IND-CCA}$ submits the challenge plaintexts $(m_0, m_1)$, $\mathcal{A}^\text{sym}_\text{IND-OTE}$ passes them to the symmetric cipher challenger and receives the symmetric challenge ciphertext $c^\ast$.
-$\mathcal{A}^\text{sym}_\text{IND-OTE}$ then randomly samples $\sigma^\ast \leftsample \mathcal{M}^\text{asym}$ and $h^\ast \leftsample \text{COIN}^\text{asym}$ and computes $e^\ast \leftarrow E^\text{asym}(\text{PK}^\text{asym}, \sigma^\ast, h^\ast)$.
-$(e^\ast, c^\ast)$ is given to $\mathcal{A}^\text{hy}_\text{IND-CCA}$ as the challenge ciphertext.
-It is easy to verify that from $\mathcal{A}^\text{hy}_\text{IND-CCA}$'s perspective, this game is exactly game 2; in addition $\mathcal{A}^\text{hy}_\text{IND-CCA}$ wins the game if and only if $\mathcal{A}^\text{sym}_\text{IND-OTE}$ wins the game.
-\end{proof}
-
-\begin{lemma}
-    For every IND-CCA adversray $\mathcal{A}^\text{hy}_\text{IND-CCA}$, there exists an OW-CPA adversary $\mathcal{A}^\text{asym}_\text{OW-CPA}$ with advantage $\epsilon^\text{asym}_\text{OW-CPA}$ such that
-
-    $$
-    \epsilon^\text{asym}_\text{OW-CPA} = P[\text{query}] \cdot \frac{1}{q_H}
-    $$
-
-    where $q_H$ is the number of hash queries $\mathcal{A}^\text{hy}_\text{IND-CCA}$ makes to either $H$ or $G$.
-\end{lemma}
-
-\begin{proof}
-    Similar to the proof of lemma \ref{fo-bound-s2}, $\mathcal{A}^\text{asym}_\text{OW-CPA}$ can perfectly simulate key generation, hash oracles, and decryption oracle.
-
-    Let $e^\ast$ denote the asymmetric challenge ciphertext $\mathcal{A}^\text{asym}_\text{OW-CPA}$ receives from the OW-CPA challenger, let $\sigma^\ast$ denote the plaintext that corresponds with $e^\ast$, and let $(m_0, m_1)$ denote the challenge plaintexts submitted by $\mathcal{A}^\text{hy}_\text{IND-CCA}$.
-$\mathcal{A}^\text{asym}_\text{OW-CPA}$ samples a random symmetric key $a^\ast \leftsample \mathcal{K}^\text{sym}$ and a coin flip $b \leftsample \{0, 1\}$, then computes $c^\ast \leftarrow E^\text{sym}_{a^\ast}(m_b)$.
-$(e^\ast, c^\ast)$ is the IND-CCA challenge ciphertext.
-
-    After the $\mathcal{A}^\text{hy}_\text{IND-CCA}$ halts, $\mathcal{A}^\text{asym}_\text{OW-CPA}$ looks through the tape of the hash oracles $\mathcal{O}^H = \{(\tilde{\sigma}, \tilde{c})\}$ and $\mathcal{O}^G = \{\tilde{\sigma}\}$, then picks a random value among all possible $\tilde{\sigma}$'s to output.
-$\mathcal{A}^\text{asym}_\text{OW-CPA}$ wins if $\mathcal{A}^\text{hy}_\text{OW-CPA}$ makes a query $\tilde{\sigma} = \sigma^\ast$ and the correct query is chosen.
-\end{proof}
-
-\section{Tweaking FO Transform for IND-CCA KEM}
-While the FO transformation can construct IND-CCA public-key encryption scheme from OW-CPA PKE and IND-OTE symmetric cipher, there are a few drawbacks to the hybrid scheme as it is:
-
-\begin{itemize}
-    \item The security proof is not tight. The security of the hybrid scheme degrades linearly with the number of hash queries.
-    \item The hybrid scheme assumes the underlying PKE to be always correct.
-This is not the case in PKEs such as schemes based on "Learning with Errors" (LWE), which can have a small but non-zero probability of decryption error
-\end{itemize}
-
-In 2017, Hofheinz et al \cite{hofheinz2017modular} improved the FO transformation by addressing the non-tight security and accounting for possible decryption errors.
-In addition, the paper proposed various IND-CCA KEM constructions from weak PKE's that does not require a symmetric cipher.
-In fact, both Kyber \cite{avanzi2019crystals} and Classic McEliece \cite{albrecht2022classic} make direct use of transformations proposed in this paper to achieve IND-CCA security for their respective KEMs.
-
-The strategy for achieving IND-CCA KEM consists of two parts
-
-\begin{itemize}
-    \item Transform a OW-CPA and/or IND-CPA PKE into an OW-PCVA PKE
-    \item Transform the OW-PCVA PKE into a IND-CCA KEM
-\end{itemize}
-
-The intermediary security definition PCVA is defined by two oracles: the \textbf{plaintext checking oracle (PCO)} and the \textbf{ciphertext validation oracle (CVO)}.
-The details of these two oracles will be discussed in the following section
-
-\subsection{From OW/IND-CPA to OW-PCVA}
-Let $(\text{KeyGen}, E, D)$ denote the input public-key encryption scheme, let $(\text{KeyGen}^T, E^T, D^T)$ denote the transformed public-key encryption scheme, let $G$ be a hash function.
-The transformed routines are as follows:
-
-\begin{algorithm}
-    \caption[keygen-t]{$\text{KeyGen}^T$}
-    \begin{algorithmic}[1]
-        \State $\text{pk}, \text{sk} \leftsample \text{KeyGen}()$
-        \State \Return $\text{pk}^T \leftarrow \text{pk}, \text{sk}^T \leftarrow \text{sk}$
-    \end{algorithmic}
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Hybrid encryption $E^\text{hy}$}
+        \KwIn{$\text{pk}, m \in \mathcal{M}^\text{sym}$}
+        $\sigma \leftsample \mathcal{M}^\text{asym}$\;
+        $a \leftarrow G(\sigma)$; $c \leftarrow E^\text{sym}_a(m)$\;
+        $r \leftarrow H(\sigma, c)$;
+        $e \leftarrow E^\text{asym}(\text{pk}, m, r)$\;
+        \Return{$(c, e)$}\;
     \end{algorithm}
 
-\begin{algorithm}
-    \caption[encrypt-t]{Encryption routine: $E^T$}
-    \begin{algorithmic}[1]
-        \Require The message $m \in \mathcal{M}$
-        \State $r \leftarrow G(m)$ \Comment $r$ serves as the coin
-        \State $c \leftarrow E(\text{pk}^T, m, r)$
-        \State \Return $c$
-    \end{algorithmic}
+    \columnbreak
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Hybrid decryption $D^\text{hy}$}
+        \KwIn{$\text{sk}, c \in \mathcal{C}^\text{sym}, e \in \mathcal{C}^\text{asym}$}
+        $\hat{\sigma} \leftarrow D^\text{asym}(\text{sk}, e)$\;
+        $\hat{r} \leftarrow H(\hat{\sigma}, c)$\;
+        \If{
+            $E^\text{asym}(\text{pk}, \hat{\sigma}, \hat{r}) \neq e$
+        }{
+            \Return{$\bot$}; \tcp*[h]{the "re-encryption"}
+        }
+        $\hat{a} \leftarrow G(\hat{\sigma})$\;
+        $\hat{m} \leftarrow D^\text{sym}_{\hat{a}}(c)$\;
+        \Return{$\hat{m}$}\;
+    \end{algorithm}
+\end{multicols}
+
+\begin{theorem}[Security result for $(E^\text{hy}, D^\text{hy})$]
+    If the input asymmetric cipher has $\gamma$ spread, then for every IND-CPA adversary $\mathcal{A}^\text{sym}_\text{IND-CPA}$ against the underlying symmetric cipher with advantage $\epsilon^\text{sym}$ and OW-CPA adversary $\mathcal{A}^\text{asym}_\text{OW-CPA}$ against the input asymmetric cipher with advantage $\epsilon^\text{asym}$, there exists an IND-CCA2 adversary with against the hybrid scheme making $q_H$ hash queries and $q_D$ decryption queries such that:
+
+    \begin{equation*}
+        \epsilon^\text{hy} 
+        \leq q_D2^{-\gamma} 
+            + \epsilon^\text{sym} 
+            + q_H\epsilon^\text{asym}
+    \end{equation*}
+\end{theorem}
+
+Hofheinz, Hovenmann, and Kiltz pointed out in their 2017 paper a number of drawbacks of the original FO transformation:
+
+\begin{enumerate}
+    \item Decryption error, which is especially possible with lattice-based schemes, is not accounted for
+    \item the $q_H\epsilon^\text{asym}$ term causes security to be non-tight, which have implications when choosing the security parameters
+\end{enumerate}
+
+In their proposals, IND-CCA2 security is achieved through a two-step transformation.
+First, an OW-CPA asymmetric scheme is transformed into an OW-PCVA asymmetric scheme.
+The OW-PCVA scheme is then transformed into an IND-CCA2 KEM.
+Here PCVA refers to the adversary's ability to access some plaintext-checking oracle and some ciphertext validation oracle.
+Although PCVA is a non-standard security definition, it is entirely contained within the two-step transformation and serves only as an intermediary step.
+
+Neither the OW-PCVA nor the IND-CCA transformation makes use of any symmetric cipher, so we ditch the distinction.
+The OW-PCVA transformation takes as input an asymmetric cipher $(E, D)$ and a hash function $G: \mathcal{M} \rightarrow \text{Coin}$.
+The transformed encryption and decryption routines $(E^T, D^T)$ are as follows:
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{OW-PCVA $E^T$}
+        \KwIn{$\text{pk}, m \in \mathcal{M}$}
+        $r \leftarrow G(m)$\;
+        $c \leftarrow E(\text{pk}, m, r)$\;
+        \Return{$c$}\;
     \end{algorithm}
 
-\begin{algorithm}
-    \caption[decrypt-t]{Decryption routine: $D^T$}
-    \begin{algorithmic}[1]
-        \Require ciphertext $c \in \mathcal{C}$
-        \State $\hat{m} \leftarrow D(\text{sk}, c)$
-        \State $\hat{r} \leftarrow G(\hat{m})$
-        \If{$E(\text{pk}, \hat{m}, \hat{r}) \neq c$}
-            \State \Return $\bot$
-        \EndIf
-        \State \Return $\hat{m}$
-    \end{algorithmic}
+    \columnbreak
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{OW-PCVA $D^T$}
+        \KwIn{$\text{sk}, c \in \mathcal{C}$}
+        $\hat{m} \leftarrow D(\text{sk}, c)$\;
+        $\hat{r} \leftarrow G(\hat{m})$\;
+        \If{
+            $E(\text{pk}, \hat{m}, \hat{r}) \neq c$
+        }{
+            \Return{$\bot$}\;
+        }
+        \Return{$\hat{m}$};
     \end{algorithm}
+\end{multicols}
 
-To discuss the security of the transformed scheme, we first need to lay out the implementations of the PCO and the CVO.
-In the vanilla PCVA game, the two oracles have access to the challenge secret key and perform decryption routines:
+Depending on whether the implementation chooses implicit or explicit rejection of invalid ciphertext for the final KEM, the second transformation has slightly different key generation and decryption routine.
+The explicit rejection KEM is denoted by $U^\bot$, and its key generation is identical to the input asymmetric cipher.
 
-\begin{algorithm}
-    \caption[pco-0]{Plaintext checking oracle $\text{PCO}_0$}
-    \begin{algorithmic}[1]
-        \Require A query plaintext-ciphertext pair $(\tilde{m}, \tilde{c})$
-        \State $\hat{m} \leftarrow D(\text{sk}, \tilde{c})$
-        \If{$\hat{m} \neq \tilde{m}$}
-            \State \Return 0 \Comment query is not valid
-        \EndIf
-        \State $\tilde{r} \leftarrow G(\tilde{m})$
-        \If{$E(\text{pk}, \tilde{m}, \tilde{r}) \neq \tilde{c}$}
-            \State \Return 0  \Comment query is also not valid
-        \EndIf
-        \State \Return 1  \Comment query is valid
-    \end{algorithmic}
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{$U^\bot$ Encap}
+        \KwIn{$\text{pk}$}
+        $m \leftsample \mathcal{M}$\;
+        $c \leftarrow E^T(\text{pk}, m)$\;
+        $K \leftarrow H(m, c)$ \tcp*[h]{H is a hash function}\;
+        \Return{$(c, K)$}
     \end{algorithm}
+    \columnbreak
 
-\begin{algorithm}
-    \caption[cvo-0]{Ciphertext validation oracle $\text{CVO}_0$}
-    \begin{algorithmic}[1]
-        \Require A query ciphertext $\tilde{c}$
-        \State $\hat{m} \leftarrow D(\text{sk}, \tilde{c})$
-        \If{$\hat{m} = \bot$}
-            \State \Return 0 \Comment ciphertext is not valid
-        \ElsIf{$E(\text{pk}, \hat{m}, G(\hat{m})) \neq \tilde{c}$}
-            \State \Return 0 \Comment ciphertext is not valid
-        \EndIf
-        \State \Return 1  \Comment ciphertext is valid
-    \end{algorithmic}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{$U^\bot$ Decap}
+        \KwIn{$c \in \mathcal{C}, \text{sk}$}
+        $\hat{m} \leftarrow D^T(\text{sk}, c)$\;
+        \If{
+            $m = \bot$
+        }{
+            \Return{$\bot$};
+        }
+        \Return{$H(\hat{m}, c)$}
     \end{algorithm}
-\pagebreak
+\end{multicols}
 
-To link the advantage of an OW-PCVA adversary, a sequence of games is devised in which PCO and CVO are incrementally modified to not require the secret key and the challenge encryption to be performed with a truly random coin.
-These modifications will allow the OW-PCVA game to be perfectly simulated by an OW/IND-CPA adversary, thus establishing the security of the transformed encryption scheme upon the security of the input encryption scheme.
+On the other hand, the implicit rejection KEM, denoted by $U^{\not\bot}$, generate a random message $m^\prime \leftsample \mathcal{M}$ as the fake input to $H$ when the ciphertext is invalid
 
-\begin{theorem}[IND-CPA to OW-PCVA]\label{cpa-to-pcva}
-    Under the random-oracle model, for every OW-PCVA adversary $\mathcal{A}^T_\text{OW-PCVA}$ against the transformed PKE with advantage $\epsilon^T_\text{OW-PCVA}$, there exists an IND-CPA adversary $\mathcal{A}_\text{IND-CPA}$ against the input PKE with advantage $\epsilon_\text{IND-CPA}$ such that
+\begin{multicols}{3}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{KeyGen}
+        $\text{pk}, \text{sk} \leftsample \operatorname{KeyGen}()$\;
+        $m^\prime \leftsample \mathcal{M}$\;
+        \Return{$(\text{pk}, (\text{sk}, m^\prime))$}\;
+    \end{algorithm}
+    \columnbreak
 
-    $$
-    \epsilon^T_\text{OW-PCVA} 
-        \leq q_V \cdot 2^{-\gamma} 
-            + q_G \cdot \delta 
-            + 3\cdot\epsilon_\text{IND-CPA}
-    $$
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Encap}
+        $m \leftsample \mathcal{M}$\;
+        $c \leftarrow E^T(\text{pk}, m)$\;
+        $K \leftarrow H(m, c)$\;
+        \Return{$(c, K)$}
+    \end{algorithm}
+    \columnbreak
 
-    Where $\gamma$ is the spread of ciphertext of the input PKE, $\delta$ is the probability of decryption error of the input PKE, $q_V$ is the number of ciphertext validation queries, and $q_G$ is the number of hash queries to $G$.
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Decap}
+        $\hat{m} \leftarrow D^T(\text{sk}, c)$\;
+        \If{
+            $m = \bot$
+        }{
+            \Return{$H(m^\prime, c)$};
+        }
+        \Return{$H(\hat{m}, c)$}
+    \end{algorithm}
+\end{multicols}
+
+The security result is similarly expressed in two theorems:
+
+\begin{theorem}
+    If the input asymmetric scheme is $\delta$-correct and $\gamma$-spread, then for every OW-PCVA adversary that issues at most $q_G$ hash queries and $q_V$ ciphertext validation queries with advantage $\epsilon^T$, there exists an IND-CPA adversary against the underlying asymmetric scheme with advantage $\epsilon$ such that
+
+    \begin{equation*}
+        \epsilon^T \leq q_G \cdot \delta + q_V \cdot 2^{-\gamma} 
+            + \frac{2q_G + 1}{\norm{\mathcal{M}}}
+            + 3\epsilon
+    \end{equation*}
 \end{theorem}
 
-\begin{proof}\label{cpa-to-pcva-proof}
-    \textit{Admittedly this is only a sketch. Some nuance is omitted for the sake of clarity, but the overall result will be off by no more than some negligible constants.}
+\begin{theorem}
+    For every IND-CCA2 adversary against the $U^\bot$ KEM with advantage $\epsilon^{U^\bot}$, there exists an OW-PCVA adversary against the input scheme $(E^T, D^T)$ with advantage $\epsilon^T$ such that
 
-    We device a sequence of games that $\mathcal{A}^T_\text{OW-PCVA}$ plays. Game 0 is the standard OW-PCVA game.
+    \begin{equation*}
+        \epsilon^{U^\bot} \leq \epsilon^T
+    \end{equation*}
+\end{theorem}
 
-    Game 1 is identical to game 0, except that CVO has been modified to check the tape of the hash function $G$ for plaintext $\tilde{m}$ that encrypts into the queried ciphertext $\tilde{c}$ under $E^T$:
+\begin{theorem}
+    For every IND-CCA2 adversary against the $U^{\not\bot}$ KEM that issues $q_H$ hash queries with advantage $\epsilon^{U^{\not\bot}}$, there exists an OW-PCVA adversary against the input scheme $(E^T, D^T)$ with advantage $\epsilon^T$ such that
 
-    \begin{algorithm}\label{cvo-1}
-        \caption{$\text{CVO}_1$}
-        \begin{algorithmic}[1]
-            \Require queried ciphertext $\tilde{c}$
-            \For{$(\tilde{m}, \tilde{r})$ on the tape of $G$}
-                \If{$E(\text{pk}, \tilde{m}, \tilde{r}) = \tilde{c}$} 
-                    \State \Return 1  \Comment queried ciphertext is valid
-                \EndIf
-            \EndFor
-            \State \Return 0  \Comment queried ciphertext is invalid
-        \end{algorithmic}
-    \end{algorithm}
+    \begin{equation*}
+        \epsilon^{U^{\not\bot}} \leq \epsilon^T + \frac{q_H}{\norm{\mathcal{M}}}
+    \end{equation*}
+\end{theorem}
 
-    Following similar logic in the FO transformation's security reduction, it's easy to see that game 0 and game 1 diverge when $\mathcal{A}^T_\text{OW-PCVA}$ makes at least one CVO query $\tilde{c}$ whose decryption encrypts correctly (aka passes CVO in game 0) but without querying $G$ (aka fails CVO in game 1)
+\subsection{A discussion of techniques}
 
-    \begin{equation}
-        P[\text{wins game 0}] - P[\text{wins game 1}] \leq q_V \cdot 2^{-\gamma}
-    \end{equation}
+\subsection{A sketch of proof}
 
-    Game 2 is identical to game 1, except that PCO is modified to check only the encryption instead of checking both encryption and decryption
-
-    \begin{algorithm}
-        \caption{$\text{PCO}_2$}
-        \begin{algorithmic}[1]
-            \Require query $(\tilde{m}, \tilde{c})$
-            \For{$(m, r)$ from the tape of $G$}
-                \If{$m = \tilde{m}$ and $E(\text{pk}, m, r) = \tilde{c}$}
-                    \State \Return 1
-                \EndIf
-            \EndFor
-            \State \Return 0
-        \end{algorithmic}
-    \end{algorithm}\pagebreak
-
-    Game 2 diverges from game 1 if $\mathcal{A}^T_\text{OW-PCVA}$ makes at least one hash query $(\tilde{m}, \tilde{r})$ such that
-
-    $$D(\text{sk}, E(\text{pk}, \tilde{m}, \tilde{r})) \neq \tilde{m}$$
-
-    Therefore
-
-    $$P[\text{wins game 1}] - P[\text{wins game 2}] \leq q_G \cdot \delta$$
-
-    Game 3 is identical to game 2, except that when encrypting the challenge ciphertext $c^\ast$, a truly random coin $(r^\ast \leftsample \text{Coin})$ is used instead of a pseudorandom coin $r^\ast \leftarrow G(m^\ast)$.
-Game 3 diverges from game 2 if $\mathcal{A}^T_\text{OW-PCVA}$ ever queries the hash of the challenge plaintext $m^\ast$:
-
-    $$
-    P[\text{wins game 2}] - P[\text{wins game 3}] \leq P[\text{query $m^\ast$}]
-    $$
-
-    In game 3, PCO and CVO can be simulated without the secret key, and the challenge ciphertext is computed using a truly random coin instead of the hash of the challenge plaintext.
-This allows game 3 to be perfectly simulated by some OW-CPA adversary $\mathcal{A}_\text{OW-CPA}$ against the input PKE:
-
-    \begin{itemize}
-        \item When $\mathcal{A}_\text{OW-CPA}$ receives $\text{pk}$, the public key is passed to $\mathcal{A}^T_\text{OW-PCVA}$
-        \item $\mathcal{A}_\text{OW-CPA}$ simulates $G$, $\text{PCO}_2$, $\text{CVO}_1$ for $\mathcal{A}^T_\text{OW-PCVA}$
-        \item When $\mathcal{A}_\text{OW-CPA}$ receives the challenge ciphertext $c^\ast$, it is passed to $\mathcal{A}^T_\text{OW-PCVA}$ as the OW-PCVA challenge ciphertext
-        \item When $\mathcal{A}^T_\text{OW-PCVA}$ outputs $m$, $\mathcal{A}_\text{OW-CPA}$ outputs the same value
-    \end{itemize}
-
-    It's easy to see that from $\mathcal{A}^T_\text{OW-PCVA}$'s perspective, it is playing game 3, and $\mathcal{A}^T_\text{OW-PCVA}$ wins if and only if $\mathcal{A}_\text{OW-CPA}$ wins:
-
-    $$
-    P[\text{$\mathcal{A}^T_\text{OW-PCVA}$ wins game 3}] = \epsilon_\text{OW-CPA} = \epsilon_\text{IND-CPA} + \frac{1}{\vert\mathcal{M}\vert}
-    $$
-
-    % TODO: bound P[query] by P[IND-CPA wins]
-
-\end{proof}
-
-\subsection{IND-CCA KEM}
-
-\section{OAEP and RSA-OAEP}
+\section{Future works on generic IND-CCA transformation}
 

### 2024-04-17 co789/project/manuscript/main.tex
Checkpoint for CO789 project

@@
 
-    \begin{algorithm}[H]
+    \begin{algorithm}[H]\label{hybrid-decryption-99}
         \SetAlgoLined
@@
 
-\begin{theorem}[Security result for $(E^\text{hy}, D^\text{hy})$]
+\begin{theorem}[Security result for $(E^\text{hy}, D^\text{hy})$]\label{fo-1999-theorem}
     If the input asymmetric cipher has $\gamma$ spread, then for every IND-CPA adversary $\mathcal{A}^\text{sym}_\text{IND-CPA}$ against the underlying symmetric cipher with advantage $\epsilon^\text{sym}$ and OW-CPA adversary $\mathcal{A}^\text{asym}_\text{OW-CPA}$ against the input asymmetric cipher with advantage $\epsilon^\text{asym}$, there exists an IND-CCA2 adversary with against the hybrid scheme making $q_H$ hash queries and $q_D$ decryption queries such that:
@@
 
-\begin{theorem}
+\begin{theorem}\label{ow-pcva-result}
     If the input asymmetric scheme is $\delta$-correct and $\gamma$-spread, then for every OW-PCVA adversary that issues at most $q_G$ hash queries and $q_V$ ciphertext validation queries with advantage $\epsilon^T$, there exists an IND-CPA adversary against the underlying asymmetric scheme with advantage $\epsilon$ such that
@@
 
-\begin{theorem}
+\begin{theorem}\label{cca-kem-result}
     For every IND-CCA2 adversary against the $U^\bot$ KEM with advantage $\epsilon^{U^\bot}$, there exists an OW-PCVA adversary against the input scheme $(E^T, D^T)$ with advantage $\epsilon^T$ such that
@@
 
-\subsection{A discussion of techniques}
+\subsection{A discussion of proof techniques}
+The 1999 Fujisaki-Okamoto transformation and its 2017 variations deployed similar strategies for their respective security reductions: start with the appropriate standard IND-CCA game, then gradually transform the game into something that can be simulated by some adversary playing another standard OW-CPA or IND-CPA game (or in the case of the IND-CCA KEM, some adversary playing the OW-PCVA game).
+
+In the standard games, the decryption oracle, the decapsulation oracle, the plaintext checking oracle, and the ciphertext validation oracles all behave as if they possess the true secret key and use the secret key to execute the decryption routine on their query inputs.
+However, thanks to the design of the transformed schemes, there are ways to simulate the behaviors of those oracles under the random oracle assumption by modeling the hash functions as probabilistic turing machines whose state is taped.
+The simulated oracles will behave differently from the true oracles in certain circumstances, so the transformed game will diverge from the original game, but because of the difference lemma, we can bound the loss of security by the probability of the diverging event.
+
+\subsubsection{Simulate decryption oracle with random oracle}\label{simulate-dec-oracle}
+The decryption oracle can be simulated without secret key by taking advantage of the fact that in an honest execution of the encryption routine, a hash of the message is used to generate the pseudorandom coin.
+Assuming that the random oracle records its queries, if the ciphertext is honestly generated, then the random oracle should contain exactly one record whose input contains the correct message.
+
+Here is an example of comparing the standard decryption oracle with the simulated decryption oracle in the 1999 Fujisaki-Okamoto transformation, where the hash function $H: \{0, 1\}^\ast \rightarrow \text{Coin}$ is modeled as a probabilistic turing machine and used to recover the plaintext
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Standard decryption oracle}
+        \KwIn{
+            $\tilde{c} \in \mathcal{C}^\text{sym}, 
+            \tilde{e} \in \mathcal{C}^\text{asym}$
+        }
+        \tcp*[h]{Check algorithm \ref{hybrid-decryption-99} for $D^\text{hy}$}
+        $\hat{m} \leftarrow D^\text{hy}(\text{sk}, (\tilde{c}, \tilde{e}))$\;
+        \Return{$\hat{m}$}
+    \end{algorithm}
+
+    \columnbreak
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Simulated decryption oracle}
+        \KwIn{
+            $\tilde{c} \in \mathcal{C}^\text{sym}
+            , \tilde{e}\in\mathcal{C}^\text{asym}$
+        }
+        \If{
+            $\exists (\sigma, c, r) \in \mathcal{O}^H$ such that
+            $c = \tilde{c}$ 
+            and $E^\text{asym}(\text{pk}, \sigma, r) = \tilde{e}$
+        }{
+            \tcp{We have found a decryption}
+            $\hat{a} \leftarrow G(\sigma)$\;
+            $\hat{m} \leftarrow D^\text{sym}_{\hat{a}}(c)$\;
+            \Return{$\hat{m}$}
+        }
+        \Return{$\bot$}\;
+    \end{algorithm}
+\end{multicols}
+
+The two oracles will behave differently if and only if the queried ciphertext is generated without querying the hash oracle, but if the hash oracle is not consulted, then the coin is indistinguishable from truly random.
+Therefore, the simulated oracle's behavior diverges from the standard oracle if and only if the adversary correctly guesses the ciphertext generated from a truly random coin.
+The probaility of such an event is bounded by the spread of the underlying encryption scheme.
+
+\subsubsection{Random until queried}\label{random-until-queried}
+In the 1999 transformation, both the symmetric key $a \leftarrow G(\sigma)$ and the asymmetric coin $r \leftarrow H(\sigma, c)$ are pseudorandom.
+In the 2017 transformation, the asymmetric coin $r \leftarrow H(m)$ is also pseudorandom.
+On the other hand, in the standard security game against any symmetric cipher, the secret key is truly random, and in the standard security game against an asymmetric cipher, the encryption coin is also truly random.
+This difference prevents an adversary playing the standard game to correctly simulate the challenge encryption in the IND-CCA2 game.
+
+Under the random oracle model, this difference can be resolved by observing that there is no distinction between querying a random oracle once and randomly sampling from the specified probability space.
+In other words, unless the IND-CCA2 adversary makes a query to the random oracle with the specific input, from the IND-CCA2 adversary's perspective there is no difference between pseudorandom and truly random values, so the sequence of games can safely replace pseudorandom values with truly random values, and the difference in advantage will be bounded by the probability that the IND-CCA2 adversary indeed makes some very specific query.
+Finally, this probability can be bounded by a standard game adversary who can take advantage of the specific value being queried.
+
+As an example, in the 1999 transformation's security reduction, consider an IND-CPA adversary against the underlying symmetric cipher $\mathcal{A}^\text{sym}_\text{IND-CPA}$.
+$\mathcal{A}^\text{sym}_\text{IND-CPA}$ tries to simulate the IND-CCA2 game for an adversary against the hybrid scheme $\mathcal{A}^\text{hy}_\text{IND-CCA}$ in the following steps:
+
+\begin{enumerate}
+    \item $\mathcal{A}^\text{sym}_\text{IND-CPA}$ runs hybrid KeyGen and passes pk to $\mathcal{A}^\text{hy}_\text{IND-CCA}$
+    \item $\mathcal{A}^\text{sym}_\text{IND-CPA}$ simulates random oracle $H, G$ and the decryption oracle $\mathcal{O}^D$ using method described in section \ref{simulate-dec-oracle}
+    \item When $\mathcal{A}^\text{hy}$ outputs the challenge messages $m_0, m_1$, $\mathcal{A}^\text{sym}$ passes these two messages to the symmetric cipher challenger.
+The symmetric challenger returns the encryption of one of them $c^\ast = E^\text{sym}_{a^\ast}(m_b)$ under a \textbf{truly random key $a^\ast$}
+    \item $\mathcal{A}^\text{sym}$ samples random $\sigma \leftarrow \mathcal{M}^\text{asym}$, computes the coin $r \leftarrow H(c^\ast, \sigma)$ and the asymmetric ciphertext $e^\ast \leftarrow E^\text{asym}(\text{pk}, \sigma, r)$
+    \item $\mathcal{A}^\text{sym}$ passes $(c^\ast, e^\ast)$ as the challenge ciphertext to $\mathcal{A}^\text{hy}$.
+When $\mathcal{A}^\text{hy}$ outputs the guess $\hat{b}$, $\mathcal{A}^\text{sym}$ passes $\hat{b}$ as its own output
+\end{enumerate}
+
+Note that in the simulation of the IND-CCA2 game above, the symmetric key for the challenge encryption in the IND-CCA2 game is truly random instead of pseudorandom.
+Let $G_a$ denote the IND-CCA2 game with truly random symmetric key (aka the simulation above), let $G_b$ denote the IND-CCA2 game with pseudorandom symmetric key, and let $\operatorname{QUERY}^\ast$ denote the event that the IND-CCA2 adversary queries $G$ with $m_b$, which is the diverging event between the two games.
+By the difference lemma:
+
+\begin{equation*}
+    \epsilon_a - \epsilon_b \leq P[\operatorname{QUERY}^\ast]
+\end{equation*}
+
+On the other hand, if the adversary indeed queries the hash $G(m_b)$, then the tape of the random oracle $G$ should contain an entry that matches one of $(m_0, m_1)$, which can then be used by another IND-CPA adversary:
+
+\begin{equation*}
+    P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}
+\end{equation*}
+
+\subsubsection{Plaintext checking, ciphertext validation, and patched decapsulation oracle}
+The concept of plaintext-checking attack (PCA) was proposed in early 2000s and first used in generic IND-CCA transformation by Coron et al in \textit{GEM: a generic chosen-ciphertext secure encryption method} \cite{jean2002gem}.
+Although PCA security is not a standard security definition, it is nonetheless useful as an intermediary step for constructing simple generic IND-CCA secure cryptosystems.
+
+The standard plaintext-checking oracle (PCO) takes as input a plaintext-ciphertext pair $(m, c)$ and uses the secret key to determine whether they correspond to each other.
+Because lattice-based cryptosystems sometimes suffer from decryption errors, and because the OW-PCVA transformation $(E^T, D^T)$ is a deterministic encryption scheme, the standard PCO checks both encryption and decryption.
+When trying to simulate PCO without using the secret key, the decryption routine is omitted.
+Thanks for $E^T$ being deterministic, the behavior of the simulated PCO diverges from the standard PCO if and only if $m$ causes decryption error.
+The probability of decryption error is bounded by $\delta$-correctness.
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Standard PCO}
+        \KwIn{$(m \in \mathcal{M}, c \in \mathcal{C})$}
+        \If{
+            $E^T(\text{pk}, m) \neq c$
+        }{
+            \Return{$\bot$}\;
+        }
+        \If{
+            $D^T(\text{sk}, c) \neq m$
+        }{
+            \Return{$\bot$}\;
+        }
+    \end{algorithm}
+
+    \columnbreak
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Simulated PCO}
+        \KwIn{$(m \in \mathcal{M}, c \in \mathcal{C})$}
+        \If{
+            $E^T(\text{pk}, m) \neq c$
+        }{
+            \Return{$\bot$}\;
+        }
+    \end{algorithm}
+\end{multicols}
+
+In addition to PCO, Hofheinz \cite{hofheinz2017modular} also introduced a ciphertext-validation oracle (CVO) that checks whether the queried ciphertext is valid or not.
+The standard CVO uses the secret key to execute the decryption routine, and the simulated CVO uses the random oracle in a similar trick as described in section \ref{simulate-dec-oracle} with a similar loss of security.
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Standard CVO}
+        \KwIn{
+            $c \in \mathcal{C}$
+        }
+        $\hat{m} \leftarrow D^T(\text{sk}, c)$\;
+        \If{
+            $\hat{m} = \bot$
+        }{
+            \Return{$\bot$}
+        }
+        \If{
+            $E^T(\text{pk}, \hat{m}) \neq c$
+        }{
+            \Return{$\bot$}
+        }
+    \end{algorithm}
+    \columnbreak
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Simulated CVO}
+        \KwIn{
+            $c \in \mathcal{C}$
+        }
+        \If{
+            $\not\exists (m, r) \in \mathcal{O}^G$ such that
+            $E^T(\text{pk}, m) = c$
+        }{
+            \Return{$\bot$}
+        }
+    \end{algorithm}
+\end{multicols}
+
+The combination of PCO and CVO allows the decapsulation oracle to be simulated by an OW-PCVA adversary having access to these two oracles.
+The simulation is achieved by making both the hash oracle $H$ and the simulated decapsulation oracle $\mathcal{O}^D$ stateful, and allowing valid queries to $H$ be added to the state of $\mathcal{O}^D$ so that the output of $\mathcal{O}^D$ stays consistent.
+In this simulation $H$ \textbf{patches} the state of $\mathcal{O}^D$, hence we call this decapsulation ``patched".
+
+\pagebreak
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Standard decap oracle}
+        \KwIn{
+            $c \in \mathcal{C}$
+        }
+        $\hat{m} \leftarrow D^T(\text{sk}, c)$\;
+        \If{
+            $\hat{m} \neq \bot$
+        }{
+            \Return{$H(\hat{m}, c)$}\;
+        }
+        \Return{$\bot$}\;
+    \end{algorithm}
+    \columnbreak
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Patched hash oracle $H$}
+        \KwIn{
+            $m \in \mathcal{M}, c \in \mathcal{C}$
+        }
+        \If{
+            $\exists (
+                \tilde{m}, \tilde{c}, \tilde{K}
+            ) \in \mathcal{O}^H$ such that $\tilde{m} = m, \tilde{c} = c$
+        }{
+            \Return{$\tilde{K}$}
+        }
+        $K \leftsample \mathcal{K}$\;
+        Append $(m, c, K)$ to $\mathcal{O}^H$\;
+        \If{
+            $\operatorname{PCO}(m, c) \neq bot$
+        }{
+            Append $(c, K)$ to $\mathcal{O}^D$
+        }
+        \Return{$K$}
+    \end{algorithm}
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{Patched decap oracle $\mathcal{O}^D$}
+        \KwIn{$c \in \mathcal{C}$}
+        \If{
+            $\exists (\tilde{c}, \tilde{K}) \in \mathcal{O}^D$
+            such that $\tilde{c} = c$
+        }{
+            \Return{$\tilde{K}$}
+        }
+        \If{
+            $\operatorname{CVO}(c) \neq \bot$
+        }{
+            $K \leftsample \mathcal{K}$\;
+            Append $(c, K)$ to $\mathcal{O}^D$\;
+            \Return{$K$}
+        }
+        \Return{$\bot$}
+        \tcp*[alignment]{Or the implicit rejection}
+    \end{algorithm}
+\end{multicols}
+
+Under the random oracle model, the combination of patched hash function and patched decapsulation oracle should behave exactly the same as the standard decapsulation oracle.
 
 \subsection{A sketch of proof}
+Here we provide a sketch of the security proof for both the 1999 transformation and the 2017 variations.
+
+\subsubsection{Proof of theorem \ref{fo-1999-theorem}}
+\begin{proof}
+
+\end{proof}
+
+\subsubsection{Proof of theorem \ref{ow-pcva-result}}
+
+\subsubsection{Proof of theorem \ref{cca-kem-result}}
 

### 2024-04-26 lattices/primitives/src/lib.rs
Kyber polynomial and its coefficient representation

@@ -1,0 +1,82 @@
+//! Primitives used in lattice-based cryptography
+#![no_std]
+
+use core::fmt::{LowerHex, UpperHex};
+
+/// A member of the polynomial ring used in Kyber:
+/// R_q = Z_q[x] / (x ** n + 1)
+/// where q is 3329 = 2 ** 8 * 13 + 1 and n is 256
+#[derive(Copy, Clone, PartialEq)]
+pub struct Poly {
+    /// Each member is uniquely determined by its coefficients
+    /// Coefficient at lower index correspond to term with lower power
+    ///
+    /// NOTE: something about using a larger uint type than strictly necessary (3329 < 2 ** 16)
+    /// because it's easier to do modular arithmetic with
+    coeffs: [u32; 256],
+}
+
+impl Poly {
+    pub fn from_coeffs(coeffs: [u32; 256]) -> Self {
+        Self { coeffs }
+    }
+
+    pub fn as_coeffs(&self) -> &[u32] {
+        &self.coeffs
+    }
+}
+
+impl core::fmt::Debug for Poly {
+    fn fmt(&self, f: &mut core::fmt::Formatter<'_>) -> core::fmt::Result {
+        write!(f, "Poly {{")?;
+        for (i, coeff) in self.as_coeffs().iter().enumerate() {
+            write!(f, "{}", coeff)?;
+            if i < self.as_coeffs().len() - 1 {
+                write!(f, ", ")?;
+            }
+        }
+        write!(f, "}}")?;
+
+        return Ok(());
+    }
+}
+
+impl core::fmt::Display for Poly {
+    fn fmt(&self, f: &mut core::fmt::Formatter<'_>) -> core::fmt::Result {
+        <Self as core::fmt::Debug>::fmt(self, f)
+    }
+}
+
+// Hexadecimal representation is still meaningful since we will be dealing with compression and
+// decompression later on
+impl UpperHex for Poly {
+    fn fmt(&self, f: &mut core::fmt::Formatter<'_>) -> core::fmt::Result {
+        write!(f, "Poly {{")?;
+        for (i, coeff) in self.as_coeffs().iter().enumerate() {
+            // 3329 < 2 ** 12, so 12 bits is sufficient for representation
+            write!(f, "0x{:03X}", coeff)?;
+            if i < self.as_coeffs().len() - 1 {
+                write!(f, ", ")?;
+            }
+        }
+        write!(f, "}}")?;
+
[…]

### 2024-04-26 lattices/primitives/src/main.rs

@@ -1,0 +1,6 @@
+use primitives::Poly;
+
+fn main() {
+    let poly = Poly::from_coeffs([3328; 256]);
+    println!("{:x}", poly);
+}

### 2024-04-28 lattices/primitives/src/lib.rs
Barret reduction

@@ -2,6 +2,49 @@
 #![no_std]
 
 use core::fmt::{LowerHex, UpperHex};
+use subtle::{Choice, ConditionallySelectable};
+
+/// word size of all arithmetics
+pub type Word = u64;
+
+/// An element of the prime field Z_q where q = 3329
+#[derive(Debug, Clone, Copy)]
+pub struct FieldElem(pub Word);
+
+impl ConditionallySelectable for FieldElem {
+    fn conditional_select(a: &Self, b: &Self, choice: Choice) -> Self {
+        let val = u64::conditional_select(&a.0, &b.0, choice);
+        return Self(val);
+    }
+}
+
+impl FieldElem {
+    pub const Q: Word = 3329;
+    /// k = 2 * (floor(log2(Q)) + 1)
+    pub const BARRETT_SHIFT: usize = 24;
+    pub const BARRETT_MULTIPLIER: Word = (1 << Self::BARRETT_SHIFT) / Self::Q;
+
+    /// Assuming that the input value satisfies 0 <= val < 2*Q, perform a simple reduction
+    fn simple_reduce(val: Word) -> Self {
+        let val = Word::conditional_select(
+            &(val.wrapping_sub(Self::Q)),
+            &val,
+            Choice::from((val < Self::Q) as u8),
+        );
+
+        Self(val)
+    }
+
+    /// Perform the Barrett reduction
+    pub fn barrett_reduce(val: Word) -> Self {
+        let quot = (val * Self::BARRETT_MULTIPLIER) >> Self::BARRETT_SHIFT;
+        return Self::simple_reduce(val - quot * Self::Q);
+    }
+
+    pub fn modmul(&self, other: &Self) -> Self {
+        Self::barrett_reduce(self.0 * other.0)
+    }
+}
 
 /// A member of the polynomial ring used in Kyber:
 /// R_q = Z_q[x] / (x ** n + 1)
@@ -78,5 +121,21 @@ impl LowerHex for Poly {
         write!(f, "}}")?;
 
         return Ok(());
+    }
+}
+
+#[cfg(test)]
+mod tests {
+    use super::*;
+
+    #[test]
+    fn field_modmul() {
+        for a in 0..FieldElem::Q {
+            for b in 0..FieldElem::Q {
+                let barret_modmul = FieldElem(a).modmul(&FieldElem(b)).0;
+                let expected = (a * b) % FieldElem::Q;
+                assert_eq!(barret_modmul, expected);
+            }
+        }
     }
 }

### 2024-04-30 lattices/primitives/src/lib.rs
Untested impl of algorithm 6

@@ -2,13 +2,17 @@
 #![no_std]
 
 use core::fmt::{LowerHex, UpperHex};
+use sha3::digest::XofReader;
 use subtle::{Choice, ConditionallySelectable};
 
 /// word size of all arithmetics
 pub type Word = u64;
 
+/// Degree of the quotient polynomial
+pub const N: usize = 256;
+
 /// An element of the prime field Z_q where q = 3329
-#[derive(Debug, Clone, Copy)]
+#[derive(Debug, Clone, Copy, PartialEq, Eq)]
 pub struct FieldElem(pub Word);
 
 impl ConditionallySelectable for FieldElem {
@@ -43,6 +47,50 @@ impl FieldElem {
 
     pub fn modmul(&self, other: &Self) -> Self {
         Self::barrett_reduce(self.0 * other.0)
+    }
+}
+
+/// A member of the polynomial ring R_q in NTT domain
+///
+/// Mathematically, the NTT representation of some polynomial in R_q consists of 128 polynomials
+/// each of degree 1, but we don't impose additional structure on the coefficients so the sampling
+/// and transformation methods can be easier to implement
+#[derive(Debug, Copy, Clone, PartialEq)]
+pub struct PolyNTT {
+    /// Entries of the NTT representation
+    coeffs: [FieldElem; N],
+}
+
+impl PolyNTT {
+    /// Algorithm 6: Sample from uniform distribution
+    ///
+    /// Since q = 3329 takes 12 bits, we try to sample 2 entries from 3 bytes at a time. Each entry
+    /// is sampled from a random 12 bit string, but if the 12-bit integer is too large, it is
+    /// rejected. Conditioned on the 12-bit integer being less than Q, the sampled coefficient
+    /// follows a uniform distribution within Z_q
+    pub fn sample(mut xof: impl XofReader) -> Self {
+        let mut j: usize = 0;
+        let mut coeffs: [FieldElem; N] = [FieldElem(0); N];
+        let mut stream = [0u8; 3];
+
+        while j < 256 {
+            xof.read(&mut stream);
+            let b1: Word = stream[0].into();
+            let b2: Word = stream[1].into();
+            let b3: Word = stream[2].into();
+            let d1 = b1 + (b2 & 0x0F) << 8;
+            let d2 = b2 >> 4 + b3 << 4;
+            if d1 < FieldElem::Q {
+                coeffs[j] = FieldElem(d1);
+                j += 1;
+            }
[…]

### 2024-04-30 lattices/primitives/src/lib.rs
refactored Poly to be an array of field elemenets

@@ -23,6 +23,7 @@ impl ConditionallySelectable for FieldElem {
 }
 
 impl FieldElem {
+    pub const ZERO: Self = Self(0);
     pub const Q: Word = 3329;
     /// k = 2 * (floor(log2(Q)) + 1)
     pub const BARRETT_SHIFT: usize = 24;
@@ -45,8 +46,19 @@ impl FieldElem {
         return Self::simple_reduce(val - quot * Self::Q);
     }
 
+    /// Modulus multiplication
     pub fn modmul(&self, other: &Self) -> Self {
         Self::barrett_reduce(self.0 * other.0)
+    }
+
+    /// Modulus subtraction. Will wrap around the modulus if self is less than other
+    pub fn modsub(&self, other: &Self) -> Self {
+        let negwrap = Choice::from((self.0 < other.0) as u8);
+        return Self::conditional_select(
+            &Self(self.0.wrapping_sub(other.0)),
+            &Self(Self::Q + self.0 - other.0),
+            negwrap,
+        );
     }
 }
 
@@ -104,15 +116,19 @@ pub struct Poly {
     ///
     /// NOTE: something about using a larger uint type than strictly necessary (3329 < 2 ** 16)
     /// because it's easier to do modular arithmetic with
-    coeffs: [u32; 256],
+    coeffs: [FieldElem; N],
 }
 
 impl Poly {
-    pub fn from_coeffs(coeffs: [u32; 256]) -> Self {
-        Self { coeffs }
+    pub fn from_words(words: [Word; N]) -> Self {
+        let mut coeffs = [FieldElem::ZERO; N];
+        (0..N).for_each(|i| {
+            coeffs[i] = FieldElem(words[i]);
+        });
+        return Self { coeffs };
     }
 
-    pub fn as_coeffs(&self) -> &[u32] {
+    pub fn as_coeffs(&self) -> &[FieldElem] {
         &self.coeffs
     }
 }
@@ -121,7 +137,7 @@ impl core::fmt::Debug for Poly {
     fn fmt(&self, f: &mut core::fmt::Formatter<'_>) -> core::fmt::Result {
         write!(f, "Poly {{")?;
         for (i, coeff) in self.as_coeffs().iter().enumerate() {
-            write!(f, "{}", coeff)?;
+            write!(f, "{}", coeff.0)?;
             if i < self.as_coeffs().len() - 1 {
                 write!(f, ", ")?;
             }
@@ -145,7 +161,7 @@ impl UpperHex for Poly {
[…]

### 2024-04-30 lattices/primitives/src/main.rs

@@ -1,6 +1,6 @@
 use primitives::Poly;
 
 fn main() {
-    let poly = Poly::from_coeffs([3328; 256]);
+    let poly = Poly::from_words([3328; 256]);
     println!("{:x}", poly);
 }

### 2024-04-30 lattices/primitives/src/lib.rs
Sample from CBD(eta=2)

@@ -29,6 +29,10 @@ impl FieldElem {
     pub const BARRETT_SHIFT: usize = 24;
     pub const BARRETT_MULTIPLIER: Word = (1 << Self::BARRETT_SHIFT) / Self::Q;
 
+    pub fn from_u8(byte: u8) -> Self {
+        Self(byte.into())
+    }
+
     /// Assuming that the input value satisfies 0 <= val < 2*Q, perform a simple reduction
     fn simple_reduce(val: Word) -> Self {
         let val = Word::conditional_select(
@@ -131,6 +135,35 @@ impl Poly {
     pub fn as_coeffs(&self) -> &[FieldElem] {
         &self.coeffs
     }
+
+    /// Sample a polynomial from CBD(eta=2)
+    pub fn sample_cbd_eta2(mut xof: impl XofReader) -> Self {
+        let mut stream = [0u8; N / 2];
+        xof.read(&mut stream);
+        let mut coeffs = [FieldElem::ZERO; N];
+
+        stream.iter().enumerate().for_each(|(i, byte)| {
+            // bit operators have even lower precedence than +
+            let d1 = FieldElem::from_u8(((byte >> 0) & 1) + ((byte >> 1) & 1))
+                .modsub(&FieldElem::from_u8(((byte >> 2) & 1) + ((byte >> 3) & 1)));
+            let d2 = FieldElem::from_u8(((byte >> 4) & 1) + ((byte >> 5) & 1))
+                .modsub(&FieldElem::from_u8(((byte >> 6) & 1) + ((byte >> 7) & 1)));
+            coeffs[2 * i] = d1;
+            coeffs[2 * i + 1] = d2;
+        });
+
+        return Self { coeffs };
+    }
+
+    /// Sample a polynomial from CBD(eta=3)
+    pub fn sample_cbd_eta3(mut xof: impl XofReader) -> Self {
+        // Every 3 bytes can output 4 samples
+        let mut stream = [0u8; N / 4 * 3];
+        xof.read(&mut stream);
+        let coeffs = [FieldElem::ZERO; N];
+
+        return Self { coeffs };
+    }
 }
 
 impl core::fmt::Debug for Poly {
@@ -201,5 +234,21 @@ mod tests {
                 assert_eq!(barret_modmul, expected);
             }
         }
+    }
+
+    #[test]
+    fn test_sample_cbd_eta2() {
+        use sha3::{
+            digest::{ExtendableOutput, Update},
+            Shake256,
[…]

### 2024-04-30 lattices/primitives/src/main.rs

@@ -1,6 +1,13 @@
 use primitives::Poly;
+use sha3::{
+    digest::{ExtendableOutput, Update},
+    Shake256,
+};
 
 fn main() {
-    let poly = Poly::from_words([3328; 256]);
-    println!("{:x}", poly);
+    let mut xof = Shake256::default();
+    xof.update(b"!!!Hello, world");
+    let xof = xof.finalize_xof();
+    let poly = Poly::sample_cbd_eta2(xof);
+    println!("{}", poly);
 }

### 2024-05-01 lattices/primitives/src/lib.rs
Implemented sample_cbd_eta3, also did some refactoring

@@ -9,7 +9,8 @@ use subtle::{Choice, ConditionallySelectable};
 pub type Word = u64;
 
 /// Degree of the quotient polynomial
-pub const N: usize = 256;
+pub const KYBER_N: usize = 256;
+pub const KYBER_Q: Word = 3329;
 
 /// An element of the prime field Z_q where q = 3329
 #[derive(Debug, Clone, Copy, PartialEq, Eq)]
@@ -24,10 +25,10 @@ impl ConditionallySelectable for FieldElem {
 
 impl FieldElem {
     pub const ZERO: Self = Self(0);
-    pub const Q: Word = 3329;
+    pub const Q: Word = KYBER_Q;
     /// k = 2 * (floor(log2(Q)) + 1)
     pub const BARRETT_SHIFT: usize = 24;
-    pub const BARRETT_MULTIPLIER: Word = (1 << Self::BARRETT_SHIFT) / Self::Q;
+    pub const BARRETT_MULTIPLIER: Word = (1 << Self::BARRETT_SHIFT) / KYBER_Q;
 
     pub fn from_u8(byte: u8) -> Self {
         Self(byte.into())
@@ -36,9 +37,9 @@ impl FieldElem {
     /// Assuming that the input value satisfies 0 <= val < 2*Q, perform a simple reduction
     fn simple_reduce(val: Word) -> Self {
         let val = Word::conditional_select(
-            &(val.wrapping_sub(Self::Q)),
+            &(val.wrapping_sub(KYBER_Q)),
             &val,
-            Choice::from((val < Self::Q) as u8),
+            Choice::from((val < KYBER_Q) as u8),
         );
 
         Self(val)
@@ -47,7 +48,7 @@ impl FieldElem {
     /// Perform the Barrett reduction
     pub fn barrett_reduce(val: Word) -> Self {
         let quot = (val * Self::BARRETT_MULTIPLIER) >> Self::BARRETT_SHIFT;
-        return Self::simple_reduce(val - quot * Self::Q);
+        return Self::simple_reduce(val - quot * KYBER_Q);
     }
 
     /// Modulus multiplication
@@ -60,7 +61,7 @@ impl FieldElem {
         let negwrap = Choice::from((self.0 < other.0) as u8);
         return Self::conditional_select(
             &Self(self.0.wrapping_sub(other.0)),
-            &Self(Self::Q + self.0 - other.0),
+            &Self(KYBER_Q + self.0 - other.0),
[…]

### 2024-05-01 lattices/primitives/src/main.rs

@@ -8,6 +8,6 @@ fn main() {
     let mut xof = Shake256::default();
     xof.update(b"!!!Hello, world");
     let xof = xof.finalize_xof();
-    let poly = Poly::sample_cbd_eta2(xof);
+    let poly = Poly::sample_cbd_eta3(xof);
     println!("{}", poly);
 }

### 2024-05-01 lattices/compact-fo-using-mac/manuscript/main.tex
Let's read Poly1305

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{multicol}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
\newcommand{\keygen}{\operatorname{KeyGen}}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{Mermory-efficient Fujisaki-Okamoto transformation}
\author{Ganyu (Bruce) Xu (g66xu)}
% \date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{OW-PCVA transformation}
Let $(\keygen, E, D)$ define a probabilistic public-key encryption scheme, then $T$ is a transformation that takes this scheme and outputs a second public-key encryption scheme. The transformation makes use of two additional components:

\begin{enumerate}
    \item A $\operatorname{MAC}: \mathcal{K}^\text{MAC} \times \{0,1\}^\ast \rightarrow \mathcal{T}$
    \item A hash function $G: \mathcal{M} \rightarrow \text{Coin} \times \mathcal{K}^\text{MAC}$
\end{enumerate}

\begin{multicols}{2}
    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{OW-PCVA encryption $E^T$}
        \KwIn{
            $\text{pk}$, $m \in \mathcal{M}$
        }
        $(r, k) \leftarrow G(m)$\;
        $c \leftarrow E(\text{pk}, m, r)$\;
        $t \leftarrow \operatorname{MAC}(k, c)$
        \tcp*[h]{"encrypt-then-mac"}\;
        \Return{$(c, t)$}\;
    \end{algorithm}

    \columnbreak

    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{OW-PCVA decryption $D^T$}
        \KwIn{
            $\text{sk}, c \in \mathcal{C}, t \in \mathcal{T}$
        }
        $\hat{m} \leftarrow D(\text{sk}, c)$\;
        $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$\;
        \If{
            $\operatorname{MAC}(\hat{k}, c) \neq t$
        }{
            \Return{$\bot$}\;
        }
        \Return{$\hat{m}$}\;
    \end{algorithm}
\end{multicols}

Instead of using re-encryption to check for the integrity of the ciphertext, a MAC tag is used. At the cost of adding a fixed number of bytes to the ciphertext, the decryption routine becomes significantly faster and memory-efficient. The memory trade-off is especially important since in Kyber, re-encryption needs to compute the entire public key $A \in R_q^{k \times k}$ from the 32-byte seed.

We claim that this transformation is still OW-PCVA. The proof will be largely similar to that of the original OW-PCVA transformation proof, but with some notable differences:

\begin{itemize}
    \item When queried on some authenticated ciphertext $(\tilde{c}, \tilde{t})$, the secret-key-less $\operatorname{CVO}_1$ searches the hash oracle $G$ for $(m, r, k)$ such that $\operatorname{MAC}(k, \tilde{c}) = \tilde{t}$. The diverging event between using $\operatorname{CVO}$ and using $\operatorname{CVO}_1$ is ``adversary produces valid authenticated ciphertext without consulting the hash oracle'', which is equivalent to selective forgery (or existential forgery, I am not sure which): $\epsilon_a - \epsilon_b \leq q_V \cdot \epsilon_\text{MAC}$
\end{itemize}

\begin{theorem}\label{mac-fo-theorem}
    Let $(\keygen, E, D)$ be a public-key encryption scheme that is $\delta$-correct and that has $\gamma$-spread. For every OW-PCVA adversary against the transformed encryption scheme $(E^T, D^T)$ that makes $q_G$ hash queries and $q_V$ ciphertext validation queries and that has advantage $\epsilon_\text{OW-PCVA}$, there exists an IND-CPA adversary against $(\keygen, E, D)$ with advantage $\epsilon_\text{IND-CPA}$ and an EF-CMA adversary against the MAC with advantage $\epsilon_\text{MAC}$ such that:

    \begin{equation*}
        \epsilon_\text{OW-PCVA}
        \leq q_G \cdot \delta 
        + q_V \cdot \epsilon_\text{MAC}
        + \frac{2q+1}{\norm{\mathcal{M}}} 
        + \epsilon_\text{IND-CPA}
    \end{equation*}
\end{theorem}

The proof will be discussed in section \ref{proof-of-mac-fo-theorem}.

\section{Proof techniques}
Similar to the 2017 paper, we will use a sequence of games to incrementally replace PCO, CVO, and other aspects of the standard OW-PCVA game with some simulation. Then we will show that the total loss of security is negligible.

\subsection{Replacing plaintext-checking oracle}\label{replacing-pco}
The plaintext-checking oracle $\operatorname{PCO}$ takes as input a plaintext-ciphertext pair $(m, (c, t))$ and return ``reject'' if and only if the the ciphertext is a valid encryption of the plaintext and vice versa under the context of some fixed keypair.

The simulated plaintext-checking oracle $\operatorname{PCO}_1$ removes the requirement for the secret key by removing the step that checks whether the queried ciphertext $(c, t)$ decrypts back to the queried plaintext $m$.

From the OW-PCVA adversary's perspective, the two oracles are indistinguishable, except for when there is a decryption error $D(\text{sk}, E(\text{pk}, m, r)) \neq m$ for the input plaintext $m$ and the corresponding coin $r$. The MAC is not involved in this argument. The loss of security is still $q_G\cdot\delta$

\begin{multicols}{2}
    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{Vanilla PCO}
        \KwIn{
            $m \in \mathcal{M}, c \in \mathcal{C}, t \in \mathcal{T}$ 
        }
        \If{
            $D^T(\text{sk}, (c, t)) \neq m$
        }{
            \Return{$\bot$}\;
            \tcp*[h]{Decryption did not match}
        }
        \If{
            $E^T(\text{pk}, m) \neq (c, t)$
        }{
            \Return{$\bot$}\;
            \tcp*[h]{Encryption did not match}
        }
    \end{algorithm}

    \columnbreak

    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{Simulated $\operatorname{PCO}_1$}
        \KwIn{
            $m \in \mathcal{M}, c \in \mathcal{C}, t \in \mathcal{T}$
        }
        \If{
            $E^T(\text{pk}, m) \neq (c, t)$
        }{
            \Return{$\bot$}\;
            \tcp*[h]{Encryption did not match}
        }
    \end{algorithm}
\end{multicols}

\subsection{Replacing ciphertext-validation oracle}\label{replacing-cvo}
This is a similar trick to what was used in the 2017 paper, as well. Where the ciphertext is honestly generated, both oracles will return accept. Where the vanilla CVO rejects the ciphertext (aka ciphertext is malformed or the tag doesn't match), the simulated CVO will also reject. Therefore, the diverging event is when the vanilla CVO accepts but the simulated CVO rejects. Since the vanilla CVO accepts the query, the tag $t$ must be valid for the queried ciphertext $c$; on the other hand, the simulated CVO's rejection means there is no matching query in the hash oracle. Under the random oracle assumption, $t$ is a valid tag for some data $c$ under some key that the adversary does not know. In other words, $(c, t)$ is some kind of forgery.

\textbf{The argument on how the security of the MAC relates to this diverging event is still a bit fuzzy}, but here are two possible ways I can think of:

\begin{itemize}
    \item For each CVO query, the adversary is trying to forge tag for that specific ciphertext. This means there is a selective forgery attack, and over all $q_V$ validation queries, the probability of having at least one selective forgery attack that works is at most $q_V\epsilon_\text{SF}$

    \item Across all CVO query, the adversary wants to forge tag for some ciphertext, meaning there is a existential forgery attack, and the probability is at most $\epsilon_\text{EF}$
\end{itemize}

\begin{multicols}{2}
    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{Vanilla CVO}
        \KwIn{
            $(c \in \mathcal{C}, t \in \mathcal{T})$
        }
        \If{
            $D^T(\text{sk}, (c, t)) = \bot$
        }{
            \Return{$\bot$}\;
        }
    \end{algorithm}

    \columnbreak

    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{Simulated $\operatorname{CVO}_1$}
        \KwIn{
            $(c \in \mathcal{C}, t \in \mathcal{T})$
        }
        \If{
            $\exists (m, r, k) \in \mathcal{O}^G$ such that $\operatorname{MAC}(k, c) = t$
        }{
            \Return{Accept}\;
        }
        \Return{$\bot$}
    \end{algorithm}
\end{multicols}

\subsection{Random until queried}\label{random-until-queried}
We then replace the pseudorandom coin and the MAC key with truly random coins and truly random keys in the challenge encryption. These two challenge encryption routines are identical, unless the adversary queries $G$ with the challenge plaintext $m^\ast$, but the probability of making such query can be bounded.

\begin{multicols}{2}
    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{Challenge encryption}
        $m^\ast \leftsample \mathcal{M}$\;
        $(c^\ast, t^\ast) \leftarrow E^T(\text{pk}, m^\ast)$\;
        \Return{$(c^\ast, t^\ast)$}
    \end{algorithm}

    \begin{algorithm}[H]
        \SetAlgoLined
        \caption{Simulated challenge encryption}
        $m^\ast \leftsample \mathcal{M}$\;
        $r^\ast \leftsample \text{Coin}, k^\ast \leftsample \mathcal{K}_\text{MAC}$\;
        $c^\ast \leftarrow E(\text{pk}, m^\ast, r^\ast)$\;
        $t^\ast \leftarrow \operatorname{MAC}(k^\ast, c^\ast)$
        \Return{$(c^\ast, t^\ast)$}
    \end{algorithm}
\end{multicols}

Let the OW-PCVA adversary's advantage using the vanilla challenge encryption be $\epsilon_0$ and its advantage using the simulated challenge encryption be $\epsilon_1$. Denote the event that the adversary queries the hash oracle $G$ with the challenge plaintext $m^\ast$ by $\operatorname{QUERY}^\ast$, then:

\begin{equation*}
    \epsilon_0 - \epsilon_1 \leq P[\operatorname{QUERY}^\ast]
\end{equation*}

We can build an IND-CPA adversary against the underlying encryption scheme to bound $P[\operatorname{QUERY}^\ast]$: if the OW-PCVA adversary makes the magic query, then the IND-CPA adversary can find it in the hash oracle's tape and win the IND-CPA game; if the OW-PCVA adversary does not make the magic query, then the IND-CPA adversary outputs a blind guess:

\begin{equation*}
    P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}
\end{equation*}

\section{Complete proof of \ref{mac-fo-theorem}}\label{proof-of-mac-fo-theorem}
\begin{proof}
    We will prove using a sequence of games. Game 0 is the standard OW-PCVA game.

    Game 1 is identical to game 0, except PCO is replaced with $\operatorname{PCO}_1$, the loss of security is described in section \ref{replacing-pco}:

    \begin{equation*}
        \epsilon_0 - \epsilon_1 \leq q_G \cdot \delta
    \end{equation*}

    Game 2 is identical to game 1, except CVO is replaced with $\operatorname{CVO}_1$, the loss of security is described in section \ref{replacing-cvo}. Here I put a placeholder $\epsilon_\text{MAC}$ to indicate that this quantity is tied to the security of the MAC, but I don't have a solid answer yet

    \begin{equation*}
        \epsilon_1 - \epsilon_2 \leq \epsilon_\text{MAC}
    \end{equation*}

    Game 3 is identical to game 2, except the challenge encryption is replaced with the simulated challenge encryption

    \begin{equation*}
        \epsilon_3 - \epsilon_2 \leq P[\operatorname{QUERY}^\ast]
    \end{equation*}

    From section \ref{random-until-queried} we know that $P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}$ for some IND-CPA adversary against the underlying PKE.

    Game 3 can be entirely simulated by an OW-CPA adversary against the underlying PKE:

    \begin{itemize}
        \item The keypairs are identical between the two PKE's
        \item The OW-CPA adversary can simulate PCO, CVO, and hash oracle $G$
        \item The OW-CPA adversary receives challenge ciphertext $c^\ast$ which is computed from a truly random coin; it can then random a truly random MAC key and compute the tag $t^\ast$
        \item The OW-CPA adversary passes $(c^\ast, t^\ast)$ to the OW-PCVA adversary and returns whatever the OW-PCVA adversary returns
    \end{itemize}

    The OW-CPA adversary wins if and only if the OW-PCVA adversary wins: $\epsilon_3 = \epsilon_\text{OW-CPA}$.

    Putting everything together we have:

    \begin{equation*}
        \epsilon_0 
        \leq q_G \cdot \delta 
        + \epsilon_\text{MAC} 
        + \epsilon_\text{IND-CPA} 
        + \epsilon_\text{OW-CPA}
    \end{equation*}
\end{proof}

\section{Open questions}
Can we get rid of the coin and just let the encryption scheme be probabilistic? Do we need to?

\end{document}

### 2024-05-03 lattices/primitives/src/lib.rs
Implemented ntt and ntt_invert

@@ -12,6 +12,155 @@ pub type Word = u64;
 pub const KYBER_N: usize = 256;
 pub const KYBER_Q: Word = 3329;
 
+/// zeta (the 256-th primitive root of 3329) and its powers, up to 127
+pub const ZETA_POWS: [FieldElem; 128] = [
+    FieldElem(1),
+    FieldElem(17),
+    FieldElem(289),
+    FieldElem(1584),
+    FieldElem(296),
+    FieldElem(1703),
+    FieldElem(2319),
+    FieldElem(2804),
+    FieldElem(1062),
+    FieldElem(1409),
+    FieldElem(650),
+    FieldElem(1063),
+    FieldElem(1426),
+    FieldElem(939),
+    FieldElem(2647),
+    FieldElem(1722),
+    FieldElem(2642),
+    FieldElem(1637),
+    FieldElem(1197),
+    FieldElem(375),
+    FieldElem(3046),
+    FieldElem(1847),
+    FieldElem(1438),
+    FieldElem(1143),
+    FieldElem(2786),
+    FieldElem(756),
+    FieldElem(2865),
+    FieldElem(2099),
+    FieldElem(2393),
+    FieldElem(733),
+    FieldElem(2474),
+    FieldElem(2110),
+    FieldElem(2580),
+    FieldElem(583),
+    FieldElem(3253),
+    FieldElem(2037),
+    FieldElem(1339),
+    FieldElem(2789),
+    FieldElem(807),
+    FieldElem(403),
+    FieldElem(193),
+    FieldElem(3281),
+    FieldElem(2513),
+    FieldElem(2773),
+    FieldElem(535),
+    FieldElem(2437),
+    FieldElem(1481),
+    FieldElem(1874),
+    FieldElem(1897),
+    FieldElem(2288),
+    FieldElem(2277),
+    FieldElem(2090),
+    FieldElem(2240),
+    FieldElem(1461),
+    FieldElem(1534),
+    FieldElem(2775),
+    FieldElem(569),
+    FieldElem(3015),
+    FieldElem(1320),
+    FieldElem(2466),
+    FieldElem(1974),
+    FieldElem(268),
+    FieldElem(1227),
+    FieldElem(885),
+    FieldElem(1729),
+    FieldElem(2761),
+    FieldElem(331),
+    FieldElem(2298),
+    FieldElem(2447),
+    FieldElem(1651),
+    FieldElem(1435),
+    FieldElem(1092),
[…]

### 2024-05-04 lattices/primitives/src/lib.rs
implemented NTT and verified correctness

@@ -13,7 +13,7 @@ pub const KYBER_N: usize = 256;
 pub const KYBER_Q: Word = 3329;
 
 /// zeta (the 256-th primitive root of 3329) and its powers, up to 127
-pub const ZETA_POWS: [FieldElem; 128] = [
+pub const ZETA_POWS: [FieldElem; 256] = [
     FieldElem(1),
     FieldElem(17),
     FieldElem(289),
@@ -142,6 +142,134 @@ pub const ZETA_POWS: [FieldElem; 128] = [
     FieldElem(641),
     FieldElem(910),
     FieldElem(2154),
+    FieldElem(3328),
+    FieldElem(3312),
+    FieldElem(3040),
+    FieldElem(1745),
+    FieldElem(3033),
+    FieldElem(1626),
+    FieldElem(1010),
+    FieldElem(525),
+    FieldElem(2267),
+    FieldElem(1920),
+    FieldElem(2679),
+    FieldElem(2266),
+    FieldElem(1903),
+    FieldElem(2390),
+    FieldElem(682),
+    FieldElem(1607),
+    FieldElem(687),
+    FieldElem(1692),
+    FieldElem(2132),
+    FieldElem(2954),
+    FieldElem(283),
+    FieldElem(1482),
+    FieldElem(1891),
+    FieldElem(2186),
+    FieldElem(543),
+    FieldElem(2573),
+    FieldElem(464),
+    FieldElem(1230),
+    FieldElem(936),
+    FieldElem(2596),
+    FieldElem(855),
+    FieldElem(1219),
+    FieldElem(749),
+    FieldElem(2746),
+    FieldElem(76),
+    FieldElem(1292),
+    FieldElem(1990),
+    FieldElem(540),
+    FieldElem(2522),
+    FieldElem(2926),
+    FieldElem(3136),
+    FieldElem(48),
+    FieldElem(816),
+    FieldElem(556),
+    FieldElem(2794),
+    FieldElem(892),
+    FieldElem(1848),
+    FieldElem(1455),
+    FieldElem(1432),
+    FieldElem(1041),
+    FieldElem(1052),
+    FieldElem(1239),
+    FieldElem(1089),
+    FieldElem(1868),
+    FieldElem(1795),
+    FieldElem(554),
+    FieldElem(2760),
+    FieldElem(314),
+    FieldElem(2009),
+    FieldElem(863),
+    FieldElem(1355),
+    FieldElem(3061),
[…]

### 2024-05-04 lattices/primitives/src/main.rs

@@ -9,5 +9,5 @@ fn main() {
     xof.update(b"!!!Hello, world");
     let xof = xof.finalize_xof();
     let poly = Poly::sample_cbd_eta3(xof);
-    println!("{}", poly);
+    println!("{}", poly.polymul(&poly));
 }

### 2024-05-06 lattices/primitives/src/lib.rs
Need to refactor CBD and uniform sampling

@@ -2,7 +2,8 @@
 #![no_std]
 
 use core::fmt::{LowerHex, UpperHex};
-use sha3::digest::XofReader;
+use sha3::digest::{ExtendableOutput, Update, XofReader};
+use sha3::{Shake128, Shake256};
 use subtle::{Choice, ConditionallySelectable};
 
 /// word size of all arithmetics
@@ -11,6 +12,12 @@ pub type Word = u64;
 /// Degree of the quotient polynomial
 pub const KYBER_N: usize = 256;
 pub const KYBER_Q: Word = 3329;
+pub const KYBER_K_512: usize = 2;
+pub const KYBER_K_768: usize = 3;
+pub const KYBER_K_1024: usize = 4;
+
+/// Most seeds are 32-bytes
+pub const SEEDSIZE: usize = 32;
 
 /// zeta (the 256-th primitive root of 3329) and its powers, up to 127
 pub const ZETA_POWS: [FieldElem; 256] = [
@@ -289,6 +296,26 @@ const fn bitrev7(val: u8) -> u8 {
     return bitrev;
 }
 
+pub fn shake256_prf<const ETA: usize>(seed: [u8; SEEDSIZE], ctr: u8) -> [u8; ETA] {
+    let mut hasher = Shake256::default();
+    hasher.update(&seed);
+    hasher.update(&[ctr]);
+    let mut xof = hasher.finalize_xof();
+
+    let mut output = [0u8; ETA];
+    xof.read(&mut output);
+    return output;
+}
+
+pub fn shake128_xof(seed: [u8; SEEDSIZE], i: u8, j: u8) -> impl XofReader {
+    let mut hasher = Shake128::default();
+    hasher.update(&seed);
+    hasher.update(&[i, j]);
+    let xof = hasher.finalize_xof();
+
+    return xof;
+}
+
 /// An element of the prime field Z_q where q = 3329
 #[derive(Debug, Clone, Copy, PartialEq, Eq)]
 pub struct FieldElem(pub Word);
@@ -362,13 +389,18 @@ pub struct PolyNTT {
 }
 
 impl PolyNTT {
+    pub const ZERO: Self = Self {
+        coeffs: [FieldElem::ZERO; KYBER_N],
+    };
+
     /// Algorithm 6: Sample from uniform distribution
     ///
     /// Since q = 3329 takes 12 bits, we try to sample 2 entries from 3 bytes at a time. Each entry
     /// is sampled from a random 12 bit string, but if the 12-bit integer is too large, it is
[…]

### 2024-05-06 lattices/primitives/src/main.rs

@@ -7,7 +7,7 @@ use sha3::{
 fn main() {
     let mut xof = Shake256::default();
     xof.update(b"!!!Hello, world");
-    let xof = xof.finalize_xof();
-    let poly = Poly::sample_cbd_eta3(xof);
+    let mut xof = xof.finalize_xof();
+    let poly = Poly::sample_cbd_eta3(&mut xof);
     println!("{}", poly.polymul(&poly));
 }

### 2024-05-07 lattices/primitives/src/algebra.rs
Implemented generic CBD sampling and byte_encode

@@ -1,0 +1,801 @@
+//! Modulus and polynomial ring operations
+use core::fmt::{Binary, LowerHex, UpperHex};
+use sha3::digest::XofReader;
+use subtle::{Choice, ConditionallySelectable};
+
+use crate::{Word, KYBER_N, KYBER_Q};
+
+/// zeta (the 256-th primitive root of 3329) and its powers, up to 127
+pub const ZETA_POWS: [FieldElem; 256] = [
+    FieldElem(1),
+    FieldElem(17),
+    FieldElem(289),
+    FieldElem(1584),
+    FieldElem(296),
+    FieldElem(1703),
+    FieldElem(2319),
+    FieldElem(2804),
+    FieldElem(1062),
+    FieldElem(1409),
+    FieldElem(650),
+    FieldElem(1063),
+    FieldElem(1426),
+    FieldElem(939),
+    FieldElem(2647),
+    FieldElem(1722),
+    FieldElem(2642),
+    FieldElem(1637),
+    FieldElem(1197),
+    FieldElem(375),
+    FieldElem(3046),
+    FieldElem(1847),
+    FieldElem(1438),
+    FieldElem(1143),
+    FieldElem(2786),
+    FieldElem(756),
+    FieldElem(2865),
+    FieldElem(2099),
+    FieldElem(2393),
+    FieldElem(733),
+    FieldElem(2474),
+    FieldElem(2110),
+    FieldElem(2580),
+    FieldElem(583),
+    FieldElem(3253),
+    FieldElem(2037),
+    FieldElem(1339),
+    FieldElem(2789),
+    FieldElem(807),
+    FieldElem(403),
+    FieldElem(193),
+    FieldElem(3281),
+    FieldElem(2513),
+    FieldElem(2773),
+    FieldElem(535),
+    FieldElem(2437),
+    FieldElem(1481),
+    FieldElem(1874),
+    FieldElem(1897),
+    FieldElem(2288),
+    FieldElem(2277),
+    FieldElem(2090),
+    FieldElem(2240),
+    FieldElem(1461),
+    FieldElem(1534),
+    FieldElem(2775),
+    FieldElem(569),
+    FieldElem(3015),
+    FieldElem(1320),
+    FieldElem(2466),
+    FieldElem(1974),
+    FieldElem(268),
+    FieldElem(1227),
+    FieldElem(885),
+    FieldElem(1729),
+    FieldElem(2761),
+    FieldElem(331),
[…]

### 2024-05-07 lattices/primitives/src/lib.rs

@@ -1,17 +1,18 @@
 //! Primitives used in lattice-based cryptography
 #![no_std]
 
-use core::fmt::{LowerHex, UpperHex};
-use sha3::digest::{ExtendableOutput, Update, XofReader};
-use sha3::{Shake128, Shake256};
-use subtle::{Choice, ConditionallySelectable};
+use crate::algebra::FieldElem;
+use sha3::{
+    digest::{ExtendableOutput, Update, XofReader},
+    Shake128, Shake256,
+};
 
 /// word size of all arithmetics
 pub type Word = u64;
 
-/// Degree of the quotient polynomial
 pub const KYBER_N: usize = 256;
 pub const KYBER_Q: Word = 3329;
+pub const KYBER_Q_BITS: usize = 12;
 pub const KYBER_K_512: usize = 2;
 pub const KYBER_K_768: usize = 3;
 pub const KYBER_K_1024: usize = 4;
@@ -19,282 +20,7 @@ pub const KYBER_K_1024: usize = 4;
 /// Most seeds are 32-bytes
 pub const SEEDSIZE: usize = 32;
 
-/// zeta (the 256-th primitive root of 3329) and its powers, up to 127
-pub const ZETA_POWS: [FieldElem; 256] = [
-    FieldElem(1),
-    FieldElem(17),
-    FieldElem(289),
-    FieldElem(1584),
-    FieldElem(296),
-    FieldElem(1703),
-    FieldElem(2319),
-    FieldElem(2804),
-    FieldElem(1062),
-    FieldElem(1409),
-    FieldElem(650),
-    FieldElem(1063),
-    FieldElem(1426),
-    FieldElem(939),
-    FieldElem(2647),
-    FieldElem(1722),
-    FieldElem(2642),
-    FieldElem(1637),
-    FieldElem(1197),
-    FieldElem(375),
-    FieldElem(3046),
-    FieldElem(1847),
-    FieldElem(1438),
-    FieldElem(1143),
-    FieldElem(2786),
-    FieldElem(756),
-    FieldElem(2865),
-    FieldElem(2099),
-    FieldElem(2393),
-    FieldElem(733),
-    FieldElem(2474),
-    FieldElem(2110),
-    FieldElem(2580),
-    FieldElem(583),
-    FieldElem(3253),
-    FieldElem(2037),
-    FieldElem(1339),
-    FieldElem(2789),
-    FieldElem(807),
-    FieldElem(403),
[…]

### 2024-05-07 lattices/primitives/src/main.rs

@@ -1,4 +1,4 @@
-use primitives::Poly;
+use primitives::algebra::PolyNTT;
 use sha3::{
     digest::{ExtendableOutput, Update},
     Shake256,
@@ -6,8 +6,15 @@ use sha3::{
 
 fn main() {
     let mut xof = Shake256::default();
-    xof.update(b"!!!Hello, world");
+    xof.update(b"test");
     let mut xof = xof.finalize_xof();
-    let poly = Poly::sample_cbd_eta3(&mut xof);
-    println!("{}", poly.polymul(&poly));
+    let poly = PolyNTT::sample_uniform(&mut xof).invert_ntt();
+    println!("{:b}", poly);
+
+    let mut buffer = [0u8; 32 * 12];
+    primitives::byte_encode(poly.as_coeffs(), 12, &mut buffer);
+
+    for byte in buffer {
+        print!("{:08b}, ", byte);
+    }
 }

### 2024-05-08 lattices/primitives/src/algebra.rs
byte_decode correctly decode byte_encode

@@ -491,6 +491,10 @@ impl Poly {
         &self.coeffs
     }
 
+    pub fn as_mut_coeffs(&mut self) -> &mut [FieldElem] {
+        &mut self.coeffs
+    }
+
     /// Sample a polynomial from CBD(-eta, -eta+1, ..., eta-1, eta)
     fn sample_cbd(eta: usize, uniform: &[u8]) -> Self {
         // a single sample from CBD(eta) requires (2 * eta) uniform bits, each polynomial has 256

### 2024-05-08 lattices/primitives/src/lib.rs

@@ -49,8 +49,8 @@ pub fn shake128_xof(seed: [u8; SEEDSIZE], i: u8, j: u8) -> impl XofReader {
 /// Vec), so the user has to provide the buffer to write the bytes to. Fortunately, it is easy to
 /// check that the input buffer has the right size
 pub fn byte_encode(coeffs: &[FieldElem], d: usize, buffer: &mut [u8]) {
-    // Encoding 256 integers where each integer takes d bits requires at least (32 * d) bytes
-    assert!(buffer.len() == coeffs.len() * d / 8);
+    // The total number of bits must match
+    assert_eq!(coeffs.len() * d, buffer.len() * (u8::BITS as usize));
 
     for coeff_loc in 0..coeffs.len() {
         let coeff = coeffs[coeff_loc].0;
@@ -68,13 +68,37 @@ pub fn byte_encode(coeffs: &[FieldElem], d: usize, buffer: &mut [u8]) {
 }
 
 /// Algorithm 5: ByteDecode
+/// Assume that the input buffer is the output of calling byte_encode with the same value for d,
+/// recover the encoding back to the integer array
 pub fn byte_decode(coeffs: &mut [FieldElem], d: usize, buffer: &[u8]) {
-    todo!();
+    // The total number of bits must match
+    assert_eq!(coeffs.len() * d, buffer.len() * (u8::BITS as usize));
+
+    for (byte_loc, byte) in buffer.iter().enumerate() {
+        for bit_loc in 0..(u8::BITS as usize) {
+            let bit = (byte >> (7 - bit_loc)) & 1;
+
+            let coeff_loc = (byte_loc * (u8::BITS as usize) + bit_loc) / d;
+            let coeff_bit_loc = (byte_loc * (u8::BITS as usize) + bit_loc) % d;
+            let coeff = if bit == 1 {
+                coeffs[coeff_loc].0 | (1 << coeff_bit_loc)
+            } else {
+                coeffs[coeff_loc].0 & (!(1 << coeff_bit_loc))
+            };
+            coeffs[coeff_loc] = FieldElem(coeff);
+        }
+    }
 }
 
 #[cfg(test)]
 mod tests {
+    use self::algebra::{Poly, PolyNTT};
+
     use super::*;
+    use sha3::{
+        digest::{ExtendableOutput, Update},
+        Shake256,
+    };
 
     #[test]
     fn byte_encode_12() {
@@ -107,5 +131,20 @@ mod tests {
[…]

### 2024-05-08 lattices/compact-fo-using-mac/manuscript/main.tex
checkpoint

@@
 \newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
+% Cryptographic primitives
+\newcommand{\pke}{\operatorname{PKE}}
+\newcommand{\enc}{\operatorname{Enc}}
+\newcommand{\dec}{\operatorname{Dec}}
+\newcommand{\mac}{\operatorname{MAC}}
+\newcommand{\cvo}{\operatorname{CVO}}
+\newcommand{\pco}{\operatorname{PCO}}
 \newcommand{\keygen}{\operatorname{KeyGen}}
+\newcommand{\llbrack}{[\![}
+\newcommand{\rrbrack}{]\!]}
 
@@
 
-\section{OW-PCVA transformation}
-Let $(\keygen, E, D)$ define a probabilistic public-key encryption scheme, then $T$ is a transformation that takes this scheme and outputs a second public-key encryption scheme.
-The transformation makes use of two additional components:
+\section{Introduction}
 
-\begin{enumerate}
-    \item A $\operatorname{MAC}: \mathcal{K}^\text{MAC} \times \{0,1\}^\ast \rightarrow \mathcal{T}$
-    \item A hash function $G: \mathcal{M} \rightarrow \text{Coin} \times \mathcal{K}^\text{MAC}$
-\end{enumerate}
+\section{Preliminaries}
+
+\subsection{Security definitions}
+
+\subsection{Fujisaki-Okamoto transformation}
+
+\section{Memory-efficient OW-PCVA transformation using MAC}
+In this section we propose an alternative OW-PCVA transformation $\operatorname{T}$.
+
+Let $\pke = (\keygen, \enc, \dec)$ be a public-key encryption scheme defined over key space $\mathcal{K}_{\pke}$, message space $\mathcal{M}$, ciphertext space $\mathcal{C}$, and coin space $\mathcal{R}$.
+Let $\mac = (\operatorname{Sign}, \operatorname{Verify})$ be a message authentication code defined over key space $\mathcal{K}_{\mac}$, message space $\mathcal{M}$ (the same message space as $\pke$), and tag space $\mathcal{T}$.
+Let $G: \mathcal{M} \rightarrow \mathcal{R} \times \mathcal{K}_{\mac}$ be a random oracle.
+The transformed public-key encryption scheme is denoted by $\pke_1 = T(\pke, \mac, G)$ and includes the sub-routines $(\keygen, \enc_1, \dec_1)$.
+
+The transformed encryption and decryption routines are as follows
 
@@
         \SetAlgoLined
-        \caption{OW-PCVA encryption $E^T$}
+        \caption{OW-PCVA encryption $E_1$}
         \KwIn{
@@
         \SetAlgoLined
-        \caption{OW-PCVA decryption $D^T$}
+        \caption{OW-PCVA decryption $D_1$}
         \KwIn{
@@
 
-Instead of using re-encryption to check for the integrity of the ciphertext, a MAC tag is used.
-At the cost of adding a fixed number of bytes to the ciphertext, the decryption routine becomes significantly faster and memory-efficient.
-The memory trade-off is especially important since in Kyber, re-encryption needs to compute the entire public key $A \in R_q^{k \times k}$ from the 32-byte seed.
+If $\pke$ is OW-CPA secure and $\mac$ is existentially unforgeable under chosen message attack, then the transformed encryption scheme $\pke_q$ is OW-PCVA:
 
-We claim that this transformation is still OW-PCVA. The proof will be largely similar to that of the original OW-PCVA transformation proof, but with some notable differences:
-
-\begin{itemize}
-    \item When queried on some authenticated ciphertext $(\tilde{c}, \tilde{t})$, the secret-key-less $\operatorname{CVO}_1$ searches the hash oracle $G$ for $(m, r, k)$ such that $\operatorname{MAC}(k, \tilde{c}) = \tilde{t}$.
-The diverging event between using $\operatorname{CVO}$ and using $\operatorname{CVO}_1$ is ``adversary produces valid authenticated ciphertext without consulting the hash oracle'', which is equivalent to selective forgery (or existential forgery, I am not sure which): $\epsilon_a - \epsilon_b \leq q_V \cdot \epsilon_\text{MAC}$
-\end{itemize}
-
-\begin{theorem}\label{mac-fo-theorem}
-    Let $(\keygen, E, D)$ be a public-key encryption scheme that is $\delta$-correct and that has $\gamma$-spread.
-For every OW-PCVA adversary against the transformed encryption scheme $(E^T, D^T)$ that makes $q_G$ hash queries and $q_V$ ciphertext validation queries and that has advantage $\epsilon_\text{OW-PCVA}$, there exists an IND-CPA adversary against $(\keygen, E, D)$ with advantage $\epsilon_\text{IND-CPA}$ and an EF-CMA adversary against the MAC with advantage $\epsilon_\text{MAC}$ such that:
+\begin{theorem}
+    Let $\pke$ be $\delta$-correct and have $\gamma$-spread.
+For every OW-PCVA adversary $\mathcal{A}_1$ against $\pke_1$ that makes $q_V$ ciphertext validation queries and $q_G$ hash queries, and that has advantage $\epsilon_1$, there exists an OW-CPA adversary against $\pke$ with advantage $\epsilon_{\pke}$ and an EF-CMA adversary against $\mac$ with advantage $\epsilon_{\mac}$ such that
 
     \begin{equation*}
-        \epsilon_\text{OW-PCVA}
-        \leq q_G \cdot \delta 
-        + q_V \cdot \epsilon_\text{MAC}
-        + \frac{2q+1}{\norm{\mathcal{M}}} 
-        + \epsilon_\text{IND-CPA}
+        \epsilon_{\pke_1} \leq q_G \cdot \delta + q_V \cdot \epsilon_{\mac} + q_G \cdot \epsilon_{\pke}
     \end{equation*}
@@
 
-The proof will be discussed in section \ref{proof-of-mac-fo-theorem}.
+The proof follows a sequence of games that begins with the standard OW-PCVA game but incrementally replaces the PCO and the CVO with alternative implementations that can be simulated by an OW-CPA adversary and/or the EF-CMA adversary who does not have the secret key.
 
-\section{Proof techniques}
-Similar to the 2017 paper, we will use a sequence of games to incrementally replace PCO, CVO, and other aspects of the standard OW-PCVA game with some simulation.
-Then we will show that the total loss of security is negligible.
+\begin{proof}
+    Let $\mathcal{A}$ denote an adversary against the OW-PCVA security of $\enc_1$. Consider the following sequence of games.
 
-\subsection{Replacing plaintext-checking oracle}\label{replacing-pco}
-The plaintext-checking oracle $\operatorname{PCO}$ takes as input a plaintext-ciphertext pair $(m, (c, t))$ and return ``reject'' if and only if the the ciphertext is a valid encryption of the plaintext and vice versa under the context of some fixed keypair.
+    Game $G_0$ is the standard OW-PCVA game.
+In the original OW-PCVA game, queries to the random oracle $G$ are stored on a tape $\mathcal{L}^G$.
+By convention, we say that $G(m) = (r, k)$ if and only if $(m, r, k) \in \mathcal{L}^G$.
 
-The simulated plaintext-checking oracle $\operatorname{PCO}_1$ removes the requirement for the secret key by removing the step that checks whether the queried ciphertext $(c, t)$ decrypts back to the queried plaintext $m$.
-
-From the OW-PCVA adversary's perspective, the two oracles are indistinguishable, except for when there is a decryption error $D(\text{sk}, E(\text{pk}, m, r)) \neq m$ for the input plaintext $m$ and the corresponding coin $r$.
-The MAC is not involved in this argument.
-The loss of security is still $q_G\cdot\delta$
+    We can be more explicit with the implementation of the PCO and the CVO:
 
@@
             \SetAlgoLined
-        \caption{Vanilla PCO}
+            \caption{
+                Standard $\pco$ in $G_0$
+            }
             \KwIn{
-            $m \in \mathcal{M}, c \in \mathcal{C}, t \in \mathcal{T}$ 
+                $m \in \mathcal{M}, (c \in \mathcal{C}, t \in \mathcal{T})$
             }
-        \If{
-            $D^T(\text{sk}, (c, t)) \neq m$
-        }{
-            \Return{$\bot$}\;
-            \tcp*[h]{Decryption did not match}
-        }
-        \If{
-            $E^T(\text{pk}, m) \neq (c, t)$
-        }{
-            \Return{$\bot$}\;
-            \tcp*[h]{Encryption did not match}
+            $\hat{m} \leftarrow \dec(\operatorname{sk}, c)$\;
+            $(\hat{r}, \hat{k}) = G(\hat{m})$\;
+
+            \Return{
+                $\llbrack m = \hat{m} \rrbrack$ 
+                and $\llbrack \mac(\hat{k}, c) = t\rrbrack$
             }
@@
 
-    \columnbreak
+        \begin{algorithm}[H]
+            \SetAlgoLined
+            \caption{
+                Standard $\cvo$ in $G_0$
+            }
+            \KwIn{
+                $c \in \mathcal{C}, t \in \mathcal{T}$
+            }
+            $\hat{m} \leftarrow \dec(\operatorname{sk}, c)$\;
+            $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$\;
+            \Return{$\llbrack \mac(\hat{k}, c) = t\rrbrack$}
+        \end{algorithm}
+    \end{multicols}
+
+    Game $G_1$ is identical to $G_0$, except that the CVO is replaced with an alternative implementation.
+Instead of executing the decryption routine $\hat{m} \leftarrow \dec(\operatorname{sk}, (c, t))$ to recover the plaintext, the alternative implementation checks previous hash queries in $\mathcal{L}^G$ to recover the plaintext.
 
@@
         \SetAlgoLined
-        \caption{Simulated $\operatorname{PCO}_1$}
-        \KwIn{
-            $m \in \mathcal{M}, c \in \mathcal{C}, t \in \mathcal{T}$
+        \caption{
+            $\cvo_1$ in $G_1$
         }
-        \If{
-            $E^T(\text{pk}, m) \neq (c, t)$
-        }{
-            \Return{$\bot$}\;
-            \tcp*[h]{Encryption did not match}
-        }
-    \end{algorithm}
-\end{multicols}
-
-\subsection{Replacing ciphertext-validation oracle}\label{replacing-cvo}
-This is a similar trick to what was used in the 2017 paper, as well.
-Where the ciphertext is honestly generated, both oracles will return accept.
-Where the vanilla CVO rejects the ciphertext (aka ciphertext is malformed or the tag doesn't match), the simulated CVO will also reject.
-Therefore, the diverging event is when the vanilla CVO accepts but the simulated CVO rejects.
-Since the vanilla CVO accepts the query, the tag $t$ must be valid for the queried ciphertext $c$; on the other hand, the simulated CVO's rejection means there is no matching query in the hash oracle.
-Under the random oracle assumption, $t$ is a valid tag for some data $c$ under some key that the adversary does not know.
-In other words, $(c, t)$ is some kind of forgery.
-
-\textbf{The argument on how the security of the MAC relates to this diverging event is still a bit fuzzy}, but here are two possible ways I can think of:
-
-\begin{itemize}
-    \item For each CVO query, the adversary is trying to forge tag for that specific ciphertext.
-This means there is a selective forgery attack, and over all $q_V$ validation queries, the probability of having at least one selective forgery attack that works is at most $q_V\epsilon_\text{SF}$
-
-    \item Across all CVO query, the adversary wants to forge tag for some ciphertext, meaning there is a existential forgery attack, and the probability is at most $\epsilon_\text{EF}$
-\end{itemize}
-
-\begin{multicols}{2}
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Vanilla CVO}
         \KwIn{
@@
         \If{
-            $D^T(\text{sk}, (c, t)) = \bot$
+            $\exists (m, r, k) \in \mathcal{L}^G$ such that
+            $\mac(k, c) = t$
         }{
+            \Return{$1$}\;
+        }
         \Return{$\bot$}\;
+    \end{algorithm}
+
+    Consider a single ciphertext query $(c, t)$.
+If the query is accepted by $\cvo_1$, then this cipehrtext query is honestly generated, so the correctness of $\pke$ and $\mac$ implies that the query will also be accepted by $\cvo$.
+Similarly, if the query is rejected by $\cvo$, then either the re-encryption does not match, or the MAC does not match; in any case, $\cvo_1$ will also reject the ciphertext query.
+Therefore, these two implementations disagree if and only if $\cvo$ accepts the query but $\cvo_1$ rejects the query.
+
+    Since $\cvo$ accepts the query, $t$ is a valid tag for $c$ under some MAC key $k$ derived by hashing the decryption $m \leftarrow \dec(\operatorname{sk}, c)$.
+Since there is no matching query in $\mathcal{L}^G$, under the random oracle assumption, the MAC key $k$ is indistinguishable from truly random.
+In other words, for some ciphertext $c$ of the adversary's choosing, $t$ is a forgery under an unknown key.
+Therefore, for a single ciphertext query, the probability that $\cvo$ and $\cvo_1$ disagree is bounded by the advantage of some existential forgery adversary: $P[\cvo(c, t) \neq \cvo_1(c, t)] \leq \epsilon_{\mac}$.
+
+    Game 1 and game 0 behave differently if one or more among all $q_V$ ciphertext validation queries satisfies the condition above, therefore:
+
+    \begin{equation*}
+        \epsilon_0 - \epsilon_1 \leq q_V \cdot \epsilon_{\mac}
+    \end{equation*}
+
+    Game 2 $G_2$ is identical to $G_1$, except that the standard $\pco$ is replaced by an alternative implementation $\pco_1$
+
+    \begin{algorithm}[H]
+        \SetAlgoLined
+        \caption{$\pco_1$ in $G_2$}
+        \KwIn{
+            $m \in \mathcal{M}, (c \in \mathcal{C}, t \in \mathcal{T})$
+        }
+        $(r, k) \leftarrow G(m)$\;
+        \Return{
+            $\llbrack \enc(\operatorname{pk}, m; r) = c \rrbrack$
+            and $\llbrack \mac(k, c) = t \rrbrack$
         }
@@
 
-    \columnbreak
+    % TODO: 
+\end{proof}
 
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Simulated $\operatorname{CVO}_1$}
-        \KwIn{
-            $(c \in \mathcal{C}, t \in \mathcal{T})$
-        }
-        \If{
-            $\exists (m, r, k) \in \mathcal{O}^G$ such that $\operatorname{MAC}(k, c) = t$
-        }{
-            \Return{Accept}\;
-        }
-        \Return{$\bot$}
-    \end{algorithm}
-\end{multicols}
-
-\subsection{Random until queried}\label{random-until-queried}
-We then replace the pseudorandom coin and the MAC key with truly random coins and truly random keys in the challenge encryption.
-These two challenge encryption routines are identical, unless the adversary queries $G$ with the challenge plaintext $m^\ast$, but the probability of making such query can be bounded.
-
-\begin{multicols}{2}
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Challenge encryption}
-        $m^\ast \leftsample \mathcal{M}$\;
-        $(c^\ast, t^\ast) \leftarrow E^T(\text{pk}, m^\ast)$\;
-        \Return{$(c^\ast, t^\ast)$}
-    \end{algorithm}
-
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Simulated challenge encryption}
-        $m^\ast \leftsample \mathcal{M}$\;
-        $r^\ast \leftsample \text{Coin}, k^\ast \leftsample \mathcal{K}_\text{MAC}$\;
-        $c^\ast \leftarrow E(\text{pk}, m^\ast, r^\ast)$\;
-        $t^\ast \leftarrow \operatorname{MAC}(k^\ast, c^\ast)$
-        \Return{$(c^\ast, t^\ast)$}
-    \end{algorithm}
-\end{multicols}
-
-Let the OW-PCVA adversary's advantage using the vanilla challenge encryption be $\epsilon_0$ and its advantage using the simulated challenge encryption be $\epsilon_1$.
-Denote the event that the adversary queries the hash oracle $G$ with the challenge plaintext $m^\ast$ by $\operatorname{QUERY}^\ast$, then:
-
-\begin{equation*}
-    \epsilon_0 - \epsilon_1 \leq P[\operatorname{QUERY}^\ast]
-\end{equation*}
-
-We can build an IND-CPA adversary against the underlying encryption scheme to bound $P[\operatorname{QUERY}^\ast]$: if the OW-PCVA adversary makes the magic query, then the IND-CPA adversary can find it in the hash oracle's tape and win the IND-CPA game; if the OW-PCVA adversary does not make the magic query, then the IND-CPA adversary outputs a blind guess:
-
-\begin{equation*}
-    P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}
-\end{equation*}
-
-\section{Complete proof of \ref{mac-fo-theorem}}\label{proof-of-mac-fo-theorem}
-\begin{proof}
-    We will prove using a sequence of games. Game 0 is the standard OW-PCVA game.
-
-    Game 1 is identical to game 0, except PCO is replaced with $\operatorname{PCO}_1$, the loss of security is described in section \ref{replacing-pco}:
-
-    \begin{equation*}
-        \epsilon_0 - \epsilon_1 \leq q_G \cdot \delta
-    \end{equation*}
-
-    Game 2 is identical to game 1, except CVO is replaced with $\operatorname{CVO}_1$, the loss of security is described in section \ref{replacing-cvo}.
-Here I put a placeholder $\epsilon_\text{MAC}$ to indicate that this quantity is tied to the security of the MAC, but I don't have a solid answer yet
-
-    \begin{equation*}
-        \epsilon_1 - \epsilon_2 \leq \epsilon_\text{MAC}
-    \end{equation*}
-
-    Game 3 is identical to game 2, except the challenge encryption is replaced with the simulated challenge encryption
-
-    \begin{equation*}
-        \epsilon_3 - \epsilon_2 \leq P[\operatorname{QUERY}^\ast]
-    \end{equation*}
-
-    From section \ref{random-until-queried} we know that $P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}$ for some IND-CPA adversary against the underlying PKE.
-
-    Game 3 can be entirely simulated by an OW-CPA adversary against the underlying PKE:
-
-    \begin{itemize}
-        \item The keypairs are identical between the two PKE's
-        \item The OW-CPA adversary can simulate PCO, CVO, and hash oracle $G$
-        \item The OW-CPA adversary receives challenge ciphertext $c^\ast$ which is computed from a truly random coin; it can then random a truly random MAC key and compute the tag $t^\ast$
-        \item The OW-CPA adversary passes $(c^\ast, t^\ast)$ to the OW-PCVA adversary and returns whatever the OW-PCVA adversary returns
-    \end{itemize}
-
-    The OW-CPA adversary wins if and only if the OW-PCVA adversary wins: $\epsilon_3 = \epsilon_\text{OW-CPA}$.
-
-    Putting everything together we have:
-
-    \begin{equation*}
-        \epsilon_0 
-        \leq q_G \cdot \delta 
-        + \epsilon_\text{MAC} 
-        + \epsilon_\text{IND-CPA} 
-        + \epsilon_\text{OW-CPA}
-    \end{equation*}
-\end{proof}
+\section{Experimental results}
 

### 2024-05-09 lattices/primitives/src/lib.rs
Implemented hash H, J, G

@@ -2,10 +2,6 @@
 #![no_std]
 
 use crate::algebra::FieldElem;
-use sha3::{
-    digest::{ExtendableOutput, Update, XofReader},
-    Shake128, Shake256,
-};
 
 /// word size of all arithmetics
 pub type Word = u64;
@@ -21,26 +17,7 @@ pub const KYBER_K_1024: usize = 4;
 pub const SEEDSIZE: usize = 32;
 
 pub mod algebra;
-
-pub fn shake256_prf<const ETA: usize>(seed: [u8; SEEDSIZE], ctr: u8) -> [u8; ETA] {
-    let mut hasher = Shake256::default();
-    hasher.update(&seed);
-    hasher.update(&[ctr]);
-    let mut xof = hasher.finalize_xof();
-
-    let mut output = [0u8; ETA];
-    xof.read(&mut output);
-    return output;
-}
-
-pub fn shake128_xof(seed: [u8; SEEDSIZE], i: u8, j: u8) -> impl XofReader {
-    let mut hasher = Shake128::default();
-    hasher.update(&seed);
-    hasher.update(&[i, j]);
-    let xof = hasher.finalize_xof();
-
-    return xof;
-}
+pub mod symmetric;
 
 /// Algorithm 4: ByteEncode
 /// Encode an integer array into a byte array by encoding each integer using d bits

### 2024-05-10 lattices/primitives/src/symmetric.rs
Implemented hash H, J, G (#6)

@@ -1,0 +1,110 @@
+//! Hash, XOF, and PRF
+use crate::SEEDSIZE;
+use sha3::{
+    digest::{ExtendableOutput, Update, XofReader},
+    Digest, Sha3_256, Sha3_512, Shake128, Shake256,
+};
+
+/// Hash function H is Sha3-256(s)
+pub fn hash_h(stream: &[u8]) -> [u8; SEEDSIZE] {
+    let mut output = [0u8; SEEDSIZE];
+    let mut hasher = Sha3_256::new();
+    Digest::update(&mut hasher, stream);
+    let result = hasher.finalize();
+    output.clone_from_slice(&result);
+
+    return output;
+}
+
+/// Hash function J is Shake256(s)
+pub fn hash_j(stream: &[u8]) -> [u8; SEEDSIZE] {
+    let mut output = [0u8; SEEDSIZE];
+
+    let mut hasher = Shake256::default();
+    hasher.update(stream);
+    let mut xof = hasher.finalize_xof();
+    xof.read(&mut output);
+
+    return output;
+}
+
+/// Hash function G uses SHA3-512 but output 2 32-bytes
+pub fn hash_g(stream: &[u8]) -> ([u8; SEEDSIZE], [u8; SEEDSIZE]) {
+    let mut hasher = Sha3_512::new();
+    Digest::update(&mut hasher, stream);
+    let hash = hasher.finalize();
+
+    let mut first = [0u8; SEEDSIZE];
+    let mut second = [0u8; SEEDSIZE];
+
+    first.clone_from_slice(hash.get(0..SEEDSIZE).expect("Unexpected out-of-bound"));
+    second.clone_from_slice(
+        hash.get(SEEDSIZE..(2 * SEEDSIZE))
+            .expect("Unexpected out-of-bound"),
+    );
+
+    return (first, second);
+}
+
+pub fn shake256_prf<const ETA: usize>(seed: [u8; SEEDSIZE], ctr: u8) -> [u8; ETA] {
+    let mut hasher = Shake256::default();
+    hasher.update(&seed);
+    hasher.update(&[ctr]);
+    let mut xof = hasher.finalize_xof();
+
+    let mut output = [0u8; ETA];
+    xof.read(&mut output);
+    return output;
+}
+
[…]

### 2024-05-10 lattices/primitives/src/algebra.rs
checkpoint

@@ -496,7 +496,7 @@ impl Poly {
     }
 
     /// Sample a polynomial from CBD(-eta, -eta+1, ..., eta-1, eta)
-    fn sample_cbd(eta: usize, uniform: &[u8]) -> Self {
+    pub fn sample_cbd(eta: usize, uniform: &[u8]) -> Self {
         // a single sample from CBD(eta) requires (2 * eta) uniform bits, each polynomial has 256
         // coefficients, hence a total of (64 * eta) bytes are needed
         // TODO: make ETA a compile-time constant instead of runtime argument

### 2024-05-15 lattices/primitives/src/algebra.rs
Mock implementation of keygen

@@ -715,6 +715,18 @@ impl PolyNTTVec {
 
         return product;
     }
+
+    /// Coordinate-wise addition between two PolyNTT vectors
+    /// TODO: consider doing an in-place implementation, as well
+    pub fn add(&self, other: &Self) -> Self {
+        let mut sum = Self::ZERO;
+
+        for i in 0..KYBER_K {
+            sum.vec[i] = self.vec[i].polyadd(&other.vec[i]);
+        }
+
+        return sum;
+    }
 }
 
 /// A matrix of Polynomials in NTT domain.

### 2024-05-15 lattices/primitives/src/lib.rs

@@ -1,7 +1,7 @@
 //! Primitives used in lattice-based cryptography
 #![no_std]
 
-use algebra::{Poly, PolyNTT};
+use algebra::{Poly, PolyNTT, PolyNTTMatrix, PolyNTTVec};
 use symmetric::shake128_xof;
 
 use crate::{algebra::FieldElem, symmetric::hash_g};
@@ -81,6 +81,39 @@ pub fn byte_decode(coeffs: &mut [FieldElem], d: usize, buffer: &[u8]) {
 // 3. The secret key is encoded to be decoded later
 // 4. Server decode public key from a byte stream
 
+/// Given a seed d, return the full keypair
+///
+/// Keypair encoding will be performed at later stage
+pub fn keygen_with_seed(seed: [u8; SEEDSIZE]) -> (PublicKey, SecretKey) {
+    let (pub_seed, sec_seed) = hash_g(&seed); // correspond to rho, sigma respectively
+    let a = PolyNTTMatrix::sample_uniform(pub_seed);
+    let mut ctr = 0;
+    let mut sols = PolyNTTVec::ZERO;
+    let mut errs = PolyNTTVec::ZERO;
+
+    for i in 0..KYBER_K {
+        sols.vec[i] = Poly::sample_cbd_eta1(&sec_seed, ctr).ntt();
+        ctr += 1;
+    }
+
+    for i in 0..KYBER_K {
+        errs.vec[i] = Poly::sample_cbd_eta1(&sec_seed, ctr).ntt();
+        ctr += 1;
+    }
+
+    let t = a.dot(&sols).add(&errs);
+
+    let pk = PublicKey {
+        seed: pub_seed,
+        a,
+        t,
+    };
+
+    let sk = SecretKey { sols, errs };
+
+    return (pk, sk);
+}
+
 /// The public key includes the noisy linear system (A, t = As + e). In addition, the seed used to
 /// pseudorandomly generate A is tracked so the public key can be compactly encoded.
 pub struct PublicKey {
@@ -88,67 +121,18 @@ pub struct PublicKey {
     pub seed: [u8; SEEDSIZE],
 
     /// The k * k public matrix A
-    pub a: [[PolyNTT; KYBER_K]; KYBER_K],
+    pub a: PolyNTTMatrix,
 
     /// The noisy samples: t = A * s + e
-    pub t: [PolyNTT; KYBER_K],
-}
-
[…]

### 2024-05-18 lattices/compact-fo-using-mac/manuscript/main.tex
getting started with revised proof
Progress on improved proof

--- 2024-05-01 lattices/compact-fo-using-mac/manuscript/main.tex
@@
 \newcommand{\keygen}{\operatorname{KeyGen}}
+\newcommand{\pk}{\operatorname{pk}}
+\newcommand{\sk}{\operatorname{sk}}
+\newcommand{\pco}{\operatorname{PCO}}
+\newcommand{\cvo}{\operatorname{CVO}}
+\newcommand{\llbrack}{[\![}
+\newcommand{\rrbrack}{]\!]}
 
@@
 \maketitle
+
+\section{OW-PCVA transformation}\label{sec:ow-pcva-transformation}
+Let $(K, E, D)$ be a public-key encryption scheme defined over $(\mathcal{K}_\text{PKE}, \mathcal{M}_\text{PKE}, \mathcal{C})$.
+Let $(S, V)$ be a message authentication code defined over $(\mathcal{K}_\text{MAC}, \mathcal{M}_\text{MAC}, \mathcal{T})$.
+Let $G: \mathcal{M}_\text{PKE} \rightarrow \mathcal{K}_\text{MAC}$ be a hash function.
+The transformation outputs a public-key encryption scheme $(K, E_1, D_1)$.
+The key generation routine and key space of the transformed scheme are identical to those of the input scheme.
+The encryption and decryption routines of the transformed scheme are as follows:
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \caption{$E_1$}
+        \SetAlgoLined
+        \KwIn{
+            $\pk, m \in \mathcal{M}_\text{PKE}$
+        }
+        $k \leftarrow G(m)$\;
+        $\sigma \leftarrow E(\pk, m)$\;
+        $t \leftarrow S(k, \sigma)$\;
+        \Return{$c = (\sigma, t)$}\;
+    \end{algorithm}
+
+    \begin{algorithm}[H]
+        \caption{$D_1$}
+        \SetAlgoLined
+        \KwIn{
+            $\sk, c = (\sigma \in \mathcal{C}, t \in \mathcal{T})$
+        }
+        $\hat{m} \leftarrow D(\sk, \sigma)$\;
+        $\hat{k} \leftarrow G(\hat{m})$\;
+        \If{
+            $V(\hat{k}, \sigma, t) \neq 1$
+        }{
+            \Return{$\bot$}\;
+        }
+        \Return{$\hat{m}$}\;
+    \end{algorithm}
+\end{multicols}
+
+If the input PKE is deterministic, OW-CPA secure, and $\delta$-correct, and the input MAC is existentially unforgeable, then the ``encrypt-then-MAC'' PKE is OW-PCVA secure. More specifically:
+
+\begin{theorem}\label{thm:if-ow-cpa-then-ow-pcva}
+    For every OW-PCVA adversary against the transformed scheme $(E_1, D_1)$ who makes $q_G$ hash queries, $q_P$ plaintext checking queries, and $q_V$ ciphertext validation queries, and who has advantage $\epsilon_\text{OW-PCVA}$, there exists an OW-CPA adversary against the underlying PKE with advantage $\epsilon_\text{OW-CPA}$, a correctness adversary against the underlying PKE with advantage $\delta$, and a existential forgery adversary against the MAC with advantage $\epsilon_\text{MAC}$ such that:
+
+    \begin{equation*}
+        \epsilon_\text{OW-PCVA} 
+        \leq (q_P + q_G) \cdot \delta
+        + q_V \cdot \epsilon_\text{MAC}
+        + 2 \cdot \epsilon_\text{OW-CPA}
+    \end{equation*}
+\end{theorem}
+
+\begin{proof}
+    To prove theorem \ref{thm:if-ow-cpa-then-ow-pcva}, we use a sequence of game.
+
+    Game 0 is the OW-PCVA game, as described in the section above.
+
+    In Game 1, the PCO is replaced with an alternative implementation $\pco_1$:
+
+    \begin{multicols}{2}
+        \begin{algorithm}[H]
+            \caption{$\pco$}
+            \SetAlgoLined
+            \KwIn{
+                $(m, c = (\sigma, t))$
+            }
+            $\hat{m} \leftarrow D(\sk, \sigma)$\;
+            $\hat{k} \leftarrow G(\hat{m})$\;
+            \Return{
+                $\llbrack \hat{m} = m \rrbrack$ 
+                and $\llbrack V(\hat{k}, \sigma, t) \rrbrack$
+            }
+        \end{algorithm}
+
+        \begin{algorithm}[H]
+            \caption{$\pco_1$}
+            \SetAlgoLined
+            \KwIn{
+                $(m, c = (\sigma, t))$
+            }
+            $k \leftarrow G(m)$\;
+            $\hat{\sigma} \leftarrow E(\pk, m)$\;
+            \Return{
+                $\llbrack \sigma = \hat{\sigma} \rrbrack$
+                and $\llbrack V(k, \sigma, t) \rrbrack$
+            }
+        \end{algorithm}
+    \end{multicols}
+
+    For any single plaintext checking query, the two oracles will disagree if and only if correctness of $(K, E, D)$ is broken, so such probability can be bounded by $\delta$.
+From the OW-PCVA adversary's perspective, the two games behave differently if and only if the two oracles disagree on at least one of the plaintext checking queries, which is at most $q_P \cdot \delta$:
+
+    \begin{equation*}
+        \epsilon_0 - \epsilon_1 \leq q_P \cdot \delta
+    \end{equation*}
+
+    In Game 2, the $\cvo$ is replaced with an alternative implementation $\cvo_1$, the latter of which checks the hash oracle's tape $\mathcal{L}^G$
+
+    \begin{multicols}{2}
+        \begin{algorithm}[H]
+            \caption{$\cvo$}
+            \SetAlgoLined
+            \KwIn{
+                $c = (\sigma, t)$
+            }
+            $\hat{m} \leftarrow D(\sk, \sigma)$\;
+            $\hat{k} \leftarrow G(\hat{m})$\;
+            \Return{
+                $\llbrack V(\hat{k}, \sigma, t) = 1 \rrbrack$
+            }
+        \end{algorithm}
+
+        \begin{algorithm}[H]
+            \caption{$\cvo_1$}
+            \SetAlgoLined
+            \KwIn{
+                $c = (\sigma, t)$
+            }
+            \If{
+                $\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G$ s.t.
+                $E(\pk, \tilde{m}) = c$ and $V(\tilde{k}, \sigma, t) = 1$
+            }{
+                \Return{$1$}\;
+            }
+            \Return{$0$}\;
+        \end{algorithm}
+    \end{multicols}
+
+    There are two scenarios in which the two oracles can disagree.
+First, there is a matching hash query $(\tilde{m}, \tilde{k}) \in \mathcal{L}^G$, but $\tilde{m}$ triggers a decryption failure, so the same $\sigma$ decrypts to $\hat{m} \neq \tilde{m}$.
+For a single hash query, the probability of decryption error is at most $\delta$, so across $q_G$ hash queries, the probability of at least one decryption error is at most $q_G \cdot \delta$.
+Second, $(\sigma, t)$ is a valid message-tag pair that will pass $\cvo$, but there is no matching hash query.
+Under the random oracle model, from the perspective of the adversary $k \leftarrow G(m)$ is a truly random and unknown MAC key, which means that $(\sigma, t)$ is a forgery.
+The probability of a single ciphertet validation query being a valid forgery is bounded by the advantage of a MAC adversary, so across all $q_V$ ciphertext validation queries, the probability of having at least one valid forgery is at most $q_V\cdot\epsilon_\text{MAC}$.
+Finally, the probability of the two implementations disagree is at most the sum of the probabilities of the two scenarios:
+
+    \begin{equation*}
+        \epsilon_1 - \epsilon_2 \leq q_G \cdot \delta + q_V \cdot \epsilon_\text{MAC}
+    \end{equation*}
+
+    % TODO: in game 3, the challenge encryption is done using a truly random MAC key
+    %   talk about QUERY*
+
+    % TODO: game 3 can be simulated by a OW-CPA adversary
+
+\end{proof}
 

### 2024-05-20 lattices/compact-fo-using-mac/manuscript/main.tex
checkpoint

@@
 
-    % TODO: in game 3, the challenge encryption is done using a truly random MAC key
-    %   talk about QUERY*
+    In game 3, the challenge encryption routine is modified.
+Instead of computing a pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$, a truly random MAC key is sampled uniformly from the message space $k^\ast \leftsample \mathcal{K}_\text{MAC}$.
+Under the random oracle model, game 3 and game 2 are indistinguishable from the adversary's perspective unless it makes a hash query on the value of $m^\ast$.
+Denote the probability that the adversary makes a hash query on $m^\ast$ by $P[\operatorname{QUERY}\ast]$, then:
+
+    \begin{equation*}
+        \epsilon_2 - \epsilon_3 \leq P[\operatorname{QUERY}\ast]
+    \end{equation*}
+
+    % TODO: Bound the probabiolity P[QUERY*]
 

### 2024-05-22 lattices/compact-fo-using-mac/manuscript/main.tex
QUERY *

@@
 
-    % TODO: Bound the probabiolity P[QUERY*]
+    we can bound $P[\operatorname{QUERY}\ast]$ by constructing an OW-CPA adversary that simulates game 3 for a OW-PCVA adversary as a sub-routine.
+After the OW-PCVA adversary halts, the OW-CPA adversary checks through the tape of the hash oracle.
+Since the encryption routine is deterministic, if $m^\ast$ exists in the hash oracle tape, then it can be identified for sure.
+In other words, if $m^\ast$ has been queried by the OW-PCVA sub-routine, then the OW-CPA adversary can be guaranteed to win its game
 
-    % TODO: game 3 can be simulated by a OW-CPA adversary
+    \begin{equation*}
+        P[\operatorname{QUERY}\ast] = \epsilon_\text{OW-CPA}
+    \end{equation*}
+
+    Finally, game 3 can be simulated by the OW-CPA
 

### 2024-05-22 lattices/compact-fo-using-mac/manuscript/main.tex
REvised proof; removed bad proof

@@
 \newcommand{\cvo}{\operatorname{CVO}}
+\newcommand{\coin}{\operatorname{COIN}}
 \newcommand{\llbrack}{[\![}
 \newcommand{\rrbrack}{]\!]}
+\newlength{\wdth}
+\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}
 
@@
     Game 0 is the OW-PCVA game, as described in the section above.
+
+    \begin{equation*}
+        \epsilon_\text{OW-PCVA} = \epsilon_0
+    \end{equation*}
 
@@
 Instead of computing a pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$, a truly random MAC key is sampled uniformly from the message space $k^\ast \leftsample \mathcal{K}_\text{MAC}$.
+
+    \begin{multicols}{2}
+        \begin{algorithm}[H]
+            \caption{OW-PCVA game 2}
+            \SetAlgoLined
+            $(\pk, \sk) \leftarrow \keygen(1^\lambda)$\;
+            $m^\ast \leftsample \mathcal{M}_\text{PKE}$\;
+            $k^\ast \leftarrow G(m^\ast)$\;
+            $\sigma^\ast \leftarrow E(\pk, m^\ast)$\;
+            $t^\ast \leftarrow S(k^\ast, \sigma^\ast)$\;
+            $c^\ast = (\sigma^\ast, t^\ast)$\;
+            $\hat{m} \leftarrow \mathcal{A}_\text{OW-PCVA}^{
+                \mathcal{O}^G, \pco_1, \cvo_1
+            }(1^\lambda, \pk, c^\ast)$\;
+            \Return{$\llbrack \hat{m} = m^\ast \rrbrack$}
+        \end{algorithm}
+
+        \begin{algorithm}[H]
+            \caption{OW-PCVA game 3}
+            \SetAlgoLined
+            $(\pk, \sk) \leftarrow \keygen(1^\lambda)$\;
+            $m^\ast \leftsample \mathcal{M}_\text{PKE}$\;
+            $k^\ast \leftsample \mathcal{K}_\text{MAC}$\;
+            $\sigma^\ast \leftarrow E(\pk, m^\ast)$\;
+            $t^\ast \leftarrow S(k^\ast, \sigma^\ast)$\;
+            $c^\ast = (\sigma^\ast, t^\ast)$\;
+            $\hat{m} \leftarrow \mathcal{A}_\text{OW-PCVA}^{
+                \mathcal{O}^G, \pco_1, \cvo_1
+            }(1^\lambda, \pk, c^\ast)$\;
+            \Return{$\llbrack \hat{m} = m^\ast \rrbrack$}
+        \end{algorithm}
+    \end{multicols}
+
     Under the random oracle model, game 3 and game 2 are indistinguishable from the adversary's perspective unless it makes a hash query on the value of $m^\ast$.
@@
 
-    Finally, game 3 can be simulated by the OW-CPA
+    Finally, game 3 can be simulated by another OW-CPA adversary.
+First, all three oracles $\pco_1, \cvo_1$ and $G$ can be simulated using the same public key.
+When the OW-CPA challenger produces the the challenge ciphertext $\sigma^\ast \in \mathcal{C}$, the OW-CPA adversary samples a random MAC key $k^\ast \leftarrow \mathcal{K}_\text{MAC}$ and computes the tag $t^\ast \leftarrow S(k^\ast, \sigma^\ast)$ before passing $c^\ast = (\sigma^\ast, t^\ast)$ to the OW-PCVA adversary as its challenge ciphertext.
+When the OW-PCVA adversary halts, the OW-CPA adversary passes OW-PCVA adversary's output as its own output.
+It's easy to see that the OW-CPA adversary wins if and only if the OW-PCVA adversary wins:
 
+    \begin{equation*}
+        \epsilon_3 = \epsilon_\text{OW-CPA}
+    \end{equation*}
+
+    Putting all the inequalities above together gives us the desired inequality.
 \end{proof}
 
-\section{OW-PCVA transformation}
-Let $(\keygen, E, D)$ define a probabilistic public-key encryption scheme, then $T$ is a transformation that takes this scheme and outputs a second public-key encryption scheme.
-The transformation makes use of two additional components:
+\subsection{Deterministic OW-CPA transformation}
+In the OW-PCVA transformation, we required the input encryption scheme to be deterministic.
+In this section, we present a transformation that takes a probabilistic PKE and output a deterministic PKE that provides identical level of OW-CPA security.
 
-\begin{enumerate}
-    \item A $\operatorname{MAC}: \mathcal{K}^\text{MAC} \times \{0,1\}^\ast \rightarrow \mathcal{T}$
-    \item A hash function $G: \mathcal{M} \rightarrow \text{Coin} \times \mathcal{K}^\text{MAC}$
-\end{enumerate}
+Let $(K, E, D)$ be a PKE defined over $(\mathcal{K}, \mathcal{M}, \mathcal{C})$.
+We assume that the encryption routine is probabilistic and its randomness is derived from some seed $r \in \coin$.
+The encryption routine accepts as an argument a coin value that will be used to pseudorandomly derive all randomness in the routine.
+Let $H$ be a hash function $H: \mathcal{M} \rightarrow \coin$.
 
-\begin{multicols}{2}
+The transformed PKE is defined over the same spaces and shares the key generation and decryption routine with the input PKE, The only difference lies with its encryption routine:
+
 \begin{algorithm}[H]
+    \caption{$E^{\strike{\$}}$}
     \SetAlgoLined
-        \caption{OW-PCVA encryption $E^T$}
     \KwIn{
-            $\text{pk}$, $m \in \mathcal{M}$
+        $(\pk, m)$
     }
-        $(r, k) \leftarrow G(m)$\;
-        $c \leftarrow E(\text{pk}, m, r)$\;
-        $t \leftarrow \operatorname{MAC}(k, c)$
-        \tcp*[h]{"encrypt-then-mac"}\;
-        \Return{$(c, t)$}\;
+    $r \leftarrow H(m)$\;
+    $\sigma \leftarrow E(\pk, m; r)$\;
+    \Return{$\sigma$}
 \end{algorithm}
 
-    \columnbreak
+The transformed PKE is deterministic, and offers comparable level of OW-CPA security as the input PKE. More specifically:
 
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{OW-PCVA decryption $D^T$}
-        \KwIn{
-            $\text{sk}, c \in \mathcal{C}, t \in \mathcal{T}$
-        }
-        $\hat{m} \leftarrow D(\text{sk}, c)$\;
-        $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$\;
-        \If{
-            $\operatorname{MAC}(\hat{k}, c) \neq t$
-        }{
-            \Return{$\bot$}\;
-        }
-        \Return{$\hat{m}$}\;
-    \end{algorithm}
-\end{multicols}
-
-Instead of using re-encryption to check for the integrity of the ciphertext, a MAC tag is used.
-At the cost of adding a fixed number of bytes to the ciphertext, the decryption routine becomes significantly faster and memory-efficient.
-The memory trade-off is especially important since in Kyber, re-encryption needs to compute the entire public key $A \in R_q^{k \times k}$ from the 32-byte seed.
-
-We claim that this transformation is still OW-PCVA. The proof will be largely similar to that of the original OW-PCVA transformation proof, but with some notable differences:
-
-\begin{itemize}
-    \item When queried on some authenticated ciphertext $(\tilde{c}, \tilde{t})$, the secret-key-less $\operatorname{CVO}_1$ searches the hash oracle $G$ for $(m, r, k)$ such that $\operatorname{MAC}(k, \tilde{c}) = \tilde{t}$.
-The diverging event between using $\operatorname{CVO}$ and using $\operatorname{CVO}_1$ is ``adversary produces valid authenticated ciphertext without consulting the hash oracle'', which is equivalent to selective forgery (or existential forgery, I am not sure which): $\epsilon_a - \epsilon_b \leq q_V \cdot \epsilon_\text{MAC}$
-\end{itemize}
-
-\begin{theorem}\label{mac-fo-theorem}
-    Let $(\keygen, E, D)$ be a public-key encryption scheme that is $\delta$-correct and that has $\gamma$-spread.
-For every OW-PCVA adversary against the transformed encryption scheme $(E^T, D^T)$ that makes $q_G$ hash queries and $q_V$ ciphertext validation queries and that has advantage $\epsilon_\text{OW-PCVA}$, there exists an IND-CPA adversary against $(\keygen, E, D)$ with advantage $\epsilon_\text{IND-CPA}$ and an EF-CMA adversary against the MAC with advantage $\epsilon_\text{MAC}$ such that:
+\begin{theorem}
+    For every OW-CPA adversary against $(K, E^{\strike{\$}}), D$ that makes $q_H$ hash queries to $H$ and has advantage $\epsilon_{\strike{\$}}$, there exists an OW-CPA adversary against $(K, E, D)$ with advantage $\epsilon_\text{OW-CPA}$ such that:
 
     \begin{equation*}
-        \epsilon_\text{OW-PCVA}
-        \leq q_G \cdot \delta 
-        + q_V \cdot \epsilon_\text{MAC}
-        + \frac{2q+1}{\norm{\mathcal{M}}} 
-        + \epsilon_\text{IND-CPA}
+        \epsilon_{\strike{\$}} \leq (1 + q_H) \cdot \epsilon_\text{OW-CPA}
     \end{equation*}
@@
 
-The proof will be discussed in section \ref{proof-of-mac-fo-theorem}.
+\begin{proof}
+    % Game 1: replace pseudorandom coin with random coin and use QUERY* argument, this time we will lose q_H * epsilon amount of security, probably
 
-\section{Proof techniques}
-Similar to the 2017 paper, we will use a sequence of games to incrementally replace PCO, CVO, and other aspects of the standard OW-PCVA game with some simulation.
-Then we will show that the total loss of security is negligible.
-
-\subsection{Replacing plaintext-checking oracle}\label{replacing-pco}
-The plaintext-checking oracle $\operatorname{PCO}$ takes as input a plaintext-ciphertext pair $(m, (c, t))$ and return ``reject'' if and only if the the ciphertext is a valid encryption of the plaintext and vice versa under the context of some fixed keypair.
-
-The simulated plaintext-checking oracle $\operatorname{PCO}_1$ removes the requirement for the secret key by removing the step that checks whether the queried ciphertext $(c, t)$ decrypts back to the queried plaintext $m$.
-
-From the OW-PCVA adversary's perspective, the two oracles are indistinguishable, except for when there is a decryption error $D(\text{sk}, E(\text{pk}, m, r)) \neq m$ for the input plaintext $m$ and the corresponding coin $r$.
-The MAC is not involved in this argument.
-The loss of security is still $q_G\cdot\delta$
-
-\begin{multicols}{2}
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Vanilla PCO}
-        \KwIn{
-            $m \in \mathcal{M}, c \in \mathcal{C}, t \in \mathcal{T}$ 
-        }
-        \If{
-            $D^T(\text{sk}, (c, t)) \neq m$
-        }{
-            \Return{$\bot$}\;
-            \tcp*[h]{Decryption did not match}
-        }
-        \If{
-            $E^T(\text{pk}, m) \neq (c, t)$
-        }{
-            \Return{$\bot$}\;
-            \tcp*[h]{Encryption did not match}
-        }
-    \end{algorithm}
-
-    \columnbreak
-
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Simulated $\operatorname{PCO}_1$}
-        \KwIn{
-            $m \in \mathcal{M}, c \in \mathcal{C}, t \in \mathcal{T}$
-        }
-        \If{
-            $E^T(\text{pk}, m) \neq (c, t)$
-        }{
-            \Return{$\bot$}\;
-            \tcp*[h]{Encryption did not match}
-        }
-    \end{algorithm}
-\end{multicols}
-
-\subsection{Replacing ciphertext-validation oracle}\label{replacing-cvo}
-This is a similar trick to what was used in the 2017 paper, as well.
-Where the ciphertext is honestly generated, both oracles will return accept.
-Where the vanilla CVO rejects the ciphertext (aka ciphertext is malformed or the tag doesn't match), the simulated CVO will also reject.
-Therefore, the diverging event is when the vanilla CVO accepts but the simulated CVO rejects.
-Since the vanilla CVO accepts the query, the tag $t$ must be valid for the queried ciphertext $c$; on the other hand, the simulated CVO's rejection means there is no matching query in the hash oracle.
-Under the random oracle assumption, $t$ is a valid tag for some data $c$ under some key that the adversary does not know.
-In other words, $(c, t)$ is some kind of forgery.
-
-\textbf{The argument on how the security of the MAC relates to this diverging event is still a bit fuzzy}, but here are two possible ways I can think of:
-
-\begin{itemize}
-    \item For each CVO query, the adversary is trying to forge tag for that specific ciphertext.
-This means there is a selective forgery attack, and over all $q_V$ validation queries, the probability of having at least one selective forgery attack that works is at most $q_V\epsilon_\text{SF}$
-
-    \item Across all CVO query, the adversary wants to forge tag for some ciphertext, meaning there is a existential forgery attack, and the probability is at most $\epsilon_\text{EF}$
-\end{itemize}
-
-\begin{multicols}{2}
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Vanilla CVO}
-        \KwIn{
-            $(c \in \mathcal{C}, t \in \mathcal{T})$
-        }
-        \If{
-            $D^T(\text{sk}, (c, t)) = \bot$
-        }{
-            \Return{$\bot$}\;
-        }
-    \end{algorithm}
-
-    \columnbreak
-
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Simulated $\operatorname{CVO}_1$}
-        \KwIn{
-            $(c \in \mathcal{C}, t \in \mathcal{T})$
-        }
-        \If{
-            $\exists (m, r, k) \in \mathcal{O}^G$ such that $\operatorname{MAC}(k, c) = t$
-        }{
-            \Return{Accept}\;
-        }
-        \Return{$\bot$}
-    \end{algorithm}
-\end{multicols}
-
-\subsection{Random until queried}\label{random-until-queried}
-We then replace the pseudorandom coin and the MAC key with truly random coins and truly random keys in the challenge encryption.
-These two challenge encryption routines are identical, unless the adversary queries $G$ with the challenge plaintext $m^\ast$, but the probability of making such query can be bounded.
-
-\begin{multicols}{2}
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Challenge encryption}
-        $m^\ast \leftsample \mathcal{M}$\;
-        $(c^\ast, t^\ast) \leftarrow E^T(\text{pk}, m^\ast)$\;
-        \Return{$(c^\ast, t^\ast)$}
-    \end{algorithm}
-
-    \begin{algorithm}[H]
-        \SetAlgoLined
-        \caption{Simulated challenge encryption}
-        $m^\ast \leftsample \mathcal{M}$\;
-        $r^\ast \leftsample \text{Coin}, k^\ast \leftsample \mathcal{K}_\text{MAC}$\;
-        $c^\ast \leftarrow E(\text{pk}, m^\ast, r^\ast)$\;
-        $t^\ast \leftarrow \operatorname{MAC}(k^\ast, c^\ast)$
-        \Return{$(c^\ast, t^\ast)$}
-    \end{algorithm}
-\end{multicols}
-
-Let the OW-PCVA adversary's advantage using the vanilla challenge encryption be $\epsilon_0$ and its advantage using the simulated challenge encryption be $\epsilon_1$.
-Denote the event that the adversary queries the hash oracle $G$ with the challenge plaintext $m^\ast$ by $\operatorname{QUERY}^\ast$, then:
-
-\begin{equation*}
-    \epsilon_0 - \epsilon_1 \leq P[\operatorname{QUERY}^\ast]
-\end{equation*}
-
-We can build an IND-CPA adversary against the underlying encryption scheme to bound $P[\operatorname{QUERY}^\ast]$: if the OW-PCVA adversary makes the magic query, then the IND-CPA adversary can find it in the hash oracle's tape and win the IND-CPA game; if the OW-PCVA adversary does not make the magic query, then the IND-CPA adversary outputs a blind guess:
-
-\begin{equation*}
-    P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}
-\end{equation*}
-
-\section{Complete proof of \ref{mac-fo-theorem}}\label{proof-of-mac-fo-theorem}
-\begin{proof}
-    We will prove using a sequence of games. Game 0 is the standard OW-PCVA game.
-
-    Game 1 is identical to game 0, except PCO is replaced with $\operatorname{PCO}_1$, the loss of security is described in section \ref{replacing-pco}:
-
-    \begin{equation*}
-        \epsilon_0 - \epsilon_1 \leq q_G \cdot \delta
-    \end{equation*}
-
-    Game 2 is identical to game 1, except CVO is replaced with $\operatorname{CVO}_1$, the loss of security is described in section \ref{replacing-cvo}.
-Here I put a placeholder $\epsilon_\text{MAC}$ to indicate that this quantity is tied to the security of the MAC, but I don't have a solid answer yet
-
-    \begin{equation*}
-        \epsilon_1 - \epsilon_2 \leq \epsilon_\text{MAC}
-    \end{equation*}
-
-    Game 3 is identical to game 2, except the challenge encryption is replaced with the simulated challenge encryption
-
-    \begin{equation*}
-        \epsilon_3 - \epsilon_2 \leq P[\operatorname{QUERY}^\ast]
-    \end{equation*}
-
-    From section \ref{random-until-queried} we know that $P[\operatorname{QUERY}^\ast] \leq \epsilon_\text{IND-CPA}$ for some IND-CPA adversary against the underlying PKE.
-
-    Game 3 can be entirely simulated by an OW-CPA adversary against the underlying PKE:
-
-    \begin{itemize}
-        \item The keypairs are identical between the two PKE's
-        \item The OW-CPA adversary can simulate PCO, CVO, and hash oracle $G$
-        \item The OW-CPA adversary receives challenge ciphertext $c^\ast$ which is computed from a truly random coin; it can then random a truly random MAC key and compute the tag $t^\ast$
-        \item The OW-CPA adversary passes $(c^\ast, t^\ast)$ to the OW-PCVA adversary and returns whatever the OW-PCVA adversary returns
-    \end{itemize}
-
-    The OW-CPA adversary wins if and only if the OW-PCVA adversary wins: $\epsilon_3 = \epsilon_\text{OW-CPA}$.
-
-    Putting everything together we have:
-
-    \begin{equation*}
-        \epsilon_0 
-        \leq q_G \cdot \delta 
-        + \epsilon_\text{MAC} 
-        + \epsilon_\text{IND-CPA} 
-        + \epsilon_\text{OW-CPA}
-    \end{equation*}
+    % Game 1 can be simulated by OW-CPA adversary
 \end{proof}
 
-\section{Open questions}
-Can we get rid of the coin and just let the encryption scheme be probabilistic? Do we need to?
-
 \end{document}

### 2024-05-23 lattices/compact-fo-using-mac/manuscript/main.tex
Completed draft proof

@@
     Under the random oracle model, game 3 and game 2 are indistinguishable from the adversary's perspective unless it makes a hash query on the value of $m^\ast$.
-Denote the probability that the adversary makes a hash query on $m^\ast$ by $P[\operatorname{QUERY}\ast]$, then:
+Denote the probability that the adversary makes a hash query on $m^\ast$ by $P[\operatorname{QUERY}^\ast]$, then:
 
     \begin{equation*}
-        \epsilon_2 - \epsilon_3 \leq P[\operatorname{QUERY}\ast]
+        \epsilon_2 - \epsilon_3 \leq P[\operatorname{QUERY}^\ast]
     \end{equation*}
 
-    we can bound $P[\operatorname{QUERY}\ast]$ by constructing an OW-CPA adversary that simulates game 3 for a OW-PCVA adversary as a sub-routine.
+    we can bound $P[\operatorname{QUERY}^\ast]$ by constructing an OW-CPA adversary that simulates game 3 for a OW-PCVA adversary as a sub-routine.
 After the OW-PCVA adversary halts, the OW-CPA adversary checks through the tape of the hash oracle.
@@
     \begin{equation*}
-        P[\operatorname{QUERY}\ast] = \epsilon_\text{OW-CPA}
+        P[\operatorname{QUERY}^\ast] = \epsilon_\text{OW-CPA}
     \end{equation*}
@@
 \begin{theorem}
-    For every OW-CPA adversary against $(K, E^{\strike{\$}}), D$ that makes $q_H$ hash queries to $H$ and has advantage $\epsilon_{\strike{\$}}$, there exists an OW-CPA adversary against $(K, E, D)$ with advantage $\epsilon_\text{OW-CPA}$ such that:
+    For every OW-CPA adversary against the deterministic PKE $(K, E^{\strike{\$}}), D$ that makes $q_H$ hash queries to $H$ and has advantage $\epsilon_{\strike{\$}}$, there exists an OW-CPA adversary against the probabilistic PKE $(K, E, D)$ with advantage $\epsilon_\$$ such that:
 
     \begin{equation*}
-        \epsilon_{\strike{\$}} \leq (1 + q_H) \cdot \epsilon_\text{OW-CPA}
+        \epsilon_{\strike{\$}} \leq (1 + q_H) \cdot \epsilon_\$
     \end{equation*}
@@
 \begin{proof}
-    % Game 1: replace pseudorandom coin with random coin and use QUERY* argument, this time we will lose q_H * epsilon amount of security, probably
+    We prove using a sequence of game.
+Denote the OW-CPA adversary against the deterministic PKE by $\mathcal{A}_{\strike{\$}}$ and its advantage by $\epsilon_{\strike{\$}}$; denote the OW-CPA adversary against the probabilistic PKE by $\mathcal{A}_\$$ and its advantage by $\epsilon_\$$
 
-    % Game 1 can be simulated by OW-CPA adversary
+    Game 0 is the standard OW-CPA game: $\epsilon_0 = \epsilon_{\strike{\$}}$
+
+    In game 1, the challenge encryption routine is modified. Instead of using a pseudorandom coin, a truly random coin is used.
+
+    \begin{multicols}{2}
+        \begin{algorithm}[H]
+            \caption{Game 0}
+            \SetAlgoLined
+            $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+            $m^\ast \leftsample \mathcal{M}$\;
+            $r^\ast \leftarrow H(m^\ast)$\;
+            $c^\ast \leftarrow E(\pk, m^\ast; r^\ast)$\;
+            $\hat{m} \leftsample \mathcal{A}_\text{OW-CPA}(1^\lambda, \pk, c^\ast)$\;
+            \Return{
+                $\llbrack \hat{m} = m^\ast\rrbrack$
+            }
+        \end{algorithm}
+
+        \begin{algorithm}[H]
+            \caption{Game 1}
+            \SetAlgoLined
+            $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+            $m^\ast \leftsample \mathcal{M}$\;
+            $r^\ast \leftsample \coin$\;
+            $c^\ast \leftarrow E(\pk, m^\ast; r^\ast)$\;
+            $\hat{m} \leftsample \mathcal{A}_\text{OW-CPA}(1^\lambda, \pk, c^\ast)$\;
+            \Return{
+                $\llbrack \hat{m} = m^\ast\rrbrack$
+            }
+        \end{algorithm}
+    \end{multicols}
+
+    Under the random oracle assumption, game 0 and game 1 are indistinguishable from the perspective of $\mathcal{A}_{\strike{\$}}$ except for when $\mathcal{A}_{\strike{\$}}$ makes a hash query $\tilde{m} = m^\ast$.
+Denote the event that $\mathcal{A}_{\strike{\$}}$ makes such a query by $\operatorname{QUERY}^\ast$, then by the difference lemma:
+
+    \begin{equation*}
+        \epsilon_0 - \epsilon_1 \leq P[\operatorname{QUERY}^\ast]
+    \end{equation*}
+
+    We can bound $P[\operatorname{QUERY}^\ast]$ by constructing a OW-CPA adversary against the probabilistic scheme that uses $\mathcal{A}_{\strike{\$}}$ as a sub-routine and simulates game 1 for $\mathcal{A}_{\strike{\$}}$.
+When $\mathcal{A}_\$$ receives the probabilistic challenge ciphertext $c^\ast$, it directly passes it to the sub-routine $\mathcal{A}_{\strike{\$}}$.
+After the sub-routine halts, $\mathcal{A}_\$$ samples from the hash oracle tape a random message as its output.
+If the sub-routine $\mathcal{A}_{\strike{\$}}$ indeed makes the correct query $m^\ast$, then the probability that a randomly chosen query value being the correct value is $\frac{1}{q_H}$:
+
+    \begin{equation*}
+        \epsilon_\$ = P[\operatorname{QUERY}^\ast] \cdot \frac{1}{q_H}
+    \end{equation*}
+
+    Game 1 can be simulated by an OW-CPA adversary against the probabilistic PKE:
+
+    \begin{equation*}
+        \epsilon_1 = \epsilon_\$
+    \end{equation*}
+
+    Putting the three equations above together gives us the desired inequality.
 \end{proof}

### 2024-05-28 lattices/compact-fo-using-mac/security-definitions/main.tex
Clarification on security definitions

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{multicol}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
\newcommand{\keygen}{\operatorname{KeyGen}}
\newcommand{\pk}{\operatorname{pk}}
\newcommand{\sk}{\operatorname{sk}}
\newcommand{\pco}{\operatorname{PCO}}
\newcommand{\cvo}{\operatorname{CVO}}
\newcommand{\coin}{\operatorname{COIN}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{Some clarification on security definitions}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Spring, 2024}

\begin{document}
%%%% TITLE %%%%%
\maketitle

The confusion about what various security definition means seems to be caused by some inconsistencies between the popular textbooks and papers. I consulted two textbooks and found that indeed their security definitions are meaningfully different from what I am familiar with.

\section{Textbook definitions}
``A graduate course in applied cryptography''\cite{boneh2020graduate} introduced the concept of \textbf{CPA security} in section 5.3 (page 181):

\begin{definition}[CPA security from Boneh and Shoup]
    For a given cipher $\mathcal{E} = (E, D)$ defined over $(\mathcal{K}, \mathcal{M}, \mathcal{C})$ and for a given adversary $\mathcal{A}$, we define experiments $b$ for $b = 0, 1$:

    \begin{enumerate}
        \item The challenge selects $k \leftsample \mathcal{K}$
        \item The adversary submits a sequence of queries to the challenge. For $i = 1, 2, \ldots$, the $i$-th query is a pair of messages $m_{i, 0}, m_{i, 1}$ of the same length. The challenger computes $c_i \leftarrow E(k, m_{i, b})$ and return $c_i$ to the adversary
        \item The adversary outputs a bit $\hat{b} \in \{0, 1\}$
    \end{enumerate}

    Let $W_b$ denote the event that $\mathcal{A}$ outputs 1 in experiment $b$. We define $\mathcal{A}$'s advantage with respect to $\mathcal{E}$ to be:

    \begin{equation*}
        \operatorname{CPAadv}[A, \mathcal{E}] = \norm{
            P[W_0] - P[W_1]
        }
    \end{equation*}

    A cipher $\mathcal{E}$ is called \textbf{semantically secure against chosen plaintext attack}, or simply \textbf{CPA secure} if for all efficient adversaries, $\operatorname{CPAadv}$ is negligible.
\end{definition}

``Introduction to modern cryptography''\cite{katz2007introduction} also introduced the concept of CPA security in the context of an adversarial game:

\begin{definition}[CPA security from Katz and Lindell]
    We first define an experiment for any encryption scheme, any adversary, and any value $\lambda$ for the security parameter:

    \begin{enumerate}
        \item A random key is generated
        \item The adversary is given oracle access to the encryption routine and outputs a pair of messages of the same length
        \item A random bit is chosen and a ciphertext computed and given to the adversary
        \item The adversary outputs a bit
        \item The adversary wins if the output bit is equal to the random bit
    \end{enumerate}

    An encryption scheme has \textbf{indistinguishable encryptions under a chosen-plaintext attack}, or is \textbf{CPA secure}, if for all PPT adversaries there exists a negligible function $\operatorname{negl}$ such that

    \begin{equation*}
        P[\hat{b} = b^\ast] \leq \frac{1}{2} + \operatorname{negl}(\lambda)
    \end{equation*}
\end{definition}

\section{Security definition in research paper}
In ``A modular analysis of the Fujisaki-Okamoto transformation''\cite{hofheinz2017modular} by Hofheinz et al, the security definitions are as follows:

\begin{definition}[OW-ATK]
    Let $\operatorname{PKE} = (\operatorname{Gen}, \operatorname{Enc}, \operatorname{Dec})$ be a public-key encryption scheme with message space $\mathcal{M}$. For $\operatorname{ATK} \in \{\operatorname{CPA}, \operatorname{PCA}, \operatorname{VA}, \operatorname{PCVA}\}$ we define OW-ATK game, where

    \begin{equation*}
        \mathcal{O}_{\operatorname{ATK}} = \begin{cases}
            - & \operatorname{ATK} = \operatorname{CPA} \\
            \operatorname{PCO} & \operatorname{ATK} = \operatorname{CPA} \\
            \operatorname{CVO} & \operatorname{ATK} = \operatorname{VA} \\
            \operatorname{PCO}, \operatorname{CVO} & \operatorname{ATK} = \operatorname{PCVA}
        \end{cases}
    \end{equation*}

    \begin{multicols}{2}
        \begin{algorithm}[H]
            \caption{$\operatorname{OW-ATK}$ game}
            $(\operatorname{pk}, \operatorname{sk}) \leftarrow \operatorname{Gen}()$\;
            $m^\ast \leftsample \mathcal{M}$\;
            $c^\ast \leftarrow E(\operatorname{pk}, m^\ast)$\;
            $\hat{m} \leftarrow \mathcal{A}^{\mathcal{O}_{\operatorname{ATK}}}(\operatorname{pk}, c^\ast)$\;
            \Return{$\llbrack \hat{m} = m^\ast \rrbrack$}
        \end{algorithm}

        \begin{algorithm}[H]
            \caption{$\pco(m \in \mathcal{M}, c)$}
            \Return{$\llbrack D(\sk, c) = m \rrbrack$}
        \end{algorithm}

        \begin{algorithm}[H]
            \caption{$\cvo(c \neq c^\ast)$}
            \Return{$\llbrack D(\sk, c) \in \mathcal{M} \rrbrack$}
        \end{algorithm}
    \end{multicols}
\end{definition}

\section{Conclusion}
In the textbooks, ``\textbf{CPA secure}'' really means ``\textbf{IND-CPA} secure''. On the other hand, in research paper, the security definition is always explicitly spelled with both the goal (to break one-wayness or to break indistinguishability) and the adversary's capabilities (access to some specified set of oracles). My guess is that we got confused last week because in public-key cryptography, CPA is a rather meaningless notion because the adversary has the public key, so ``one-wayness'' automatically implies ``one-way security under CPA''.

For an example, here is textbook RSA:

\begin{multicols}{2}
    \begin{algorithm}[H]
        \caption{RSA KeyGen}
        $p, q \leftsample \operatorname{PrimeGen}()$\;
        $N \leftarrow p \cdot q$\;
        $\phi \leftarrow (p-1)\cdot(q-1)$\;
        $e \leftarrow 3$\;
        $d \leftarrow e^{-1} \mod \phi$\;
        \Return{
            $\pk = (N, e)$, $\sk = d$
        }
    \end{algorithm}

    \begin{algorithm}[H]
        \caption{Encryption $E(\pk, m)$}
        \Return{
            $m^e \mod N$
        }
    \end{algorithm}

    \begin{algorithm}[H]
        \caption{Decryption $D(\sk, c)$}
        \Return{
            $c^d \mod N$
        }
    \end{algorithm}
\end{multicols}

From the textbooks, \textbf{textbook RSA achieves one-wayness but is not CPA secure}, because its encryption is deterministic. On the other hand, using explicit game definitions, we say that RSA is OW-CPA secure but not IND-CPA secure.

OW-CPA and IND-CPA security are commonly accepted standard security notions in. Both Hofheinz\cite{hofheinz2017modular} and the Kyber team\cite{bos2018crystals} make use of OW-CPA and IND-CPA in their papers. In any case, since I will introduce non-standard security notions, I will explicitly spell out the security games, including the adversary's goal and the adversary's capabilities, so there should be no confusion about the meaning of any terms.

\bibliographystyle{plain}
\bibliography{./references.bib}

\end{document}

### 2024-05-29 lattices/compact-fo-using-mac/manuscript/main.tex
Rough draft

@@
 \newcommand{\coin}{\operatorname{COIN}}
+\newcommand{\encap}{\operatorname{Encap}}
+\newcommand{\decap}{\operatorname{Decap}}
 \newcommand{\llbrack}{[\![}
@@
 
-\title{Mermory-efficient Fujisaki-Okamoto transformation}
+\title{Mermory-efficient FO transform using ``encrypt-then-MAC''}
 \author{Ganyu (Bruce) Xu (g66xu)}
@@
 
+\section{Preliminaries}
+\subsection{Security definitions}
+The security of public-key encryption schemes (PKE) and message authentication code (MAC) is defined using an number of adversarial games, where each game is defined by the adversary's goal and objectives.
+
+For PKE, we define the $\operatorname{OW-ATK}$ and $\operatorname{IND-ATK}$ security game as follows:
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \caption{OW-ATK security game}
+        \KwIn{
+            Security parameter $\lambda$, adversary $\mathcal{A}$
+        }
+        \SetAlgoLined
+        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+        $m^\ast \leftsample \mathcal{M}$\;
+        $c^\ast \leftsample E(\pk, m^\ast)$\;
+        $\hat{m} \leftarrow \mathcal{A}(1^\lambda, \pk, c^\ast, \mathcal{O}_{\operatorname{ATK}})$\;
+        \Return{
+            $\llbrack m = \hat{m} \rrbrack$
+        }
+    \end{algorithm}
+
+    \begin{algorithm}[H]
+        \caption{IND-ATK security game}
+        \KwIn{
+            Security param $\lambda$, adversary $\mathcal{A}$
+        }
+        \SetAlgoLined
+        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+        $(m_0, m_1) \leftsample \mathcal{A}(1^\lambda, \pk, \mathcal{O}_{\operatorname{ATK}})$\;
+        $b \leftsample \{0, 1\}$\;
+        $c^\ast \leftarrow E(\pk, m_b)$\;
+        $\hat{b} \leftarrow \mathcal{A}(1^\lambda, \pk, c^\ast, \mathcal{O}_{\operatorname{ATK}})$\;
+        \Return{
+            $\llbrack \hat{b} = b \rrbrack$
+        }
+    \end{algorithm}
+\end{multicols}
+
+Where $\mathcal{O}_{\operatorname{ATK}}$ depends on the choice of $\operatorname{ATK}$:
+
+\begin{equation*}
+    \mathcal{O}_{\operatorname{ATK}} =\begin{cases}
+        - & \operatorname{ATK} = \operatorname{CPA} \\
+        \pco & \operatorname{ATK} = \operatorname{PCA} \\
+        \cvo & \operatorname{ATK} = \operatorname{VA} \\
+        (\pco, \cvo) & \operatorname{ATK} = \operatorname{PCVA} \\
+    \end{cases}
+\end{equation*}
+
+From a high level, $\pco(m, c)$ returns 1 if and only if $m$ is a valid decryption of $c$, and $\cvo(c)$ returns 1 if and only if $c$ is a valid ciphertext.
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \caption{$\pco(m, c)$}
+        \Return{
+            $\llbrack D(\sk, c) = m \rrbrack$
+        }
+    \end{algorithm}
+
+    \begin{algorithm}[H]
+        \caption{$\cvo(c)$}
+        \Return{
+            $\llbrack D(\sk, c) \in \mathcal{M} \rrbrack$
+        }
+    \end{algorithm}
+\end{multicols}
+
+\subsection{$\delta$-correctness}
+We define the correctness of a PKE using an adversarial game in which an adversary tries to produce plaintext that triggers decryption failures.
+
+\begin{algorithm}[H]
+    \caption{Correctness game $\operatorname{CORS}$}
+    \SetAlgoLined
+    $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+    $m \leftarrow \mathcal{A}(1^\lambda, \pk)$\;
+    \Return{
+        $\llbrack D(\sk, E(\pk, m)) \neq m \rrbrack$\;
+    }
+\end{algorithm}
+
+\begin{definition}
+    A public-key encryption scheme $\mathcal{E}=(E, D)$ is $\delta$-correct if for any efficient adversary $\mathcal{A}$, the probability of winning the correctness game is bounded by $\delta$:
+
+    \begin{equation*}
+        P[\operatorname{CORS}(\mathcal{A}) \rightarrow 1] \leq \delta
+    \end{equation*}
+\end{definition}
+
+\subsection{The IND-CCA KEM transformation}
+A key encapsulation mechanism (KEM) is defined by three routines $(\keygen, \encap, \decap)$, where $\keygen$ outputs a keypair $(\pk, \sk)$, $\encap$ takes the public key and outputs a pair of ciphertext $c$ and shared secret $K$, and $\decap$ takes the secret key $\sk$ and ciphertext $c$ and outputs the shared secret $K$.
+
+The security of a KEM is defined by an adversarial game.
+In the IND-CCA game, the adversary has access to a decapsulation oracle $\mathcal{O}^{\decap}$ and tries to distinguish whether the challenge shared secret is pseudorandom or truly random:
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]
+        \caption{IND-CCA KEM game}
+        $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+        $(c^\ast, K_0) \leftarrow \encap(\pk)$\;
+        $K_1 \leftsample \mathcal{K}$\;
+        $b \leftsample \{0, 1\}$\;
+        $\hat{b} \leftarrow \mathcal{A}^{\mathcal{O}^{\decap}}(1^\lambda, \pk, c^\ast, K_b)$\;
+        \Return{
+            $\llbrack \hat{b} = b \rrbrack$
+        }
+    \end{algorithm}
+
+    \begin{algorithm}[H]
+        \caption{Decapsulation oracle $O^{\decap}$}
+        \KwIn{$c \in \mathcal{C}$}
+        \Return{
+            $\decap(\sk, c)$
+        }
+    \end{algorithm}
+\end{multicols}
+
+The advantage of an adversary in the IND-CCA KEM game is defined by
+
+$$\epsilon = \norm{P[\operatorname{IND-CCA-KEM}(\mathcal{A}) = 1] - \frac{1}{2}}$$
+
+A KEM is IND-CCA secure if for any efficient adversary, the advantage is negligible with respect to $\lambda$.
+
+A key result from Hofheinz, Hovelmann, and Kiltz\cite{hofheinz2017modular} is a generic transformation from a PKE to a KEM.
+If the PKE is OW-PCVA secure, then the KEM is IND-CCA secure with tight security bounds.
+In this paper, we focus on the OW-PCVA transformation and use the transformation from OW-PCVA PKE to IND-CCA KEM as described by Hofheinz et al.
+
 \section{OW-PCVA transformation}\label{sec:ow-pcva-transformation}
+
+\subsection{From deterministic encryption scheme to OW-PCVA security}
 Let $(K, E, D)$ be a public-key encryption scheme defined over $(\mathcal{K}_\text{PKE}, \mathcal{M}_\text{PKE}, \mathcal{C})$.
@@
 
-\subsection{Deterministic OW-CPA transformation}
+\subsection{From probabilistic encryption to deterministic encryption}
 In the OW-PCVA transformation, we required the input encryption scheme to be deterministic.
@@
 
+        \columnbreak
+
         \begin{algorithm}[H]
@@
 
+\bibliographystyle{plain}
+\bibliography{./references.bib}
+
 \end{document}

### 2024-05-30 lattices/compact-fo-using-mac/manuscript/main.tex
Checkpoint for Friday meeting

@@
 \newcommand{\decap}{\operatorname{Decap}}
+\newcommand{\pke}{\operatorname{PKE}}
+\newcommand{\mac}{\operatorname{MAC}}
+\newcommand{\cors}{\operatorname{CORS}}
 \newcommand{\llbrack}{[\![}
@@
 \section{Preliminaries}
-\subsection{Security definitions}
-The security of public-key encryption schemes (PKE) and message authentication code (MAC) is defined using an number of adversarial games, where each game is defined by the adversary's goal and objectives.
+\subsection{Public-key encryption}
+A public key encryption scheme $\pke$ consists of three algorithms $(\keygen, E, D)$ defined over the collection of key space, message space, and ciphertext space $(\mathcal{K}, \mathcal{E}, \mathcal{D})$.
+The $\keygen$ routines takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk)$.
+The encryption routine $E$ takes a public key $\pk$ and a plaintext message $m \in \mathcal{M}$, and outputs a ciphertext $c \in \mathcal{C}$.
+Where the encryption routine is probabilistic, we denote the source of randomness by a coin sampled from the coin space $r \in \mathcal{R}$ such that $c \leftsample E(\pk, m; r)$.
+When $m, r$ is both specified, $E$ must be deterministic and injective.
+The decryption routine $D(\sk, c)$ takes a secret key and a ciphertext, and outputs the corresponding plaintext.
+Decryption routine is always deterministic.
+Where appropriate, the decryption routine could also output the rejection symbol $\bot$ to indicate that the ciphertext is invalid.
 
-For PKE, we define the $\operatorname{OW-ATK}$ and $\operatorname{IND-ATK}$ security game as follows:
+Correctness is commonly required of all $\pke$'s, which means that for all possible key pairs $(\pk, \sk)$ and messages $m \in \mathcal{M}$, $D(\sk, E(\sk, m)) = m$.
+However, with lattice-based schemes (such as CRYSTALS-Kyber), for randomly sampled keypairs and messages, there is a non-zero probability that $D(\sk, E(\pk, m)) \neq m$.
+Furthermore, decryption failures could be an attack vector through which an adversary learns partial information about the keys or ciphertext\cite{bos2018crystals}.
+To account for this possibility, we define a correctness game, then model the probability of decryption failure on this game.
+
+In the correctness game $\cors$, an adversary is given the security parameter and some public key and outputs a plaintext message and a coin.
+The adversary wins the $\cors$ game if and only if $m$ triggers a decryption failure for the specified keypair.
+Note that this game can be trivially adapted for deterministic encryption scheme, so we will skip the discussion here.
+
+\begin{algorithm}[H]
+    \caption{The correctness game $\cors$}
+    \KwIn{
+        The correctness adversary $\mathcal{A}$
+    }
+    $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
+    $(m, r) \leftsample \mathcal{A}(1^\lambda, \pk)$\;
+    \Return{
+        $\llbrack D(\sk, E(\pk, m; r)) \neq m \rrbrack$
+    }
+\end{algorithm}
+
+\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
+    A public-key encryption scheme $\pke$ is $\delta$-correct if no (possibly unbounded) adversary can win the $\cors$ game with probability greater than $\delta$.
+\end{definition}
+
+For some concrete instantiation, the value of $\delta$ for Kyber\cite{avanzi2019crystals} is shown in table \ref{table:kyber-delta-vals}:
+
+\begin{table}
+    \centering
+    \begin{tabular}{cc}
+        Security level & $\delta$ \\
+        Kyber512 & $2^{-139}$ \\
+        Kyber768 & $2^{-164}$ \\
+        Kyber1024 & $2^{-174}$ \\
+    \end{tabular}
+    \caption{Concrete $\delta$ value for CRYSTAL-Kyber}
+    \label{table:kyber-delta-vals}
+\end{table}
+
+The security of a $\pke$ is defined using an number of adversarial games, where each game is defined by the adversary's goal and objectives.
+We define the $\operatorname{OW-ATK}$ and $\operatorname{IND-ATK}$ security game in algorithm \ref{alg:ow-atk-game} and algorithm \ref{alg:ind-atk-game}:
 
 \begin{multicols}{2}
-    \begin{algorithm}[H]
+    \begin{algorithm}[H]\label{alg:ow-atk-game}
         \caption{OW-ATK security game}
@@
 
-    \begin{algorithm}[H]
+    \begin{algorithm}[H]\label{alg:ind-atk-game}
         \caption{IND-ATK security game}
@@
 
-\subsection{$\delta$-correctness}
-We define the correctness of a PKE using an adversarial game in which an adversary tries to produce plaintext that triggers decryption failures.
+\subsection{Message authentication code}
+A message authentication code $\mac$ consists of a pair of routines $(S, V)$ defined over a collection $(\mathcal{K}, \mathcal{M}, \mathcal{T})$ of key space, message, space, and tag space.
+The signing routine $S(k, m)$ takes the secret key $k$ and a message $m$, and outputs a tag $t$.
+The verfiying routine $V(k, m, t)$ takes the triplet of secret key $k$, message $m$, and tag $t$, and outputs 1 if the message-tag pair is valid for the given secret key or 0 otherwise.
 
-\begin{algorithm}[H]
-    \caption{Correctness game $\operatorname{CORS}$}
-    \SetAlgoLined
-    $(\pk, \sk) \leftsample \keygen(1^\lambda)$\;
-    $m \leftarrow \mathcal{A}(1^\lambda, \pk)$\;
+One common notion of security for $\mac$ is ``existential forgery under chosen message attack'' (EF-CMA), which is defined using an adversarial game.
+In the EF-CMA game, an adversary is given access to a signing oracle $\mathcal{O}^S$ and tries to forge valid message-tag pairs that have not be recorded in the signing oracle before.
+
+\begin{multicols}{2}
+    \begin{algorithm}[H]\label{alg:euf-cma-game}
+        \caption{
+            EF-CMA game
+        }
+        $k^\ast \leftsample \mathcal{K}$\;
+        $(m, t) \leftarrow \mathcal{A}^{\mathcal{O}^S}()$\;
         \Return{
-        $\llbrack D(\sk, E(\pk, m)) \neq m \rrbrack$\;
+            $\llbrack (m, t) \not\in \mathcal{O}^{S} \rrbrack$
+            and $\llbrack V(k^\ast, m, t) \rrbrack$
         }
@@
 
-\begin{definition}
-    A public-key encryption scheme $\mathcal{E}=(E, D)$ is $\delta$-correct if for any efficient adversary $\mathcal{A}$, the probability of winning the correctness game is bounded by $\delta$:
+    \begin{algorithm}[H]\label{alg:mac-signing-oracle}
+        \caption{The signing oracle $\mathcal{O}^S$}
+        \KwIn{
+            $m \in \mathcal{M}$
+        }
+        $t \leftarrow S(k^\ast, m)$\;
+        \If{$(m, t) \not\in \mathcal{O}^S$}{
+            Append $(m, t)$ to $\mathcal{O}^S$
+        }
+        \Return{$t$}
+    \end{algorithm}
+\end{multicols}
 
-    \begin{equation*}
-        P[\operatorname{CORS}(\mathcal{A}) \rightarrow 1] \leq \delta
-    \end{equation*}
+\begin{definition}[Existential unforgeability]\label{def:euf-cma}
+    A $\mac$ is existentially unforgeable under chosen message attack if for all efficient adversary, the probability of winning the EF-CMA game is negligible with respect to the security parameter.
 \end{definition}

### 2024-06-01 computational-number-theory/src/lib.rs
Simple arithmetics

@@ -1,0 +1,154 @@
+//! My implementation of computational number theory
+//! Follows Henri Cohen's [A Course in Computational Algebraic Number Theory]
+
+/// 32-bit architecture
+pub type Word = u32;
+/// Twice the size of Word, useful for widening multiplication
+pub type LongWord = u64;
+
+/// Equivalent to carrying_add but I have to implement it because carrying_add is not stable
+const fn overflowing_adc(a: Word, b: Word, carry: bool) -> (Word, bool) {
+    // There are two overflowing addition to perform
+    let (mut sum, mut carry_next) = a.overflowing_add(b);
+    if carry {
+        let (carried_sum, carry_2) = sum.overflowing_add(1);
+        carry_next = carry_next || carry_2;
+        sum = carried_sum;
+    }
+
+    return (sum, carry_next);
+}
+
+/// Equivalent to borrowing_sub but borrowing_sub is unstable
+const fn overflowing_sbb(a: Word, b: Word, borrow: bool) -> (Word, bool) {
+    let (mut diff, mut borrow_next) = a.overflowing_sub(b);
+    if borrow {
+        let (borrowed_diff, borrow_2) = diff.overflowing_sub(1);
+        borrow_next = borrow_next || borrow_2;
+        diff = borrowed_diff;
+    }
+
+    return (diff, borrow_next);
+}
+
+fn widening_mul(a: Word, b: Word) -> (Word, Word) {
+    let a: LongWord = a.into();
+    let b: LongWord = b.into();
+    let prod: LongWord = a.wrapping_mul(b);
+    let low: Word = prod as Word;
+    let high: Word = (prod >> 32) as Word;
+
+    return (high, low);
+}
+
+/// A big integer is an array of words
+/// We will tentatively use little-endian: lower index encodes less significant digits
+/// num = sum_{i=0}^{L-1}(Uint[i] * (Word ** i))
+#[derive(Debug, Copy, Clone, PartialEq)]
+pub struct Uint<const L: usize>([Word; L]);
+
+impl<const L: usize> Uint<L> {
+    pub const ZERO: Self = Self([0; L]);
+    pub const MAX: Self = Self([Word::MAX; L]);
+
+    /// There is no easy way to specify a "1" using just the array notation, so we use a const fn
+    pub const fn one() -> Self {
+        let mut arr = [0; L];
+        arr[0] = 1;
+        return Self(arr);
+    }
+
[…]

### 2024-06-01 computational-number-theory/src/lib.rs
Faulty implementation of Uint::widening_mul

@@ -97,9 +97,28 @@ impl<const L: usize> Uint<L> {
         let mut self_loc = 0;
 
         while self_loc < L {
+            let self_val = self.0[self_loc];
+
             let mut other_loc = 0;
             while other_loc < L {
-                todo!("need Word::widening_mul");
+                let other_val = other.0[other_loc];
+                let (tmp_high, tmp_low) = widening_mul(self_val, other_val);
+                let tmp_high_loc = self_loc + other_loc + 1;
+                let tmp_low_loc = self_loc + other_loc;
+
+                if tmp_high_loc >= L {
+                    // TODO: this will cause overflow in many cases
+                    // We should have a dedicated "add word and pow to big integer"
+                    high[tmp_high_loc - L] += tmp_high;
+                } else {
+                    low[tmp_high_loc] += tmp_high;
+                }
+
+                if tmp_low_loc >= L {
+                    high[tmp_low_loc - L] += tmp_low;
+                } else {
+                    low[tmp_low_loc] += tmp_low;
+                }
 
                 other_loc += 1;
             }
@@ -150,5 +169,19 @@ mod tests {
         let expected_prod = (Word::MAX as LongWord) * (Word::MAX as LongWord);
         assert_eq!(expected_prod as Word, low);
         assert_eq!((expected_prod >> 32) as Word, high);
+    }
+
+    #[test]
+    fn u256_widening_mul() {
+        let (high, low) = U256::ZERO.widening_mul(&U256::MAX);
+        assert_eq!(high, U256::ZERO);
+        assert_eq!(low, U256::ZERO);
+
+        let (high, low) = U256::one().widening_mul(&U256::one());
+        assert_eq!(high, U256::ZERO);
+        assert_eq!(low, U256::one());
+
+        // TODO: this doesn't work
+        // let (high, low) = U256::MAX.widening_mul(&U256::MAX);
     }
 }

### 2024-06-09 computational-number-theory/src/lib.rs
partial ordering

@@ -1,7 +1,7 @@
 //! My implementation of computational number theory
 //! Follows Henri Cohen's [A Course in Computational Algebraic Number Theory]
 
-use std::{fmt::UpperHex, num::ParseIntError};
+use std::{cmp::Ordering, fmt::UpperHex, num::ParseIntError};
 
 /// 32-bit architecture
 pub type Word = u32;
@@ -46,7 +46,7 @@ fn widening_mul(a: Word, b: Word) -> (Word, Word) {
 /// A big integer is an array of words
 /// We will tentatively use little-endian: lower index encodes less significant digits
 /// num = sum_{i=0}^{L-1}(Uint[i] * (Word ** i))
-#[derive(Debug, Copy, Clone, PartialEq)]
+#[derive(Debug, Copy, Clone, PartialEq, Eq)]
 pub struct Uint<const L: usize>([Word; L]);
 
 impl<const L: usize> Uint<L> {
@@ -194,7 +194,6 @@ impl<const L: usize> Uint<L> {
         return (shifted, shift >= (Self::BITS as usize));
     }
 
-    #[allow(unused_variables, unused_mut, unreachable_code)]
     pub fn overflowing_shr(&self, shift: usize) -> (Self, bool) {
         let mut shifted = Self::ZERO;
 
@@ -219,6 +218,26 @@ impl<const L: usize> Uint<L> {
 
         return (shifted, shift >= (Self::BITS as usize));
     }
+
+    /// Comparison, but running through all digits
+    pub fn cmp(&self, other: &Self) -> Ordering {
+        let mut i = L - 1;
+        let mut ordering = Ordering::Equal;
+        loop {
+            let local_cmp = self.0[i].cmp(&other.0[i]);
+            if ordering == Ordering::Equal {
+                ordering = local_cmp;
+            }
+
+            if i == 0 {
+                break;
+            } else {
+                i -= 1;
+            }
+        }
+
+        return ordering;
+    }
 }
 
 pub type U256 = Uint<8>;
@@ -229,6 +248,12 @@ impl<const L: usize> UpperHex for Uint<L> {
             write!(f, "{:08X}", self.0[L - i - 1])?;
         }
         return Ok(());
+    }
+}
+
+impl<const L: usize> PartialOrd for Uint<L> {
+    fn partial_cmp(&self, other: &Self) -> Option<Ordering> {
[…]

### 2024-06-09 computational-number-theory/src/lib.rs
Simplified implementation of cmp, implemented Uint::leading_zeros

@@ -43,6 +43,26 @@ fn widening_mul(a: Word, b: Word) -> (Word, Word) {
     return (high, low);
 }
 
+/// Count the number of leading zero bits
+/// TODO: there is probably more efficient implementation but oh well
+const fn leading_zeros(word: Word) -> u32 {
+    let mut cnt = 0;
+    let mask = 1 << 31;
+    let mut shift = 0;
+
+    while shift < 32 {
+        if word & (mask >> shift) == 0 {
+            cnt += 1;
+        } else {
+            return cnt;
+        }
+
+        shift += 1;
+    }
+
+    return cnt;
+}
+
 /// A big integer is an array of words
 /// We will tentatively use little-endian: lower index encodes less significant digits
 /// num = sum_{i=0}^{L-1}(Uint[i] * (Word ** i))
@@ -224,9 +244,8 @@ impl<const L: usize> Uint<L> {
         let mut i = L - 1;
         let mut ordering = Ordering::Equal;
         loop {
-            let local_cmp = self.0[i].cmp(&other.0[i]);
             if ordering == Ordering::Equal {
-                ordering = local_cmp;
+                ordering = self.0[i].cmp(&other.0[i]);
             }
 
             if i == 0 {
@@ -237,6 +256,24 @@ impl<const L: usize> Uint<L> {
         }
 
         return ordering;
+    }
+
+    /// The number of leading zero bits
+    pub const fn leading_zeros(&self) -> u32 {
+        let mut cnt = 0;
+        let mut i = 0;
+
+        while i < L {
+            let word = self.0[L - i - 1];
+            let word_leading_zeros = leading_zeros(word);
+            cnt += word_leading_zeros;
+            if word != 0 {
+                return cnt;
+            }
+            i += 1;
+        }
+
+        return cnt;
     }
 }
 
@@ -403,5 +440,20 @@ mod tests {
 
         assert_eq!(U256::ONE.cmp(&U256::ZERO), Ordering::Greater);
         assert!(U256::ZERO < U256::ONE);
+    }
+
+    #[test]
+    fn word_leading_zeros() {
+        for cnt in 0..32 {
+            let (shifted, _) = Word::MAX.overflowing_shr(cnt);
+            assert_eq!(leading_zeros(shifted), cnt);
+        }
+    }
+
+    #[test]
+    fn uint_leading_zeros() {
+        assert_eq!(U256::MAX.leading_zeros(), 0);
[…]

### 2024-06-09 computational-number-theory/src/main.rs

@@ -4,7 +4,7 @@ fn main() {
     let mut val = U256::ONE;
     for _ in 0..40 {
         println!("{:?}", val);
-        let (next_val, overflowing) = val.overflowing_add(&val);
+        let (next_val, _) = val.overflowing_add(&val);
         val = next_val;
     }
 }

### 2024-06-09 computational-number-theory/src/lib.rs
Implemented Euclidean division with remainder

@@ -275,6 +275,34 @@ impl<const L: usize> Uint<L> {
 
         return cnt;
     }
+
+    /// Euclidean division (self / other), returning the quotient and the remainder such that
+    /// `self = other * quotient + remainder`
+    /// There is no constant-time guarantee of any sort
+    /// It is probably also very inefficient
+    pub fn div_rem_vartime(&self, other: &Self) -> (Self, Self) {
+        let mut quo = Self::ZERO;
+        let mut rem = self.clone();
+        // TODO: this is the source of timing variability
+        let leading_zeros: usize = other.leading_zeros() as usize;
+        let mut i = 0;
+
+        while i <= leading_zeros {
+            let other_shift = leading_zeros - i;
+            let (shifted, _) = other.overflowing_shl(other_shift);
+            let (multiplier, _) = Self::ONE.overflowing_shl(other_shift);
+            if rem >= shifted {
+                let (diff, _) = rem.overflowing_sub(&shifted);
+                let (sum, _) = quo.overflowing_add(&multiplier);
+                rem = diff;
+                quo = sum;
+            }
+
+            i += 1;
+        }
+
+        return (quo, rem);
+    }
 }
 
 pub type U256 = Uint<8>;
@@ -455,5 +483,18 @@ mod tests {
         assert_eq!(U256::MAX.leading_zeros(), 0);
         assert_eq!(U256::ONE.leading_zeros(), 255);
         assert_eq!(U256::ZERO.leading_zeros(), 256);
+    }
+
+    #[test]
+    fn uint_div_rem() {
+        let (quo, rem) = U256::MAX.div_rem_vartime(&U256::ONE);
+        assert_eq!(quo, U256::MAX);
+        assert_eq!(rem, U256::ZERO);
+
+        let (two, _) = U256::ONE.overflowing_shl(1);
+        let (quo, rem) = U256::MAX.div_rem_vartime(&two);
+        let (half, _) = U256::MAX.overflowing_shr(1);
+        assert_eq!(quo, half);
+        assert_eq!(rem, U256::ONE);
     }
 }

### 2024-06-09 computational-number-theory/src/main.rs

@@ -1,10 +1,6 @@
 use compnt::U256;
 
 fn main() {
-    let mut val = U256::ONE;
-    for _ in 0..40 {
-        println!("{:?}", val);
-        let (next_val, _) = val.overflowing_add(&val);
-        val = next_val;
-    }
+    let (quo, rem) = U256::MAX.div_rem_vartime(&U256::ONE);
+    println!("{:X}, {:X}", quo, rem);
 }

### 2024-06-11 lattices/compact-fo-using-mac/impl-considerations/main.tex
Implementation considerations for encrypt-then-mac

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{multicol}

% Common shortcuts
\newcommand{\round}[1]{\left\lfloor {#1} \right\rceil}
\newcommand{\abs}[1]{\left\lvert {#1} \right\rvert}
\newcommand{\norm}[1]{\left\lvert {#1} \right\rvert}
\newcommand{\Norm}[1]{\left\lVert {#1} \right\rVert}
\newcommand{\monospace}{\texttt}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
% \newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\pke}{\monospace{PKE}}
\newcommand{\keygen}{\monospace{KeyGen}}
\newcommand{\encrypt}{\monospace{E}}
\newcommand{\decrypt}{\monospace{D}}
\newcommand{\kem}{\monospace{KEM}}
\newcommand{\encap}{\monospace{Encap}}
\newcommand{\decap}{\monospace{Decap}}
\newcommand{\mac}{\monospace{MAC}}
\newcommand{\sign}{\monospace{S}}
\newcommand{\verify}{\monospace{V}}
\newcommand{\pk}{\monospace{pk}}
\newcommand{\sk}{\monospace{sk}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{Some implementation considerations}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{June, 2024}

\begin{document}
%%%% TITLE %%%%%
\maketitle

Here are some considerations when implementing Kyber using ``encrypt-then-mac''

\begin{enumerate}
    \item Strong candidates for $\mac$ are HMAC and KMAC. From an efficiency standpoint KMAC is preferable because $\monospace{HMAC}(k, m) = H(k \Vert H(k \Vert m))$ makes two calls to the hash function, but $\monospace{KMAC}$ makes only one call to $\monospace{Shake}$. From a security and usability standpoint, HMAC is more widely used than KMAC. We can simply try both and report the results.

    \item Since the $\mac$ key is derived from the message, and in Kyber, the message $m \in \mathcal{M} = \{0,1\}^{256}$ is exactly 256-bit in length, we should choose the $\mac$ key to also be exactly 256-bit in length. Any longer key is unnecessary, since it then because easier to simply search for the message.

    \item There is argument about how HMAC is not affected by weakness in collision resistance in the underlying hash function\cite{krawczyk1997hmac} (the idea is that an adversary can only obtain a tag online, which is much slower than computing hash offline). In addition, collision resistance is less affected by quantum computing than preimage resistance is. The best classical attack on collision resistance is the birthday attack, with the expected number of steps being $2^{\frac{n}{2}}$ for an $n$-bit hash; the best quantum attack\cite{brassard1997quantum} is expected to take $2^{\frac{n}{3}}$ steps for an $n$-bit hash.
\end{enumerate}

With the three points above, I came up with this table for the recommended parameters. For classical 128-bit security and 256-bit security, I chose a 256-bit and 384-bit hash for the tag because TLS 1.2 pairs SHA-256 with AES-128 and SHA-384 with AES-256. For quantum computing, we can choose more conservative parameters (such as using 384-bit hash for 128-bit security and 512-bit hash for 256-bit security).

\begin{table}[h]
    \begin{center}
        \begin{tabular}{c|c|c|c|c}
            & key size & tag size & pre-image resistance & collision resistance \\
            \hline
            classical, 128-bit security & 256 & 256 & 256 & 128 \\
            classical, 256-bit security & 256 & 384 & 384 & 192 \\
            quantum, optimistic & 256 & 256 & 128 & $\approx 86$\\
            quantum, moderate & 256 & 384 & 192 & 128\\
            quantum, conservative & 256 & 512 & 256 & $\approx 171$\\
        \end{tabular}
    \end{center}
    \caption{Parameter sizes and estimated security levels, all in bits}
\end{table}

\bibliographystyle{plain}
\bibliography{./references.bib}
\end{document}

### 2024-06-11 computational-number-theory/src/lib.rs
Getting started with integer modulus

@@ -305,8 +305,6 @@ impl<const L: usize> Uint<L> {
     }
 }
 
-pub type U256 = Uint<8>;
-
 impl<const L: usize> UpperHex for Uint<L> {
     fn fmt(&self, f: &mut std::fmt::Formatter<'_>) -> std::fmt::Result {
         for i in 0..L {
@@ -321,6 +319,10 @@ impl<const L: usize> PartialOrd for Uint<L> {
         Some(self.cmp(other))
     }
 }
+
+pub mod ring;
+
+pub type U256 = Uint<8>;
 
 #[cfg(test)]
 mod tests {

### 2024-06-11 computational-number-theory/src/ring.rs

@@ -1,0 +1,63 @@
+//! The integer modulus ring
+use std::error::Error;
+
+use crate::Uint;
+
+/// Various ways integer ring arithmetics can go wrong
+#[derive(Debug, Eq, PartialEq, Copy, Clone)]
+pub enum RingError {
+    /// The modulus of a ring cannot be zero
+    ModulusIsZero,
+}
+
+impl std::fmt::Display for RingError {
+    fn fmt(&self, f: &mut std::fmt::Formatter<'_>) -> std::fmt::Result {
+        write!(f, "{:?}", self)
+    }
+}
+
+impl Error for RingError {}
+
+/// The integer ring is defined by the modulus, which cannot be zero
+#[derive(Debug, Eq, PartialEq, Copy, Clone)]
+pub struct Ring<const L: usize> {
+    pub modulus: Uint<L>,
+}
+
+impl<const L: usize> Ring<L> {
+    pub fn new(modulus: Uint<L>) -> Result<Self, RingError> {
+        if modulus == Uint::<L>::ZERO {
+            return Err(RingError::ModulusIsZero);
+        }
+
+        Ok(Self { modulus })
+    }
+}
+
+/// An element of the modulus
+pub struct RingElem<const L: usize> {
+    pub val: Uint<L>,
+    pub ring: Ring<L>,
+}
+
+impl<const L: usize> RingElem<L> {
+    pub fn new(val: Uint<L>, ring: Ring<L>) -> Self {
+        todo!();
+    }
+
+    pub fn add(&self, other: &Self) -> Self {
+        todo!();
+    }
+
+    pub fn sub(&self, other: &Self) -> Self {
+        todo!();
+    }
+
+    pub fn mul(&self, other: &Self) -> Self {
+        todo!();
+    }
+
+    pub fn modexp(&self, exp: &Uint<L>) -> Self {
+        todo!();
+    }
+}

### 2024-06-12 lattices/compact-fo-using-mac/impl-considerations/main.tex
Updated MAC security recommendations

@@
 
+\section{June 12, 2024 considerations}
+Since ML-KEM as defined in FIPS 203 uses the SHA3 family of hash functions and XOFs, we will continue using SHA3 instead of introducing new families of hash functions so the code size can be reduced (and so the codebase can be more manageable).
+
+Using HMAC with SHA3 does not make much sense from an efficiency perspective, especially given that KMAC exists and is approved by NIST\cite{kelsey2016sha}.
+We can use the SHA2 family in \monospace{HMAC-SHA256} for 128-bit classical security and \monospace{HMAC-SHA384} for 256-bit classical security to modify the \textit{90s-variant} of Kyber, but I doubt how meaningful this modification can be, given that NIST gave up on it anyways.
+
+NIST SP.800.185\cite{kelsey2016sha} specified two approved variations of KMAC.
+\monospace{KMAC-128} uses \monospace{Shake128} as the underlying XOF, and \monospace{KMAC-256} uses \monospace{Shake256} as the underlying XOF.
+\monospace{Shake128} and \monospace{Shake256} respectively claim 128-bit and 256-bit (classical) collision resistance, though \textbf{citation is needed}.
+I also don't know about \textbf{quantum attacks on Shake}.
+If time permits I will look into security under quantum attacks.
+
+So here are the updated parameter recommendations
+
+\begin{table}[h]
+    \begin{center}
+        \begin{tabular}{c|c|c|c|c|c}
+            Security level 
+                & \monospace{MAC} 
+                & Key size
+                & Tag size
+                & Preimage resistance
+                & Collision resistance \\
+                \hline
+            \monospace{Kyber512}
+                & \monospace{KMAC-128}
+                & 256
+                & 256
+                & 256 (c), 128 (q)
+                & 128 (c), 85.3 (q) \\
+                \hline
+            \monospace{Kyber768}
+                & \monospace{KMAC-256}
+                & 256
+                & 384
+                & 384 (c), 192 (q)
+                & 192 (c), 128 (q) \\
+                \hline
+            \monospace{Kyber1024}
+                & \monospace{KMAC-256}
+                & 256
+                & 512
+                & 512 (c), 256(q)
+                & 256 (c), 170.6 (q) \\
+                \hline
+        \end{tabular}
+    \end{center}
+    \caption{MAC recommendations}
+\end{table}
+
+\section{June 11, 2024 consdierations}
 Here are some considerations when implementing Kyber using ``encrypt-then-mac''
@@
 
-\begin{table}[h]
+\begin{table}
     \begin{center}

### 2024-06-13 lattices/compact-fo-using-mac/rounds.py
Checkpoint

@@ -1,0 +1,26 @@
+"""Estimate the cost of various subroutines of Kyber, measured in the number of Keccak permutations
+and the number of NTT operations (including NTT and invert_NTT)
+"""
+from enum import Enum
+
+class SecurityLevels(Enum):
+    Kyber512 = 1
+    Kyber768 = 2
+    Kyber1024 = 3
+
+class KyberParameters:
+    KYBER_Q = 3329
+    KYBER_Q_BITS = 12
+    KYBER_N = 256
+    # Defaults to Kyber512 but will be modified
+    KYBER_K = 2
+    KYBER_ETA1 = 3
+    KYBER_ETA2 = 2
+
+    def __init__(self, level: SecurityLevels):
+        if level == SecurityLevels.Kyber512:
+            self.KYBER_K = 2
+            self.KYBER_ETA1 = 3
+            self.KYBER_ETA2 = 
+
+

### 2024-06-18 lattices/compact-fo-using-mac/impl-considerations/main.tex
Some consideration for choice of MAC

--- 2024-06-11 lattices/compact-fo-using-mac/impl-considerations/main.tex
@@
 % Common shortcuts
-\newcommand{\round}[1]{\left\lfloor {#1} \right\rceil}
-\newcommand{\abs}[1]{\left\lvert {#1} \right\rvert}
-\newcommand{\norm}[1]{\left\lvert {#1} \right\rvert}
-\newcommand{\Norm}[1]{\left\lVert {#1} \right\rVert}
 \newcommand{\monospace}{\texttt}
-\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
-% \newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}
-\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
 \newcommand{\pke}{\monospace{PKE}}
@@
 \newcommand{\decap}{\monospace{Decap}}
+\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
 \newcommand{\mac}{\monospace{MAC}}
-\newcommand{\sign}{\monospace{S}}
-\newcommand{\verify}{\monospace{V}}
+\newcommand{\sign}{\monospace{MAC}}
+\newcommand{\verify}{\monospace{MAC.Verify}}
 \newcommand{\pk}{\monospace{pk}}
 \newcommand{\sk}{\monospace{sk}}
+\newcommand{\pco}{\monospace{PCO}}
+\newcommand{\cvo}{\monospace{CVO}}
+\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
 \newcommand{\llbrack}{[\![}
 \newcommand{\rrbrack}{]\!]}
-\newlength{\wdth}
-\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}
+\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
 
@@
 \author{Ganyu (Bruce) Xu (g66xu)}
-\date{June, 2024}
+% \date{June, 2024}
 
@@
 
-Here are some considerations when implementing Kyber using ``encrypt-then-mac''
+\section{One-time MAC}
+The challenge encryption $c^\ast$ in the \monospace{OW-PCVA} game is computed using a uniformly randomly sampled plaintext message $m^\ast \leftsample \mathcal{M}_\pke$, and a $\mac$ key $k_\mac$ derived from said plaintext $m^\ast$.
+Assuming that $m^\ast$ is not already known (otherwise the \monospace{OW-CPA} security of the underlying $\pke$ is already broken), then $k^\ast_\mac$ is also not known, so the \monospace{OW-PCVA} adversary has no way to evaluate the $\mac$ under $k^\ast_\mac$ offline.
 
-\begin{enumerate}
-    \item Strong candidates for $\mac$ are HMAC and KMAC.
-From an efficiency standpoint KMAC is preferable because $\monospace{HMAC}(k, m) = H(k \Vert H(k \Vert m))$ makes two calls to the hash function, but $\monospace{KMAC}$ makes only one call to $\monospace{Shake}$.
-From a security and usability standpoint, HMAC is more widely used than KMAC.
-We can simply try both and report the results.
+Furthermore, I propose that the \monospace{OW-PCVA} adversary also has no way of issuing signing queries for $k^\ast_\mac$.
+The adversary can run $\encrypt$ or $\encap$ offline, but the tags generated from $\encrypt$ or $\encap$ will be associated with a chosen message or a random message, which has $\frac{1}{\norm{\mathcal{M}_\pke}}$ probability of being $m^\ast$.
+In other words, for the challenge $\mac$ key, the adversary will only see one valid message-tag pair, which is the challenge message $(c^\ast, t^\ast)$.
 
-    \item Since the $\mac$ key is derived from the message, and in Kyber, the message $m \in \mathcal{M} = \{0,1\}^{256}$ is exactly 256-bit in length, we should choose the $\mac$ key to also be exactly 256-bit in length.
-Any longer key is unnecessary, since it then because easier to simply search for the message.
+\section{Choice of MAC}
+When thinking about choosing a MAC there are two parameters that we need to care about: the key size and the tag size.
 
-    \item There is argument about how HMAC is not affected by weakness in collision resistance in the underlying hash function\cite{krawczyk1997hmac} (the idea is that an adversary can only obtain a tag online, which is much slower than computing hash offline).
-In addition, collision resistance is less affected by quantum computing than preimage resistance is.
-The best classical attack on collision resistance is the birthday attack, with the expected number of steps being $2^{\frac{n}{2}}$ for an $n$-bit hash; the best quantum attack\cite{brassard1997quantum} is expected to take $2^{\frac{n}{3}}$ steps for an $n$-bit hash.
-\end{enumerate}
+Whether working with Round 3 Kyber \cite{avanzi2019crystals} or ML-KEM \cite{key2023mechanism}, the message is always 256 bits in size, and since the MAC key is derived from the message, the MAC key should also always be 256 bits.
+A 256-bit MAC key is great since both GCM and Poly1305 accepts 256-bit MAC keys.
 
-With the three points above, I came up with this table for the recommended parameters.
-For classical 128-bit security and 256-bit security, I chose a 256-bit and 384-bit hash for the tag because TLS 1.2 pairs SHA-256 with AES-128 and SHA-384 with AES-256.
-For quantum computing, we can choose more conservative parameters (such as using 384-bit hash for 128-bit security and 512-bit hash for 256-bit security).
+Reasoning about the security implication of the tag size is more nuanced.
 
-\begin{table}[h]
-    \begin{center}
-        \begin{tabular}{c|c|c|c|c}
-            & key size & tag size & pre-image resistance & collision resistance \\
-            \hline
-            classical, 128-bit security & 256 & 256 & 256 & 128 \\
-            classical, 256-bit security & 256 & 384 & 384 & 192 \\
-            quantum, optimistic & 256 & 256 & 128 & $\approx 86$\\
-            quantum, moderate & 256 & 384 & 192 & 128\\
-            quantum, conservative & 256 & 512 & 256 & $\approx 171$\\
-        \end{tabular}
-    \end{center}
-    \caption{Parameter sizes and estimated security levels, all in bits}
-\end{table}
+\subsection{Generic collision resistance}
+We can model a MAC has a keyed hash function, which means that for a fixed MAC key, we can argue about the seurity of a MAC using the standard security notions for hash function: pre-image resistance, 2nd pre-image resistance, and collision resistance.
+
+However, there is an important distinction between the security of a MAC as a keyed hash function and the security of a hash function.
+With MAC as a keyed hash function, assuming that the adversary doesn't already have a way to evaluating this function offline (otherwise forgery is trivial), \textbf{it has to query the signing oracle to evaluate the hash function}.
+Given an n-bit tag, a generic birthday attack takes $2^{\frac{n}{2}}$ evaluations to find a collision, and the BHT algorithm \cite{brassard1997quantum} takes $2^{\frac{n}{3}}$ evaluations to find a collision, but neither attack is relevant, because "evaluation" means that the message-tag pair is queried from the signing oracle and \textbf{does not count as forgery}.
+
+This means that we only need to care about preimage resistance for our MAC (as a keyed hash function),
+
+\subsection{Offline versus online attacks}
+There should be a distinction between attacks that involve interaction with another party (online) versus attacks that can be done without such interaction (offline).
+When we measure the security of a scheme by the number of computational steps needed to carry out an attack, we specifically mean an offline attack.
+\textbf{An online attack with substantial number of interactions can be trivially thwarted} because the other party can implement rate limiting with exponential backoff, so after only a few (think fewer than 10 password tries) invalid interactions, the other party will effectively block the adversary from further interactions, thus stopping the attack in the track.
+
+\textbf{Adversary obtains $(c, t)$}, tampers with $c$ to obtain $c^\prime$ which still correspond with the same plaintext $m$, which means that $t^\prime$ should be created under the same key $(c, t)$ is created.
 

### 2024-06-18 lattices/compact-fo-using-mac/impl-considerations/main.tex
I want to argue that 128-bit tag is not a security concern

@@
 
+This is the literal definition of a one-time MAC.
+All many-time MAC are valid one-time MAC, but there are specific one-time MAC constructions that might be even more computationally efficient.
+\textbf{GMAC and Poly1305 with 256-bit keys and 128-bit tag} are great candidates, as are any Carter-Wegman constructions $t = h(m) \oplus \monospace{PRF}(r, k)$.
+
+We need 256 bits for the $\mac$ key, because checking a $\mac$ key can be done offline (the adversary takes a guess at the key, computes the tag using the guessed key and see if it matches the known tag) and is therefore vulnerable to Grover's algorithm.
+
+On the other hand, a 128-bit tag is not a security concern.
+This is because for one, the adversary has no way to issuing additional signing queries, so there is no concern for collision.
+The adversary can use the decapsulation oracle to check whether a forgery $(\hat{c}, \hat{t})$ is valid, but assuming that the $\mac$ is an ideal PRF, the adversary will have no better way of forgery than random guesses at the tag value.
+In an actual setting, it is easy for the peer to implement some rate limiting such that after only a small number of invalid ciphertext-tag pairs the adversary will simply be blocked.
+
 \section{Choice of MAC}

### 2024-06-18 lattices/compact-fo-using-mac/ml-kem/src/algebra.rs
Copied source code from RustCrypto/KEMs

@@ -1,0 +1,692 @@
+use core::ops::{Add, Mul, Sub};
+use hybrid_array::{typenum::U256, Array};
+use sha3::digest::XofReader;
+
+use crate::crypto::{PrfOutput, PRF, XOF};
+use crate::encode::Encode;
+use crate::param::{ArraySize, CbdSamplingSize};
+use crate::util::{Truncate, B32};
+
+pub type Integer = u16;
+
+/// An element of GF(q).  Although `q` is only 16 bits wide, we use a wider uint type to so that we
+/// can defer modular reductions.
+#[derive(Copy, Clone, Debug, Default, PartialEq)]
+pub struct FieldElement(pub Integer);
+
+impl FieldElement {
+    pub const Q: Integer = 3329;
+    pub const Q32: u32 = Self::Q as u32;
+    pub const Q64: u64 = Self::Q as u64;
+    const BARRETT_SHIFT: usize = 24;
+    #[allow(clippy::integer_division_remainder_used)]
+    const BARRETT_MULTIPLIER: u64 = (1 << Self::BARRETT_SHIFT) / Self::Q64;
+
+    // A fast modular reduction for small numbers `x < 2*q`
+    fn small_reduce(x: u16) -> u16 {
+        if x < Self::Q {
+            x
+        } else {
+            x - Self::Q
+        }
+    }
+
+    fn barrett_reduce(x: u32) -> u16 {
+        let product = u64::from(x) * Self::BARRETT_MULTIPLIER;
+        let quotient = (product >> Self::BARRETT_SHIFT).truncate();
+        let remainder = x - quotient * Self::Q32;
+        Self::small_reduce(remainder.truncate())
+    }
+
+    // Algorithm 11. BaseCaseMultiply
+    //
+    // This is a hot loop.  We promote to u64 so that we can do the absolute minimum number of
+    // modular reductions, since these are the expensive operation.
+    fn base_case_multiply(a0: Self, a1: Self, b0: Self, b1: Self, i: usize) -> (Self, Self) {
+        let a0 = u32::from(a0.0);
+        let a1 = u32::from(a1.0);
+        let b0 = u32::from(b0.0);
+        let b1 = u32::from(b1.0);
+        let g = u32::from(GAMMA[i].0);
+
+        let b1g = u32::from(Self::barrett_reduce(b1 * g));
+
+        let c0 = Self::barrett_reduce(a0 * b0 + a1 * b1g);
[…]

### 2024-06-18 lattices/compact-fo-using-mac/ml-kem/src/crypto.rs

@@ -1,0 +1,210 @@
+#![allow(dead_code)]
+
+use hybrid_array::{Array, ArraySize};
+use rand_core::CryptoRngCore;
+use sha3::{
+    digest::{ExtendableOutput, Update, XofReader},
+    Digest, Sha3_256, Sha3_512, Shake128, Shake256,
+};
+
+use crate::param::{CbdSamplingSize, EncodedPolynomial};
+use crate::util::B32;
+
+pub fn rand<L: ArraySize>(rng: &mut impl CryptoRngCore) -> Array<u8, L> {
+    let mut val = Array::default();
+    rng.fill_bytes(&mut val);
+    val
+}
+
+pub fn G(inputs: &[impl AsRef<[u8]>]) -> (B32, B32) {
+    let mut h = Sha3_512::new();
+    for x in inputs {
+        Digest::update(&mut h, x);
+    }
+    let out = h.finalize();
+
+    let mut a = B32::default();
+    let mut b = B32::default();
+
+    a.copy_from_slice(&out[..32]);
+    b.copy_from_slice(&out[32..]);
+    (a, b)
+}
+
+pub fn H(x: impl AsRef<[u8]>) -> B32 {
+    let mut h = Sha3_256::new();
+    Digest::update(&mut h, x);
+
+    // This odd conversion is needed because the `sha3` crate links against an old version of
+    // the `generic-array` crate.  It should be pretty cheap though, since there's only one
+    // allocation / no copies.
+    let mut out = B32::default();
+    h.finalize_into(out.as_mut_slice().into());
+    out
+}
+
+pub fn J(inputs: &[impl AsRef<[u8]>]) -> B32 {
+    let mut h = Shake256::default();
+    for x in inputs {
+        h.update(x.as_ref());
+    }
+    let mut r = h.finalize_xof();
+
+    let mut out = B32::default();
+    r.read(&mut out);
+    out
+}
+
+pub type PrfOutput<Eta> = EncodedPolynomial<<Eta as CbdSamplingSize>::SampleSize>;
+
+pub fn PRF<Eta>(s: &B32, b: u8) -> PrfOutput<Eta>
+where
+    Eta: CbdSamplingSize,
+{
+    let mut h = Shake256::default();
+    h.update(s.as_ref());
+    h.update(&[b]);
+    let mut r = h.finalize_xof();
+
[…]

### 2024-06-18 lattices/compact-fo-using-mac/ml-kem/src/kem.rs

@@ -1,0 +1,284 @@
+use core::marker::PhantomData;
+use hybrid_array::typenum::U32;
+use rand_core::CryptoRngCore;
+
+use crate::crypto::{rand, G, H, J};
+use crate::param::{DecapsulationKeySize, EncapsulationKeySize, EncodedCiphertext, KemParams};
+use crate::pke::{DecryptionKey, EncryptionKey};
+use crate::util::B32;
+use crate::{Encoded, EncodedSizeUser};
+
+/// A shared key resulting from an ML-KEM transaction
+pub(crate) type SharedKey = B32;
+
+/// A `DecapsulationKey` provides the ability to generate a new key pair, and decapsulate an
+/// encapsulated shared key.
+#[derive(Clone, Debug, PartialEq)]
+pub struct DecapsulationKey<P>
+where
+    P: KemParams,
+{
+    dk_pke: DecryptionKey<P>,
+    ek: EncapsulationKey<P>,
+    z: B32,
+}
+
+impl<P> EncodedSizeUser for DecapsulationKey<P>
+where
+    P: KemParams,
+{
+    type EncodedSize = DecapsulationKeySize<P>;
+
+    #[allow(clippy::similar_names)] // allow dk_pke, ek_pke, following the spec
+    fn from_bytes(enc: &Encoded<Self>) -> Self {
+        let (dk_pke, ek_pke, h, z) = P::split_dk(enc);
+        let ek_pke = EncryptionKey::from_bytes(ek_pke);
+
+        // XXX(RLB): The encoding here is redundant, since `h` can be computed from `ek_pke`.
+        // Should we verify that the provided `h` value is valid?
+
+        Self {
+            dk_pke: DecryptionKey::from_bytes(dk_pke),
+            ek: EncapsulationKey {
+                ek_pke,
+                h: h.clone(),
+            },
+            z: z.clone(),
+        }
+    }
+
+    fn as_bytes(&self) -> Encoded<Self> {
+        let dk_pke = self.dk_pke.as_bytes();
+        let ek = self.ek.as_bytes();
+        P::concat_dk(dk_pke, ek, self.ek.h.clone(), self.z.clone())
+    }
+}
+
+// 0xff if x == y, 0x00 otherwise
+fn constant_time_eq(x: u8, y: u8) -> u8 {
+    let diff = x ^ y;
+    let is_zero = !diff & diff.wrapping_sub(1);
+    0u8.wrapping_sub(is_zero >> 7)
+}
+
[…]

### 2024-06-18 lattices/compact-fo-using-mac/ml-kem/src/lib.rs

@@ -1,0 +1,229 @@
+#![no_std]
+#![cfg_attr(docsrs, feature(doc_auto_cfg))]
+#![doc = include_str!("../README.md")]
+#![doc(
+    html_logo_url = "https://raw.githubusercontent.com/RustCrypto/meta/master/logo.svg",
+    html_favicon_url = "https://raw.githubusercontent.com/RustCrypto/meta/master/logo.svg"
+)]
+#![warn(clippy::pedantic)] // Be pedantic by default
+#![warn(clippy::integer_division_remainder_used)] // Be judicious about using `/` and `%`
+#![allow(non_snake_case)] // Allow notation matching the spec
+#![allow(clippy::clone_on_copy)] // Be explicit about moving data
+#![deny(missing_docs)] // Require all public interfaces to be documented
+
+//! # Usage
+//!
+//! This crate implements the Module-Latice-based Key Encapsulation Method (ML-KEM) algorithm
+//! being standardized by NIST in FIPS 203.  ML-KEM is a KEM in the sense that it creates an
+//! (decapsulation key, encapsulation key) pair, such that anyone can use the encapsulation key to
+//! establish a shared key with the holder of the decapsulation key.  ML-KEM is the first KEM
+//! algorithm standardized by NIST that is designed to be resistant to attacks using quantum
+//! computers.
+//!
+//! ```
+//! # use ml_kem::*;
+//! # use ::kem::{Decapsulate, Encapsulate};
+//! let mut rng = rand::thread_rng();
+//!
+//! // Generate a (decapsulation key, encapsulation key) pair
+//! let (dk, ek) = MlKem768::generate(&mut rng);
+//!
+//! // Encapsulate a shared key to the holder of the decapsulation key, receive the shared
+//! // secret `k_send` and the encapsulated form `ct`.
+//! let (ct, k_send) = ek.encapsulate(&mut rng).unwrap();
+//!
+//! // Decapsulate the shared key and verify that it was faithfully received.
+//! let k_recv = dk.decapsulate(&ct).unwrap();
+//! assert_eq!(k_send, k_recv);
+//! ```
+//!
+//! [RFC 9180]: https://www.rfc-editor.org/info/rfc9180
+
+/// The inevitable utility module
+mod util;
+
+/// Section 2.4. Interpreting the Pseudocode
+/// Section 4.2.2. Sampling algorithms
[…]

### 2024-06-19 faster-fo-transform/manuscript/main.tex
Cleaned up research project manuscript

% TODO: instead of signing the ciphertext, how about signing a hash of the ciphertext?
% TODO: Kyber round 3's implementation already contains H(c), we can just use that for the MAC
% TODO: need to look into using GCM as a more efficient MAC
% TODO: what is the quantum security of the various hash functions and how do they impact the strength of the FO transform?
% TODO: Need to read the rest of the U transformation since Kyber round 3 uses U^{\not\bot} but ML-KEM (FIPS 203) uses U^{\not\bot}_m, which removed the hash of the ciphertext

%%%% IACR Transactions TEMPLATE %%%%
% This file shows how to use the iacrtrans class to write a paper.
% Written by Gaetan Leurent gaetan.leurent@inria.fr (2020)
% Public Domain (CC0)

%%%% 1. DOCUMENTCLASS %%%%
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[floatrow,journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

%%%% 2. PACKAGES %%%%
\usepackage{algorithm}
\usepackage{algpseudocode}
% \usepackage[demo]{graphicx}

%%%% 3. AUTHOR, INSTITUTE %%%%
\author{
    Ganyu Xu\inst{1}
    \and Guang Gong\inst{1}
    \and Kalikinkar Mandal\inst{2}
}
\institute{
  University of Waterloo, Waterloo, Canada, \email{{g66xu,ggong}@uwaterloo.ca}
  \and
  University of New Brunswick, New Brunswick, Canada, \email{kmandal@unb.ca}
}
%%%% NOTES:
% - We need a city name for indexation purpose, even if it is redundant
%   (eg: University of Atlantis, Atlantis, Atlantis)
% - \inst{} can be omitted if there is a single institute,
%   or exactly one institute per author

% Custom commands
\newcommand{\monospace}{\texttt}
\newcommand{\pke}{\monospace{PKE}}
\newcommand{\keygen}{\monospace{KeyGen}}
\newcommand{\encrypt}{\monospace{E}}
\newcommand{\decrypt}{\monospace{D}}
\newcommand{\kem}{\monospace{KEM}}
\newcommand{\encap}{\monospace{Encap}}
\newcommand{\decap}{\monospace{Decap}}
\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\monospace{MAC}}
\newcommand{\sign}{\monospace{MAC}}
\newcommand{\verify}{\monospace{MAC.Verify}}
\newcommand{\pk}{\monospace{pk}}
\newcommand{\sk}{\monospace{sk}}
\newcommand{\pco}{\monospace{PCO}}
\newcommand{\cvo}{\monospace{CVO}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\fotplus}{$\monospace{FOT}^+$}

%%%% 4. TITLE %%%%
\title{
Fast Fujisaki-Okamoto transformation  using encrypt-then-mac and applications to Kyber}
%%%% NOTES:
% - If the title is too long, or includes special macro, please
%   provide a "running title" as optional argument: \title[Short]{Long}
% - You can provide an optional subtitle with \subtitle.

\begin{document}

\maketitle

%%%% 5. KEYWORDS %%%%
\keywords{
    Key encapsulation mechanism, 
    post-quantum cryptography,
    lattice cryptography,
    Fujisaki-Okamoto transformation
}

%%%% 6. ABSTRACT %%%%
\begin{abstract}
  The modular Fujisaki-Okamoto (FO) transformation takes public-key encryption with weaker security and constructs a key encapsulation mechanism (KEM) with indistinguishability under adaptive chosen ciphertext attacks. While the modular FO transform enjoys tight security bound and quantum resistance, it also suffers from computational inefficiency due to using de-randomization and re-encryption for providing ciphertext integrity. In this work, we propose an alternative modular FO transformation that replaces re-encryption with a message authentication code (MAC) and prove the security bound of our construction. We then instantiate a concrete instance with ML-KEM and show that when re-encryption incurs significant computational cost, our construction provides substantial runtime speedup and reduced memory footprint.
\end{abstract}

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks. Later works extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (standardized by NIST into ML-KEM \cite{key2023mechanism}).

The current state of the FO transformation enjoys tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many open questions. One such problem is the use of re-encryption for providing ciphertext integrity \cite{bernstein2018towards}, which requires the decryption/decapsulation to run the encryption routine as a subroutine. In many post-quantum schemes, such as Kyber, the encryption routine is substantially computationally more expensive than the decryption routine.

The problem of ciphertext integrity was solved in symmetric cryptography. Given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}. We took inspiration from this strategy and applied a similar technique to provide ciphertext integrity for a public-key encryption scheme, which then translates to an \monospace{IND-CCA} secure $\kem$. Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of a tag, which should offer significant performance improvements while maintaining comparable level of security.

The main challenge in applying ``encrypt-then-mac'' to public-key cryptography is the lack of a pre-shared $\mac$ key. We proposed to derive the shared $\mac$ key by hashing the plaintext message. We will prove in section \ref{sec:main-results} that, under the random oracle model, the $\mac$ key is securely hidden behind the hash function, and producing a valid pair of ciphertext and tag without full knowledge of the plaintext constitutes a forgery attack on the message authentication code. Thanks to the modular construction in \cite{hofheinz2017modular}, providing ciphertext integrity in the underlying encryption scheme gives us an \monospace{IND-CCA} secure $\kem$ for free.

In section \ref{sec:exp-results}, we instantiate concrete instances of our proposed transformation by modifying ML-KEM. We will demonstrate that, at the cost of small increase in encryption runtime and ciphertext size, our construction reduces both the runtime and memory footprint of the decryption routine. % TODO: need to write more?

\section{Preliminaries and previous results}

\subsection{Public-key encryption scheme}
We define a public key encryption scheme $\pke$ to be a collection of three routines $(\keygen, \encrypt, \decrypt)$ defined over a finite message space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$. Many encryption routines are probabilistic, and we define their source of randomness to come from some coin space $\mathcal{R}$.

The encryption routine $\encrypt(\pk, m)$ takes a public key, a plaintext message, and outputs a ciphertext $c \in \mathcal{C}$. Where the encryption routine is probabilistic, specifying a pseudorandom seed $r \in \mathcal{R}$ will make the encryption routine behave deterministically. The decryption routine $\decrypt(\sk, c)$ takes a secret key, a ciphertext, and outputs the decryption $\hat{m}$ if the ciphertext is valid under the given secret key, or the rejection symbol $\bot$ if the ciphertext is invalid.

\subsubsection{Correctness}
It is common to require a $\pke$ to be perfectly correct, meaning that for all possible keypairs $(\pk, \sk)$ and plaintext messages $m \in \mathcal{M}$, $\decrypt(\sk, \encrypt(\pk, m)) = m$ at all times. However, some encryption schemes, including many popular lattice-based schemes, admit a non-zero probability of decryption failure: $\decrypt(\sk, \encrypt(\pk, m)) \neq m$. Furthermore, \cite{hofheinz2017modular} and \cite{avanzi2019crystals} explained how decryption failure played a role in an adversary's advantage. In this paper, we inherit the definition for correctness from \cite{hofheinz2017modular}:

\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
    A public key encryption scheme $\pke$ is $\delta$-correct if

    \begin{equation*}
        \mathbf{E}[
            \max_{m \in \mathcal{M}} P[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)]
        ] \leq \delta
    \end{equation*}

    Where the expectation is taken over the probability distribution of keypairs $(\pk, \sk) \leftsample \keygen()$
\end{definition}

\subsubsection{Security}
We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}. Specifically, we first define the $\monospace{OW-ATK}$ and the $\monospace{IND-CPA}$ game as they pertain to a public key encryption scheme. In later section we will define the $\monospace{IND-CCA}$ game as it pertains to a key encapsulation mechanism.

In the \monospace{OW-ATK} game, an adversary's goal is to recover the decryption of a randomly generated ciphertext.

\begin{figure}
    \centering

    \begin{algorithm}[H]
        \caption{
            The $\monospace{OW-ATK}$ game
        }\label{alg:pke-ow-atk-game}
        \begin{algorithmic}[1]
            \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
            \State $m^\ast \leftsample \mathcal{M}$
            \State $c^\ast \leftsample \encrypt(\pk, m)^\ast$
            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\monospace{ATK}}(1^\lambda, \pk, c^\ast)$
            \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
        \end{algorithmic}
    \end{algorithm}

    \caption{The \monospace{OW-ATK} game}
    \label{fig:pke-ow-atk-game}
\end{figure}

The adversary $\mathcal{A}$ with access to oracle(s) $\mathcal{O}_\monospace{ATK}$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$. The \textit{advantage} $\epsilon_\monospace{OW-ATK}$ of an adversary in this game is the probability that it wins the game.

The choice of oracle(s) $\mathcal{O}_\monospace{ATK}$ depends on the choice of $\monospace{ATK}$. Specifically:

\begin{equation*}
    \mathcal{O}_\monospace{ATK} = \begin{cases}
        - & \monospace{ATK} = \monospace{CPA} \\
        \monospace{PCO} & \monospace{ATK} = \monospace{PCA} \\
        \monospace{CVO} & \monospace{ATK} = \monospace{VA} \\
        \monospace{PCO, CVO} & \monospace{ATK} = \monospace{PCVA}
    \end{cases}
\end{equation*}

Where the definitions of plaintext-checking oracle $\monospace{PCO}$ and the ciphertext-validation oracle $\monospace{CVO}$ are inherited from \cite{hofheinz2017modular}

\begin{figure}
    \begin{floatrow}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\pco(m \in \mathcal{M}, c \in \mathcal{C})$}\label{alg:pco}
                \begin{algorithmic}[1]
                    \State \Return $\llbrack \decrypt(\sk, c) = m \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }
        {\caption{The Plaintext-Checking Oracle $\pco$}}
        \ffigbox{
            \begin{algorithm}[H]
                \caption{$\cvo(c \in \mathcal{C})$}\label{alg:cvo}
                \begin{algorithmic}[1]
                    \State \Return $\llbrack \decrypt(\sk, c) \in \mathcal{M} \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        }
        {\caption{the Ciphertext-Validation Oracle $\cvo$}}
    \end{floatrow}
    \label{fig:pco-and-cvo}
\end{figure}

In the \monospace{IND-CPA} game, an adversary's goal is to distinguish the encryption of one message from the encryption of another message. Given the public key, the adversary outputs two adversarially chosen messages and obtains the encryption of a random choice between these two messages. The adversary wins the \monospace{IND-CPA} game if it correctly identifies which message the encryption is obtained from.

\begin{algorithm}
    \caption{The $\monospace{IND-CPA}$ game}\label{alg:pke-ind-cpa-game}

    \begin{algorithmic}[1]
        \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
        \State $(m_0, m_1) \leftsample \mathcal{A}(a^\lambda, \pk)$
        \State $b \leftsample \{0, 1\}$
        \State $c^\ast \leftsample \encrypt(\pk, m_b)$
        \State $\hat{b} \leftsample \mathcal{A}(1^\lambda, \pk, c^\ast)$
        \State \Return $\llbrack b = \hat{b} \rrbrack$
    \end{algorithmic}
\end{algorithm}

The \textit{advantage} $\epsilon_\monospace{IND-CPA}$ of an \monospace{IND-CPA} adversary $\mathcal{A}$ is defined by

\begin{equation*}
    \epsilon_\monospace{IND-CPA} = \norm{
        P[\hat{b} = b] - \frac{1}{2}
    }
\end{equation*}

\subsection{Key encapsulation mechanism}
A key encapsulation mechanism $\kem$ is a collection of three routines $(\keygen, \encap, \decap)$ defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$. The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$. $\encap(\pk)$ is a probabilistic routine that takes a public key $\pk$ and outputs a pair of values $(c, K)$ where $c \in \mathcal{C}$ is the encapsulation (or ciphertext) of the shared secret $k \in \mathcal{K}$. $\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $k$ if the ciphertext is valid, or the rejection symbol $\bot$ if the ciphertext is invalid.

The IND-CCA security of a $\kem$ is defined by an adversarial game in which an adversary's goal is to distinguish pseudorandom shared secret (generated by running the $\encap$ routine) and a truly random value.

\begin{algorithm}
    \caption{
        \monospace{IND-CCA} game for \kem
    }\label{alg:kem-ind-cca-game}

    \begin{algorithmic}[1]
        \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
        \State $(c^\ast, k_0) \leftsample \encap(\pk)$
        \State $k_1 \leftsample \mathcal{K}$
        \State $b \leftsample \{0, 1\}$
        \State $\hat{b} \leftsample \mathcal{A}_\monospace{IND-CCA}^{\mathcal{O}_\decap}(
            1^\lambda, \pk, c^\ast, k_b
        )$
        \State \Return $\llbrack \hat{b} = b \rrbrack$
    \end{algorithmic}
\end{algorithm}

The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key. The advantage $\epsilon_\monospace{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\monospace{IND-CCA}$ is defined by

\begin{equation*}
    \epsilon_\monospace{IND-CCA} = \norm{
        P[\hat{b} = b] - \frac{1}{2}
    }
\end{equation*}

\subsection{Message authentication code}
A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$. The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$. The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\monospace{1}$ if the message-tag pair is valid under the secret key, or $\monospace{0}$ otherwise.

The security of a MAC is defined in an adversarial game in which an adversary, with access to some signing oracle $\mathcal{O}_\sign(m)$, tries to forge a new valid message-tag pair that has never been queried before. The existential unforgeability under chosen message attack (EUF-CMA) game is shown below:

\begin{algorithm}
    \caption{The EUF-CMA game}\label{alg:mac-euf-cma-game}

    \begin{algorithmic}[1]
        \State $k^\ast \leftsample \mathcal{K}$
        \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
        \State \Return $
            \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
            \monospace{ and } (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
            \rrbrack
        $
    \end{algorithmic}
\end{algorithm}

The advantage $\epsilon_\monospace{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.

\subsection{Modular Fujisaki-Okamoto transformation}
The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \monospace{OW-CPA} or \monospace{IND-CPA}) and outputs a $\pke$ with stronger security. A later variation \cite{hofheinz2017modular} improved the original construction in \cite{fujisaki1999secure} by accounting for decryption failures, tightening security bounds, and providing a modular construction that first transforms \monospace{OW-CPA}/\monospace{IND-CPA} $\pke$ into \monospace{OW-PCVA} $\pke$ by providing ciphertext integrity through re-encryption (the \monospace{T} transformation), then transforming the \monospace{OW-PCVA} $\pke$ into an \monospace{IND-CCA} $\kem$ (the \monospace{U} transformation).

% TODO: Restating the algorithm seems redundant

Particularly relevant to our results are two variations of the \monospace{U} transformation: $\monospace{U}^\bot$ (KEM with explicit rejection) and $\monospace{U}^{\not\bot}$ (KEM with implicit rejection). If $\pke$ is \monospace{OW-PCVA} secure, then $\monospace{U}^\bot$ transforms $\pke$ into an \monospace{IND-CCA} secure $\kem^\bot$:

\begin{theorem}\label{thm:u-bot}
    For any \monospace{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^\bot$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \monospace{OW-PCVA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ that makes at most $q_H$ queries to $\pco$ and $\cvo$ such that

    \begin{equation*}
        \epsilon_\kem \leq \epsilon_\pke
    \end{equation*}
\end{theorem}

Similarly, if $\pke$ is \monospace{OW-PCA} secure, then $\monospace{U}^{\not\bot}$ transforms $\pke$ into an \monospace{IND-CCA} secure $\kem^{\not\bot}$

\begin{theorem}\label{thm:u-notbot}
    For any \monospace{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^{\not\bot}$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \monospace{OW-CPA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ issuing at most $q_H$ queries to $\pco$ such that:

    \begin{equation*}
        \epsilon_\kem \leq \frac{q_H}{\norm{\mathcal{M}_\pke}} + \epsilon_\pke
    \end{equation*}
\end{theorem}

The modularity of the \monospace{T} and \monospace{U} transformation allows us to tweak only the \monospace{T} transformation (see section \ref{sec:main-results}), obtain \monospace{OW-PCVA} security, then automatically get \monospace{IND-CCA} security for free. This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:exp-results}).

\section{FO transform with encrypt-then-mac for achieving ciphertext integrity}\label{sec:main-results}
In this section, we present two different encrypt-then-mac modes using symmetric-key approaches   to achieve ciphertext integrity, which result in two different fast FOTs.  The first method is to apply an MAC over ciphertext [???], referred to as \emph{standard Enc-then-MAC}, and the second method is adopted from GCM [??] mode in authenticated encryption (AE) using polynomial hash, referred to as \emph{poly Enc-then-MAC}. In other words, we will replace the public-key re-encryption  in FOT  by symmetric-key MAC approaches, and the resulting transform is referred to as \emph{FOT+}.

\subsection{FOT+ in standard encrypt-then-MAC}

Let $\pke(\keygen, \encrypt, \decrypt)$ be a probabilistic public-key encryption scheme defined over message space $\mathcal{M}_\pke$, ciphertext space $\mathcal{C}$, and coin space $\mathcal{R}$. Let $\mac$ be a deterministic and perfectly correct message authentication code defined over key space $\mathcal{K}_\mac$, message space $\mathcal{M}_\mac$, and tag space $\mathcal{T}_\mac$. Let $G: \mathcal{M}_\pke \rightarrow \mathcal{R} \times \mathcal{K}_\mac$ be a hash function that hashes a plaintext message into a pseudorandom coin and a $\mac$ key. The $\monospace{T}_\etm[\pke, \mac, G]$ transformation takes the input $\pke$, $\mac$, and hash function $G$, and outputs a public-key encryption scheme $\pke_\etm(\keygen, \encrypt_\etm, \decrypt_\etm)$ where as the key generation routine remains unchanged, and the encryption/decryption routines are as follows:

\begin{algorithm}
    \caption{$\encrypt_\etm$(\pk, m)}\label{alg:encrypt-etm}

    \begin{algorithmic}[1]
        \State $(r, k) \leftarrow G(m)$
        \State $c \leftarrow \encrypt(\pk, m; r)$
        \State $t \leftarrow \sign(k, c)$
        \State \Return $(c, t)$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}
    \caption{$\decrypt_\etm$(\sk, (c, t))}\label{alg:decrypt-etm}

    \begin{algorithmic}[1]
        \State $\hat{m} \leftarrow \decrypt(\sk, c)$
        \State $(\hat{r}, \hat{k}) \leftarrow G(m)$
        \If {$\verify(\hat{k}, c, t) = 0$}
            \State \Return $\bot$
        \EndIf
        \State \Return $\hat{m}$
    \end{algorithmic}
\end{algorithm}

We claim that if the input $\pke$ is OW-CPA secure and $\mac$ is existentially unforgeable, then under the random oracle model, $\pke_\etm$ is OW-PCVA secure with non-tight security reduction.

\begin{theorem}\label{thm:if-owcpa-then-owpcva}
    If $\pke$ is $\delta$-correct, then $\pke_\etm$ is $\delta$-correct. In addition, for every \monospace{OW-PCVA} adversary $\mathcal{A}_\monospace{OW-PCVA}$ against $\pke_\etm$ that makes $q_P$ PCO queries, $q_V$ CVO queries, $q_G$ hash queries to $G$, and that has advantage $\epsilon_\monospace{OW-PCVA}$ there exists an existential forgery adversary $\mathcal{A}_\mac$ against the underlying $\mac$ with advantage $\epsilon_\mac$ and some $\monospace{OW-CPA}$ adversary $\mathcal{A}_\monospace{OW-CPA}$ against the underlying $\pke$ with advantage $\epsilon_\monospace{OW-CPA}$ such that

    \begin{equation*}
        \epsilon_\monospace{OW-PCVA}
        \leq (q_G + q_P) \cdot \delta
            + q_V \cdot \epsilon_\mac
            + (q_G + q_P + 1) \cdot \epsilon_\monospace{OW-CPA}
    \end{equation*}
\end{theorem}

Furthermore, if the input $\pke$ is additionally IND-CPA secure, then $\pke_\etm$ is OW-PCVA secure with tight security reduction.

\begin{corollary}\label{thm:if-indcpa-then-owpcva}
    For every \monospace{OW-PCVA} adversary $\mathcal{A}_\monospace{OW-PCVA}$ against $\pke_\etm$ that makes $q_P$ PCO queries, $q_V$ CVO queries, $q_G$ hash queries to $G$, and that has advantage $\epsilon_\monospace{OW-PCVA}$ there exists an existential forgery adversary $\mathcal{A}_\mac$ against the underlying $\mac$ with advantage $\epsilon_\mac$ and some $\monospace{IND-CPA}$ adversary $\mathcal{A}_\monospace{IND-CPA}$ against the underlying $\pke$ with advantage $\epsilon_\monospace{IND-CPA}$ such that

    \begin{equation*}
        \epsilon_\monospace{OW-PCVA}
        \leq (q_G + q_P) \cdot \delta
            + q_V \cdot \epsilon_\mac
            + \frac{1 + 2q_G}{\norm{\mathcal{M}_\pke}}
            + 3\epsilon_\monospace{IND-CPA}
    \end{equation*}
\end{corollary}

\begin{proof}
    Since no modification was made to the internals of the input $\pke$, the correctness of the transformed scheme is trivially identical to the correctness of the input scheme.

    We will prove the security claim using a sequence of games \cite{shoup2004sequences}, then prove the security claim in the corollary \ref{thm:if-indcpa-then-owpcva} by making a few modifications. This proof borrows heavily from the proof presented in \cite{hofheinz2017modular}.

    \begin{figure}
        \centering

        \begin{algorithm}[H]
            \caption{Sequence of games}\label{alg:sequence-of-games}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $m^\ast \leftsample \mathcal{M}_\pke$
                \State $(r^\ast, k^\ast) \leftarrow G(m^\ast)$
                    \Comment{Game 0-2}
                \State $r^\ast \leftsample \mathcal{R}, k^\ast \leftsample \mathcal{K}_\mac$
                    \Comment{Game 3}
                \State $c^\ast \leftarrow \encrypt(\pk, m^\ast; r^\ast)$
                \State $t^\ast \leftarrow \sign(k^\ast, c^\ast)$
                \State $
                    \hat{m} \leftarrow \mathcal{A}^{\mathcal{O}^G, \pco, \cvo}_\monospace{OW-PCVA}(
                        1^\lambda, \pk, (c^\ast, t^\ast)
                    )
                $
                    \Comment{Game 0}
                \State $
                    \hat{m} \leftarrow \mathcal{A}^{\mathcal{O}^G, \pco_1, \cvo}_\monospace{OW-PCVA}(
                        1^\lambda, \pk, (c^\ast, t^\ast)
                    )
                $
                    \Comment{Game 1}
                \State $
                    \hat{m} \leftarrow \mathcal{A}^{\mathcal{O}^G, \pco_1, \cvo_1}_\monospace{OW-PCVA}(
                        1^\lambda, \pk, (c^\ast, t^\ast)
                    )
                $
                    \Comment{Game 2-3}
                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
            \end{algorithmic}
        \end{algorithm}

        \caption{Sequence of \monospace{Game 0} to \monospace{Game 3} in proof of theorem \ref{thm:if-owcpa-then-owpcva}}
        \label{fig:enter-label}
    \end{figure}

    \begin{algorithm}
        \caption{$\pco(m, (c, t))$}\label{alg:etm-pco}
        \begin{algorithmic}[1]
            \State $\hat{m} \leftarrow \decrypt(\sk, c)$
            \State $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$
            \If{$\verify(\hat{k}, c, t) = 0$}
                \State \Return $0$
            \EndIf
            \State \Return $\llbrack \hat{m} = m \rrbrack$
        \end{algorithmic}
    \end{algorithm}

    \begin{algorithm}
        \caption{$\pco_1(m, (c, t))$}\label{alg:etm-pco1}
        \begin{algorithmic}[1]
            \State $(r, k) \leftarrow G(\hat{m})$
            \State \Return $\llbrack 
                \encrypt(\pk, m; r) = c
            \rrbrack \monospace{ and } \llbrack
                \verify(k, c, t) = 1
            \rrbrack$
        \end{algorithmic}
    \end{algorithm}

    \begin{algorithm}
        \caption{$\cvo(c, t)$}\label{alg:etm-cvo}
        \begin{algorithmic}[1]
            \State $\hat{m} \leftarrow \decrypt(\sk, c)$
            \State $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$
            \State \Return $\verify(\hat{k}, c, t)$
        \end{algorithmic}
    \end{algorithm}

    \begin{algorithm}
        \caption{$\cvo_1(c, t)$}\label{alg:etm-cvo1}

        \begin{algorithmic}[1]
            \If{
                $\exists (\tilde{m}, \tilde{r}, \tilde{k}) \in \mathcal{O}^G$ such that
                $\encrypt(\pk, \tilde{m}; \tilde{r}) = c$
                and $\verify(\tilde{k}, c, t)$
            }
                \State \Return 1
            \EndIf
            \State \Return 0
        \end{algorithmic}
    \end{algorithm}

    \monospace{Game 0} is the \monospace{OW-PCVA} game. Let $\epsilon_0$ denote the \monospace{OW-PCVA} adversary's advantage in \monospace{Game 0}, then $\epsilon_0 = \epsilon_\monospace{OW-PCVA}$

    \monospace{Game 1} is identical to \monospace{Game 0}, except $\pco$ is replaced with $\pco_1$. Because $\encrypt_\etm$ is a deterministic encryption routine, the two games differ from the adversary's perspective if and only if any of the $\pco$ query $(m, (c, t))$ causes decryption failure $\decrypt(\sk, \encrypt(\pk, m; r)) \neq m$. The probability of decryption failure for any single $\pco$ query is bounded by $\delta$, so the overall probability of having at least one query causing decryption failure is at most $q_P \cdot \delta$. Let $\epsilon_0$ and $\epsilon_1$ respectively denote $\mathcal{A}_\monospace{OW-PCVA}$'s advantage in \monospace{Game 0} and \monospace{Game 1} respectively, then by the difference lemma \cite{shoup2004sequences}:

    \begin{equation}\label{eq:ep0-ep1}
        \epsilon_0 - \epsilon_1 \leq q_P \cdot \delta
    \end{equation}

    \monospace{Game 2} is identical to \monospace{Game 1} except $\cvo$ is replaced with $\cvo_1$. $\cvo_1$ replaces the decryption routine with the deterministic encryption routine and the $\mac$ key derivation with ``searching through hash oracle records''. Therefore, there are exactly two scenarios in which \monospace{Game 2} differ from \monospace{Game 1} from the adversary's perspective.

    In the first scenario, the queried ciphertext $(c, t)$ has a matching hash query $(\tilde{m}, \tilde{r}, \tilde{k})$, but $(\tilde{m}, \tilde{r})$ causes decryption failure: $\decrypt(\sk, \encrypt(\pk, m; r)) \neq m$. For each hash query, the probability that $(\tilde{m}, \tilde{r})$ causes decryption failure is bounded by $\delta$, so the probability of having at least one such hash query is at most $q_G \cdot \delta$.

    In the second scenario, the queried ciphertext $(c, t)$ has no matching hash query. Under the random oracle model, this means that the $\mac$ key $k$ used to sign $c$ is an unknown and uniformly random key from the adversary's perspective. In other words, $(c, t)$ is an existential forgery. The probability of producing a single forgery is bounded by the advantage of a MAC adversary, so the probability of having at least one dishonest $\cvo$ query is at most $q_V \cdot \epsilon_\mac$.

    Denote the $\monospace{OW-PCVA}$ adversary's advantage in game 2 by $\epsilon_2$, then by the difference lemma:

    \begin{equation}\label{eq:ep1-ep2}
        \epsilon_1 - \epsilon_2 \leq q_G \cdot \delta + q_V \cdot \epsilon_\mac
    \end{equation}

    In \monospace{Game 3}, the challenge encryption routine is modified. Instead of pseudorandomly deriving the coin $r^\ast$ and the $\mac$ key $k^\ast$ from $G$, the coin and the $\mac$ key are uniformly sampled from their respective domain. Under the random oracle model, \monospace{Game 3} and \monospace{Game 2} are indistinguishable from the adversary's perspective unless the adversary queries $G$ or $\pco$ with the value $m^\ast$. Denote the probability of ``adversary querying $G$ or $\pco$ with $m^\ast$'' by $P[\monospace{QUERY}^\ast]$, and the adversary's advantage in \monospace{Game 3} by $\epsilon_3$, then by the difference lemma:

    \begin{equation}\label{eq:ep2-ep3}
        \epsilon_2 - \epsilon_3 \leq P[\monospace{QUERY}^\ast]
    \end{equation}

    A standard \monospace{OW-CPA} adversary against the underlying $\pke$ can simulate \monospace{Game 3} for an \monospace{OW-PCVA} adversary, since $\pco_1$ and $\cvo_1$ only make use of the public key $\pk$ and the hash oracle $\mathcal{O}^G$, the challenge encryption $c^\ast$ is obtained using a truly random coin, and the $\mac$ key can be uniformly sampled. After the \monospace{OW-PCVA} adversary outputs a guess, the \monospace{OW-CPA} adversary can simply pass \monospace{OW-PCVA}'s output. The \monospace{OW-CPA} adversary wins if and only if the \monospace{OW-CPA} adversary wins \monospace{Game 3}:

    \begin{equation}\label{eq:ep3-owcpa}
        \epsilon_3 = \epsilon_\monospace{OW-CPA}
    \end{equation}

    We can construct another \monospace{OW-CPA} adversary that simulates \monospace{Game 3} for an \monospace{OW-PCVA} adversary. After the \monospace{OW-PCVA} adversary halts, this \monospace{OW-CPA} adversary picks and outputs a random value $\tilde{m}$ from all possible values recorded on the tape of the hash oracle $\mathcal{O}^G$ and $\pco_1$. If $\monospace{QUERY}^\ast$ occurs, then the \monospace{OW-CPA} adversary wins the game with probability $\frac{1}{q_G + q_P}$. In other words:

    \begin{equation}\label{eq:querystar-owcpa}
        \epsilon_\monospace{OW-CPA} = \frac{1}{q_G + q_P} \cdot P[\monospace{QUERY}^\ast]
    \end{equation}

    Combining equations \ref{eq:ep0-ep1}, \ref{eq:ep1-ep2}, \ref{eq:ep2-ep3}, \ref{eq:ep3-owcpa}, and \ref{eq:querystar-owcpa} gives the security bound in theorem \ref{thm:if-owcpa-then-owpcva}.

    We make two modifications to prove corollary \ref{thm:if-indcpa-then-owpcva}. First, we invoke a well-known result that the IND-CPA security of a $\pke$ with sufficiently large message space implies its OW-CPA security.

    \begin{lemma}
        For every \monospace{OW-CPA} adversary with advantage $\epsilon_\monospace{OW-CPA}$ there exists an \monospace{IND-CPA} adversary with advantage $\epsilon_\monospace{IND-CPA}$ such that

        \begin{equation}\label{eq:indcpa-implies-owcpa}
            \epsilon_\monospace{OW-CPA} \leq \frac{1}{\norm{\mathcal{M}_\pke}} + \epsilon_\monospace{IND-CPA}
        \end{equation}
    \end{lemma}

    Second, we borrow results from \cite{hofheinz2017modular} and construct an \monospace{IND-CPA} adversary to bound $P[\monospace{QUERY}^\ast]$:

    \begin{equation}\label{eq:querystar-indcpa}
        \frac{1}{2}P[\monospace{QUERY}^\ast] 
        \leq \epsilon_\monospace{IND-CPA} + \frac{q_G}{\norm{\mathcal{M}_\pke}}
    \end{equation}

    Combining equations \ref{eq:ep0-ep1}, \ref{eq:ep1-ep2}, \ref{eq:ep2-ep3}, \ref{eq:ep3-owcpa}, \ref{eq:indcpa-implies-owcpa}, and \ref{eq:querystar-indcpa} into theorem \ref{thm:if-owcpa-then-owpcva} completes the proof of the corollary.
\end{proof}

\subsection{FO Transform+ in poly Enc-then-MAC}

\section{Applications to Kyber}

\subsection{Kyber using FO Transform+}
\subsection{Performance comparisons}\label{sec:exp-results}
% TODO: Count how many Keccak permutations
% TODO: Count how many NTT operations
% 
Comparisons will do for Kyber and Kyber$^+$, define it.

\subsection{Experimental results}

\section{Conclusions and future works}\label{sec:future-works}

%%%% 8. BILBIOGRAPHY %%%%
% TODO: figure out how to use bibliography
\bibliographystyle{alpha}
\bibliography{abbrev3,crypto,biblio}
%%%% NOTES
% - Download abbrev3.bib and crypto.bib from https://cryptobib.di.ens.fr/
% - Use bilbio.bib for additional references not in the cryptobib database.
%   If possible, take them from DBLP.

\end{document}

### 2024-06-19 lattices/compact-fo-using-mac/rounds.py

@@ -1,26 +1,0 @@
-"""Estimate the cost of various subroutines of Kyber, measured in the number of Keccak permutations
-and the number of NTT operations (including NTT and invert_NTT)
-"""
-from enum import Enum
-
-class SecurityLevels(Enum):
-    Kyber512 = 1
-    Kyber768 = 2
-    Kyber1024 = 3
-
-class KyberParameters:
-    KYBER_Q = 3329
-    KYBER_Q_BITS = 12
-    KYBER_N = 256
-    # Defaults to Kyber512 but will be modified
-    KYBER_K = 2
-    KYBER_ETA1 = 3
-    KYBER_ETA2 = 2
-
-    def __init__(self, level: SecurityLevels):
-        if level == SecurityLevels.Kyber512:
-            self.KYBER_K = 2
-            self.KYBER_ETA1 = 3
-            self.KYBER_ETA2 = 
-
-

### 2024-06-23 faster-fo-transform/randomized-etm/main.tex
Start thinking about randomized EtM

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
% \usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}

% \usepackage{floatrow}

\newcommand{\monospace}{\texttt}
\newcommand{\pke}{\monospace{PKE}}
\newcommand{\keygen}{\monospace{KeyGen}}
\newcommand{\encrypt}{\monospace{E}}
\newcommand{\decrypt}{\monospace{D}}
\newcommand{\kem}{\monospace{KEM}}
\newcommand{\encap}{\monospace{Encap}}
\newcommand{\decap}{\monospace{Decap}}
\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\monospace{MAC}}
\newcommand{\sign}{\monospace{MAC}}
\newcommand{\verify}{\monospace{MAC.Verify}}
\newcommand{\pk}{\monospace{pk}}
\newcommand{\sk}{\monospace{sk}}
\newcommand{\pco}{\monospace{PCO}}
\newcommand{\cvo}{\monospace{CVO}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newlength{\wdth}
\newcommand{\strike}[1]{\settowidth{\wdth}{#1}\rlap{\rule[.5ex]{\wdth}{.4pt}}#1}
\newcommand{\notrand}{{\strike{\$}}}
\newcommand{\rand}{\$}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{Randomized encrypt-then-MAC}
\author{Ganyu (Bruce) Xu (g66xu)}
% \date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{Introduction}
In the modular FO transform\cite{hofheinz2017modular}, the authors constructed IND-CCA KEM using two steps. The first step is adding ciphertext integrity to a one-way secure encryption scheme (the \monospace{T} transformation). The second step is constructing the IND-CCA KEM out of the OW-PCVA secure encryption scheme (the \monospace{U} transformation). The \monospace{T} transformation uses two techniques for achieving ciphertext integrity, which the authors called ``\emph{de-randomization}'' and ``\emph{re-encryption}''.

Our current proposal is to replace re-encryption with a message authentication code (MAC), but we maintained the deterministic encryption routine. For clarity we will call this ``\emph{de-randomized encrypt-then-MAC}''. The algorithm is specified in figure \ref{fig:derand-etm}.

\begin{figure}[h]
  \begin{minipage}{.49\textwidth}  % Adjust width as needed
    \begin{algorithm}[H]
      \caption{$\encrypt^\notrand_\etm(\pk, m)$}
      \begin{algorithmic}[1]
        \State $(r, k_\mac) \leftarrow G(m)$
        \State $c \leftarrow \encrypt(\pk, m; r)$
        \State $t \leftarrow \sign(k_\mac, c)$
        \State \Return $(c, t)$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
  \hfill  % Add space between algorithms
  \begin{minipage}{.49\textwidth}  % Adjust width as needed
    \begin{algorithm}[H]
      \caption{$\decrypt^\notrand_\etm(\sk, (c, t))$}
      \begin{algorithmic}[1]
        \State $\hat{m} \leftarrow \decrypt(\sk, c)$
        \State $(\hat{r}, \hat{k}_\mac) \leftarrow G(\hat{m})$
        \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
          \State \Return $\bot$
        \EndIf
        \State \Return $\hat{m}$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
  \caption{Derandomized encrypt-then-MAC}\label{fig:derand-etm}
\end{figure}

For the past week I have been thinking about removing the de-randomization, as well. This means that we no longer derive the pseudorandom coin from the message. If the input $\pke$ has a randomized encryption routine, then the transformed encrypt-then-mac scheme will also be randomized. We call this ``\emph{randomized encrypt-then-MAC}''. The algorithm is specified in figure \ref{fig:rand-etm}

\begin{figure}[h]
    \center
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\encrypt^\rand_\etm(\pk, m)$}\label{alg:rand-etm-encrypt}
            \begin{algorithmic}[1]
                \State $k_\mac \leftarrow G(m)$
                \State $c \leftarrow \encrypt(\pk, m)$
                \State $t \leftarrow \sign(k_\mac, c)$
                \State \Return $(c, t)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decrypt^\rand_\etm(\sk, (c, t))$}\label{alg:rand-etm-decrypt}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
                \State $\hat{k}_\mac \leftarrow G(\hat{m})$
                \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
                    \State \Return $\bot$
                \EndIf
                \State \Return $\hat{m}$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Randomized encrypt-then-MAC}\label{fig:rand-etm}
\end{figure}

\section{Security of MAC}
Intuitively, if the unauthenticated ciphertext is tempered with, then the tag should change unpredictably. Since the MAC key is derived from the plaintext message, and the plaintext message is hidden behind the OW-CPA security of the input scheme, it is safe to assume that the adversary does not have information on the MAC key. If the adversary does not have the MAC key, then it will not be able to produce the correct tag. Therefore, the adversary cannot produce valid authenticated ciphertexts without knowing the corresponding plaintext in the first place.

However, formalizing this idea into a security proof is surprisingly difficult. The most popular methods of formal security proof is

\bibliographystyle{plain}
\bibliography{./references.bib}

\end{document}

### 2024-06-24 faster-fo-transform/etm-transform/main.tex
Reorganized etm-transform section

--- 2024-06-23 faster-fo-transform/randomized-etm/main.tex
@@
 
-\section{Introduction}
-In the modular FO transform\cite{hofheinz2017modular}, the authors constructed IND-CCA KEM using two steps.
-The first step is adding ciphertext integrity to a one-way secure encryption scheme (the \monospace{T} transformation).
-The second step is constructing the IND-CCA KEM out of the OW-PCVA secure encryption scheme (the \monospace{U} transformation).
-The \monospace{T} transformation uses two techniques for achieving ciphertext integrity, which the authors called ``\emph{de-randomization}'' and ``\emph{re-encryption}''.
+\section{Encrypt-then-MAC transformation}
+In this section, we introduce two ``encrypt-then-MAC'' ($\etm$ for short) transformations.
+For a given input public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$, the de-randomized $\etm$ transformation returns a a public-key encryption scheme $\pke^\notrand_\etm = (\keygen, \encrypt^\notrand_\etm, \decrypt^\notrand_\etm)$ such that $\encrypt^\notrand_\etm$ is deterministic.
+On the other hand, the randomized $\etm$ transformation returns a public-key encryption scheme $\pke_\etm = (\keygen, \encrypt_\etm, \decrypt_\etm)$ whose encryption routine preserves the randomization of the input encryption routine.
 
-Our current proposal is to replace re-encryption with a message authentication code (MAC), but we maintained the deterministic encryption routine.
-For clarity we will call this ``\emph{de-randomized encrypt-then-MAC}''.
-The algorithm is specified in figure \ref{fig:derand-etm}.
+The encryption routine of both transformations uses the plaintext $m \in \mathcal{M}_\pke$ to derive a MAC key $k_\mac$, then after using the input encryption routine $\encrypt$ to produce an unauthenticated ciphertext $c \in \mathcal{C}_\pke$, the MAC key is used to compute a tag $t$ on the unauthenticated ciphertext.
+The tuple $(c, t)$ is the authenticated ciphertext that the transformed encryption routine outputs.
+The decryption routine of both transformations first decrypt the unauthenticated ciphertext to recover the plaintext, then uses the decryption to derive a MAC key, which is used to verify the tag.
+
+Because the tag is computed on the unauthenticated ciphertext, if the unauthenticated ciphertext is tempered with, then the tag must be recomputed.
+Because of the one-way security of the underlying encryption scheme, it is safe to assume that no adversary can use the unauthenticated ciphertext to recover the message, which means that producing valid tags for new message constitutes a forgery attack.
+Therefore, mounting any meaningful chosen ciphertext attacks is at least as hard as producing a forgery against the underlying $\mac$ scheme.
+
+The transformed routines $(\encrypt_\etm, \decrypt_\etm)$ and $(\encrypt^\notrand_\etm, \decrypt^\notrand_\etm)$ are described in figure \ref{fig:rand-etm} and \ref{fig:derand-etm} respectively.
 
 \begin{figure}[h]
-  \begin{minipage}{.49\textwidth}  % Adjust width as needed
+    \center
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encrypt_\etm(\pk, m)$}\label{alg:rand-etm-encrypt}
+            \begin{algorithmic}[1]
+                \State $k_\mac \leftarrow G(m)$  % Need to distinguish from the other G
+                \State $c \leftarrow \encrypt(\pk, m)$
+                    \Comment{If $\encrypt$ is randomized, then $\encrypt_\etm$ is randomized}
+                \State $t \leftarrow \sign(k_\mac, c)$
+                \State \Return $(c, t)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decrypt_\etm(\sk, (c, t))$}\label{alg:rand-etm-decrypt}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
+                \State $\hat{k}_\mac \leftarrow G(\hat{m})$
+                \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
+                    \State \Return $\bot$
+                \EndIf
+                \State \Return $\hat{m}$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{
+        $\etm$ transformation.
+    }\label{fig:rand-etm}
+\end{figure}
+
+\begin{figure}[h]
+  \begin{minipage}{0.49\textwidth}  % Adjust width as needed
     \begin{algorithm}[H]
@@
       \begin{algorithmic}[1]
-        \State $(r, k_\mac) \leftarrow G(m)$
+        \State $k_\mac \leftarrow G(m)$
+        \State $r \leftarrow H(m)$
         \State $c \leftarrow \encrypt(\pk, m; r)$
@@
         \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-        \State $(\hat{r}, \hat{k}_\mac) \leftarrow G(\hat{m})$
+        \State $\hat{k}_\mac \leftarrow G(\hat{m})$
         \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
@@
 
-For the past week I have been thinking about removing the de-randomization, as well.
-This means that we no longer derive the pseudorandom coin from the message.
-If the input $\pke$ has a randomized encryption routine, then the transformed encrypt-then-mac scheme will also be randomized.
-We call this ``\emph{randomized encrypt-then-MAC}''.
-The algorithm is specified in figure \ref{fig:rand-etm}
-
-\begin{figure}[h]
-    \center
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\encrypt^\rand_\etm(\pk, m)$}\label{alg:rand-etm-encrypt}
-            \begin{algorithmic}[1]
-                \State $k_\mac \leftarrow G(m)$
-                \State $c \leftarrow \encrypt(\pk, m)$
-                \State $t \leftarrow \sign(k_\mac, c)$
-                \State \Return $(c, t)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\decrypt^\rand_\etm(\sk, (c, t))$}\label{alg:rand-etm-decrypt}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-                \State $\hat{k}_\mac \leftarrow G(\hat{m})$
-                \If{$\verify(\hat{k}_\mac, c, t) \neq 1$}
-                    \State \Return $\bot$
-                \EndIf
-                \State \Return $\hat{m}$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{Randomized encrypt-then-MAC}\label{fig:rand-etm}
-\end{figure}
-
-\section{Security of MAC}
-Intuitively, if the unauthenticated ciphertext is tempered with, then the tag should change unpredictably.
-Since the MAC key is derived from the plaintext message, and the plaintext message is hidden behind the OW-CPA security of the input scheme, it is safe to assume that the adversary does not have information on the MAC key.
-If the adversary does not have the MAC key, then it will not be able to produce the correct tag.
-Therefore, the adversary cannot produce valid authenticated ciphertexts without knowing the corresponding plaintext in the first place.
-
-However, formalizing this idea into a security proof is surprisingly difficult. The most popular methods of formal security proof is
-
-\bibliographystyle{plain}
-\bibliography{./references.bib}
+% \bibliographystyle{plain}
+% \bibliography{./references.bib}
 

### 2024-06-26 faster-fo-transform/etm-transform/main.tex
Listed all algorithms so far, but pco1 and cvo1 are incorrect

@@
 \newcommand{\mac}{\monospace{MAC}}
-\newcommand{\sign}{\monospace{MAC}}
-\newcommand{\verify}{\monospace{MAC.Verify}}
+\newcommand{\sign}{\monospace{Sign}}
+\newcommand{\verify}{\monospace{Verify}}
 \newcommand{\pk}{\monospace{pk}}
@@
 
-\title{Randomized encrypt-then-MAC}
+\title{The ``encrypt-then-MAC'' transformations}
 \author{Ganyu (Bruce) Xu (g66xu)}
@@
 
-\section{Encrypt-then-MAC transformation}
-In this section, we introduce two ``encrypt-then-MAC'' ($\etm$ for short) transformations.
-For a given input public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$, the de-randomized $\etm$ transformation returns a a public-key encryption scheme $\pke^\notrand_\etm = (\keygen, \encrypt^\notrand_\etm, \decrypt^\notrand_\etm)$ such that $\encrypt^\notrand_\etm$ is deterministic.
-On the other hand, the randomized $\etm$ transformation returns a public-key encryption scheme $\pke_\etm = (\keygen, \encrypt_\etm, \decrypt_\etm)$ whose encryption routine preserves the randomization of the input encryption routine.
+\section{Encrypt-then-MAC transformations}
+In this section we describe two ``encrypt-then-MAC'' transformations and discuss their security properties.
 
-The encryption routine of both transformations uses the plaintext $m \in \mathcal{M}_\pke$ to derive a MAC key $k_\mac$, then after using the input encryption routine $\encrypt$ to produce an unauthenticated ciphertext $c \in \mathcal{C}_\pke$, the MAC key is used to compute a tag $t$ on the unauthenticated ciphertext.
-The tuple $(c, t)$ is the authenticated ciphertext that the transformed encryption routine outputs.
-The decryption routine of both transformations first decrypt the unauthenticated ciphertext to recover the plaintext, then uses the decryption to derive a MAC key, which is used to verify the tag.
-
-Because the tag is computed on the unauthenticated ciphertext, if the unauthenticated ciphertext is tempered with, then the tag must be recomputed.
-Because of the one-way security of the underlying encryption scheme, it is safe to assume that no adversary can use the unauthenticated ciphertext to recover the message, which means that producing valid tags for new message constitutes a forgery attack.
-Therefore, mounting any meaningful chosen ciphertext attacks is at least as hard as producing a forgery against the underlying $\mac$ scheme.
-
-The transformed routines $(\encrypt_\etm, \decrypt_\etm)$ and $(\encrypt^\notrand_\etm, \decrypt^\notrand_\etm)$ are described in figure \ref{fig:rand-etm} and \ref{fig:derand-etm} respectively.
-
-\begin{figure}[h]
+\begin{figure}[H]
     \center
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
   \begin{minipage}{0.49\textwidth}  % Adjust width as needed
@@
 
+\begin{figure}[H]
+    \center
+
+    \begin{algorithm}[H]
+        \caption{Sequence of games}\label{alg:sequence-of-games}
+
+        \begin{algorithmic}
+            \State $(\pk, sk) \leftsample \keygen(1^\lambda)$
+            \State $m^\ast \leftsample \mathcal{M}_\pke$
+            \State $k^\ast_\mac \leftarrow H(m^\ast)$
+                \Comment{Game 0-2}
+            \State $k^\ast_\mac \leftsample \mathcal{K}_\mac$
+                \Comment{Game 3}
+            \State $c^\ast \leftsample \encrypt(\pk, m)$
+            \State $t^\ast \leftarrow \sign(k^\ast_\mac, c^\ast)$
+            \State $
+                \hat{m} \leftarrow 
+                \mathcal{A}^{
+                    \pco, \cvo, \mathcal{O}^H
+                }(1^\lambda, \pk, (c^\ast, t^\ast))
+            $
+                \Comment{Game 0}
+            \State $
+                \hat{m} \leftarrow 
+                \mathcal{A}^{
+                    \pco_1, \cvo, \mathcal{O}^H
+                }(1^\lambda, \pk, (c^\ast, t^\ast))
+            $
+                \Comment{Game 1}
+            \State $
+                \hat{m} \leftarrow 
+                \mathcal{A}^{
+                    \pco_1, \cvo_1, \mathcal{O}^H
+                }(1^\lambda, \pk, (c^\ast, t^\ast))
+            $
+                \Comment{Game 2-3}
+            \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+        \end{algorithmic}
+    \end{algorithm}
+
+    \caption{Sequence of games 0-3}\label{fig:sequence-of-games}
+\end{figure}
+
+\begin{figure}[H]
+    \center
+
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\pco(m, (c, t))$}\label{alg:pco-etm}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
+                \State $\hat{k}_\mac \leftarrow H(\hat{m})$
+                \If{$\sign(\hat{k}_\mac, c) \neq t$}
+                    \State \Return 0
+                \EndIf
+                \State \Return $\llbrack \hat{m} = m \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\pco_1(m, (c, t))$}\label{alg:pco1-etm}
+
+            \begin{algorithmic}[1]
+                \If{
+                    $\exists (\tilde{m}, \tilde{k}) \in \mathcal{O}^H :
+                        \tilde{m} = m \land \sign(\tilde{k}, c) = t
+                    $
+                }
+                    \State \Return $1$
+                \EndIf
+                \State \Return $0$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{True $\pco$ and simulated $\pco$}\label{fig:pco-and-pco1}
+\end{figure}
+
+\begin{figure}[H]
+    \center
+
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\cvo(c, t)$}\label{alg:cvo-etm}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
+                \State $\hat{k}_\mac \leftarrow H(\hat{m})$
+                \If{$\sign(\hat{k}_\mac, c) \neq t$}
+                    \State \Return 0
+                \EndIf
+                \State \Return 1
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\cvo_1(c, t)$}\label{alg:cvo1-etm}
+
+            \begin{algorithmic}[1]
+                \If{
+                    $\exists (\tilde{m}, \tilde{k}) \in \mathcal{O}^H :
+                        \sign(\tilde{k}, c) = t
+                    $
+                }
+                    \State \Return $1$
+                \EndIf
+                \State \Return $0$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{True $\cvo$ and simulated $\cvo_1$}\label{fig:cvo-and-cvo1}
+\end{figure}
+
 % \bibliographystyle{plain}

### 2024-06-27 faster-fo-transform/etm-transform/main.tex
plaintext-awareness is very hard to do

@@
 \newcommand{\cvo}{\monospace{CVO}}
+\newcommand{\lwe}{\monospace{LWE}}
+\newcommand{\cbd}{\monospace{CBD}}
+\newcommand{\oracle}[1]{\mathcal{O}^{#1}}
 \newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
@@
 \newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
+\newcommand{\round}[1]{\lceil #1 \rfloor}
 \newlength{\wdth}
@@
 %%%% TITLE %%%%%
-\maketitle
+% \maketitle
+\section{Encrypt-then-MAC transformations}
+Let $\pke(\keygen, \encrypt, \decrypt)$ be a probabilistic public-key encryption scheme defined over message space $\mathcal{M}_\pke$, ciphertext space $\mathcal{C}_\pke$, and coin space $\mathcal{R}_\pke$.
+Where the encryption routine is deterministic, we simply set the coin space to contain a single element $\mathcal{R} = \{r\}$.
+Let $\mac(\sign, \verify)$ be a message authentication code defined over key space $\mathcal{K}_\mac$.
+The message space of the $\mac$ should contain the ciphertext space of the $\pke$: $\mathcal{C}_\pke \subseteq \mathcal{M}_\mac$.
+Let $G: \mathcal{M}_\pke \rightarrow \mathcal{R}_\pke$ and $H: \mathcal{M}_\pke \rightarrow \mathcal{K}_\mac$ be hash functions.
 
-\section{Encrypt-then-MAC transformations}
-In this section we describe two ``encrypt-then-MAC'' transformations and discuss their security properties.
+The ``encrypt-then-MAC'' transformation $\pke_\etm(\keygen, \encrypt_\etm, \decrypt_\etm) = T_\etm(\pke, \mac, H)$ outputs a public-key encryption scheme where the key generation routine is identical to the input $\pke$'s key generation routine.
+The de-randomized ``encrypt-then-MAC'' transformation $\pke^\notrand_\etm(\keygen, \encrypt_\etm^\notrand, \decrypt^\notrand_\etm) = T^\notrand_\etm(\pke, \mac, G, H)$ similarly outputs a public-key encryption scheme.
+In both transformations, the key generation routine remains unchanged.
+The modified encryption and decryption routines are described in figure \ref{fig:rand-etm} and \ref{fig:derand-etm}.
 
@@
     \center
-    \begin{minipage}{0.49\textwidth}
+    \begin{minipage}{0.59\textwidth}
         \begin{algorithm}[H]
@@
             \begin{algorithmic}[1]
-                \State $k_\mac \leftarrow G(m)$  % Need to distinguish from the other G
-                \State $c \leftarrow \encrypt(\pk, m)$
+                \State $r \leftsample \mathcal{R}_\pke$
                     \Comment{If $\encrypt$ is randomized, then $\encrypt_\etm$ is randomized}
+                \State $k_\mac \leftarrow H(m)$  % Need to distinguish from the other G
+                \State $c \leftarrow \encrypt(\pk, m; r)$
                 \State $t \leftarrow \sign(k_\mac, c)$
@@
     \hfill
-    \begin{minipage}{0.49\textwidth}
+    \begin{minipage}{0.39\textwidth}
         \begin{algorithm}[H]
@@
     \caption{
-        $\etm$ transformation.
+        ``encrypt-then-MAC'' transformation
     }\label{fig:rand-etm}
@@
       \begin{algorithmic}[1]
-        \State $k_\mac \leftarrow G(m)$
-        \State $r \leftarrow H(m)$
+        \State $k_\mac \leftarrow H(m)$
+        \State $r \leftarrow G(m)$
         \State $c \leftarrow \encrypt(\pk, m; r)$
@@
   \end{minipage}
-  \caption{Derandomized encrypt-then-MAC}\label{fig:derand-etm}
+  \caption{
+    de-randomized ``encrypt-then-MAC'' transformation
+  }\label{fig:derand-etm}
 \end{figure}
+
+\emph{Intuition} the design goal of this transformation is to achieve plaintext awareness (PA).
+According to \cite{bellare1998relations}, PA implies resistance to adaptive chosen ciphertext attacks.
+For an adversary to produce a valid authenticated ciphertext $(c, t)$, it must be able to produce a valid tag $t$ for the unauthenticated ciphertext $c$ under the correct MAC key $k_\mac = H(\decrypt(\sk, c))$, which means that the adversary either knows $k_\mac$ or performs a forgery against the MAC.
+Since $k_\mac \leftarrow H(m)$ is derived from the plaintext $m$, under the random oracle model, I argue that knowing $k_\mac$ implies knowing $m$.
+In other words, if the adversary cannot perform forgery against the MAC nor break the one-way security of the encryption routine, then it cannot produce valid ciphertext for which it does not not the plaintext, thus rendering the advantage of a decryption oracle (or any ciphertext-processing oracles, such PCO or CVO) negligible.
+
+Let $A$ be an one-way adversary against the transformed scheme with access to a decryption oracle.
+I want to show that under the random oracle model, \textbf{we can construct a second adversary $B$ that uses the public key $\pk$ and the tapes of the hash oracles $\oracle{G}, \oracle{H}$ to simulate a decryption oracle for $A$} such that $A$ can only distinguish the true decryption oracle from the simulated decryption oracle with negligible advantage.
+Note that in \cite{hofheinz2017modular}, the first transformation asks for OW-PCVA security, but since the true PCO and CVO are both implemented using the decryption routine, it is easy to build simulated PCO and CVO using the simulated decryption oracle.
+
+Unfortunately, constructing a simulated decryption oracle is annoyingly elusive. Here are the things I have tried so far.
+
+\subsection{Identify plaintext using the coin and the key}
+With derandomized encrypt-then-MAC, if an authenticated  ciphertext $(c, t)$ is honestly generated, then there should be a pair of matching hash queries $(\tilde{m}, \tilde{r}) \in \oracle{G}$ and $(\tilde{m}, \tilde{k}) \in \oracle{H}$ such that $\encrypt(\pk, \tilde{m}; r) = c$ and $\verify(\tilde{k}, c, t) = 1$.
+See the algorithm below:
 
@@
     \center
-
     \begin{algorithm}[H]
-        \caption{Sequence of games}\label{alg:sequence-of-games}
-
-        \begin{algorithmic}
-            \State $(\pk, sk) \leftsample \keygen(1^\lambda)$
-            \State $m^\ast \leftsample \mathcal{M}_\pke$
-            \State $k^\ast_\mac \leftarrow H(m^\ast)$
-                \Comment{Game 0-2}
-            \State $k^\ast_\mac \leftsample \mathcal{K}_\mac$
-                \Comment{Game 3}
-            \State $c^\ast \leftsample \encrypt(\pk, m)$
-            \State $t^\ast \leftarrow \sign(k^\ast_\mac, c^\ast)$
-            \State $
-                \hat{m} \leftarrow 
-                \mathcal{A}^{
-                    \pco, \cvo, \mathcal{O}^H
-                }(1^\lambda, \pk, (c^\ast, t^\ast))
+        \caption{$\oracle{\decrypt}_\monospace{sim}(c, t)$ attempt 1}
+        \begin{algorithmic}[1]
+            \If{
+                $\exists (\tilde{m}, \tilde{r}) \in \oracle{G}
+                    , (\tilde{m}, \tilde{k}) \in \oracle{H} :
+                    \encrypt(\pk, \tilde{m}; \tilde{r}) = c
+                    \land \verify(\tilde{k}, c, t) = 1
                 $
-                \Comment{Game 0}
-            \State $
-                \hat{m} \leftarrow 
-                \mathcal{A}^{
-                    \pco_1, \cvo, \mathcal{O}^H
-                }(1^\lambda, \pk, (c^\ast, t^\ast))
-            $
-                \Comment{Game 1}
-            \State $
-                \hat{m} \leftarrow 
-                \mathcal{A}^{
-                    \pco_1, \cvo_1, \mathcal{O}^H
-                }(1^\lambda, \pk, (c^\ast, t^\ast))
-            $
-                \Comment{Game 2-3}
-            \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+            }
+                \State \Return $\tilde{m}$
+            \EndIf
+            \State \Return $\bot$
         \end{algorithmic}
     \end{algorithm}
-
-    \caption{Sequence of games 0-3}\label{fig:sequence-of-games}
+    \caption{Simulate decryption oracle using coin and MAC key}
+    \label{fig:sim-decrypt-}
 \end{figure}
 
+This simulated decryption oracle can be efficiently distinguished from the true decryption oracle if the ciphertext can be predictably perturbed while still decrpyting to the same plaintext.
+This ciphertext perturbation is possible in most LWE schemes: one can add a small amount of noise to the unauthenticated ciphertext, and the resulting ``new ciphertext'' will still decrypt to the same plaintext.
+To distinguish simulation from true decryption oracle:
+
+\begin{enumerate}
+    \item Generate an honest plaintext-ciphertext pair $(m, (c, t))$. In this process, store the coin $r \leftarrow G(m)$ and MAC key $k_\mac \leftarrow H(m)$
+    \item Perturb the unauthenticated ciphertext. Denote the result by $c^\prime$.
+    \item Compute $t^\prime = \sign(k_\mac, c^\prime)$
+    \item Submit $(c^\prime, t^\prime)$ as a decryption query. The true decryption oracle will return $m$ (which is correct), but the simulation will reject this ciphertext (which is too strict).
+\end{enumerate}
+
+A similar attempt to simulate decryption oracle can be made on the (possibly randomized) ``encrypt-then-MAC'' scheme, as well.
+We need to replace the hash oracle $\oracle{G}$ with a ``randomness oracle'' $\mathcal{O}^\mathcal{R}$: when adversary $A$ wants to sample $r \leftsample \mathcal{R}$, $A$ needs to query this randomness oracle, which will return a random value at this query, and store the query output on a tape:
+
 \begin{figure}[H]
-    \center
-
-    \begin{minipage}{0.49\textwidth}
     \begin{algorithm}[H]
-            \caption{$\pco(m, (c, t))$}\label{alg:pco-etm}
+        \caption{$\mathcal{O}^\decrypt_\monospace{sim}(c, t)$ attempt 1.1}
         \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-                \State $\hat{k}_\mac \leftarrow H(\hat{m})$
-                \If{$\sign(\hat{k}_\mac, c) \neq t$}
-                    \State \Return 0
+            \If{
+                $\exists \tilde{r} \in \mathcal{O}^\mathcal{R}
+                    , (\tilde{m}, \tilde{k}) \in \oracle{H} :
+                    \encrypt(\pk, \tilde{m}; \tilde{r}) = c
+                    \land \verify(\tilde{k}, c, t) = 1$
+            }
+                \State \Return $\tilde{m}$
             \EndIf
-                \State \Return $\llbrack \hat{m} = m \rrbrack$
+            \State \Return $\bot$ 
         \end{algorithmic}
     \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.49\textwidth}
+    \caption{Simulate decryption oracle in randomized $\etm$}
+\end{figure}
+
+\subsection{Identify plaintext using only the key}
+My second attempt at recovering the plaintext from the hash oracles uses only the MAC key:
+
+\begin{figure}[H]
     \begin{algorithm}[H]
-            \caption{$\pco_1(m, (c, t))$}\label{alg:pco1-etm}
+        \caption{$\mathcal{O}^\decrypt_\monospace{sim}(c, t)$ attempt 2}
+        \begin{algorithmic}[1]
+            \If{$\exists (\tilde{m}, \tilde{k}) \in \oracle{H} : 
+                \verify(\tilde{k}, c, t) = 1
+            $}
+                \State \Return $\tilde{m}$
+            \EndIf
+            \State \Return $\bot$
+        \end{algorithmic}
+    \end{algorithm}
+\end{figure}
 
+This allows the adversary to sign unauthenticated ciphertext produced from perturbing honestly generated unauthenticated ciphertext, but then creates the problem that the adversary can now sign unauthenticated ciphertext with MAC keys derived from arbitrary plaintext:
+
+\begin{enumerate}
+    \item Sample a random plaintext and run the encryption routine. This creates an honest plaintext-ciphertext pair $(m, (c, t))$
+    \item Sample a second, uncorrelated plaintext $m^\prime$ and derive the corresponding MAC key $k_\mac^\prime = H(m^\prime)$
+    \item Sign $c$ with $k^\prime_\mac$: $t^\prime \leftarrow \sign(k^\prime_\mac, c)$
+    \item Submit $(c, t^\prime)$ for a decryption query. The true decryption oracle will reject this query, but the simulation will return $m^\prime$, which is just wrong.
+\end{enumerate}
+
+\subsection{Identify matching plaintext-ciphertext using only the public key}
+Let $\monospace{match}(\pk, m, c)$ be an efficient algorithm that takes the public key $\pk$ and a plaintext-ciphertext pair $(m, c)$ as input.
+$\monospace{match}$ returns 1 if $m$ is the decryption of $c$ under the keypair $(\pk, \sk)$ and 0 otherwise.
+
+If such an algorithm exists, then it is easy to implement the simulated decryption oracle.
+
+\begin{figure}[H]
+    \begin{algorithm}[H]
+        \caption{$\mathcal{O}^\decrypt_\monospace{sim}(c, t)$ if \monospace{match} exists}
         \begin{algorithmic}[1]
             \If{
-                    $\exists (\tilde{m}, \tilde{k}) \in \mathcal{O}^H :
-                        \tilde{m} = m \land \sign(\tilde{k}, c) = t
+                $\exists (\tilde{m}, \tilde{k}) \in \oracle{H} 
+                : \monospace{match}(\tilde{m}, c) = 1
+                \land \verify(\tilde{k}, c, t)
                 $
             }
-                    \State \Return $1$
+                \State \Return $\tilde{m}$
             \EndIf
-                \State \Return $0$
+            \State \Return $\bot$
         \end{algorithmic}
     \end{algorithm}
-    \end{minipage}
-
-    \caption{True $\pco$ and simulated $\pco$}\label{fig:pco-and-pco1}
 \end{figure}
 
-\begin{figure}[H]
-    \center
+On the other hand, if such an algorithm exists, then the encryption scheme is necessarily not semantically secure.
+Since Kyber begins with a semantically secure encryption scheme before transforming into a key encapsulation mechanism, such an algorithm must not exist, which means that we need to build the capabilities of ``identifying matching plaintext-ciphertext pair'' into the $\etm$ transformation(s).
 
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\cvo(c, t)$}\label{alg:cvo-etm}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-                \State $\hat{k}_\mac \leftarrow H(\hat{m})$
-                \If{$\sign(\hat{k}_\mac, c) \neq t$}
-                    \State \Return 0
-                \EndIf
-                \State \Return 1
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\cvo_1(c, t)$}\label{alg:cvo1-etm}
+\subsection{Questioning the premise}
+It is of course possible the reason why I cannot come up with a security reduction is that $\etm$ transformation is insecure, but so far I also could not come up with any meaningful attack that breaks the one-way security of the transformed schemes.
 
-            \begin{algorithmic}[1]
-                \If{
-                    $\exists (\tilde{m}, \tilde{k}) \in \mathcal{O}^H :
-                        \sign(\tilde{k}, c) = t
-                    $
-                }
-                    \State \Return $1$
-                \EndIf
-                \State \Return $0$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
+Note that the two distinguishing algorithms in the previous sections only reveal flaws in the simulated decryption oracle.
+They do not lead to meaningful weakness (or meaningful advantages) of the $\etm$ transformation under chosen ciphertext attacks
 
-    \caption{True $\cvo$ and simulated $\cvo_1$}\label{fig:cvo-and-cvo1}
-\end{figure}
-
-% \bibliographystyle{plain}
-% \bibliography{./references.bib}
+\bibliographystyle{alpha}
+\bibliography{./references.bib}
 

### 2024-06-29 faster-fo-transform/etm-transform/main.tex
Trying a new appriach

@@
 
+\section{Plaintext Awareness: a new approach}
+A new approach I can think of is as follows:
+
+\begin{enumerate}
+    \item prove that $\etm$ transformed schemes are plaintext aware
+    \item prove that if a scheme is plaintext aware, then it is OW-CCA2
+\end{enumerate}
+
+Both steps are non-trivial, though the second step can be heavily borrowed from \cite{bellare1998relations}.
+
+\subsection{$\etm$ is plaintext aware}
+
+\subsection{Plaintext-awareness implies OW-CCA2}
+
 \bibliographystyle{alpha}

### 2024-06-29 faster-fo-transform/etm-transform/main.tex
Defined plaintext awareness

--- 2024-06-26 faster-fo-transform/etm-transform/main.tex
@@
 %%%% TITLE %%%%%
-\maketitle
+% \maketitle
+
+\section{Preliminaries}
+\subsection{Plaintext awareness}
+\emph{Plaintext awareness (PA)}\cite{bellare1995optimal,bellare1998relations} describes the idea that no efficient algorithm can produce a valid ciphertext without knowing the corresponding decryption.
+
+Let $\pke = (\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme.
+Let $H$ be a hash function.
+Let $\mathcal{E}^H_\pk$ be an encryption oracle that takes no argument and returns valid ciphertexts when queried.
+Under the random oracle model, hash queries made to the oracle $\mathcal{O}^H$ can be logged to a tape $\mathcal{L}^H = \{(h_i, H(h_i))\}$, and ciphertexts obtained from the encryption oracle are also logged to a separate tape $C = \{c : c \leftarrow \mathcal{E}^H_\pk(\cdot)\}$.
+
+A plaintext-awareness adversary $B$ is a probabilistic algorithm that is given some public key and access to the two oracles, then output some ciphertext $c$.
+Note that the encryption oracle here is not redundant because obtaining ciphertexts from the encryption oracle will not log any corresponding hash queries in the hash oracle.
+This models a PA adversary's ability to obtain valid ciphertexts without running the encryption routine, such as by eavesdropping.
+The PA adversary $B$ outputs a ciphertext $c$ and the transcript $\mathcal{L}^H, C$ of its interactions with the oracles.
+
+Let $K$ be some algorithm that outputs a decryption of $c$ using the corresponding transcript $\mathcal{L}^H, C$.
+We restrict $c \not \in C$ to prevent trivially turning $K$ into a decryption oracle.
+The plaintext awareness game is defined in figure \ref{fig:pa-game}.
+
+\begin{figure}[H]
+    \begin{algorithm}[H]
+        \caption{Plaintext awareness game}\label{alg:pa-game}
+        \begin{algorithmic}[1]
+            \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+            \State $(\mathcal{L}^H, C, c) \leftarrow B^{H, \mathcal{E}^H_\pk}(1^\lambda, \pk)$
+            \State $m \leftarrow K(1^\lambda, \pk, \mathcal{L}^H, C, c)$
+            \State \Return $\llbrack c \not\in C \land m = D(\sk, c) \rrbrack$
+        \end{algorithmic}
+    \end{algorithm}
+    \caption{The plaintext awareness game}\label{fig:pa-game}
+\end{figure}
+
+\begin{definition}
+    A public-key encryption scheme is \emph{plaintext-aware} if there exists an efficient $K$ such that for all efficient PA adversaries $B$, the probability of $K$ failing to extract the correct decryption is negligible.
+
+    \begin{equation*}
+        P[\monospace{PA(K, B)} \neq 1] \leq \monospace{negl}(\lambda)
+    \end{equation*}
+\end{definition}
+
+Note that our definition of plaintext awareness deviates from \cite{bellare1998relations} in that we do not require the $\pke$ to also be IND-CPA secure.
+This is because in constructing a key encapsulation mechanism using the Fujisaki-Okamoto transformation, we only require the input $\pke$ to be one-way secure, though under a stronger attack model.
+Restricting the definition to only the plaintext-awareness game allows us to combine PA with other security definition, such as OW-CPA, to prove stronger security result.
 
 \section{Encrypt-then-MAC transformations}
-In this section we describe two ``encrypt-then-MAC'' transformations and discuss their security properties.
+Let $\pke(\keygen, \encrypt, \decrypt)$ be a probabilistic public-key encryption scheme defined over message space $\mathcal{M}_\pke$, ciphertext space $\mathcal{C}_\pke$, and coin space $\mathcal{R}_\pke$.
+Where the encryption routine is deterministic, we simply set the coin space to contain a single element $\mathcal{R} = \{r\}$.
+Let $\mac(\sign, \verify)$ be a message authentication code defined over key space $\mathcal{K}_\mac$.
+The message space of the $\mac$ should contain the ciphertext space of the $\pke$: $\mathcal{C}_\pke \subseteq \mathcal{M}_\mac$.
+Let $G: \mathcal{M}_\pke \rightarrow \mathcal{R}_\pke$ and $H: \mathcal{M}_\pke \rightarrow \mathcal{K}_\mac$ be hash functions.
+
+The ``encrypt-then-MAC'' transformation $\pke_\etm(\keygen, \encrypt_\etm, \decrypt_\etm) = T_\etm(\pke, \mac, H)$ outputs a public-key encryption scheme where the key generation routine is identical to the input $\pke$'s key generation routine.
+The de-randomized ``encrypt-then-MAC'' transformation $\pke^\notrand_\etm(\keygen, \encrypt_\etm^\notrand, \decrypt^\notrand_\etm) = T^\notrand_\etm(\pke, \mac, G, H)$ similarly outputs a public-key encryption scheme.
+In both transformations, the key generation routine remains unchanged.
+The modified encryption and decryption routines are described in figure \ref{fig:rand-etm} and \ref{fig:derand-etm}.
 
@@
   \end{minipage}
-  \caption{Derandomized encrypt-then-MAC}\label{fig:derand-etm}
+  \caption{
+    de-randomized ``encrypt-then-MAC'' transformation
+  }\label{fig:derand-etm}
 \end{figure}
 
-\begin{figure}[H]
-    \center
-
-    \begin{algorithm}[H]
-        \caption{Sequence of games}\label{alg:sequence-of-games}
-
-        \begin{algorithmic}
-            \State $(\pk, sk) \leftsample \keygen(1^\lambda)$
-            \State $m^\ast \leftsample \mathcal{M}_\pke$
-            \State $k^\ast_\mac \leftarrow H(m^\ast)$
-                \Comment{Game 0-2}
-            \State $k^\ast_\mac \leftsample \mathcal{K}_\mac$
-                \Comment{Game 3}
-            \State $c^\ast \leftsample \encrypt(\pk, m)$
-            \State $t^\ast \leftarrow \sign(k^\ast_\mac, c^\ast)$
-            \State $
-                \hat{m} \leftarrow 
-                \mathcal{A}^{
-                    \pco, \cvo, \mathcal{O}^H
-                }(1^\lambda, \pk, (c^\ast, t^\ast))
-            $
-                \Comment{Game 0}
-            \State $
-                \hat{m} \leftarrow 
-                \mathcal{A}^{
-                    \pco_1, \cvo, \mathcal{O}^H
-                }(1^\lambda, \pk, (c^\ast, t^\ast))
-            $
-                \Comment{Game 1}
-            \State $
-                \hat{m} \leftarrow 
-                \mathcal{A}^{
-                    \pco_1, \cvo_1, \mathcal{O}^H
-                }(1^\lambda, \pk, (c^\ast, t^\ast))
-            $
-                \Comment{Game 2-3}
-            \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
-        \end{algorithmic}
-    \end{algorithm}
-
-    \caption{Sequence of games 0-3}\label{fig:sequence-of-games}
-\end{figure}
-
-\begin{figure}[H]
-    \center
-
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\pco(m, (c, t))$}\label{alg:pco-etm}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-                \State $\hat{k}_\mac \leftarrow H(\hat{m})$
-                \If{$\sign(\hat{k}_\mac, c) \neq t$}
-                    \State \Return 0
-                \EndIf
-                \State \Return $\llbrack \hat{m} = m \rrbrack$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\pco_1(m, (c, t))$}\label{alg:pco1-etm}
-
-            \begin{algorithmic}[1]
-                \If{
-                    $\exists (\tilde{m}, \tilde{k}) \in \mathcal{O}^H :
-                        \tilde{m} = m \land \sign(\tilde{k}, c) = t
-                    $
-                }
-                    \State \Return $1$
-                \EndIf
-                \State \Return $0$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-
-    \caption{True $\pco$ and simulated $\pco$}\label{fig:pco-and-pco1}
-\end{figure}
-
-\begin{figure}[H]
-    \center
-
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\cvo(c, t)$}\label{alg:cvo-etm}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-                \State $\hat{k}_\mac \leftarrow H(\hat{m})$
-                \If{$\sign(\hat{k}_\mac, c) \neq t$}
-                    \State \Return 0
-                \EndIf
-                \State \Return 1
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\cvo_1(c, t)$}\label{alg:cvo1-etm}
-
-            \begin{algorithmic}[1]
-                \If{
-                    $\exists (\tilde{m}, \tilde{k}) \in \mathcal{O}^H :
-                        \sign(\tilde{k}, c) = t
-                    $
-                }
-                    \State \Return $1$
-                \EndIf
-                \State \Return $0$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-
-    \caption{True $\cvo$ and simulated $\cvo_1$}\label{fig:cvo-and-cvo1}
-\end{figure}
-
-% \bibliographystyle{plain}
-% \bibliography{./references.bib}
+\bibliographystyle{alpha}
+\bibliography{./references.bib}
 

### 2024-06-30 faster-fo-transform/etm-transform/main.tex
Laid out proof strategy

@@
 
+\begin{theorem}\label{thm:etm-is-ow-cca}
+    If the $\mac$ is one-time unforgeable and $\pke$ is \monospace{OW-CPA} secure, then $\etm$ and $\etm^\notrand$ are both \monospace{OW-CCA} secure
+\end{theorem}
+
+We will prove theorem \ref{thm:etm-is-ow-cca} in two steps.
+First we show that $\etm$ and $\etm^\notrand$ are both plaintext-aware in section \ref{sec:etm-is-pa}, then we show that plaintext awareness and \monospace{OW-CPA} security imply \monospace{OW-CCA} security in section \ref{sec:pa-and-owcpa-imply-owcca}.
+
+The KEM transformations from \cite{hofheinz2017modular} require the input $\pke$ to be OW-PCVA secure, although OW-PCVA security is implied by OW-CCA security
+
+\begin{theorem}\label{thm:ow-cca-implies-ow-pcva}
+    For every \monospace{OW-PCVA} adversary $\mathcal{A}_\monospace{PCVA}$ against some $\pke$, there exists an \monospace{OW-CCA} adversary $\mathcal{A}_\monospace{CCA}$ with identical advantage:
+
+    \begin{equation*}
+        \epsilon_\monospace{OW-PCVA} = \epsilon_\monospace{OW-CCA}
+    \end{equation*}
+\end{theorem}
+
+\begin{proof}
+    The \monospace{OW-CCA} adversary can perfectly simulate both the PCO and the CVO by replacing ``running decryption routine'' with ``querying the decryption oracle''.
+When the \monospace{OW-PCVA} adversary halts, the \monospace{OW-CCA} adversary passes \monospace{OW-PCVA} adversary's output as its own.
+Thus, \monospace{OW-CCA} adversary wins if and only if \monospace{OW-PCVA} adversary wins.
+\end{proof}
+
+\subsection{$\etm$ is plaintext aware}\label{sec:etm-is-pa}
+
+\subsection{Plaintext awareness and OW-CPA imply OW-CCA}\label{sec:pa-and-owcpa-imply-owcca}
+
 \bibliographystyle{alpha}

### 2024-07-02 faster-fo-transform/etm-transform/main.tex
Proof that OW-CPA and PA implies OW-CCA

@@
 \subsection{Plaintext awareness and OW-CPA imply OW-CCA}\label{sec:pa-and-owcpa-imply-owcca}
+In this section we show that if a $\pke$ is both plaintext aware and \monospace{OW-CPA} secure, then it is \monospace{OW-CCA} secure.
+
+\begin{theorem}\label{thm:pa-implies-cca}
+    For every \monospace{OW-CCA} adversary $\mathcal{A}_\monospace{OW-CCA}$ with advantage $\epsilon_\monospace{OW-CCA}$, there exists an $\epsilon_K$ knowledge extractor $K$ and an \monospace{OW-CPA} adversary with advantage $\epsilon_\monospace{OW-CPA}$ such that
+
+    \begin{equation*}
+        \epsilon_\monospace{OW-CCA} \leq \epsilon_\monospace{OW-CPA} + q_D \cdot \epsilon_K
+    \end{equation*}
+\end{theorem}
+
+\begin{proof}
+    We will prove theorem \ref{thm:pa-implies-cca} by constructing an \monospace{OW-CPA} adversary $\mathcal{A}_\monospace{OW-CPA}$ that runs the \monospace{OW-CCA} adversary $\mathcal{A}_\monospace{OW-CCA}$ as a sub-routine.
+$\mathcal{A}_\monospace{OW-CPA}$ will answer $\mathcal{A}_\monospace{OW-CCA}$'s decryption query using the knowledge extractor.
+
+    After receiving the security parameter $1^\lambda$ and public key $\pk$, $\mathcal{A}_\monospace{OW-CPA}$ initializes the hash tape $\mathcal{L}^H = \{\}$ to be an empty list, the runs $\mathcal{A}_\monospace{OW-CCA}$ until the sub-routine is waiting for the challenge ciphertext.
+During this phase:
+
+    \begin{enumerate}
+        \item When $\mathcal{A}_\monospace{OW-CCA}$ makes a hash query $h$, $\mathcal{A}_\monospace{OW-CPA}$ queries $\mathcal{O}^H$ to obtain the hash value $H(h)$, appends $(h, H(h))$ to $\mathcal{L}^H$, then answers with $H(h)$
+        \item When $\mathcal{A}_\monospace{OW-CCA}$ makes a decryption query $c$, $\mathcal{A}_\monospace{OW-CPA}$ runs the knowledge extractor $m \leftarrow K(\mathcal{L}^H, C = \{\}, c)$, then answers with $m$
+    \end{enumerate}
+
+    After receiving $c^\ast$, $\mathcal{A}_\monospace{OW-CPA}$ passes $c^\ast$ to $\mathcal{A}_\monospace{OW-CCA}$, then continues simulating the \monospace{OW-CCA} game for the sub-routine.
+
+    \begin{enumerate}
+        \item Hash queries are answered in the same way as in the first phase
+        \item When $\mathcal{A}_\monospace{OW-CCA}$ makes a decryption query $c$: \begin{itemize}
+            \item If $c = c^\ast$, answer with $\bot$
+            \item If $c \neq c^\ast$, run the knowledge extractor $m \leftarrow K(\mathcal{L}^H, C = \{c^\ast\}, c)$, then answer with $m$
+        \end{itemize}
+    \end{enumerate}
+
+    We borrow the argument from \cite{bellare1998relations} to analyze how well $\mathcal{A}_\monospace{OW-CPA}$ simulates the decryption oracle.
+Let $q_D$ be the total number of decryption queries made throughout both phases and let $q_1$ denote the number of decryption queries made in phase 1.
+We can thus label the decryption queries by $(c_1, c_2, \ldots, c_{q_1}, c_{q_1 + 1}, \ldots, c_{q_D})$.
+Each decryption query corresponds to a plaintext awareness adversary $B_i$ playing an instance of the plaintext awareness game using $\mathcal{A}_\monospace{OW-CCA}$ as a sub-routine.
+Each $B_i$ answers hash queries from $\mathcal{A}_\monospace{OW-CCA}$ in the same way as described above.
+For $j < i$, $B_i$ answers the decryption query $c_j$ using what $K$ returns on $B_j$'s output.
+When $\mathcal{A}_\monospace{OW-CCA}$ makes the i-th decryption query $c_i$, $B_i$ outputs $c_i$ and halts.
+For $1 \leq j \leq q_1$, there is no ciphertext obtained without running the encryption routine, so $K$ is run with $C = \{\}$.
+For $q_1 < j \leq q_D$, $\mathcal{A}_\monospace{OW-CCA}$ has access to $c^\ast$, which is obtained without running the encryption routine, so $K$ is run with $C = \{c^\ast\}$.
+
+    The simulated \monospace{OW-CCA} game deviates from the true \monospace{OW-CCA} game if and only if the knowledge extractor returns an incorrect decryption at one or more decryption queries.
+Since the probability of any one of decryption query being incorrect is bounded by $\epsilon_K$, the probability that at least one decryption query being incorrect is bounded by $q_D \cdot \epsilon_K$:
+
+    \begin{equation*}
+        \epsilon_\monospace{OW-CCA} - \epsilon_\monospace{OW-CPA}
+        \leq q_D \cdot \epsilon_K
+    \end{equation*}
+\end{proof}
 

### 2024-07-02 faster-fo-transform/etm-transform/main.tex
Open questions on proving EtM to be plaintext aware

@@
 \newcommand{\notrand}{{\strike{\$}}}
+\newcommand{\negl}[1]{\operatorname{negl}\left(#1\right)}
 \newcommand{\rand}{\$}
@@
 \subsection{$\etm$ is plaintext aware}\label{sec:etm-is-pa}
+I want to prove that $\etm$ is plaintext aware by showing that if $\etm$ is not plaintext aware, then the underlying $\mac$ is broken.
+For clearer notations we denote unauthenticated ciphertexts by $\sigma \in \mathcal{C}_\pke$ and authenticated ciphertexts by $c = (\sigma, t)$.
+
+If there exists a PA adversary $B$ with output $(\mathcal{L}^{G, H, \mathcal{R}}, C, c=(\sigma, t)) \leftarrow B(\cdot)$ such that for all knowledge extractor $K$
+
+\begin{equation*}
+    P \left\lbrack
+    K(\mathcal{L}^{G, H, \mathcal{R}}, C, c=(\sigma, t)) \neq \decrypt_\etm(\sk, c)
+    \right\rbrack > \negl{\lambda}
+\end{equation*}
+
+If $B$ produces a valid ciphertext $c = (\sigma, t)$, then $t$ must be a valid tag of $\sigma$ under $k_\mac = H(m)$ where $m = \decrypt(\sk, \sigma)$.
+
+\begin{enumerate}
+    \item How can I express that ``If $B$ knows $m$, then some knowledge extractor can recover $m$''?
+    \item If $B$ does not know $m$, then $k_\mac$ is unknown and uniformly random. How can I construct a forgery adversary using $B$?
+    \item How do I account for the scenario in which $c$ is invalid?
+\end{enumerate}
 

### 2024-07-03 faster-fo-transform/cca-transform-survey/main.tex
correctness and ow-atk game

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Custom commands
\usepackage{mystyle}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    A survey of generic IND-CCA2 transformations
}
\author{
    Ganyu Xu
}
% Leave the date field empty to display the date of compilation
% \date{}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{Preliminaries}
\subsection{Public-key encryption schemes}
A public key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$ is a collection of three routines. $\keygen(1^\lambda)$ takes the security parameteras input and returns a keypair $(\pk, \sk)$. $\encrypt(\pk, m)$ takes some public key and some plaintext message $m \in \mathcal{M}_\pke$ and output a ciphertext $c \in \mathcal{C}_\pke$. Where the encryption routine is probabilistic, we model the randomness using a coin $r \in \mathcal{R}$ such that $E(\pk, m; r)$ is deterministic with an explicit $r$. Finally, $\decrypt(\sk, c)$ uses the secret key to decrypt the ciphertext.

\subsubsection{Correctness}
Conventionally we require a $\pke$ to be perfectly correct. This means that for all possible key pairs $(\pk, \sk)$ and plaintexts $m$, the decryption routine always correctly inverts the encryption routine: $\decrypt(\sk, \encrypt(\pk, m)) = m$.

Where perfect correctness is not achieved, such as with most lattice-based encryption schemes, we need to account for the possiblity that decryption can fail. If under some keypair $(\pk, \sk)$, a plaintext-ciphertext pair $(m, c)$ is such that $c \leftsample \encrypt(\pk, m)$ is obtained from encrypting $m$ (probabilistically) but $m \neq \decrypt(\sk, c)$, we call it a decryption failure. For probabilistic encryption routines where the coin is uniformly sampled from the coin space, we can quantify the probability that $m$ triggers a decryption failure. From here, we can take the distribution of all keypairs and quantify the ``correctness'' of a (possibly imperfectly correct) encryption scheme. The following definition of $\delta$-correctness is directly taken from \cite{bos2018crystals}.

\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
    A probabilistic public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$ is $\delta$-correct if the expected maximal probability of decryption failure taken across the distribution of keypairs is at most $\delta$:

    \begin{equation*}
        E\left\lbrack\max_{m \in \mathcal{M}} P\left\lbrack
            \decrypt(\sk, \encrypt(\pk, m))
        \right\rbrack
        \right\rbrack \leq \delta
    \end{equation*}

    where the expectation is taken over the distribution of keypairs and the probability is taken over the distribution of coins.
\end{definition}

\cite{hofheinz2017modular} also defined an adversarial game in which the adversary's goal is to find some plaintext message to trigger a decryption failure. This adversarial game meaningfully models the real-world scenario in which decryption failure can reveal information about the secret key. Notice that when evaluating the win condition, the coin is uniformly random instead of being chosen by the adversary.

\begin{figure}[H]
    \center

    \begin{minipage}{0.5\textwidth}
        \begin{algorithm}[H]
            \caption{\monospace{CORS}}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $m \leftsample A_\monospace{CORS}(1^\lambda, \pk, \sk)$
                \State \Return $\llbrack \decrypt(\sk, \encrypt(\pk, m)) \neq m \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{The correctness game \monospace{CORS}}\label{fig:cors-game}
\end{figure}

The definition of $\delta$-correctness sets an explicit upper bound on the probability that any plaintext triggers decryption failure. This means that even if the \monospace{CORS} adversary actually finds the message $m$ that is the most likely to trigger decryption failure, the probability of winning the \monospace{CORS} game is still upper-bounded by $\delta$.

\begin{lemma}
    If $\pke$ is $\delta$-correct, then for all \monospace{CORS} adversaries $A$, even computationally unbounded ones, the probability of winning the \monospace{CORS} game is at most $\delta$
\end{lemma}

The values of $\delta$ for Kyber are taken directly from \cite{avanzi2019crystals}

\begin{table}[H]
    \center
    \begin{tabular}{c|c}
        security level & $\delta$ \\
        \hline
        Kyber512 & $2^{-139}$ \\
        \hline
        Kyber768 & $2^{-164}$ \\
        \hline
        Kyber1024 & $2^{-174}$ \\
    \end{tabular}
    \caption{Concrete $\delta$ for Kyber}\label{tbl:kyber-delta-vals}
\end{table}

\subsubsection{Security}
An one-way adversary $A = (A_1, A_2)$ consists of two sub-routines $A_1$ and $A_2$. $s \leftsample A_1^{\mathcal{O}}(1^\lambda, \pk)$ takes the security parameter, some public-key, access to some oracle(s) $\mathcal{O}$, and outputs some intermediate state $s$. $\hat{m} \leftarrow A_2^\mathcal{O}(1^\lambda, \pk, c^\ast)$ resumes from the output of $A_1$ and takes some challenge ciphertext, then tries to guess the correspodning decryption.

\begin{figure}[H]
    \begin{algorithm}[H]
        \caption{\monospace{OW-ATK} game}\label{alg:ow-atk-game}
        \begin{algorithmic}[1]
            \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
            \State $s \leftsample A_1^\mathcal{O}(1^\lambda, \pk)$
            \State $m^\ast \leftsample \mathcal{M}$
            \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
            \State $\hat{m} \leftarrow A_2^\mathcal{O}(1^\lambda, \pk, s, c^\ast)$
            \State \Return $\llbrack \hat{m} = m^\ast\rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \caption{The one-way security game}\label{fig:ow-atk-game}
\end{figure}

\section{Modular Fujisaki-Okamoto transformation}
The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} and its KEM variations \cite{hofheinz2017modular}.

\bibliographystyle{alpha}
\bibliography{./references.bib}

\end{document}

### 2024-07-04 faster-fo-transform/cca-transform-survey/main.tex
Described security games

@@
 An one-way adversary $A = (A_1, A_2)$ consists of two sub-routines $A_1$ and $A_2$.
-$s \leftsample A_1^{\mathcal{O}}(1^\lambda, \pk)$ takes the security parameter, some public-key, access to some oracle(s) $\mathcal{O}$, and outputs some intermediate state $s$.
-$\hat{m} \leftarrow A_2^\mathcal{O}(1^\lambda, \pk, c^\ast)$ resumes from the output of $A_1$ and takes some challenge ciphertext, then tries to guess the correspodning decryption.
+$s \leftsample A_1^{\mathcal{O}_1}(1^\lambda, \pk)$ takes the security parameter, some public-key, access to some oracle(s) $\mathcal{O}_1$, and outputs some intermediate state $s$.
+$\hat{m} \leftarrow A_2^{\mathcal{O}_2}(1^\lambda, \pk, c^\ast)$ resumes from the output of $A_1$ and takes some challenge ciphertext, then tries to guess the corresponding decryption.
+
+The advantage $\monospace{Adv}_\monospace{OW-ATK}(A)$ of an \monospace{OW-ATK} adversary is the probability that its guess is correct.
+
+\begin{definition}
+    A $\pke$ is \monospace{OW-ATK} secure if for all efficient adversaries $A$, the advantage in the \monospace{OW-ATK} game is neligigble with respect to the security parameter:
+
+    \begin{equation*}
+        \monospace{Adv}_\monospace{OW-ATK}(A) \leq \operatorname{negl}(\lambda)
+    \end{equation*}
+\end{definition}
+
+An \textbf{indistinguishability} adversary $A = (A_1, A_2)$ similarly consists of two sub-routines.
+The first sub-routine adversarially chooses two distinct plaintext messages, and the second sub-routine tries to distinguish which of the two plaintext messages is the decryptino of the challenge encryption.
+The advantage of an indistinguishability adversary is defined by $\monospace{Adv}_\monospace{IND-ATK}(A) = P[\hat{b} = b] - \frac{1}{2}$
+
+\begin{definition}
+    A $\pke$ is \monospace{IND-ATK} secure if for all efficient adversaries $A$, the advantage in the indistinguishability game is negligible with respect to the security parameter
+
+    \begin{equation*}
+        \monospace{Adv}_\monospace{IND-ATK}(A) \leq \operatorname{negl}(\lambda)
+    \end{equation*}
+\end{definition}
+
+The security games are described in details in figure \ref{fig:sec-games}
 
 \begin{figure}[H]
+    \center
+    \begin{minipage}{0.49\textwidth}
         \begin{algorithm}[H]
@@
                 \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-            \State $s \leftsample A_1^\mathcal{O}(1^\lambda, \pk)$
+                \State $s \leftsample A_1^{\mathcal{O}_1}(1^\lambda, \pk)$
                 \State $m^\ast \leftsample \mathcal{M}$
                 \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
-            \State $\hat{m} \leftarrow A_2^\mathcal{O}(1^\lambda, \pk, s, c^\ast)$
+                \State $\hat{m} \leftarrow A_2^{\mathcal{O}_2}(1^\lambda, \pk, s, c^\ast)$
                 \State \Return $\llbrack \hat{m} = m^\ast\rrbrack$
@@
         \end{algorithm}
-    \caption{The one-way security game}\label{fig:ow-atk-game}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.5\textwidth}
+        \begin{algorithm}[H]
+            \caption{\monospace{IND-ATK} game}\label{alg:ind-atk-game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+                \State $(m_0, m_1) \leftsample A_1^{\mathcal{O}_1}(1^\lambda, \pk)$
+                \State $b \leftsample \{0,1\}$
+                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
+                \State $\hat{b} \leftarrow A_2^{\mathcal{O}_2}(1^\lambda, \pk, s, c^\ast)$
+                \State \Return $\llbrack \hat{b} = b \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{The indistinguishability security game}\label{fig:sec-games}
 \end{figure}
+
+The capabilities of the adveraries are modeled using different collections of oracles.
+In standard security requirements, adversaries with access to no additional oracles can only mount chosen plaintex attacks (CPA), adversaries with access to decryption oracle $\mathcal{O}^\decrypt$ only before the receiving challenge ciphertext can mount non-adaptive chosen ciphertext attacks (CCA1), adversaries with access to decryption oracle both before and after receiving the challenge ciphertext can mount adaptive chosen ciphertext attacks (CCA2).
+
+\cite{hofheinz2017modular} also defined two non-standard oracles and the corresponding attacks.
+The plaintext checking oracle $\pco(m, c)$ returns $1$ if $m$ is a decryption of $c$ and $0$ otherwise.
+The ciphertext validation oracle $\cvo(c)$ returns $1$ if $c$ is a valid ciphertext and $0$ otherwise.
+
+\begin{figure}[H]
+    \center
+
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\pco(m, c)$}\label{alg:pco}
+            \begin{algorithmic}
+                \State \Return $\llbrack \decrypt(\sk, c) = m \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\cvo(c)$}\label{alg:cvo}
+            \begin{algorithmic}[1]
+                \State \Return $\llbrack \decrypt(\sk, c) \in \mathcal{M} \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{$\pco$ and $\cvo$}\label{fig:pco-and-cvo}
+\end{figure}
+
+Here is an overview of the various kinds of attacks and their associated oracles
+
+\begin{table}[H]
+    \center
+    \begin{tabular}{c|c|c}
+        \monospace{ATK} & $\mathcal{O}_1$ & $\mathcal{O}_2$ \\
+        \hline
+        \monospace{CPA} & \multicolumn{2}{|c|}{$-$} \\
+        \hline
+        \monospace{CCA1} & $\mathcal{O}^\decrypt$ & - \\
+        \hline
+        \monospace{CCA2} & \multicolumn{2}{|c|}{$\mathcal{O}^\decrypt$} \\
+        \hline
+        \monospace{PCVA} & \multicolumn{2}{|c|}{$\pco, \cvo$} \\
+        \hline
+        \monospace{PCA} & \multicolumn{2}{|c|}{$\pco$} \\
+        \hline
+        \monospace{VA} & \multicolumn{2}{|c|}{$\cvo$} \\
+    \end{tabular}
+    \caption{Attacks and associated oracle access}
+\end{table}
 

### 2024-07-05 faster-fo-transform/cca-transform-survey/main.tex
described gamma-spread and rigidity

@@
 
+\subsubsection{Spread and rigidity}
+The spread of a public key encryption scheme measures the diffusion the encryption routine's output. The higher the spread, the lower the probability of obtaining any specific ciphertext.
+
+\begin{definition}[$\gamma$-spread]\label{def:gamma-spread}
+    For a given keypair $(\pk, \sk)$ and plaintext message $m \in \mathcal{M}$, the \emph{min-entropy} of the encryption routine is:
+
+    $$\operatorname{min-entropy}(\pk, m) := -\log_2\left( 
+        \max_{c \in \mathcal{C}} P\left\lbrack
+            c = \encrypt(\pk, m)
+        \right\rbrack
+    \right) $$
+
+    A $\pke$ has $\gamma$-spread if for all keypairs $(\pk, \sk)$ and plaintext $m \in \mathcal{M}$:
+
+    $$
+        \operatorname{min-entropy}(\pk, m) \leq \gamma
+    $$
+\end{definition}
+
+Having $\gamma$ sperad means that for any keypair $(\pk, \sk)$, plaintext $m$, and ciphertext $c$:
+
+$$
+P\left\lbrack c = \encrypt(\pk, m) \right\rbrack \leq 2^{-\lambda}
+$$
+
+Finally, \emph{rigidity} conveys the idea that a ciphertext cannot be perturbed without becoming either invalid or decrypting to another plaintext
+
+\begin{definition}[rigidity]\label{def:rigidity}
+    $\pke(\keygen, \encrypt, \decrypt)$ is \emph{rigid} if for all keypairs $(\pk, \sk)$ and ciphertext $c$, either $\decrypt(\sk, c) = \bot$ or $\encrypt\left(\pk, \decrypt(\sk, c)\right) = c$
+\end{definition}
+
 \section{Modular Fujisaki-Okamoto transformation}

### 2024-07-07 faster-fo-transform/cca-transform-survey/main.tex
well known result: ind-cpa security implies ow-cpa security

@@
 
+\cite{hofheinz2017modular} stated a ``well-known'' result that the \monospace{IND-CPA} security of a scheme with a large message space implies \monospace{OW-CPA} security:
+
+\begin{theorem}\label{thm:ind-cpa-implies-ow-cpa}
+    For every \monospace{IND-CPA} adversary $B$ against some $\pke$, there exists an \monospace{OW-CPA} adversary $A$ against the same $\pke$ such that:
+
+    \begin{equation*}
+        \monospace{Adv}_\monospace{OW-CPA}(A) = \frac{1}{\norm{\mathcal{M}}} + \monospace{Adv}_\monospace{IND-CPA}(B)
+    \end{equation*}
+\end{theorem}
+
 \subsubsection{Spread and rigidity}
@@
 \section{Modular Fujisaki-Okamoto transformation}
-The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} and its KEM variations \cite{hofheinz2017modular}.
+The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} and its KEM variations \cite{hofheinz2017modular} achieve security against adaptive chosen-ciphertext attacks through \emph{de-randomization} and \emph{re-encryption}.
+In particular, the modular FO transformation contains two steps.
+First, the
 

### 2024-07-07 faster-fo-transform/cca-transform-survey/main.tex
Checkpoint

@@
 
-\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
+\begin{definition}[$\delta$-correctness]\label{def:delta-correctness-pke}
     A probabilistic public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$ is $\delta$-correct if the expected maximal probability of decryption failure taken across the distribution of keypairs is at most $\delta$:
@@
     \center
-    \begin{tabular}{c|c|c}
+    \begin{tabular}{|c|c|c|}
+        \hline
         \monospace{ATK} & $\mathcal{O}_1$ & $\mathcal{O}_2$ \\
@@
         \monospace{VA} & \multicolumn{2}{|c|}{$\cvo$} \\
+        \hline
     \end{tabular}
@@
 
+\subsection{Key encapsulation mechanism}
+A key encapsulation mechanism $\kem = (\keygen, \encap, \decap)$ is a collection of three routines.
+The key generation routine $(\pk, \sk) \leftsample \keygen(1^\lambda)$ takes the security parameter $1^\lambda$ and returns a keypair.
+The encapsulation routine $(c, K) \leftsample \encap(\pk)$ takes the public key and outputs some ciphertext $c \in \mathcal{C}_\kem$ and some shared secret $K \in \mathcal{K}_\kem$.
+Finally, the decapsulation routine $K \leftarrow \decap(\sk, c)$ takes the secret key and a ciphertext and outputs the correponding shared secret.
+
+\emph{Correctness}: similar to a $\pke$, key encapsulation mechanisms are usually required to be perfectly correct, meaning that for all keypairs $(\pk, \sk)$, decapsulation always outputs the same shared secret as the encapsulation
+
+\begin{equation*}
+    \Prob{(c, K_1) \leftsample \encap(\pk); K_2 \leftarrow \decap(\sk, c); K_1 = K_2} = 1
+\end{equation*}
+
+However, where decapsulation failure has a non-zero probability, we need to upperbound this quantity.
+
+\begin{definition}[$\delta$-correctness]\label{def:delta-correctness-kem}
+    A $\kem = (\keygen, \encap, \decap)$ is $\delta$-correct if the probability of decapsulation failure taken across the keypair distribution is at most $\delta$:
+
+    \begin{equation*}
+        P\lrbrack{
+            \decap(\sk, c) \neq K 
+            \mid (\pk, \sk) \leftsample \keygen()
+                ; (c, K) \leftsample \encap(\pk)
+        } \leq \delta
+    \end{equation*}
+\end{definition}
+
+\emph{Security}: the security of a $\kem$ is defined in an adversarial game in which the adversary's goal is to distinguish between shared secret derived from encapsulation and uniformly random samples.
+An adversary $A = (A_1, A_2)$ contains two sub-routines with access to some oracle $\mathcal{O}$ depending on game.
+
+\begin{figure}[H]
+    \center
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{\monospace{IND-CCA2} game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+                \State $s \leftsample A_1^{\mathcal{O}^\decap}(1^\lambda, \pk)$
+                \State $(c^\ast, K_0) \leftsample \encap(\pk)$
+                \State $K_1 \leftsample \mathcal{K}_\kem$
+                \State $b \leftsample \{0, 1\}$
+                \State $\hat{b} \leftsample A_2^{\mathcal{O}^\decap}(1^\lambda, \pk, s, c^\ast, K_b)$
+                \State \Return $\llbrack \hat{b} = b \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap$ oracle $\mathcal{O}^\decap(c \neq c^\ast)$}
+            \begin{algorithmic}[1]
+                \State \Return $\decap(\sk, c)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{The \monospace{IND-CCA2} game for $\kem$}\label{fig:kem-ind-cca2-game}
+\end{figure}
+
+\begin{definition}[\monospace{IND-CCA2 security}]\label{def:kem-ind-cca2-security}
+    A $\kem = (\keygen, \encap, \decap)$ is \monospace{IND-CCA2} secure if no efficient adversary has non-negligible advantage in the \monospace{IND-CCA2} game.
+\end{definition}
+
 \section{Modular Fujisaki-Okamoto transformation}
@@
 In particular, the modular FO transformation contains two steps.
-First, the
+The first step (denoted the $T$ transformation) takes a $\pke = (\keygen, \encrypt, \decrypt)$ and outputs a $\pke_1 = (\keygen, \encrypt_1, \decrypt_1)$.
+The key generation remains unchanged, but the encryption routine is \emph{de-randomized} by deriving the coin from the plaintext using a hash function $G: \mathcal{M}_\pke \rightarrow \mathcal{R}_\pke$, and the decryption routine uses \emph{re-encryption} to reject invalid ciphertexts.
+
+\begin{figure}[H]
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encrypt_1(\pk, m)$}\label{alg:encrypt-1}
+            \begin{algorithmic}[1]
+                \State $r \leftarrow G(m)$
+                \State $c \leftarrow \encrypt(\pk, m; r)$
+                \State \Return $c$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decrypt_1(\sk, c)$}\label{alg:decrypt-1}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow D(\sk, c)$
+                \If{
+                    $\hat{m} \in \mathcal{M}_\pke 
+                    \;\land\; \encrypt(\pk, \hat{m}; G(\hat{m})) = c$
+                }
+                    \State \Return $\hat{m}$
+                \EndIf
+                \State \Return $\bot$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{T transformation}\label{fig:t-transformation}
+\end{figure}
+
+\cite{hofheinz2017modular} states the security property of $\pke_1$ as follows:
+
+\begin{theorem}\label{thm:owcpa-implies-owpcva}
+    If $\pke$ is $\delta$-correct and has $\gamma$ spread, then for every \monospace{OW-PCVA} adversary $B$ against $\pke_1$ who makes $q_G$ hash queries, $q_V$ ciphertext validation queries, and $q_P$ plaintext checking queries, there exists an \monospace{OW-CPA} adversary $A$ such that:
+
+    \begin{equation*}
+        \monospace{Adv}(B) \leq q_V \cdot 2^{-\gamma}
+            + (q_G + q_P) \cdot \delta
+            + (1 + q_G + q_P) \cdot \monospace{Adv}(A)
+    \end{equation*}
+\end{theorem}
+
+In other words, if $\pke$ is \monospace{OW-CPA} secure, then $\pke_1$ is \monospace{OW-PCVA} secure, though the security is non-tight.
+On the other hand, if $\pke$ is \monospace{IND-CPA} secure, the \monospace{OW-PCVA} security is tight:
+
+\begin{theorem}\label{thm:indcpa-implies-owpcva}
+    for every \monospace{OW-PCVA} adversary $B$ against $\pke_1$ who makes $q_G$ hash queries, $q_V$ ciphertext validation queries, and $q_P$ plaintext checking queries, there exists an \monospace{IND-CPA} adversary $A$ such that:
+
+    \begin{equation*}
+        \monospace{Adv}(B) \leq q_V \cdot 2^{-\gamma}
+            + (q_G + q_P) \cdot \delta
+            + \frac{2 \cdot q_G + 1}{\norm{\mathcal{M}_\pke}}
+            + 3 \cdot \monospace{Adv}(A)
+    \end{equation*}
+\end{theorem}
+
+\begin{proof}
+    We will provide a sketch of proof for theorem \ref{thm:owcpa-implies-owpcva} and \ref{thm:indcpa-implies-owpcva}.
+The main idea is to construct an \monospace{OW-CPA}/\monospace{IND-CPA} adversary $A$ who can simulate the \monospace{OW-PCVA} game and use the \monospace{OW-PCVA} adversary $B$ as a sub-routine.
+However, there are three main difficulties with constructing a ``convincing'' simulation:
+
+    \begin{enumerate}
+        \item $A$ has no access to $\pco$
+        \item $A$ has no access to $\cvo$
+        \item The challenge ciphertext $A$ receives is obtained using a truly random coin, but the challenge ciphertext $B$ expects is obtained using a pseudorandom coin $r \leftarrow G(m)$
+    \end{enumerate}
+\end{proof}
 

### 2024-07-08 faster-fo-transform/cca-transform-survey/main.tex
Explained U-bot KEM security

@@
     \end{enumerate}
+
+    % TODO: finish writing this section
 \end{proof}
+
+\subsection{From PKE to KEM}
+The second part of the modular FO transform (denoted the $U$ transform) takes a $\pke$ and outputs a $\kem$.
+Depending on whether the input $\pke$ is rigid and whether the output $\kem$ rejects invalid ciphertext implicitly or explicitly, the security requirements for the input $\pke$ and the construction of the $\kem$ will vary.
+\cite{hofheinz2017modular} presents four variations of the $U$ transform, which are listed in table \ref{tbl:u-transform-summary}
+
+\begin{table}[H]
+    \center
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        Name & $\pke$ requirements & rejection & shared secret\\
+        \hline
+        $U^\bot$ & \monospace{OW-PCVA} & $\bot$ & $H(m, c)$ \\
+        \hline
+        $U^{\not\bot}$ & \monospace{OW-PCA} & $H(s, c)$ & $H(m, c)$ \\
+        \hline
+        $U^\bot_m$ & rigidity + \monospace{OW-VA} & $\bot$ & $H(m)$ \\
+        \hline
+        $U^{\not\bot}_m$ & rigidity + \monospace{OW-CPA} & $H(s, c)$ & $H(m)$ \\
+        \hline
+    \end{tabular}
+    \caption{A summary of variants of $U$ transformations}\label{tbl:u-transform-summary}
+\end{table}
+
+For the remaining of this section we will present the four transformations and sketch proofs of their securities.
+
+\subsubsection{KEM with explicit rejection from randomized PKE}
+Let $\pke = (\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme, and $H: \mathcal{M}_\pke \times \mathcal{C}_\pke \rightarrow \mathcal{K}_\kem$ be a hash function.
+The $U^\bot$ transformation outputs a $\kem^\bot = (\keygen^\bot, \encap^\bot, \decap^\bot)$, where the key generation routine remains unchanged: $\keygen^\bot = \keygen$.
+The encapsulation and decapsulation routines are described in figure \ref{fig:u-bot-routines}
+
+\begin{figure}[H]
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap^\bot(\pk)$}\label{alg:u-bot-encap}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}_\pke$
+                \State $c \leftsample \encrypt(\pk, m)$
+                \State $K \leftarrow H(m, c)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap^\bot(\sk, c)$}\label{alg:u-bot-decap}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
+                \If{$\hat{m} \in \mathcal{M}_\pke$}
+                    \State $K \leftarrow H(\hat{m}, c)$
+                    \State \Return $K$
+                \EndIf
+                \State \Return $\bot$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{$U^\bot$ routines}\label{fig:u-bot-routines}
+\end{figure}
+
+Under the random oracle model, the security of $\kem^\bot$ depends on the security of the input $\pke$.
+
+\begin{theorem}\label{thm:owpcva-implies-indcca2}
+    For every \monospace{IND-CCA2} adversary $B$ against $\kem^\bot$, there exists an \monospace{OW-PCVA} adversary $A$ against the underlying $\pke$ such that
+
+    \begin{equation*}
+        \monospace{Adv}(B) \leq \monospace{Adv}(A)
+    \end{equation*}
+\end{theorem}
+
+\begin{proof}
+    For a sketch of proof, we argue that under the random oracle model, $H(m, c)$ is indistinguishable from a uniformly random sample from $\mathcal{K}_\kem$ from the adversary $B$'s perspective unless $B$ somehow queries $H$ on $(m, c)$.
+This means that an \monospace{OW-PCVA} adversary can perfectly simulate the decapsulation oracle using uniformly random samples, as long as the outputs from the simulated decapsulation oracle are consistent with the outputs from the hash functions.
+
+    The only flaw in the simulated decapsulated oracle lies in the case when $B$ queries $H(m^\ast, c^\ast)$, but because $A$ has access to $\pco$, $A$ can detect that $B$ has successfully recovered $m^\ast$ and uses the recovered $m^\ast$ to win the \monospace{OW-PCVA} game.
+In other words, $A$ keeps running $B$ in a simulation until $B$ makes the special hash query, at which time $A$ simply terminates $B$ and outputs the answer.
+
+    To make the simulated decapsulation oracle and the hash oracle consistent, $A$ maintains two tapes $\mathcal{L}^\decap$ and $\mathcal{L}^H$ that record the queries made to the decapsulation oracle and the hash oracles respectively:
+
+    $$
+    \begin{aligned}
+        (c, K) \in \mathcal{L}^\decap &\Leftrightarrow K = \mathcal{O}^\decap(c) \\
+        (m, c, K) \in \mathcal{L}^H &\Leftrightarrow K = H(m, c)
+    \end{aligned}
+    $$
+
+    If $\mathcal{O}^\decap$ has been queried with some valid ciphertext $c$, then when $H$ is queried with a corresponding input $(m, c)$ where $m$ is the decryption of $c$, $H$ needs to output the same value as $\mathcal{O}^\decap$.
+Similarly, if $H$ has been queried with some valid plaintext-ciphertext pair $(m, c)$, then when $\mathcal{O}^\decap$ is queried with $c$, the two oracles need to output the same value.
+The details of the simulation are described in figure \ref{fig:u-bot-patched-oracles}
+
+    \begin{figure}[H]
+        \begin{minipage}{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:u-bot-patched-decap}
+                \begin{algorithmic}[1]
+                    \If{$\exists (\tilde{c}, \tilde{K}) \in \mathcal{L}^\decap : \tilde{c} = c$}
+                        \State \Return $\tilde{K}$
+                    \EndIf
+                    \If{$\cvo(c) = 1$}
+                        \State $K \leftsample \mathcal{K}_\kem$
+                        \State Append $(c, K)$ to $\mathcal{L}^\decap$
+                        \State \Return $K$
+                    \EndIf
+                    \State \Return $\bot$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \hfill
+        \begin{minipage}{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption{$H_1(m, c)$}\label{alg:u-bot-patched-hash}
+                \begin{algorithmic}[1]
+                    \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H:(\tilde{m}, \tilde{c}) = (m, c)$}
+                        \State \Return $\tilde{K}$
+                    \EndIf
+                    \State $K \leftsample \mathcal{K}_\kem$
+                    \If{$\pco(m, c) = 1$}
+                        \If{$\exists (\tilde{c}, \tilde{K}) \in \mathcal{L}^\decap : \tilde(c) = c$}
+                            \State \Return $\tilde{K}$
+                        \Else
+                            \State Append $(c, K)$ to $\mathcal{L}^\decap$
+                        \EndIf
+                    \EndIf
+                    \State Append $(m, c, K)$ to $\mathcal{L}^H$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \caption{Patched oracles in $U^\bot$}\label{fig:u-bot-patched-oracles}
+    \end{figure}
+\end{proof}
+
+\subsubsection{KEM with implicit rejection from randomized PKE}
+Given $\pke = (\keygen, \encrypt, \decrypt)$, the $U^{\not\bot}$ transformation outputs $\kem = (\keygen^{\not\bot}, \encap^{\not\bot}, \decap^{\not\bot})$.
+With implicit rejection, decapsulation routine returns some output from $H$ even when the input ciphertext $c$ is malformed.
+Specifically, the implicit rejection value depends on the ciphertext and some secret value $s$ that is a part of the secret key.
+
+\begin{figure}[H]
+    \center
+    \begin{minipage}{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\keygen^{\not\bot}$}
+
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.29\textwidth}
+
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.3\textwidth}
+
+    \end{minipage}
+    \caption{$U^{\not\bot}$ routines}
+\end{figure}
 

### 2024-07-09 faster-fo-transform/cca-transform-survey/main.tex
Need to write out the implicit rejection + rigid PKE

@@
     \center
-    \begin{minipage}{0.3\textwidth}
+    \begin{minipage}{0.33\textwidth}
         \begin{algorithm}[H]
-            \caption{$\keygen^{\not\bot}$}
-
+            \caption{$\keygen^{\not\bot}$}\label{alg:u-notbot-keygen}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \keygen(1^\lambda)$
+                \State $s \leftsample \mathcal{M}_\pke$
+                \State $\sk = (\sk^\prime, s)$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
         \end{algorithm}
@@
     \hfill
-    \begin{minipage}{0.29\textwidth}
-
+    \begin{minipage}{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap^{\not\bot}(\pk)$}\label{alg:u-notbot-encap}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}_\pke$
+                \State $c \leftsample \encrypt(\pk, m)$
+                \State $K \leftarrow H(m, c)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
     \end{minipage}
     \hfill
-    \begin{minipage}{0.3\textwidth}
-
+    \begin{minipage}{0.32\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap^{\not\bot}(\sk, c)$}\label{alg:u-notbot-decap}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
+                \If{$\hat{m} \in \mathcal{M}_\pke$}
+                    \State $K \leftarrow H(\hat{m}, c)$
+                    \State \Return $K$
+                \EndIf
+                \State \Return $H(s, c)$
+            \end{algorithmic}
+        \end{algorithm}
     \end{minipage}
@@
 \end{figure}
+
+The security of $\kem^{\not\bot}$, similar to the security of $\kem^\bot$, derives from the indistinguishability between pseudorandom $K \leftarrow H(m, c)$ and truly random $K \leftsample \mathcal{K}_\kem$ under the random oracle model, unless the $\kem$ adversary somehow queries $H$ on $(m^\ast, c^\ast)$, in which case an adversary against the underlying $\pke$ with access to a $\pco$ can detect this special query use it to win the \monospace{OW-PCA} game.
+In addition, because $\decap^{\not\bot}$ will always return random-looking values regardless of the validity of the input ciphertext, the decapsulation oracle can be simulated without needing a ciphertext validation oracle.
+The $\kem$ adversary will be able to distinguish between a true decapsulation oracle and a simulated decapsulation oracle, but only by querying $H$ on $(s, c)$ for some $c$.
+Since $s$ is uniformly random, each query has a $\norm{\mathcal{M}_\pke}^{-1}$ chance of hitting $s$, so across $q_H$ hash queries to $H$, the probability of hitting $s$ at least once is at most $\frac{q_H}{\norm{\mathcal{M}_\pke}}$.
+
+\begin{theorem}
+    For every \monospace{IND-CCA} adversary $B$ against $\kem^{\not\bot}$ that makes $q_H$ hash queries to $H$, there exists an \monospace{OW-PCA} adversary $A$ against the underyling $\pke$ such that:
+
+    \begin{equation*}
+        \monospace{Adv}(B) \leq \frac{q_H}{\norm{\mathcal{M}_\pke}} + \monospace{Adv}(A)
+    \end{equation*}
+\end{theorem}
+
+\subsubsection{KEM with explicit rejection from rigid PKE}
+If the input $\pke$ is rigid (definition \ref{def:rigidity}), then the $\pco$ can be simulated with indistinguishability from the true $\pco$ except for when decryption failure occurs.
+Given $\pke = (\keygen, \encrypt, \decrypt)$ and hash function $H: \mathcal{M}_\pke \rightarrow \mathcal{K}_\kem$, the $U^\bot_m$ transformation outputs a $\kem^\bot_m = (\keygen, \encap^\bot_m, \decap^\bot_m)$ where the key generation routine remains unchanged.
+The encapsulation and decapsulation routines are described in figure \ref{fig:u-bot-m-routines}
+
+\begin{figure}[H]
+    \center
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap^\bot_m(\pk)$}\label{alg:encap-bot-m}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}_\pke$
+                \State $c \leftarrow \encrypt(\pk, m)$
+                \State $K \leftarrow H(m)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap^\bot_m(\sk, c)$}\label{alg:decap-bot-m}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk, c)$
+                \If{$\hat{m} \in \mathcal{M}_\pke$}
+                    \State \Return $H(m)$
+                \EndIf
+                \State \Return $\bot$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{$\kem^\bot_m$ routines}\label{fig:u-bot-m-routines}
+\end{figure}
+
+Under the random oracle model, no adversary $B$ can distinguish between the pseudorandom $K_0 \leftarrow H(m)$ and the truly random $K_1 \leftsample \mathcal{K}_\kem$ unless $B$ queries $H$ on $m^\ast$.
+Because rigidity means $\encrypt(\pk, m^\ast) = c^\ast$ is equivalent to $\decrypt(\sk, c^\ast) = m^\ast$, if $H$ is queried on $m^\ast$, a second adversary $A$ against the $\pke$ can detect the special query and use it to win the \monospace{OW-VA} game.
+
+\begin{theorem}
+    For every \monospace{IND-CCA2} adversary $B$ against $\kem^\bot_m$, there exists an \monospace{OW-VA} adversary $A$ against the underlying $\pke$ such that
+
+    \begin{equation*}
+        \monospace{Adv}(B) \leq \monospace{Adv}(A) + \delta \cdot (q_H + q_D)
+    \end{equation*}
+\end{theorem}
+
+\subsubsection{KEM with implicit rejection from rigid PKE}
+% TODO: write out this section
 

### 2024-07-10 faster-fo-transform/cca-transform-survey/main.tex
Described Kyber's PKE

@@
 
+\subsection{Applications}
+Kyber's round 3 submission \cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} uses \emph{de-randomization}, \emph{re-encryption}, and some variation of the $U$ transformation to convert the \monospace{IND-CPA} $\pke$ into an \monospace{IND-CCA2} $\kem$.
+Here we recall the \monospace{IND-CPA} $\pke$ first, then describe and discuss their respective KEM construction.
+
+In the $\pke$, $R_q = \mathbb{Z}_q[x]/\langle x^n + 1 \rangle$ denotes the quotient ring, $\mathcal{X}_\eta$ denotes the centered binomial distribution that spans $\{-\eta, \eta+1, \ldots, \eta-1, \eta\}$.
+The plaintext space is the set of all polynomials in $R_q$ with coefficients $0, 1$.
+$k$ denotes the dimensions of the secret.
+The full set of parameters is listed in table \ref{tbl:kyber-params}
+
+\begin{table}[H]
+    \center
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        Security level & n & q & k & $\eta_1$ & $\eta_2$ \\
+        \hline
+        Kyber512 & 256 & 3329 & 2 & 3 & 2 \\
+        \hline
+        Kyber768 & 256 & 3329 & 3 & 2 & 2 \\
+        \hline
+        Kyber1024 & 256 & 3329 & 4 & 3 & 2 \\
+        \hline
+    \end{tabular}
+    \caption{Kyber/ML-KEM parameter set}\label{tbl:kyber-params}
+\end{table}
+
+The \monospace{IND-CPA} $\pke(\keygen, \encrypt, \decrypt)$ is identical between round 3 Kyber and ML-KEM. The details of the $\pke$ routines are described in figure \ref{fig:kyber-pke-routines}
+
+\begin{figure}[H]
+    \begin{minipage}{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\keygen_\pke$}\label{alg:kyber-pke-keygen}
+            \begin{algorithmic}[1]
+                \State $A \leftsample R_q^{k \times k}
+                    , \mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{t} \leftarrow A \cdot \mathbf{s}$
+                \State $\pk \leftarrow (A, \mathbf{t}), \sk \leftarrow \mathbf{s}$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encrypt_\pke(\pk, m)$}\label{alg:kyber-pke-encrypt}
+            \begin{algorithmic}[1]
+                \State $(A, \mathbf{t}) \leftarrow \pk$
+                \State $\mathbf{r}_1 \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k,
+                    e_2 \leftsample \mathcal{X}_{\eta_2}$
+                \State $\mathbf{c}_1 \leftarrow A^\intercal \cdot \mathbf{r}_1 + \mathbf{e}_1$
+                \State $c_2 \leftarrow \mathbf{t}^\intercal \cdot \mathbf{r}_1 + e_2 + m \cdot \round{\frac{q}{2}}$
+                \State \Return $(\mathbf{c}_1, c_2)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decrypt_\pke(\sk, c)$}\label{alg:kyber-pke-decrypt}
+            \begin{algorithmic}[1]
+                \State $(\mathbf{c}_1, c_2) \leftarrow c$
+                \State $\mathbf{s} \leftarrow \sk$
+                \State $\hat{m} = c_2 - \mathbf{s}^\intercal \cdot \mathbf{c}_1$
+                \State $\hat{m} \leftarrow \operatorname{Round}(\hat{m})$
+                \State \Return $\hat{m}$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{$\pke$ routines}\label{fig:kyber-pke-routines}
+\end{figure}
+
 \bibliographystyle{alpha}

### 2024-07-10 faster-fo-transform/cca-transform-survey/main.tex
Described kyber and ML-KEM's KEM construction

@@
     \end{minipage}
-    \caption{$U^{\not\bot}$ routines}
+    \caption{$U^{\not\bot}$ routines}\label{fig:u-notbot-routines}
 \end{figure}
@@
 \subsubsection{KEM with implicit rejection from rigid PKE}
+\begin{figure}[H]
+    \caption{$U^{\not\bot}_m$ routines}\label{fig:u-notbot-m-routines}
+\end{figure}
 % TODO: write out this section
@@
 
+For constructing $\kem$ from $\pke$, Kyber uses the $U^{\not\bot}$ transformation while ML-KEM uses the $U^{\not\bot}_m$ transformation.
+In implementation, however, there are two deviations from algorithm listed in figure \ref{fig:u-notbot-routines} and \ref{fig:u-notbot-m-routines}.
+
+The first deviation is the inclusion of the public key $\pk$ when deriving the coin.
+Instead of deriving coin from the plaintext only $r \leftarrow G(m)$, Kyber and ML-KEM derive the coin from both the plaintext and (a hash of) the public key $(K, r) \leftarrow G(m, H(\pk))$.
+The main rationale \cite{bos2018crystals} behind this design choice is to prevent adversaries from precomputing $m$ values that have a high likelihood of triggering decryption failure, which can reveal information about the secret key.
+
+Recall from the decryption routine $\decrypt_\pke$:
+
+\begin{equation*}
+    \begin{aligned}
+        \hat{m} &= \mathbf{s}^\intercal \cdot \mathbf{c}_1 - c_2 \\
+        &= \mathbf{s}^\intercal \cdot (
+            A^\intercal \cdot \mathbf{r}_1 + \mathbf{e}_1
+        ) - (
+            \mathbf{t}^\intercal \cdot \mathbf{r}_1 + e_2 
+            + m \cdot \round{\frac{q}{2}}
+        ) \\
+        &= \mathbf{s}^\intercal \cdot (
+            A^\intercal \cdot \mathbf{r}_1 + \mathbf{e}_1
+        ) - (
+            (
+                A\cdot\mathbf{s} + \mathbf{e}
+            )^\intercal \cdot \mathbf{r}_1 + e_2 
+            + m \cdot \round{\frac{q}{2}}
+        ) \\
+        &= \mathbf{s}^\intercal \cdot \mathbf{e}_1 
+            - \mathbf{e}^\intercal \cdot \mathbf{r}_1 
+            + e_2 + m \round{\frac{q}{2}}
+    \end{aligned}
+\end{equation*}
+
+A decryption failure occurs if the noise term $\mathbf{s}^\intercal\cdot\mathbf{e}_1 - \mathbf{e}^\intercal\cdot\mathbf{r}_1 + e_2$ has large coefficient, which roughly translates to $\mathbf{s}^\intercal\cdot\mathbf{e}_1$ having large coefficients.
+If the coin $r \leftarrow G(m)$ is entirely derived from the adversary, then the adversary can compute a large number of $\mathbf{r}_1, \mathbf{e}_1, e_2$ offline, then observe which set of $\mathbf{r}_1, \mathbf{e}_1, e_2$ trigger decryption failure, which then can be used to infer information about $\mathbf{s}$ and $\mathbf{e}$.
+On the other hand, if the coin is derived from both the message and the public key, then the adversary needs to compute a large set of value per keypair, which pushes this attack into infeasibility even for adversaries with large quantum computers.
+
+A second deviation is that instead of deriving the shared secret from the plaintext and the ciphertext $K \leftarrow H(m, c)$, Kyber derives the shared secret from the plaintext and \emph{a hash of the ciphertext}.
+Since ML-KEM uses the $U^{\not\bot}_m$ transformation, this deviation does not apply.
+\cite{bos2018crystals} cites benchmarking for non-incremental hash APIs to be the reason for this deviation, which I do not fully understand, but has likely less to do with the security of the scheme and more with verification and compliance.
+
+The details of Kyber and ML-KEM's KEM construction are described in figures \ref{fig:kyber-kem-routines} and \ref{fig:mlkem-kem-routines} respectively.
+
+\begin{figure}[H]
+    \begin{minipage}{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\keygen_\monospace{KyberKEM}$}\label{alg:kyber-kem-keygen}
+            \begin{algorithmic}[1]
+                \State $(\pk_\pke, \sk_\pke) \leftsample \keygen_\pke()$
+                \State $z \leftsample R_{0,1}$
+                \State $h \leftarrow H(\pk_\pke)$
+                \State $\sk_\kem \leftarrow (\sk_\pke, z, h, \pk_\pke)$
+                \State $\pk_\kem \leftarrow \pk_\pke$
+                \State \Return $(\pk_\kem, \sk_\kem)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap_\monospace{KyberKEM}(\pk)$}\label{alg:kyber-kem-encap}
+            \begin{algorithmic}[1]
+                \State $m \leftsample R_2$
+                \State $(K, r) \leftarrow G(H(\pk), m)$
+                \State $c \leftarrow \encrypt_\pke(\pk, m; r)$
+                \State $K \leftarrow \monospace{KDF}(K, c)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap_\monospace{KyberKEM}(\sk, c)$}\label{alg:kyber-kem-decap}
+            \begin{algorithmic}[1]
+                \State $(\sk_\pke, z, h, \pk_\pke) \leftarrow \sk$
+                \State $\hat{m} \leftarrow \decrypt_\pke(\sk_\pke, c)$
+                \State $(\hat{K}, \hat{r}) \leftarrow G(h, \hat{m})$
+                \State $\hat{c} \leftarrow \encrypt_\pke(\pk, \hat{m}; \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State \Return $\monospace{KDF}(\hat{K}, H(\hat{c}))$
+                \Else
+                    \State \Return $\monospace{KDF}(z, H(\hat{c}))$
+                \EndIf
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{Kyber's KEM routines}\label{fig:kyber-kem-routines}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{minipage}{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\keygen_\monospace{ML-KEM}$}\label{alg:mlkem-keygen}
+            \begin{algorithmic}[1]
+                \State $(\pk_\pke, \sk_\pke) \leftsample \keygen_\pke()$
+                \State $z \leftsample R_2$
+                \State $h \leftarrow H(\pke_\pke)$
+                \State $\sk_\kem \leftarrow (\sk_\pke, z, h, \pk_\pke)$
+                \State $\pk_\kem \leftarrow \pke_\pke$
+                \State \Return $(\pk_\kem, \sk_\kem)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap_\monospace{ML-KEM}(\pk_\kem)$}\label{alg:mlkem-encap}
+            \begin{algorithmic}[1]
+                \State $m \leftsample R_2$
+                \State $(K, r) \leftarrow G(m, H(\pk))$
+                \State $c \leftarrow \encrypt_\pke(\pk, m; r)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap_\monospace{ML-KEM}(\sk, c)$}\label{alg:mlkem-decap}
+            \begin{algorithmic}[1]
+                \State $(\sk_\pke, z, h, \pk_\pke) \leftarrow \sk$
+                \State $\hat{m} \leftarrow \decrypt_\pke(\sk_\pke, c)$
+                \State $(\hat{K}, \hat{r}) \leftarrow G(\hat{m}, h)$
+                \State $K^\prime \leftarrow \monospace{KDF}(z, c)$
+                \State $\hat{c} \leftarrow \encrypt_\pke(\pk_\pke, \hat{m}; \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State \Return $\hat{K}$
+                \Else
+                    \State \Return $K^\prime$
+                \EndIf
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{ML-KEM's KEM routines}\label{fig:mlkem-kem-routines}
+\end{figure}
+
 \bibliographystyle{alpha}

### 2024-07-11 faster-fo-transform/cca-transform-survey/main.tex
Described Plaintext Awareness

@@
 
+\section{Plaintext awareness and OAEP}
+Another successful strategy for constructing \monospace{IND-CCA2} secure encryption scheme is \emph{Optimal Asymmetric Encryption Padding}\cite{bellare1995optimal}.
+OAEP achieves security against chosen ciphertext attacks by achieving a special property called \emph{Plaintext Awareness}.
+We will introduce plaintext awareness in \ref{sec:plaintext-awareness} and discuss its relationship with \monospace{IND-CCA} security, then describe OAEP and show that it is plaintext aware under the ROM.
+
+\subsection{Plaintext awareness}\label{sec:plaintext-awareness}
+From a high level, plaintext awareness describes a public-key encryption scheme in which an adversary cannnot produce valid ciphertexts without ``knowing'' the corresponding plaintexts.
+The ``knowledge of plaitnext'' is modeled by a universal knowledge extractor $K$, which uses the public key $\pk$ and the random oracle tape $\mathcal{L}$ to recover the decryptin of some ciphertext $c$ without using the secret key.
+Bellare noted in \cite{bellare1998relations} that such a knowledge extractor can only exist within the ROM, or the one-way security of the encryption scheme is trivially broken.
+
+Plaintext awareness also models the adversary's ability to eavesdrop using an ``encryption oracle $\mathcal{O}^\mathcal{E}$'', from which the adversary can query valid ciphertexts without leaving the corresponding hash queries in the random oracle.
+We rule out trivials wins in which an adversary samples a random ciphertexts to which nobody knows the answer by requiring the adversary's output to not be in the encryption oracle's tape.
+
+We can now formally define the \monospace{PA} game.
+Let $\pke(\keygen, \encrypt, \decrypt)$ be a public key encryption scheme.
+Let $H$ be a hash function.
+Let $\mathcal{O}^\mathcal{E}$ be the an encryption oracle.
+Let $\mathcal{L}^H$ and $\mathcal{L}^\mathcal{E}$ respectively denote the tape of the random oracle and the tape of the encryption oracle.
+
+\begin{figure}[H]
+    \center
+    \begin{minipage}{0.5\textwidth}
+        \begin{algorithm}[H]
+            \caption{The \monospace{PA} game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen()$
+                \State $c \leftsample A^{\mathcal{O}^H, \mathcal{O}^\mathcal{E}}(\pk)$
+                \State $\hat{m} \leftarrow K(\pk, c, \mathcal{L}^H, \mathcal{L}^\mathcal{E})$
+                \State \Return $\llbrack c \not\in \mathcal{L}^\mathcal{E} \land \hat{m} \neq D(\sk, c)\rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{The \monospace{PA} game}\label{fig:pa-game}
+\end{figure}
+
+\monospace{PA} adversary wins the game if $c$ is not in $\mathcal{L}^\mathcal{E}$ and the knowledge extractor fails to recover the correct decryption.
+
+\begin{definition}[$\epsilon$-Plaintext Awareness]\label{def:pa}
+    A $\pke$ is $\epsilon$-Plaintext Aware if there exists a knowledge extractor $K$ such that for all efficient $\monospace{PA}$ adversary $A$, the probability of winning the game is at most $\epsilon$.
+\end{definition}
+
+If a $\pke$ is $\epsilon$-plaintext aware with a negligible $\epsilon$, then a decryption oracle can be simulated using the universal knowledge extractor with negligible difference in behaviors.
+
+\begin{theorem}[\cite{bellare1998relations} theorem 4.2]\label{thm:pa-implies-cca}
+    If $\pke$ is $\epsilon$-plaintext aware, then for every \monospace{IND-CCA2} adversary $B$ against $\pke$ making at most $q_D$ decryption queries, there exists an \monospace{IND-CPA} adversary $A$ such that
+
+    \begin{equation*}
+        \monospace{Adv}(B) \leq \monospace{Adv}(A) + q_D \cdot \epsilon
+    \end{equation*}
+\end{theorem}
+
+\subsection{Optimal Asymmetric Encryption Padding}\label{sec:oaep}
+
 \bibliographystyle{alpha}

### 2024-07-12 faster-fo-transform/etm-kyber-is-insecure/main.tex
EtM is sadly insecure

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Custom commands
\usepackage{mystyle}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    ``Encrypt-then-MAC'' with Kyber/ML-KEM is insecure
}
\author{
    Ganyu Xu
}
% Leave the date field empty to display the date of compilation
% \date{}

\begin{document}
%%%% TITLE %%%%%
\maketitle

After some additional thoughts I found a chosen-ciphertext attack against the ``Encrypt-then-MAC'' transformation when combined with Kyber/ML-KEM. This attack takes advantage of the general structure of LWE-based cryptosystem, and I have no immediate ways to patch this problem. Unfortunately I think this means that $\etm$ is a dead end with lattice-based schemes.

\section{A plaintext-checking attack against Kyber}
Recall the construction of \monospace{Kyber.CPAPKE}

\begin{figure}[H]
    \begin{minipage}{0.3\textwidth}
        \begin{algorithm}[H]
            \caption{$\keygen_\pke$}\label{alg:kyber-pke-keygen}
            \begin{algorithmic}[1]
                \State $A \leftsample R_q^{k \times k}
                    , \mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
                \State $\mathbf{t} \leftarrow A \cdot \mathbf{s}$
                \State $\pk \leftarrow (A, \mathbf{t}), \sk \leftarrow \mathbf{s}$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.3\textwidth}
        \begin{algorithm}[H]
            \caption{$\encrypt_\pke(\pk, m)$}\label{alg:kyber-pke-encrypt}
            \begin{algorithmic}[1]
                \State $(A, \mathbf{t}) \leftarrow \pk$
                \State $\mathbf{r}_1 \leftsample \mathcal{X}_{\eta_1}^k$
                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k,
                    e_2 \leftsample \mathcal{X}_{\eta_2}$
                \State $\mathbf{c}_1 \leftarrow A^\intercal \cdot \mathbf{r}_1 + \mathbf{e}_1$
                \State $c_2 \leftarrow \mathbf{t}^\intercal \cdot \mathbf{r}_1 + e_2 + m \cdot \round{\frac{q}{2}}$
                \State \Return $(\mathbf{c}_1, c_2)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.3\textwidth}
        \begin{algorithm}[H]
            \caption{$\decrypt_\pke(\sk, c)$}\label{alg:kyber-pke-decrypt}
            \begin{algorithmic}[1]
                \State $(\mathbf{c}_1, c_2) \leftarrow c$
                \State $\mathbf{s} \leftarrow \sk$
                \State $\hat{m} = c_2 - \mathbf{s}^\intercal \cdot \mathbf{c}_1$
                \State $\hat{m} \leftarrow \operatorname{Round}(\hat{m})$
                \State \Return $\hat{m}$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$\pke$ routines}\label{fig:kyber-pke-routines}
\end{figure}

This construction has no ciphertext integrity, meaning that an adversary can submit well-formed AND malformed ciphertexts and recover the secret key by observing the behavior of the decryption routine. Here we present a plaintext-checking attack, which uses a plaintext checking oracle:

\begin{figure}[H]
    \center
    \begin{minipage}{0.3\textwidth}
        \begin{algorithm}[H]
            \caption{$\pco(m, c)$}
            \begin{algorithmic}[1]
                \State \Return $\llbrack \decrypt(\sk, c) = m \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
\end{figure}

Also recall that with Kyber/ML-KEM, each polynomial can be transformed into the NTT domain:

\begin{equation*}
    R_q = \frac{\mathbb{Z}_{3329}[x]}{\left\langle x^{256} + 1 \right\rangle}
    \cong \frac{\mathbb{Z}_{3329}[x]}{\left\langle x^2 + \zeta \right\rangle}
    \times \frac{\mathbb{Z}_{3329}[x]}{\left\langle x^2 + \zeta^3 \right\rangle}
    \times \ldots
    \times \frac{\mathbb{Z}_{3329}[x]}{\left\langle x^2 + \zeta^{255} \right\rangle}
\end{equation*}

Where $\zeta$ is any solution to $\zeta^{128} + 1 \equiv 0 \mod 3329$ (Kyber picked $\zeta=17$). We denote the NTT representation by:

\begin{equation*}
    \ntt(y \in R_q) = (\hat{y}_1, \hat{y}_2, \ldots, \hat{y}_{128})
\end{equation*}

Where each $\hat{y}_j$ for $1 \leq j \leq 128$ is a degree-1 polynomial.

For each of $i \in \{1, 2, \ldots, k\}$ and each of $j \in \{1, 2, \ldots, 128\}$, an adversary can craft a maliciously malformed ciphertext $c = (\mathbf{c}_1, c_2)$ such that:

\begin{itemize}
    \item $\ntt(c_2) = (0, 0, \ldots, \hat{c}_{2, j}, \ldots, 0)$ is all 0's except for the j-th entry, which is chosen by the adversary
    \item $\mathbf{c}_1 = (0, 0, \ldots, c_{1, i}, \ldots, 0)$ is all 0's except for the i-th entry $c_{1, i}$, whose NTT representation $\ntt(c_{1, i}) = (0, 0, \ldots, \hat{c}_{1, i, j} = 1, \ldots, 0)$ is all 0's except for the j-th entry, which is 1.
\end{itemize}

In line 3 of the decryption routine (algorithm \ref{alg:kyber-pke-decrypt}):

\begin{equation*}
    \begin{aligned}
        \ntt(c_2 - \mathbf{s}^\intercal \cdot \mathbf{c}_1)
        &= \ntt(c_2) - \ntt(\mathbf{s}^\intercal \cdot \mathbf{c}_1) \\
        &= (0, 0, \ldots, \hat{c}_{2, j}, \ldots, 0)
            - \ntt(s_i \cdot c_{1, i}) \\
        &= (0, 0, \ldots, \hat{c}_{2, j}, \ldots, 0)
            - \ntt(s_i) \circ \ntt(c_{1, i}) \\
        &= (0, 0, \ldots, \hat{c}_{2, j}, \ldots, 0)
            - (\hat{s}_{i, 1}, \hat{s}_{i, 2}, \ldots, \hat{s}_{i, 128})
            \circ (0, 0, \ldots, c_{1, i, j} = 1, \ldots, 0) \\
        &= (0, 0, \ldots, \hat{c}_{2, j} - \hat{s}_{i, j}, \ldots, 0)
    \end{aligned}
\end{equation*}

I will make an unverified but probably correct claim: \emph{if $\hat{c}_{2, j} - \hat{s}_{i, j} \neq 0$ then with very high probability $\ntt^{-1}((0, 0, \ldots, \hat{c}_{2, j} - \hat{s}_{i, j}, \ldots, 0))$ will not round to $0$}. This means that with very high probability, $(\mathbf{c}_1, c_2)$ will not decrypt to $0$ if $\hat{c}_{2, j} \neq \hat{s}_{i, j}$, which is equivalent to $\pco(m=0, c=(\mathbf{c}_1, c_2)) = 0$.

The adversary can thus iterate through all $q^2$ possible degree-1 polynomials to find the correct value for $\hat{s}_{i, j}$, then repeat it for all $i, j$. In $q^2 \cdot k \cdot \frac{n}{2}$ operations, the adversary can recover the secret key.

\section{$\etm$ is vulnerable to the key recovery attack above}
Suppose an adversary crafts a malicious ciphertext using the strategy described above $c = (\mathbf{c}_1, c_2)$, if the secret key value is such that $\hat{c}_{2, j} = \hat{s}_{i, j}$, then $m = 0$ should be the correct decryption, which means that $k = G(0)$ should be the correct MAC key, so the adversary computes the tag $t = \sign(G(0), c)$.

If $(c, t)$ is rejected, then the adversary learns that $0$ is not the correct decryption. From here the key recovery attack described above can be executed. A similar attack can be executed using $\pco$ against $\etm$. This means that at least with Kyber/ML-KEM, $\etm$ is not one-way secure with either $\pco$ or $\cvo$.

\end{document}

### 2024-07-14 archives/co789/homework-01/q01/main.tex
What's next

--- 2024-01-27 co789/homework-01/main.tex
@@
 \subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
+Notice from the probability formula above:
 
@@
 \begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
+\prod_{i=0}^{n-1} (1 - q^{i-n})
+&= (1-q^{-n}) \cdot \prod_{i=1}^{n-1}(1 - q^{i-n}) \\
+&= (1-q^{-n}) \cdot \prod_{i=0}^{n-2}(1 - q^{i - (n - 1)}) \\
+\end{aligned}
+$$
+
+The product in the R.H.S.
+is the probability of drawing $n-1$ linearly independent vectors from $\mathbf{Z}_q^{n-1}$.
+Since $1 - q^{-n} < 1$ for all $q, n > 0$, the value of this probability strictly decreases as $n$ increases.
+Therefore, we can bound the probability from below by setting $n$ to the maximal value $1024$:
+
+$$
+\begin{aligned}
+\prod_{i=0}^{n-1} (1 - q^{i-n}) 
+&\geq \prod_{i=0}^{1023} (1 - q^{i-1024}) \\
+&= (1-q^{-1024})(1-q^{-1023}) \ldots (1-q^{-1})
+&\geq (1 - q^{-1})^{1024} \\
+&= (1 - 3329^{-1})^{1024} \\
+&\approx 0.735175 > \frac{2}{3}
 \end{aligned}
@@
 
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-07-14 archives/co789/homework-01/q02/main.tex
What's next

--- 2024-01-27 co789/homework-01/main.tex
@@
 
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
 \section*{Q2}
@@
 
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
+P.S.
+Given two independent random variables $A, B \in G$ whose support lies in some finite group, if $A$ is uniformly random, then $A + B$ is also uniformly random regardless of the distribution of $B$:
 
@@
 \begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
+P(A + B = c)
+&= \sum_{a \in G}P(A=a)P(B=c - a) \\
+&= \sum_{a \in G}\frac{1}{\vert G \vert}P(B=c - a) \\
+&= \frac{1}{\vert G \vert}\sum_{a \in G}P(B=c - a) \\
+&= \frac{1}{\vert G \vert} \cdot 1 = \frac{1}{\vert G \vert} \\
 \end{aligned}
@@
 
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-07-14 archives/co789/homework-01/q07/main.tex
What's next

--- 2024-01-27 co789/homework-01/main.tex
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
 
@@
 
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
-
-\section*{Q9}
-Let $(A, \mathbf{b}) \leftarrow \operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-In other words, $A$ is uniformly randomly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, $\mathbf{s} \leftarrow \chi_e^n$, $\mathbf{e} \leftarrow \chi_s^n$, and $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$.
-
-I claim that matrix inversion $A \mapsto A^{-1}$ on the set of full-rank matrices is a bijection.
-This is true because the inverse of two distinct matrices is necessarily distinct (injectivity), and every full-rank matrix is the inverse of its inverse (surjectivity).
-
-Because matrix inversion is a bijection from the set of invertible matrices onto itself, if $A$ is uniformly sampled from all full-rank matrices $\mathbb{F}_q^{n \times n}$, then $A^{-1}$ is also a uniformly randomly sampled matrix from the set all full-rank matrices.
-
-Notice that $A^{-1}\mathbf{b} = A^{-1}A\mathbf{s} + \mathbf{e} = A^{-1}\mathbf{e} + \mathbf{s}$.
-Since $\mathbf{e}$ is sampled from the secret distribution and $\mathbf{s}$ is sampled from the error distribution, $A^{-1}$ is a uniformly random sample, $(A^{-1}, A^{-1}\mathbf{b}) = (A^{-1}, A^{-1}\mathbf{e} + \mathbf{s})$ is a sample from $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$.
-If there exists an oracle for $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$, then this oracle can recover $\mathbf{e}$, and from here we can recover $\mathbf{s}$ in $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$.
-
-The argument above did not assume anything specific to $\chi_s$ or $\chi_e$, so they can be swapped, and we will arrive at the inverse conclusion that $\operatorname{LWE}(n, n, q, \chi_s, \chi_e)$ reduces to $\operatorname{LWE}(n, n, q, \chi_e, \chi_s)$, as well.
-Therefore, the two problems are equivalent.
-
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-07-14 archives/co789/homework-01/q09/main.tex
What's next

--- 2024-01-27 co789/homework-01/main.tex
@@
 % \maketitle
-
-\section*{Q1}
-\subsection*{(1)}
-The probability formula will be proved by induction. The base case is trivial: an empty set is always linearly independent, so the probability of drawing a linearly independent empty set is exactly 1.
-
-For the induction, assume that $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1} \in \mathbb{F}_q^n$ are linearly independent, then consider the probability of drawing a t-th vector uniformly from $\mathbb{F}_q^n$ such that it is not in the linear span of the previous $t-1$ vectors: $\mathbf{a}_t \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})$.
-
-There are a total of $q^n$ possible values for $\mathbf{a}_t$ to draw from.
-On the other hand, there are a total of $q^{t-1}$ possible combinations of coefficients (including all zeros) for $t-1$ vectors.
-Singe $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$ are linearly independent, each unique combination of coefficients corresponds to a unique element in the linear span.
-Thus, there are a total of $q^n - q^{t-1}$ possible values that are outside the linear span of $\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1}$, and the probability of uniformly drawing a $\mathbf{a}_t$ that's outside the linear span is $1 - \frac{q^{t-1}}{q^n}$.
-
-In other words:
-
-$$
-P(
-    \mathbf{a}_t 
-    \notin \operatorname{Span}(\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_{t-1})
-    \mid \{\mathbf{a}_i\}_{i=1}^{t-1} \text{ is linearly independent}
-) = 1 - q^{t-1-n}
-$$
-
-From here, we can recursively compute the probability of drawing $m$ linearly independent vectors:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1} P(
-    \mathbf{a}_{j+1} 
-    \notin \operatorname{Span}(\{\mathbf{a}_i\}_{i=1}^{j})
-    \mid \{\mathbf{a}_i\}_{i=1}^{j} \text{ is linearly independent}
-) \\
-&= \prod_{j=0}^{m-1}(1 - q^{(j+1) - 1 - n}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n})
-\end{aligned}
-$$
-
-\subsection*{(2)}
-% TODO: this isn't quite right
-We will bound the probability formula above from below by bounding each $1 - q^{j-n}$ term:
-
-$$
-\forall 0 \leq j < n, \; 1 - q^{j-n} \geq 1 - q^{-n}
-$$
-
-And since $m = n \leq 1024$, we have $1 - q^n \leq 1 - q^{1024}$
-
-Thus:
-
-$$
-\begin{aligned}
-&P(\{\mathbf{a}_i\}_{i=1}^m \text{ is linearly independent}) \\
-&= \prod_{j=0}^{m-1}(1 - q^{j - n}) \\
-&\geq \prod_{j=0}^{m-1}(1 - q^{-n}) \\
-&\geq \prod_{j=0}^{1024 - 1}(1 - q^{-1}) \\
-&= (1 - 3329^{-1})^{1024} \approx 0.7351 > \frac{2}{3}
-\end{aligned}
-$$
-
-\subsection*{(3)}
-For Kyber-512, the parameters are defined with $n=256, q=3329$. Plugging them into the formula:
-
-\lstinputlisting[language=Python]{q1_3.py}
-
-The result is $0.999699519257883$
-
-\newpage
-
-\section*{Q2}
-Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$.
-Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$.
-As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise.
-Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.
-
-\newpage
-
-\section*{Q3}
-Recall that a centered binomial distribution is a binomial distribution left-shifted by its mean.
-Let $X \leftarrow \mathcal{B}(n, p)$ be some binomial distribution, then the centered binomial distribution is described by $Y = X - E[X]$:
-
-$$
-P[Y=y] = P[X = y + \mu] = C(n, y + \mu)p^{y + \mu}(1-p)^{n - y - \mu}
-$$
-
-\subsection*{(1)}
-The probability mass function (PMF) of a centered binomial distribution $X \leftarrow \mathcal{B}(n=6, p=0.5)$ is given by:
-
-$$
-P(X=x) = {n \choose {x+np}}p^{x + np}(1-p)^{n - x - np} = {6 \choose {x+3}}2^{-6}
-$$
-
-On the other hand, the PMF of a discrete Gaussian with $N(\mu=3, \sigma^2=\frac{3}{2})$ is given by:
-
-$$
-P(X=x) = \frac{\rho(x)}{\sum_{j=0}^{q-1}\rho(y)}
-$$
-
-I used some Python code to approximate the statistical distance:
-
-\lstinputlisting[language=Python]{q3_1.py}
-
-The result is $0.017725703977230414$.
-
-\subsection*{(2)}
-I claim without proof that the most likely error $\mathbf{s} \leftarrow \chi_e^m$, is obtained by sampling the most likely value for each of the entry in.
-Assuming individual entries of $\mathbf{s}$ are independently sampled from identical distribution $\chi_e$ (a centered binomial distribution), the most likely value for a single entry is $0$.
-Therefore, the most likely secret is $\mathbf{s} = \mathbf{0} \in \mathbb{F}_q^n$.
-The probability of drawing $\mathbf{0} \leftarrow \mathcal{B}(6, \frac{1}{2})$ is the product of drawing 512 $0$'s:
-
-$$
-P(\mathbf{s} = \mathbf{0}) = (\frac{5}{16})^{512}
-$$
-
-\subsection*{(3)}
-Assume that $\mathbf{s} \leftarrow \mathbb{F}_q^n$ where $n = 512$. The probability of drawing a single $0$ from a centered binomial distribution $\mathcal{B}(6, \frac{1}{2})$ is:
-
-$$
-P(Y=0) = P(X = 0 + 3) = C(6, 3)(\frac{1}{2})^3(\frac{1}{2})^3 = \frac{5}{16}
-$$
-
-Since each entry of $\mathbf{s} \leftarrow \mathbb{F}_q^{512}$ is independently sampled from this centered binomial distribution, the count of $0$'s in $\mathbf{s}$ also follows a binomial distribution $\mathcal{B}(512, \frac{5}{16})$.
-The most likely number of $0$ in the secret is thus $512 \cdot \frac{5}{16} = 160$.
-
-In similar fashion, it can be computed that the probability of drawing $1$ from the centered binomial distribution is $C(6, 4)(\frac{1}{2})^6 = \frac{15}{64}$, so the most likely number of $\pm 1$ in the secret is $512 \cdot \frac{15}{64} = 120$, of $\pm 2$ is 48, of $\pm 3$ is 8.
-
-\subsection*{(4)}
-\subsubsection*{(a)}
-A guess $\hat{\mathbf{s}} \leftarrow \mathbb{F}_q^n$ is correct if the corresponding error term $\hat{\mathbf{e}} \leftarrow \mathbf{b} - A\hat{\mathbf{s}}$ is bounded by the centered binomial distribution: $\hat{\mathbf{e}} \in \{-3, -2, \ldots, 2, 3\}^n$.
-
-\subsubsection*{(b)}
-The total number of distinct keys with 160 entries being 0, 120 entries being -1, 48 entries being -2, 8 entries being -3, ... is as follows:
-
-$$
-n = \frac{512!}{160!120!120!48!48!8!8!}
-$$
-
-Assuming the uniqueness of the secret, there is exactly one correct value for for $\mathbf{s}$.
-The random process of drawing from $n$ distinct keys without replacement, among which exactly 1 key is considered "success", is modeled by the negative hypergeometric distribution with $N = n, K = 1, r = (N - K) = n - 1$.
-The expectation of such a distribution is $\frac{n-1}{n}$
-
-\newpage
-
-\section*{Q4}
-Using the Kyber described in the definition sheet, the LWE parameters are as follows: $n = m$, $q = 3329$, $\chi_s = \mathcal{B}(n=6, p=0.5)$, $\chi_e = \mathcal{B}(n=4, p=0.5)$, where $\mathcal{B}(n, p)$ denotes the centered binomial distributions.
-
-Since $\mathbf{s} \leftarrow \chi_s^{n}$ is independently sampled from identical distributions, we can describe $\norm{\mathbf{s}}^2$ as the sum of I.I.D. random variables:
-
-$$
-\norm{\mathbf{s}}^2 = \sum_{i=1}^{n} S_i^2
-$$
-
-Therefore:
-
-$$
-E[\norm{\mathbf{s}}^2] = E\bigg[\sum_{i=1}^{n} S_i^2\bigg] = \sum_{i=1}^{n}E[S_i^2]
-$$
-
-Because $S_i$ follows the \textbf{centered} binomial distribution, $E[S_i] = 0$, so $E[S_i^2] = \var{S_i}$.
-On the other hand, the variance of the centered binomial distribution is identical to that of the corresponding binomial distribution: $\var{S_i} = 6 \cdot p(1-p) = \frac{3}{2}$.
-This is true because shifting a random variable by a constant does not change its variability.
-
-Putting everything together:
-
-$$
-E[\norm{\mathbf{s}}^2] = \sum_{i=1}^{n}E[S_i^2] = \frac{3}{2} n
-$$
-
-On the other hand, for calculating the variance of $\norm{\mathbf{s}}^2$, we take advantage of the fact that the entries of $\mathbf{s}$ are independently drawn, and the variance of sum of independent random variables is the sum of variances:
-
-$$
-\begin{aligned}
-\var{\norm{s}^2} &= \var{\sum_{i=1}^n S_i^2} \\
-&= \sum_{i=1}^n \var{S_i^2} \\
-&= \sum_{i=1}^n (E[S_i^4] - E[S_i^2]^2) \\
-&= \sum_{i=1}^n \bigg(
-    (\sum_{j=-3}^3(j^4 \cdot {6 \choose j} \cdot 2^{-6}) 
-    - (\frac{3}{2})^2 
-\bigg) \\
-&= \sum_{i=1}^n \bigg( 6  - \frac{9}{4} \bigg) \\
-&= \frac{15}{4}n
-\end{aligned}
-$$
-
-Replacing the secret distribution with the error distribution, we can compute the expectation and variance of the norm square of the error term in similar fashion. In conclusion:
-
-$$
-\begin{aligned}
-E[\norm{\mathbf{s}}^2] &= \frac{3}{2} n \\
-\var{\norm{s}^2} &= \frac{15}{4}n \\
-E[\norm{\mathbf{e}}^2] &= n \\
-\var{\norm{e}^2} &= \frac{3}{2}n
-\end{aligned}
-$$
-
-\newpage
-
-\section*{Q5}
-\subsection*{(1)}
-Denote the columns of $A$ by $A = [\mathbf{a}_1, \mathbf{a}_2, \ldots, \mathbf{a}_n]$.
-Without loss of generality, let $\mathbf{a}_n$ be a non-zero linear combination of the other $n-1$ columns: $\mathbf{a}_n = A^\prime \mathbf{z}^\prime$ for some $\mathbf{z} \in \mathbb{Z}^{n-1}$.
-
-It is easy to see that because $A^\prime$ contains only a subset of columns of $A$, so $A^\prime\mathbb{Z}^{n-1} \subseteq A\mathbb{Z}^n$. It naturally follows that
-
-$$
-A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m \subseteq A\mathbb{Z}^n + q\mathbb{Z}^m
-$$
-
-On the other hand, let $\mathbf{v} \in A \mathbb{Z}^{n} + q\mathbb{Z}^m$, then there exist $\mathbf{x}_1 \in \mathbb{Z}^n, \mathbf{x}_2 \in \mathbb{Z}^m$ such that
-
-$$
-\begin{aligned}
-\mathbf{v} &= A\mathbf{x}_1 + q\mathbf{x}_2 \\
-&= \sum_{i=1}^n(\mathbf{a}_i \mathbf{x}_{(1, i)}) + q\mathbf{x}_2 \\
-&= (\sum_{i=1}^{n-1}\mathbf{a}_i x_{(1, i)}) + \mathbf{a}_n x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime \cdot (x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + A^\prime\mathbf{z}^\prime x_{(1, n)} + q\mathbf{x}_2 \\
-&= A^\prime((x_{(1, 1)}, x_{(1, 2)}, \ldots, x_{(1, n-1)}) + \mathbf{z}^\prime x_{(1, n)}) + q\mathbf{x}_2 \in A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m
-\end{aligned}
-$$
-
-Therefore we have $A\mathbb{Z}^{n} + q\mathbb{Z}^m \subseteq A^\prime\mathbb{Z}^{n-1} + q\mathbb{Z}^m$, and the two lattices are indeed equal.
-
-\subsection*{(2)}
-Let $(A, \mathbf{b})$ be a sample from generic (aka potentially not full-rank) $\text{LWE}(m, n, q, U_s, \chi_e)$.
-Without loss of generality, assume that $A = [A_1 \mid A_2] \in \mathbb{Z}_q^{m \times (n_1 + n_2)}$ where $A_1$ is full-rank, and $A_2 = A_1B$ for some non-zero $B \in \mathbb{Z}_q^{n_1 \times n_2}$.
-Denote the secret by $\mathbf{s} = [\mathbf{s}_1 \mid \mathbf{s}_2]$ where $\mathbf{s}_1 \leftarrow \chi_s^{n_1}, \mathbf{s}_2 \leftarrow \chi_s^{n_2}$, then:
-
-$$
-\begin{aligned}
-\mathbf{b} &= A\mathbf{s} + \mathbf{e} \\
-&= (A_1\mathbf{s}_1 + A_2\mathbf{s}_2) + \mathbf{e} \\
-&= A_1\mathbf{s}_1 + A_1B\mathbf{s}_2 + \mathbf{e} \\
-&= A_1(\mathbf{s}_1 + B\mathbf{s}_2) + \mathbf{e}
-\end{aligned}
-$$
-
-Because secrets are sampled from a uniform distribution, $\mathbf{s}_1 + B\mathbf{s}_2$ is uniformly random.
-Therefore $(A_1, \mathbf{b})$ is a valid sample from full-rank $\text{LWE}(m, n, q, U_s, \chi_e)$.
-We can thus feed $(A_1, \mathbf{b})$ to a full-rank LWE oracle, who will output some $\mathbf{s}^\prime$.
-Assuming that the full-rank LWE problem has unique solution, we know $\mathbf{s}^\prime = \mathbf{s}_1 + B\mathbf{s}_2$
-
-% TODO: how to recover s_1, s_2 from s_1 + Bs_2?
-
-\newpage
-
-\section*{Q6}
-\subsection*{(1)}
-Let $X \in \mathcal{B}(n, p)$ be a random variable that follows binomial distribution.
-Recall from the definition of a binomial distribution that $X = I_1 + I_2 + \ldots + I_n$ where each of $I_i$ is an independent coin toss with PMF:
-
-$$
-\begin{cases}
-P(I_i=1) = p \\
-P(I_i=0) = 1-p
-\end{cases}
-$$
-
-Let $X_1 \in \mathcal{B}(n_1, p), X_2 \in \mathcal{B}(n_2, p)$ be two independent random variables following binomial distributions, then:
-
-$$
-\begin{aligned}
-X_1 &= \sum_{i=1}^{n_1} I_i \\
-X_2 &= \sum_{i=n_1 + 1}^{n_1 + n_2} I_i \\
-\end{aligned}
-$$
-
-Therefore $X_1 + X_2 = \sum_{i=1}^{n_1 + n_2} I_i$, which is a binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-Recall that centered binomial distribution is defined by subtracting the corresponding binomial distribution by a constant (the expectation of said binomial distribution): $C_i = X_i - E(X_i)$.
-Therefore, given centered binomial distributions $C_1 = X_1 - E[X_1], C_2 = X_2 - E[X_2]$:
-
-$$
-\begin{aligned}
-C_1 + C_2 &= X_1 + X_2 - E[X_1] - E[X_2] \\
-&= (X_1 + X_2) - E[X_1 + X_2]
-\end{aligned}
-$$
-
-From the results above we know that because $X_1, X_2$ are independent binomial distributions, $X_1 + X_2$ follows binomial distribution $\mathcal{B}(n_1 + n_2, p)$, thus $C_1 + C_2$ follows centered binomial distribution $\mathcal{B}(n_1 + n_2, p)$.
-
-\subsection*{(2)}
-From part (a) we know that the sum of $k$ i.i.d. random variables following binomial $\mathcal{B}(n, p)$ is a random variable following binomial $\mathcal{B}(kn, p)$.
-
-Let $I^m = \{0, 1\}^m$ denote the set of bit-masks with length $m$ and $K \subseteq I$ denote the subset of vectors with exactly $k$ entries being $1$, then $\vert K \vert = {m\choose k}$.
-Thus we iterate through all possible values in $\mathbf{k} \in K$ and compute the inner product $\mathbf{e}^\intercal \mathbf{k}$.
-Since $\mathbf{k}$ has exactly $k$ entries being 1 and $\mathbf{e}$ contains $m$ independent samples from centered binomial $(n, p)$, $\mathbf{e}^\intercal\mathbf{k}$ is the sum of $k$ i.i.d.
-centered binomial with parameters $(kn, p)$
-
-\newpage
-
-\section*{Q7}
-We can set one of the base vectors to be extremely short and all other base vectors extremely long. For ease of constructions we will build orthogonal basis.
-
-Denote the columns of the identity matrix $I_n \in \mathbb{R}^{n \times n}$ by $I = [\mathbf{e}_1, \mathbf{e}_2, \ldots, \mathbf{e}_n]$.
-
-Let $\mathcal{L}(\{\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n\}) \subset \mathbb{R}^n$ be spanned by the basis $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$, where $\mathbf{b}_1 = \mathbf{e}_1$, and $\mathbf{b}_i = 3R\mathbf{e}_i$ for $i > 1$.
-Finally, set $\mathbf{v} = (0, 1.1R, 0, 0, \ldots, 0)$.
-
-It is easy to see that $\mathbf{b}_1, \mathbf{b}_2, \ldots, \mathbf{b}_n$ form an orthogonal basis.
-The shortest vector in $\mathcal{L}$ is $\mathbf{b}_1$, so $\lambda_1(\mathcal{L}) = 1$.
-The closest lattice point to $\mathbf{v}$ is $\mathbf{0}$, so the distance between $\mathbf{v}$ and $\mathcal{L}$ is $\norm{\mathcal{L} - \mathbf{v}} = 1.1R > R$
-
-\newpage
-
-\section*{Q8}
-Recall that the decryption is computed with the following routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1 \cdot \mathbf{s} \\
-&= ({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime} + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Where $\mathbf{s}, \mathbf{s}^\prime$ are coordinate-wise drawn from the secret distribution and $\mathbf{e}, \mathbf{e}^\prime, e^{\prime\prime}$ are coordinate-wise drawn from the error distribution.
-
-According to the decryption routine, a decryption error occurs if and only if the "noise" $({\mathbf{s}^\prime}^\intercal\mathbf{e} - {\mathbf{e}^\prime}^\intercal\mathbf{s}) + e^{\prime\prime}$ exceeds $\round{\frac{q}{4}}$.
-So to find a modulus that guarantees correct decryption all the time, we need to find the upper bound of noise.
-
-With (baby) Kyber-512, $\chi_s = \mathcal{B}(n=6, p=0.5), \chi_e = \mathcal{B}(n=4, p=0.5)$.
-Noise is maximized when all entries of $\mathbf{s}, \mathbf{e}$ reach the extremes of the support of their respective distributions.
-For example, with $\mathbf{s} = \mathbf{s}^\prime = (3, 3, \ldots, 3)$, $\mathbf{e} = (2, 2, \ldots, 2)$, $\mathbf{e}^\prime = (-2, -2, \ldots, -2)$, $e^{\prime\prime} = 2$, the noise term evaluates to $(6 \cdot 512 + 6 \cdot 512) + 2 = 6146$.
-The smallest prime $q$ such that $\round{\frac{q}{4}} \geq 6146$ is $24593$.
-
-\newpage
 
@@
 
-\newpage
-
-\section*{Q10}
-Recall from textbook LWE the decryption routine:
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2))
-&= c_2 - \mathbf{c}_1\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{b} + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= (\mathbf{s}_1^\intercal(A\mathbf{s} + \mathbf{e}) + e^\prime + m\round{\frac{q}{2}}) 
-    - (\mathbf{s}_1^\intercal A\mathbf{s} + \mathbf{e}_1^\intercal\mathbf{s}) \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + m\round{\frac{q}{2}} - \mathbf{e}_1^\intercal\mathbf{s} \\
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}}
-\end{aligned}
-$$
-
-Also recall the definition of the rounding operator: $\round{a} = a + \delta$ for some $-\frac{1}{2} \leq \delta < \frac{1}{2}$.
-
-\subsection*{(1)}
-Where $\mathbf{e}_1$ is maliciously set to all zeros except for the i-th component, $\mathbf{e}_1^\intercal \mathbf{s}$ evaluates to $s_i \cdot \round{\frac{q}{2}}$, where $s_i$ is the i-th component of the LWE secret $\mathbf{s}$.
-
-If $s_i = 2k$ is even, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k(\frac{q}{2} + \delta) + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime - 2k\delta + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Thus, where $k$ is sufficiently small (corresponding to $s_i$ being small), $\mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta$ has a high probability of being less than $\frac{q}{4}$, so the ciphertext will be correctly decrypted.
-
-If $s_i = 2k + 1$ is odd, then
-
-$$
-\begin{aligned}
-D(\text{sk}, (\mathbf{c}_1, c_2)) 
-&= (\mathbf{s}_1^\intercal\mathbf{e} - \mathbf{e}_1^\intercal\mathbf{s}) + e^\prime + m\round{\frac{q}{2}} \\
-&= \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + (2k + 1)\round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-&\equiv  \mathbf{s}_1^\intercal\mathbf{e} + e^\prime + 2k\delta + \round{\frac{q}{2}} + m\round{\frac{q}{2}} \\
-\end{aligned}
-$$
-
-Notice the additional $\round{\frac{q}{2}}$ term in the R.H.S., which will cause decryption to be incorrect with high probability.
-Thus, with high probability, decryption will be correct if and only if $s_i$ is even.
-
-\subsection*{(2)}
-Eve can recover the parity of all entries of the secret $\mathbf{s} \in \chi_s^n$ by preparing $n$ emails $\{(\mathbf{c}_{(i, 1)}, c_{(i, 2)})\}_{i=1}^{n}$ where $(\mathbf{c}_{(i, 1)}, c_{(i, 2)})$ is generated using the procedure described in part (1): $\mathbf{e}_{(i, 1)}$ is all 0's except for the i-th entry, which is set to $\round{\frac{q}{2}}$.
-For each of these $n$ emails, Eve sends it to Alice's server and receives the auto-response.
-The quoted email in the auto-response is checked against the original email: if they are identical, then Alice's server correctly decrypted Eve's email, so the corresponding entry in $\mathbf{s}$ is even; otherwise, the corresponding entry in $\mathbf{s}$ is odd.
-
-\subsection*{(3)}
-We will describe an algorithm for recovering the value of a single component of the secret key $\mathbf{s}$.
-From here, the algorithm can be repeated $n$ times to recover the value of all components of the secret key.
-If the algorithm for recovering a single value is efficient, then repeating it $n$ times is also efficient.
-
-Consider the ciphertext $(\mathbf{c}_1, c_2)$ where all entries of $\mathbf{c}_1$ is set to 0 except for the i-th entry, which we denote by $\mathbf{c}_{(1, i)}$.
-Recall from the decryption procedure the fact that decryption outputs $0$ if and only if $-\round{\frac{q}{4}} \leq m^\prime \leq \round{\frac{q}{4}}$, which is equivalent to:
-
-$$
--\round{\frac{q}{4}} + c_2 \leq c_{(1, i)} \cdot s_i \leq \round{\frac{q}{4}} + c_2
-$$
-
-Where $s_i$ is the i-th entry of the secret $\mathbf{s}$.
-
-In other words, we can check whether $s_i$ falls within a certain range of values given some values of $c_2$ and $c_{(1, i)}$.
-In addition, changing the value of $c_2$ will translate the range, while changing the value of $c_{(1, i)}$ will scale the range.
-Thus, we can perform a binary search by adjusting the values of $c_2, c_{(1, i)}$ and pinpoint the value of $s_i$ in $O(\log{q})$ steps.
-
-\subsection*{(4)}
-Converting textbook LWE to be resistant to chosen-ciphertext attack is non-trivial.
-In Kyber/Dilithium a tweaked version of Fujisaki-Okamoto transformation is applied to the textbook LWE to obtain a CCA2-secure cryptosystem.
-Due to time constraint, the details will not be covered in this write-up.
-
 \end{document}

### 2024-07-14 archives/co789/homework-03/q2/main.tex
What's next

--- 2024-02-21 co789/homework-03/q2/main.tex
@@
 \subsection*{(2)}
-Observe that when $m = \mathbf{0} \in R_{\{0,1\}}$, $c_2 = \mathbf{r}^\intercal \mathbf{b}$ is divisible by $\mathbf{b}$.
-On the other hand, if $m$ is non-zero, then with very high probability, $c_2$ will not be divisible by $\mathbf{b}$.
+When the second error term $e^{\prime\prime}$ is omitted, in the IND-CPA RLWE encryption scheme, the second part of each ciphertext is:
 
-Thus, an IND-CPA can challenge plaintexts as follows: $m_0 = \mathbf{0}$, $m_1$ is some arbitrary non-zero polynomial.
-Upon receiving the challenge ciphertext $c^\ast$, check if $c_2$ is divisible by $\mathbf{b}$ (which is part of the public key and hence accessible).
-If yes, then $c^\ast$ is the encryption of $m_0$; otherwise $c^\ast$ is the encryption of $m_1$.
+$$
+c_2 = rb + m \round{\frac{q}{2}}
+$$
+
+Assuming that $b$ is an invertible polynomial in $R_q = \mathbb{Z}_q / \langle p(x) \rangle$, an adversary who knows $m$ can recover $r$:
+
+$$
+r = (c_2 - m \round{\frac{q}{2}})b^{-1}
+$$
+
+Where $m = 0$ is the zero polynomial, then $c_2b^{-1} = r$ must have small coefficients.
+On the other hand, assuming that decisional RLWE holds, $b$ is computationally indistinguishable from truly random, which implies that $b^{-1}$ is also computationally indistinguishable from truly random.
+Therefore, it is safe to assume that it is easy to find some $m \in R_{\{0, 1\}}$ such that $m\round{\frac{q}{2}}b^{-1}$ has coefficients that are outside the bound of the secret distribution $\chi_s$.
+
+From here, the IND-CPA adversary proceeds as follows.
+Pick $m_0 = 0$, and $m_1$ such that $m_1 \round{\frac{q}{2}}b^{-1}$ has large coefficients.
+Upon receving the challenge ciphertext $c^\ast  =(c_1^\ast, c_2^\ast)$, compute $c_2^\ast b^{-1}$.
+Where $m_0$ is chosen by the challenger, $c_2^\ast b^{-1}$ will have small coefficients.
+Where $m_1$ is chosen by the challenger, then $c_2^\ast b^{-1}$ will have large coefficients.
 

### 2024-07-14 archives/co789/homework-03/q3/main.tex
What's next

--- 2024-02-21 co789/homework-03/q3/main.tex
@@
 
-Observe the following:
+Without rounding $\mathbf{w}$, the public key is $(A, \mathbf{t} = A\mathbf{s} + \mathbf{e})$ the signature contains three components:
 
@@
 \begin{aligned}
-A\mathbf{z} &= A(\mathbf{y} + c\mathbf{s}) \\
-&= \mathbf{w} + cA\mathbf{s} \\
-&= \mathbf{w} + c(\mathbf{t} - \mathbf{e})
+    \mathbf{z} = \mathbf{y} + c \mathbf{s} \\
+    c = H(\mathbf{w}, m) \\
+    \mathbf{w} = A\mathbf{y}
 \end{aligned}
@@
 
-Assuming that $c$ is invertible, re-arranging the equation above gives us:
+Multiply $\mathbf{z}$ by $A$:
+
+$$
+\begin{aligned}
+    A\mathbf{z} &= A\mathbf{y} + c A\mathbf{s} \\
+    &= \mathbf{w} + c \cdot (\mathbf{t} - \mathbf{e})
+\end{aligned}
+$$
+
+Assuming that $c$ is invertible, re-arranging the equation above allows us to recover $\mathbf{e}$ from the publis key $A, \mathbf{t}$ and a single pair of message and signature $(m, (\mathbf{w}, c, \mathbf{z}))$:
 
@@
 From here we can attempt to recover the secret key $\mathbf{s}$ by solving $A\mathbf{s} = \mathbf{t} - \mathbf{e}$.
-While this is an instance of inhomogeneous SIS, with proto-Dilithium $A$ is a wide matrix, which makes SIS easier to solve.
-In the lecture notes, we simply assume that wide SIS is "too easy".
+While this is a non-trivial instance of inhomogeneous SIS, with proto-Dilithium $A$ is a wide matrix, which makes SIS easier to solve.
+In the lecture notes, we simply assume that for the choice of parameters in proto-Dilihtium, such wide ISIS is "too easy".
 

### 2024-07-14 archives/co789/homework-03/q4/main.tex
What's next

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\adv}{\mathcal{A}_\text{EF-CMA}}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 4}
\subsection*{(1)}
We denote the augmented signature scheme's parameters and functions with a star to differentiate them from the parameters and routines of the input signature scheme.

For key generation, set $\operatorname{sk}^\ast = (\operatorname{pk}, \operatorname{sk})$ and $\operatorname{pk}^\ast = H(\operatorname{pk})$, where $\operatorname{pk}, \operatorname{sk} \leftarrow \operatorname{KeyGen}$ is generated from the input signature scheme, and $H$ is the input collision-resistant hash function.

For $\operatorname{Sign}^\ast(\operatorname{sk}^\ast, m)$, first compute $\sigma = \operatorname{Sign}(\operatorname{sk}, m)$ using the input signature scheme's signing routine, then output $\sigma^\ast = (\operatorname{pk}, \sigma)$ as the signature.

For $\operatorname{Verify}^\ast(\operatorname{pk}^\ast, \sigma^\ast, m)$, first unpack the signature $(\hat{\operatorname{pk}}, \hat{\sigma}) = \sigma^\ast$ and check that $H(\hat{\operatorname{pk}})$ is equal to $\operatorname{pk}^\ast$. Then, run the input signature scheme's verification routine $\operatorname{Verify}(\hat{\operatorname{pk}}, \hat{\sigma}, m)$. The verification passes if and only if both checks pass.

\subsection*{(2)}
We show that the modified signature scheme is EUF-CMA by showing that if there exists an EF-CMA adversary for the modified scheme $\adv^\ast$, then we can build an EF-CMA adversary for the original scheme $\adv$ with equal advantage.

In the EF-CMA game of the input scheme, key generation outputs the keypair $(\operatorname{pk}, \operatorname{sk})$. $\adv$ computes $\operatorname{pk}^\ast = H(\operatorname{pk})$ and passes $\operatorname{pk}^\ast$ to $\adv^\ast$.

When $\adv^\ast$ queries the signature of some message $m_i$, $\adv$ queries the signature $\sigma_i$ of $m_i$ from the signing oracle for the input signature scheme. $\adv$ then gives $\sigma^\ast_i = (\operatorname{pk}, \sigma_i)$ back to $\adv^\ast$ as the answer to the query.

When $\adv^\ast$ outputs the forgery $\hat{\sigma^\ast} = (\hat{\operatorname{pk}}, \hat{\sigma}, \hat{m})$, we claim that $\hat{\operatorname{pk}} = \operatorname{pk}$, because otherwise we will have found collision $\hat{\operatorname{pk}} \neq \operatorname{pk}$ such that $H(\hat{\operatorname{pk}}) = H(\operatorname{pk})$. Thus $\adv^\ast$ is valid if and only if $\hat{\sigma}, \hat{m}$ pass the verification of the original signature scheme. Therefore, $\adv$ has the same advantage as $\adv^\ast$.

\end{document}

### 2024-07-14 archives/co789/homework-03/q5/main.tex
What's next

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage{algorithm}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 5}
Recall from lecture notes the proto-Dilithium:

\begin{algorithm}
\caption{proto-Dilithium KeyGen}
\begin{algorithmic}[1]
    \Require $k, l \in \mathbb{Z}$, polynomial ring $R_q$, secret distribution $\chi_s = \mathcal{B}(n, p)$, $y$ distribution $\chi_y = R_{[-\gamma, \gamma]}$, and hash function $H$ that hashes from the message space to $R_{0,1}$ with exactly $\tau$ coefficients being ones

    \State $\mathbf{s} \leftarrow \chi_s^l$
    \State $A \leftarrow R_q^{k \times l}$
    \State $\mathbf{t} \leftarrow A\mathbf{s}$

    \Return $\text{sk} \leftarrow \mathbf{s}, \text{pk} \leftarrow (A, \mathbf{t})$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
    \caption{Dilithium.Sign}
    \begin{algorithmic}[1]
        \Require $\text{pk}, \text{sk}, m$
        \State $\sigma \leftarrow \bot$
        \While{$\sigma = \bot$}
            \State $\mathbf{y} \leftarrow \chi_y^l, \mathbf{w} \leftarrow A\mathbf{y}$
            \State $c \leftarrow H(m)$
            \State $\mathbf{z} \leftarrow \mathbf{y} + c \cdot \mathbf{s}$
            \If{$\norm{\mathbf{z}}_\infty \geq \gamma - np\tau$}
                \State $\sigma \leftarrow \bot$
            \Else
                \State $\sigma \leftarrow (\mathbf{z}, c, \mathbf{w})$
            \EndIf
        \EndWhile
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}
    \caption{Dilithium.Verify}
    \begin{algorithmic}[1]
        \Require $\text{pk}, m, \sigma$
        \State assert $\norm{\mathbf{z}}_\infty < \gamma - np\tau$
        \State assert $A\mathbf{z} = \mathbf{w} + c \mathbf{t}$ 
    \end{algorithmic}
\end{algorithm}

\subsection*{(1)}
Observe that $A, \mathbf{t}$ are both public information. In addition, any adversary can sample $\mathbf{y} \leftarrow \chi_y$ on its own, and produce a valid $\mathbf{w}$. For any chosen message $m$, an adversary can sample $\mathbf{w}$  and compute $c \leftarrow H(\mathbf{w}, m)$, then pass $(A, \mathbf{w} + c \mathbf{t})$ to the provided module-ISIS solver. The output from this solver, denoted by $\mathbf{z}$, is such that

$$
\begin{cases}
    A\mathbf{z} = \mathbf{w} + c\mathbf{t} \\
    \norm{\mathbf{z}}_\infty < \gamma - np\tau
\end{cases}
$$

In other words, $(m, (\mathbf{w}, c, \mathbf{z}))$ will pass the verification for the challenge keypair, making it a valid forgery.

\subsection*{(2)}
Knowing that $\mathbf{y}$'s infinity norm is bounded by $\gamma$, we can use the given module-ISIS solver to recover $\mathbf{y}$ from the relation $\mathbf{w} = A \mathbf{y}$. Once $\mathbf{y}$ is recovered, it is possible to recover the secret key $\mathbf{s}$ using the relation $\mathbf{z} = \mathbf{y} + c \mathbf{s}$.

Note that because $A$ is wide, $A\mathbf{y} = \mathbf{w}$ might not have unique solution. We can speculate the condition on which solutions are unique. One sufficient condition is that the linear mapping is injective, which requires that the domain is smaller than the co-domain. The domain of the linear mapping is $R_{[-\gamma, \gamma]}^l$ ($l$ polynomials of degree $n$ each having coefficients in the range $[-\gamma, \gamma]$), whose size is approximately:

$$
\vert R_{[-\gamma, \gamma]}^l \vert = (2\gamma + 1)^{ln}
$$

On the other hand, there is no restriction on $\mathbf{w} \in R_q^k$:

$$
\vert R_q^k \vert = q^{nk}
$$

For the linear mapping to be injective, $\vert R_{[-\gamma, \gamma]}^l \vert < \vert R_q^k \vert$, which results in a bound on $\gamma$:

$$
\gamma < \frac{1}{2}(q^\frac{k}{l} - 1)
$$

Finally, here is the key recovery attack:

\begin{algorithm}
    \caption{Key-recovery attack under CMA}
    \begin{algorithmic}[1]
        \Require $\text{pk}$, module-ISIS solver
        \State Generate random message $m$, and query its signature $\sigma = (\mathbf{w}, c, \mathbf{z})$
        \State Use module-ISIS solver to solve for $\mathbf{y}$ in $A\mathbf{y} = \mathbf{w}$
        \State Compute $\mathbf{s} \leftarrow (\mathbf{z} - \mathbf{y}) \cdot c^{-1}$

        \Return $\mathbf{s}$ is the secret key
    \end{algorithmic}
\end{algorithm}

\end{document}

### 2024-07-14 archives/co789/homework-04/q3/main.tex
What's next

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage{algorithm}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 3}
Recall that the FORS scheme uses $k$ Merkle trees $T_1, T_2, \ldots, T_k$ each with $n$ leaf nodes. At signing $m$ is hashed into $1 \leq h_1, h_2, \ldots, h_k \leq n$, and the signature are the authentication paths $\sigma_j$ of $h_j$ in tree $T_j$ for $1 \leq j \leq k$.

Let $\mathcal{A}$ denote a EF-CMA adversary. Suppose that $\mathcal{A}$ makes $N$ queries: $m_1, m_2, \ldots, m_N$, then the signing oracles will return the authentication paths for:

$$
\begin{bmatrix}
    h_{1, 1}, &h_{1, 2}, &\ldots, &h_{1, N}, &\text{from } T_1 \\
    h_{2, 1}, &h_{2, 2}, &\ldots, &h_{2, N}, &\text{from } T_2 \\
    \ldots \\
    h_{k, 1}, &h_{k, 2}, &\ldots, &h_{k, N}, &\text{from } T_k \\
\end{bmatrix}
$$

Let $m^\ast$ be some randomly chosen message, and $h_1^\ast, h_2^\ast, \ldots, h_k^\ast = H(m)$. $m^\ast$ is forgeable if the authentication paths in all $k$ trees are forgeable. An authentication path in tree $T_i$ is forgeable if any of $h_{i,1}, h_{i,2}, \ldots, h_{i, N}$ collides with $h_i^\ast$. Therefore:

$$
\begin{aligned}
    P[\text{$T_i$ auth path forgeable}] 
    &= 1 - P[\text{$T_i$ auth path unforgeable}] \\
    &= 1 - P[h_{i,j} \neq h_i^\ast \text{ for all } 1 \leq j \leq N] \\
    &= 1 - \prod_{j=1}^N P[h_{i,j} \neq h_i^\ast] \\
    &= 1 - (1 - \frac{1}{n})^N
\end{aligned}
$$

Where $P[h_{i,j} \neq h_i^\ast] = 1 - \frac{1}{n}$ because each hash $H(m)$ must be in the range $1, 2, \ldots, n$, and we assume that hash function to be an ideal pseudorandom function.

Finally:

$$
\begin{aligned}
P[\text{FORS forgeable}]
&= P[\text{$T_i$  authentication path forgeable for all $1 \leq i \leq k$}] \\
&= \prod_{i=1}^k P[\text{$T_i$ auth path forgeable}] \\
&= (1 - (1 - \frac{1}{n})^N)^k
\end{aligned}
$$

\end{document}

### 2024-07-14 archives/co789/homework-04/q4/main.tex
What's next

--- 2024-07-14 archives/co789/homework-04/q3/main.tex
@@
 
-\section*{Question 3}
-Recall that the FORS scheme uses $k$ Merkle trees $T_1, T_2, \ldots, T_k$ each with $n$ leaf nodes.
-At signing $m$ is hashed into $1 \leq h_1, h_2, \ldots, h_k \leq n$, and the signature are the authentication paths $\sigma_j$ of $h_j$ in tree $T_j$ for $1 \leq j \leq k$.
+\section*{Question 4}
+With a ternary Merkle tree (3MT for short), each piece of data being committed will have 2 sibling nodes at each layer in the 3MT.
+If the 3MT has $3^t$ leaf nodes, then the authentication paths will contain $2t$ hashes (compared to $t$ hashes in regular MT).
 
-Let $\mathcal{A}$ denote a EF-CMA adversary. Suppose that $\mathcal{A}$ makes $N$ queries: $m_1, m_2, \ldots, m_N$, then the signing oracles will return the authentication paths for:
+In the SPX hypertree, each layer produces one WOTS signature and one Merkle tree authentication path. If we use 3MT for SPX then:
 
 $$
-\begin{bmatrix}
-    h_{1, 1}, &h_{1, 2}, &\ldots, &h_{1, N}, &\text{from } T_1 \\
-    h_{2, 1}, &h_{2, 2}, &\ldots, &h_{2, N}, &\text{from } T_2 \\
-    \ldots \\
-    h_{k, 1}, &h_{k, 2}, &\ldots, &h_{k, N}, &\text{from } T_k \\
-\end{bmatrix}
+\vert \sigma^\text{SPX} \vert = d(
+    \vert \sigma^\text{WOTS} \vert + 2t \vert H \vert
+) 
++ \vert \sigma^\text{FORS} \vert
 $$
 
-Let $m^\ast$ be some randomly chosen message, and $h_1^\ast, h_2^\ast, \ldots, h_k^\ast = H(m)$.
-$m^\ast$ is forgeable if the authentication paths in all $k$ trees are forgeable.
-An authentication path in tree $T_i$ is forgeable if any of $h_{i,1}, h_{i,2}, \ldots, h_{i, N}$ collides with $h_i^\ast$.
-Therefore:
+where $H$ is the hash function used in the Merkle tree.
 
-$$
-\begin{aligned}
-    P[\text{$T_i$ auth path forgeable}] 
-    &= 1 - P[\text{$T_i$ auth path unforgeable}] \\
-    &= 1 - P[h_{i,j} \neq h_i^\ast \text{ for all } 1 \leq j \leq N] \\
-    &= 1 - \prod_{j=1}^N P[h_{i,j} \neq h_i^\ast] \\
-    &= 1 - (1 - \frac{1}{n})^N
-\end{aligned}
-$$
-
-Where $P[h_{i,j} \neq h_i^\ast] = 1 - \frac{1}{n}$ because each hash $H(m)$ must be in the range $1, 2, \ldots, n$, and we assume that hash function to be an ideal pseudorandom function.
-
-Finally:
-
-$$
-\begin{aligned}
-P[\text{FORS forgeable}]
-&= P[\text{$T_i$  authentication path forgeable for all $1 \leq i \leq k$}] \\
-&= \prod_{i=1}^k P[\text{$T_i$ auth path forgeable}] \\
-&= (1 - (1 - \frac{1}{n})^N)^k
-\end{aligned}
-$$
+Note that if we use 3MT, then each XMSS will have $3^t$ leaf nodes, and the entire hypertree will be able to sign $3^{dt}$ distinct messages.
+With a fixed security level $\lambda$ and $t$, the value of $d$ will decreased, meaning that we will need fewer layers in the hypertree (so the signature size is not strictly increasing).
 

### 2024-07-14 archives/co789/homework-05/q4/main.tex
What's next

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 4}

\subsection*{4.1}
We show equivalence by showing that having an oracle solver for one problem allows us to solve the other problem.

In the forward direction, let $G \in \mathbb{F}^{n \times k}_2$ but a generator matrix, $c \in \mathbb{F}^n_2$ be some partially corrupted codeword, and $\mathcal{O}^S$ be the syndrome decoding oracle. If we can find a corresponding parity-check matrix $H$, then $y \leftarrow Hc \in \mathbb{F}_2^{n-k}$ is a syndrome. We can then feed $H, y$ to $\mathcal{O}^S$ and obtain the error term $e$. Solving the linear system $Gm = c - e$ allows us to recover the message $m$.

To compute the parity check matrix, we assume that the generator matrix $G$ has the following form:

$$
G = \begin{bmatrix}
    I_k \\
    G^\prime
\end{bmatrix}
$$

where $G^\prime \in \mathbb{F}_2^{(n-k) \times k}$. This is a reasonable assumption because if $G$ does not have this form, we can column-reduce $G$ to have this form. Column-reducing $G$ is guaranteed to produce $I_k$ on the top $G$ has column rank $k$ (otherwise the code will map distinct messages onto the same codeword, which cannot happen). Column-reducing $G$ is equivalent to right-multiplication by an invertible matrix, which does not affect the set of codewords in the code $\mathcal{C}$.

Let $H = [-G^\prime \mid I_{n-k}]$, then $H$ is a rank-$(n-k)$ matrix such that $HG = 0$, meaning that $H$ is a parity check matrix that corresponds with $G$.

$$ $$

In the backward direction, let $H \in \mathbb{F}_2^{(n-k) \times n}$ be a parity check matrix, $y \in \mathbb{F}_2^{n-k}$ be a syndrome, and $\mathcal{O}^C$ be a codeword decoding oracle. We can row reduce $H$ to have the form $H = [I_{n-k} \mid H_0]$ where $H_0 \in \mathbb{F}_2^{(n-k) \times k}$. Since row reducing $H$ is equivalent to left multiplication by an invertible matrix, the row-reduced $[I_{n-k} \mid H_0]$ is also a parity check matrix for the same code.

Let $G$ be as follows:

\begin{equation*}
    G = \begin{bmatrix}
        -H_0 \\ I_k
    \end{bmatrix}
\end{equation*}

Then $G \in \mathbb{F}_2^{n \times k}$ is a rank-$k$ matrix such that $HG = 0$. In other words, $H$ is the parity check matrix of a linear code generated by $G$. If we solve the linear system $Hz = y$ for $z$, then $z$ is some partially corrupted codeword in the linear code. Let $m \leftarrow \mathcal{O}^C(G, z)$. By the definition of the codeword decoding problem $e = z - Gm$ is the error term.

\subsection*{4.2}
Again, we will show equivalence by showing that a solver for one problem allows us to solve the other problem.

In the forward direction, let $A \in \mathbb{F}_2^{n \times k} = PGS$ and $c \in \mathbb{F}_2^n = Am + e$ be what they are in the McEliece codeword decoding problem. Let $\mathcal{O}^S$ be a McEliece syndrome decoding oracle. If we can find the corresponding $H_0$ such that $[I_{n-k} \mid H_0] = SHP$ for some parity check matrix $H$, then we can compute $y \leftarrow [I_{n-k} \mid H_0]c$, feed $H_0, y$ to $\mathcal{O}^S$ to obtain the error term $e$, then solve the linear system $Am = c - e$ to recover the message $m$, which is the solution to the McEliece codeword decoding problem.

To compute $H_0$, we first column-reduce $A$ such that

$$
A = \begin{bmatrix}
    I_k \\ A^\prime
\end{bmatrix}
$$

where $A^\prime \in \mathbb{F}_2^{(n-k) \times k}$. Column-reducing $A$ is equivalent to right multiplication by an invertible matrix, so the column-reduced $A$ still corresponds to the same generator matrix $G$. In addition, both $P, S$ are invertible, so $A$ is guaranteed to have rank $k$.

Let $\tilde{H} = [-A^\prime \mid I_{n-k}]$, then $\tilde{H}A = \tilde{H}PGS = 0$, which means that $H = \tilde{H}P$ is a parity check matrix for $G$. If we then row reduce $\tilde{H} = HP^{-1}$:

\begin{equation*}
    \tilde{S}\tilde{H} = \tilde{S}HP^{-1} = [I_{n-k} \mid H_0]
\end{equation*}

Since $\tilde{S}$ is invertible and $P^{-1}$ is also a permutation, $H_0$ is indeed the matrix used in the McEliece syndrome decoding.

$$ $$

In the other direction, let $H_0 \in \mathbb{F}_2^{(n-k) \times k}$ and $y \in \mathbb{F}_2^{n-k}$ be what they are in the McEliece syndrome decoding problem. Let $\mathcal{O}^C$ be a McEliece codeword decoding oracle. Let $\tilde{H} = [I_{n-k} \mid H_0]$, then solve the linear system $\tilde{H}z = y$ for $z$.

Let $\tilde{A}$ be as follows:

\begin{equation*}
    \tilde{A} = \begin{bmatrix}
        -H_0 \\ I_k
    \end{bmatrix}
\end{equation*}

Then $\tilde{H}\tilde{A} = 0$. By the definition of the McEliece syndrome decoding problem, we know that $\tilde{H} = SHP$ for some permutation $P$ and some invertible matrix $S$, which means that $SHP\tilde{A} = 0$. Since $\tilde{A}$ has k-rank and $P$ is invertible, $P\tilde{A}$ has rank-k. In other words, $P\tilde{A} = G$ for some generator matrix $G$ that corresponds with $H$. Re-arranging this equation: $\tilde{A} = P^{-1}GI_k$ is a valid McEliece codeword decoding public key.

Let $m \leftarrow O^C(\tilde{A}, z)$, then $e = z - Gm$ is the error term and the solution to the McEliece syndrome decoding problem.

\end{document}

### 2024-07-14 archives/co789/homework-05/q7/main.tex
What's next

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 7}
Recall the Beaver triple protocol for computing $xy$ given $[x], [y]$:

\begin{enumerate}
    \item A trusted third party generates random $a, b \leftsample \mathbb{Z}_n^\ast$ and compute $c = ab$
    \item This third party distributes $[a], [b], [c]$ to the computation parties. We claim that since the circuit of the computation is known ahead of time, the trusted parties can generate and distributes $[a], [b], [c]$ before the MPC actually starts. Therefore, distributing $[a], [b], [c]$ does not count toward online communication cost during the MPC
    \item Each party computes $[x-a], [y-b]$, broadcasts their shares, which publicizes $\epsilon = x-a, \delta = y-b$. \textbf{This invokes 2 units of online communication}
    \item Each party computes $[z] = \delta[x] + \epsilon[y] - \epsilon\delta + [c]$ which evaluates to $[xy]$
    \item Each party opens $[z]$, which assembles back into $xy$
\end{enumerate}

When computing the square of a number, we don't need to obfuscate two distinct operands $x, y$; instead, we only need to obfuscate one operand $x$, so we only need to generate a single one-time pad $a$ and its square $c = a^2$. This means that individual parties only need to broadcast $[x-a]$ during their computation, bring the units of online communication during MPC from 2 to 1.

Here is the full protocol

\begin{enumerate}
    \item A trusted third party generates random $a \leftsample \mathbb{Z}_n^\ast$ and computes $c = a^2$
    \item The shares $[a], [c]$ are pre-distributed to the computation parties.
    \item Each party computes $[x-a]$ and boardcasts the value of their share, which publicizes the value of $x-a$. \textbf{This invokes 1 unit of online communication}.
    \item Each party computes $[z] = 2\cdot (x-a)[x] - (x-a)^2 + [c]$, which evaluates to $[x^2]$
    \item Each party opens $[z]$, which assembles back into $x^2$
\end{enumerate}

Proof of step 4:

$$
\begin{aligned}
    \lbrack z \rbrack &= 2\cdot (x-a)[x] - (x-a)^2 + [c] \\
    &= [2x^2 - 2ax] - (x^2 - 2ax + a^2) + [a^2] \\
    &= [2x^2 - 2ax - (x^2 - 2ax + a^2)] + [a^2] \\
    &= [x^2 - a^2] + [a^2] \\
    &= [x^2]
\end{aligned}
$$

\end{document}

### 2024-07-14 archives/co789/project/manuscript/main.tex
What's next

--- 2024-04-17 co789/project/manuscript/main.tex
@@
 
-\title{A survey of IND-CCA constructions}
+\title{An introduction to Fujisaki-Okamoto transformation}
 \author{Ganyu (Bruce) Xu (g66xu)}
@@
 
-\subsection{IND-CPA to OW-CPA}
+\subsection{IND-CPA to OW-CPA}\label{ind-cpa-to-ow-cpa}
 It is trivially true that if a public-key encryption scheme is not one-way secure, then it is not indistinguishable.
@@
 \section{Related works}
+\subsection{OAEP and plaintext awareness}
+    One of the early attempts at an IND-CCA secure public-key encryption scheme is RSA.
+Unfortunately, the first deployment of RSA in PKCS\#1 v1.5 \cite{kaliski1998rfc2313} is proved vulnerable to chosen-ciphertext attacks by Bleichenbacher \cite{bleichenbacher1998chosen}.
+Subsequent work by Bellare and Rogaway \cite{bellare1995optimal} proposed the Optimal Asymmetric Encryption Padding (OAEP) as a generic IND-CCA secure transformation.
+Bellare and Rogaway argued for the IND-CCA security of OAEP through the framework of ``plaintext awareness'', which states that the ciphertext cannot be partially decrypted without inverting some cryptographic trapdoor functions.
+Later on OAEP is shown to be only IND-CCA1 secure, although Fujisaki, et all showed that RSA-OAEP is indeed IND-CCA2 secure \cite{fujisaki2001rsa}.
+
+\subsection{In the context of key exchange}
     Key encapsulation mechanisms (KEM) usually have weaker security requirements than public-key encryption schemes (PKE) thanks to the fact that properly designed KEMs obfuscate the plaintext through a key-derivation function (which is usually a hash function and thus a random oracle under the RO assumption) in the decryption routine.
@@
 
-    \begin{multicols}{3}
-        \begin{algorithm}[H]
-            \SetAlgoLined
-            \caption{KeyGen}
-            $(\text{pk}, \text{sk}) \leftsample \operatorname{KeyGen}()$ \;
-            \Return{$(\text{pk}, \text{sk})$}
-        \end{algorithm}
+    % \begin{multicols}{3}
+    %     \begin{algorithm}[H]
+    %         \SetAlgoLined
+    %         \caption{KeyGen}
+    %         $(\text{pk}, \text{sk}) \leftsample \operatorname{KeyGen}()$ \;
+    %         \Return{$(\text{pk}, \text{sk})$}
+    %     \end{algorithm}
 
-        \begin{algorithm}[H]
-            \SetAlgoLined
-            \caption{Encap}
-            \KwIn{$\text{pk}$}
-            $m \leftsample \mathcal{M}$\;
-            $c \leftsample E(\text{pk}, m)$\;
-            $K \leftarrow H(m, c)$\;
-            \Return{$(c, K)$}\;
-        \end{algorithm}
+    %     \begin{algorithm}[H]
+    %         \SetAlgoLined
+    %         \caption{Encap}
+    %         \KwIn{$\text{pk}$}
+    %         $m \leftsample \mathcal{M}$\;
+    %         $c \leftsample E(\text{pk}, m)$\;
+    %         $K \leftarrow H(m, c)$\;
+    %         \Return{$(c, K)$}\;
+    %     \end{algorithm}
 
-        \begin{algorithm}[H]
-            \SetAlgoLined
-            \caption{Decap}
-            \KwIn{$\text{sk}, c \in \mathcal{C}$}
-            $\hat{m} \leftarrow D(\text{sk}, c)$\;
-            $\hat{K} \leftarrow H(\hat{m}, c)$\;
-            \Return{$\hat{K}$}\;
-        \end{algorithm}
-    \end{multicols}
+    %     \begin{algorithm}[H]
+    %         \SetAlgoLined
+    %         \caption{Decap}
+    %         \KwIn{$\text{sk}, c \in \mathcal{C}$}
+    %         $\hat{m} \leftarrow D(\text{sk}, c)$\;
+    %         $\hat{K} \leftarrow H(\hat{m}, c)$\;
+    %         \Return{$\hat{K}$}\;
+    %     \end{algorithm}
+    % \end{multicols}
 
@@
 The encryption and decryption routines are as follows:
+
+% \pagebreak
 
@@
 
-\subsubsection{Plaintext checking, ciphertext validation, and patched decapsulation oracle}
+\subsubsection{Plaintext checking, ciphertext validation, and patched decapsulation oracle}\label{patched-decap}
 The concept of plaintext-checking attack (PCA) was proposed in early 2000s and first used in generic IND-CCA transformation by Coron et al in \textit{GEM: a generic chosen-ciphertext secure encryption method} \cite{jean2002gem}.
@@
 \subsection{A sketch of proof}
-Here we provide a sketch of the security proof for both the 1999 transformation and the 2017 variations.
+Here we provide a sketch of the security proof for both the 1999 transformation and the 2017 variations. All three proofs follow the framework of sequence of games.
 
@@
 \begin{proof}
+    Game 0 is the standard IND-CCA2 game.
 
+    Game 1 is identical to Game 0, except for that the standard decryption oracle is replaced with the simulated decryption oracle described in section \ref{simulate-dec-oracle}.
+From the IND-CCA adversary's perspective, game 1 and game 0 are indistinguishable, except when the adversary makes one or more ciphertext query $(\tilde{c}, \tilde{e})$ to the decryption oracle without matching hash query $(m, c, r) \in \mathcal{O}^H$ such that $c = \tilde{c}$ and $E^\text{asym}(\text{pk}, m, r) = \tilde{c}$.
+Given that the adversary makes exactly $q_D$ decryption queries, the probability of ``at least one divergent query'' is:
+
+    \begin{equation*}
+        \begin{aligned}
+            P[\text{at least 1 diverging decryption query}]
+            &= P[\cup_{i=1}^{q_D} \text{ query $i$ is divergent}] \\
+            &\leq \sum_{i=1}^{q_D}P[\text{query $i$ is divergent}] \\
+            &= q_D2^{-\gamma}
+        \end{aligned}
+    \end{equation*}
+
+    By the difference lemma, we know:
+
+    \begin{equation}
+        \epsilon_0 - \epsilon_1 
+        \leq P[\text{at least 1 diverging decryption query}] 
+        \leq q_D2^{-\gamma}
+    \end{equation}
+
+    Where $\epsilon_0$ and $\epsilon_1$ is the adversary's advantage in game 0 and game 1, respectively.
+
+    Game 2 is identical to game 1, except for that in the challenge encryption, the both the symmetric key and the asymmetric coin are truly random instead of pseudorandom:
+
+    \begin{multicols}{2}
+        \begin{algorithm}[H]
+            \SetAlgoLined
+            \caption{$G_1$ challenge encryption}
+            \KwIn{$\text{pk}, m \in \mathcal{M}^\text{sym}$}
+            $\sigma \leftsample \mathcal{M}^\text{asym}$\;
+            $a \leftarrow G(\sigma)$; $c \leftarrow E^\text{sym}_a(m)$\;
+            $r \leftarrow H(\sigma, c)$;
+            $e \leftarrow E^\text{asym}(\text{pk}, m, r)$\;
+            \Return{$(c, e)$}\;
+        \end{algorithm}
+        \columnbreak
+
+        \begin{algorithm}[H]
+            \SetAlgoLined
+            \caption{$G_2$ challenge encryption}
+            \KwIn{$\text{pk}, m \in \mathcal{M}^\text{sym}$}
+            $\sigma \leftsample \mathcal{M}^\text{asym}$\;
+            $a \leftsample \mathcal{K}^\text{sym}$; $c \leftarrow E^\text{sym}_a(m)$\;
+            $r \leftsample \text{Coin}$;
+            $e \leftarrow E^\text{asym}(\text{pk}, m, r)$\;
+            \Return{$(c, e)$}\;
+        \end{algorithm}
+    \end{multicols}
+
+    From the adversary's perspective, game 2 is identical to game 1 except for when the adversary queries either $G$ or $H$ with the $\sigma$ used in the challenge encryption:
+
+    \begin{equation}
+        \epsilon_1 - \epsilon_2 \leq P[\operatorname{QUERY}^\ast]
+    \end{equation}
+
+    Game 2 can be perfectly simulated by an IND-CPA adversary against the underlying symmetric cipher using the technique discussed in section \ref{random-until-queried}:
+
+    \begin{equation}
+        \epsilon_2 = \epsilon^\text{sym}_\text{IND-CPA}
+    \end{equation}
+
+    Game 2 can also be perfectly simulated by an OW-CPA adversary against the underlying asymmetric cipher.
+The OW-CPA adversary will pass the OW-CPA challenge ciphertext as the asymmetric part of the IND-CCA challenge ciphertext.
+After the IND-CCA adversary halts, the OW-CPA adversary will check the tape of the hash oracles and pick a random queried value $\tilde{\sigma}$ as its output.
+The OW-CPA adversary wins if and only if the IND-CCA adversary makes the divergent hash query and that the randomly picked query is the correct one:
+
+    \begin{equation}
+        P[\operatorname{QUERY}^\ast] \cdot \frac{1}{q_H} = \epsilon^\text{asym}_\text{OW-CPA}
+    \end{equation}
+
+    Summing the four equations above will bound the IND-CCA adversary's advantage in game 0.
 \end{proof}
@@
 \subsubsection{Proof of theorem \ref{ow-pcva-result}}
+\begin{proof}
+    Game 0 is the OW-PCVA game, which is identical to the OW-CPA game, but the adversary also has access to the PCO and the CVO.
+
+    Game 1 is identical to game 0, except for that the PCO is replaced with the simulated PCO.
+From the perspective of the OW-PCVA adversary, the two games are indistinguishable except for when the adversary produces a ciphertext $c$ from a plaintext $m$ that causes decryption error.
+Since each call to the encryption routine involves a call to the hash oracle, and assuming that the asymmetric cipher is $\delta$-correct, we can bound the change in the adversary's advantage:
+
+    \begin{equation*}
+        \epsilon_0 - \epsilon_1 
+        \leq P[\text{at least one decryption error}]
+        \leq q_G \cdot \delta
+    \end{equation*}
+
+    Game 2 is identical to game 1, except for that the CVO is replaced with the simulated CVO. Using the technique discussed in section \ref{simulate-dec-oracle}, we have:
+
+    \begin{equation*}
+        \epsilon_1 - \epsilon_2 \leq q_V \cdot 2^{-\gamma}
+    \end{equation*}
+
+    Game 3 is identical to game 2, except for that the challenge encryption is performed using a truly random coin $r^\ast \leftsample \text{Coin}$ instead of the pseudorandom coin $r^\ast \leftarrow G(m)$.
+Using the argument in section \ref{random-until-queried}:
+
+    \begin{equation*}
+        \epsilon_2 - \epsilon_3 \leq P[\operatorname{QUERY}^\ast]
+    \end{equation*}
+
+    Game 3 can be perfectly simulated by a OW-CPA adversary against the underlying asymmetric cipher:
+
+    \begin{equation*}
+        \epsilon_3 = \epsilon_\text{OW-CPA}
+    \end{equation*}
+
+    Game 3 can also be simulated by an IND-CPA adversary against the underlying asymmetric cipher who passes the IND-CPA challenge ciphertext to the OW-PCVA adversary.
+If the OW-PCVA adversary queries $G$ with the value of $m^\ast$, then the IND-CPA adversary will find in the tape of the hash oracle $G$ a value that matches one its own challenge plaintexts.
+If no match is found, the IND-CPA adversary will output a random coin toss.
+Thus, the advantage of this IND-CPA adversary is exactly the probability $P[\operatorname{QUERY}^\ast]$:
+
+    \begin{equation*}
+        P[\operatorname{QUERY}^\ast] = \epsilon_\text{IND-CPA}
+    \end{equation*}
+
+    The four equations above, combined with the result in section \ref{ind-cpa-to-ow-cpa}, will bound the advantage of the OW-PCVA adversary
+\end{proof}
 
 \subsubsection{Proof of theorem \ref{cca-kem-result}}
+\begin{proof}
+    Game 0 is the IND-CCA KEM game.
+
+    Game 1 is identical to game 0, except for that the decapsulation oracle is replaced with the patched hash oracle plus the patched decapsulation oracle.
+Using the argument in section \ref{patched-decap}, from the perspective of the adversary, the two games are indistinguishable:
+
+    \begin{equation*}
+        \epsilon_0 = \epsilon_1
+    \end{equation*}
+
+    Game 2 is identical to game 1, except for that in the challenge encapsulation, the shared secret is truly random $K^\ast \leftsample \mathcal{K}$ instead of pseudorandom $K^\ast \leftarrow H(m, c)$.
+Game 2 diverges from game 1 in ways described in section \ref{random-until-queried}
+
+    \begin{equation*}
+        \epsilon_1 - \epsilon_2 = P[\operatorname{QUERY}^\ast]
+    \end{equation*}
+
+    If the IND-CCA KEM adversary indeed makes a hash query on the challenge values $(m^\ast, c^\ast)$, then an OW-PCVA adversary can check the hash oracle's tape to decrypt a challenge encryption:
+
+    \begin{equation*}
+        P[\operatorname{QUERY}^\ast] = \epsilon_\text{OW-PCVA}
+    \end{equation*}
+
+    Finally, since $K^\ast$ is always truly random, no adversary can have any advantage in game 2:
+
+    \begin{equation*}
+        \epsilon_2 = 0
+    \end{equation*}
+\end{proof}
 
 \section{Future works on generic IND-CCA transformation}
+The 2017 transformations $U^\bot$ and $U^{\not\bot}$ have been largely successful in its simple constructions and robust security result, and the latter is adopted by NIST PQC winners such as Kyber.
+However, there are still continuous work on further improving the Fujisaki-Okamoto transformation.
+Here are a few open questions in the field:
+
+\subsection{Security in the quantum random oracle model}
+The weakest link in all of the transformations and their security reduction above is the random oracle model.
+In addition, as the prospect of quantum computer looms bigger, there is an increasingly urgent need to re-consider the random oracle model in the quantum setting.
+For example, a quantum computer can compute the hash values of a superposition of inputs, although due to the nuance of quantum mechanics that are outside this author's capacity, quantum random oracles are not a straightforward upgrade of computing capabilities from the non-quantum random oracle model.
+Taking a step back, the entire premise of quantum security remains an open question.
+
+\subsection{Better handling of decryption error}
+In the 2017 transformation, decryption error is handled with a shotgun approach: the authors simply assumed that the probability of an adversary coming up with a message $m$ that would cause decryption error is negligible, but in practice this probability is still an non-trivial factor, and some adversary may choose such message $m$ adversarially, thus making the loss of security potentially non-trivial.
+
+\subsection{Removing re-encryption}
+The re-encryption step in the decryption routine is the source of ciphertext integrity and thus an essential component to the IND-CCA security of the Fujisaki-Okamoto transformation.
+Unfortunately, this step is also very computationally expensive (this is especially true in Kyber, whose encryption routine involves computing and storing the entire public-key matrix $A \in R_q^{k \times k}$ but whose decryption routine involves much less computation and memory requirements).
+There might be other ways to provide integrity with better performance characteristics.
+For example, classic McEliece achieves integrity through the unique properties of Goppa code and the syndrome decoding problem, although this approach lacks generality and is not readily applicable to lattice-based schemes like Kyber.
+Symmetric integrity primitives like message authentication code (MAC) are also promising candidates, though the security implications remain unproven.
 

### 2024-07-14 computational-number-theory/src/lib.rs

@@ -1,502 +1,0 @@
-//! My implementation of computational number theory
-//! Follows Henri Cohen's [A Course in Computational Algebraic Number Theory]
-
-use std::{cmp::Ordering, fmt::UpperHex, num::ParseIntError};
-
-/// 32-bit architecture
-pub type Word = u32;
-/// Twice the size of Word, useful for widening multiplication
-pub type LongWord = u64;
-
-/// Equivalent to carrying_add but I have to implement it because carrying_add is not stable
-const fn overflowing_adc(a: Word, b: Word, carry: bool) -> (Word, bool) {
-    // There are two overflowing addition to perform
-    let (mut sum, mut carry_next) = a.overflowing_add(b);
-    if carry {
-        let (carried_sum, carry_2) = sum.overflowing_add(1);
-        carry_next = carry_next || carry_2;
-        sum = carried_sum;
-    }
-
-    return (sum, carry_next);
-}
-
-/// Equivalent to borrowing_sub but borrowing_sub is unstable
-const fn overflowing_sbb(a: Word, b: Word, borrow: bool) -> (Word, bool) {
-    let (mut diff, mut borrow_next) = a.overflowing_sub(b);
-    if borrow {
-        let (borrowed_diff, borrow_2) = diff.overflowing_sub(1);
-        borrow_next = borrow_next || borrow_2;
-        diff = borrowed_diff;
-    }
-
-    return (diff, borrow_next);
-}
-
-fn widening_mul(a: Word, b: Word) -> (Word, Word) {
-    let a: LongWord = a.into();
-    let b: LongWord = b.into();
-    let prod: LongWord = a.wrapping_mul(b);
-    let low: Word = prod as Word;
-    let high: Word = (prod >> 32) as Word;
-
-    return (high, low);
-}
-
-/// Count the number of leading zero bits
-/// TODO: there is probably more efficient implementation but oh well
-const fn leading_zeros(word: Word) -> u32 {
-    let mut cnt = 0;
-    let mask = 1 << 31;
-    let mut shift = 0;
-
-    while shift < 32 {
-        if word & (mask >> shift) == 0 {
-            cnt += 1;
-        } else {
-            return cnt;
-        }
-
-        shift += 1;
-    }
-
-    return cnt;
-}
-
-/// A big integer is an array of words
-/// We will tentatively use little-endian: lower index encodes less significant digits
[…]

### 2024-07-14 computational-number-theory/src/main.rs

@@ -1,6 +1,0 @@
-use compnt::U256;
-
-fn main() {
-    let (quo, rem) = U256::MAX.div_rem_vartime(&U256::ONE);
-    println!("{:X}, {:X}", quo, rem);
-}

### 2024-07-14 computational-number-theory/src/ring.rs

@@ -1,63 +1,0 @@
-//! The integer modulus ring
-use std::error::Error;
-
-use crate::Uint;
-
-/// Various ways integer ring arithmetics can go wrong
-#[derive(Debug, Eq, PartialEq, Copy, Clone)]
-pub enum RingError {
-    /// The modulus of a ring cannot be zero
-    ModulusIsZero,
-}
-
-impl std::fmt::Display for RingError {
-    fn fmt(&self, f: &mut std::fmt::Formatter<'_>) -> std::fmt::Result {
-        write!(f, "{:?}", self)
-    }
-}
-
-impl Error for RingError {}
-
-/// The integer ring is defined by the modulus, which cannot be zero
-#[derive(Debug, Eq, PartialEq, Copy, Clone)]
-pub struct Ring<const L: usize> {
-    pub modulus: Uint<L>,
-}
-
-impl<const L: usize> Ring<L> {
-    pub fn new(modulus: Uint<L>) -> Result<Self, RingError> {
-        if modulus == Uint::<L>::ZERO {
-            return Err(RingError::ModulusIsZero);
-        }
-
-        Ok(Self { modulus })
-    }
-}
-
-/// An element of the modulus
-pub struct RingElem<const L: usize> {
-    pub val: Uint<L>,
-    pub ring: Ring<L>,
-}
-
-impl<const L: usize> RingElem<L> {
-    pub fn new(val: Uint<L>, ring: Ring<L>) -> Self {
-        todo!();
-    }
-
-    pub fn add(&self, other: &Self) -> Self {
-        todo!();
-    }
-
-    pub fn sub(&self, other: &Self) -> Self {
-        todo!();
-    }
-
-    pub fn mul(&self, other: &Self) -> Self {
-        todo!();
-    }
-
-    pub fn modexp(&self, exp: &Uint<L>) -> Self {
-        todo!();
-    }
-}

### 2024-07-14 lattices/primitives/src/algebra.rs

@@ -1,934 +1,0 @@
-//! Modulus and polynomial ring operations
-use core::fmt::{Binary, LowerHex, UpperHex};
-use sha3::{
-    digest::{ExtendableOutput, Update, XofReader},
-    Shake256,
-};
-use subtle::{Choice, ConditionallySelectable};
-
-use crate::{Word, KYBER_ETA_1, KYBER_ETA_2, KYBER_K, KYBER_N, KYBER_Q, SEEDSIZE};
-
-/// zeta (the 256-th primitive root of 3329) and its powers, up to 127
-pub const ZETA_POWS: [FieldElem; 256] = [
-    FieldElem(1),
-    FieldElem(17),
-    FieldElem(289),
-    FieldElem(1584),
-    FieldElem(296),
-    FieldElem(1703),
-    FieldElem(2319),
-    FieldElem(2804),
-    FieldElem(1062),
-    FieldElem(1409),
-    FieldElem(650),
-    FieldElem(1063),
-    FieldElem(1426),
-    FieldElem(939),
-    FieldElem(2647),
-    FieldElem(1722),
-    FieldElem(2642),
-    FieldElem(1637),
-    FieldElem(1197),
-    FieldElem(375),
-    FieldElem(3046),
-    FieldElem(1847),
-    FieldElem(1438),
-    FieldElem(1143),
-    FieldElem(2786),
-    FieldElem(756),
-    FieldElem(2865),
-    FieldElem(2099),
-    FieldElem(2393),
-    FieldElem(733),
-    FieldElem(2474),
-    FieldElem(2110),
-    FieldElem(2580),
-    FieldElem(583),
-    FieldElem(3253),
-    FieldElem(2037),
-    FieldElem(1339),
-    FieldElem(2789),
-    FieldElem(807),
-    FieldElem(403),
-    FieldElem(193),
-    FieldElem(3281),
-    FieldElem(2513),
-    FieldElem(2773),
-    FieldElem(535),
-    FieldElem(2437),
-    FieldElem(1481),
-    FieldElem(1874),
-    FieldElem(1897),
-    FieldElem(2288),
-    FieldElem(2277),
-    FieldElem(2090),
-    FieldElem(2240),
-    FieldElem(1461),
-    FieldElem(1534),
-    FieldElem(2775),
-    FieldElem(569),
-    FieldElem(3015),
-    FieldElem(1320),
-    FieldElem(2466),
-    FieldElem(1974),
[…]

### 2024-07-14 lattices/primitives/src/lib.rs

@@ -1,201 +1,0 @@
-//! Primitives used in lattice-based cryptography
-#![no_std]
-
-use algebra::{Poly, PolyNTT, PolyNTTMatrix, PolyNTTVec};
-use symmetric::shake128_xof;
-
-use crate::{algebra::FieldElem, symmetric::hash_g};
-
-/// word size of all arithmetics
-pub type Word = u64;
-
-pub const KYBER_N: usize = 256;
-pub const KYBER_Q: Word = 3329;
-pub const KYBER_Q_BITS: usize = 12;
-pub const KYBER_K_512: usize = 2;
-pub const KYBER_K_768: usize = 3;
-pub const KYBER_K_1024: usize = 4;
-
-/// TODO: will only implement ML_KEM_768 for now
-pub const KYBER_K: usize = KYBER_K_768;
-pub const KYBER_ETA_1: usize = 2;
-pub const KYBER_ETA_2: usize = 2;
-
-/// Most seeds are 32-bytes
-pub const SEEDSIZE: usize = 32;
-
-pub mod algebra;
-pub mod symmetric;
-
-/// Algorithm 4: ByteEncode
-/// Encode an integer array into a byte array by encoding each integer using d bits
-///
-/// There is no easy way to define the right size of output array (and with no_std I can't use
-/// Vec), so the user has to provide the buffer to write the bytes to. Fortunately, it is easy to
-/// check that the input buffer has the right size
-pub fn byte_encode(coeffs: &[FieldElem], d: usize, buffer: &mut [u8]) {
-    // The total number of bits must match
-    assert_eq!(coeffs.len() * d, buffer.len() * (u8::BITS as usize));
-
-    for coeff_loc in 0..coeffs.len() {
-        let coeff = coeffs[coeff_loc].0;
-        for coeff_bit_loc in 0..d {
-            let coeff_bit = (coeff >> coeff_bit_loc) & 1;
-
-            let enc_loc = (coeff_loc * d + coeff_bit_loc) / 8;
-            let enc_bit_loc = (coeff_loc * d + coeff_bit_loc) % 8;
-
-            if coeff_bit == 1 {
-                buffer[enc_loc] = buffer[enc_loc] | (0b1000_0000 >> enc_bit_loc);
-            }
-        }
-    }
-}
-
-/// Algorithm 5: ByteDecode
-/// Assume that the input buffer is the output of calling byte_encode with the same value for d,
-/// recover the encoding back to the integer array
[…]

### 2024-07-14 lattices/primitives/src/main.rs

@@ -1,20 +1,0 @@
-use primitives::algebra::PolyNTT;
-use sha3::{
-    digest::{ExtendableOutput, Update},
-    Shake256,
-};
-
-fn main() {
-    let mut xof = Shake256::default();
-    xof.update(b"test");
-    let mut xof = xof.finalize_xof();
-    let poly = PolyNTT::sample_uniform(&mut xof).invert_ntt();
-    println!("{:b}", poly);
-
-    let mut buffer = [0u8; 32 * 12];
-    primitives::byte_encode(poly.as_coeffs(), 12, &mut buffer);
-
-    for byte in buffer {
-        print!("{:08b}, ", byte);
-    }
-}

### 2024-07-14 lattices/primitives/src/symmetric.rs

@@ -1,110 +1,0 @@
-//! Hash, XOF, and PRF
-use crate::SEEDSIZE;
-use sha3::{
-    digest::{ExtendableOutput, Update, XofReader},
-    Digest, Sha3_256, Sha3_512, Shake128, Shake256,
-};
-
-/// Hash function H is Sha3-256(s)
-pub fn hash_h(stream: &[u8]) -> [u8; SEEDSIZE] {
-    let mut output = [0u8; SEEDSIZE];
-    let mut hasher = Sha3_256::new();
-    Digest::update(&mut hasher, stream);
-    let result = hasher.finalize();
-    output.clone_from_slice(&result);
-
-    return output;
-}
-
-/// Hash function J is Shake256(s)
-pub fn hash_j(stream: &[u8]) -> [u8; SEEDSIZE] {
-    let mut output = [0u8; SEEDSIZE];
-
-    let mut hasher = Shake256::default();
-    hasher.update(stream);
-    let mut xof = hasher.finalize_xof();
-    xof.read(&mut output);
-
-    return output;
-}
-
-/// Hash function G uses SHA3-512 but output 2 32-bytes
-pub fn hash_g(stream: &[u8]) -> ([u8; SEEDSIZE], [u8; SEEDSIZE]) {
-    let mut hasher = Sha3_512::new();
-    Digest::update(&mut hasher, stream);
-    let hash = hasher.finalize();
-
-    let mut first = [0u8; SEEDSIZE];
-    let mut second = [0u8; SEEDSIZE];
-
-    first.clone_from_slice(hash.get(0..SEEDSIZE).expect("Unexpected out-of-bound"));
-    second.clone_from_slice(
-        hash.get(SEEDSIZE..(2 * SEEDSIZE))
-            .expect("Unexpected out-of-bound"),
-    );
-
-    return (first, second);
-}
-
-pub fn shake256_prf<const ETA: usize>(seed: [u8; SEEDSIZE], ctr: u8) -> [u8; ETA] {
-    let mut hasher = Shake256::default();
-    hasher.update(&seed);
-    hasher.update(&[ctr]);
-    let mut xof = hasher.finalize_xof();
-
-    let mut output = [0u8; ETA];
-    xof.read(&mut output);
-    return output;
-}
-
[…]

### 2024-07-16 faster-fo-transform/manuscript/main.tex
Updated formatting
Described sequence of games for the proof

@@
 \newcommand{\pke}{\monospace{PKE}}
-\newcommand{\keygen}{\monospace{KeyGen}}
-\newcommand{\encrypt}{\monospace{E}}
-\newcommand{\decrypt}{\monospace{D}}
+\newcommand{\keygen}{\monospace{Gen}}
+\newcommand{\encrypt}{\monospace{Enc}}
+\newcommand{\decrypt}{\monospace{Dec}}
 \newcommand{\kem}{\monospace{KEM}}
@@
 \newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
-\newcommand{\fotplus}{$\monospace{FOT}^+$}
+\newcommand{\fotplus}{\monospace{FOT+}}
 
@@
 
-\section{FO transform with encrypt-then-mac for achieving ciphertext integrity}\label{sec:main-results}
-In this section, we present two different encrypt-then-mac modes using symmetric-key approaches   to achieve ciphertext integrity, which result in two different fast FOTs.
-The first method is to apply an MAC over ciphertext [???], referred to as \emph{standard Enc-then-MAC}, and the second method is adopted from GCM [??] mode in authenticated encryption (AE) using polynomial hash, referred to as \emph{poly Enc-then-MAC}.
-In other words, we will replace the public-key re-encryption  in FOT  by symmetric-key MAC approaches, and the resulting transform is referred to as \emph{FOT+}.
+\section{The ``encrypt-then-MAC'' transformation}\label{sec:main-results}
+Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme.
+Let $\mac$ be a deterministic message authentication code.
+Let $G: \mathcal{M}_\pke \rightarrow \mathcal{K}_\mac$ and $H: \{0,1\}^\ast \rightarrow \mathcal{K}_\kem$ be hash functions, where $\mathcal{K}_\kem$ denote the set of all possible session keys.
+The $\etm$ transformation outputs a key encapsulation mechanism $\kem_\etm(\keygen_\etm, \encap_\etm, \decap_\etm)$.
+The three routines are described in figure \ref{fig:etm-routines}.
 
-\subsection{FOT+ in standard encrypt-then-MAC}
-
-Let $\pke(\keygen, \encrypt, \decrypt)$ be a probabilistic public-key encryption scheme defined over message space $\mathcal{M}_\pke$, ciphertext space $\mathcal{C}$, and coin space $\mathcal{R}$.
-Let $\mac$ be a deterministic and perfectly correct message authentication code defined over key space $\mathcal{K}_\mac$, message space $\mathcal{M}_\mac$, and tag space $\mathcal{T}_\mac$.
-Let $G: \mathcal{M}_\pke \rightarrow \mathcal{R} \times \mathcal{K}_\mac$ be a hash function that hashes a plaintext message into a pseudorandom coin and a $\mac$ key.
-The $\monospace{T}_\etm[\pke, \mac, G]$ transformation takes the input $\pke$, $\mac$, and hash function $G$, and outputs a public-key encryption scheme $\pke_\etm(\keygen, \encrypt_\etm, \decrypt_\etm)$ where as the key generation routine remains unchanged, and the encryption/decryption routines are as follows:
-
-\begin{algorithm}
-    \caption{$\encrypt_\etm$(\pk, m)}\label{alg:encrypt-etm}
-
+\begin{figure}[H]
+    \center
+    \begin{minipage}{0.29\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\keygen_\etm$}\label{alg:etm-keygen}
             \begin{algorithmic}[1]
-        \State $(r, k) \leftarrow G(m)$
-        \State $c \leftarrow \encrypt(\pk, m; r)$
-        \State $t \leftarrow \sign(k, c)$
-        \State \Return $(c, t)$
+                \State $(\pk, \sk_\pke) \leftsample \keygen(1^\lambda)$
+                \State $z \leftsample \mathcal{M}_\pke$
+                \State $\sk \leftarrow (\sk_\pke, z)$
+                \State \Return $(\pk, \sk)$
             \end{algorithmic}
         \end{algorithm}
-
-\begin{algorithm}
-    \caption{$\decrypt_\etm$(\sk, (c, t))}\label{alg:decrypt-etm}
-
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.32\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap_\etm(\pk)$}\label{alg:etm-encap}
             \begin{algorithmic}[1]
-        \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-        \State $(\hat{r}, \hat{k}) \leftarrow G(m)$
-        \If {$\verify(\hat{k}, c, t) = 0$}
-            \State \Return $\bot$
-        \EndIf
-        \State \Return $\hat{m}$
+                \State $m \leftsample \mathcal{M}_\pke$
+                \State $k \leftarrow G(m)$
+                \State $c \leftsample \encrypt(\pk, m)$
+                \State $t \leftarrow \mac(k, c)$
+                \State $K \leftarrow H(m, c)$
+                \State \Return $((c, t), K)$
             \end{algorithmic}
         \end{algorithm}
+    \end{minipage}
+    \hfill
+    \begin{minipage}{0.37\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap_\etm(\sk, (c, t))$}\label{alg:etm-decap}
+            \begin{algorithmic}[1]
+                \State $(\sk_\pke, z) \leftarrow \sk$
+                \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c)$
+                \State $\hat{k} \leftarrow G(\hat{m})$
+                \If{$\mac(\hat{k}, c) \neq t$}
+                    \State \Return $H(z, c)$
+                \EndIf
+                \State \Return $H(\hat{m}, c)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
+\end{figure}
 
-We claim that if the input $\pke$ is OW-CPA secure and $\mac$ is existentially unforgeable, then under the random oracle model, $\pke_\etm$ is OW-PCVA secure with non-tight security reduction.
+If the input $\pke$ is \monospace{OW-PCA} secure, then $\kem_\etm$ is \monospace{IND-CCA2} secure.
 
-\begin{theorem}\label{thm:if-owcpa-then-owpcva}
-    If $\pke$ is $\delta$-correct, then $\pke_\etm$ is $\delta$-correct.
-In addition, for every \monospace{OW-PCVA} adversary $\mathcal{A}_\monospace{OW-PCVA}$ against $\pke_\etm$ that makes $q_P$ PCO queries, $q_V$ CVO queries, $q_G$ hash queries to $G$, and that has advantage $\epsilon_\monospace{OW-PCVA}$ there exists an existential forgery adversary $\mathcal{A}_\mac$ against the underlying $\mac$ with advantage $\epsilon_\mac$ and some $\monospace{OW-CPA}$ adversary $\mathcal{A}_\monospace{OW-CPA}$ against the underlying $\pke$ with advantage $\epsilon_\monospace{OW-CPA}$ such that
+\begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
+    For every \monospace{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes at most $q_D$ decapsulation queries, there exists an \monospace{OW-PCA} adversary $B$ against the underlying $\pke$ such that
 
     \begin{equation*}
-        \epsilon_\monospace{OW-PCVA}
-        \leq (q_G + q_P) \cdot \delta
-            + q_V \cdot \epsilon_\mac
-            + (q_G + q_P + 1) \cdot \epsilon_\monospace{OW-CPA}
+        \monospace{Adv}_\monospace{IND-CCA2}(A) \leq q_D \cdot \epsilon_\mac + 2 \cdot \monospace{Adv}_\monospace{OW-PCA}(B)
     \end{equation*}
@@
 
-Furthermore, if the input $\pke$ is additionally IND-CPA secure, then $\pke_\etm$ is OW-PCVA secure with tight security reduction.
+\begin{proof}
+    We will prove using a sequence of games. The complete sequence of games is shown in figure \ref{fig:sequence-of-games}
 
-\begin{corollary}\label{thm:if-indcpa-then-owpcva}
-    For every \monospace{OW-PCVA} adversary $\mathcal{A}_\monospace{OW-PCVA}$ against $\pke_\etm$ that makes $q_P$ PCO queries, $q_V$ CVO queries, $q_G$ hash queries to $G$, and that has advantage $\epsilon_\monospace{OW-PCVA}$ there exists an existential forgery adversary $\mathcal{A}_\mac$ against the underlying $\mac$ with advantage $\epsilon_\mac$ and some $\monospace{IND-CPA}$ adversary $\mathcal{A}_\monospace{IND-CPA}$ against the underlying $\pke$ with advantage $\epsilon_\monospace{IND-CPA}$ such that
-
-    \begin{equation*}
-        \epsilon_\monospace{OW-PCVA}
-        \leq (q_G + q_P) \cdot \delta
-            + q_V \cdot \epsilon_\mac
-            + \frac{1 + 2q_G}{\norm{\mathcal{M}_\pke}}
-            + 3\epsilon_\monospace{IND-CPA}
-    \end{equation*}
-\end{corollary}
-
-\begin{proof}
-    Since no modification was made to the internals of the input $\pke$, the correctness of the transformed scheme is trivially identical to the correctness of the input scheme.
-
-    We will prove the security claim using a sequence of games \cite{shoup2004sequences}, then prove the security claim in the corollary \ref{thm:if-indcpa-then-owpcva} by making a few modifications.
-This proof borrows heavily from the proof presented in \cite{hofheinz2017modular}.
-
-    \begin{figure}
-        \centering
-
+    \begin{figure}[H]
+        \center
+        \begin{minipage}{0.55\textwidth}
             \begin{algorithm}[H]
-            \caption{Sequence of games}\label{alg:sequence-of-games}
+                \caption{Sequence of games $G_0 - G_3$}\label{alg:sequence-of-games}
                 \begin{algorithmic}[1]
                     \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                \State $m^\ast \leftsample \mathcal{M}_\pke$
-                \State $(r^\ast, k^\ast) \leftarrow G(m^\ast)$
+                    \State $(m^\ast, z) \leftsample \mathcal{M}_\pke$
+                    \State $k^\ast \leftarrow G(m^\ast)$
+                        \Comment{Game 0-1}
+                    \State $k^\ast \leftsample \mathcal{K}_\mac$
+                        \Comment{Game 2-3}
+                    \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
+                    \State $t^\ast \leftarrow \mac(k^\ast, c^\ast)$
+                    \State $K_0 \leftarrow H(m^\ast, c^\ast)$
                         \Comment{Game 0-2}
-                \State $r^\ast \leftsample \mathcal{R}, k^\ast \leftsample \mathcal{K}_\mac$
+                    \State $K_0 \leftsample \mathcal{K}_\kem$
                         \Comment{Game 3}
-                \State $c^\ast \leftarrow \encrypt(\pk, m^\ast; r^\ast)$
-                \State $t^\ast \leftarrow \sign(k^\ast, c^\ast)$
-                \State $
-                    \hat{m} \leftarrow \mathcal{A}^{\mathcal{O}^G, \pco, \cvo}_\monospace{OW-PCVA}(
-                        1^\lambda, \pk, (c^\ast, t^\ast)
-                    )
-                $
+                    \State $K_1 \leftsample \mathcal{K}_\kem$
+                    \State $b \leftsample \{0,1\}$
+                    \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(
+                        1^\lambda, \pk, (c^\ast, t^\ast), K_b)$
                         \Comment{Game 0}
-                \State $
-                    \hat{m} \leftarrow \mathcal{A}^{\mathcal{O}^G, \pco_1, \cvo}_\monospace{OW-PCVA}(
-                        1^\lambda, \pk, (c^\ast, t^\ast)
-                    )
-                $
-                    \Comment{Game 1}
-                \State $
-                    \hat{m} \leftarrow \mathcal{A}^{\mathcal{O}^G, \pco_1, \cvo_1}_\monospace{OW-PCVA}(
-                        1^\lambda, \pk, (c^\ast, t^\ast)
-                    )
-                $
-                    \Comment{Game 2-3}
-                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+                    \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(
+                        1^\lambda, \pk, (c^\ast, t^\ast), K_b)$
+                        \Comment{Game 1-3}
+                    \State \Return $\llbrack \hat{b} = b \rrbrack$
                 \end{algorithmic}
             \end{algorithm}
-
-        \caption{Sequence of \monospace{Game 0} to \monospace{Game 3} in proof of theorem \ref{thm:if-owcpa-then-owpcva}}
-        \label{fig:enter-label}
+        \end{minipage}
+        \hfill
+        \begin{minipage}{0.4\textwidth}
+            \begin{algorithm}[H]
+                \caption{$\mathcal{O}^\decap(c, t)$}\label{alg:etm-decap-oracle}
+                \begin{algorithmic}[1]
+                    \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c)$
+                    \State $\hat{k} \leftarrow G(\hat{m})$
+                    \If{$\mac(\hat{k}, c) = t$}
+                        \State \Return $H(\hat{m}, c)$
+                    \EndIf
+                    \State \Return $H(z, c)$
+                \end{algorithmic}
+            \end{algorithm}
+            \begin{algorithm}[H]
+                \caption{$\mathcal{O}^\decap_1(c, t)$}\label{alg:etm-decap-oracle-1}
+                \begin{algorithmic}[1]
+                    % TODO: this is ugly
+                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G :$ \\
+                        $\;\;\;\;\decrypt(\sk_\pke, c) = \tilde{m}$ \\
+                        $\;\;\;\;\;\land\; \mac(\tilde{k}, c) = t$}
+                        \State \Return $H(\tilde{m}, c)$
+                    \EndIf
+                    \State \Return $H(z, c)$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \caption{Sequence of games}\label{fig:sequence-of-games}
     \end{figure}
 
-    \begin{algorithm}
-        \caption{$\pco(m, (c, t))$}\label{alg:etm-pco}
-        \begin{algorithmic}[1]
-            \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-            \State $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$
-            \If{$\verify(\hat{k}, c, t) = 0$}
-                \State \Return $0$
-            \EndIf
-            \State \Return $\llbrack \hat{m} = m \rrbrack$
-        \end{algorithmic}
-    \end{algorithm}
-
-    \begin{algorithm}
-        \caption{$\pco_1(m, (c, t))$}\label{alg:etm-pco1}
-        \begin{algorithmic}[1]
-            \State $(r, k) \leftarrow G(\hat{m})$
-            \State \Return $\llbrack 
-                \encrypt(\pk, m; r) = c
-            \rrbrack \monospace{ and } \llbrack
-                \verify(k, c, t) = 1
-            \rrbrack$
-        \end{algorithmic}
-    \end{algorithm}
-
-    \begin{algorithm}
-        \caption{$\cvo(c, t)$}\label{alg:etm-cvo}
-        \begin{algorithmic}[1]
-            \State $\hat{m} \leftarrow \decrypt(\sk, c)$
-            \State $(\hat{r}, \hat{k}) \leftarrow G(\hat{m})$
-            \State \Return $\verify(\hat{k}, c, t)$
-        \end{algorithmic}
-    \end{algorithm}
-
-    \begin{algorithm}
-        \caption{$\cvo_1(c, t)$}\label{alg:etm-cvo1}
-
-        \begin{algorithmic}[1]
-            \If{
-                $\exists (\tilde{m}, \tilde{r}, \tilde{k}) \in \mathcal{O}^G$ such that
-                $\encrypt(\pk, \tilde{m}; \tilde{r}) = c$
-                and $\verify(\tilde{k}, c, t)$
-            }
-                \State \Return 1
-            \EndIf
-            \State \Return 0
-        \end{algorithmic}
-    \end{algorithm}
-
-    \monospace{Game 0} is the \monospace{OW-PCVA} game. Let $\epsilon_0$ denote the \monospace{OW-PCVA} adversary's advantage in \monospace{Game 0}, then $\epsilon_0 = \epsilon_\monospace{OW-PCVA}$
-
-    \monospace{Game 1} is identical to \monospace{Game 0}, except $\pco$ is replaced with $\pco_1$.
-Because $\encrypt_\etm$ is a deterministic encryption routine, the two games differ from the adversary's perspective if and only if any of the $\pco$ query $(m, (c, t))$ causes decryption failure $\decrypt(\sk, \encrypt(\pk, m; r)) \neq m$.
-The probability of decryption failure for any single $\pco$ query is bounded by $\delta$, so the overall probability of having at least one query causing decryption failure is at most $q_P \cdot \delta$.
-Let $\epsilon_0$ and $\epsilon_1$ respectively denote $\mathcal{A}_\monospace{OW-PCVA}$'s advantage in \monospace{Game 0} and \monospace{Game 1} respectively, then by the difference lemma \cite{shoup2004sequences}:
-
-    \begin{equation}\label{eq:ep0-ep1}
-        \epsilon_0 - \epsilon_1 \leq q_P \cdot \delta
-    \end{equation}
-
-    \monospace{Game 2} is identical to \monospace{Game 1} except $\cvo$ is replaced with $\cvo_1$.
-$\cvo_1$ replaces the decryption routine with the deterministic encryption routine and the $\mac$ key derivation with ``searching through hash oracle records''.
-Therefore, there are exactly two scenarios in which \monospace{Game 2} differ from \monospace{Game 1} from the adversary's perspective.
-
-    In the first scenario, the queried ciphertext $(c, t)$ has a matching hash query $(\tilde{m}, \tilde{r}, \tilde{k})$, but $(\tilde{m}, \tilde{r})$ causes decryption failure: $\decrypt(\sk, \encrypt(\pk, m; r)) \neq m$.
-For each hash query, the probability that $(\tilde{m}, \tilde{r})$ causes decryption failure is bounded by $\delta$, so the probability of having at least one such hash query is at most $q_G \cdot \delta$.
-
-    In the second scenario, the queried ciphertext $(c, t)$ has no matching hash query.
-Under the random oracle model, this means that the $\mac$ key $k$ used to sign $c$ is an unknown and uniformly random key from the adversary's perspective.
-In other words, $(c, t)$ is an existential forgery.
-The probability of producing a single forgery is bounded by the advantage of a MAC adversary, so the probability of having at least one dishonest $\cvo$ query is at most $q_V \cdot \epsilon_\mac$.
-
-    Denote the $\monospace{OW-PCVA}$ adversary's advantage in game 2 by $\epsilon_2$, then by the difference lemma:
-
-    \begin{equation}\label{eq:ep1-ep2}
-        \epsilon_1 - \epsilon_2 \leq q_G \cdot \delta + q_V \cdot \epsilon_\mac
-    \end{equation}
-
-    In \monospace{Game 3}, the challenge encryption routine is modified.
-Instead of pseudorandomly deriving the coin $r^\ast$ and the $\mac$ key $k^\ast$ from $G$, the coin and the $\mac$ key are uniformly sampled from their respective domain.
-Under the random oracle model, \monospace{Game 3} and \monospace{Game 2} are indistinguishable from the adversary's perspective unless the adversary queries $G$ or $\pco$ with the value $m^\ast$.
-Denote the probability of ``adversary querying $G$ or $\pco$ with $m^\ast$'' by $P[\monospace{QUERY}^\ast]$, and the adversary's advantage in \monospace{Game 3} by $\epsilon_3$, then by the difference lemma:
-
-    \begin{equation}\label{eq:ep2-ep3}
-        \epsilon_2 - \epsilon_3 \leq P[\monospace{QUERY}^\ast]
-    \end{equation}
-
-    A standard \monospace{OW-CPA} adversary against the underlying $\pke$ can simulate \monospace{Game 3} for an \monospace{OW-PCVA} adversary, since $\pco_1$ and $\cvo_1$ only make use of the public key $\pk$ and the hash oracle $\mathcal{O}^G$, the challenge encryption $c^\ast$ is obtained using a truly random coin, and the $\mac$ key can be uniformly sampled.
-After the \monospace{OW-PCVA} adversary outputs a guess, the \monospace{OW-CPA} adversary can simply pass \monospace{OW-PCVA}'s output.
-The \monospace{OW-CPA} adversary wins if and only if the \monospace{OW-CPA} adversary wins \monospace{Game 3}:
-
-    \begin{equation}\label{eq:ep3-owcpa}
-        \epsilon_3 = \epsilon_\monospace{OW-CPA}
-    \end{equation}
-
-    We can construct another \monospace{OW-CPA} adversary that simulates \monospace{Game 3} for an \monospace{OW-PCVA} adversary.
-After the \monospace{OW-PCVA} adversary halts, this \monospace{OW-CPA} adversary picks and outputs a random value $\tilde{m}$ from all possible values recorded on the tape of the hash oracle $\mathcal{O}^G$ and $\pco_1$.
-If $\monospace{QUERY}^\ast$ occurs, then the \monospace{OW-CPA} adversary wins the game with probability $\frac{1}{q_G + q_P}$.
-In other words:
-
-    \begin{equation}\label{eq:querystar-owcpa}
-        \epsilon_\monospace{OW-CPA} = \frac{1}{q_G + q_P} \cdot P[\monospace{QUERY}^\ast]
-    \end{equation}
-
-    Combining equations \ref{eq:ep0-ep1}, \ref{eq:ep1-ep2}, \ref{eq:ep2-ep3}, \ref{eq:ep3-owcpa}, and \ref{eq:querystar-owcpa} gives the security bound in theorem \ref{thm:if-owcpa-then-owpcva}.
-
-    We make two modifications to prove corollary \ref{thm:if-indcpa-then-owpcva}.
-First, we invoke a well-known result that the IND-CPA security of a $\pke$ with sufficiently large message space implies its OW-CPA security.
-
-    \begin{lemma}
-        For every \monospace{OW-CPA} adversary with advantage $\epsilon_\monospace{OW-CPA}$ there exists an \monospace{IND-CPA} adversary with advantage $\epsilon_\monospace{IND-CPA}$ such that
-
-        \begin{equation}\label{eq:indcpa-implies-owcpa}
-            \epsilon_\monospace{OW-CPA} \leq \frac{1}{\norm{\mathcal{M}_\pke}} + \epsilon_\monospace{IND-CPA}
-        \end{equation}
-    \end{lemma}
-
-    Second, we borrow results from \cite{hofheinz2017modular} and construct an \monospace{IND-CPA} adversary to bound $P[\monospace{QUERY}^\ast]$:
-
-    \begin{equation}\label{eq:querystar-indcpa}
-        \frac{1}{2}P[\monospace{QUERY}^\ast] 
-        \leq \epsilon_\monospace{IND-CPA} + \frac{q_G}{\norm{\mathcal{M}_\pke}}
-    \end{equation}
-
-    Combining equations \ref{eq:ep0-ep1}, \ref{eq:ep1-ep2}, \ref{eq:ep2-ep3}, \ref{eq:ep3-owcpa}, \ref{eq:indcpa-implies-owcpa}, and \ref{eq:querystar-indcpa} into theorem \ref{thm:if-owcpa-then-owpcva} completes the proof of the corollary.
+    % TODO: finish the proof
 \end{proof}
-
-\subsection{FO Transform+ in poly Enc-then-MAC}
 
@@
 
-\subsection{Kyber using FO Transform+}
+\subsection{Kyber using \fotplus}
 \subsection{Performance comparisons}\label{sec:exp-results}

### 2024-07-17 faster-fo-transform/manuscript/main.tex
Checkpoint

@@
 
-If the input $\pke$ is \monospace{OW-PCA} secure, then $\kem_\etm$ is \monospace{IND-CCA2} secure.
+The security of $\kem_\etm$ depends on the one-way security of the input $\pke$ and the unforgeability of the input $\mac$.
+Specifically, we require the input $\pke$ to be \monospace{OW-PCA} secure.
+If the input $\pke$ is \monospace{OW-PCA} and the input $\mac$ is existentially one-time unforgeable, then $\kem_\etm$ is \monospace{IND-CCA2} secure.
+On the other hand, if there exists an efficient \monospace{OW-PCA} adversary against the input $\pke$, then there exists an efficient \monospace{IND-CCA2} adversary against $\kem_\etm$.
+
+\subsection{If PKE is OW-PCA secure, then EtM is IND-CCA2 secure}
 
@@
 
-\section{Applications to Kyber}
+\subsection{If PKE is not OW-PCA, then EtM is not IND-CCA2 secure}\label{sec:not-owpca-implies-not-indcca2}
+More specifically, if there exists an efficient \monospace{OW-PCA} adversary $B$ against the underlying $\pke$, then we can build an efficient \monospace{IND-CCA2} adversary $A$ who uses $B$ as a sub-routine and wins the \monospace{IND-CCA2} game.
 
-\subsection{Kyber using \fotplus}
-\subsection{Performance comparisons}\label{sec:exp-results}
-% TODO: Count how many Keccak permutations
-% TODO: Count how many NTT operations
-% 
-Comparisons will do for Kyber and Kyber$^+$, define it.
+To run $B$ as a sub-routine, $A$ needs to be able to simulate a plaintext-checking oracle, which can be done using the decapsulation oracle.
+When presented with a plaintext-chekcing query $(\tilde{m}, \tilde{c})$, $A$ derives the corresponding $\mac$ key $\tilde{k} \leftarrow G(\tilde{m})$ and signs the ciphertext $\tilde{t} \leftarrow \mac(\tilde{k}, \tilde{c})$.
+$A$ then queries the decapsulation oracle on the ciphertext-tag pair $\tilde{K} \leftarrow \mathcal{O}^\decap(\tilde{c}, \tilde{t})$.
 
-\subsection{Experimental results}
+Because the correct session key is deterministically derived from the public key, the plaintext, the ciphertext, and some other values derived from these three values, for each set of $(\tilde{m}, \tilde{c}, \tilde{t}, \tilde{K})$, $A$ can correctly derive the session key if $\tilde{m}$ is indeed the decryption of $\tilde{c}$.
+On the other hand, if $\tilde{m} \neq \decrypt_\pke(\sk_\pke, \tilde{c})$, then the decapsulation oracle will return the implicit rejection value, which will not match the expected session key value.
+
+By comparing the decapsulation orcale's output and the expected session key, $A$ can correctly distinguish whether $\tilde{m}$ is the decryption of $\tilde{c}$ or not.
+Thus $A$ can correctly simulate the plaintext-checking oracle for $B$.
+
+When $A$ receives the challenge ciphertext $(c^\ast, t^\ast)$ and unknown session key $K_b \leftsample \{K_0, K_1\}$, $A$ passes $c^\ast$ as the challenge ciphertext to $B$ as $B$'s challenge ciphertext.
+After $B$ returns the guess $\hat{m}$, $A$ computes the expected session key using $\hat{m}$ and $c^\ast$.
+If $\hat{m}$ is the correct decryption of $c^\ast$, then the expected session key should match the correct session key, in which case $A$ will correctly distinguish the true session key from the random session key.
+If $\hat{m}$ is not the correct decryption of $c^\ast$, then the expected session key will probably not match anything, so $A$ will always claim $K_b$ to be a random key.
+In other words, \emph{if $B$ wins the \monospace{OW-PCA} game, then $A$ wins the \monospace{IND-CCA2} game; if $B$ does not win, then $A$'s chance of winning is exactly $\frac{1}{2}$}.
+Therefore, the advantage of $A$ is at least that of $B$.
+
+\section{Application to Kyber}
+\subsection{``encrypt-then-MAC'' is not secure with Kyber}
+From section \ref{sec:not-owpca-implies-not-indcca2} we know that if there exists an efficient \monospace{OW-PCA} adversary $B$ against the underlying $\pke$, then there exists an efficient \monospace{IND-CCA2} adversary $A$ against the transformed $\kem$.
+In this section we'll review Kyber's construction and discuss how to construct efficient plaintext checking attack, which demonstrates that $\etm$ cannot be securely combined with Kyber.
+
+Kyber's construction as shown in \cite{avanzi2019crystals} contains two components: an \monospace{IND-CPA} secure $\pke$ (denoted by \monospace{CPAPKE}), and an \monospace{IND-CCA2} secure $\kem$ (denoted by \monospace{CCAKEM}).
+\monospace{CPAPKE} contains three routines: key generation, encryption, and decryption.
+Their details are described in figure \ref{fig:kyber-cpapke-routines}.
+
+\begin{figure}[H]
+    \begin{algorithm}[H]
+        \caption{\monospace{CPAPKE.Gen}}\label{alg:kyber-cpapke-keygen}
+        \begin{algorithmic}[1]
+            \State $d \leftsample \mathcal{B}^{32}$
+            \State $(\rho, \sigma) \leftarrow \monospace{SHA3-512}(d)$
+            \State $N = 0$
+            \For{$i \in \{0, 1, \ldots, k-1\}$}
+                \Comment{Generate public matrix $A$}
+                \For{$j \in \{0, 1, \ldots, k-1\}$}
+                    \State $\hat{A}\monospace{[i][j]} = \monospace{Parse}(
+                        \monospace{Shake128}(\rho, j, i)
+                    )$
+                        \Comment{$\hat{A}$ is in NTT-domain}
+                \EndFor
+            \EndFor
+            \For{$i \in \{0, 1, \ldots, k-1\}$}
+                \Comment{Sample secret $\mathbf{s}$}
+                \State $\mathbf{s}\monospace{[i]} = \monospace{CBD}_{\eta_1}(
+                    \monospace{Shake256}((\rho, N))
+                )$
+                \State $N \leftarrow N + 1$
+            \EndFor
+            \For{$i \in \{0, 1, \ldots, k-1\}$}
+                \Comment{Sample secret $\mathbf{e}$}
+                \State $\mathbf{e}\monospace{[i]} = \monospace{CBD}_{\eta_1}(
+                    \monospace{Shake256}((\rho, N))
+                )$
+                \State $N \leftarrow N + 1$
+            \EndFor
+            \State $\mathbf{\hat{s}} \leftarrow \monospace{NTT}(\mathbf{s})$
+            \State $\mathbf{\hat{e}} \leftarrow \monospace{NTT}(\mathbf{e})$
+            \State $\mathbf{\hat{t}} \leftarrow \hat{A} \cdot \mathbf{\hat{s}} + \mathbf{\hat{e}}$
+            \State $\pk = (\mathbf{\hat{t}} \Vert \rho)$
+            \State $\sk = \mathbf{\hat{s}}$
+            \State \Return $(\pk, \sk)$
+        \end{algorithmic}
+    \end{algorithm}
+    \begin{algorithm}[H]
+        \caption{$\monospace{CPAPKE.Enc}(\pk, m, r)$}\label{alg:kyber-cpapke-encrypt}
+        \begin{algorithmic}[1]
+            \State $N \leftarrow 0$
+            \State $(\rho, \mathbf{\hat{t}}) \leftarrow \pk$
+                \Comment{Unpack and decode the public key}
+            \For{}
+                \For{}
+                \EndFor{}
+            \EndFor{}
+        \end{algorithmic}
+    \end{algorithm}
+    \caption{Kyber's \monospace{CPAPKE} routines}\label{fig:kyber-cpapke-routines}
+\end{figure}
+
+\subsection{Experimental results}\label{sec:exp-results}
+We claim that the input $\mac$ only needs to be one-time existentially unforgeable.
+This is because besides the challenge ciphertext $(c^\ast, t^\ast)$, the adversary has no external resources from which it can obtain authenticated ciphertexts for which it does not know the decryption.
+Here we compare the performance of a variety of $\mac$ instantiations.
+Some of them are many-time secure while others are one-time secure.
+The standalone performance results are listed in table \ref{tbl:standalone-mac-perf}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        Name & Security & 768 bytes & 1088 bytes & 1568 bytes \\
+        \hline
+        CMAC & many-time & 5022 & 5442 & 6090 \\
+        & & 5131 & 5578 & 6154 \\
+        \hline
+        GMAC & one-time & 2778 & 2756 & 2762 \\
+        & & 2843 & 2780 & 2919 \\
+        \hline
+        KMAC-256 & many-time & 7934 & 9862 & 11742 \\
+        & & 8594 & 10693 & 12319 \\
+        \hline
+        Poly1305 & one-time & 7934 & 9862 & 11742 \\
+        & & 8594 & 10693 & 12319 \\
+        \hline
+    \end{tabular}
+    \caption{Standalone $\mac$ performances\\
+    Top number is median CPU cycle, bottom number is mean CPU cycle
+    }\label{tbl:standalone-mac-perf}
+\end{table}
 

### 2024-07-18 faster-fo-transform/manuscript/main.tex
Use case description

@@
 
-In section \ref{sec:exp-results}, we instantiate concrete instances of our proposed transformation by modifying ML-KEM.
+In section \ref{sec:mac-performance}, we instantiate concrete instances of our proposed transformation by modifying ML-KEM.
 We will demonstrate that, at the cost of small increase in encryption runtime and ciphertext size, our construction reduces both the runtime and memory footprint of the decryption routine.
@@
 
-In the \monospace{IND-CPA} game, an adversary's goal is to distinguish the encryption of one message from the encryption of another message.
+In the \monospace{IND-CPA} game (algorithm \ref{alg:ind-cpa-game}), an adversary's goal is to distinguish the encryption of one message from the encryption of another message.
 Given the public key, the adversary outputs two adversarially chosen messages and obtains the encryption of a random choice between these two messages.
@@
 
-\begin{algorithm}
-    \caption{The $\monospace{IND-CPA}$ game}\label{alg:pke-ind-cpa-game}
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{The $\monospace{IND-CPA}$ game}\label{alg:ind-cpa-game}
 
@@
         \end{algorithm}
+        }{\caption{The \monospace{IND-CPA} game}}
+    \end{floatrow}
+\end{figure}
 
-The \textit{advantage} $\epsilon_\monospace{IND-CPA}$ of an \monospace{IND-CPA} adversary $\mathcal{A}$ is defined by
+The \textit{advantage} $\epsilon_\monospace{IND-CPA}$ of an \monospace{IND-CPA} adversary $A$ is defined by
 
-\begin{equation*}
-    \epsilon_\monospace{IND-CPA} = \norm{
+$$
+    \monospace{Adv}_\monospace{IND-CPA}(A) = \norm{
         P[\hat{b} = b] - \frac{1}{2}
     }
-\end{equation*}
+$$
 
@@
 
-\begin{algorithm}
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
                 \caption{
@@
             \end{algorithm}
+        }{\caption{The \monospace{KEM-IND-CCA2} game}}
+    \end{floatrow}
+\end{figure}
 
@@
 
-\begin{algorithm}
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox[250px]{
+            \begin{algorithm}[H]
                 \caption{The EUF-CMA game}\label{alg:mac-euf-cma-game}
@@
                         \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
-            \monospace{ and } (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
+                        \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                         \rrbrack
@@
             \end{algorithm}
+        }{\caption{The \monospace{EUF-CMA} game}}
+    \end{floatrow}
+\end{figure}
 
@@
 The modularity of the \monospace{T} and \monospace{U} transformation allows us to tweak only the \monospace{T} transformation (see section \ref{sec:main-results}), obtain \monospace{OW-PCVA} security, then automatically get \monospace{IND-CCA} security for free.
-This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:exp-results}).
+This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:mac-performance}).
 
@@
                 \State $m \leftsample \mathcal{M}_\pke$
-                \State $k \leftarrow G(m)$
-                \State $c \leftsample \encrypt(\pk, m)$
-                \State $t \leftarrow \mac(k, c)$
-                \State $K \leftarrow H(m, c)$
-                \State \Return $((c, t), K)$
+                \State $k_\mac \leftarrow G(\pk, m)$
+                \State $c_\pke \leftsample \encrypt(\pk, m)$
+                \State $t \leftarrow \mac(k, c_\pke)$
+                \State $K \leftarrow H(\pk, m, c_\pke)$
+                \State $c \leftarrow (c_\pke, K)$
+                \State \Return $(c, K)$
             \end{algorithmic}
@@
         \begin{algorithm}[H]
-            \caption{$\decap_\etm(\sk, (c, t))$}\label{alg:etm-decap}
+            \caption{$\decap_\etm(\sk, c)$}\label{alg:etm-decap}
             \begin{algorithmic}[1]
+                \State $(c_\pke, t) \leftarrow c$
                 \State $(\sk_\pke, z) \leftarrow \sk$
-                \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c)$
-                \State $\hat{k} \leftarrow G(\hat{m})$
-                \If{$\mac(\hat{k}, c) \neq t$}
-                    \State \Return $H(z, c)$
+                \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
+                \State $\hat{k}_\pk \leftarrow G(\pk, \hat{m})$
+                \If{$\mac(\hat{k}, c_\pke) \neq t$}
+                    \State \Return $H(\pk, z, c_\pke)$
                 \EndIf
-                \State \Return $H(\hat{m}, c)$
+                \State \Return $H(\pk, \hat{m}, c_\pke)$
             \end{algorithmic}
@@
 
-The security of $\kem_\etm$ depends on the one-way security of the input $\pke$ and the unforgeability of the input $\mac$.
-Specifically, we require the input $\pke$ to be \monospace{OW-PCA} secure.
-If the input $\pke$ is \monospace{OW-PCA} and the input $\mac$ is existentially one-time unforgeable, then $\kem_\etm$ is \monospace{IND-CCA2} secure.
-On the other hand, if there exists an efficient \monospace{OW-PCA} adversary against the input $\pke$, then there exists an efficient \monospace{IND-CCA2} adversary against $\kem_\etm$.
+Here are a few design rationale:
+\begin{enumerate}
+    \item \emph{Deriving MAC key and session key}.
+We choose to include the public key $\pk$ when deriving the MAC key $k_\mac$ and the session key $K$ for similar reason stated in \cite{bos2018crystals}.
+If the MAC key is derived solely from the message, then an adversary can pre-compute a large lookup table mapping MAC key to the source plaintext that can applied to \emph{all sessions}.
+When the adversary intercepts a ciphertext $c = (c_\pke, t)$, it can brute-force all possible MAC keys to recover the plaintext.
+Since brute-forcing MAC key on a known ciphertext-tag pair is an offline search, given a sufficiently large lookup table and a sufficiently large quantum computer, this search might be feasible.
+On the other hand, if the MAC key is derived from both the public key and the plaintext, then the adversary will need to compute the large lookup table and run the key search \emph{per session}, which greatly increases the cost of the attack.
+    \item \emph{Not hashing unauthenticated ciphertext $c_\pke$}: \cite{avanzi2019crystals}\cite{bos2018crystals} derives the session key from a hash of the ciphertext $K \leftarrow \monospace{KDF}(\overline{K} \Vert H(c_\pke))$ to \emph{``simplify implementation with non-incremental hash APIs''}.
+Since ``hashing the ciphertext separately'' vs ``deriving session key directly from ciphertext''  does not affect the security nor the performance (since the ciphertext will be hashed somewhere) of the scheme, we opt to not complicate the design with unnecessary hashes and leave the implementation considerations to concrete instantiations.
+    \item \emph{Not adding key confirmation $\sigma \leftarrow \mac(\cdot, K)$}: within the security model of a KEM, there is no interaction between the KEM adversary and any other party (be it the challenger or any oracle) that involves the adversary sending a session key.
+In other words, there is no scenario in which tempering with any session key is meaningful, which means that the session key does not need protection.
+\end{enumerate}
 
@@
     % TODO: finish the proof
+    \textbf{Need to finish the rest of the proof}
 \end{proof}
@@
 \subsection{``encrypt-then-MAC'' is not secure with Kyber}
-From section \ref{sec:not-owpca-implies-not-indcca2} we know that if there exists an efficient \monospace{OW-PCA} adversary $B$ against the underlying $\pke$, then there exists an efficient \monospace{IND-CCA2} adversary $A$ against the transformed $\kem$.
-In this section we'll review Kyber's construction and discuss how to construct efficient plaintext checking attack, which demonstrates that $\etm$ cannot be securely combined with Kyber.
+CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \monospace{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem.
+For the construction of the \monospace{IND-CCA2} secure KEM, Kyber first constructs an \monospace{IND-CPA} public-key encryption scheme (which we will call \monospace{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \monospace{CCAKEM}).
+Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation.
+The routines of \monospace{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.
 
-Kyber's construction as shown in \cite{avanzi2019crystals} contains two components: an \monospace{IND-CPA} secure $\pke$ (denoted by \monospace{CPAPKE}), and an \monospace{IND-CCA2} secure $\kem$ (denoted by \monospace{CCAKEM}).
-\monospace{CPAPKE} contains three routines: key generation, encryption, and decryption.
-Their details are described in figure \ref{fig:kyber-cpapke-routines}.
+\emph{\monospace{CPAPKE} is not \monospace{OW-PCA} secure}: What's important to know about \monospace{CPAPKE} is that it is not secure against plaintext-checking attack.
 
-\begin{figure}[H]
-    \begin{algorithm}[H]
-        \caption{\monospace{CPAPKE.Gen}}\label{alg:kyber-cpapke-keygen}
-        \begin{algorithmic}[1]
-            \State $d \leftsample \mathcal{B}^{32}$
-            \State $(\rho, \sigma) \leftarrow \monospace{SHA3-512}(d)$
-            \State $N = 0$
-            \For{$i \in \{0, 1, \ldots, k-1\}$}
-                \Comment{Generate public matrix $A$}
-                \For{$j \in \{0, 1, \ldots, k-1\}$}
-                    \State $\hat{A}\monospace{[i][j]} = \monospace{Parse}(
-                        \monospace{Shake128}(\rho, j, i)
-                    )$
-                        \Comment{$\hat{A}$ is in NTT-domain}
-                \EndFor
-            \EndFor
-            \For{$i \in \{0, 1, \ldots, k-1\}$}
-                \Comment{Sample secret $\mathbf{s}$}
-                \State $\mathbf{s}\monospace{[i]} = \monospace{CBD}_{\eta_1}(
-                    \monospace{Shake256}((\rho, N))
-                )$
-                \State $N \leftarrow N + 1$
-            \EndFor
-            \For{$i \in \{0, 1, \ldots, k-1\}$}
-                \Comment{Sample secret $\mathbf{e}$}
-                \State $\mathbf{e}\monospace{[i]} = \monospace{CBD}_{\eta_1}(
-                    \monospace{Shake256}((\rho, N))
-                )$
-                \State $N \leftarrow N + 1$
-            \EndFor
-            \State $\mathbf{\hat{s}} \leftarrow \monospace{NTT}(\mathbf{s})$
-            \State $\mathbf{\hat{e}} \leftarrow \monospace{NTT}(\mathbf{e})$
-            \State $\mathbf{\hat{t}} \leftarrow \hat{A} \cdot \mathbf{\hat{s}} + \mathbf{\hat{e}}$
-            \State $\pk = (\mathbf{\hat{t}} \Vert \rho)$
-            \State $\sk = \mathbf{\hat{s}}$
-            \State \Return $(\pk, \sk)$
-        \end{algorithmic}
-    \end{algorithm}
-    \begin{algorithm}[H]
-        \caption{$\monospace{CPAPKE.Enc}(\pk, m, r)$}\label{alg:kyber-cpapke-encrypt}
-        \begin{algorithmic}[1]
-            \State $N \leftarrow 0$
-            \State $(\rho, \mathbf{\hat{t}}) \leftarrow \pk$
-                \Comment{Unpack and decode the public key}
-            \For{}
-                \For{}
-                \EndFor{}
-            \EndFor{}
-        \end{algorithmic}
-    \end{algorithm}
-    \caption{Kyber's \monospace{CPAPKE} routines}\label{fig:kyber-cpapke-routines}
-\end{figure}
+The main reason why a plaintext-checking attack works against \monospace{CPAPKE} is that for each known honest plaintext-ciphertext pair $(m, c)$, one can perturb the ciphertext and obtain a new ciphertext $c^\prime \leftarrow \monospace{addNoise}(c)$.
+Where the perturbation is small, the perturbed ciphertext can still correctly decrypt to the original plaintext $\decrypt_\monospace{CPAPKE}(\sk, c^\prime) = \decrypt_\monospace{CPAPKE}(\sk, c) = m$.
+Where the perturbation is large, the perturbed ciphertext will not decrypt to the original plaintext.
+The boundary between "small perturbation" and "large perturbation" depends on the values of the secret $\mathbf{s}$, so an adversary can probe the boundary in each of the $k \times n$ dimensions and learn the value of $\mathbf{s}$ coefficient-by-coefficient.
 
-\subsection{Experimental results}\label{sec:exp-results}
+As described in section \ref{sec:not-owpca-implies-not-indcca2}, when the integrity of the unauthenticated ciphertext is protected under a MAC key derived from the corresponding plaintext, \textbf{the MAC will not prevent an adversary from perturbing an honest ciphertext} since the adversary can use the same MAC key to compute tag $t^\prime$ on the perturbed unauthenticated ciphertext $c^\prime$.
+
+On the other hand, \cite{hofheinz2017modular} prevents tempering by the combination of \emph{de-randomization} and \emph{re-encryption}.
+Even if $c^\prime \leftarrow \monospace{addNoise}(c)$ still decrypts back to the same plaintext, \emph{de-randomization} and \emph{re-encryption} will reveal that $c^\prime$ has been tempered with because $c^\prime \neq \encrypt(\pk, m; r\leftarrow G(m))$.
+Unlike $\kem_\etm$ where an adversary can temper with the unauthenticated ciphertext and still produce some valid authenticated ciphertexts, with \emph{re-encryption} \emph{there is no ``other valid ciphertext'' that correspond to the same plaintext}, thus preventing all tempering.
+
+\subsection{Is ``encrypt-then-MAC'' useful?}
+From \cite{hofheinz2017modular} we know that if a $\pke$ is rigid and \monospace{OW-CPA} secure, then the $U^{\not\bot}_m$ transformation is sufficient for constructing an \monospace{IND-CCA2} secure $\kem$ with tight security reduction.
+Furtherfore, the performance overhead of $U^{\not\bot}_m$ over the rigid $\pke$ is minimal.
+Note that rigidity and \monospace{OW-CPA} automatically implies \monospace{OW-PCA} security, so $\kem_\etm$ will also construct an \monospace{IND-CCA2} secure $\kem$, but with extra runtime and communication overhead, so there is practical point of using $\etm$ over the $U^{\not\bot}_m$ transformation.
+On the other hand, if the $\pke$ is not \monospace{OW-PCA}, then $\kem_\etm$ is not \monospace{IND-CCA2} secure.
+
+If the $\pke$ is not rigid but still \monospace{OW-PCA} secure, \emph{de-randomization + re-encryption} is still an option, so we need to consider the performance trade-offs.
+\emph{de-randomization + re-encryption} adds one addition hash (for deriving coin) into the encryption routine and add one hash call and one call to the input encryption routine in the decryption routine.
+\emph{``encrypt-then-MAC''} introduces one hash call and one MAC call to the encryption routine, adds a tag to the ciphertext size, and adds one hash call and one MAC call to the decryption routine.
+Using a one-time MAC like GMAC or Poly1305, the computational cost of producing a tag and the communication cost of the tag should be minimal (1000-2000 CPU cycles + 128 bits of tag).
+Where appropriate, replacing \emph{re-encryption} in the decryption routine with a MAC computation can lead to substantial saving in computational cost.
+This is especially true in protocols where the client (often constrained environments) is responsible for running decryption routine, such as hybrid key exchanges used in CECPQ2.
+
+\begin{table}[H]
+    \begin{tabular}{|c|c|}
+        \hline
+        Input $\pke$ & what should I use to build $\kem$ \\
+        \hline
+        rigid & $U^{\not\bot}_m$ \\
+        \hline
+        \monospace{OW-PCA} but not rigid & $\etm$ \\
+        \hline
+        \monospace{OW-CPA} but not \monospace{OW-PCA} & \emph{de-randomization + re-encryption} \\
+        \hline
+    \end{tabular}
+    \caption{Use cases}
+\end{table}
+
+\subsection{MAC Performance}\label{sec:mac-performance}
 We claim that the input $\mac$ only needs to be one-time existentially unforgeable.
@@
 Some of them are many-time secure while others are one-time secure.
-The standalone performance results are listed in table \ref{tbl:standalone-mac-perf}
+The standalone performance results are listed in table \ref{tbl:standalone-mac-perf}.
+For each choice of MAC, we checked the median (top) and average (bottom) CPU cycles (run on a 2019 MacBook Pro 16-inch) needed to sign 768, 1088, and 1568 bytes of data (respectively the ciphertext size for Kyber512, Kyber768, and Kyber1024).
 
@@
         \hline
-        Poly1305 & one-time & 7934 & 9862 & 11742 \\
-        & & 8594 & 10693 & 12319 \\
+        Poly1305 & one-time & 1128 & 1218 & 1338 \\
+        & & 1435 & 1504 & 1625 \\
         \hline
     \end{tabular}
-    \caption{Standalone $\mac$ performances\\
-    Top number is median CPU cycle, bottom number is mean CPU cycle
-    }\label{tbl:standalone-mac-perf}
+    \caption{Standalone $\mac$ performances}\label{tbl:standalone-mac-perf}
 \end{table}

### 2024-07-18 clangover/entrypoint.sh
Containerized clangover

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+./clangover

### 2024-07-21 etm-transform/insecure/main.tex
Security analysis

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}

% Custom commands
\usepackage{crypto-primitives}
\usepackage{kyber-algos}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    Document title
}
\author{
    Author 1
}
% Leave the date field empty to display the date of compilation
% \date{}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\begin{algorithm}[H]
    \caption{\texttt{Kyber.CCAKEM.KeyGen()}}\label{alg:ccakem-keygen}
    \begin{algorithmic}[1]
        \State $z \leftsample \mathcal{B}^{32}$
            \Comment{Randomly sample 32 bytes (256 bits)}
        \State $(\pk, \sk^\prime) \leftsample \texttt{Kyber.CPAPKE.KeyGen()}$
        \State $\sk = (\sk^\prime, \pk, H(\pk), z)$
            \Comment{H is instantiated with SHA3-256}
        \State \Return $(\pk, \sk)$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}\label{alg:ccakem-encap+}
    \begin{algorithmic}[1]
        \State $m \leftsample \mathcal{B}^{32}$
        \State $m = H(m)$
            \Comment{Do not output system RNG directly}
        \State $(\bar{K}, K_\mac, r) = G(m \Vert H(\pk))$
            \Comment{G is instantiated with SHA3-512}
        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc(\pk, m, r)}$
            \Comment{Because $r$ is set, \texttt{CPAPKE} is deterministic}
            \State $t_1 = \mac(K_\mac, c^\prime)$
        \State $K = \texttt{KDF}(\bar{K} \Vert t_1)$
        \State $t_2=\mac(K, t_1)$
            \Comment{\texttt{KDF} is instantiated with Shake256}
        \State $c \leftarrow (c^\prime, t_1, t_2)$
        \State \Return $(c, K)$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c)$}\label{alg:ccakem-decap+}
    \begin{algorithmic}[1]
        \Require Secret key $\sk = (\sk^\prime, \pk, H(\pk), z)$
        \Require Ciphertext $c = (c^\prime, t_1, t_2)$
        \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
            \Comment{Unpack the secret key; $h$ is the hash of $\pk$}
        \State $(c^\prime, t_1, t_2) \leftarrow c$
        \State $m^\prime = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c^\prime)$ 
        \State $(\overline{K},  K_\mac, r) =   G(m^\prime \Vert h)$
        \State $t_1^\prime=\mac(K_\mac, c^\prime)$
        \If{$t_1' = t_1$}
            \State $K  = \texttt{KDF}(\bar{K} \Vert t_1)$ 
            \State $t_2'= \mac(K, t_1)$
        \EndIf
        \If{$t_2'=t_2$}
            \State \Return $K$
        \Else
            \State Abort
        \EndIf
    \end{algorithmic}
\end{algorithm}

Unfortunately, the \texttt{Kyber+} construction is not \texttt{IND-CCA2} secure when combined with Kyber's \texttt{CPAPKE}. This is because there exists an efficient plaintext-checking attack against \texttt{CPAPKE} that can recover the secret key, and an \texttt{IND-CCA2} adversary against \texttt{KYBER+} can use the decapsulation oracle to simulate the plaintext-checking oracle, which means that an \texttt{IND-CCA2} adversary can run the plaintext-checking attack as a sub-routine and efficiently recover the secret key.

Let $B$ denote the plaintext-checking attack routine. It is given the public key $\pk$ (since the public key is identical between \texttt{CPAPKE} and \texttt{CCAKEM} there is no need to disambiguate) and access to the plaintext-checking oracle $\pco$ (see figure \ref{fig:pco}). After a number of plaintext checking queries, $B$ will output the $\sk^\prime$ that corresponds with the $\pk$ that it receives.

\begin{figure}[H]
    \centering
    \begin{minipage}{0.5\textwidth}
        \begin{algorithm}[H]
            \caption{$\pco(m, c^\prime)$}\label{alg:pco}
            \begin{algorithmic}
                \State \Return $\llbrack m = \texttt{Kyber.CPAPKE.Dec}(
                    \sk^\prime, c^\prime)\rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Plaintext checking oracle}\label{fig:pco}
\end{figure}

We now construct a chosen-ciphertext attack against \texttt{Kyber+}, which we will denote by $A$. $A$ will use $B$ as a sub-routine, which means that $A$ needs to \begin{enumerate}
    \item give $B$ a public key
    \item service $B$'s plaintext checking query
\end{enumerate}

Because \texttt{Kyber+}'s public key is identical to \texttt{Kyber.CPAPKE}'s public key, requirement 1 is trivial: $A$ just gives its own public key to $B$. To service $B$'s plaintext checking query $(\tilde{m}, \tilde{c}^\prime)$, $A$ will use the decapsulation oracle $\decap(c)$ (figure \ref{fig:decap-oracle}).

\begin{figure}[H]
    \centering
    \begin{minipage}{0.5\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap(c)$}
            \begin{algorithmic}[1]
                \State \Return $\texttt{Kyber.CCAKEM.Decap}^+(\sk, c)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The decapsulation oracle}\label{fig:decap-oracle}
\end{figure}

$A$ can run through the steps of $\texttt{Kyber.CCAKEM.Encap}^+$ using ($\tilde{m}$, $\tilde{c}^\prime)$ and arrive at some (authenticated) ciphertext $c = (\tilde{c}^\prime, t_1, t_2)$ and session key $K$. $A$ then sends the $c$ to the decapsulation oracle, which might return some session key is $c$ is valid, or abort if $c$ is not valid. If $\tilde{m}$ is the decryption of $\tilde{c}^\prime$, then $A$'s output of $t_1, t_2$ will be correct, so the decapsulation oracle will accept. If $\tilde{m}$ is not the decryption of $\tilde{c}^\prime$, then $(\hat{K}, K_\mac, r) \leftarrow G(\tilde{m} \Vert H(\pk))$ does not produce the correct $\mac$ key, which means that $t_1, t_2$ will be incorrect, and the decapsulation oracle will abort. See figure \ref{fig:simulated-pco}

\begin{figure}[H]
    \centering
    \begin{algorithm}[H]
        \caption{$\pco_1(m, c^\prime)$}
        \begin{algorithmic}[1]
            \State $(\hat{K}, K_\mac, r) \leftarrow G(m \Vert H(\pk))$
                \Comment{Line 3 of algorithm \ref{alg:ccakem-encap+}}
            \State $t_1 \leftarrow \mac(K_\mac, c^\prime)$
                \Comment{Line 5}
            \State $K \leftarrow \texttt{KDF}(\hat{K} \Vert t_1)$
                \Comment{Line 6}
            \State $t_2 \leftarrow \mac(K, t_1)$
                \Comment{Line 7}
            \State $c \leftarrow (c^\prime, t_1, t_2)$
                \Comment{Line 8}
            \State \Return $\llbrack \decap(c) = K \rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \caption{Simulated plaintext-checking oracle}\label{fig:simulated-pco}
\end{figure}

Here are the complete steps of the chosen-ciphertext attack \begin{enumerate}
    \item Challenger runs \texttt{CCAKEM.KeyGen} and gets $(\pk, \sk)$ where $\sk = (\sk^\prime, \pk, H(\pk), z)$
    \item Challenger gives $\pk$ to $A$, $A$ then gives $\pk$ to $B$
    \item For each of $B$'s plaintext checking query, $A$ runs $\pco_1$ and returns the result back to $B$
    \item After $B$ halts, it returns the recovered secret key $\sk^\prime$
    \item Challenger generates the challenge ciphertext $(c^\ast, K_0) \leftsample \texttt{Kyber.CCAKEM.Encap}^+(\pk)$, samples a random session key $K_1 \leftsample \{0,1\}^{256}$, flips a coin $b \leftsample \{0, 1\}$, then gives $(c^\ast, K_b)$ to $A$
    \item $A$ unpacks $c^\ast = (c^\prime, t_1, t_2)$, uses the recovered secret key $\sk^\prime$ to decrypt $c^\prime$. From here it is trivial to distinguish whether $K_b$ is the correct session key or the random session key
\end{enumerate}

\end{document}

### 2024-07-21 etm-transform/manuscript/main.tex
Security analysis

--- 2024-07-18 faster-fo-transform/manuscript/main.tex
@@
     \center
-    \begin{minipage}{0.29\textwidth}
     \begin{algorithm}[H]
@@
     \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.32\textwidth}
     \begin{algorithm}[H]
@@
     \end{algorithm}
-    \end{minipage}
-    \hfill
-    \begin{minipage}{0.37\textwidth}
     \begin{algorithm}[H]
@@
     \end{algorithm}
-    \end{minipage}
     \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
@@
 \section{Application to Kyber}
-\subsection{``encrypt-then-MAC'' is not secure with Kyber}
+
 CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \monospace{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem.
@@
 
-\emph{\monospace{CPAPKE} is not \monospace{OW-PCA} secure}: What's important to know about \monospace{CPAPKE} is that it is not secure against plaintext-checking attack.
+The routines of \monospace{CCAKEM} can be found in Algorithm 7, 8, 9 in \cite{avanzi2019crystals}. We recap them here:
+
+\begin{algorithm}[H]
+    \caption{\monospace{Kyber.CCAKEM.KeyGen()}}
+    \begin{algorithmic}[1]
+        \State $z \leftsample \mathcal{B}^{32}$
+            \Comment{Randomly sample 32 bytes (256 bits)}
+        \State $(\pk, \sk^\prime) \leftsample \monospace{Kyber.CPAPKE.KeyGen()}$
+        \State $\sk = (\sk^\prime, \pk, H(\pk), z)$
+            \Comment{H is instantiated with SHA3-256}
+        \State \Return $(\pk, \sk)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{\monospace{Kyber.CCAKEM.Encap(\pk)}}
+    \begin{algorithmic}[1]
+        \State $m \leftsample \mathcal{B}^{32}$
+        \State $m \leftarrow H(m)$
+            \Comment{Do not output system RNG directly}
+        \State $(\bar{K}, r) \leftarrow G(m \Vert H(\pk))$
+            \Comment{G is instantiated with SHA3-512}
+        \State $c \leftarrow \monospace{Kyber.CPAPKE.Enc(\pk, m, r)}$
+            \Comment{Because $r$ is set, \monospace{CPAPKE} is deterministic}
+        \State $K \leftarrow \monospace{KDF}(\bar{K} \Vert H(c))$
+            \Comment{\monospace{KDF} is instantiated with Shake256}
+        \State \Return $(c, K)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{\monospace{Kyber.CCAKEM.Decap(\sk, c)}}
+    \begin{algorithmic}[1]
+        \Require Secret key $\sk = (\sk^\prime, \pk, H(\pk), z)$
+        \Require \monospace{Kyber.CPAPKE} Ciphertext $c$
+        \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
+            \Comment{Unpack the secret key; $h$ is the hash of $\pk$}
+        \State $m^\prime \leftarrow \monospace{Kyber.CPAPKE.Dec}(\sk^\prime, c)$
+        \State $(\overline{K}^\prime, r^\prime) \leftarrow G(m^\prime \Vert h)$
+        \State $c^\prime = \monospace{Kyber.CPAPKE.Enc}(\pk, m^\prime, r^\prime)$
+        \If{$c = c^\prime$}
+            \State $K \leftarrow \monospace{KDF}(\bar{K}^\prime \Vert H(c))$
+        \Else
+            \State $K \leftarrow \monospace{KDF}(z \Vert H(c))$
+        \EndIf
+        \State \Return $K$
+    \end{algorithmic}
+\end{algorithm}
+
+\subsection{``encrypt-then-MAC'' is not secure with Kyber}
+\begin{lemma}
+    There exists efficient \monospace{OW-PCA} adversary against \monospace{CPAPKE}
+\end{lemma}
 

### 2024-08-02 etm-transform/manuscript/main.tex
update

--- 2024-07-18 faster-fo-transform/manuscript/main.tex
@@
 % Custom commands
-\newcommand{\monospace}{\texttt}
-\newcommand{\pke}{\monospace{PKE}}
-\newcommand{\keygen}{\monospace{Gen}}
-\newcommand{\encrypt}{\monospace{Enc}}
-\newcommand{\decrypt}{\monospace{Dec}}
-\newcommand{\kem}{\monospace{KEM}}
-\newcommand{\encap}{\monospace{Encap}}
-\newcommand{\decap}{\monospace{Decap}}
-\newcommand{\etm}{\monospace{EtM}}  % encrypt-then-mac
-\newcommand{\mac}{\monospace{MAC}}
-\newcommand{\sign}{\monospace{Sign}}
-\newcommand{\verify}{\monospace{Verify}}
-\newcommand{\pk}{\monospace{pk}}
-\newcommand{\sk}{\monospace{sk}}
-\newcommand{\pco}{\monospace{PCO}}
-\newcommand{\cvo}{\monospace{CVO}}
+\newcommand{\pke}{\texttt{PKE}}
+\newcommand{\keygen}{\texttt{Gen}}
+\newcommand{\encrypt}{\texttt{Enc}}
+\newcommand{\decrypt}{\texttt{Dec}}
+\newcommand{\kem}{\texttt{KEM}}
+\newcommand{\encap}{\texttt{Encap}}
+\newcommand{\decap}{\texttt{Decap}}
+\newcommand{\etm}{\texttt{EtM}}  % encrypt-then-mac
+\newcommand{\mac}{\texttt{MAC}}
+\newcommand{\sign}{\texttt{Sign}}
+\newcommand{\verify}{\texttt{Verify}}
+\newcommand{\pk}{\texttt{pk}}
+\newcommand{\sk}{\texttt{sk}}
+\newcommand{\pco}{\texttt{PCO}}
+\newcommand{\cvo}{\texttt{CVO}}
 \newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
@@
 \newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
-\newcommand{\fotplus}{\monospace{FOT+}}
+\newcommand{\fotplus}{\texttt{FOT+}}
 
@@
 Given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}.
-We took inspiration from this strategy and applied a similar technique to provide ciphertext integrity for a public-key encryption scheme, which then translates to an \monospace{IND-CCA} secure $\kem$.
+We took inspiration from this strategy and applied a similar technique to provide ciphertext integrity for a public-key encryption scheme, which then translates to an \texttt{IND-CCA} secure $\kem$.
 Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of a tag, which should offer significant performance improvements while maintaining comparable level of security.
@@
 We will prove in section \ref{sec:main-results} that, under the random oracle model, the $\mac$ key is securely hidden behind the hash function, and producing a valid pair of ciphertext and tag without full knowledge of the plaintext constitutes a forgery attack on the message authentication code.
-Thanks to the modular construction in \cite{hofheinz2017modular}, providing ciphertext integrity in the underlying encryption scheme gives us an \monospace{IND-CCA} secure $\kem$ for free.
+Thanks to the modular construction in \cite{hofheinz2017modular}, providing ciphertext integrity in the underlying encryption scheme gives us an \texttt{IND-CCA} secure $\kem$ for free.
 
@@
 We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
-Specifically, we first define the $\monospace{OW-ATK}$ and the $\monospace{IND-CPA}$ game as they pertain to a public key encryption scheme.
-In later section we will define the $\monospace{IND-CCA}$ game as it pertains to a key encapsulation mechanism.
+Specifically, we first define the $\texttt{OW-ATK}$ and the $\texttt{IND-CPA}$ game as they pertain to a public key encryption scheme.
+In later section we will define the $\texttt{IND-CCA}$ game as it pertains to a key encapsulation mechanism.
 
-In the \monospace{OW-ATK} game, an adversary's goal is to recover the decryption of a randomly generated ciphertext.
+In the \texttt{OW-ATK} game, an adversary's goal is to recover the decryption of a randomly generated ciphertext.
 
@@
         \caption{
-            The $\monospace{OW-ATK}$ game
+            The $\texttt{OW-ATK}$ game
         }\label{alg:pke-ow-atk-game}
@@
             \State $c^\ast \leftsample \encrypt(\pk, m)^\ast$
-            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\monospace{ATK}}(1^\lambda, \pk, c^\ast)$
+            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
             \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
@@
 
-    \caption{The \monospace{OW-ATK} game}
+    \caption{The \texttt{OW-ATK} game}
     \label{fig:pke-ow-atk-game}
@@
 
-The adversary $\mathcal{A}$ with access to oracle(s) $\mathcal{O}_\monospace{ATK}$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$.
-The \textit{advantage} $\epsilon_\monospace{OW-ATK}$ of an adversary in this game is the probability that it wins the game.
+The adversary $\mathcal{A}$ with access to oracle(s) $\mathcal{O}_\texttt{ATK}$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$.
+The \textit{advantage} $\epsilon_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game.
 
-The choice of oracle(s) $\mathcal{O}_\monospace{ATK}$ depends on the choice of $\monospace{ATK}$. Specifically:
+The choice of oracle(s) $\mathcal{O}_\texttt{ATK}$ depends on the choice of $\texttt{ATK}$. Specifically:
 
 \begin{equation*}
-    \mathcal{O}_\monospace{ATK} = \begin{cases}
-        - & \monospace{ATK} = \monospace{CPA} \\
-        \monospace{PCO} & \monospace{ATK} = \monospace{PCA} \\
-        \monospace{CVO} & \monospace{ATK} = \monospace{VA} \\
-        \monospace{PCO, CVO} & \monospace{ATK} = \monospace{PCVA}
+    \mathcal{O}_\texttt{ATK} = \begin{cases}
+        - & \texttt{ATK} = \texttt{CPA} \\
+        \texttt{PCO} & \texttt{ATK} = \texttt{PCA} \\
+        \texttt{CVO} & \texttt{ATK} = \texttt{VA} \\
+        \texttt{PCO, CVO} & \texttt{ATK} = \texttt{PCVA}
     \end{cases}
@@
 
-Where the definitions of plaintext-checking oracle $\monospace{PCO}$ and the ciphertext-validation oracle $\monospace{CVO}$ are inherited from \cite{hofheinz2017modular}
+Where the definitions of plaintext-checking oracle $\texttt{PCO}$ and the ciphertext-validation oracle $\texttt{CVO}$ are inherited from \cite{hofheinz2017modular}
 
@@
 
-In the \monospace{IND-CPA} game (algorithm \ref{alg:ind-cpa-game}), an adversary's goal is to distinguish the encryption of one message from the encryption of another message.
+In the \texttt{IND-CPA} game (algorithm \ref{alg:ind-cpa-game}), an adversary's goal is to distinguish the encryption of one message from the encryption of another message.
 Given the public key, the adversary outputs two adversarially chosen messages and obtains the encryption of a random choice between these two messages.
-The adversary wins the \monospace{IND-CPA} game if it correctly identifies which message the encryption is obtained from.
+The adversary wins the \texttt{IND-CPA} game if it correctly identifies which message the encryption is obtained from.
 
@@
             \begin{algorithm}[H]
-                \caption{The $\monospace{IND-CPA}$ game}\label{alg:ind-cpa-game}
+                \caption{The $\texttt{IND-CPA}$ game}\label{alg:ind-cpa-game}
 
@@
         \end{algorithm}
-        }{\caption{The \monospace{IND-CPA} game}}
+        }{\caption{The \texttt{IND-CPA} game}}
     \end{floatrow}
@@
 
-The \textit{advantage} $\epsilon_\monospace{IND-CPA}$ of an \monospace{IND-CPA} adversary $A$ is defined by
+The \textit{advantage} $\epsilon_\texttt{IND-CPA}$ of an \texttt{IND-CPA} adversary $A$ is defined by
 
 $$
-    \monospace{Adv}_\monospace{IND-CPA}(A) = \norm{
+    \texttt{Adv}_\texttt{IND-CPA}(A) = \norm{
         P[\hat{b} = b] - \frac{1}{2}
@@
                 \caption{
-                    \monospace{IND-CCA} game for \kem
+                    \texttt{IND-CCA} game for \kem
                 }\label{alg:kem-ind-cca-game}
@@
                     \State $b \leftsample \{0, 1\}$
-                    \State $\hat{b} \leftsample \mathcal{A}_\monospace{IND-CCA}^{\mathcal{O}_\decap}(
+                    \State $\hat{b} \leftsample \mathcal{A}_\texttt{IND-CCA}^{\mathcal{O}_\decap}(
                         1^\lambda, \pk, c^\ast, k_b
@@
             \end{algorithm}
-        }{\caption{The \monospace{KEM-IND-CCA2} game}}
+        }{\caption{The \texttt{KEM-IND-CCA2} game}}
     \end{floatrow}
@@
 The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key.
-The advantage $\epsilon_\monospace{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\monospace{IND-CCA}$ is defined by
+The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by
 
 \begin{equation*}
-    \epsilon_\monospace{IND-CCA} = \norm{
+    \epsilon_\texttt{IND-CCA} = \norm{
         P[\hat{b} = b] - \frac{1}{2}
@@
 The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$.
-The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\monospace{1}$ if the message-tag pair is valid under the secret key, or $\monospace{0}$ otherwise.
+The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
 
@@
             \end{algorithm}
-        }{\caption{The \monospace{EUF-CMA} game}}
+        }{\caption{The \texttt{EUF-CMA} game}}
     \end{floatrow}
@@
 
-The advantage $\epsilon_\monospace{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
+The advantage $\epsilon_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
 
 \subsection{Modular Fujisaki-Okamoto transformation}
-The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \monospace{OW-CPA} or \monospace{IND-CPA}) and outputs a $\pke$ with stronger security.
-A later variation \cite{hofheinz2017modular} improved the original construction in \cite{fujisaki1999secure} by accounting for decryption failures, tightening security bounds, and providing a modular construction that first transforms \monospace{OW-CPA}/\monospace{IND-CPA} $\pke$ into \monospace{OW-PCVA} $\pke$ by providing ciphertext integrity through re-encryption (the \monospace{T} transformation), then transforming the \monospace{OW-PCVA} $\pke$ into an \monospace{IND-CCA} $\kem$ (the \monospace{U} transformation).
+The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \texttt{OW-CPA} or \texttt{IND-CPA}) and outputs a $\pke$ with stronger security.
+A later variation \cite{hofheinz2017modular} improved the original construction in \cite{fujisaki1999secure} by accounting for decryption failures, tightening security bounds, and providing a modular construction that first transforms \texttt{OW-CPA}/\texttt{IND-CPA} $\pke$ into \texttt{OW-PCVA} $\pke$ by providing ciphertext integrity through re-encryption (the \texttt{T} transformation), then transforming the \texttt{OW-PCVA} $\pke$ into an \texttt{IND-CCA} $\kem$ (the \texttt{U} transformation).
 
@@
 
-Particularly relevant to our results are two variations of the \monospace{U} transformation: $\monospace{U}^\bot$ (KEM with explicit rejection) and $\monospace{U}^{\not\bot}$ (KEM with implicit rejection).
-If $\pke$ is \monospace{OW-PCVA} secure, then $\monospace{U}^\bot$ transforms $\pke$ into an \monospace{IND-CCA} secure $\kem^\bot$:
+Particularly relevant to our results are two variations of the \texttt{U} transformation: $\texttt{U}^\bot$ (KEM with explicit rejection) and $\texttt{U}^{\not\bot}$ (KEM with implicit rejection).
+If $\pke$ is \texttt{OW-PCVA} secure, then $\texttt{U}^\bot$ transforms $\pke$ into an \texttt{IND-CCA} secure $\kem^\bot$:
 
 \begin{theorem}\label{thm:u-bot}
-    For any \monospace{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^\bot$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \monospace{OW-PCVA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ that makes at most $q_H$ queries to $\pco$ and $\cvo$ such that
+    For any \texttt{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^\bot$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \texttt{OW-PCVA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ that makes at most $q_H$ queries to $\pco$ and $\cvo$ such that
 
@@
 
-Similarly, if $\pke$ is \monospace{OW-PCA} secure, then $\monospace{U}^{\not\bot}$ transforms $\pke$ into an \monospace{IND-CCA} secure $\kem^{\not\bot}$
+Similarly, if $\pke$ is \texttt{OW-PCA} secure, then $\texttt{U}^{\not\bot}$ transforms $\pke$ into an \texttt{IND-CCA} secure $\kem^{\not\bot}$
 
 \begin{theorem}\label{thm:u-notbot}
-    For any \monospace{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^{\not\bot}$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \monospace{OW-CPA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ issuing at most $q_H$ queries to $\pco$ such that:
+    For any \texttt{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^{\not\bot}$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \texttt{OW-CPA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ issuing at most $q_H$ queries to $\pco$ such that:
 
@@
 
-The modularity of the \monospace{T} and \monospace{U} transformation allows us to tweak only the \monospace{T} transformation (see section \ref{sec:main-results}), obtain \monospace{OW-PCVA} security, then automatically get \monospace{IND-CCA} security for free.
+The modularity of the \texttt{T} and \texttt{U} transformation allows us to tweak only the \texttt{T} transformation (see section \ref{sec:main-results}), obtain \texttt{OW-PCVA} security, then automatically get \texttt{IND-CCA} security for free.
 This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:mac-performance}).
@@
             \State $m \leftsample \mathcal{M}_\pke$
-                \State $k_\mac \leftarrow G(\pk, m)$
+            \State $k \leftarrow G(m)$
             \State $c_\pke \leftsample \encrypt(\pk, m)$
             \State $t \leftarrow \mac(k, c_\pke)$
-                \State $K \leftarrow H(\pk, m, c_\pke)$
+            \State $K \leftarrow H(m, c_\pke)$
             \State $c \leftarrow (c_\pke, K)$
@@
             \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
-                \State $\hat{k}_\pk \leftarrow G(\pk, \hat{m})$
+            \State $\hat{k} \leftarrow G(\hat{m})$
             \If{$\mac(\hat{k}, c_\pke) \neq t$}
-                    \State \Return $H(\pk, z, c_\pke)$
+                \State \Return $H(z, c_\pke)$
             \EndIf
-                \State \Return $H(\pk, \hat{m}, c_\pke)$
+            \State \Return $H(\hat{m}, c_\pke)$
         \end{algorithmic}
     \end{algorithm}
-    \end{minipage}
     \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
@@
 
-Here are a few design rationale:
-\begin{enumerate}
-    \item \emph{Deriving MAC key and session key}.
-We choose to include the public key $\pk$ when deriving the MAC key $k_\mac$ and the session key $K$ for similar reason stated in \cite{bos2018crystals}.
-If the MAC key is derived solely from the message, then an adversary can pre-compute a large lookup table mapping MAC key to the source plaintext that can applied to \emph{all sessions}.
-When the adversary intercepts a ciphertext $c = (c_\pke, t)$, it can brute-force all possible MAC keys to recover the plaintext.
-Since brute-forcing MAC key on a known ciphertext-tag pair is an offline search, given a sufficiently large lookup table and a sufficiently large quantum computer, this search might be feasible.
-On the other hand, if the MAC key is derived from both the public key and the plaintext, then the adversary will need to compute the large lookup table and run the key search \emph{per session}, which greatly increases the cost of the attack.
-    \item \emph{Not hashing unauthenticated ciphertext $c_\pke$}: \cite{avanzi2019crystals}\cite{bos2018crystals} derives the session key from a hash of the ciphertext $K \leftarrow \monospace{KDF}(\overline{K} \Vert H(c_\pke))$ to \emph{``simplify implementation with non-incremental hash APIs''}.
-Since ``hashing the ciphertext separately'' vs ``deriving session key directly from ciphertext''  does not affect the security nor the performance (since the ciphertext will be hashed somewhere) of the scheme, we opt to not complicate the design with unnecessary hashes and leave the implementation considerations to concrete instantiations.
-    \item \emph{Not adding key confirmation $\sigma \leftarrow \mac(\cdot, K)$}: within the security model of a KEM, there is no interaction between the KEM adversary and any other party (be it the challenger or any oracle) that involves the adversary sending a session key.
-In other words, there is no scenario in which tempering with any session key is meaningful, which means that the session key does not need protection.
-\end{enumerate}
-
-\subsection{If PKE is OW-PCA secure, then EtM is IND-CCA2 secure}
-
 \begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
-    For every \monospace{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes at most $q_D$ decapsulation queries, there exists an \monospace{OW-PCA} adversary $B$ against the underlying $\pke$ such that
+    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes at most $q_D$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes $q_P$ plaintext-checking queries against the underlying $\pke$ such that
 
     \begin{equation*}
-        \monospace{Adv}_\monospace{IND-CCA2}(A) \leq q_D \cdot \epsilon_\mac + 2 \cdot \monospace{Adv}_\monospace{OW-PCA}(B)
+        \texttt{Adv}_\texttt{IND-CCA2}(A) \leq q_D \cdot \epsilon_\mac + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
     \end{equation*}
@@
     \begin{figure}[H]
-        \center
-        \begin{minipage}{0.55\textwidth}
+        \centering
         \begin{algorithm}[H]
@@
                 \State $k^\ast \leftarrow G(m^\ast)$
-                        \Comment{Game 0-1}
+                    \Comment{$G_0$-$G_1$}
                 \State $k^\ast \leftsample \mathcal{K}_\mac$
-                        \Comment{Game 2-3}
-                    \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
-                    \State $t^\ast \leftarrow \mac(k^\ast, c^\ast)$
-                    \State $K_0 \leftarrow H(m^\ast, c^\ast)$
-                        \Comment{Game 0-2}
+                    \Comment{$G_2$-$G_3$}
+                \State $c^\ast_\pke \leftsample \encrypt(\pk, m^\ast)$
+                \State $t^\ast \leftarrow \mac(k^\ast, c^\ast_\pke)$
+                \State $c^\ast \leftarrow (c^\ast_\pke, t^\ast)$
+                \State $K_0 \leftarrow H(m^\ast, c^\ast_\pke)$
+                    \Comment{$G_0$-$G_2$}
                 \State $K_0 \leftsample \mathcal{K}_\kem$
-                        \Comment{Game 3}
+                    \Comment{$G_3$}
                 \State $K_1 \leftsample \mathcal{K}_\kem$
@@
                 \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(
-                        1^\lambda, \pk, (c^\ast, t^\ast), K_b)$
-                        \Comment{Game 0}
+                    1^\lambda, \pk, c^\ast, K_b)$
+                    \Comment{$G_0$}
                 \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(
-                        1^\lambda, \pk, (c^\ast, t^\ast), K_b)$
-                        \Comment{Game 1-3}
+                    1^\lambda, \pk, c^\ast, K_b)$
+                    \Comment{$G_1$-$G_3$}
                 \State \Return $\llbrack \hat{b} = b \rrbrack$
@@
         \end{algorithm}
-        \end{minipage}
-        \hfill
-        \begin{minipage}{0.4\textwidth}
         \begin{algorithm}[H]
-                \caption{$\mathcal{O}^\decap(c, t)$}\label{alg:etm-decap-oracle}
+            \caption{$\mathcal{O}^\decap(c)$}\label{alg:etm-decap-oracle}
             \begin{algorithmic}[1]
-                    \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c)$
+                \State $(c_\pke, t) \leftarrow c$
+                \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
                 \State $\hat{k} \leftarrow G(\hat{m})$
-                    \If{$\mac(\hat{k}, c) = t$}
-                        \State \Return $H(\hat{m}, c)$
+                \If{$\mac(\hat{k}, c_\pke) = t$}
+                    \State \Return $H(\hat{m}, c_\pke)$
                 \EndIf
-                    \State \Return $H(z, c)$
+                \State \Return $H(z, c_\pke)$
             \end{algorithmic}
@@
         \begin{algorithm}[H]
-                \caption{$\mathcal{O}^\decap_1(c, t)$}\label{alg:etm-decap-oracle-1}
+            \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:etm-decap-oracle-1}
             \begin{algorithmic}[1]
-                    % TODO: this is ugly
-                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G :$ \\
-                        $\;\;\;\;\decrypt(\sk_\pke, c) = \tilde{m}$ \\
-                        $\;\;\;\;\;\land\; \mac(\tilde{k}, c) = t$}
-                        \State \Return $H(\tilde{m}, c)$
+                \State $(c_\pke, t) \leftarrow c$
+                \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G 
+                    : \decrypt(\sk_\pke, c_\pke) = \tilde{m} 
+                    \land \mac(\tilde{k}, c_\pke) = t$}
+                    \State \Return $H(\tilde{m}, c_\pke)$
                 \EndIf
-                    \State \Return $H(z, c)$
+                \State \Return $H(z, c_\pke)$
             \end{algorithmic}
         \end{algorithm}
-        \end{minipage}
-        \caption{Sequence of games}\label{fig:sequence-of-games}
+        \caption{Sequence of games, true decap oracle $\mathcal{O}^\decap$ and simulated oracle $\mathcal{O}^\decap_1$}\label{fig:sequence-of-games}
     \end{figure}
 
-    % TODO: finish the proof
-    \textbf{Need to finish the rest of the proof}
+    \emph{Game 0} is the standard \texttt{IND-CCA2} game for a key encapsulation mechanism.
+
+    \emph{Game 1} is identical to \emph{Game 0} except for that the decapsulation oracle $\mathcal{O}^\decap$ (algorithm \ref{alg:etm-decap-oracle}) is replaced with a simulated decapsulation oracle $\mathcal{O}^\decap_1$ (algorithm \ref{alg:etm-decap-oracle-1}).
+If $\mathcal{O}^\decap_1$ accepts the queried ciphertext $c = (c_\pke, t)$ and outputs the true session key $K \leftarrow H(\tilde{m}, c_\pke)$, then the queried ciphertext must be honestly generated, which means that $\mathcal{O}^\decap$ must also accept the queried ciphertext and output the true session key.
+If $\mathcal{O}^\decap$ rejects the queried ciphertext $c = (c_\pke, t)$ and outputs the implicit rejection $K \leftarrow H(z, c_\pke)$, then the tag $t$ is invalid under the MAC key $k \leftarrow G(\decrypt(\sk_\pke, c_\pke))$.
+Since for a given ciphertext $c_\pke$, the correct MAC key is fixed, there could not be a matching hash query $(m, k)$ such that $m$ is the correct decryption and $k$ can validate the incorrect tag.
+Therefore, $\mathcal{O}^\decap_1$ must also reject the queried ciphertext and output the implicit rejection.
+
+    This means that game 0 and game 1 differ when $\mathcal{O}^\decap$ accepts the queried ciphertext $c = (c_\pke, t)$ but $\mathcal{O}^\decap_1$ rejects it, which means that $t$ is a valid tag for $c_\pke$ under the correct MAC key $k \leftarrow G(\decrypt(\sk_\pke, c_\pke))$ but such key is never queried by the adversary.
+Under the random oracle model, from the adversary's perspective, such $k$ is an unknown and uniformly random key, so producing a valid tag under such key constitutes a forgery against the MAC.
+Denote the probability of forgery against unknown uniformly random MAC key by $\epsilon_\mac$, then the probability that the two decapsulation oracles disagree on one or more queries is at most $q_D \cdot \epsilon_\mac$.
+Finally, by the difference lemma,
+
+    \begin{equation*}
+        \texttt{Adv}_0(A) - \texttt{Adv}_1(A) \leq q_D \cdot \epsilon_\mac
+    \end{equation*}
+
 \end{proof}
@@
 \subsection{If PKE is not OW-PCA, then EtM is not IND-CCA2 secure}\label{sec:not-owpca-implies-not-indcca2}
-More specifically, if there exists an efficient \monospace{OW-PCA} adversary $B$ against the underlying $\pke$, then we can build an efficient \monospace{IND-CCA2} adversary $A$ who uses $B$ as a sub-routine and wins the \monospace{IND-CCA2} game.
+More specifically, if there exists an efficient \texttt{OW-PCA} adversary $B$ against the underlying $\pke$, then we can build an efficient \texttt{IND-CCA2} adversary $A$ who uses $B$ as a sub-routine and wins the \texttt{IND-CCA2} game.
 
 To run $B$ as a sub-routine, $A$ needs to be able to simulate a plaintext-checking oracle, which can be done using the decapsulation oracle.
-When presented with a plaintext-chekcing query $(\tilde{m}, \tilde{c})$, $A$ derives the corresponding $\mac$ key $\tilde{k} \leftarrow G(\tilde{m})$ and signs the ciphertext $\tilde{t} \leftarrow \mac(\tilde{k}, \tilde{c})$.
+When presented with a plaintext-checking query $(\tilde{m}, \tilde{c})$, $A$ derives the corresponding $\mac$ key $\tilde{k} \leftarrow G(\tilde{m})$ and signs the ciphertext $\tilde{t} \leftarrow \mac(\tilde{k}, \tilde{c})$.
 $A$ then queries the decapsulation oracle on the ciphertext-tag pair $\tilde{K} \leftarrow \mathcal{O}^\decap(\tilde{c}, \tilde{t})$.
@@
 If $\hat{m}$ is not the correct decryption of $c^\ast$, then the expected session key will probably not match anything, so $A$ will always claim $K_b$ to be a random key.
-In other words, \emph{if $B$ wins the \monospace{OW-PCA} game, then $A$ wins the \monospace{IND-CCA2} game; if $B$ does not win, then $A$'s chance of winning is exactly $\frac{1}{2}$}.
+In other words, \emph{if $B$ wins the \texttt{OW-PCA} game, then $A$ wins the \texttt{IND-CCA2} game; if $B$ does not win, then $A$'s chance of winning is exactly $\frac{1}{2}$}.
 Therefore, the advantage of $A$ is at least that of $B$.
@@
 \section{Application to Kyber}
+
+\subsection{Kyber with AE}
+CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \texttt{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem.
+For the construction of the \texttt{IND-CCA2} secure KEM, Kyber first constructs an \texttt{IND-CPA} public-key encryption scheme (which we will call \texttt{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \texttt{CCAKEM}).
+Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation.
+The routines of \texttt{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.
+
+The routines of \texttt{CCAKEM} can be found in Algorithm 7, 8, 9 in \cite{avanzi2019crystals}.
+We modify Algorithms 8 and 9 using authenticated encryption (AE)  mode as follows where Algorithm 7 is unchanged.
+
+\begin{algorithm}[H]
+    \caption{\texttt{Kyber.CCAKEM.KeyGen()}}
+    \begin{algorithmic}[1]
+        \State $z \leftsample \mathcal{B}^{32}$
+            \Comment{Randomly sample 32 bytes (256 bits)}
+        \State $(\pk, \sk^\prime) \leftsample \texttt{Kyber.CPAPKE.KeyGen()}$
+        \State $\sk = (\sk^\prime, \pk, H(\pk), z)$
+            \Comment{H is instantiated with SHA3-256}
+        \State \Return $(\pk, \sk)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}
+    \begin{algorithmic}[1]
+        \State $m \leftsample \mathcal{B}^{32}$
+       % \State $m'= H(m)$
+            \Comment{Do not output system RNG directly}
+        \State $(\bar{K}, r, K_{mac}) = G(m' \Vert H(\pk))$
+            \Comment{G is instantiated with SHA3-512}
+        \State $c  \leftarrow \texttt{Kyber.CPAPKE.Enc(\pk, m', r)}$
+            \Comment{Because $r$ is set, \texttt{CPAPKE} is deterministic}
+            \State $t_1=MAC(K_{mac}, c)$
+        \State $K = \texttt{KDF}(\bar{K} \Vert t_1)$
+       % \State $t_2=MAC(K, t_1)$
+            \Comment{\texttt{KDF} is instantiated with Shake256}
+        %\State $(c, t_1, t_2)$
+        \State $c \leftarrow (c, t_1)$
+        \State \Return $(c, K)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c, t_1, t_2)$}
+    \begin{algorithmic}[1]
+        \Require Secret key $\sk = (\sk^\prime, \pk, H(\pk), z)$
+        \Require \texttt{Kyber.CPAPKE} Ciphertext $c$ and Tags $t_1, t_2$
+        \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
+            \Comment{Unpack the secret key; $h$ is the hash of $\pk$}
+        \State $(c, t_1) \leftarrow c$
+        \State $ m = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c)$ 
+       % \State $m'=H(m) $ 
+        \State $(\overline{K},   r, K_{mac}) =   G(m^\prime \Vert h)$
+        \State $t_1'=\mac(K_\mac, c)$
+        \If{$t_1' = t_1$}
+            \State $K  = \texttt{KDF}(\bar{K} \Vert t_1)$ 
+            %\State $t_2'= \mac(K, t_1)$
+            \State \Return $K$
+         \Else 
+            \State Abort
+        \EndIf
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{remark} If $c$ is manipulated, then the verification of  $t_1$  will be failed.
+In this case, there is no $K$ outputted from the decap$^+$.
+So the attacks described in the following subsections won't work.
+We also added the key confirmation, which is tag $t_2$.
+\end{remark}
+
+{\color{blue} Note for authenticated encryption, tags $t_i$'s are necessary for the inputs.
+
+In fact, this is authenticated encryption instead of EtM. So we should change that, called authenticated encryption. 
+Please do not change my notation. They have their meanings in AE mode. }
+
 \subsection{``encrypt-then-MAC'' is not secure with Kyber}
-CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \monospace{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem.
-For the construction of the \monospace{IND-CCA2} secure KEM, Kyber first constructs an \monospace{IND-CPA} public-key encryption scheme (which we will call \monospace{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \monospace{CCAKEM}).
-Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation.
-The routines of \monospace{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.
+\begin{lemma}
+    There exists efficient \texttt{OW-PCA} adversary against \texttt{CPAPKE}
+\end{lemma}
 
-\emph{\monospace{CPAPKE} is not \monospace{OW-PCA} secure}: What's important to know about \monospace{CPAPKE} is that it is not secure against plaintext-checking attack.
-
-The main reason why a plaintext-checking attack works against \monospace{CPAPKE} is that for each known honest plaintext-ciphertext pair $(m, c)$, one can perturb the ciphertext and obtain a new ciphertext $c^\prime \leftarrow \monospace{addNoise}(c)$.
-Where the perturbation is small, the perturbed ciphertext can still correctly decrypt to the original plaintext $\decrypt_\monospace{CPAPKE}(\sk, c^\prime) = \decrypt_\monospace{CPAPKE}(\sk, c) = m$.
+The main reason why a plaintext-checking attack works against \texttt{CPAPKE} is that for each known honest plaintext-ciphertext pair $(m, c)$, one can perturb the ciphertext and obtain a new ciphertext $c^\prime \leftarrow \texttt{addNoise}(c)$.
+Where the perturbation is small, the perturbed ciphertext can still correctly decrypt to the original plaintext $\decrypt_\texttt{CPAPKE}(\sk, c^\prime) = \decrypt_\texttt{CPAPKE}(\sk, c) = m$.
 Where the perturbation is large, the perturbed ciphertext will not decrypt to the original plaintext.
@@
 On the other hand, \cite{hofheinz2017modular} prevents tempering by the combination of \emph{de-randomization} and \emph{re-encryption}.
-Even if $c^\prime \leftarrow \monospace{addNoise}(c)$ still decrypts back to the same plaintext, \emph{de-randomization} and \emph{re-encryption} will reveal that $c^\prime$ has been tempered with because $c^\prime \neq \encrypt(\pk, m; r\leftarrow G(m))$.
+Even if $c^\prime \leftarrow \texttt{addNoise}(c)$ still decrypts back to the same plaintext, \emph{de-randomization} and \emph{re-encryption} will reveal that $c^\prime$ has been tempered with because $c^\prime \neq \encrypt(\pk, m; r\leftarrow G(m))$.
 Unlike $\kem_\etm$ where an adversary can temper with the unauthenticated ciphertext and still produce some valid authenticated ciphertexts, with \emph{re-encryption} \emph{there is no ``other valid ciphertext'' that correspond to the same plaintext}, thus preventing all tempering.
@@
 \subsection{Is ``encrypt-then-MAC'' useful?}
-From \cite{hofheinz2017modular} we know that if a $\pke$ is rigid and \monospace{OW-CPA} secure, then the $U^{\not\bot}_m$ transformation is sufficient for constructing an \monospace{IND-CCA2} secure $\kem$ with tight security reduction.
+From \cite{hofheinz2017modular} we know that if a $\pke$ is rigid and \texttt{OW-CPA} secure, then the $U^{\not\bot}_m$ transformation is sufficient for constructing an \texttt{IND-CCA2} secure $\kem$ with tight security reduction.
 Furtherfore, the performance overhead of $U^{\not\bot}_m$ over the rigid $\pke$ is minimal.
-Note that rigidity and \monospace{OW-CPA} automatically implies \monospace{OW-PCA} security, so $\kem_\etm$ will also construct an \monospace{IND-CCA2} secure $\kem$, but with extra runtime and communication overhead, so there is practical point of using $\etm$ over the $U^{\not\bot}_m$ transformation.
-On the other hand, if the $\pke$ is not \monospace{OW-PCA}, then $\kem_\etm$ is not \monospace{IND-CCA2} secure.
+Note that rigidity and \texttt{OW-CPA} automatically implies \texttt{OW-PCA} security, so $\kem_\etm$ will also construct an \texttt{IND-CCA2} secure $\kem$, but with extra runtime and communication overhead, so there is practical point of using $\etm$ over the $U^{\not\bot}_m$ transformation.
+On the other hand, if the $\pke$ is not \texttt{OW-PCA}, then $\kem_\etm$ is not \texttt{IND-CCA2} secure.
 
-If the $\pke$ is not rigid but still \monospace{OW-PCA} secure, \emph{de-randomization + re-encryption} is still an option, so we need to consider the performance trade-offs.
+If the $\pke$ is not rigid but still \texttt{OW-PCA} secure, \emph{de-randomization + re-encryption} is still an option, so we need to consider the performance trade-offs.
 \emph{de-randomization + re-encryption} adds one addition hash (for deriving coin) into the encryption routine and add one hash call and one call to the input encryption routine in the decryption routine.
@@
         \hline
-        \monospace{OW-PCA} but not rigid & $\etm$ \\
+        \texttt{OW-PCA} but not rigid & $\etm$ \\
         \hline
-        \monospace{OW-CPA} but not \monospace{OW-PCA} & \emph{de-randomization + re-encryption} \\
+        \texttt{OW-CPA} but not \texttt{OW-PCA} & \emph{de-randomization + re-encryption} \\
         \hline

### 2024-08-02 etm-transform/manuscript/main.tex
Proof that EtM is IND-CCA2 if PKE is OW-PCA

@@
 \begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
-    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes at most $q_D$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes $q_P$ plaintext-checking queries against the underlying $\pke$ such that
+    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes $q_D$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes at least $q_D$ plaintext-checking queries against the underlying $\pke$ such that
 
@@
 
+    \emph{Game 2} is identical to \emph{Game 1}, except for that when the challenger generates the challenge ciphertext $c^\ast = (c^\ast_\pke, t^\ast)$, the tag $t^\ast$ is computed using a uniformly random key $k^\ast \leftarrow \mathcal{K}_\mac$ instead of a pseudorandom key derived from hashing the challenge plaintext.
+
+    Under the random oracle model, game 2 and game 1 are statistically identical to the adversary $A$, unless $A$ queries $G$ with $m^\ast$.
+Denote the probability that $A$ queries $G$ with $m^\ast$ by $P[\texttt{QUERY G}^\ast]$, then:
+
+    \begin{equation*}
+        \texttt{Adv}_1(A) - \texttt{Adv}_2(A) \leq P[\texttt{QUERY G}^\ast]
+    \end{equation*}
+
+    \emph{Game 3} is identical to \emph{Game 2}, except for that $K_0$ is a uniformly random session key instead of a pseudorandom session key derived from the challenge plaintext-ciphertext pair.
+Under the random oracle model, game 3 and game 2 are statistically identical unless the adversary $A$ queries $H$ with $(m^\ast, \cdot)$.
+Denote the probability that $A$ makes such $H$ query by $P[\texttt{QUERY H}^\ast]$, then:
+
+    \begin{equation*}
+        \texttt{Adv}_2(A) - \texttt{Adv}_3(A) \leq P[\texttt{QUERY H}^\ast]
+    \end{equation*}
+
+    In game 3, both $K_0$ and $K_1$ are uniformly random. There is no statistical difference between the two session keys, so no adversary can have any advantage: $\texttt{Adv}_3(A) = 0$.
+
+    Now consider an \texttt{OW-PCA} adversary $B$ simulating game 3 for $A$: \begin{enumerate}
+        \item When $B$ receives its public key $\pk$, $B$ passes $\pk$ to $A$
+        \item $B$ can sample the implicit rejection $z$ by itself
+        \item $B$ can simulate both hash oracles $G$ and $H$ for $A$
+        \item $B$ can simulate $\mathcal{O}^\decap_1$ for $A$.
+Instead of checking if $\decrypt(\sk_\pke, c) = \tilde{m}$, $B$ can use its access to the plaintext-checking oracle and check if $\pco(\tilde{m}, c) = 1$.
+This means that for every decapsulation query $B$ services, $B$ needs to make at least one plaintext-checking query.
+Therefore, $B$ needs to make at least $q_D$ plaintext-checking query.
+        \item When $B$ receives its challenge ciphertext $c_\pke^\ast$, it can sample a uniformly random key $k^\ast \leftsample \mathcal{K}_\mac$, produce the corresponding tag $t^\ast \leftarrow \mac(k^\ast, c^\ast_\pke)$, and sample a uniformly random session keys $K_0, K_1 \leftsample \mathcal{K}_\kem$.
+$B$ then passes $c^\ast = (c^\ast_\pke, t^\ast)$ as the challenge ciphertext and a coin-flip $K_b$ as the challenge session key.
+    \end{enumerate}
+
+    If $A$ ever queries $G$ or $H$ with the decryption of $c^\ast_\pke$, $B$ will be able to detect it using the plaintext-checking oracle.
+From where $B$ is guaranteed to win the \texttt{OW-PCA} game.
+Therefore:
+
+    \begin{equation*}
+        \begin{aligned}
+            P[\texttt{QUERY G}^\ast] &\leq \texttt{Adv}_\texttt{OW-PCA}(B) \\
+            P[\texttt{QUERY H}^\ast] &\leq \texttt{Adv}_\texttt{OW-PCA}(B)
+        \end{aligned}
+    \end{equation*}
+
+    Combining all inequalities above, we have:
+
+    \begin{equation*}
+        \texttt{Adv}_0(A) \leq q_D \cdot \epsilon_\mac + 2 \texttt{Adv}_\texttt{OW-PCA}(B)
+    \end{equation*}
 \end{proof}
 
-\subsection{If PKE is not OW-PCA, then EtM is not IND-CCA2 secure}\label{sec:not-owpca-implies-not-indcca2}
-More specifically, if there exists an efficient \texttt{OW-PCA} adversary $B$ against the underlying $\pke$, then we can build an efficient \texttt{IND-CCA2} adversary $A$ who uses $B$ as a sub-routine and wins the \texttt{IND-CCA2} game.
-
-To run $B$ as a sub-routine, $A$ needs to be able to simulate a plaintext-checking oracle, which can be done using the decapsulation oracle.
-When presented with a plaintext-checking query $(\tilde{m}, \tilde{c})$, $A$ derives the corresponding $\mac$ key $\tilde{k} \leftarrow G(\tilde{m})$ and signs the ciphertext $\tilde{t} \leftarrow \mac(\tilde{k}, \tilde{c})$.
-$A$ then queries the decapsulation oracle on the ciphertext-tag pair $\tilde{K} \leftarrow \mathcal{O}^\decap(\tilde{c}, \tilde{t})$.
-
-Because the correct session key is deterministically derived from the public key, the plaintext, the ciphertext, and some other values derived from these three values, for each set of $(\tilde{m}, \tilde{c}, \tilde{t}, \tilde{K})$, $A$ can correctly derive the session key if $\tilde{m}$ is indeed the decryption of $\tilde{c}$.
-On the other hand, if $\tilde{m} \neq \decrypt_\pke(\sk_\pke, \tilde{c})$, then the decapsulation oracle will return the implicit rejection value, which will not match the expected session key value.
-
-By comparing the decapsulation orcale's output and the expected session key, $A$ can correctly distinguish whether $\tilde{m}$ is the decryption of $\tilde{c}$ or not.
-Thus $A$ can correctly simulate the plaintext-checking oracle for $B$.
-
-When $A$ receives the challenge ciphertext $(c^\ast, t^\ast)$ and unknown session key $K_b \leftsample \{K_0, K_1\}$, $A$ passes $c^\ast$ as the challenge ciphertext to $B$ as $B$'s challenge ciphertext.
-After $B$ returns the guess $\hat{m}$, $A$ computes the expected session key using $\hat{m}$ and $c^\ast$.
-If $\hat{m}$ is the correct decryption of $c^\ast$, then the expected session key should match the correct session key, in which case $A$ will correctly distinguish the true session key from the random session key.
-If $\hat{m}$ is not the correct decryption of $c^\ast$, then the expected session key will probably not match anything, so $A$ will always claim $K_b$ to be a random key.
-In other words, \emph{if $B$ wins the \texttt{OW-PCA} game, then $A$ wins the \texttt{IND-CCA2} game; if $B$ does not win, then $A$'s chance of winning is exactly $\frac{1}{2}$}.
-Therefore, the advantage of $A$ is at least that of $B$.
+%%%%%%%%%%%%%%%% uncomment to see %%%%%%%%%%%%%%%%%%%%
+% \include{not-pca-means-not-indcca}
 
@@
 In fact, this is authenticated encryption instead of EtM. So we should change that, called authenticated encryption. 
-Please do not change my notation. They have their meanings in AE mode. }
+Please do not change my notation. They have their meanings in AE mode.}
 
-\subsection{``encrypt-then-MAC'' is not secure with Kyber}
-\begin{lemma}
-    There exists efficient \texttt{OW-PCA} adversary against \texttt{CPAPKE}
-\end{lemma}
-
-The main reason why a plaintext-checking attack works against \texttt{CPAPKE} is that for each known honest plaintext-ciphertext pair $(m, c)$, one can perturb the ciphertext and obtain a new ciphertext $c^\prime \leftarrow \texttt{addNoise}(c)$.
-Where the perturbation is small, the perturbed ciphertext can still correctly decrypt to the original plaintext $\decrypt_\texttt{CPAPKE}(\sk, c^\prime) = \decrypt_\texttt{CPAPKE}(\sk, c) = m$.
-Where the perturbation is large, the perturbed ciphertext will not decrypt to the original plaintext.
-The boundary between "small perturbation" and "large perturbation" depends on the values of the secret $\mathbf{s}$, so an adversary can probe the boundary in each of the $k \times n$ dimensions and learn the value of $\mathbf{s}$ coefficient-by-coefficient.
-
-As described in section \ref{sec:not-owpca-implies-not-indcca2}, when the integrity of the unauthenticated ciphertext is protected under a MAC key derived from the corresponding plaintext, \textbf{the MAC will not prevent an adversary from perturbing an honest ciphertext} since the adversary can use the same MAC key to compute tag $t^\prime$ on the perturbed unauthenticated ciphertext $c^\prime$.
-
-On the other hand, \cite{hofheinz2017modular} prevents tempering by the combination of \emph{de-randomization} and \emph{re-encryption}.
-Even if $c^\prime \leftarrow \texttt{addNoise}(c)$ still decrypts back to the same plaintext, \emph{de-randomization} and \emph{re-encryption} will reveal that $c^\prime$ has been tempered with because $c^\prime \neq \encrypt(\pk, m; r\leftarrow G(m))$.
-Unlike $\kem_\etm$ where an adversary can temper with the unauthenticated ciphertext and still produce some valid authenticated ciphertexts, with \emph{re-encryption} \emph{there is no ``other valid ciphertext'' that correspond to the same plaintext}, thus preventing all tempering.
-
-\subsection{Is ``encrypt-then-MAC'' useful?}
-From \cite{hofheinz2017modular} we know that if a $\pke$ is rigid and \texttt{OW-CPA} secure, then the $U^{\not\bot}_m$ transformation is sufficient for constructing an \texttt{IND-CCA2} secure $\kem$ with tight security reduction.
-Furtherfore, the performance overhead of $U^{\not\bot}_m$ over the rigid $\pke$ is minimal.
-Note that rigidity and \texttt{OW-CPA} automatically implies \texttt{OW-PCA} security, so $\kem_\etm$ will also construct an \texttt{IND-CCA2} secure $\kem$, but with extra runtime and communication overhead, so there is practical point of using $\etm$ over the $U^{\not\bot}_m$ transformation.
-On the other hand, if the $\pke$ is not \texttt{OW-PCA}, then $\kem_\etm$ is not \texttt{IND-CCA2} secure.
-
-If the $\pke$ is not rigid but still \texttt{OW-PCA} secure, \emph{de-randomization + re-encryption} is still an option, so we need to consider the performance trade-offs.
-\emph{de-randomization + re-encryption} adds one addition hash (for deriving coin) into the encryption routine and add one hash call and one call to the input encryption routine in the decryption routine.
-\emph{``encrypt-then-MAC''} introduces one hash call and one MAC call to the encryption routine, adds a tag to the ciphertext size, and adds one hash call and one MAC call to the decryption routine.
-Using a one-time MAC like GMAC or Poly1305, the computational cost of producing a tag and the communication cost of the tag should be minimal (1000-2000 CPU cycles + 128 bits of tag).
-Where appropriate, replacing \emph{re-encryption} in the decryption routine with a MAC computation can lead to substantial saving in computational cost.
-This is especially true in protocols where the client (often constrained environments) is responsible for running decryption routine, such as hybrid key exchanges used in CECPQ2.
-
-\begin{table}[H]
-    \begin{tabular}{|c|c|}
-        \hline
-        Input $\pke$ & what should I use to build $\kem$ \\
-        \hline
-        rigid & $U^{\not\bot}_m$ \\
-        \hline
-        \texttt{OW-PCA} but not rigid & $\etm$ \\
-        \hline
-        \texttt{OW-CPA} but not \texttt{OW-PCA} & \emph{de-randomization + re-encryption} \\
-        \hline
-    \end{tabular}
-    \caption{Use cases}
-\end{table}
+% \include{when-is-etm-useful}
 

### 2024-08-06 etm-transform/cca-elgamal/main.tex
cca elgamal

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}

% Custom commands
\usepackage{crypto-primitives}
\usepackage{kyber-algos}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    ElGamal cryptosystem
}
\author{
    Ganyu (Bruce) Xu
}
% Leave the date field empty to display the date of compilation
% \date{}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{The ElGamal cryptosystem}
The ElGamal cryptosystem is a public key encryption scheme that mainly operates on the discrete log problem. Each instance of the encryption scheme is parameterized by a cyclic group $G$ with prime order $q$, a generator $g$ of this cyclic group. The routines of the encryption scheme is shown in figure \ref{fig:cpa-elgamal-routines}

\begin{figure}[H]
    \begin{algorithm}[H]
        \caption{$\keygen$}
        \begin{algorithmic}[1]
            \State $x \leftsample \mathbb{Z}_q$
            \State $u \leftarrow g^x$
            \State $\pk \leftarrow u, \sk \leftarrow x$
            \State \Return $(\pk, \sk)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\encrypt(\pk = u, m \in G)$}
        \begin{algorithmic}[1]
            \State $y \leftsample \mathbb{Z}_q$
            \State $v \leftarrow g^y$
            \State $w \leftarrow u^y$
                \Comment{$w = g^{xy}$}
            \State $c \leftarrow (v, m \cdot w)$
            \State \Return $c$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\decrypt(\sk = x, c)$}
        \begin{algorithmic}[1]
            \State $(c_1, c_2) \leftarrow c$
            \State $\hat{w} \leftarrow c_1^x$
            \State $\hat{m} \leftarrow c_2 \cdot \hat{w}^{-1}$
            \State \Return $\hat{m}$
        \end{algorithmic}
    \end{algorithm}
    \caption{ElGamal encryption scheme is IND-CPA secure if DDH holds}\label{fig:cpa-elgamal-routines}
\end{figure}

The IND-CPA security of the ElGamal cryptosystem depends on the hardness of the following two problems:

\begin{definition}[Computational Diffie-Hellman Problem]
    Let $G$ be a cyclic group with prime order $q$ and generator $g$. Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $g, g^x,g^y$, compute $g^{xy}$
\end{definition}

\begin{definition}[Decisional Diffie-Hellman Problem]
    Let $G$ be a cyclic group with prime order $q$ and generator $g$. Let $x, y, z \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $g, g^x, g^y$, distinguish $g^{xy}$ from $g^z$
\end{definition}

\begin{theorem}
    For every IND-CPA adversary $A$ against the ElGamal cryptosystem, there exists an adversary $B$ against the DDH game such that

    \begin{equation*}
        \texttt{Adv}(A) = 2 \cdot \texttt{Adv}(B)
    \end{equation*}
\end{theorem}

Because ElGamal ciphertexts are malleable, this encryption scheme is not secure against chosen-ciphertext attacks. However, a hybrid encryption scheme can be used to achieve chosen-ciphertext attack security \cite{boneh2020graduate}. Denote this construction by ``ElGamal HPKE''.

To construct the HPKE, we need the cyclic group $G$ of prime order $q$ and generator $g$. We also need a symmetric cipher $(\encrypt_S, \decrypt_S)$ defined over $(\mathcal{K}, \mathcal{M}, \mathcal{C})$, and a hash function $H: G \rightarrow \mathcal{K}$. The routines are listed in figure \ref{fig:cca-elgamal-routines}.

\begin{figure}[H]
    \begin{algorithm}[H]
        \caption{$\keygen$}
        \begin{algorithmic}[1]
            \State $x \leftarrow \mathbb{Z}_q$
            \State $u \leftarrow g^x$
            \State $\pk \leftarrow u$
            \State $\sk \leftarrow x$
            \State \Return $(\pk, \sk)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\encrypt(\pk = u, m \in \mathcal{M})$}
        \begin{algorithmic}[1]
            \State $y \leftsample \mathbb{Z}_q$
            \State $v \leftarrow g^y$
            \State $w \leftarrow u^y$
                \Comment{$w = g^{xy}$}
            \State $k \leftarrow H(w)$
            \State $c^\prime \leftarrow \encrypt_S(k, m)$
            \State $c \leftarrow (v, c^\prime)$
            \State \Return $c$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\decrypt(\sk = x, c)$}
        \begin{algorithmic}[1]
            \State $(v, c^\prime) \leftarrow c$
            \State $\hat{w} \leftarrow v^x$
            \State $\hat{k} \leftarrow H(\hat{w})$
            \State $\hat{m} \leftarrow \decrypt_S(\hat{k}, c^\prime)$
            \State \Return $\hat{m}$
        \end{algorithmic}
    \end{algorithm}
    \caption{ElGamal HPKE}\label{fig:cca-elgamal-routines}
\end{figure}

\begin{theorem}
    For every IND-CCA adversary $A$ against the HPKE, there exists an \emph{interactive computational Diffie-Hellman problem} adversary $B$ and an IND-CPA adversary $C$ against the symmetric encryption scheme such that

    \begin{equation*}
        \texttt{Adv}(A) \leq \text{NEED TO WRITE THIS PART}
    \end{equation*}
\end{theorem}

While having a decryption oracle breaks the decisional Diffie-Hellman assumption, we still feel confident that the computational Diffie-Hellman remains hard, which is how we can reason about the security of the HPKE under chosen-ciphertext attacks.

Unfortunately, \emph{this is not the case in Kyber}. Having a decapsulation oracle that can take arbitrary number of decapsulation queries will allows an adversary to complete recover the secret key, unlike ElGamal HPKE where having a decryption oracle does not give away the secret key. \emph{There is no immediate parallel loosening of security assumption we can make in Kyber}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}

### 2024-08-13 etm-transform/tls-12-vs-13/main.tex
Explanation of TLS 1.2 vs 1.3

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}

% Custom commands
\usepackage{crypto-primitives}
\usepackage{kyber-algos}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    Handshake workflow in TLS 1.2 vs 1.3
}
\author{
    Ganyu (Bruce) Xu
}
% Leave the date field empty to display the date of compilation
% \date{}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{Handshake workflow in TLS 1.2 (RFC 5246)}
TLS 1.2 supports three main key exchange workflows: long-term Diffie-Hellman (DH and ECDH), RSA, and ephemeral Diffie-Hellman (DHE and ECDHE). Depending on the choice of key exchange method, the handshake workflow may vary, but there are three invariants: \begin{itemize}
    \item Some key exchange so client and server can establish shared secret over insecure channel
    \item Server needs to authenticate itself by presenting some kind of public key signed by a certificate authority, then prove posession of the corresponding secret key
    \item Some kind of key confirmation to confirm that client and server indeed have the same secret
\end{itemize}

\subsection{With long-term Diffie-Hellman}
\begin{enumerate}
    \item \textbf{Client sends \texttt{ClientHello}}: \\ in \texttt{ClientHello}, client specifies that it supports \texttt{DH\_*} for key exchange
    \item \textbf{Server responds with \texttt{ServerHello}} \\ in \texttt{ServerHello}, server chooses \texttt{DH\_*} for key exchange
    \item \textbf{Server sends \texttt{Certificate}} \\ \texttt{Certificate} contains server's Diffie-Hellman parameters $g^y$ signed by the certificate authority (RFC 5246 A.5). Also note that with a \texttt{DH\_*} key exchange, \textit{server must not send \texttt{ServerKeyExchange}}, or the client will abort the handshake (7.4.3) 
    \item \textbf{Client sends \texttt{ClientKeyExchange}} \\ \texttt{ClientKeyExchange} contains client's Diffie-Hellman parameters. \\ After server receives \texttt{ClientKeyExchange}, client and server have agreed on a pre-master secret, and the handshake traffic keys are derived from pre-master secret. 
    \item \textbf{Client and server send each other \texttt{Finished}} \\ \texttt{Finished} contains a MAC over all previous messages signed using a secret key derived from the pre-master secret.
\end{enumerate}

With long-term Diffie-Hellman, \textbf{server authentication is implicit}. Server demonstrates possession of the Diffie-Hellman secret key that correspond to the public key in \texttt{Certificate} by correctly computing the shared secret, then use the shared secret to compute a valid tag, which is presented in \texttt{Finished}. If the server does not have the corresponding secret key, then it cannot produce a valid tag, so the client will abort the handshake after receiving an invalid \texttt{Finished} from the server.

\subsection{With RSA}
\begin{enumerate}
    \item \textbf{Client sends \texttt{ClientHello}}: \\ in \texttt{ClientHello}, client specifies that it supports \texttt{RSA\_*} for key exchange
    \item \textbf{Server responds with \texttt{ServerHello}} \\ in \texttt{ServerHello}, server chooses \texttt{RSA\_*} for key exchange
    \item \textbf{Server sends \texttt{Certificate}} \\ \texttt{Certificate} contains server's RSA public key (an encryption key) signed by the certificate authority (RFC 5246 A.5). Similar to long-term Diffie-Hellman key exchange, \textit{server must not send \texttt{ServerKeyExchange}}.
    \item \textbf{Client sends \texttt{ClientKeyExchange}} \\ \texttt{ClientKeyExchange} contains the RSA ciphertext generated by the client using the server's public key in the certificate. \\ After server receives \texttt{ClientKeyExchange}, client and server have agreed on a pre-master secret.
    \item \textbf{Client and server send each other \texttt{Finished}} \\ \texttt{Finished} contains a MAC over all previous messages signed using a secret key derived from the pre-master secret.
\end{enumerate}

In RSA key exchange, \textbf{server authentication} is also implicit just like with long-term Diffie-Hellman key exchange.

\subsection{With ephemeral Diffie-Hellman}
\begin{enumerate}
    \item \textbf{Client sends \texttt{ClientHello}}: \\ in \texttt{ClientHello}, client specifies that it supports \texttt{DHE\_*} for key exchange
    \item \textbf{Server responds with \texttt{ServerHello}} \\ in \texttt{ServerHello}, server chooses \texttt{DHE\_*} for key exchange
    \item \textbf{Server sends \texttt{Certificate}} \\ \texttt{Certificate} contains server's public key (a digital signature signing key) signed by the certificate authority
    \item \textbf{Server sends \texttt{ServerKeyExchange}} \\ \texttt{ServerKeyExchange} contains server's ephemeral Diffie-Hellman parameters and a digital signature of these server parameters signed using server's secret key corresponding to the public key in server certificate
    \item \textbf{Client sends \texttt{ClientKeyExchange}} \\ contains client's ephemeral Diffie-Hellman parameters 
    \item \textbf{CLient and server send each other \texttt{Finished}}
\end{enumerate}

In DHE, \textbf{server authentication is explicit}: server presents its public key (a verification key) in the server certificate, which is signed by the the certificate authority and thus bound to the server's identity. Server sends demonstrates possession of the corresponding secret key (a signing key) by signing client secret, server secret, and server DHE parameters in \texttt{ServerKeyExchange}.

\section{Key exchange workflow in TLS 1.3}
TLS 1.3 differs from 1.2 in that \textbf{TLS 1.3 deprecated all long-term Diffie-Hellman and RSA key exchange methods; ephemeral Diffie-Hellman key exchange is the only supported key exchange}. The handshake workflow thus differs from TLS 1.2.

\begin{enumerate}
    \item \textbf{Client sends \texttt{ClientHello}} \begin{itemize}
        \item The \texttt{cipher\_suites} field contains the supported set of cipher suites, but \textit{this field no longer specifies the key exchange methods}, only the AEAD scheme (AES-128-GCM, AES-256-GCM or ChaCha20-Poly1305) and HKDF (SHA256 or SHA384)
        \item The \texttt{random} field contains client's random values
        \item The \texttt{signature\_algorithms} extension specifies the type of signature the client supports for \texttt{CertificateVerify} (aka server's signature).
        \item The \texttt{signature\_algorithms\_sert} extension specifies the type of signature the client supports for \texttt{Certificate} (aka certificate authority's signature)
        \item The \texttt{supported\_group} extension specifies the Diffie-Hellman groups that the client supports
        \item The \texttt{key\_share} extension specifies client's ephemeral Diffie-Hellman parameters ($g^x$)
    \end{itemize}

    \item \textbf{Server responds with \texttt{ServerHello}} \begin{itemize}
        \item The \texttt{random} field contains server's random values
        \item The \texttt{key\_share} extension contains server's chosen Diffie-Hellman group and the corresponding ephemeral parameters
    \end{itemize}

    \item \textbf{Server sends \texttt{Certificate}} \\ In TLS 1.3, \textit{server's certificate must contain a digital signature's verification key}. This public key is bound to the server's identity by certificate authority's signature

    \item \textbf{Server sends \texttt{CertificateVerify}} \\ This message contains a signature over all handshake messages up to this point, signed using the server's secret key. \textbf{This authenticates the server explicitly}.

    \item Client authentication is optional in TLS. If client authentication is needed, then server requests client's certificate with \texttt{CertificateRequest}, and client respond with client's authentication messages \texttt{Certificate} and \texttt{CertificateVerify}.
    \item Client and server send each other \texttt{Finished}, which contains a MAC over the handshake transcript signed using the shared secret
\end{enumerate}

\section{Incorporating post-quantum KEM into TLS 1.3}
While there is no standardization on how post-quantum KEMs will be incorporated into TLS 1.3, Google and Cloudflare have done some \href{https://blog.cloudflare.com/the-tls-post-quantum-experiment/}{experimental implementations} in 2019. At the time the chosen KEMs were \texttt{HRSSS} and \texttt{SIKE}, though there is nothing special about these two KEMs, so any other KEM can be a drop-in replacement. The handshake workflow will be as follows:

\begin{enumerate}
    \item \texttt{ClientHello} \begin{itemize}
        \item In the \texttt{supported\_groups} extensions, add a group that indicates client's support for post-quantum (hybrid) key exchange, such as \texttt{Kyber768X25519}
        \item In the \texttt{key\_share} extension, include a key share entry whose value is the concatenation of classical and PQ public keys \texttt{key\_share = X25519\_public\_value || ephemeral\_KEM\_public\_key}
    \end{itemize}

    \item \texttt{ServerHello} \begin{itemize}
        \item In the \texttt{key\_share} extension, server's response is the concatenation of server's ephemeral Diffie-Hellman parameters and the ciphertext produced by running the encapsulation routine using client's ephemeral encapsulation key
    \end{itemize}
\end{enumerate}

The pre-master secret is derived from the concatenation of the Diffie-Hellman shared secret and the KEM shared secret.

In this setup, a fresh KEM key pair is generated by the client at each handshake, which means that each decapsulation key is used exactly once. If the decapsulation routine rejects the ciphertext explicitly, or if the client and the server's pre-master secret disagree, then somewhere along the handshake protocol the client will abort the handshake. When the handshake is restarted, another key pair will be generated. Therefore, at least in the context of TLS, \textbf{limiting a CCA adversary to exactly one decryption query} is a reasonable security goal.

\end{document}

### 2024-08-19 etm-transform/manuscript/main.tex
things I need to work out

@@
 \section{Application to Kyber}
+CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism that first constructs an IND-CPA secure public key encryption scheme whose security is based on the Module Learning with Error Problem (MLWE), then applies a generic transformation using \emph{de-randomization} and \emph{re-encryption} \cite{hofheinz2017modular}.
+The use of \emph{re-encryption} for providing rigidity means that the decapsulation routine needs to run the encryption routine for verifying ciphertext integrity.
+Unfortunately for Kyber, the encryption routine consumes both more CPU cycles and carries a larger memory footprint.
+Therefore, applying an alternate transformation that replaces \emph{re-encryption} with a MAC verification will bring substantial performance enhancement.
+This is particularly applicable to communication protocols such as TLS 1.3, where clients, often constrained environments, need to run the decapsulation routine.
 
-\subsection{Kyber with AE}
-CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} and ML-KEM \cite{key2023mechanism} are \texttt{IND-CCA2} secure key encapsulation mechanisms whose security depends on the hardness of the Module Learning with Error (MLWE) problem.
-For the construction of the \texttt{IND-CCA2} secure KEM, Kyber first constructs an \texttt{IND-CPA} public-key encryption scheme (which we will call \texttt{CPAPKE}), then applies the modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} to construct the key encapsulation mechanism (which we will call \texttt{CCAKEM}).
-Specifically, Kyber's round-3 submission uses the $U^{\not\bot}$ transformation, while ML-KEM uses the $U^{\not\bot}_m$ transformation.
-The routines of \texttt{CPAPKE} can be found in Algorithm 4, 5, 6 in \cite{avanzi2019crystals} and are largely identical between Kyber and ML-KEM.
+The \texttt{IND-CPA} PKE of Kyber (algorithms 4, 5, 6 in \cite{avanzi2019crystals}) is not \texttt{OW-PCA} secure.
+A plaintext-checking attack \cite{ravi2019generic} can recover the secret key using a few maliciously constructed plaintext-checking queries for each coefficient of a Kyber secret key (there are 512, 768, and 1024 coefficients in the secret key depending on the desired security level).
+However, we propose mitigation for plaintext-checking attack at protocol level by requiring each keypair to be used for decryption only once.
+If a decryption fails, then the secret key should be discarded and the key exchange terminated, and a new keypair should be generated to restart the key exchange.
+In fact, such ephemeral key exchange is already required in TLS 1.3 for forward secrecy, so requiring the one-time use of keypair does not introduce additional operational cost for protocols such as TLS 1.3.
 
@@
 
-\subsection{MAC Performance}\label{sec:mac-performance}
-We claim that the input $\mac$ only needs to be one-time existentially unforgeable.
-This is because besides the challenge ciphertext $(c^\ast, t^\ast)$, the adversary has no external resources from which it can obtain authenticated ciphertexts for which it does not know the decryption.
-Here we compare the performance of a variety of $\mac$ instantiations.
-Some of them are many-time secure while others are one-time secure.
-The standalone performance results are listed in table \ref{tbl:standalone-mac-perf}.
-For each choice of MAC, we checked the median (top) and average (bottom) CPU cycles (run on a 2019 MacBook Pro 16-inch) needed to sign 768, 1088, and 1568 bytes of data (respectively the ciphertext size for Kyber512, Kyber768, and Kyber1024).
+\subsection{Choosing message authentication code}\label{sec:mac-performance}
+Given that each keypair is only used once, an adversary can only obtain one authenticated ciphertext, which means that the corresponding MAC key is also only used once per key exchange.
+This allows us to use MAC constructions that are only one-time secure for substantial computation efficiency gains.
+We compared the median and mean number of CPU cycles (Intel i7-9750H) used to sign Kyber's IND-CPA ciphertexts.
+The results are listed in \ref{tbl:standalone-mac-perf}.
 

### 2024-08-19 etm-transform/manuscript/main.tex
Scaffolding for experimental data

@@
 \begin{algorithm}[H]
-    \caption{\texttt{Kyber.CCAKEM.KeyGen()}}
+    \caption{\texttt{Kyber.CCAKEM.KeyGen()}}\label{alg:kyber-ae-keygen}
     \begin{algorithmic}[1]
@@
 \begin{algorithm}[H]
-    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}
+    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}\label{alg:kyber-ae-encap}
     \begin{algorithmic}[1]
@@
 \begin{algorithm}[H]
-    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c, t_1, t_2)$}
+    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c, t_1, t_2)$}\label{alg:kyber-ae-decap}
     \begin{algorithmic}[1]
@@
 
-\subsection{Choosing message authentication code}\label{sec:mac-performance}
-Given that each keypair is only used once, an adversary can only obtain one authenticated ciphertext, which means that the corresponding MAC key is also only used once per key exchange.
-This allows us to use MAC constructions that are only one-time secure for substantial computation efficiency gains.
-We compared the median and mean number of CPU cycles (Intel i7-9750H) used to sign Kyber's IND-CPA ciphertexts.
-The results are listed in \ref{tbl:standalone-mac-perf}.
+\subsection{MAC performance}\label{sec:mac-performance}
+When instantiating an instance $\kem(\keygen, \encap, \decap)$, there are a variety of possible MAC's to choose from.
+For each of the security level and choice of MAC, we measured the number of CPU cycles needed to produce a digest of the unauthenticated ciphertext.
+The median (top) and mean (bottom) measurements are reported in table \ref{tbl:standalone-mac-perf}.
 
@@
     \centering
-    \begin{tabular}{|c|c|c|c|c|}
+    \begin{tabular}{|c|c|c|c|c|c|}
         \hline
-        Name & Security & 768 bytes & 1088 bytes & 1568 bytes \\
+        Name & Security & measurement & 768 bytes & 1088 bytes & 1568 bytes \\
         \hline
-        CMAC & many-time & 5022 & 5442 & 6090 \\
-        & & 5131 & 5578 & 6154 \\
+        CMAC & many-time & median & 5022 & 5442 & 6090 \\
+        & & mean & 5131 & 5578 & 6154 \\
         \hline
-        GMAC & one-time & 2778 & 2756 & 2762 \\
-        & & 2843 & 2780 & 2919 \\
+        KMAC-256 & many-time & median & 7934 & 9862 & 11742 \\
+        & & mean & 8594 & 10693 & 12319 \\
         \hline
-        KMAC-256 & many-time & 7934 & 9862 & 11742 \\
-        & & 8594 & 10693 & 12319 \\
+        GMAC & one-time & median & 2778 & 2756 & 2762 \\
+        & & mean & 2843 & 2780 & 2919 \\
         \hline
-        Poly1305 & one-time & 1128 & 1218 & 1338 \\
-        & & 1435 & 1504 & 1625 \\
+        Poly1305 & one-time & median & 1128 & 1218 & 1338 \\
+        & & mean & 1435 & 1504 & 1625 \\
         \hline
@@
     \caption{Standalone $\mac$ performances}\label{tbl:standalone-mac-perf}
+\end{table}
+
+Based on the security reduction in section \ref{sec:main-results} we chose standalone GMAC and Poly1305 for their substantial performance advantage over other constructions such as CBC-MAC and KMAC.
+We then modified the reference implementation (TODO: citation needed https://github.com/pq-crystals/kyber) according to algorithms \ref{alg:kyber-ae-keygen}, \ref{alg:kyber-ae-encap}, and \ref{alg:kyber-ae-decap}.
+We measured the median and mean CPU cycles needed to run each of the routines.
+The measurements are listed in table \ref{tbl:kyber-ae-perf}
+
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|}
+        \hline
+        Name & Security level & measurement & $\keygen$ & $\encap$ & $\decap$ \\
+        \hline
+        Kyber512 & 128 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        Kyber768 & 192 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        Kyber1024 & 256 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        KyberAE512 w/ GMAC& 128 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        KyberAE768 w/ GMAC& 192 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        KyberAE1024 w/ GMAC& 256 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        KyberAE512 w/ Poly1305& 128 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        KyberAE768 w/ Poly1305& 192 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+        KyberAE1024 w/ Poly1305& 256 bits & median &  & &  \\
+        & & mean &  & &  \\
+        \hline
+    \end{tabular}
+    \caption{Kyber-AE performance measurements}\label{tbl:kyber-ae-perf}
 \end{table}

### 2024-08-19 etm-transform/manuscript/main.tex
short discussion on probability of blind forgery

@@
 
+    Note that $\epsilon_\mac$ quantifies the probability that an adversary can produce forgery for a unknown key without access to a signing oracles.
+While this is not a standard security definition for MAC, this probability is straightforward to estimate for some classes of MACs.
+As will be discussed in section \ref{sec:mac-performance}, with a Carter-Wegman-like one-time MAC instantiated with message length $L$ and a finite field with $F$ elements, such probability is at most $\epsilon_\mac \leq \frac{L+1}{F}$
+
     \emph{Game 2} is identical to \emph{Game 1}, except for that when the challenger generates the challenge ciphertext $c^\ast = (c^\ast_\pke, t^\ast)$, the tag $t^\ast$ is computed using a uniformly random key $k^\ast \leftarrow \mathcal{K}_\mac$ instead of a pseudorandom key derived from hashing the challenge plaintext.

### 2024-08-20 etm-transform/manuscript/main.tex
Revised Kyber-AE routines
Some improvements on intro and preliminaries

@@
-% TODO: instead of signing the ciphertext, how about signing a hash of the ciphertext?
-% TODO: Kyber round 3's implementation already contains H(c), we can just use that for the MAC
-% TODO: need to look into using GCM as a more efficient MAC
-% TODO: what is the quantum security of the various hash functions and how do they impact the strength of the FO transform?
-% TODO: Need to read the rest of the U transformation since Kyber round 3 uses U^{\not\bot} but ML-KEM (FIPS 203) uses U^{\not\bot}_m, which removed the hash of the ciphertext
-
-%%%% IACR Transactions TEMPLATE %%%%
-% This file shows how to use the iacrtrans class to write a paper.
-% Written by Gaetan Leurent gaetan.leurent@inria.fr (2020)
-% Public Domain (CC0)
-
 %%%% 1. DOCUMENTCLASS %%%%
@@
 \newcommand{\pke}{\texttt{PKE}}
-\newcommand{\keygen}{\texttt{Gen}}
+\newcommand{\keygen}{\texttt{KeyGen}}
 \newcommand{\encrypt}{\texttt{Enc}}
@@
 \newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
+\newcommand{\adv}{\texttt{Adv}}
 \newcommand{\fotplus}{\texttt{FOT+}}
@@
 %%%% 4. TITLE %%%%
-\title{
+\title[IND-CCA2 secure KEM using encrypt-then-MAC]{
 Fast Fujisaki-Okamoto transformation  using encrypt-then-mac and applications to Kyber}
@@
   The modular Fujisaki-Okamoto (FO) transformation takes public-key encryption with weaker security and constructs a key encapsulation mechanism (KEM) with indistinguishability under adaptive chosen ciphertext attacks.
-While the modular FO transform enjoys tight security bound and quantum resistance, it also suffers from computational inefficiency due to using de-randomization and re-encryption for providing ciphertext integrity.
-In this work, we propose an alternative modular FO transformation that replaces re-encryption with a message authentication code (MAC) and prove the security bound of our construction.
-We then instantiate a concrete instance with ML-KEM and show that when re-encryption incurs significant computational cost, our construction provides substantial runtime speedup and reduced memory footprint.
+While the modular FO transform enjoys tight security bound and quantum resistance, it also suffers from computational inefficiency and vulnerabilities to side-channel attacks due to using de-randomization and re-encryption for providing ciphertext integrity.
+In this work, we propose an alternative KEM construction that achieves ciphertext integrity using a message authentication code (MAC) and instantiate a concrete instance using ML-KEM.
+Our experimental results showed that where the encryption routine incurs heavy computational cost, replacing re-encryption with MAC provides substantial performance improvements at comparable security level.
 \end{abstract}
@@
 The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks.
-Later works extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (standardized by NIST into ML-KEM \cite{key2023mechanism}).
+Later works \cite{hofheinz2017modular} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals}, FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
 
-The current state of the FO transformation enjoys tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many open questions.
-One such problem is the use of re-encryption for providing ciphertext integrity \cite{bernstein2018towards}, which requires the decryption/decapsulation to run the encryption routine as a subroutine.
-In many post-quantum schemes, such as Kyber, the encryption routine is substantially computationally more expensive than the decryption routine.
+The current state of the FO transformation enjoys tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many deficiencies to be improved on.
+One such shortcoming is the use of re-encryption for providing ciphertext integrity \cite{bernstein2018towards}, which requires the decapsulation routine to run the encryption routine as a subroutine.
+In many post-quantum schemes, such as Kyber, the encryption routine is substantially more expensive than the decryption routine, so running the encryption routine in the decapsulation routine inflates the computational cost of the decapsulator.
+In addition, running the encryption as a subroutine introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key \cite{ravi2019generic}\cite{ueno2022curse}.
 
-The problem of ciphertext integrity was solved in symmetric cryptography.
-Given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}.
-We took inspiration from this strategy and applied a similar technique to provide ciphertext integrity for a public-key encryption scheme, which then translates to an \texttt{IND-CCA} secure $\kem$.
-Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of a tag, which should offer significant performance improvements while maintaining comparable level of security.
+The problem of ciphertext integrity was solved in symmetric cryptography: given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}.
+We took inspiration from this strategy and applied a similar technique to transform an IND-CPA secure public-key encryption scheme into an IND-CCA2 secure key encapsulation mechanism.
+Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of an authenticator, which offers significant performance improvements while maintaining comparable level of security.
 
-The main challenge in applying ``encrypt-then-mac'' to public-key cryptography is the lack of a pre-shared $\mac$ key.
-We proposed to derive the shared $\mac$ key by hashing the plaintext message.
-We will prove in section \ref{sec:main-results} that, under the random oracle model, the $\mac$ key is securely hidden behind the hash function, and producing a valid pair of ciphertext and tag without full knowledge of the plaintext constitutes a forgery attack on the message authentication code.
-Thanks to the modular construction in \cite{hofheinz2017modular}, providing ciphertext integrity in the underlying encryption scheme gives us an \texttt{IND-CCA} secure $\kem$ for free.
+The main challenge in applying ``encrypt-then-mac'' to public-key cryptography is the lack of a pre-shared symmetric key.
+We proposed to derive the symmetric key by hashing the plaintext message.
+In section \ref{sec:main-results}, we prove that under the random oracle model, if the input public-key encryption scheme is one-way secure against plaintext-checking attack and the input message authentication code is one-time existentially unforgeable, then the transformed key encapsulation mechanism is IND-CCA2 secure.
 
-In section \ref{sec:mac-performance}, we instantiate concrete instances of our proposed transformation by modifying ML-KEM.
-We will demonstrate that, at the cost of small increase in encryption runtime and ciphertext size, our construction reduces both the runtime and memory footprint of the decryption routine.
-% TODO: need to write more?
+In section \ref{sec:experimental-results}, we instantiate concrete instances of our constructions by combining Kyber with GMAC and Poly1305.
+Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal overhead in the encapsulation routine.
 
@@
 \subsection{Public-key encryption scheme}
-We define a public key encryption scheme $\pke$ to be a collection of three routines $(\keygen, \encrypt, \decrypt)$ defined over a finite message space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
-Many encryption routines are probabilistic, and we define their source of randomness to come from some coin space $\mathcal{R}$.
+A public key encryption scheme $\pke$ is a collection of three routines $(\keygen, \encrypt, \decrypt)$ defined over some message space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
+Where the encryption routine is probabilistic, the source of randomness is denoted by the coin space $\mathcal{R}$.
 
@@
 
-\subsubsection{Correctness}
-It is common to require a $\pke$ to be perfectly correct, meaning that for all possible keypairs $(\pk, \sk)$ and plaintext messages $m \in \mathcal{M}$, $\decrypt(\sk, \encrypt(\pk, m)) = m$ at all times.
-However, some encryption schemes, including many popular lattice-based schemes, admit a non-zero probability of decryption failure: $\decrypt(\sk, \encrypt(\pk, m)) \neq m$.
-Furthermore, \cite{hofheinz2017modular} and \cite{avanzi2019crystals} explained how decryption failure played a role in an adversary's advantage.
-In this paper, we inherit the definition for correctness from \cite{hofheinz2017modular}:
-
-\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
-    A public key encryption scheme $\pke$ is $\delta$-correct if
-
-    \begin{equation*}
-        \mathbf{E}[
-            \max_{m \in \mathcal{M}} P[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)]
-        ] \leq \delta
-    \end{equation*}
-
-    Where the expectation is taken over the probability distribution of keypairs $(\pk, \sk) \leftsample \keygen()$
-\end{definition}
-
-\subsubsection{Security}
 We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
-Specifically, we first define the $\texttt{OW-ATK}$ and the $\texttt{IND-CPA}$ game as they pertain to a public key encryption scheme.
+Specifically, we first define the $\texttt{OW-ATK}$ as they pertain to a public key encryption scheme.
 In later section we will define the $\texttt{IND-CCA}$ game as it pertains to a key encapsulation mechanism.
 
-In the \texttt{OW-ATK} game, an adversary's goal is to recover the decryption of a randomly generated ciphertext.
-
-\begin{figure}
-    \centering
-
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
             \begin{algorithm}[H]
@@
                     \State $m^\ast \leftsample \mathcal{M}$
-            \State $c^\ast \leftsample \encrypt(\pk, m)^\ast$
+                    \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
                     \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
@@
             \end{algorithm}
-
-    \caption{The \texttt{OW-ATK} game}
-    \label{fig:pke-ow-atk-game}
-\end{figure}
-
-The adversary $\mathcal{A}$ with access to oracle(s) $\mathcal{O}_\texttt{ATK}$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$.
-The \textit{advantage} $\epsilon_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game.
-
-The choice of oracle(s) $\mathcal{O}_\texttt{ATK}$ depends on the choice of $\texttt{ATK}$. Specifically:
-
-\begin{equation*}
-    \mathcal{O}_\texttt{ATK} = \begin{cases}
-        - & \texttt{ATK} = \texttt{CPA} \\
-        \texttt{PCO} & \texttt{ATK} = \texttt{PCA} \\
-        \texttt{CVO} & \texttt{ATK} = \texttt{VA} \\
-        \texttt{PCO, CVO} & \texttt{ATK} = \texttt{PCVA}
-    \end{cases}
-\end{equation*}
-
-Where the definitions of plaintext-checking oracle $\texttt{PCO}$ and the ciphertext-validation oracle $\texttt{CVO}$ are inherited from \cite{hofheinz2017modular}
-
-\begin{figure}
-    \begin{floatrow}
+        }
+        {\caption{The \texttt{OW-ATK} game}}
         \ffigbox{
@@
         }
-        {\caption{The Plaintext-Checking Oracle $\pco$}}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\cvo(c \in \mathcal{C})$}\label{alg:cvo}
-                \begin{algorithmic}[1]
-                    \State \Return $\llbrack \decrypt(\sk, c) \in \mathcal{M} \rrbrack$
-                \end{algorithmic}
-            \end{algorithm}
-        }
-        {\caption{the Ciphertext-Validation Oracle $\cvo$}}
+        {\caption{Plaintext-checking oracle}}
     \end{floatrow}
@@
 
-In the \texttt{IND-CPA} game (algorithm \ref{alg:ind-cpa-game}), an adversary's goal is to distinguish the encryption of one message from the encryption of another message.
-Given the public key, the adversary outputs two adversarially chosen messages and obtains the encryption of a random choice between these two messages.
-The adversary wins the \texttt{IND-CPA} game if it correctly identifies which message the encryption is obtained from.
+In the \texttt{OW-ATK} game (see figure \ref{alg:pke-ow-atk-game}), an adversary's goal is to recover the decryption of a randomly generated ciphertext.
+A challenger randomly samples a keypair and a challenge plaintext $m^\ast$, encrypts the challenge plaintext $c^\ast \leftsample \encrypt(\pk, m^\ast)$, then gives $\pk$ and $c^\ast$ to the adversary $A$.
+The adversary $A$, with access to some oracle $\mathcal{O}_\texttt{ATK}$, outputs a guess decryption $\hat{m}$.
+$A$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$.
+The \textit{advantage} $\adv_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game:
 
-\begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{The $\texttt{IND-CPA}$ game}\label{alg:ind-cpa-game}
+\begin{equation*}
+    \adv_\texttt{OW-ATK}(A) = P\left[
+        A(\pk, c^\ast) = m^\ast \vert
+        (\pk, \sk) \leftsample \keygen() ;
+        m^\ast \leftsample \mathcal{M};
+        c^\ast \leftsample \encrypt(\pk, m^\ast)
+    \right]
+\end{equation*}
 
-                \begin{algorithmic}[1]
-                    \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                    \State $(m_0, m_1) \leftsample \mathcal{A}(a^\lambda, \pk)$
-                    \State $b \leftsample \{0, 1\}$
-                    \State $c^\ast \leftsample \encrypt(\pk, m_b)$
-                    \State $\hat{b} \leftsample \mathcal{A}(1^\lambda, \pk, c^\ast)$
-                \State \Return $\llbrack b = \hat{b} \rrbrack$
-            \end{algorithmic}
-        \end{algorithm}
-        }{\caption{The \texttt{IND-CPA} game}}
-    \end{floatrow}
-\end{figure}
+The capabilities of the oracle $\mathcal{O}_\texttt{ATK}$ depends on the choice of security goal $\texttt{ATK}$.
+Particularly relevant to our result is security against plaintext-checking attack (PCA), for which the adversary has access to a plaintext-checking oracle (PCO) (see figure \ref{alg:pco}).
+A PCO takes as input a plaintext-ciphertext pair $(m, c)$ and returns \texttt{True} if $m$ is the decryption of $c$ or \texttt{False} otherwise.
 
-The \textit{advantage} $\epsilon_\texttt{IND-CPA}$ of an \texttt{IND-CPA} adversary $A$ is defined by
+\subsection{Key encapsulation mechanism (KEM)}
+A key encapsulation mechanism is a collection of three routines $(\keygen, \encap, \decap)$ defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
+The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$.
+$\encap(\pk)$ is a probabilistic routine that takes a public key $\pk$ and outputs a pair of values $(c, K)$ where $c \in \mathcal{C}$ is the ciphertext (also called encapsulation) and $K \in \mathcal{K}$ is the shared secret (also called session key).
+$\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $K$ if the ciphertext is valid.
+Some KEM constructions use explicit rejection, where if $c$ is invalid then $\decap$ will return a rejection symbol $\bot$; other KEM constructions use implicit rejection, where if $c$ is invalid then $\decap$ will return a ``fake'' session key that depends on the ciphertext and some other secret values.
 
-$$
-    \texttt{Adv}_\texttt{IND-CPA}(A) = \norm{
-        P[\hat{b} = b] - \frac{1}{2}
-    }
-$$
-
-\subsection{Key encapsulation mechanism}
-A key encapsulation mechanism $\kem$ is a collection of three routines $(\keygen, \encap, \decap)$ defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
-The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$.
-$\encap(\pk)$ is a probabilistic routine that takes a public key $\pk$ and outputs a pair of values $(c, K)$ where $c \in \mathcal{C}$ is the encapsulation (or ciphertext) of the shared secret $k \in \mathcal{K}$.
-$\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $k$ if the ciphertext is valid, or the rejection symbol $\bot$ if the ciphertext is invalid.
-
-The IND-CCA security of a $\kem$ is defined by an adversarial game in which an adversary's goal is to distinguish pseudorandom shared secret (generated by running the $\encap$ routine) and a truly random value.
+The IND-CCA security of a KEM is defined by an adversarial game in which an adversary's goal is to distinguish pseudorandom shared secret (generated by running the $\encap$ routine) and a truly random value.
 
@@
                     \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                    \State $(c^\ast, k_0) \leftsample \encap(\pk)$
-                    \State $k_1 \leftsample \mathcal{K}$
+                    \State $(c^\ast, K_0) \leftsample \encap(\pk)$
+                    \State $K_1 \leftsample \mathcal{K}$
                     \State $b \leftsample \{0, 1\}$
-                    \State $\hat{b} \leftsample \mathcal{A}_\texttt{IND-CCA}^{\mathcal{O}_\decap}(
-                        1^\lambda, \pk, c^\ast, k_b
+                    \State $\hat{b} \leftsample A^{\mathcal{O}_\decap}(
+                        1^\lambda, \pk, c^\ast, K_b
                     )$
@@
         }{\caption{The \texttt{KEM-IND-CCA2} game}}
+        \ffigbox{
+            \begin{algorithm}[H]
+            \caption{
+                $\mathcal{O}_\decap(c)$
+            }
+            \begin{algorithmic}[1]
+                \State \Return $\decap(\sk, c)$
+            \end{algorithmic}
+            \end{algorithm}
+        }{\caption{Decapsulation oracle}}
     \end{floatrow}
@@
 \begin{equation*}
-    \epsilon_\texttt{IND-CCA} = \norm{
-        P[\hat{b} = b] - \frac{1}{2}
+    \adv_\texttt{IND-CCA}(A) = \norm{
+        P[A^{\mathcal{O}_\decap}(a^\lambda, \pk, c^\ast, K_b) = b] - \frac{1}{2}
     }
@@
 
-\subsection{Message authentication code}
+\subsection{Message authentication code (MAC)}
 A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
@@
 The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
+Many MAC constructions are deterministic.
+For these constructions it is simpler to denote the signing routine by $t \leftarrow \mac(k, m)$ and perform verification using a simple comparison.
 
@@
 
-The advantage $\epsilon_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
+The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
 
-\subsection{Modular Fujisaki-Okamoto transformation}
+We are also interested in one-time MAC, whose security goals are identical to normal MAC, except the signing oracle will only answer one signing query.
+In practice this means that the adversary only gets to see one message-tag pair, and the owner of the secret key only signs one message.
+One way to construct an unforgeable one-time MAC is to use a universal hash function (UHF), which computes the digest using finite field arithmetic.
+Each instance of the one-time MAC is parameterized by some maximal message length $L$ and the finite field $\mathbb{F}$ in which arithmetics are performed.
+The tag is computed by evaluating a polynomial whose coefficients are derived from the message:
+
+\begin{equation*}
+    \begin{aligned}
+    &H(k_1, (m_1, m_2, \ldots, m_l)) = k_1^{l+1} + k_1^{l}m_1 + k_1^{l-1}m_2 + \ldots + k_1m_l \\
+    &\mac((k_1, k_2), m) = H(k_1, m) + k_2
+    \end{aligned}
+\end{equation*}
+
+We claim without proof (though proof can be found in \cite{boneh2020graduate}) that the MAC construction above is one-time existentially unforgeable.
+
+\subsection{Related works}
 The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \texttt{OW-CPA} or \texttt{IND-CPA}) and outputs a $\pke$ with stronger security.
@@
 
-\section{Application to Kyber}
+\section{Application to Kyber}\label{sec:experimental-results}
 CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism that first constructs an IND-CPA secure public key encryption scheme whose security is based on the Module Learning with Error Problem (MLWE), then applies a generic transformation using \emph{de-randomization} and \emph{re-encryption} \cite{hofheinz2017modular}.
@@
         \State $z \leftsample \mathcal{B}^{32}$
-            \Comment{Randomly sample 32 bytes (256 bits)}
         \State $(\pk, \sk^\prime) \leftsample \texttt{Kyber.CPAPKE.KeyGen()}$
@@
         \State $m \leftsample \mathcal{B}^{32}$
-       % \State $m'= H(m)$
+        \State $m'= H(m)$
             \Comment{Do not output system RNG directly}
-        \State $(\bar{K}, r, K_{mac}) = G(m' \Vert H(\pk))$
+        \State $(\bar{K}, K_\mac) = G(m' \Vert H(\pk))$
             \Comment{G is instantiated with SHA3-512}
-        \State $c  \leftarrow \texttt{Kyber.CPAPKE.Enc(\pk, m', r)}$
-            \Comment{Because $r$ is set, \texttt{CPAPKE} is deterministic}
-            \State $t_1=MAC(K_{mac}, c)$
-        \State $K = \texttt{KDF}(\bar{K} \Vert t_1)$
-       % \State $t_2=MAC(K, t_1)$
+        \State $r \leftsample \mathcal{R}$
+        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m', r)$
+            \State $t=\mac(K_\mac, c^\prime)$
+        \State $K = \texttt{KDF}(\bar{K} \Vert t)$
             \Comment{\texttt{KDF} is instantiated with Shake256}
-        %\State $(c, t_1, t_2)$
-        \State $c \leftarrow (c, t_1)$
+        \State $c \leftarrow (c^\prime, t)$
         \State \Return $(c, K)$
@@
 \begin{algorithm}[H]
-    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c, t_1, t_2)$}\label{alg:kyber-ae-decap}
+    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c)$}\label{alg:kyber-ae-decap}
     \begin{algorithmic}[1]
         \Require Secret key $\sk = (\sk^\prime, \pk, H(\pk), z)$
-        \Require \texttt{Kyber.CPAPKE} Ciphertext $c$ and Tags $t_1, t_2$
+        \Require Ciphertext $c = (c^\prime, t)$
         \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
-            \Comment{Unpack the secret key; $h$ is the hash of $\pk$}
-        \State $(c, t_1) \leftarrow c$
-        \State $ m = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c)$ 
-       % \State $m'=H(m) $ 
-        \State $(\overline{K},   r, K_{mac}) =   G(m^\prime \Vert h)$
-        \State $t_1'=\mac(K_\mac, c)$
-        \If{$t_1' = t_1$}
-            \State $K  = \texttt{KDF}(\bar{K} \Vert t_1)$ 
-            %\State $t_2'= \mac(K, t_1)$
+        \State $(c^\prime, t_1) \leftarrow c$
+        \State $\hat{m} = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c^\prime)$ 
+        \State $(\overline{K}, K_\mac) =   G(m^\prime \Vert h)$
+        \State $\hat{t}=\mac(K_\mac, c)$
+        \If{$\hat{t} = t$}
+            \State $K  = \texttt{KDF}(\bar{K} \Vert t)$ 
+         \Else 
+            \State $K  = \texttt{KDF}(z \Vert t)$ 
+        \EndIf
             \State \Return $K$
-         \Else 
-            \State Abort
-        \EndIf
     \end{algorithmic}

### 2024-08-20 etm-transform/manuscript/main.tex
Major revision, improved preliminaries, experimental results, and future works

@@
 
-\subsection{Message authentication code (MAC)}
+\subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
 A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
@@
 
-We are also interested in one-time MAC, whose security goals are identical to normal MAC, except the signing oracle will only answer one signing query.
-In practice this means that the adversary only gets to see one message-tag pair, and the owner of the secret key only signs one message.
-One way to construct an unforgeable one-time MAC is to use a universal hash function (UHF), which computes the digest using finite field arithmetic.
-Each instance of the one-time MAC is parameterized by some maximal message length $L$ and the finite field $\mathbb{F}$ in which arithmetics are performed.
-The tag is computed by evaluating a polynomial whose coefficients are derived from the message:
+We are specifically interested in one-time MAC, whose security goal is identical to EUF-CMA described above, except for the constraint that each secret key can be used to sign exactly one distinct message.
+This translates to an attack model in which the signing oracle will only answer one signing query.
+Restricting to one-time usage allows for more efficient MAC constructions.
+One popular way to build one-time MAC is with universal hash functions (UHF), where each instance is parameterized by a finite field $\mathbb{F}$ and a maximal message length $L \geq 0$.
+The secret key is a pair of field elements $(k_1, k_2) \in \mathbb{F} \times \mathbb{F}$, and each message is a tuple of up to $L$ field elements $m = (m_1, m_2, \ldots, m_l) \in \mathbb{F}^{\leq L}$.
+To compute the tag:
 
 \begin{equation*}
-    \begin{aligned}
-    &H(k_1, (m_1, m_2, \ldots, m_l)) = k_1^{l+1} + k_1^{l}m_1 + k_1^{l-1}m_2 + \ldots + k_1m_l \\
-    &\mac((k_1, k_2), m) = H(k_1, m) + k_2
-    \end{aligned}
+    \mac((k_1, k_2), m) = H_\text{xpoly}(k_1, m) + k_2
 \end{equation*}
 
-We claim without proof (though proof can be found in \cite{boneh2020graduate}) that the MAC construction above is one-time existentially unforgeable.
+Where the $H_\text{xpoly}$ is a universal hash function:
+
+\begin{equation*}
+    H_\text{xpoly}(k_1, (m_1, m_2, \ldots, m_l)) = k_1^l \cdot m_1 + k_1^{l-1} \cdot m_2 + \ldots + k_1 \cdot m_l
+\end{equation*}
+
+\begin{lemma}
+    For all adversaries (including unbounded ones) against the MAC described above, the probability of winning the one-time EUF-CMA game is at most:
+
+    \begin{equation*}
+        \adv_\text{OT-EUF-CMA}(A) \leq \frac{L+1}{\norm{\mathbb{F}}}
+    \end{equation*}
+\end{lemma}
+\begin{proof}
+    See \cite{boneh2020graduate} lemma 7.11
+\end{proof}
 
 \subsection{Related works}
-The Fujisaki-Okamoto transformation (FOT) \cite{fujisaki1999secure} is a generic transformation that takes a $\pke$ with weaker security (such as \texttt{OW-CPA} or \texttt{IND-CPA}) and outputs a $\pke$ with stronger security.
-A later variation \cite{hofheinz2017modular} improved the original construction in \cite{fujisaki1999secure} by accounting for decryption failures, tightening security bounds, and providing a modular construction that first transforms \texttt{OW-CPA}/\texttt{IND-CPA} $\pke$ into \texttt{OW-PCVA} $\pke$ by providing ciphertext integrity through re-encryption (the \texttt{T} transformation), then transforming the \texttt{OW-PCVA} $\pke$ into an \texttt{IND-CCA} $\kem$ (the \texttt{U} transformation).
+The Fujisaki-Okamoto transformation \cite{fujisaki1999secure}\cite{hofheinz2017modular} is a family of generic transformations that takes as input a PKE with weaker security, such as OW-CPA, and outputs a PKE or KEM with IND-CCA2 security.
+The key ingredient in achieving ciphertext non-malleability is with \emph{de-randomization} and \emph{re-encryption}, which first transform a OW-CPA PKE into a \emph{rigid} PKE, then transform the rigid PKE into a KEM.
+More specifically: \begin{enumerate}
+    \item \emph{de-randomization} means that a randomized encryption routine $c \leftsample \encrypt(\pk, m)$ is made into a deterministic encryption routine by deriving randomization coin pseudorandomly: $c \leftarrow \encrypt(\pk, m, r = H(m))$ for some hash function $H$
+    \item \emph{re-encryption} means that the transformed decryption routine will run the transformed encryption routine to verify the integrity of the ciphertext.
+Because after \emph{de-randomization}, each plaintext strictly corresponds exacty one ciphertext, tempering with a ciphertext means that even if the ciphertext decrypts back to the same plaintext, the re-encryption will detect that the ciphertext has been tempered with.
+    \item \emph{rigidity} means that the decryption routine is a perfect inverse of the encryption routine: $c = \encrypt(\pk, m) \Leftrightarrow m = \decrypt(\sk, c)$.
+Converting a one-way secure rigid PKE (which is essentially a trapdoor function) into a IND-CCA2 KEM is well solved problem.
+We refer readers to \cite{boneh2020graduate} for details on such constructions.
+\end{enumerate}
 
-% TODO: Restating the algorithm seems redundant
+let $\pke = (\keygen_\pke, \encrypt, \decrypt)$ be defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$.
+Let $G: \mathcal{M} \rightarrow \mathcal{R}$ hash plaintexts into coincs, and let $H: \{0,1\}^\ast \rightarrow \{0,1\}^\ast$ hash byte stream into session keys.
+Depending on whether the constructed KEM uses implicit or explicit rejection, and the security property of the $\pke$, \cite{hofheinz2017modular} described four variations.
+They are summarized in table \ref{tbl:fo-variations} and figure \ref{fig:fo-routines}.
 
-Particularly relevant to our results are two variations of the \texttt{U} transformation: $\texttt{U}^\bot$ (KEM with explicit rejection) and $\texttt{U}^{\not\bot}$ (KEM with implicit rejection).
-If $\pke$ is \texttt{OW-PCVA} secure, then $\texttt{U}^\bot$ transforms $\pke$ into an \texttt{IND-CCA} secure $\kem^\bot$:
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|c|}
+        \hline
+        name & rejection & PKE security \\
+        \hline
+        $U^\bot$ & explicit & OW-PCVA \\
+        \hline
+        $U^{\not\bot}$ & implicit & OW-PCA \\
+        \hline
+        $U^\bot_m$ & explicit & OW-VA + rigid \\
+        \hline
+        $U^{\not\bot}_m$ & implicit & OW-CPA + rigid \\
+        \hline
+    \end{tabular}
+    \caption{Variants of modular FO transforms}\label{tbl:fo-variations}
+\end{table}
 
-\begin{theorem}\label{thm:u-bot}
-    For any \texttt{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^\bot$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \texttt{OW-PCVA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ that makes at most $q_H$ queries to $\pco$ and $\cvo$ such that
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.40\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\keygen()$}\label{alg:fo-keygen}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \keygen_\pke()$
+                \State $z \leftsample \mathcal{M}$
+                \State $\sk \leftarrow (\sk^\prime, z)$
+                    \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
+                \State $\sk \leftarrow \sk^\prime$
+                    \Comment{$U^{\bot}, U^{\bot}_m$}
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \hspace{0.05\textwidth}
+    \begin{minipage}{0.45\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\encap(\pk)$}\label{alg:fo-encap}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}$
+                \State $r \leftarrow G(m)$
+                \State $c \leftarrow \encrypt(\pk, m, r)$
+                \State $K \leftarrow H(m, c)$
+                    \Comment{$U^\bot, U^{\not\bot}$}
+                \State $K \leftarrow H(m)$
+                    \Comment{$U_m^\bot, U_m^{\not\bot}$}
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \vspace{0.5cm}
+    \begin{minipage}{\textwidth}
+        \begin{algorithm}[H]
+            \caption{$\decap(\sk=(\sk^\prime, z), c)$}\label{alg:fo-decap}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \decrypt(\sk^\prime, c)$
+                \State $\hat{r} \leftarrow G(\hat{m})$
+                \State $\hat{c} \leftarrow \encrypt(\pk, \hat{m}, \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State $K \leftarrow H(\hat{m})$
+                        \Comment{$U_m^\bot, U_m^{\not\bot}$}
+                    \State $K \leftarrow H(\hat{m}, c)$
+                        \Comment{$U^\bot, U^{\not\bot}$}
+                \Else
+                    \State $K \leftarrow H(z, c)$
+                        \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
+                    \State $K \leftarrow \bot$
+                        \Comment{$U^{\bot}, U^{\bot}_m$}
+                \EndIf 
+                \State \Return $K$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{Summary of the modular Fujisaki-Okamoto transformation variations}\label{fig:fo-routines}
+\end{figure}
 
-    \begin{equation*}
-        \epsilon_\kem \leq \epsilon_\pke
-    \end{equation*}
-\end{theorem}
-
-Similarly, if $\pke$ is \texttt{OW-PCA} secure, then $\texttt{U}^{\not\bot}$ transforms $\pke$ into an \texttt{IND-CCA} secure $\kem^{\not\bot}$
-
-\begin{theorem}\label{thm:u-notbot}
-    For any \texttt{IND-CCA} adversary $\mathcal{A}_\kem$ against $\kem^{\not\bot}$ with advantage $\epsilon_\kem$ issuing at most $q_D$ decapsulation queries and at most $q_H$ hash queries, there exists an \texttt{OW-CPA} adversary $\mathcal{A}_\pke$ against the underlying $\pke$ with advantage $\epsilon_\pke$ issuing at most $q_H$ queries to $\pco$ such that:
-
-    \begin{equation*}
-        \epsilon_\kem \leq \frac{q_H}{\norm{\mathcal{M}_\pke}} + \epsilon_\pke
-    \end{equation*}
-\end{theorem}
-
-The modularity of the \texttt{T} and \texttt{U} transformation allows us to tweak only the \texttt{T} transformation (see section \ref{sec:main-results}), obtain \texttt{OW-PCVA} security, then automatically get \texttt{IND-CCA} security for free.
-This means that we can directly apply our contribution to existing KEM's already using this modular transformation, such as ML-KEM \cite{key2023mechanism}, and obtain performance improvements while maintaining comparable levels of security (see section \ref{sec:mac-performance}).
+The modular FO transformations enjoy tight security bounds and proven quantum resistance.
+Variations have been deployed to many post-quantum KEMs submitted to NIST's post-quantum cryptography competition.
+Kyber, one of the round 3 finalists, uses the $U^{\not\bot}$ transformation.
+When it was later standardized into FIPS-203, it changed to use the $U^{\not\bot}_m$ transformation for computational efficiencies.
 
@@
 While this is not a standard security definition for MAC, this probability is straightforward to estimate for some classes of MACs.
-As will be discussed in section \ref{sec:mac-performance}, with a Carter-Wegman-like one-time MAC instantiated with message length $L$ and a finite field with $F$ elements, such probability is at most $\epsilon_\mac \leq \frac{L+1}{F}$
+As described in section \ref{sec:message-authentication-code}, with a Carter-Wegman-like one-time MAC instantiated with message length $L$ and a finite field with $F$ elements, such probability is at most $\epsilon_\mac \leq \frac{L+1}{F}$
 
@@
 
-%%%%%%%%%%%%%%%% uncomment to see %%%%%%%%%%%%%%%%%%%%
-% \include{not-pca-means-not-indcca}
+\section{Application to Kyber}\label{sec:experimental-results}
+CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism whose security is based on the conjecture hardness of the decisional Module Learning with Error problem.
+To achieve the the IND-CCA2 security, Kyber first constructs an IND-CPA secure public key encryption scheme based on \cite{lpr2013ideal}, then apply a slightly modified variation of the Fujisaki-Okamoto transformation described in \cite{hofheinz2017modular}.
+The resulting decapsulation routine is especially inefficient because Kyber's IND-CPA encryption routine incurs significantly more computational cost than the decryption routine.
+This makes Kyber a prime target for demonstrating the performance improvements enjoyed by the ``encrypt-then-MAC'' KEM construction.
 
-\section{Application to Kyber}\label{sec:experimental-results}
-CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism that first constructs an IND-CPA secure public key encryption scheme whose security is based on the Module Learning with Error Problem (MLWE), then applies a generic transformation using \emph{de-randomization} and \emph{re-encryption} \cite{hofheinz2017modular}.
-The use of \emph{re-encryption} for providing rigidity means that the decapsulation routine needs to run the encryption routine for verifying ciphertext integrity.
-Unfortunately for Kyber, the encryption routine consumes both more CPU cycles and carries a larger memory footprint.
-Therefore, applying an alternate transformation that replaces \emph{re-encryption} with a MAC verification will bring substantial performance enhancement.
-This is particularly applicable to communication protocols such as TLS 1.3, where clients, often constrained environments, need to run the decapsulation routine.
-
-The \texttt{IND-CPA} PKE of Kyber (algorithms 4, 5, 6 in \cite{avanzi2019crystals}) is not \texttt{OW-PCA} secure.
-A plaintext-checking attack \cite{ravi2019generic} can recover the secret key using a few maliciously constructed plaintext-checking queries for each coefficient of a Kyber secret key (there are 512, 768, and 1024 coefficients in the secret key depending on the desired security level).
-However, we propose mitigation for plaintext-checking attack at protocol level by requiring each keypair to be used for decryption only once.
-If a decryption fails, then the secret key should be discarded and the key exchange terminated, and a new keypair should be generated to restart the key exchange.
-In fact, such ephemeral key exchange is already required in TLS 1.3 for forward secrecy, so requiring the one-time use of keypair does not introduce additional operational cost for protocols such as TLS 1.3.
-
-The routines of \texttt{CCAKEM} can be found in Algorithm 7, 8, 9 in \cite{avanzi2019crystals}.
-We modify Algorithms 8 and 9 using authenticated encryption (AE)  mode as follows where Algorithm 7 is unchanged.
+We took the IND-CPA PKE routines (algorithms 4, 5, 6 in \cite{avanzi2019crystals}) and applied the ``encrypt-then-MAC'' transformation.
+The resulting KEM routines are described in algorithms \ref{alg:kyber-ae-keygen}, \ref{alg:kyber-ae-encap}, and \ref{alg:kyber-ae-decap}.
 
@@
     \begin{algorithmic}[1]
-        \State $z \leftsample \mathcal{B}^{32}$
+        \State $z \leftsample \{0,1\}^{256}$
         \State $(\pk, \sk^\prime) \leftsample \texttt{Kyber.CPAPKE.KeyGen()}$
@@
     \begin{algorithmic}[1]
-        \State $m \leftsample \mathcal{B}^{32}$
+        \State $m \leftsample \{0,1\}^{256}$
         \State $m'= H(m)$
@@
         \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
-        \State $(c^\prime, t_1) \leftarrow c$
+        \State $(c^\prime, t) \leftarrow c$
         \State $\hat{m} = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c^\prime)$ 
@@
 
-\begin{remark} If $c$ is manipulated, then the verification of  $t_1$  will be failed.
-In this case, there is no $K$ outputted from the decap$^+$.
-So the attacks described in the following subsections won't work.
-We also added the key confirmation, which is tag $t_2$.
+\begin{remark}
+    We derive the MAC key from both the plaintext $m$ and the public key $\pk$ for the same reason \cite{avanzi2019crystals} derives the pseudorandom coin from both $m$ and $\pk$.
+One is to allow both parties to participate in the encryption process, the other is to prevent a quantum adversary from pre-computing a large table of MAC keys that can then be brute-forced.
+\end{remark}
+\begin{remark}
+    We chose to derive the shared secret $K$ from the ciphertext digest $t$ instead of the ciphertext itself, which saves a few Keccak permutations since $t$ is only 128 bits while the full ciphertext could span more than a thousand bytes.
+This should not impact the security of the scheme since finding collision for an unknown MAC key constitutes a forgery attack on the MAC.
+\end{remark}
+\begin{remark}
+    We constructed the ``encrypt-then-MAC'' transformation to use implicit rejection so that we can directly use existing implementation of Kyber with minimal modification.
+In principle, a construction with explicit rejection should also be equally secure and efficient.
+\end{remark}
+\begin{remark}
+    We chose to use MAC with 256-bit key size and 128-bit tag size.
+When modeling the security threats, we assumed that the adversary may have access to quantum computers, making it necessary to use the maximal key size of common MACs while maintaining the minimum 128-bit security.
+On the other hand, the tag size can be relative small because the decapsulator (aka the decapsulation oracle) is assumed to be a classical computer, so there is no quantum speedup on brute-forcing a valid tag.
 \end{remark}
 
-{\color{blue} Note for authenticated encryption, tags $t_i$'s are necessary for the inputs.
-
-In fact, this is authenticated encryption instead of EtM. So we should change that, called authenticated encryption. 
-Please do not change my notation. They have their meanings in AE mode.}
-
-% \include{when-is-etm-useful}
-
-\subsection{MAC performance}\label{sec:mac-performance}
 When instantiating an instance $\kem(\keygen, \encap, \decap)$, there are a variety of possible MAC's to choose from.
-For each of the security level and choice of MAC, we measured the number of CPU cycles needed to produce a digest of the unauthenticated ciphertext.
-The median (top) and mean (bottom) measurements are reported in table \ref{tbl:standalone-mac-perf}.
+For each of the security level and choice of MAC, we measured the number of CPU cycles needed to produce a digest of Kyber's ciphertext.
+The MAC implementations are taken from OpenSSL and run on an Intel MacBook Pro with i7-9700h (TODO: run it on a server).
+The median measurements are reported in table \ref{tbl:standalone-mac-perf}.
 
@@
     \centering
-    \begin{tabular}{|c|c|c|c|c|c|}
+    \begin{tabular}{|c|c|c|c|c|}
         \hline
-        Name & Security & measurement & 768 bytes & 1088 bytes & 1568 bytes \\
+        Name & Security & 768 bytes & 1088 bytes & 1568 bytes \\
         \hline
-        CMAC & many-time & median & 5022 & 5442 & 6090 \\
-        & & mean & 5131 & 5578 & 6154 \\
+        CMAC & many-time & 5022 & 5442 & 6090 \\
         \hline
-        KMAC-256 & many-time & median & 7934 & 9862 & 11742 \\
-        & & mean & 8594 & 10693 & 12319 \\
+        KMAC-256 & many-time & 7934 & 9862 & 11742 \\
         \hline
-        GMAC & one-time & median & 2778 & 2756 & 2762 \\
-        & & mean & 2843 & 2780 & 2919 \\
+        GMAC & one-time & 2778 & 2756 & 2762 \\
         \hline
-        Poly1305 & one-time & median & 1128 & 1218 & 1338 \\
-        & & mean & 1435 & 1504 & 1625 \\
+        Poly1305 & one-time & 1128 & 1218 & 1338 \\
         \hline
@@
 
-Based on the security reduction in section \ref{sec:main-results} we chose standalone GMAC and Poly1305 for their substantial performance advantage over other constructions such as CBC-MAC and KMAC.
-We then modified the reference implementation (TODO: citation needed https://github.com/pq-crystals/kyber) according to algorithms \ref{alg:kyber-ae-keygen}, \ref{alg:kyber-ae-encap}, and \ref{alg:kyber-ae-decap}.
-We measured the median and mean CPU cycles needed to run each of the routines.
-The measurements are listed in table \ref{tbl:kyber-ae-perf}
+Based on the security reduction in section \ref{sec:main-results} we chose standalone GMAC and Poly1305 for their substantial performance advantage over other constructions.
+We modified the reference implementation (https://github.com/pq-crystals/kyber) to implement the proposed construction, then measured the CPU cycles needed to run each of routines.
+The median CPU cycles are reported in table \ref{tbl:kyber-ae-perf}.
 
@@
     \centering
-    \begin{tabular}{|c|c|c|c|c|c|}
+    \begin{tabular}{|c|c|c|c|c|}
         \hline
-        Name & Security level & measurement & $\keygen$ & $\encap$ & $\decap$ \\
+        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
         \hline
-        Kyber512 & 128 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber512 & 128 bits &  & &  \\
         \hline
-        Kyber768 & 192 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber768 & 192 bits &  & &  \\
         \hline
-        Kyber1024 & 256 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber1024 & 256 bits &  & &  \\
         \hline
-        KyberAE512 w/ GMAC& 128 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber512 + GMAC& 128 bits &  & &  \\
         \hline
-        KyberAE768 w/ GMAC& 192 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber768 + GMAC& 192 bits &  & &  \\
         \hline
-        KyberAE1024 w/ GMAC& 256 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber1024 + GMAC& 256 bits &  & &  \\
         \hline
-        KyberAE512 w/ Poly1305& 128 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber512 + Poly1305& 128 bits &  & &  \\
         \hline
-        KyberAE768 w/ Poly1305& 192 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber768 + Poly1305& 192 bits &  & &  \\
         \hline
-        KyberAE1024 w/ Poly1305& 256 bits & median &  & &  \\
-        & & mean &  & &  \\
+        Kyber1024 + Poly1305& 256 bits &  & &  \\
         \hline
     \end{tabular}
-    \caption{Kyber-AE performance measurements}\label{tbl:kyber-ae-perf}
+    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
 \end{table}
+
+Compared to Kyber using the FO transform, our proposed construction adds a small amount of runtime overhead (for computing a digest) in the encapsulation routine and a small increase in ciphertext size (for the 128-bit tag).
+In exchange, we see significant performance runtime savings in the decapsulation routines.
+This trade-off is especially meaningful in a key exchange protocol where one party has substantially more computationl resource than the other.
+For example, in many experimental implementation of TLS with post-quantum KEMs (such as \href{https://blog.cloudflare.com/the-tls-post-quantum-experiment/}{CECPQ2}), the client (might be IoT devices) runs $\keygen$ and $\decap$ while the server (usually data centers) runs $\encap$.
 
@@
 
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox[300px]{
+            \begin{algorithm}[H]
+                \caption{$\pco(m, c)$}\label{alg:pco-using-decap-oracle}
+                \begin{algorithmic}[1]
+                    \State $k \leftarrow G(m)$
+                    \State $t \leftarrow \mac(k, c)$
+                    \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
+                \end{algorithmic}
+            \end{algorithm}
+        }{
+            \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}
+            \label{fig:pco-using-decap-oracle}
+        }
+    \end{floatrow}
+\end{figure}
+
+We applied the ``encrypt-then-MAC'' transformation to Kyber, which we speculated to be safe to use in ephemeral key exchange where each decapsulation key can only be used to decapsulate one ciphertext, and saw meaningful performance improvements over using the Fujisaki-Okamoto transformation.
+Unfortunately, the resulting KEM does not achieve the desired full IND-CCA2 security.
+This is because every chosen-ciphertext attack agaisnt the ```encrypt-then-MAC''' KEM can be converted into a plaintext-checking attack against the the underlying PKE (see algorithm \ref{alg:pco-using-decap-oracle}), and the IND-CPA subroutines of Kyber are known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}.
+This observation, combined with theorem \ref{thm:ow-pca-implies-kem-ind-cca2}, results in the following security equivalence:
+
+\begin{lemma}
+    $\kem_\etm$ is IND-CCA2 secure if and only if the input $\pke$ is OW-PCA secure
+\end{lemma}
+
+However, if the input $\pke$ is OW-CPA secure and rigid (which implies OW-PCA security), then the $U^\bot_m, U^{\not\bot}_m$ transformations from \cite{hofheinz2017modular} can build an IND-CCA2 secure KEM with almost no overhead, as the encapsulation and decapsulation routine each only adds a hash of the plaintext on top of the input encryption and decryption routine.
+This leaves a gap between OW-PCA security and \emph{OW-CPA + rigidity} in which the ``encrypt-then-MAC'' transformation is a superior strategy for building KEM from a PKE.
+It remains an open problem to find a suitable PKE that is OW-PCA but not rigid.
+
 %%%% 8. BILBIOGRAPHY %%%%
-% TODO: figure out how to use bibliography
 \bibliographystyle{alpha}
 \bibliography{abbrev3,crypto,biblio}
-%%%% NOTES
 % - Download abbrev3.bib and crypto.bib from https://cryptobib.di.ens.fr/

### 2024-08-21 etm-transform/manuscript/main.tex
Reworked section 3; section 5 needs more work

@@
 \section{The ``encrypt-then-MAC'' transformation}\label{sec:main-results}
-Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme.
-Let $\mac$ be a deterministic message authentication code.
-Let $G: \mathcal{M}_\pke \rightarrow \mathcal{K}_\mac$ and $H: \{0,1\}^\ast \rightarrow \mathcal{K}_\kem$ be hash functions, where $\mathcal{K}_\kem$ denote the set of all possible session keys.
-The $\etm$ transformation outputs a key encapsulation mechanism $\kem_\etm(\keygen_\etm, \encap_\etm, \decap_\etm)$.
-The three routines are described in figure \ref{fig:etm-routines}.
+Let $\mathcal{B}^\ast$ denote the set of finite bit strings.
+Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$.
+Let $\mac: \mathcal{K}_\mac \times \mathcal{B}^\ast \rightarrow \mathcal{T}$ be a deterministic message authentication code that takes a key $k \in \mathcal{K}_\mac$, some message $m \in \mathcal{B}^\ast$, and outputs a digest $t \in \mathcal{T}$.
+Let $G: \mathcal{M} \rightarrow \mathcal{K}_\mac$ be a hash function that maps from $\pke$'s plaintext space to $\mac$'s key space.
+Let $H: \mathcal{B}^\ast \rightarrow \mathcal{K}_\kem$ be a hash function that maps bit strings into the set of possible shared secrets.
+The ``encrypt-then-MAC'' transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen_\kem, \encap, \decap)$, whose routines are described in figure \ref{fig:etm-routines}.
 
@@
         \begin{algorithmic}[1]
-            \State $(\pk, \sk_\pke) \leftsample \keygen(1^\lambda)$
-            \State $z \leftsample \mathcal{M}_\pke$
-            \State $\sk \leftarrow (\sk_\pke, z)$
+            \State $(\pk, \sk^\prime) \leftsample \keygen(1^\lambda)$
+            \State $z \leftsample \mathcal{M}$
+            \State $\sk \leftarrow (\sk^\prime, z)$
             \State \Return $(\pk, \sk)$
@@
     \begin{algorithm}[H]
-        \caption{$\encap_\etm(\pk)$}\label{alg:etm-encap}
+        \caption{$\encap(\pk)$}\label{alg:etm-encap}
         \begin{algorithmic}[1]
-            \State $m \leftsample \mathcal{M}_\pke$
+            \State $m \leftsample \mathcal{M}$
             \State $k \leftarrow G(m)$
-            \State $c_\pke \leftsample \encrypt(\pk, m)$
-            \State $t \leftarrow \mac(k, c_\pke)$
-            \State $K \leftarrow H(m, c_\pke)$
-            \State $c \leftarrow (c_\pke, K)$
+            \State $c^\prime \leftsample \encrypt(\pk, m)$
+            \State $t \leftarrow \mac(k, c^\prime)$
+            \State $K \leftarrow H(m, c^\prime)$
+            \State $c \leftarrow (c^\prime, t)$
             \State \Return $(c, K)$
@@
     \begin{algorithm}[H]
-        \caption{$\decap_\etm(\sk, c)$}\label{alg:etm-decap}
+        \caption{$\decap(\sk, c)$}\label{alg:etm-decap}
         \begin{algorithmic}[1]
-            \State $(c_\pke, t) \leftarrow c$
-            \State $(\sk_\pke, z) \leftarrow \sk$
-            \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
+            \State $(c^\prime, t) \leftarrow c$
+            \State $(\sk^\prime, z) \leftarrow \sk$
+            \State $\hat{m} \leftarrow \decrypt(\sk^\prime, c^\prime)$
             \State $\hat{k} \leftarrow G(\hat{m})$
-            \If{$\mac(\hat{k}, c_\pke) \neq t$}
-                \State \Return $H(z, c_\pke)$
+            \If{$\mac(\hat{k}, c^\prime) \neq t$}
+                \State $K \leftarrow H(z, c^\prime)$
+            \Else
+                \State $K \leftarrow H(\hat{m}, c^\prime)$
             \EndIf
-            \State \Return $H(\hat{m}, c_\pke)$
+            \State \Return $K$
         \end{algorithmic}
@@
 
+The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol.
+In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext $k \leftarrow G(m)$, then used to sign the unauthenticated ciphertext $c^\prime$.
+Because the encryption routine might be randomized, the session key is derived from both the message and the ciphertext.
+Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$, which is transmitted to the peer.
+In the decapsulation routine, the decryption $\hat{m}$ of the unauthenticated ciphertext is used to re-derive the MAC key $\hat{k}$, which is then used to re-compute the tag $\hat{t}$.
+The ciphertext is considered valid if and only if the recomputed tag is identical to the input tag.
+
+For an adversary $A$ to produce a valid tag $t$ for some unauthenticated ciphertext $c^\prime$ under the symmetric key $k \leftarrow G(\decrypt(\sk^\prime, c^\prime))$ implies that $A$ must either know the symmetric key $k$ or produce a forgery.
+Under the random oracle model, $A$ also cannot know $k$ without also knowing $\decrypt(\sk^\prime, c^\prime)$, so $A$ must either have produced $c^\prime$ honestly, or have broken the one-way security of $\pke$.
+This means that the decapsulation oracle will not give out information on decryptions that the adversary does not already know.
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox[300px]{
+            \begin{algorithm}[H]
+                \caption{$\pco(m, c)$}\label{alg:pco-using-decap-oracle}
+                \begin{algorithmic}[1]
+                    \State $k \leftarrow G(m)$
+                    \State $t \leftarrow \mac(k, c)$
+                    \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
+                \end{algorithmic}
+            \end{algorithm}
+        }{
+            \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}
+            \label{fig:pco-using-decap-oracle}
+        }
+    \end{floatrow}
+\end{figure}
+
+However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed.
+This means that a decapsulation oracle can be converted into a plaintext-checking oracle, so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
+
+On the other hand, if the underlying PKE is one-way secure against plaintext-checking attack that makes $q$ plaintext-checking queries, then ``encrypt-then-MAC'' KEM is semantically secure under chosen ciphertext attacks making the same number of decapsulation queries:
+
 \begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
-    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes $q_D$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes at least $q_D$ plaintext-checking queries against the underlying $\pke$ such that
+    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes $q$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes at least $q$ plaintext-checking queries against the underlying $\pke$, and an one-time existential forgery adversary $C$ against the underlying $\mac$ such that
 
     \begin{equation*}
-        \texttt{Adv}_\texttt{IND-CCA2}(A) \leq q_D \cdot \epsilon_\mac + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
+        \texttt{Adv}_\texttt{IND-CCA2}(A) \leq q \cdot \adv_\texttt{OT-MAC}(C) + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
     \end{equation*}
@@
 
+Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} naturally flows into an equivalence relationship between the security of the KEM and the security of the PKE:
+
+\begin{lemma}
+    $\kem_\etm$ is IND-CCA2 secure if and only if the input $\pke$ is OW-PCA secure
+\end{lemma}
+
+\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
 \begin{proof}
-    We will prove using a sequence of games. The complete sequence of games is shown in figure \ref{fig:sequence-of-games}
+    We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of games.
 
     \begin{figure}[H]
-        \centering
+        \begin{floatrow}
+            \ffigbox{
                 \begin{algorithm}[H]
-            \caption{Sequence of games $G_0 - G_3$}\label{alg:sequence-of-games}
+                    \caption{IND-CCA2 game for KEM}\label{alg:game-0}
                     \begin{algorithmic}[1]
-                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                \State $(m^\ast, z) \leftsample \mathcal{M}_\pke$
+                        \State $(\pk, \sk) \leftsample \keygen_\etm()$
+                        \State $m^\ast \leftsample \mathcal{M}$
+                        \State $c^\prime \leftsample \encrypt(\pk, m^\ast)$
                         \State $k^\ast \leftarrow G(m^\ast)$
-                    \Comment{$G_0$-$G_1$}
-                \State $k^\ast \leftsample \mathcal{K}_\mac$
-                    \Comment{$G_2$-$G_3$}
-                \State $c^\ast_\pke \leftsample \encrypt(\pk, m^\ast)$
-                \State $t^\ast \leftarrow \mac(k^\ast, c^\ast_\pke)$
-                \State $c^\ast \leftarrow (c^\ast_\pke, t^\ast)$
-                \State $K_0 \leftarrow H(m^\ast, c^\ast_\pke)$
-                    \Comment{$G_0$-$G_2$}
-                \State $K_0 \leftsample \mathcal{K}_\kem$
-                    \Comment{$G_3$}
+                        \State $t \leftarrow \mac(k, c^\prime)$
+                        \State $c^\ast \leftarrow (c^\prime, t)$
+                        \State $K_0 \leftarrow H(m^\ast, c^\prime)$
                         \State $K_1 \leftsample \mathcal{K}_\kem$
                         \State $b \leftsample \{0,1\}$
-                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(
-                    1^\lambda, \pk, c^\ast, K_b)$
-                    \Comment{$G_0$}
-                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(
-                    1^\lambda, \pk, c^\ast, K_b)$
-                    \Comment{$G_1$-$G_3$}
+                        \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(\pk, c^\ast, K_b)$
                         \State \Return $\llbrack \hat{b} = b \rrbrack$
@@
                 \end{algorithm}
+            }{}
+            \ffigbox{
                 \begin{algorithm}[H]
-            \caption{$\mathcal{O}^\decap(c)$}\label{alg:etm-decap-oracle}
+                    \caption{$\mathcal{O}^\decap(c)$}\label{alg:true-decap-oracle}
                     \begin{algorithmic}[1]
-                \State $(c_\pke, t) \leftarrow c$
-                \State $\hat{m} \leftarrow \decrypt(\sk_\pke, c_\pke)$
+                        \State $(c^\prime, t) \leftarrow c$
+                        \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
                         \State $\hat{k} \leftarrow G(\hat{m})$
-                \If{$\mac(\hat{k}, c_\pke) = t$}
-                    \State \Return $H(\hat{m}, c_\pke)$
+                        \If{$\mac(\hat{k}, c^\prime) = t$}
+                            \State $K \leftarrow H(\hat{m}, c^\prime)$
+                        \Else 
+                            \State $K \leftarrow H(z, c^\prime)$
                         \EndIf 
-                \State \Return $H(z, c_\pke)$
+                        \State \Return $K$
                     \end{algorithmic}
                 \end{algorithm}
+            }{}
+        \end{floatrow}
+        \begin{floatrow}
+            \ffigbox{
                 \begin{algorithm}[H]
-            \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:etm-decap-oracle-1}
+                    \caption{$\mathcal{O}^G(m)$}\label{alg:hash-oracle-g}
                     \begin{algorithmic}[1]
-                \State $(c_\pke, t) \leftarrow c$
-                \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G 
-                    : \decrypt(\sk_\pke, c_\pke) = \tilde{m} 
-                    \land \mac(\tilde{k}, c_\pke) = t$}
-                    \State \Return $H(\tilde{m}, c_\pke)$
+                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
+                            \State \Return $\tilde{k}$
                         \EndIf
-                \State \Return $H(z, c_\pke)$
+                        \State $k \leftsample \mathcal{K}_\mac$
+                        \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
+                        \State \Return $k$
                     \end{algorithmic}
                 \end{algorithm}
-        \caption{Sequence of games, true decap oracle $\mathcal{O}^\decap$ and simulated oracle $\mathcal{O}^\decap_1$}\label{fig:sequence-of-games}
+            }{}
+            \ffigbox{
+                \begin{algorithm}[H]
+                    \caption{$\mathcal{O}^H(m, c)$}\label{alg:hash-oracle-h}
+                    \begin{algorithmic}[1]
+                        \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
+                            \State \Return $\tilde{K}$
+                        \EndIf
+                        \State $K \leftsample \mathcal{K}_\kem$
+                        \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
+                        \State \Return $K$
+                    \end{algorithmic}
+                \end{algorithm}
+            }{}
+        \end{floatrow}
     \end{figure}
 
-    \emph{Game 0} is the standard \texttt{IND-CCA2} game for a key encapsulation mechanism.
+    \emph{Game 0} is the standard IND-CCA2 game for KEMs.
+The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
+The queries made to the hash oracles $\mathcal{O}^G, \mathcal{O}^H$ are recorded to their respective tapes $\mathcal{L}^G, \mathcal{L}^H$.
 
-    \emph{Game 1} is identical to \emph{Game 0} except for that the decapsulation oracle $\mathcal{O}^\decap$ (algorithm \ref{alg:etm-decap-oracle}) is replaced with a simulated decapsulation oracle $\mathcal{O}^\decap_1$ (algorithm \ref{alg:etm-decap-oracle-1}).
-If $\mathcal{O}^\decap_1$ accepts the queried ciphertext $c = (c_\pke, t)$ and outputs the true session key $K \leftarrow H(\tilde{m}, c_\pke)$, then the queried ciphertext must be honestly generated, which means that $\mathcal{O}^\decap$ must also accept the queried ciphertext and output the true session key.
-If $\mathcal{O}^\decap$ rejects the queried ciphertext $c = (c_\pke, t)$ and outputs the implicit rejection $K \leftarrow H(z, c_\pke)$, then the tag $t$ is invalid under the MAC key $k \leftarrow G(\decrypt(\sk_\pke, c_\pke))$.
-Since for a given ciphertext $c_\pke$, the correct MAC key is fixed, there could not be a matching hash query $(m, k)$ such that $m$ is the correct decryption and $k$ can validate the incorrect tag.
-Therefore, $\mathcal{O}^\decap_1$ must also reject the queried ciphertext and output the implicit rejection.
+    \begin{figure}[H]
+        \begin{floatrow}
+            \ffigbox[270px]{
+                \begin{algorithm}[H]
+                    \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:sim-decap-oracle}
+                    \begin{algorithmic}[1]
+                        \State $(c^\prime, t) \leftarrow c$
+                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
+                            \tilde{m} = \decrypt(\sk^\prime, c^\prime)
+                            \land \mac(\tilde{k}, c^\prime) = t
+                        $}
+                            \State $K \leftarrow H(\tilde{m}, c^\prime)$
+                        \Else
+                            \State $K \leftarrow H(z, c^\prime)$
+                        \EndIf
+                        \State \Return $K$
+                    \end{algorithmic}
+                \end{algorithm}
+            }{
+                \caption{Simulated decapsulation oracle}
+            }
+        \end{floatrow}
+    \end{figure}
 
-    This means that game 0 and game 1 differ when $\mathcal{O}^\decap$ accepts the queried ciphertext $c = (c_\pke, t)$ but $\mathcal{O}^\decap_1$ rejects it, which means that $t$ is a valid tag for $c_\pke$ under the correct MAC key $k \leftarrow G(\decrypt(\sk_\pke, c_\pke))$ but such key is never queried by the adversary.
-Under the random oracle model, from the adversary's perspective, such $k$ is an unknown and uniformly random key, so producing a valid tag under such key constitutes a forgery against the MAC.
-Denote the probability of forgery against unknown uniformly random MAC key by $\epsilon_\mac$, then the probability that the two decapsulation oracles disagree on one or more queries is at most $q_D \cdot \epsilon_\mac$.
-Finally, by the difference lemma,
+    \emph{Game 1} is identical to game 0 except that the true decapsulation oracle $\mathcal{O}^\decap$ is replaced with a simulated oracle $\mathcal{O}^\decap_1$.
+Instead of directly decrypting $c^\prime$ as in the decapsulation routine, the simulated oracle searches through the tape $\mathcal{L}^G$ to find a matching query $(\tilde{m}, \tilde{k})$ such that $\tilde{m}$ is the decryption of $c^\prime$.
+The simulated oracle then uses $\tilde{k}$ to validate the tag $t$ against $c^\prime$.
+
+    If the simulated oracle accepts the queried ciphertext as valid, then there is a matching query that also validates the tag, which means that the queried ciphertext is honestly generated.
+Therefore, the true oracle must also accept the queried ciphertext.
+On the other hand, if the true oracle rejects the queried ciphertext (and output the implicit rejection $H(z, c^\prime)$), then the tag is simply invalid under the MAC key $k = G(\decrypt(\sk^\prime, c^\prime))$.
+Therefore, there could not have been a matching query that also validates the tag, and the simulated oracle must also rejects the queried ciphertext.
+
+    This means that from the adversary $A$'s perspective, game 1 and game 0 differ only when the true oracle accepts while the simulated oracle rejects, which means that $t$ is a valid tag for $c^\prime$ under $k = G(\decrypt(\sk^\prime, c^\prime))$, but $k$ has never been queried.
+Under the random oracle model, such $k$ is a uniformly random sample of $\mathcal{K}_\mac$ that the adversary does not know, so for $A$ to produce a valid tag is to produce a forgery against the $\mac$ under an unknown and uniformly random key.
+Furthermore, the security game does not include a signing oracle, so this is a zero-time forgery.
+While zero-time forgery is not a standard security definition for a MAC, we can bound it by the advantage of a one-time forgery adversary $C$:
 
     \begin{equation*}
-        \texttt{Adv}_0(A) - \texttt{Adv}_1(A) \leq q_D \cdot \epsilon_\mac
+        P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]
+        \leq \adv_\texttt{OT-MAC}(C)
     \end{equation*}
 
-    Note that $\epsilon_\mac$ quantifies the probability that an adversary can produce forgery for a unknown key without access to a signing oracles.
-While this is not a standard security definition for MAC, this probability is straightforward to estimate for some classes of MACs.
-As described in section \ref{sec:message-authentication-code}, with a Carter-Wegman-like one-time MAC instantiated with message length $L$ and a finite field with $F$ elements, such probability is at most $\epsilon_\mac \leq \frac{L+1}{F}$
-
-    \emph{Game 2} is identical to \emph{Game 1}, except for that when the challenger generates the challenge ciphertext $c^\ast = (c^\ast_\pke, t^\ast)$, the tag $t^\ast$ is computed using a uniformly random key $k^\ast \leftarrow \mathcal{K}_\mac$ instead of a pseudorandom key derived from hashing the challenge plaintext.
-
-    Under the random oracle model, game 2 and game 1 are statistically identical to the adversary $A$, unless $A$ queries $G$ with $m^\ast$.
-Denote the probability that $A$ queries $G$ with $m^\ast$ by $P[\texttt{QUERY G}^\ast]$, then:
+    Across all $q$ decapsulation queries, the probability that at least one query is a forgery is thus at most $q \cdot \adv_\texttt{OT-MAC}(C)$. By the difference lemma:
 
     \begin{equation*}
-        \texttt{Adv}_1(A) - \texttt{Adv}_2(A) \leq P[\texttt{QUERY G}^\ast]
+        \adv_{G_0}(A) - \adv_{G_1}(A) \leq q\cdot  \adv_\texttt{OT-MAC}(C)
     \end{equation*}
 
-    \emph{Game 3} is identical to \emph{Game 2}, except for that $K_0$ is a uniformly random session key instead of a pseudorandom session key derived from the challenge plaintext-ciphertext pair.
-Under the random oracle model, game 3 and game 2 are statistically identical unless the adversary $A$ queries $H$ with $(m^\ast, \cdot)$.
-Denote the probability that $A$ makes such $H$ query by $P[\texttt{QUERY H}^\ast]$, then:
+    \emph{Game 2} is identical to game 1, except that on line 4 of algorithm \ref{alg:game-0}, the challenger samples a uniformly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$ instead of deriving it from $m$.
+From $A$'s perspective the two games are indistinguishable, unless $A$ queries $G$ with the value of $m^\ast$.
+Denote the probability that $A$ queries $G$ with $m^\ast$ by $P[\texttt{QUERY G}]$, then:
 
     \begin{equation*}
-        \texttt{Adv}_2(A) - \texttt{Adv}_3(A) \leq P[\texttt{QUERY H}^\ast]
+        \adv_{G_1}(A) - \adv_{G_2}(A) \leq P\left[\texttt{QUERY G}\right]
     \end{equation*}
 
-    In game 3, both $K_0$ and $K_1$ are uniformly random. There is no statistical difference between the two session keys, so no adversary can have any advantage: $\texttt{Adv}_3(A) = 0$.
+    \emph{Game 3} is identical to game 2, except that on line 7 of algorithm \ref{alg:game-0}, the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $c^\prime$.
+From $A$'s perspective the two games are indistinguishable, unless $A$ queries $H$ with $(m^\ast, \cdot)$.
+Denote the probability that $A$ queries $H$ with $(m^\ast, \cdot)$ by $P[\texttt{QUERY H}]$, then:
 
-    Now consider an \texttt{OW-PCA} adversary $B$ simulating game 3 for $A$: \begin{enumerate}
-        \item When $B$ receives its public key $\pk$, $B$ passes $\pk$ to $A$
-        \item $B$ can sample the implicit rejection $z$ by itself
-        \item $B$ can simulate both hash oracles $G$ and $H$ for $A$
-        \item $B$ can simulate $\mathcal{O}^\decap_1$ for $A$.
-Instead of checking if $\decrypt(\sk_\pke, c) = \tilde{m}$, $B$ can use its access to the plaintext-checking oracle and check if $\pco(\tilde{m}, c) = 1$.
-This means that for every decapsulation query $B$ services, $B$ needs to make at least one plaintext-checking query.
-Therefore, $B$ needs to make at least $q_D$ plaintext-checking query.
-        \item When $B$ receives its challenge ciphertext $c_\pke^\ast$, it can sample a uniformly random key $k^\ast \leftsample \mathcal{K}_\mac$, produce the corresponding tag $t^\ast \leftarrow \mac(k^\ast, c^\ast_\pke)$, and sample a uniformly random session keys $K_0, K_1 \leftsample \mathcal{K}_\kem$.
-$B$ then passes $c^\ast = (c^\ast_\pke, t^\ast)$ as the challenge ciphertext and a coin-flip $K_b$ as the challenge session key.
-    \end{enumerate}
+    \begin{equation*}
+        \adv_{G_2}(A) - \adv_{G_3}(A) \leq P\left[\texttt{QUERY H}\right]
+    \end{equation*}
 
-    If $A$ ever queries $G$ or $H$ with the decryption of $c^\ast_\pke$, $B$ will be able to detect it using the plaintext-checking oracle.
-From where $B$ is guaranteed to win the \texttt{OW-PCA} game.
-Therefore:
+    Since in game 3, both $K_0$ and $K_1$ are uniformly random and independent of all other variables, no adversary can have any advantage: $\adv_{G_3}(A) = 0$.
+
+    \begin{figure}[H]
+        \begin{floatrow}
+            \ffigbox{
+                \begin{algorithm}[H]
+                    \caption{$B(\pk, {c^\prime}^\ast)$}\label{alg:ow-pca-adversary}
+                    \begin{algorithmic}[1]
+                        \State $k \leftsample \mathcal{K}_\mac$
+                        \State $t \leftarrow \mac(k, {c^\prime}^\ast)$
+                        \State $c^\ast \leftarrow ({c^\prime}^\ast, t)$
+                        \State $K \leftsample \mathcal{K}_\kem$
+                        \State $\hat{b} \leftarrow A^{
+                            \mathcal{O}^\decap_B, \mathcal{O}^G_B, \mathcal{O}^H_B
+                        }(\pk, c^\ast, K)$
+                        \If{$\texttt{ABORT}(m)$}
+                            \State \Return $m$
+                        \EndIf
+                    \end{algorithmic}
+                \end{algorithm}
+            }{}
+            \ffigbox{
+                \begin{algorithm}[H]
+                    \caption{$\mathcal{O}^\decap_B(c)$}\label{alg:sim-decap-oracle-b}
+                    \begin{algorithmic}[1]
+                        \State $(c^\prime, t) \leftarrow c$
+                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
+                            % \tilde{m} = \decrypt(\sk^\prime, c^\prime)
+                            \pco(c^\prime, \tilde{m}) = 1
+                            \land \mac(\tilde{k}, c^\prime) = t
+                        $}
+                            \State $K \leftarrow H(\tilde{m}, c^\prime)$
+                        \Else
+                            \State $K \leftarrow H(z, c^\prime)$
+                        \EndIf
+                        \State \Return $K$
+                    \end{algorithmic}
+                \end{algorithm}
+            }{}
+        \end{floatrow}
+        \begin{floatrow}
+            \ffigbox{
+                \begin{algorithm}[H]
+                    \caption{$\mathcal{O}^G_B(m)$}\label{alg:hash-oracle-g-b}
+                    \begin{algorithmic}[1]
+                        \If{$\pco(m, {c^\prime}^\ast) = 1$}
+                            \State $\texttt{ABORT}(m)$
+                        \EndIf
+                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
+                            \State \Return $\tilde{k}$
+                        \EndIf
+                        \State $k \leftsample \mathcal{K}_\mac$
+                        \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
+                        \State \Return $k$
+                    \end{algorithmic}
+                \end{algorithm}
+            }{}
+            \ffigbox{
+                \begin{algorithm}[H]
+                    \caption{$\mathcal{O}^H_B(m, c)$}\label{alg:hash-oracle-h-b}
+                    \begin{algorithmic}
+                        \If{$\pco(m, {c^\prime}^\ast) = 1$}
+                            \State $\texttt{ABORT}(m)$
+                        \EndIf
+                        \If{$
+                            \exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H 
+                            : \tilde{m} = m \land \tilde{c} = c
+                        $}
+                            \State \Return $\tilde{K}$
+                        \EndIf
+                        \State $K \leftsample \mathcal{K}_\kem$
+                        \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
+                        \State \Return $K$
+                    \end{algorithmic}
+                \end{algorithm}
+            }{}
+        \end{floatrow}
+    \end{figure}
+    We will bound $P[\texttt{QUERY G}]$ and $P[\texttt{QUERY H}]$ by constructing a OW-PCA adversary $B$ against the underlying PKE that uses $A$ as a sub-routine.
+$B$'s behaviors are described in algorithms \ref{alg:ow-pca-adversary}, \ref{alg:sim-decap-oracle-b}, \ref{alg:hash-oracle-g-b}, and \ref{alg:hash-oracle-h-b}.
+
+    $B$ simulates game 3 for $A$: receiving the public key $\pk$ and challenge encryption ${c^\prime}^\ast$, $B$ samples random MAC key and session key to produce the challenge encapsulation, then feeds it to $A$.
+When simulating the decapsulation oracle, $B$ uses the plaintext-checking oracle to look for matching queries in $\mathcal{L}^G$.
+When simulaitng the hash oracles, $B$ uses the plaintext-checking oracle to detect when $m^\ast = \decrypt(\sk^\prime, {c^\prime}^\star)$ has been queried.
+When $m^\ast$ is queried, $B$ terminates $A$ and returns $m^\ast$ to win the OW-PCA game.
+In other words:
 
@@
         \begin{aligned}
-            P[\texttt{QUERY G}^\ast] &\leq \texttt{Adv}_\texttt{OW-PCA}(B) \\
-            P[\texttt{QUERY H}^\ast] &\leq \texttt{Adv}_\texttt{OW-PCA}(B)
+            P\left[\texttt{QUERY G}\right] &\leq \adv_\texttt{OW-PCA}(B) \\
+            P\left[\texttt{QUERY H}\right] &\leq \adv_\texttt{OW-PCA}(B) \\
         \end{aligned}
@@
 
-    Combining all inequalities above, we have:
-
-    \begin{equation*}
-        \texttt{Adv}_0(A) \leq q_D \cdot \epsilon_\mac + 2 \texttt{Adv}_\texttt{OW-PCA}(B)
-    \end{equation*}
+    Combining all equations above produce the desired security bound.
 \end{proof}
@@
 \section{Conclusions and future works}\label{sec:future-works}
-
-\begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox[300px]{
-            \begin{algorithm}[H]
-                \caption{$\pco(m, c)$}\label{alg:pco-using-decap-oracle}
-                \begin{algorithmic}[1]
-                    \State $k \leftarrow G(m)$
-                    \State $t \leftarrow \mac(k, c)$
-                    \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
-                \end{algorithmic}
-            \end{algorithm}
-        }{
-            \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}
-            \label{fig:pco-using-decap-oracle}
-        }
-    \end{floatrow}
-\end{figure}
+% TODO: need to rework this
 
@@
 This observation, combined with theorem \ref{thm:ow-pca-implies-kem-ind-cca2}, results in the following security equivalence:
-
-\begin{lemma}
-    $\kem_\etm$ is IND-CCA2 secure if and only if the input $\pke$ is OW-PCA secure
-\end{lemma}
 

### 2024-08-22 etm-transform/manuscript/main.tex
section 5 revision

@@
 %%%% 4. TITLE %%%%
-\title[IND-CCA2 secure KEM using encrypt-then-MAC]{
-Fast Fujisaki-Okamoto transformation  using encrypt-then-mac and applications to Kyber}
+\title{
+Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''}
 %%%% NOTES:
@@
 While the modular FO transform enjoys tight security bound and quantum resistance, it also suffers from computational inefficiency and vulnerabilities to side-channel attacks due to using de-randomization and re-encryption for providing ciphertext integrity.
-In this work, we propose an alternative KEM construction that achieves ciphertext integrity using a message authentication code (MAC) and instantiate a concrete instance using ML-KEM.
+In this work, we propose an alternative KEM construction that achieves ciphertext integrity using a message authentication code (MAC) and instantiate a concrete instance using Kyber.
 Our experimental results showed that where the encryption routine incurs heavy computational cost, replacing re-encryption with MAC provides substantial performance improvements at comparable security level.
@@
 
-The current state of the FO transformation enjoys tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many deficiencies to be improved on.
+The current state of the FO transformation enjoys proven tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many deficiencies to be improved on.
 One such shortcoming is the use of re-encryption for providing ciphertext integrity \cite{bernstein2018towards}, which requires the decapsulation routine to run the encryption routine as a subroutine.
@@
 The problem of ciphertext integrity was solved in symmetric cryptography: given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}.
-We took inspiration from this strategy and applied a similar technique to transform an IND-CPA secure public-key encryption scheme into an IND-CCA2 secure key encapsulation mechanism.
+We took inspiration from this strategy and applied a similar technique to transform an OW-PCA secure public-key encryption scheme into an IND-CCA2 secure key encapsulation mechanism.
 Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of an authenticator, which offers significant performance improvements while maintaining comparable level of security.
@@
 In section \ref{sec:experimental-results}, we instantiate concrete instances of our constructions by combining Kyber with GMAC and Poly1305.
-Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal overhead in the encapsulation routine.
+Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal runtime overhead in the encapsulation routine and a small increase in ciphertext size.
 
@@
 Where the encryption routine is probabilistic, specifying a pseudorandom seed $r \in \mathcal{R}$ will make the encryption routine behave deterministically.
-The decryption routine $\decrypt(\sk, c)$ takes a secret key, a ciphertext, and outputs the decryption $\hat{m}$ if the ciphertext is valid under the given secret key, or the rejection symbol $\bot$ if the ciphertext is invalid.
+The decryption routine $\decrypt(\sk, c)$ takes a secret key, a ciphertext, and outputs the decryption $\hat{m}$ if the ciphertext is valid.
+Some PKE will explicitly reject invalid ciphertext, in which case the decryption routine will output the rejection symbol $\bot$
 
@@
 $\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $K$ if the ciphertext is valid.
-Some KEM constructions use explicit rejection, where if $c$ is invalid then $\decap$ will return a rejection symbol $\bot$; other KEM constructions use implicit rejection, where if $c$ is invalid then $\decap$ will return a ``fake'' session key that depends on the ciphertext and some other secret values.
+Some KEM constructions use explicit rejection, where if $c$ is invalid then $\decap$ will return a rejection symbol $\bot$; other KEM constructions use implicit rejection, where if $c$ is invalid then $\decap$ will return a fake session key that depends on the ciphertext and some other secret values.
 
@@
 Because the encryption routine might be randomized, the session key is derived from both the message and the ciphertext.
-Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$, which is transmitted to the peer.
+Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$ that would be transmitted to the peer.
 In the decapsulation routine, the decryption $\hat{m}$ of the unauthenticated ciphertext is used to re-derive the MAC key $\hat{k}$, which is then used to re-compute the tag $\hat{t}$.
@@
 For an adversary $A$ to produce a valid tag $t$ for some unauthenticated ciphertext $c^\prime$ under the symmetric key $k \leftarrow G(\decrypt(\sk^\prime, c^\prime))$ implies that $A$ must either know the symmetric key $k$ or produce a forgery.
-Under the random oracle model, $A$ also cannot know $k$ without also knowing $\decrypt(\sk^\prime, c^\prime)$, so $A$ must either have produced $c^\prime$ honestly, or have broken the one-way security of $\pke$.
+Under the random oracle model, $A$ also cannot know $k$ without knowing its preimage $\decrypt(\sk^\prime, c^\prime)$, so $A$ must either have produced $c^\prime$ honestly, or have broken the one-way security of $\pke$.
 This means that the decapsulation oracle will not give out information on decryptions that the adversary does not already know.
@@
 However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed.
-This means that a decapsulation oracle can be converted into a plaintext-checking oracle, so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
+This means that a decapsulation oracle can be converted into a plaintext-checking oracle (algorithm \ref{alg:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
 
@@
 
-    Across all $q$ decapsulation queries, the probability that at least one query is a forgery is thus at most $q \cdot \adv_\texttt{OT-MAC}(C)$. By the difference lemma:
+    Across all $q$ decapsulation queries, the probability that at least one query is a forgery is thus at most $q \cdot P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]$.
+By the difference lemma:
 
@@
                     \begin{algorithmic}[1]
+                        \State $z \leftsample \mathcal{M}$
                         \State $k \leftsample \mathcal{K}_\mac$
@@
 When simulating the decapsulation oracle, $B$ uses the plaintext-checking oracle to look for matching queries in $\mathcal{L}^G$.
-When simulaitng the hash oracles, $B$ uses the plaintext-checking oracle to detect when $m^\ast = \decrypt(\sk^\prime, {c^\prime}^\star)$ has been queried.
+When simulating the hash oracles, $B$ uses the plaintext-checking oracle to detect when $m^\ast = \decrypt(\sk^\prime, {c^\prime}^\star)$ has been queried.
 When $m^\ast$ is queried, $B$ terminates $A$ and returns $m^\ast$ to win the OW-PCA game.
@@
             \Comment{Do not output system RNG directly}
-        \State $(\bar{K}, K_\mac) = G(m' \Vert H(\pk))$
+        \State $(\bar{K}, k) = G(m' \Vert H(\pk))$
             \Comment{G is instantiated with SHA3-512}
@@
         \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m', r)$
-            \State $t=\mac(K_\mac, c^\prime)$
+            \State $t=\mac(k, c^\prime)$
         \State $K = \texttt{KDF}(\bar{K} \Vert t)$
@@
         \State $\hat{m} = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c^\prime)$ 
-        \State $(\overline{K}, K_\mac) =   G(m^\prime \Vert h)$
-        \State $\hat{t}=\mac(K_\mac, c)$
+        \State $(\overline{K}, k) =   G(m^\prime \Vert h)$
+        \State $\hat{t}=\mac(k, c)$
         \If{$\hat{t} = t$}
@@
 \section{Conclusions and future works}\label{sec:future-works}
-% TODO: need to rework this
+We applied the ``encrypt-then-MAC'' transformation to Kyber and saw meaningful performance improvements over using de-randomization and re-encryption.
+Unfortunately the resulting KEM does not achieve the desired full IND-CCA2 security, because Kyber is known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}.
+We speculate that while Kyber with ``encrypt-then-MAC'' could not achieve the full IND-CCA2 security, it can still be safe for use in ephemeral key exchange, where each secret key is used to decrypt at most one ciphertext (the KR-PCA requires a few hundred decryption queries to recover the secret key).
 
-We applied the ``encrypt-then-MAC'' transformation to Kyber, which we speculated to be safe to use in ephemeral key exchange where each decapsulation key can only be used to decapsulate one ciphertext, and saw meaningful performance improvements over using the Fujisaki-Okamoto transformation.
-Unfortunately, the resulting KEM does not achieve the desired full IND-CCA2 security.
-This is because every chosen-ciphertext attack agaisnt the ```encrypt-then-MAC''' KEM can be converted into a plaintext-checking attack against the the underlying PKE (see algorithm \ref{alg:pco-using-decap-oracle}), and the IND-CPA subroutines of Kyber are known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}.
-This observation, combined with theorem \ref{thm:ow-pca-implies-kem-ind-cca2}, results in the following security equivalence:
-
-However, if the input $\pke$ is OW-CPA secure and rigid (which implies OW-PCA security), then the $U^\bot_m, U^{\not\bot}_m$ transformations from \cite{hofheinz2017modular} can build an IND-CCA2 secure KEM with almost no overhead, as the encapsulation and decapsulation routine each only adds a hash of the plaintext on top of the input encryption and decryption routine.
-This leaves a gap between OW-PCA security and \emph{OW-CPA + rigidity} in which the ``encrypt-then-MAC'' transformation is a superior strategy for building KEM from a PKE.
-It remains an open problem to find a suitable PKE that is OW-PCA but not rigid.
+In section \ref{sec:main-results}, we showed that if the input PKE is OW-PCA secure, then the resulting KEM is IND-CCA2 secure.
+One sufficient condition for OW-PCA security is one-way security plus rigidity.
+If the input PKE is rigid, then $m = \decrypt(\sk, c)$ is equivalent to $c = \encrypt(\pk, m)$, so a plaintext-checking oracle can be simulated without any secret information.
+However, the $U_m^{\not\bot}$ transformation in \cite{hofheinz2017modular} can already transform an OW-CPA secure and rigid PKE into an IND-CCA2 secure KEM with minimal overhead: the encapsulation and decapsulation routines each adds a hash of the plaintext to the encryption and decryption routine.
+In other words, where the input PKE is rigid, ``encrypt-then-MAC'' doesn't offer any performance advantage.
+It remains an open problem whether there exists a PKE that is OW-PCA secure but not rigid.
+If such a PKE exists, then ``encrypt-then-MAC'' would be a preferable strategy for constructing an IND-CCA2 KEM.
 

### 2024-08-27 etm-transform/manuscript/main.tex
Added experimental data

@@
 
-When instantiating an instance $\kem(\keygen, \encap, \decap)$, there are a variety of possible MAC's to choose from.
-For each of the security level and choice of MAC, we measured the number of CPU cycles needed to produce a digest of Kyber's ciphertext.
-The MAC implementations are taken from OpenSSL and run on an Intel MacBook Pro with i7-9700h (TODO: run it on a server).
-The median measurements are reported in table \ref{tbl:standalone-mac-perf}.
-
-\begin{table}[H]
-    \centering
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        Name & Security & 768 bytes & 1088 bytes & 1568 bytes \\
-        \hline
-        CMAC & many-time & 5022 & 5442 & 6090 \\
-        \hline
-        KMAC-256 & many-time & 7934 & 9862 & 11742 \\
-        \hline
-        GMAC & one-time & 2778 & 2756 & 2762 \\
-        \hline
-        Poly1305 & one-time & 1128 & 1218 & 1338 \\
-        \hline
-    \end{tabular}
-    \caption{Standalone $\mac$ performances}\label{tbl:standalone-mac-perf}
-\end{table}
-
-Based on the security reduction in section \ref{sec:main-results} we chose standalone GMAC and Poly1305 for their substantial performance advantage over other constructions.
-We modified the reference implementation (https://github.com/pq-crystals/kyber) to implement the proposed construction, then measured the CPU cycles needed to run each of routines.
-The median CPU cycles are reported in table \ref{tbl:kyber-ae-perf}.
+When instantiating an instance of $\kem(\keygen, \encap, \decap)$, there are a variety of MAC's to chose from.
+We experimented with Poly1305, AES-256-GCM (aka GMAC), AES-256-CBC (aka CMAC), and KMAC-256.
+We instantiated an intance of the ``encrypt-then-MAC'' KEM using each of the chosen MAC, then measured the median number of CPU cycles needed to perform the key generation, encapsulation, and decapsulation routines among 10000 runs.
+The Kyber implementation is taken from the reference implementation (https://github.com/pq-crystals/kyber).
+MAC implementations are taken from OpenSSL 3.3.1.
+The source code is compiled with GCC 11.4.1 on Amazon Linux 2.
+Performance measurements were taken from a c7a.medium AWS EC2 instance with a AMD EPYC 9R14 (1) @ 3.700GHz.
+The experimental results are listed in table \ref{tbl:kyber-ae-perf}.
 
@@
         \hline
-        Kyber512 & 128 bits &  & &  \\
+        Kyber512 & 128 bits & 75945 & 91467& 121185 \\
         \hline
-        Kyber768 & 192 bits &  & &  \\
+        Kyber512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
         \hline
-        Kyber1024 & 256 bits &  & &  \\
+        Kyber512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
         \hline
-        Kyber512 + GMAC& 128 bits &  & &  \\
+        Kyber512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
         \hline
-        Kyber768 + GMAC& 192 bits &  & &  \\
+        Kyber512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
         \hline
-        Kyber1024 + GMAC& 256 bits &  & &  \\
+        Kyber768 & 192 bits & 129895 & 146405& 186445 \\
         \hline
-        Kyber512 + Poly1305& 128 bits &  & &  \\
+        Kyber768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
         \hline
-        Kyber768 + Poly1305& 192 bits &  & &  \\
+        Kyber768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
         \hline
-        Kyber1024 + Poly1305& 256 bits &  & &  \\
+        Kyber768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
+        \hline
+        Kyber768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
+        \hline
+        Kyber1024 & 256 bits & 194921 & 199185 & 246245 \\
+        \hline
+        Kyber1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
+        \hline
+        Kyber1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
+        \hline
+        Kyber1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
+        \hline
+        Kyber1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
         \hline

### 2024-08-27 etm-transform/manuscript/main.tex
What needs to be done next

@@
     See \cite{boneh2020graduate} lemma 7.11
+    % TODO: need more detailed citations
 \end{proof}
@@
         \State $m \leftsample \{0,1\}^{256}$
-        \State $m'= H(m)$
-            \Comment{Do not output system RNG directly}
-        \State $(\bar{K}, k) = G(m' \Vert H(\pk))$
-            \Comment{G is instantiated with SHA3-512}
-        \State $r \leftsample \mathcal{R}$
-        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m', r)$
+        % \State $m'= H(m)$
+        %     \Comment{Do not output system RNG directly}
+        \State $(\bar{K}, r, k) = G(m \Vert H(\pk))$
+            \Comment{G is instantiated with SHA3-768}
+
+        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m, r)$
             \State $t=\mac(k, c^\prime)$
@@
         \hline
-        Kyber512 & 128 bits & 75945 & 91467& 121185 \\
+        % TODO: change to ML-KEM-512/768/1024
+        ML-KEM-512 & 128 bits & 75945 & 91467& 121185 \\
         \hline
-        Kyber512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
+        ML-KEM-512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
         \hline
-        Kyber512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
+        ML-KEM-512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
         \hline
-        Kyber512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
+        ML-KEM-512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
         \hline
-        Kyber512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
+        ML-KEM-512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
         \hline
-        Kyber768 & 192 bits & 129895 & 146405& 186445 \\
+        ML-KEM-768 & 192 bits & 129895 & 146405& 186445 \\
         \hline
-        Kyber768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
+        ML-KEM-768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
         \hline
-        Kyber768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
+        ML-KEM-768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
         \hline
-        Kyber768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
+        ML-KEM-768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
         \hline
-        Kyber768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
+        ML-KEM-768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
         \hline
-        Kyber1024 & 256 bits & 194921 & 199185 & 246245 \\
+        ML-KEM-1024 & 256 bits & 194921 & 199185 & 246245 \\
         \hline
-        Kyber1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
+        ML-KEM-1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
         \hline
-        Kyber1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
+        ML-KEM-1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
         \hline
-        Kyber1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
+        ML-KEM-1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
         \hline
-        Kyber1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
+        ML-KEM-1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
         \hline
+        % TODO: also do a short comparison with Kyber Round 3
+        % TODO: KeyGen, Encap, and Decap CPU cycles should be separate tables
+        % TODO: also include MAC performances on different ciphertext sizes
     \end{tabular}
@@
 
+\section{Experimental Evaluation}
+We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available \cite{}.
+We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits.
+Our experiments were run on a desktop with a 2.3 GHz Intel Core i9 laptop (Coffee Lake) with 16 GB RAM.
+The codes were compiled using \texttt{clange 14.0.0}.
+%with \texttt{-std=c99 -O1 -fomit-frame-pointer flag}. 
+Table~\ref{tab:ExpResults} reports the cycle counts for three different instances of Keyber, namely .
+The cycle counts reported in Table~\ref{tab:ExpResults} are the average of the cycle counts of 10000 executions of all algorithms.
+\begin{table}[h]
+\centering
+\begin{tabular}{|c|c||c|c| |c|c| |c|c|} \hline
+\multicolumn{4}{|c|}{Kyber512}  & \multicolumn{2}{|c|}{Kyber512+Poly1305} & \multicolumn{2}{|c|}{Kyber512+GMAC} \\ \hline
+\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
+{\bf sk:} & 1632 (or 32) & {\bf $\keygen$:} &  & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
+{\bf pk:} & 800 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
+{\bf ct:} & 768 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
+\hline 
+\multicolumn{4}{|c|}{Kyber768}  & \multicolumn{2}{|c|}{Kyber768+Poly1305} & \multicolumn{2}{|c|}{Kyber768+GMAC} \\ \hline
+\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
+{\bf sk:} & 2400 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
+{\bf pk:} & 1184 & {\bf $\encap$:} & & {\bf $\encap^+$:} &  & {\bf $\encap^+$:} & \\ \hline 
+{\bf ct:} & 1088 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
+\hline 
+\multicolumn{4}{|c|}{Kyber1024}  & \multicolumn{2}{|c|}{Kyber1024+Poly1305} & \multicolumn{2}{|c|}{Kyber1024+GMAC} \\ \hline
+\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
+{\bf sk:} & 3168 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
+{\bf pk:} & 1568 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
+{\bf ct:} & 1568 & {\bf $\decap$:} & & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
+\end{tabular}
+\caption{Cycle counts in AVX2 on a Coffee Lake laptop}
+\label{tab:ExpResults}
+\end{table}
+
+\begin{itemize}
+\item Forward secrecy: Mutual authentication protocol using KEM, computing the number of key-excahnge per second for both Kyber and ours. 
+\item Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
+\item Look at OpenSSL for GMAC with no AES-256. 
+\item Also setup networking experiment on AWS
+\end{itemize}
+
 \section{Conclusions and future works}\label{sec:future-works}
-We applied the ``encrypt-then-MAC'' transformation to Kyber and saw meaningful performance improvements over using de-randomization and re-encryption.
+\emph{Comparison with Fujisaki-Okamoto transformation:} We applied the ``encrypt-then-MAC'' transformation to Kyber and saw meaningful performance improvements over using de-randomization and re-encryption.
 Unfortunately the resulting KEM does not achieve the desired full IND-CCA2 security, because Kyber is known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}.

### 2024-09-10 etm-transform/manuscript/main.tex
Entered KEX performance data, but need nicer formatting

@@
 \newcommand{\fotplus}{\texttt{FOT+}}
+\newcommand{\us}{\mu s}
 
@@
 
+\subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
+A common deployment of key encapsulation mechanism is key exchange.
+\cite{bos2018crystals} described three key exchange protocols: unauthenticated key exchange, unilaterally authenticated key exchange, and mutually authenticated key exchange.
+We instantiated an implementation of each of the key exchange protocol with different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with instantiations using ML-KEM with identical security levels.
+For clarity, we denote the party who initiates the key exchange to be the client, and the other party to be the server.
+
+\subsubsection{Unauthenticated key exchange (UKEX)}\label{sec:unauthenticated-key-exchange}
+In unauthenticated key exchange, the client first generates an ephemeral keypair, the transmits the ephemeral public key to the server.
+The server uses the ephemeral public key to run the encapsulation routine, which returns an ephemeral shared secret and an ephemeral ciphertext.
+The server transmits the ephemeral ciphertext to the client, whereupon the client decapsulates the ephemeral ciphertext to recover the ephemeral shared secret.
+Finally, client and server both runs a key-derivation function on the shared secret to derive the session key.
+
+Note that although the two parties already have a shared secret after client decapsulation of the ephemeral ciphertext, a key derivation function is still applied.
+This is done so that the implementation of the unauthenticated key exchange is consistent with the implementation of other authenticated key exchange, where the session key is derived from multiple shared secrets.
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{UKEX}_\texttt{C}$}\label{alg:unauthenticated-kex-client}
+                \begin{algorithmic}[1]
+                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
+                    \State $\texttt{send}(\pk_e)$
+                    \State $c_e \leftarrow \texttt{read}()$
+                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{UKEX Client}}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{UKEX}_\texttt{S}$}\label{alg:unauthenticated-kex-server}
+                \begin{algorithmic}[1]
+                    \State $\pk_e \leftarrow \texttt{read}()$
+                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
+                    \State $\texttt{send}(c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K_e)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{UKEX Server}}
+    \end{floatrow}
+\end{figure}
+
+The round trip time measures the time between the moment before client generates the ephemeral keypair and the moment after client derives the session key.
+A total of 10000 rounds of key exchanges were performed, with the median and average RTT recorded below:
+
+\begin{table}[H]
+    \caption{Unauthenticated key exchange performance}\label{tbl:unauthenticated-kex}
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-512} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{92$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{97$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-512 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-512 + GMAC} \\
+        \hline
+        Median RTT & 70$\us$ & Median RTT & 73$\us$ \\
+        \hline
+        Average RTT & 72$\us$ & Average RTT & 76$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-512 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-512 + KMAC} \\
+        \hline
+        Median RTT & 75$\us$ & Median RTT & 76$\us$ \\
+        \hline
+        Average RTT & 79$\us$ & Average RTT & 78$\us$ \\
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-768} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{135$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{140$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-768 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-768 + GMAC} \\
+        \hline
+        Median RTT & 99$\us$ & Median RTT & 101$\us$ \\
+        \hline
+        Average RTT & 104$\us$ & Average RTT & 105$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-768 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-768 + KMAC} \\
+        \hline
+        Median RTT & 103$\us$ & Median RTT & 103$\us$ \\
+        \hline
+        Average RTT & 109$\us$ & Average RTT & 107$\us$ \\
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-1024} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{193$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{199$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-1024 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-1024 + GMAC} \\
+        \hline
+        Median RTT & 138$\us$ & Median RTT & 140$\us$ \\
+        \hline
+        Average RTT & 141$\us$ & Average RTT & 145$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-1024 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-1024 + KMAC} \\
+        \hline
+        Median RTT & 143$\us$ & Median RTT & 144$\us$ \\
+        \hline
+        Average RTT & 148$\us$ & Average RTT & 149$\us$ \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\subsubsection{Unilaterally authenticated key exchange(UAKE)}\label{sec:uakex}
+Common communication protocols, including TLS 1.3, perform unilaterally authenticated key exchange, where the server authenticates itself to the client, but client authentication is optional.
+In this scenario, the client possesses the server's long-term public key, while the server possesses the corresponding secret key.
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{UAKE}_\texttt{C}(\pk_s)$}\label{alg:uakex-client}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\pk_s$
+                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
+                    \State $(c_s, K_s) \leftsample \encap(\pk_s)$
+                    \State $\texttt{send}(\pk_e, c_s)$
+                    \State $c_e \leftarrow \texttt{read}()$
+                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_s)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{UAKE Client}}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{UAKE}_\texttt{S}(\sk_s)$}\label{alg:uakex-server}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\sk_s$
+                    \State $(\pk_e, c_s) \leftarrow \texttt{read}()$
+                    \State $K_s \leftarrow \decap(\sk_s, c_s)$
+                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
+                    \State $\texttt{send}(c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_s)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{UKEX Server}}
+    \end{floatrow}
+\end{figure}
+
+\begin{table}[H]
+    \caption{UAKE performance}\label{tbl:uake}
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-512} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{145$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{151$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-512 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-512 + GMAC} \\
+        \hline
+        Median RTT & 103$\us$ & Median RTT & 106$\us$ \\
+        \hline
+        Average RTT & 106$\us$ & Average RTT & 110$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-512 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-512 + KMAC} \\
+        \hline
+        Median RTT & 108$\us$ & Median RTT & 109$\us$ \\
+        \hline
+        Average RTT & 112$\us$ & Average RTT & 113$\us$ \\
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-768} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{215$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{222$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-768 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-768 + GMAC} \\
+        \hline
+        Median RTT & 144$\us$ & Median RTT & 149$\us$ \\
+        \hline
+        Average RTT & 150$\us$ & Average RTT & 156$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-768 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-768 + KMAC} \\
+        \hline
+        Median RTT & 153$\us$ & Median RTT & 154$\us$ \\
+        \hline
+        Average RTT & 160$\us$ & Average RTT & 159$\us$ \\
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-1024} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{310$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{318$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-1024 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-1024 + GMAC} \\
+        \hline
+        Median RTT & 202$\us$ & Median RTT & 212$\us$ \\
+        \hline
+        Average RTT & 209$\us$ & Average RTT & 228$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-1024 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-1024 + KMAC} \\
+        \hline
+        Median RTT & 212$\us$ & Median RTT & 213$\us$ \\
+        \hline
+        Average RTT & 218$\us$ & Average RTT & 220$\us$ \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\subsubsection{Mutually authenticated key exchange (AKE)}\label{sec:akex}
+Mutually authenticated key exchange is largely identical to unilaterally authenticated key exchange, except for that client authentication is required.
+This means that client possesses server's long-term public key and its own long-term secret key, while the server possesses client's long-term public key and its own long-term secret key.
+The session key is derived from three shared secrets.
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{AKE}_\texttt{C}(\pk_s, \sk_c)$}\label{alg:akex-client}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\pk_s$
+                    \Require Client's long-term $\sk_s$
+                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
+                    \State $(c_s, K_s) \leftsample \encap(\pk_s)$
+                    \State $\texttt{send}(\pk_e, c_s)$
+                    \State $(c_e, c_c) \leftarrow \texttt{read}()$
+                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
+                    \State $K_c \leftarrow \decap(\sk_e, c_c)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_s \Vert K_c)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{AKE Client}}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{AKE}_\texttt{S}(\sk_s, \pk_c)$}\label{alg:akex-server}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\sk_s$
+                    \Require Client's long-term $\pk_c$
+                    \State $(\pk_e, c_s) \leftarrow \texttt{read}()$
+                    \State $K_s \leftarrow \decap(\sk_s, c_s)$
+                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
+                    \State $(c_c, K_c) \leftsample \encap(\pk_c)$
+                    \State $\texttt{send}(c_e, c_c)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_s \Vert K_c)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{AKE Server}}
+    \end{floatrow}
+\end{figure}
+
+\begin{table}[H]
+    \caption{AKE performance}\label{tbl:ake}
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-512} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{200$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{213$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-512 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-512 + GMAC} \\
+        \hline
+        Median RTT & 133$\us$ & Median RTT & 139$\us$ \\
+        \hline
+        Average RTT & 138$\us$ & Average RTT & 143$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-512 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-512 + KMAC} \\
+        \hline
+        Median RTT & 143$\us$ & Median RTT & 145$\us$ \\
+        \hline
+        Average RTT & 148$\us$ & Average RTT & 151$\us$ \\
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-768} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{294$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{301$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-768 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-768 + GMAC} \\
+        \hline
+        Median RTT & 190$\us$ & Median RTT & 197$\us$ \\
+        \hline
+        Average RTT & 196$\us$ & Average RTT & 210$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-768 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-768 + KMAC} \\
+        \hline
+        Median RTT & 202$\us$ & Median RTT & 204$\us$ \\
+        \hline
+        Average RTT & 208$\us$ & Average RTT & 210$\us$ \\
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-1024} \\
+        \hline
+        \multicolumn{2}{|c|}{Median RTT}
+        & \multicolumn{2}{|c|}{512$\us$} \\
+        \hline
+        \multicolumn{2}{|c|}{Average RTT}
+        & \multicolumn{2}{|c|}{511$\us$} \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-1024 + Poly1305}
+        & \multicolumn{2}{|c|}{ML-KEM-1024 + GMAC} \\
+        \hline
+        Median RTT & 266$\us$ & Median RTT & 273$\us$ \\
+        \hline
+        Average RTT & 273$\us$ & Average RTT & 282$\us$ \\
+        \hline\hline
+        \multicolumn{2}{|c|}{ML-KEM-1024 + CMAC}
+        & \multicolumn{2}{|c|}{ML-KEM-1024 + KMAC} \\
+        \hline
+        Median RTT & 280$\us$ & Median RTT & 282$\us$ \\
+        \hline
+        Average RTT & 287$\us$ & Average RTT & 288$\us$ \\
+        \hline
+    \end{tabular}
+\end{table}
+
 \section{Experimental Evaluation}
-We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available \cite{}.
+We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available.
 We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits.

### 2024-09-11 etm-transform/manuscript/main.tex
Section 4 needs extensive rewrite

--- 2024-08-27 etm-transform/manuscript/main.tex
@@
 \section{Application to Kyber}\label{sec:experimental-results}
+% TODO: organize into three subsection:
+% 1. Choosing a MAC
+% 2. Comparing CPU cycles
+% 3. Comparing KEX RTT
+
 CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism whose security is based on the conjecture hardness of the decisional Module Learning with Error problem.
@@
         \hline
-        % TODO: change to ML-KEM-512/768/1024
         ML-KEM-512 & 128 bits & 75945 & 91467& 121185 \\
@@
 
+\subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
+A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel.
+\cite{bos2018crystals} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE).
+We instantiated an implementation for each of the three key exchange protocols using different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with implementations instantiated using ML-KEM.
+
+For clarity, we denote the party who sends the first message to be the client and the other party to be the server.
+Round trip time (RTT) is defined to be the time interval between the moment before the client starts generating ephemeral keypairs and the moment after the client derives the final session key.
+All experiements are run on a pair of AWS c7a.medium instances both located in the \texttt{us-west-2} region.
+For each experiment, a total of 10,000 rounds of key exchange are performed, with the median and average round trip time (measured in microsecond) recorded.
+
+\subsubsection{Unauthenticated key exchange (KE)}\label{sec:unauthenticated-key-exchange}
+In unauthenticated key exchange, a single pair of ephemeral keypair $(\pk_e, \sk_e) \leftsample \keygen()$ is generated by the client.
+The client transmits the ephemeral public key $\pk_e$ to the server, who runs the encapsulation routine $(c_e, K_e) \leftsample \encap(\pk_e)$ and transmits the ciphertext $c_e$ back to the client.
+The client finally decapsulates the ciphertext to recover the shared secret $K_e \leftarrow \decap(\sk_e, c_e)$.
+The specific steps are described in algorithms \ref{alg:ke-client}, \ref{alg:ke-server}.
+
+Note that in our implementation, a key derivation function (\texttt{KDF}) is applied to the ephemeral shared secret to derive the final session key.
+This step is added to maintain consistency with other authenticated key exchange protocols, where the final session key is derived from multiple shared secrets.
+The key derivation function is instantiated using Shake256, and the final session key is 256 bits in length.
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{KE}_\texttt{C}$}\label{alg:ke-client}
+                \begin{algorithmic}[1]
+                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
+                    \State $\texttt{send}(\pk_e)$
+                    \State $c_e \leftarrow \texttt{read}()$
+                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{KE Client}}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{KE}_\texttt{S}$}\label{alg:ke-server}
+                \begin{algorithmic}[1]
+                    \State $\pk_e \leftarrow \texttt{read}()$
+                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
+                    \State $\texttt{send}(c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K_e)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{KE Server}}
+    \end{floatrow}
+\end{figure}
+
+The RTT comparison is summarized in table \ref{tbl:ke-rtt}
+
+\begin{table}[H]
+    \caption{Unauthenticated key exchange RTT comparison}\label{tbl:ke-rtt}
+    \begin{tabular}{|c|c|c|}
+        \hline
+        KEM & median RTT ($\us$) & average RTT ($\us$) \\
+        \hline
+        ML-KEM-512 & 92 & 97 \\
+        \hline
+        ML-KEM-512 + Poly1305 & 70 & 72 \\
+        \hline
+        ML-KEM-512 + GMAC & 73 & 76 \\
+        \hline
+        ML-KEM-512 + CMAC & 75 & 79 \\
+        \hline
+        ML-KEM-512 + KMAC & 76 & 78 \\
+        \hline 
+        ML-KEM-768 & 135 & 140 \\
+        \hline
+        ML-KEM-768 + Poly1305 & 99 & 104 \\
+        \hline
+        ML-KEM-768 + GMAC & 101 & 105 \\
+        \hline
+        ML-KEM-768 + CMAC & 103 & 109 \\
+        \hline
+        ML-KEM-768 + KMAC & 103 & 107 \\
+        \hline
+        ML-KEM-1024 & 193 & 199 \\
+        \hline
+        ML-KEM-1024 + Poly1305 & 138 & 141 \\
+        \hline
+        ML-KEM-1024 + GMAC & 140 & 145 \\
+        \hline
+        ML-KEM-1024 + CMAC & 143 & 148 \\
+        \hline
+        ML-KEM-1024 + KMAC & 144 & 149 \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\subsubsection{Unilaterally authenticated key exchange (UAKE)}\label{sec:uakex}
+In unilaterally authenticated key exchange, the authenticating party proves its identity to the other party by demonstrating possession of a secret key that corresponds to a published long-term public key.
+In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by demonstrating possession of the corresponding long-term secret key $\sk_S$.
+Details are described in algorithms \ref{alg:uake-client} and \ref{alg:uake-server}.
+
+In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key.
+This helps the key exchange to achieve weak forward secrecy (citation needed).
+
+Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
+KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage.
+(citation needed).
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{UAKE}_\texttt{C}(\pk_S)$}\label{alg:uake-client}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\pk_S$
+                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
+                    \State $(c_S, K_S) \leftsample \encap(\pk_S)$
+                    \State $\texttt{send}(\pk_e, c_S)$
+                    \State $c_e \leftarrow \texttt{read}()$
+                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{UAKE Client}}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{UAKE}_\texttt{S}(\sk_S)$}\label{alg:uake-server}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\sk_S$
+                    \State $(\pk_e, c_S) \leftarrow \texttt{read}()$
+                    \State $K_S \leftarrow \decap(\sk_S, c_S)$
+                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
+                    \State $\texttt{send}(c_e)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{UAKE Server}}
+    \end{floatrow}
+\end{figure}
+
+\begin{table}[H]
+    \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
+    \begin{tabular}{|c|c|c|}
+        \hline
+        KEM & median RTT ($\us$) & average RTT ($\us$) \\
+        \hline
+        ML-KEM-512 & 145 & 151 \\
+        \hline
+        ML-KEM-512 + Poly1305 & 103 & 106 \\
+        \hline
+        ML-KEM-512 + GMAC & 106 & 110 \\
+        \hline
+        ML-KEM-512 + CMAC & 108 & 112 \\
+        \hline
+        ML-KEM-512 + KMAC & 109 & 113 \\
+        \hline 
+        ML-KEM-768 & 215 & 222 \\
+        \hline
+        ML-KEM-768 + Poly1305 & 144 & 150 \\
+        \hline
+        ML-KEM-768 + GMAC & 149 & 156 \\
+        \hline
+        ML-KEM-768 + CMAC & 153 & 160 \\
+        \hline
+        ML-KEM-768 + KMAC & 154 & 159 \\
+        \hline
+        ML-KEM-1024 & 310 & 318 \\
+        \hline
+        ML-KEM-1024 + Poly1305 & 202 & 209 \\
+        \hline
+        ML-KEM-1024 + GMAC & 212 & 228 \\
+        \hline
+        ML-KEM-1024 + CMAC & 212 & 218 \\
+        \hline
+        ML-KEM-1024 + KMAC & 213 & 220 \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\subsubsection{Mutually authenticated key exchange (AKE)}\label{sec:akex}
+Mutually authenticated key exchange is largely identical to unilaterally authenticated key exchange, except for that client authentication is required.
+This means that client possesses server's long-term public key and its own long-term secret key, while the server possesses client's long-term public key and its own long-term secret key.
+The session key is derived by applying \texttt{KDF} onto the concatenation of shared secrets produced under the ephemeral keypair, server's long-term keypair, and client's long-term keypair, in this order.
+
+\begin{figure}[H]
+    \begin{floatrow}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{AKE}_\texttt{C}(\pk_S, \sk_C)$}\label{alg:ake-client}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\pk_S$
+                    \Require Client's long-term $\sk_C$
+                    \State $(\pk_e, \sk_e) \leftsample \keygen()$
+                    \State $(c_S, K_S) \leftsample \encap(\pk_S)$
+                    \State $\texttt{send}(\pk_e, c_S)$
+                    \State $(c_e, c_C) \leftarrow \texttt{read}()$
+                    \State $K_e \leftarrow \decap(\sk_e, c_e)$
+                    \State $K_C \leftarrow \decap(\sk_e, c_C)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S \Vert K_C)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{AKE Client}}
+        \ffigbox{
+            \begin{algorithm}[H]
+                \caption{$\texttt{AKE}_\texttt{S}(\sk_S, \pk_C)$}\label{alg:akex-server}
+                \begin{algorithmic}[1]
+                    \Require Server's long-term $\sk_S$
+                    \Require Client's long-term $\pk_C$
+                    \State $(\pk_e, c_S) \leftarrow \texttt{read}()$
+                    \State $K_S \leftarrow \decap(\sk_S, c_S)$
+                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
+                    \State $(c_C, K_C) \leftsample \encap(\pk_C)$
+                    \State $\texttt{send}(c_e, c_C)$
+                    \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S \Vert K_C)$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        }{\caption{AKE Server}}
+    \end{floatrow}
+\end{figure}
+
+\begin{table}[H]
+    \caption{AKE RTT comparison}\label{tbl:ake-rtt}
+    \begin{tabular}{|c|c|c|}
+        \hline
+        KEM & median RTT ($\us$) & average RTT ($\us$) \\
+        \hline
+        ML-KEM-512 & 200 & 213 \\
+        \hline
+        ML-KEM-512 + Poly1305 & 133 & 138 \\
+        \hline
+        ML-KEM-512 + GMAC & 139 & 143 \\
+        \hline
+        ML-KEM-512 + CMAC & 143 & 148 \\
+        \hline
+        ML-KEM-512 + KMAC & 145 & 151 \\
+        \hline 
+        ML-KEM-768 & 294 & 301 \\
+        \hline
+        ML-KEM-768 + Poly1305 & 190 & 196 \\
+        \hline
+        ML-KEM-768 + GMAC & 197 & 210 \\
+        \hline
+        ML-KEM-768 + CMAC & 202 & 208 \\
+        \hline
+        ML-KEM-768 + KMAC & 204 & 210 \\
+        \hline
+        ML-KEM-1024 & 512 & 511 \\
+        \hline
+        ML-KEM-1024 + Poly1305 & 266 & 273 \\
+        \hline
+        ML-KEM-1024 + GMAC & 273 & 282 \\
+        \hline
+        ML-KEM-1024 + CMAC & 280 & 287 \\
+        \hline
+        ML-KEM-1024 + KMAC & 282 & 288 \\
+        \hline
+    \end{tabular}
+\end{table}
+
 \section{Experimental Evaluation}
-We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available \cite{}.
+We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available.
 We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits.
@@
 \begin{itemize}
-\item Forward secrecy: Mutual authentication protocol using KEM, computing the number of key-excahnge per second for both Kyber and ours. 
 \item Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables

### 2024-09-12 etm-transform/manuscript/main.tex
Need to extensively rewrite sectino 4

@@
 
+\section{Implementation}\label{sec:implementation}
+ML-KEM (based on \cite{bos2018crystals}\cite{avanzi2019crystals}) is an IND-CCA2 secure key encapsulation mechanism standardized by NIST in FIPS 203.
+ML-KEM achieves IND-CCA2 security first by constructing a public-key encryption scheme whose IND-CPA security is derived from the hardness of the module learning with error (MLWE) problems, then applying the Fujisaki-Okamoto transformation.
+We extended the reference C implementation of ML-KEM (https://github.com/pq-crystals/kyber) to implement our ``encrypt-then-MAC'' transformation and compared the performance characteristics of our instantiation with ML-KEM.
+
+\subsection{Choosing a message authenticator}
+
+\subsection{KEM performance}
+
+\subsection{Key exchange performance}
+
 \section{Application to Kyber}\label{sec:experimental-results}

### 2024-09-12 etm-transform/manuscript/main.tex
Some table

--- 2024-09-11 etm-transform/manuscript/main.tex
@@
 
+\section{Implementation}\label{sec:implementation}
+Originally known as Kyber \cite{bos2018crystals}\cite{avanzi2019crystals}, ML-KEM is an \texttt{IND-CCA2} secure key encapsulation mechanism standardized in FIPS 203 by NIST.
+The \texttt{IND-CCA2} security of ML-KEM is achieved in two steps.
+First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries.
+Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}.\keygen, \texttt{K-PKE}.\encrypt, \texttt{K-PKE}.\decrypt$ as sub-routines.
+Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine suffers from computational inefficiency.
+
+We implemented the ``encrypt-then-MAC'' KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios.
+The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
+
+A detailed description of $\texttt{K-PKE}$'s routines can be found in FIPS 203 (TODO: citation).
+The ``encrypt-then-MAC'' routines are listed in algorithms \ref{alg:ml-kem-etm-keygen}, \ref{alg:ml-kem-etm-encap}, and \ref{alg:ml-kem-etm-decap}.
+
+\begin{algorithm}[H]
+    \caption{\texttt{ML-KEM-ETM.KeyGen()}}\label{alg:ml-kem-etm-keygen}
+    \begin{algorithmic}[1]
+        \State $z \leftsample \{0,1\}^{256}$
+        \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE}.\keygen()$
+        \State $h \leftarrow H(\pk)$
+            \Comment{$H$ is \texttt{SHA3-256}}
+        \State $\sk \leftarrow (sk^\prime \Vert \pk \Vert h \Vert z)$
+        \State \Return $(\pk, \sk)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{\texttt{ML-KEM-ETM.Encap}(\pk)}\label{alg:ml-kem-etm-encap}
+    \begin{algorithmic}[1]
+        \Require Public key $\pk$
+        \State $m \leftsample \{0,1\}^{256}$
+        \State $(\overline{K}, r, k) \leftarrow \texttt{XOF}(m \Vert H(\pk))$
+            \Comment{\texttt{XOF} is \texttt{Shake256}, outputting 768 bits}
+        \State $c^\prime \leftarrow \texttt{K-PKE}.\encrypt(\pk, m, r)$
+        \State $t \leftarrow \mac(k, c^\prime)$
+        \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
+            \Comment{\texttt{KDF} is \texttt{Shake256}, outputting 256 bits}
+        \State $c \leftarrow (c^\prime, c)$
+        \State \Return $(c, K)$
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+    \caption{$\texttt{ML-KEM-ETM.Decap}(\sk, c)$}\label{alg:ml-kem-etm-decap}
+    \begin{algorithmic}[1]
+        \Require Secret key $\sk = (\sk^\prime \Vert \pk \Vert h \Vert z)$
+        \Require Ciphertext $c = (c^\prime \Vert t)$
+        \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
+        \State $(c^\prime, t) \leftarrow c$
+        \State $\hat{m} \leftarrow \texttt{K-PKE}.\decrypt(\sk^\prime, c^\prime)$
+        \State $(\overline{K}, \hat{r}, \hat{k}) \leftarrow \texttt{XOF}(\hat{m} \Vert h)$
+        \State $\hat{t} \leftarrow \mac(\hat{k}, c^\prime)$
+        \If{$\hat{t} = t$}
+            \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
+        \Else
+            \State $K \leftarrow \texttt{KDF}(z \Vert t)$
+        \EndIf
+        \State \Return $K$
+    \end{algorithmic}
+\end{algorithm}
+
+Our implementation extended from the reference implementation by the PQCrystals team (https://github.com/pq-crystals/kyber).
+All C code is compiled with \texttt{GCC 11.4.1} and \texttt{OpenSSL 3.0.8}.
+All binaries are executed on an AWS c7a.medium instance with an AMD EPYC 9R14 CPU at 3.7 GHz and 1 GB of RAM.
+
+\subsection{Choosing a message authenticator}
+When instantiating \texttt{ML-KEM-ETM}, there are a variety of message authentication codes to choose from. We selected four categories that cover a wide range of designs: \begin{enumerate}
+    \item Poly1305, a Carter-Wegman style MAC operating on a prime field
+    \item GMAC (AES-256-GCM), a Carter-Wegman style MAC operating on a binary field
+    \item CMAC (AES-256-CBC), a CBC-MAC
+    \item KMAC-256, which is a keyed hash function based on Keccak
+\end{enumerate}
+
+We tested each MAC's throughput by measuring the CPU cycles needed to compute a tag on the ciphertext returned by \texttt{K-PKE}.
+The length of the ciphertext varies depending on the security level.
+The measurements are summarized in table \ref{tbl:mac-performance}.
+
+\begin{table}
+    \caption{MAC performance}\label{tbl:mac-performance}
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        \multicolumn{4}{|c|}{ML-KEM-512} \\
+        \hline
+        \multicolumn{2}{|c|}{Ciphertext size (bytes): }
+        & \multicolumn{2}{|c|}{768} \\
+        \hline
+        \multicolumn{2}{|c|}{Poly1305}
+        & \multicolumn{2}{|c|}{GMAC} \\
+        \hline
+        Median & 909 & Median & 3899 \\
+        \hline
+        Average & 2823 & Average & 4859 \\
+        \hline
+        \multicolumn{2}{|c|}{CMAC}
+        & \multicolumn{2}{|c|}{KMAC} \\
+        \hline
+        Median & 6291 & Median & 7565 \\
+        \hline
+        Average & 6373 & Average & 7791 \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\subsection{KEM performance}
+
+\subsection{Key exchange performance}
+
 \section{Application to Kyber}\label{sec:experimental-results}

### 2024-09-13 etm-transform/manuscript/main.tex
Widened table; look like so much more stuff!
Drafted the Implementation section

@@
 \subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
+% TODO: write some intuition first?
 \begin{proof}
@@
 \subsection{Choosing a message authenticator}
-When instantiating \texttt{ML-KEM-ETM}, there are a variety of message authentication codes to choose from. We selected four categories that cover a wide range of designs: \begin{enumerate}
+When instantiating \texttt{ML-KEM-ETM}, there are a variety of message authentication codes to choose from. We selected four instances covering a wide range of designs: \begin{enumerate}
     \item Poly1305, a Carter-Wegman style MAC operating on a prime field
@@
     \item CMAC (AES-256-CBC), a CBC-MAC
-    \item KMAC-256, which is a keyed hash function based on Keccak
+    \item KMAC-256, a keyed hash function based on Keccak
 \end{enumerate}
@@
 
-\begin{table}
+\begin{table}[H]
     \caption{MAC performance}\label{tbl:mac-performance}
-    \begin{tabular}{|c|c|c|c|}
-        \hline
-        \multicolumn{4}{|c|}{ML-KEM-512} \\
-        \hline
-        \multicolumn{2}{|c|}{Ciphertext size (bytes): }
-        & \multicolumn{2}{|c|}{768} \\
+    \begin{tabular}{|c|c|c|c|c|c|c|c|}
+        \hline
+        \multicolumn{8}{|c|}{ML-KEM-512} \\
+        \hline
+        \multicolumn{4}{|c|}{Ciphertext size (bytes): }
+        & \multicolumn{4}{|c|}{768} \\
         \hline
         \multicolumn{2}{|c|}{Poly1305} 
-        & \multicolumn{2}{|c|}{GMAC} \\
-        \hline
-        Median & 909 & Median & 3899 \\
-        \hline
-        Average & 2823 & Average & 4859 \\
-        \hline
-        \multicolumn{2}{|c|}{CMAC}
+        & \multicolumn{2}{|c|}{GMAC} 
+        & \multicolumn{2}{|c|}{CMAC} 
         & \multicolumn{2}{|c|}{KMAC} \\
         \hline
-        Median & 6291 & Median & 7565 \\
-        \hline
-        Average & 6373 & Average & 7791 \\
+        Median & 909 
+        & Median & 3899 
+        & Median & 6291 
+        & Median & 6373\\
+        \hline
+        Average & 2823 
+        & Average & 4859 
+        & Average & 6373 
+        & Average & 7791 \\
+        \hline
+        \hline
+        \multicolumn{8}{|c|}{ML-KEM-768} \\
+        \hline
+        \multicolumn{4}{|c|}{Ciphertext size (bytes): }
+        & \multicolumn{4}{|c|}{1088} \\
+        \hline
+        \multicolumn{2}{|c|}{Poly1305} 
+        & \multicolumn{2}{|c|}{GMAC} 
+        & \multicolumn{2}{|c|}{CMAC} 
+        & \multicolumn{2}{|c|}{KMAC} \\
+        \hline
+        Median & 961 
+        & Median & 3899 
+        & Median & 7305 
+        & Median & 9697 \\
+        \hline
+        Average & 2704 
+        & Average & 4827 
+        & Average & 7588 
+        & Average & 9928 \\
+        \hline
+        \hline
+        \multicolumn{8}{|c|}{ML-KEM-1024} \\
+        \hline
+        \multicolumn{4}{|c|}{Ciphertext size (bytes): }
+        & \multicolumn{4}{|c|}{1568} \\
+        \hline
+        \multicolumn{2}{|c|}{Poly1305} 
+        & \multicolumn{2}{|c|}{GMAC} 
+        & \multicolumn{2}{|c|}{CMAC} 
+        & \multicolumn{2}{|c|}{KMAC} \\
+        \hline
+        Median & 1065 
+        & Median & 4055 
+        & Median & 8735 
+        & Median & 11647 \\
+        \hline
+        Average & 1809 
+        & Average & 5026 
+        & Average & 8772 
+        & Average & 12186 \\
         \hline
@@
 \subsection{KEM performance}
-
-\subsection{Key exchange performance}
-
-\section{Application to Kyber}\label{sec:experimental-results}
-% TODO: organize into three subsection:
-% 1. Choosing a MAC
-% 2. Comparing CPU cycles
-% 3. Comparing KEX RTT
-
-CRYSTALS-Kyber \cite{bos2018crystals}\cite{avanzi2019crystals} is an IND-CCA2 secure key encapsulation mechanism whose security is based on the conjecture hardness of the decisional Module Learning with Error problem.
-To achieve the the IND-CCA2 security, Kyber first constructs an IND-CPA secure public key encryption scheme based on \cite{lpr2013ideal}, then apply a slightly modified variation of the Fujisaki-Okamoto transformation described in \cite{hofheinz2017modular}.
-The resulting decapsulation routine is especially inefficient because Kyber's IND-CPA encryption routine incurs significantly more computational cost than the decryption routine.
-This makes Kyber a prime target for demonstrating the performance improvements enjoyed by the ``encrypt-then-MAC'' KEM construction.
-
-We took the IND-CPA PKE routines (algorithms 4, 5, 6 in \cite{avanzi2019crystals}) and applied the ``encrypt-then-MAC'' transformation.
-The resulting KEM routines are described in algorithms \ref{alg:kyber-ae-keygen}, \ref{alg:kyber-ae-encap}, and \ref{alg:kyber-ae-decap}.
-
-\begin{algorithm}[H]
-    \caption{\texttt{Kyber.CCAKEM.KeyGen()}}\label{alg:kyber-ae-keygen}
-    \begin{algorithmic}[1]
-        \State $z \leftsample \{0,1\}^{256}$
-        \State $(\pk, \sk^\prime) \leftsample \texttt{Kyber.CPAPKE.KeyGen()}$
-        \State $\sk = (\sk^\prime, \pk, H(\pk), z)$
-            \Comment{H is instantiated with SHA3-256}
-        \State \Return $(\pk, \sk)$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-    \caption{\texttt{Kyber.CCAKEM.Encap$^+$(\pk)}}\label{alg:kyber-ae-encap}
-    \begin{algorithmic}[1]
-        \State $m \leftsample \{0,1\}^{256}$
-        % \State $m'= H(m)$
-        %     \Comment{Do not output system RNG directly}
-        \State $(\bar{K}, r, k) = G(m \Vert H(\pk))$
-            \Comment{G is instantiated with SHA3-768}
-
-        \State $c^\prime  \leftarrow \texttt{Kyber.CPAPKE.Enc}(\pk, m, r)$
-            \State $t=\mac(k, c^\prime)$
-        \State $K = \texttt{KDF}(\bar{K} \Vert t)$
-            \Comment{\texttt{KDF} is instantiated with Shake256}
-        \State $c \leftarrow (c^\prime, t)$
-        \State \Return $(c, K)$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-    \caption{$\texttt{Kyber.CCAKEM.Decap}^+ (\sk, c)$}\label{alg:kyber-ae-decap}
-    \begin{algorithmic}[1]
-        \Require Secret key $\sk = (\sk^\prime, \pk, H(\pk), z)$
-        \Require Ciphertext $c = (c^\prime, t)$
-        \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
-        \State $(c^\prime, t) \leftarrow c$
-        \State $\hat{m} = \texttt{Kyber.CPAPKE.Dec}(\sk^\prime, c^\prime)$ 
-        \State $(\overline{K}, k) =   G(m^\prime \Vert h)$
-        \State $\hat{t}=\mac(k, c)$
-        \If{$\hat{t} = t$}
-            \State $K  = \texttt{KDF}(\bar{K} \Vert t)$ 
-         \Else 
-            \State $K  = \texttt{KDF}(z \Vert t)$ 
-        \EndIf
-            \State \Return $K$
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{remark}
-    We derive the MAC key from both the plaintext $m$ and the public key $\pk$ for the same reason \cite{avanzi2019crystals} derives the pseudorandom coin from both $m$ and $\pk$.
-One is to allow both parties to participate in the encryption process, the other is to prevent a quantum adversary from pre-computing a large table of MAC keys that can then be brute-forced.
-\end{remark}
-\begin{remark}
-    We chose to derive the shared secret $K$ from the ciphertext digest $t$ instead of the ciphertext itself, which saves a few Keccak permutations since $t$ is only 128 bits while the full ciphertext could span more than a thousand bytes.
-This should not impact the security of the scheme since finding collision for an unknown MAC key constitutes a forgery attack on the MAC.
-\end{remark}
-\begin{remark}
-    We constructed the ``encrypt-then-MAC'' transformation to use implicit rejection so that we can directly use existing implementation of Kyber with minimal modification.
-In principle, a construction with explicit rejection should also be equally secure and efficient.
-\end{remark}
-\begin{remark}
-    We chose to use MAC with 256-bit key size and 128-bit tag size.
-When modeling the security threats, we assumed that the adversary may have access to quantum computers, making it necessary to use the maximal key size of common MACs while maintaining the minimum 128-bit security.
-On the other hand, the tag size can be relative small because the decapsulator (aka the decapsulation oracle) is assumed to be a classical computer, so there is no quantum speedup on brute-forcing a valid tag.
-\end{remark}
-
-When instantiating an instance of $\kem(\keygen, \encap, \decap)$, there are a variety of MAC's to chose from.
-We experimented with Poly1305, AES-256-GCM (aka GMAC), AES-256-CBC (aka CMAC), and KMAC-256.
-We instantiated an intance of the ``encrypt-then-MAC'' KEM using each of the chosen MAC, then measured the median number of CPU cycles needed to perform the key generation, encapsulation, and decapsulation routines among 10000 runs.
-The Kyber implementation is taken from the reference implementation (https://github.com/pq-crystals/kyber).
-MAC implementations are taken from OpenSSL 3.3.1.
-The source code is compiled with GCC 11.4.1 on Amazon Linux 2.
-Performance measurements were taken from a c7a.medium AWS EC2 instance with a AMD EPYC 9R14 (1) @ 3.700GHz.
-The experimental results are listed in table \ref{tbl:kyber-ae-perf}.
-
-\begin{table}[H]
-    \centering
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        Name & Security level & $\keygen$ & $\encap$ & $\decap$ \\
-        \hline
-        ML-KEM-512 & 128 bits & 75945 & 91467& 121185 \\
-        \hline
-        ML-KEM-512 + Poly1305& 128 bits & 76907 & 93157 & 33733 \\
-        \hline
-        ML-KEM-512 + GMAC& 128 bits & 76917 & 95419 & 37725 \\
-        \hline
-        ML-KEM-512 + CMAC& 128 bits & 76907 & 99839 & 40117 \\
-        \hline
-        ML-KEM-512 + KMAC-256& 128 bits & 76387 & 101009 & 40741 \\
-        \hline
-        ML-KEM-768 & 192 bits & 129895 & 146405& 186445 \\
-        \hline
-        ML-KEM-768 + Poly1305& 192 bits & 128205 & 146405& 43315 \\
-        \hline
-        ML-KEM-768 + GMAC& 192 bits & 127997 & 149525 & 46513 \\
-        \hline
-        ML-KEM-768 + CMAC& 192 bits & 129167 & 151007 & 49841 \\
-        \hline
-        ML-KEM-768 + KMAC-256& 192 bits & 128829 & 155219 & 52415 \\
-        \hline
-        ML-KEM-1024 & 256 bits & 194921 & 199185 & 246245 \\
-        \hline
-        ML-KEM-1024 + Poly1305& 256 bits & 196013 & 205763 & 51375 \\
-        \hline
-        ML-KEM-1024 + GMAC& 256 bits & 196039 & 208805 & 54573 \\
-        \hline
-        ML-KEM-1024 + CMAC& 256 bits & 195389 & 213667 & 59175 \\
-        \hline
-        ML-KEM-1024 + KMAC-256& 256 bits & 196117 & 216761 & 62269 \\
-        \hline
-        % TODO: also do a short comparison with Kyber Round 3
-        % TODO: KeyGen, Encap, and Decap CPU cycles should be separate tables
-        % TODO: also include MAC performances on different ciphertext sizes
+Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the ``encrypt-then-MAC'' transformation the following trade-off when given the same input sub-routines: \begin{enumerate}
+    \item Both encapsulation and decapsulation add a small amount of overhead for needing to hash both the PKE plaintext and the PKE ciphertext when deriving the shared secret, where as the $U^{\not\bot}_m$ transformation only needs to hash the PKE plaintext.
+    \item The encapsulation routine adds a small amount of run-time overhead for computing the authenticator
+    \item The decapsulation routine enjoys substantial runtime speedup because \emph{re-encryption} is replaced with computing an authenticator
+    \item Ciphertext size increases by the size of an authenticator
+\end{enumerate}
+
+Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the ``encrypt-then-MAC'' transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
+A comparison between ML-KEM and variations of the ML-KEM-ETM can be found in table \ref{tbl:kem-performance}
+
+\begin{table}
+    \caption{CPU cycles of each KEM routine}\label{tbl:kem-performance}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|}
+        \hline\hline
+        \multicolumn{8}{|c|}{\bf ML-KEM-512} \\
+        \hline
+        \multicolumn{2}{|c|}{public key size} 
+        & \multicolumn{2}{|c|}{800 bytes}
+        & \multicolumn{4}{|c|}{ \keygen} \\
+        \hline
+        \multicolumn{2}{|c|}{secret key size} 
+        & \multicolumn{2}{|c|}{1632 bytes} 
+        & \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{75945} \\
+        \hline
+        \multicolumn{2}{|c|}{ciphertext size} 
+        & \multicolumn{2}{|c|}{768 bytes} 
+        & \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{76171} \\
+        \hline
+        \multicolumn{4}{|c|}{ \encap}
+        & \multicolumn{4}{|c|}{\decap} \\
+        \hline
+        \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{91467} 
+        & \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{121185} \\
+        \hline
+        \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{92065} 
+        & \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{121650} \\
+        \hline
+        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + Poly1305}
+        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + GMAC} \\
+        \hline
+        \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$}
+        & \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$} \\
+        \hline
+        Median & 93157 & Median & 33773 
+        & Median & 97369 & Median & 37725 \\
+        \hline
+        Average & 93626 & Average & 33908 
+        & Average & 97766 & Average & 37831 \\
+        \hline
+        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + CMAC}
+        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + KMAC} \\
+        \hline
+        \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$}
+        & \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$} \\
+        \hline
+        Median & 99839 & Median & 40117 & Median & 101009 & Median & 40741 \\
+        \hline
+        Average & 99959 & Average & 39943 & Average & 101313 & Average & 40916 \\
+        \hline\hline
+        \multicolumn{8}{|c|}{\bf ML-KEM-768} \\
+        \hline
+        \multicolumn{2}{|c|}{public key size} 
+        & \multicolumn{2}{|c|}{1184 bytes}
+        & \multicolumn{4}{|c|}{ \keygen} \\
+        \hline
+        \multicolumn{2}{|c|}{secret key size} 
+        & \multicolumn{2}{|c|}{2400 bytes} 
+        & \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{129895} \\
+        \hline
+        \multicolumn{2}{|c|}{ciphertext size} 
+        & \multicolumn{2}{|c|}{1088 bytes} 
+        & \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{130650} \\
+        \hline
+        \multicolumn{4}{|c|}{ \encap}
+        & \multicolumn{4}{|c|}{\decap} \\
+        \hline
+        \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{146405} 
+        & \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{186445} \\
+        \hline
+        \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{147400} 
+        & \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{187529} \\
+        \hline
+        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + Poly1305}
+        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + GMAC} \\
+        \hline
+        \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$}
+        & \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$} \\
+        \hline
+        Median & 146405 & Median & 43315 
+        & Median & 149525 & Median & 46513 \\
+        \hline
+        Average & 146860 & Average & 43463 
+        & Average & 150128 & Average & 46706 \\
+        \hline
+        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + CMAC}
+        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + KMAC} \\
+        \hline
+        \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$}
+        & \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$} \\
+        \hline
+        Median & 153139 & Median & 49841 & Median & 155219 & Median & 52415 \\
+        \hline
+        Average & 153735 & Average & 50074 & Average & 155848 & Average & 52611 \\
+        \hline\hline
+        \multicolumn{8}{|c|}{\bf ML-KEM-1024} \\
+        \hline
+        \multicolumn{2}{|c|}{public key size} 
+        & \multicolumn{2}{|c|}{1568 bytes}
+        & \multicolumn{4}{|c|}{ \keygen} \\
+        \hline
+        \multicolumn{2}{|c|}{secret key size} 
+        & \multicolumn{2}{|c|}{3168 bytes} 
+        & \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{194921} \\
+        \hline
+        \multicolumn{2}{|c|}{ciphertext size} 
+        & \multicolumn{2}{|c|}{1568 bytes} 
+        & \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{195465} \\
+        \hline
+        \multicolumn{4}{|c|}{ \encap}
+        & \multicolumn{4}{|c|}{\decap} \\
+        \hline
+        \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{199185} 
+        & \multicolumn{2}{|c|}{Median} 
+        & \multicolumn{2}{|c|}{246245} \\
+        \hline
+        \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{199903} 
+        & \multicolumn{2}{|c|}{Average} 
+        & \multicolumn{2}{|c|}{247320} \\
+        \hline
+        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + Poly1305}
+        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + GMAC} \\
+        \hline
+        \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$}
+        & \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$} \\
+        \hline
+        Median & 205763 & Median & 51375 
+        & Median & 208805 & Median & 54573 \\
+        \hline
+        Average & 206499 & Average & 51562 
+        & Average & 209681 & Average & 54780 \\
+        \hline
+        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + CMAC}
+        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + KMAC} \\
+        \hline
+        \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$}
+        & \multicolumn{2}{|c|}{ $\encap$}
+        & \multicolumn{2}{|c|}{$\decap$} \\
+        \hline
+        Median & 213667 & Median & 59175 & Median & 216761 & Median & 62269 \\
+        \hline
+        Average & 214483 & Average & 59408 & Average & 217468 & Average & 62516 \\
+        \hline\hline
     \end{tabular}
-    \caption{Performance measurements}\label{tbl:kyber-ae-perf}
 \end{table}
-
-Compared to Kyber using the FO transform, our proposed construction adds a small amount of runtime overhead (for computing a digest) in the encapsulation routine and a small increase in ciphertext size (for the 128-bit tag).
-In exchange, we see significant performance runtime savings in the decapsulation routines.
-This trade-off is especially meaningful in a key exchange protocol where one party has substantially more computationl resource than the other.
-For example, in many experimental implementation of TLS with post-quantum KEMs (such as \href{https://blog.cloudflare.com/the-tls-post-quantum-experiment/}{CECPQ2}), the client (might be IoT devices) runs $\keygen$ and $\decap$ while the server (usually data centers) runs $\encap$.
 
@@
 
-\section{Experimental Evaluation}
-We implement our scheme in C on top of {\sf CRYSTALS-Kyber}'s implementation available.
-We use the Poly1305 MAC from {\sf OpenSSL} with a key of size 256 bits and a tag size of 128 bits.
-Our experiments were run on a desktop with a 2.3 GHz Intel Core i9 laptop (Coffee Lake) with 16 GB RAM.
-The codes were compiled using \texttt{clange 14.0.0}.
-%with \texttt{-std=c99 -O1 -fomit-frame-pointer flag}. 
-Table~\ref{tab:ExpResults} reports the cycle counts for three different instances of Keyber, namely .
-The cycle counts reported in Table~\ref{tab:ExpResults} are the average of the cycle counts of 10000 executions of all algorithms.
-\begin{table}[h]
-\centering
-\begin{tabular}{|c|c||c|c| |c|c| |c|c|} \hline
-\multicolumn{4}{|c|}{Kyber512}  & \multicolumn{2}{|c|}{Kyber512+Poly1305} & \multicolumn{2}{|c|}{Kyber512+GMAC} \\ \hline
-\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
-{\bf sk:} & 1632 (or 32) & {\bf $\keygen$:} &  & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
-{\bf pk:} & 800 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
-{\bf ct:} & 768 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
-\hline 
-\multicolumn{4}{|c|}{Kyber768}  & \multicolumn{2}{|c|}{Kyber768+Poly1305} & \multicolumn{2}{|c|}{Kyber768+GMAC} \\ \hline
-\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
-{\bf sk:} & 2400 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
-{\bf pk:} & 1184 & {\bf $\encap$:} & & {\bf $\encap^+$:} &  & {\bf $\encap^+$:} & \\ \hline 
-{\bf ct:} & 1088 & {\bf $\decap$:} &  & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
-\hline 
-\multicolumn{4}{|c|}{Kyber1024}  & \multicolumn{2}{|c|}{Kyber1024+Poly1305} & \multicolumn{2}{|c|}{Kyber1024+GMAC} \\ \hline
-\multicolumn{2}{|l|}{Size in bytes} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} & \multicolumn{2}{|l|}{Clock cycles (AVX2)} \\ \hline 
-{\bf sk:} & 3168 (or 32) & {\bf $\keygen$:} & & {\bf $\keygen$:} & & {\bf $\keygen$:} & \\ \hline 
-{\bf pk:} & 1568 & {\bf $\encap$:} & & {\bf $\encap^+$:} & & {\bf $\encap^+$:} & \\ \hline 
-{\bf ct:} & 1568 & {\bf $\decap$:} & & {\bf $\decap^+$:} & & {\bf $\decap^+$:} & \\ \hline 
-\end{tabular}
-\caption{Cycle counts in AVX2 on a Coffee Lake laptop}
-\label{tab:ExpResults}
-\end{table}
-
-\begin{itemize}
-\item Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
-\item Look at OpenSSL for GMAC with no AES-256. 
-\item Also setup networking experiment on AWS
-\end{itemize}
+%TODO: Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
+%TODO: Look at OpenSSL for GMAC with no AES-256. 
+%TODO: Also setup networking experiment on AWS
 

### 2024-09-16 etm-transform/manuscript/main.tex
Stuff I need to do

@@
 %%%% 1. DOCUMENTCLASS %%%%
+%
+% TODO: Algo 9, 10, 11 should be one algorithms
+%   similarly agorithm 13-21, 22-24, 25-30
+% TODO: the implementation should be called ML-KEM+
+% TODO: Table 2 should be called ML-KEM+, as well
+% TODO: Table 3 should be broken up: pk/sk/keygen measurements should be one table, ct/encap/decap measurements should be one table for clearer comparison "
+% TODO: Table 4-6 need double lines to separate security level
+% TODO: need more thorough references
+% TODO: need better intro and conclusion (but later)
 % \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
@@
 Later works \cite{hofheinz2017modular} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals}, FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
-
-The current state of the FO transformation enjoys proven tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many deficiencies to be improved on.
-One such shortcoming is the use of re-encryption for providing ciphertext integrity \cite{bernstein2018towards}, which requires the decapsulation routine to run the encryption routine as a subroutine.
-In many post-quantum schemes, such as Kyber, the encryption routine is substantially more expensive than the decryption routine, so running the encryption routine in the decapsulation routine inflates the computational cost of the decapsulator.
-In addition, running the encryption as a subroutine introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key \cite{ravi2019generic}\cite{ueno2022curse}.
-
-The problem of ciphertext integrity was solved in symmetric cryptography: given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using ``encrypt-then-mac'' provides authenticated encryption \cite{bellare2000authenticated}.
-We took inspiration from this strategy and applied a similar technique to transform an OW-PCA secure public-key encryption scheme into an IND-CCA2 secure key encapsulation mechanism.
-Using a message authentication code for ciphertext integrity replaces the re-encryption step in decryption with the computation of an authenticator, which offers significant performance improvements while maintaining comparable level of security.
-
-The main challenge in applying ``encrypt-then-mac'' to public-key cryptography is the lack of a pre-shared symmetric key.
-We proposed to derive the symmetric key by hashing the plaintext message.
+% TODO: Mention and cite ML-KEM
+
+The current state of the FO transformation enjoys proven tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many open problems.
+One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} \cite{bernstein2018towards} for providing ciphertext integrity.
+In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine dominates the runtime cost of the decapsulation routine.
+While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.
+
+We are inspired by how ciphertext integrity is achieved in symmetric cryptography: given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using in a pattern called ``encrypt-then-MAC'' provides proven authenticated encryption \cite{bellare2000authenticated}.
+``encrypt-then-MAC'' is now the most widely accepted method for doing symmetric encryption as AES-GCM \cite{mcgrew2004galois} and ChaCha20-Poly1305 \cite{nir2018chacha20}.
+
+The main challenge in applying ``encrypt-then-MAC'' to public-key cryptography is the lack of a pre-shared symmetric key.
+We took inspiration from hybrid public-key encryption schemes (HPKE) such as .We proposed to derive the symmetric key by hashing the plaintext message.
 In section \ref{sec:main-results}, we prove that under the random oracle model, if the input public-key encryption scheme is one-way secure against plaintext-checking attack and the input message authentication code is one-time existentially unforgeable, then the transformed key encapsulation mechanism is IND-CCA2 secure.
 
-In section \ref{sec:experimental-results}, we instantiate concrete instances of our constructions by combining Kyber with GMAC and Poly1305.
+In section \ref{sec:implementation}, we instantiate concrete instances of our constructions by combining Kyber with GMAC and Poly1305.
 Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal runtime overhead in the encapsulation routine and a small increase in ciphertext size.

### 2024-09-17 etm-transform/manuscript/main.tex
define algorithms as short commands
Reformatted figures and tables

@@
 % \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
-\documentclass[floatrow,journal=tches,submission]{iacrtrans}
+\documentclass[journal=tches,submission]{iacrtrans}
 %%%% NOTES:
@@
 \usepackage{algpseudocode}
+\usepackage{caption}
 % \usepackage[demo]{graphicx}
@@
 
-\begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{
-                    The $\texttt{OW-ATK}$ game
-                }\label{alg:pke-ow-atk-game}
+\def\algowatkgame{
+    \begin{algorithm}[H]
+        \caption*{The $\texttt{OW-ATK}$ game}
         \begin{algorithmic}[1]
@@
             \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
-                    \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
+            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(
+                1^\lambda, \pk, c^\ast
+            )$
             \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
@@
 }
-        {\caption{The \texttt{OW-ATK} game}}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\pco(m \in \mathcal{M}, c \in \mathcal{C})$}\label{alg:pco}
+\def\algpco{
+    \begin{algorithm}[H]
+        \caption*{$\pco(m \in \mathcal{M}, c \in \mathcal{C})$}
         \begin{algorithmic}[1]
@@
 }
-        {\caption{Plaintext-checking oracle}}
-    \end{floatrow}
-    \label{fig:pco-and-cvo}
+\begin{figure}[H]
+    \centering
+    \begin{minipage}[b]{0.49\textwidth}
+        \algowatkgame
+    \end{minipage}
+    \begin{minipage}[b]{0.49\textwidth}
+        \algpco
+    \end{minipage}
+    \caption{
+        One-way security game of PKE (left) and plaintext-checking oracle (right)
+    }\label{fig:ow-game-pco}
 \end{figure}
 
-In the \texttt{OW-ATK} game (see figure \ref{alg:pke-ow-atk-game}), an adversary's goal is to recover the decryption of a randomly generated ciphertext.
+In the \texttt{OW-ATK} game (see figure \ref{fig:ow-game-pco}), an adversary's goal is to recover the decryption of a randomly generated ciphertext.
 A challenger randomly samples a keypair and a challenge plaintext $m^\ast$, encrypts the challenge plaintext $c^\ast \leftsample \encrypt(\pk, m^\ast)$, then gives $\pk$ and $c^\ast$ to the adversary $A$.
@@
 The capabilities of the oracle $\mathcal{O}_\texttt{ATK}$ depends on the choice of security goal $\texttt{ATK}$.
-Particularly relevant to our result is security against plaintext-checking attack (PCA), for which the adversary has access to a plaintext-checking oracle (PCO) (see figure \ref{alg:pco}).
+Particularly relevant to our result is security against plaintext-checking attack (PCA), for which the adversary has access to a plaintext-checking oracle (PCO) (see figure \ref{fig:ow-game-pco}).
 A PCO takes as input a plaintext-ciphertext pair $(m, c)$ and returns \texttt{True} if $m$ is the decryption of $c$ or \texttt{False} otherwise.
@@
 \begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{
-                    \texttt{IND-CCA} game for \kem
-                }\label{alg:kem-ind-cca-game}
-
+    \centering
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{KEM-IND-CCA2 game}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{The \texttt{KEM-IND-CCA2} game}}
-        \ffigbox{
-            \begin{algorithm}[H]
-            \caption{
-                $\mathcal{O}_\decap(c)$
-            }
+    \end{minipage}\hfill
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+        \caption*{$\mathcal{O}_\decap(c)$}
         \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{Decapsulation oracle}}
-    \end{floatrow}
+    \end{minipage}
+    \caption{IND-CCA2 game for KEM (left) and decapsulation oracle (right)}\label{fig:kem-game}
 \end{figure}
@@
 \begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox[250px]{
-            \begin{algorithm}[H]
-                \caption{The EUF-CMA game}\label{alg:mac-euf-cma-game}
+    \centering
+    \begin{minipage}{0.7\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{EUF-CMA} game}
 
@@
         \end{algorithm}
-        }{\caption{The \texttt{EUF-CMA} game}}
-    \end{floatrow}
+    \end{minipage}
+    \caption{The existential forgery game}\label{fig:euf-cma-game}
 \end{figure}
@@
     \centering
-    \begin{minipage}{0.40\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\keygen()$}\label{alg:fo-keygen}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk^\prime) \leftsample \keygen_\pke()$
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem\texttt{.}\keygen()$}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                 \State $z \leftsample \mathcal{M}$
@@
         \end{algorithm}
-    \end{minipage}
-    \hspace{0.05\textwidth}
-    \begin{minipage}{0.45\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\encap(\pk)$}\label{alg:fo-encap}
+        \begin{algorithm}[H]
+            \caption*{$\kem\texttt{.}\encap(\pk)$}
             \begin{algorithmic}[1]
@@
                 \State $r \leftarrow G(m)$
-                \State $c \leftarrow \encrypt(\pk, m, r)$
+                \State $c \leftarrow \pke\texttt{.}\encrypt(\pk, m, r)$
                 \State $K \leftarrow H(m, c)$
@@
     \end{minipage}
-    \vspace{0.5cm}
-    \begin{minipage}{\textwidth}
-        \begin{algorithm}[H]
-            \caption{$\decap(\sk=(\sk^\prime, z), c)$}\label{alg:fo-decap}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \decrypt(\sk^\prime, c)$
+    \begin{minipage}[t]{0.5\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem\texttt{.}\decap(\sk, c)$}
+            \begin{algorithmic}[1]
+                \Ensure $\sk = (\sk^\prime, z)$
+                \Ensure $\sk^\prime$ is valid PKE secret key
+                \Ensure $z$ is valid PKE plaintext
+                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c)$
                 \State $\hat{r} \leftarrow G(\hat{m})$
-                \State $\hat{c} \leftarrow \encrypt(\pk, \hat{m}, \hat{r})$
+                \State $\hat{c} \leftarrow \pke\texttt{.}\encrypt(\pk, \hat{m}, \hat{r})$
                 \If{$\hat{c} = c$}
@@
 \begin{figure}[H]
-    \center
-    \begin{algorithm}[H]
-        \caption{$\keygen_\etm$}\label{alg:etm-keygen}
-        \begin{algorithmic}[1]
-            \State $(\pk, \sk^\prime) \leftsample \keygen(1^\lambda)$
+    \centering
+    \begin{minipage}[t]{0.5\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem_\etm.\keygen()$}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                 \State $z \leftsample \mathcal{M}$
-            \State $\sk \leftarrow (\sk^\prime, z)$
+                \State $\sk \leftarrow \sk^\prime \Vert z$
                 \State \Return $(\pk, \sk)$
@@
         \begin{algorithm}[H]
-        \caption{$\encap(\pk)$}\label{alg:etm-encap}
-        \begin{algorithmic}[1]
+            \caption*{$\kem_\etm.\encap(\pk)$}
+            \begin{algorithmic}[1]
+                \Ensure $\pk$ is some $\pke$ public key
                 \State $m \leftsample \mathcal{M}$
                 \State $k \leftarrow G(m)$
-            \State $c^\prime \leftsample \encrypt(\pk, m)$
+                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                 \State $t \leftarrow \mac(k, c^\prime)$
                 \State $K \leftarrow H(m, c^\prime)$
-            \State $c \leftarrow (c^\prime, t)$
+                \State $c \leftarrow c^\prime \Vert t$
                 \State \Return $(c, K)$
@@
         \end{algorithm}
-    \begin{algorithm}[H]
-        \caption{$\decap(\sk, c)$}\label{alg:etm-decap}
-        \begin{algorithmic}[1]
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem_\etm.\decap(\sk, c)$}
+            \begin{algorithmic}[1]
+                \Require $c = c^\prime \Vert t, \sk = sk^\prime \Vert z$
+                \Ensure $c^\prime$ is some $\pke$ ciphertext
+                \Ensure $t$ is some $\mac$ tag
+                \Ensure $\sk^\prime$ is some $\pke$ secret key
+                \Ensure $z$ is some $\pke$ plaintext
                 \State $(c^\prime, t) \leftarrow c$
                 \State $(\sk^\prime, z) \leftarrow \sk$
-            \State $\hat{m} \leftarrow \decrypt(\sk^\prime, c^\prime)$
+                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                 \State $\hat{k} \leftarrow G(\hat{m})$
@@
         \end{algorithm}
+    \end{minipage}
     \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
@@
 \begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox[300px]{
-            \begin{algorithm}[H]
-                \caption{$\pco(m, c)$}\label{alg:pco-using-decap-oracle}
+    \centering
+    \begin{minipage}{0.6\textwidth}
+    \begin{algorithm}[H]
+        \caption*{$\pco(m, c)$}
         \begin{algorithmic}[1]
@@
     \end{algorithm}
-        }{
-            \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}
-            \label{fig:pco-using-decap-oracle}
-        }
-    \end{floatrow}
+    \end{minipage}
+    \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}\label{fig:pco-using-decap-oracle}
 \end{figure}
@@
 \begin{proof}
-    We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of games.
-
-    \begin{figure}[H]
-        \begin{floatrow}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{IND-CCA2 game for KEM}\label{alg:game-0}
-                    \begin{algorithmic}[1]
-                        \State $(\pk, \sk) \leftsample \keygen_\etm()$
+    We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of games. The sequence of games is summarized in figure \ref{fig:etm-ind-cc2-sequence-of-games}.
+
+    \begin{figure}[h]
+        \centering
+        \begin{minipage}[t]{0.5\textwidth}
+            \begin{algorithm}[H]
+                \caption*{\texttt{IND-CCA2} game for $\kem_\etm$}
+                \begin{algorithmic}[1]
+                    \State $(\pk, \sk) \leftsample \kem_\etm\texttt{.}\keygen()$
                     \State $m^\ast \leftsample \mathcal{M}$
-                        \State $c^\prime \leftsample \encrypt(\pk, m^\ast)$
+                    \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m^\ast)$
                     \State $k^\ast \leftarrow G(m^\ast)$
-                        \State $t \leftarrow \mac(k, c^\prime)$
-                        \State $c^\ast \leftarrow (c^\prime, t)$
+                        \Comment{Game 0-1}
+                    \State $k^\ast \leftsample \mathcal{K}_\mac$
+                        \Comment{Game 2-3}
+                    \State $t \leftarrow \mac(k^\ast, c^\prime)$
+                    \State $c^\ast \leftarrow c^\prime \Vert t$
                     \State $K_0 \leftarrow H(m^\ast, c^\prime)$
+                        \Comment{Game 0-2}
+                    \State $K_0 \leftsample \mathcal{K}_\kem$
+                        \Comment{Game 3}
                     \State $K_1 \leftsample \mathcal{K}_\kem$
@@
                     \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(\pk, c^\ast, K_b)$
+                        \Comment{Game 0}
+                    \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(\pk, c^\ast, K_b)$
+                        \Comment{Game 1-3}
                     \State \Return $\llbrack \hat{b} = b \rrbrack$
@@
             \end{algorithm}
-            }{}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^\decap(c)$}\label{alg:true-decap-oracle}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^G(m)$}
+                \begin{algorithmic}[1]
+                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
+                        \State \Return $\tilde{k}$
+                    \EndIf
+                    \State $k \leftsample \mathcal{K}_\mac$
+                    \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
+                    \State \Return $k$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \begin{minipage}[t]{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^\decap(c)$}
                 \begin{algorithmic}[1]
@@
             \end{algorithm}
-            }{}
-        \end{floatrow}
-        \begin{floatrow}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^G(m)$}\label{alg:hash-oracle-g}
-                    \begin{algorithmic}[1]
-                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
-                            \State \Return $\tilde{k}$
-                        \EndIf
-                        \State $k \leftsample \mathcal{K}_\mac$
-                        \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
-                        \State \Return $k$
-                    \end{algorithmic}
-                \end{algorithm}
-            }{}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^H(m, c)$}\label{alg:hash-oracle-h}
-                    \begin{algorithmic}[1]
-                        \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
-                            \State \Return $\tilde{K}$
-                        \EndIf
-                        \State $K \leftsample \mathcal{K}_\kem$
-                        \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
-                        \State \Return $K$
-                    \end{algorithmic}
-                \end{algorithm}
-            }{}
-        \end{floatrow}
-    \end{figure}
-
-    \emph{Game 0} is the standard IND-CCA2 game for KEMs.
-The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
-The queries made to the hash oracles $\mathcal{O}^G, \mathcal{O}^H$ are recorded to their respective tapes $\mathcal{L}^G, \mathcal{L}^H$.
-
-    \begin{figure}[H]
-        \begin{floatrow}
-            \ffigbox[270px]{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^\decap_1(c)$}\label{alg:sim-decap-oracle}
+            \vspace{-0.5cm}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^\decap_1(c)$}
                 \begin{algorithmic}[1]
@@
             \end{algorithm}
-            }{
-                \caption{Simulated decapsulation oracle}
-            }
-        \end{floatrow}
+            \vspace{-0.5cm}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^H(m, c)$}
+                \begin{algorithmic}[1]
+                    \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
+                        \State \Return $\tilde{K}$
+                    \EndIf
+                    \State $K \leftsample \mathcal{K}_\kem$
+                    \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \caption{Sequence of games}\label{fig:etm-ind-cc2-sequence-of-games}
     \end{figure}
+
+    \emph{Game 0} is the standard IND-CCA2 game for KEMs.
+The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
+The queries made to the hash oracles $\mathcal{O}^G, \mathcal{O}^H$ are recorded to their respective tapes $\mathcal{L}^G, \mathcal{L}^H$.
 
@@
 
-    \begin{figure}[H]
-        \begin{floatrow}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$B(\pk, {c^\prime}^\ast)$}\label{alg:ow-pca-adversary}
+    \begin{figure}[h]
+        \begin{minipage}[t]{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$B(\pk, {c^\prime}^\ast)$}
                 \begin{algorithmic}[1]
@@
             \end{algorithm}
-            }{}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^\decap_B(c)$}\label{alg:sim-decap-oracle-b}
-                    \begin{algorithmic}[1]
-                        \State $(c^\prime, t) \leftarrow c$
-                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
-                            % \tilde{m} = \decrypt(\sk^\prime, c^\prime)
-                            \pco(c^\prime, \tilde{m}) = 1
-                            \land \mac(\tilde{k}, c^\prime) = t
-                        $}
-                            \State $K \leftarrow H(\tilde{m}, c^\prime)$
-                        \Else
-                            \State $K \leftarrow H(z, c^\prime)$
-                        \EndIf
-                        \State \Return $K$
-                    \end{algorithmic}
-                \end{algorithm}
-            }{}
-        \end{floatrow}
-        \begin{floatrow}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^G_B(m)$}\label{alg:hash-oracle-g-b}
-                    \begin{algorithmic}[1]
-                        \If{$\pco(m, {c^\prime}^\ast) = 1$}
-                            \State $\texttt{ABORT}(m)$
-                        \EndIf
-                        \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
-                            \State \Return $\tilde{k}$
-                        \EndIf
-                        \State $k \leftsample \mathcal{K}_\mac$
-                        \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
-                        \State \Return $k$
-                    \end{algorithmic}
-                \end{algorithm}
-            }{}
-            \ffigbox{
-                \begin{algorithm}[H]
-                    \caption{$\mathcal{O}^H_B(m, c)$}\label{alg:hash-oracle-h-b}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^H_B(m, c)$}
                 \begin{algorithmic}
@@
             \end{algorithm}
-            }{}
-        \end{floatrow}
+        \end{minipage}\hfill
+        \begin{minipage}[t]{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^\decap_B(c)$}
+                \begin{algorithmic}[1]
+                    \State $(c^\prime, t) \leftarrow c$
+                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
+                        % \tilde{m} = \decrypt(\sk^\prime, c^\prime)
+                        \pco(c^\prime, \tilde{m}) = 1
+                        \land \mac(\tilde{k}, c^\prime) = t
+                    $}
+                        \State $K \leftarrow H(\tilde{m}, c^\prime)$
+                    \Else
+                        \State $K \leftarrow H(z, c^\prime)$
+                    \EndIf
+                    \State \Return $K$
+                \end{algorithmic}
+            \end{algorithm}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^G_B(m)$}
+                \begin{algorithmic}[1]
+                    \If{$\pco(m, {c^\prime}^\ast) = 1$}
+                        \State $\texttt{ABORT}(m)$
+                    \EndIf
+                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
+                        \State \Return $\tilde{k}$
+                    \EndIf
+                    \State $k \leftsample \mathcal{K}_\mac$
+                    \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
+                    \State \Return $k$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \caption{OW-PCA adversary $B$ simulates game 3 for IND-CCA2 adversary $A$}\label{fig:ow-pca-simulates-game-3}
     \end{figure}
+
     We will bound $P[\texttt{QUERY G}]$ and $P[\texttt{QUERY H}]$ by constructing a OW-PCA adversary $B$ against the underlying PKE that uses $A$ as a sub-routine.
-$B$'s behaviors are described in algorithms \ref{alg:ow-pca-adversary}, \ref{alg:sim-decap-oracle-b}, \ref{alg:hash-oracle-g-b}, and \ref{alg:hash-oracle-h-b}.
+$B$'s behaviors are summarized in figure \ref{fig:ow-pca-simulates-game-3}.
 
@@
 
-A detailed description of $\texttt{K-PKE}$'s routines can be found in FIPS 203 (TODO: citation).
-The ``encrypt-then-MAC'' routines are listed in algorithms \ref{alg:ml-kem-etm-keygen}, \ref{alg:ml-kem-etm-encap}, and \ref{alg:ml-kem-etm-decap}.
-
-\begin{algorithm}[H]
-    \caption{\texttt{ML-KEM-ETM.KeyGen()}}\label{alg:ml-kem-etm-keygen}
+A detailed description of $\texttt{K-PKE}$'s routines can be found in FIPS 203 (TODO: citation). The ``encrypt-then-MAC'' routines are described in figure \ref{fig:ml-kem-plus-routines}.
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.5\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\keygen\texttt{()}$}
             \begin{algorithmic}[1]
@@
                 \State $h \leftarrow H(\pk)$
-            \Comment{$H$ is \texttt{SHA3-256}}
                 \State $\sk \leftarrow (sk^\prime \Vert \pk \Vert h \Vert z)$
@@
             \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[H]
-    \caption{\texttt{ML-KEM-ETM.Encap}(\pk)}\label{alg:ml-kem-etm-encap}
+        \end{algorithm}\vspace{-0.5cm}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\encap\texttt{(\pk)}$}
             \begin{algorithmic}[1]
@@
                 \State $(\overline{K}, r, k) \leftarrow \texttt{XOF}(m \Vert H(\pk))$
-            \Comment{\texttt{XOF} is \texttt{Shake256}, outputting 768 bits}
                 \State $c^\prime \leftarrow \texttt{K-PKE}.\encrypt(\pk, m, r)$
@@
                 \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
-            \Comment{\texttt{KDF} is \texttt{Shake256}, outputting 256 bits}
                 \State $c \leftarrow (c^\prime, c)$
@@
         \end{algorithm}
-
-\begin{algorithm}[H]
-    \caption{$\texttt{ML-KEM-ETM.Decap}(\sk, c)$}\label{alg:ml-kem-etm-decap}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\decap\texttt{(\sk, c)}$}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
+    \end{minipage}
+
+    \caption{$\texttt{ML-KEM}^+$ routines}\label{fig:ml-kem-plus-routines}
+\end{figure}
 
@@
 \begin{table}[H]
+    \centering
     \caption{MAC performance}\label{tbl:mac-performance}
@@
 \begin{table}
+    \centering
     \caption{CPU cycles of each KEM routine}\label{tbl:kem-performance}
@@
 The client finally decapsulates the ciphertext to recover the shared secret $K_e \leftarrow \decap(\sk_e, c_e)$.
-The specific steps are described in algorithms \ref{alg:ke-client}, \ref{alg:ke-server}.
+The key exchange routines are summarized in figure \ref{fig:ke-routines}.
 
@@
 \begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\texttt{KE}_\texttt{C}$}\label{alg:ke-client}
+    \centering
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{KE}_\texttt{C}()$}\label{alg:ke-client}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{KE Client}}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\texttt{KE}_\texttt{S}$}\label{alg:ke-server}
+    \end{minipage}\hfill
+    \begin{minipage}[b]{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$\texttt{KE}_\texttt{S}()$}\label{alg:ke-server}
                 \begin{algorithmic}[1]
@@
             \end{algorithm}
-        }{\caption{KE Server}}
-    \end{floatrow}
+    \end{minipage}
+
+    \caption{Unauthenticated key exchange (KE) routines}\label{fig:ke-routines}
 \end{figure}
@@
 \begin{table}[H]
+    \centering
     \caption{Unauthenticated key exchange RTT comparison}\label{tbl:ke-rtt}
@@
 In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by demonstrating possession of the corresponding long-term secret key $\sk_S$.
-Details are described in algorithms \ref{alg:uake-client} and \ref{alg:uake-server}.
+UAKE routines are summarized in figure \ref{fig:uake-routines}.
 
@@
 \begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\texttt{UAKE}_\texttt{C}(\pk_S)$}\label{alg:uake-client}
+    \centering
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{UAKE}_\texttt{C}(\pk_S)$}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{UAKE Client}}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\texttt{UAKE}_\texttt{S}(\sk_S)$}\label{alg:uake-server}
+    \end{minipage}\hfill
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{UAKE}_\texttt{S}(\sk_S)$}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{UAKE Server}}
-    \end{floatrow}
+    \end{minipage}
+
+    \caption{Unilaterally authenticated key exchange (UAKE) routines}\label{fig:uake-routines}
 \end{figure}
@@
 \begin{table}[H]
+    \centering
     \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
@@
 \begin{figure}[H]
-    \begin{floatrow}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\texttt{AKE}_\texttt{C}(\pk_S, \sk_C)$}\label{alg:ake-client}
+    \centering
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{AKE}_\texttt{C}(\pk_S, \sk_C)$}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{AKE Client}}
-        \ffigbox{
-            \begin{algorithm}[H]
-                \caption{$\texttt{AKE}_\texttt{S}(\sk_S, \pk_C)$}\label{alg:akex-server}
+    \end{minipage}
+    \begin{minipage}[b]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{AKE}_\texttt{S}(\sk_S, \pk_C)$}
             \begin{algorithmic}[1]
@@
         \end{algorithm}
-        }{\caption{AKE Server}}
-    \end{floatrow}
+    \end{minipage}
+
+    \caption{Mutually authenticated key exchange (AKE) routines}\label{fig:ake-routines}
 \end{figure}
@@
 \begin{table}[H]
+    \centering
     \caption{AKE RTT comparison}\label{tbl:ake-rtt}

### 2024-09-18 etm-transform/manuscript/main.tex
Prettier formatting for KEM routine comparisons

@@
 \usepackage{caption}
+\usepackage{multirow}
 % \usepackage[demo]{graphicx}
@@
 
-\begin{table}
-    \centering
+\begin{table}[h]
+    \centering
+    \footnotesize
     \caption{CPU cycles of each KEM routine}\label{tbl:kem-performance}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|}
-        \hline\hline
-        \multicolumn{8}{|c|}{\bf ML-KEM-512} \\
-        \hline
-        \multicolumn{2}{|c|}{public key size} 
-        & \multicolumn{2}{|c|}{800 bytes}
-        & \multicolumn{4}{|c|}{ \keygen} \\
-        \hline
-        \multicolumn{2}{|c|}{secret key size} 
-        & \multicolumn{2}{|c|}{1632 bytes} 
-        & \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{75945} \\
-        \hline
-        \multicolumn{2}{|c|}{ciphertext size} 
-        & \multicolumn{2}{|c|}{768 bytes} 
-        & \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{76171} \\
-        \hline
-        \multicolumn{4}{|c|}{ \encap}
-        & \multicolumn{4}{|c|}{\decap} \\
-        \hline
-        \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{91467} 
-        & \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{121185} \\
-        \hline
-        \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{92065} 
-        & \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{121650} \\
-        \hline
-        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + Poly1305}
-        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + GMAC} \\
-        \hline
-        \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$}
-        & \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$} \\
-        \hline
-        Median & 93157 & Median & 33773 
-        & Median & 97369 & Median & 37725 \\
-        \hline
-        Average & 93626 & Average & 33908 
-        & Average & 97766 & Average & 37831 \\
-        \hline
-        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + CMAC}
-        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-512 + KMAC} \\
-        \hline
-        \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$}
-        & \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$} \\
-        \hline
-        Median & 99839 & Median & 40117 & Median & 101009 & Median & 40741 \\
-        \hline
-        Average & 99959 & Average & 39943 & Average & 101313 & Average & 40916 \\
-        \hline\hline
-        \multicolumn{8}{|c|}{\bf ML-KEM-768} \\
-        \hline
-        \multicolumn{2}{|c|}{public key size} 
-        & \multicolumn{2}{|c|}{1184 bytes}
-        & \multicolumn{4}{|c|}{ \keygen} \\
-        \hline
-        \multicolumn{2}{|c|}{secret key size} 
-        & \multicolumn{2}{|c|}{2400 bytes} 
-        & \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{129895} \\
-        \hline
-        \multicolumn{2}{|c|}{ciphertext size} 
-        & \multicolumn{2}{|c|}{1088 bytes} 
-        & \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{130650} \\
-        \hline
-        \multicolumn{4}{|c|}{ \encap}
-        & \multicolumn{4}{|c|}{\decap} \\
-        \hline
-        \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{146405} 
-        & \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{186445} \\
-        \hline
-        \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{147400} 
-        & \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{187529} \\
-        \hline
-        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + Poly1305}
-        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + GMAC} \\
-        \hline
-        \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$}
-        & \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$} \\
-        \hline
-        Median & 146405 & Median & 43315 
-        & Median & 149525 & Median & 46513 \\
-        \hline
-        Average & 146860 & Average & 43463 
-        & Average & 150128 & Average & 46706 \\
-        \hline
-        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + CMAC}
-        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-768 + KMAC} \\
-        \hline
-        \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$}
-        & \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$} \\
-        \hline
-        Median & 153139 & Median & 49841 & Median & 155219 & Median & 52415 \\
-        \hline
-        Average & 153735 & Average & 50074 & Average & 155848 & Average & 52611 \\
-        \hline\hline
-        \multicolumn{8}{|c|}{\bf ML-KEM-1024} \\
-        \hline
-        \multicolumn{2}{|c|}{public key size} 
-        & \multicolumn{2}{|c|}{1568 bytes}
-        & \multicolumn{4}{|c|}{ \keygen} \\
-        \hline
-        \multicolumn{2}{|c|}{secret key size} 
-        & \multicolumn{2}{|c|}{3168 bytes} 
-        & \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{194921} \\
-        \hline
-        \multicolumn{2}{|c|}{ciphertext size} 
-        & \multicolumn{2}{|c|}{1568 bytes} 
-        & \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{195465} \\
-        \hline
-        \multicolumn{4}{|c|}{ \encap}
-        & \multicolumn{4}{|c|}{\decap} \\
-        \hline
-        \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{199185} 
-        & \multicolumn{2}{|c|}{Median} 
-        & \multicolumn{2}{|c|}{246245} \\
-        \hline
-        \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{199903} 
-        & \multicolumn{2}{|c|}{Average} 
-        & \multicolumn{2}{|c|}{247320} \\
-        \hline
-        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + Poly1305}
-        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + GMAC} \\
-        \hline
-        \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$}
-        & \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$} \\
-        \hline
-        Median & 205763 & Median & 51375 
-        & Median & 208805 & Median & 54573 \\
-        \hline
-        Average & 206499 & Average & 51562 
-        & Average & 209681 & Average & 54780 \\
-        \hline
-        \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + CMAC}
-        & \multicolumn{4}{|c|}{\bf ML-KEM-ETM-1024 + KMAC} \\
-        \hline
-        \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$}
-        & \multicolumn{2}{|c|}{ $\encap$}
-        & \multicolumn{2}{|c|}{$\decap$} \\
-        \hline
-        Median & 213667 & Median & 59175 & Median & 216761 & Median & 62269 \\
-        \hline
-        Average & 214483 & Average & 59408 & Average & 217468 & Average & 62516 \\
-        \hline\hline
+
+    \begin{tabular}[t]{|cc|}
+        \hline
+        \multicolumn{2}{|c|}{\bf ML-KEM-512} \\
+        \hline
+        \multicolumn{2}{|c|}{size parameters (bytes)} \\
+        pubkey size & 800 \\
+        seckey size & 1632 \\
+        ciphertext size & 768 \\
+        \hline
+        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+        Median & 75945 \\
+        Average & 76171 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}[t]{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant} 
+        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+        \cline{2-5}
+        & Median & Average & Median & Average \\
+        \hline
+        \texttt{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
+        \hline
+        \texttt{Kyber512} & ??? & ??? & ??? & ??? \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+
+    \begin{tabular}[t]{|cc|}
+        \hline
+        \multicolumn{2}{|c|}{\bf ML-KEM-768} \\
+        \hline
+        \multicolumn{2}{|c|}{size parameters (bytes)} \\
+        pubkey size & 1184 \\
+        seckey size & 2400 \\
+        ciphertext size & 1088 \\
+        \hline
+        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+        Median & 129895 \\
+        Average & 130650 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}[t]{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant} 
+        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+        \cline{2-5}
+        & Median & Average & Median & Average \\
+        \hline
+        \texttt{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
+        \hline
+        \texttt{Kyber768} & ??? & ??? & ??? & ??? \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+
+    \begin{tabular}[t]{|cc|}
+        \hline
+        \multicolumn{2}{|c|}{\bf ML-KEM-1024} \\
+        \hline
+        \multicolumn{2}{|c|}{size parameters (bytes)} \\
+        pubkey size & 1568 \\
+        seckey size & 3168 \\
+        ciphertext size & 1568 \\
+        \hline
+        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+        Median & 194921 \\
+        Average & 195465 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}[t]{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant} 
+        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+        \cline{2-5}
+        & Median & Average & Median & Average \\
+        \hline
+        \texttt{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
+        \hline
+        \texttt{Kyber1024} & ??? & ??? & ??? & ??? \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
+        \hline
     \end{tabular}

### 2024-09-18 etm-transform/manuscript/main.tex
Improved KEX RTT comparison table
updated with Kyber round 3 data

@@
 %
-% TODO: Algo 9, 10, 11 should be one algorithms
-%   similarly agorithm 13-21, 22-24, 25-30
-% TODO: the implementation should be called ML-KEM+
-% TODO: Table 2 should be called ML-KEM+, as well
-% TODO: Table 3 should be broken up: pk/sk/keygen measurements should be one table, ct/encap/decap measurements should be one table for clearer comparison "
-% TODO: Table 4-6 need double lines to separate security level
 % TODO: need more thorough references
@@
 \subsection{Choosing a message authenticator}
-When instantiating \texttt{ML-KEM-ETM}, there are a variety of message authentication codes to choose from. We selected four instances covering a wide range of designs: \begin{enumerate}
-    \item Poly1305, a Carter-Wegman style MAC operating on a prime field
-    \item GMAC (AES-256-GCM), a Carter-Wegman style MAC operating on a binary field
-    \item CMAC (AES-256-CBC), a CBC-MAC
-    \item KMAC-256, a keyed hash function based on Keccak
-\end{enumerate}
-
-We tested each MAC's throughput by measuring the CPU cycles needed to compute a tag on the ciphertext returned by \texttt{K-PKE}.
-The length of the ciphertext varies depending on the security level.
-The measurements are summarized in table \ref{tbl:mac-performance}.
+For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{kelsey2016sha}.
+
+% TODO: What is Poly1305
+
+% TODO: What is GMAC
+
+% TODO: What is CMAC
+
+% TODO: What is KMAC
+
+% TODO: We benchmarked MAC performance computing tags on various sizes of ML-KEM ciphertexts
 
@@
 
+% TODO: How is ML-KEM different from Kyber round 3?
+
 \begin{table}[h]
@@
         \hline
-        \texttt{Kyber512} & ??? & ??? & ??? & ??? \\
+        \texttt{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
         \hline
@@
         \hline
-        \texttt{Kyber768} & ??? & ??? & ??? & ??? \\
+        \texttt{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
         \hline
@@
         \hline
-        \texttt{Kyber1024} & ??? & ??? & ??? & ??? \\
+        \texttt{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
         \hline
@@
     \centering
-    \caption{Unauthenticated key exchange RTT comparison}\label{tbl:ke-rtt}
-    \begin{tabular}{|c|c|c|}
-        \hline
-        KEM & median RTT ($\us$) & average RTT ($\us$) \\
-        \hline
-        ML-KEM-512 & 92 & 97 \\
-        \hline
-        ML-KEM-512 + Poly1305 & 70 & 72 \\
-        \hline
-        ML-KEM-512 + GMAC & 73 & 76 \\
-        \hline
-        ML-KEM-512 + CMAC & 75 & 79 \\
-        \hline
-        ML-KEM-512 + KMAC & 76 & 78 \\
-        \hline 
-        ML-KEM-768 & 135 & 140 \\
-        \hline
-        ML-KEM-768 + Poly1305 & 99 & 104 \\
-        \hline
-        ML-KEM-768 + GMAC & 101 & 105 \\
-        \hline
-        ML-KEM-768 + CMAC & 103 & 109 \\
-        \hline
-        ML-KEM-768 + KMAC & 103 & 107 \\
-        \hline
-        ML-KEM-1024 & 193 & 199 \\
-        \hline
-        ML-KEM-1024 + Poly1305 & 138 & 141 \\
-        \hline
-        ML-KEM-1024 + GMAC & 140 & 145 \\
-        \hline
-        ML-KEM-1024 + CMAC & 143 & 148 \\
-        \hline
-        ML-KEM-1024 + KMAC & 144 & 149 \\
+    \small
+    \caption{KE RTT comparison}\label{tbl:ke-rtt}
+
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-512} & 800 & 768 & 92 & 97 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 800 & 784 & 70 & 72 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ GMAC & 800 & 784 & 73 & 76 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ CMAC & 800 & 784 & 75 & 79 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ KMAC & 800 & 784 & 76 & 78 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-768} & 1184 & 1088 & 135 & 140 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 1184 & 1104 & 99 & 104 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ GMAC & 1184 & 1104 & 101 & 105 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ CMAC & 1184 & 1104 & 103 & 109 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ KMAC & 1184 & 1104 & 103 & 107 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-1024} & 1568 & 1568 & 193 & 199 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 1568 & 1584 & 138 & 141 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 1568 & 1584 & 140 & 145 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 1568 & 1584 & 143 & 148 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 1568 & 1584 & 144 & 149 \\
         \hline
@@
     \centering
+    \small
     \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
-    \begin{tabular}{|c|c|c|}
-        \hline
-        KEM & median RTT ($\us$) & average RTT ($\us$) \\
-        \hline
-        ML-KEM-512 & 145 & 151 \\
-        \hline
-        ML-KEM-512 + Poly1305 & 103 & 106 \\
-        \hline
-        ML-KEM-512 + GMAC & 106 & 110 \\
-        \hline
-        ML-KEM-512 + CMAC & 108 & 112 \\
-        \hline
-        ML-KEM-512 + KMAC & 109 & 113 \\
-        \hline 
-        ML-KEM-768 & 215 & 222 \\
-        \hline
-        ML-KEM-768 + Poly1305 & 144 & 150 \\
-        \hline
-        ML-KEM-768 + GMAC & 149 & 156 \\
-        \hline
-        ML-KEM-768 + CMAC & 153 & 160 \\
-        \hline
-        ML-KEM-768 + KMAC & 154 & 159 \\
-        \hline
-        ML-KEM-1024 & 310 & 318 \\
-        \hline
-        ML-KEM-1024 + Poly1305 & 202 & 209 \\
-        \hline
-        ML-KEM-1024 + GMAC & 212 & 228 \\
-        \hline
-        ML-KEM-1024 + CMAC & 212 & 218 \\
-        \hline
-        ML-KEM-1024 + KMAC & 213 & 220 \\
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-512} & 1568 & 768 & 145 & 151 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 784 & 103 & 106 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 784 & 106 & 110 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 784 & 108 & 112 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 784 & 109 & 113 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-768} & 2272 & 1088 & 215 & 222 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 1104 & 149 & 156 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 1104 & 153 & 160 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 1104 & 154 & 159 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-1024} & 3136 & 1568 & 310 & 318 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 1584 & 212 & 228 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 1584 & 212 & 218 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 1584 & 213 & 220 \\
         \hline
@@
     \centering
+    \small
     \caption{AKE RTT comparison}\label{tbl:ake-rtt}
-    \begin{tabular}{|c|c|c|}
-        \hline
-        KEM & median RTT ($\us$) & average RTT ($\us$) \\
-        \hline
-        ML-KEM-512 & 200 & 213 \\
-        \hline
-        ML-KEM-512 + Poly1305 & 133 & 138 \\
-        \hline
-        ML-KEM-512 + GMAC & 139 & 143 \\
-        \hline
-        ML-KEM-512 + CMAC & 143 & 148 \\
-        \hline
-        ML-KEM-512 + KMAC & 145 & 151 \\
-        \hline 
-        ML-KEM-768 & 294 & 301 \\
-        \hline
-        ML-KEM-768 + Poly1305 & 190 & 196 \\
-        \hline
-        ML-KEM-768 + GMAC & 197 & 210 \\
-        \hline
-        ML-KEM-768 + CMAC & 202 & 208 \\
-        \hline
-        ML-KEM-768 + KMAC & 204 & 210 \\
-        \hline
-        ML-KEM-1024 & 512 & 511 \\
-        \hline
-        ML-KEM-1024 + Poly1305 & 266 & 273 \\
-        \hline
-        ML-KEM-1024 + GMAC & 273 & 282 \\
-        \hline
-        ML-KEM-1024 + CMAC & 280 & 287 \\
-        \hline
-        ML-KEM-1024 + KMAC & 282 & 288 \\
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-512} & 1568 & 1536 & 220 & 213 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 1568 & 133 & 138 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 1568 & 139 & 143 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 1568 & 143 & 148 \\
+        \hline
+        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 1568 & 145 & 151 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-768} & 2272 & 2176 & 294 & 301 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 2208 & 197 & 210 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 2208 & 202 & 208 \\
+        \hline
+        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 2208 & 204 & 210 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        \texttt{ML-KEM-1024} & 3136 & 3136 & 512 & 511 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 3168 & 273 & 282 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 3168 & 280 & 287 \\
+        \hline
+        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 3168 & 282 & 288 \\
         \hline
@@
 %TODO: Look at OpenSSL for GMAC with no AES-256.
-%TODO: Also setup networking experiment on AWS
 

### 2024-09-19 etm-transform/manuscript/main.tex
Started some explanation of various kinds of MAC

@@
 
-We are specifically interested in one-time MAC, whose security goal is identical to EUF-CMA described above, except for the constraint that each secret key can be used to sign exactly one distinct message.
-This translates to an attack model in which the signing oracle will only answer one signing query.
-Restricting to one-time usage allows for more efficient MAC constructions.
-One popular way to build one-time MAC is with universal hash functions (UHF), where each instance is parameterized by a finite field $\mathbb{F}$ and a maximal message length $L \geq 0$.
-The secret key is a pair of field elements $(k_1, k_2) \in \mathbb{F} \times \mathbb{F}$, and each message is a tuple of up to $L$ field elements $m = (m_1, m_2, \ldots, m_l) \in \mathbb{F}^{\leq L}$.
-To compute the tag:
-
-\begin{equation*}
-    \mac((k_1, k_2), m) = H_\text{xpoly}(k_1, m) + k_2
-\end{equation*}
-
-Where the $H_\text{xpoly}$ is a universal hash function:
-
-\begin{equation*}
-    H_\text{xpoly}(k_1, (m_1, m_2, \ldots, m_l)) = k_1^l \cdot m_1 + k_1^{l-1} \cdot m_2 + \ldots + k_1 \cdot m_l
-\end{equation*}
-
-\begin{lemma}
-    For all adversaries (including unbounded ones) against the MAC described above, the probability of winning the one-time EUF-CMA game is at most:
-
-    \begin{equation*}
-        \adv_\text{OT-EUF-CMA}(A) \leq \frac{L+1}{\norm{\mathbb{F}}}
-    \end{equation*}
-\end{lemma}
-\begin{proof}
-    See \cite{boneh2020graduate} lemma 7.11
-    % TODO: need more detailed citations
-\end{proof}
-
 \subsection{Related works}
@@
 However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed.
-This means that a decapsulation oracle can be converted into a plaintext-checking oracle (algorithm \ref{alg:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
+This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
 
@@
 \subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
-% TODO: write some intuition first?
-\begin{proof}
-    We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of games. The sequence of games is summarized in figure \ref{fig:etm-ind-cc2-sequence-of-games}.
+We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game.
+A summary of the the sequence of games can be found in figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
+From a high level we made three incremental modifications to the IND-CCA2 game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
+A OW-PCA adversary can then simulate the modified IND-CCA2 game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.
 
@@
 
+\begin{proof}
     \emph{Game 0} is the standard IND-CCA2 game for KEMs.
@@
 
-    \emph{Game 2} is identical to game 1, except that on line 4 of algorithm \ref{alg:game-0}, the challenger samples a uniformly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$ instead of deriving it from $m$.
+    \emph{Game 2} is identical to game 1, except that the challenger samples a uniformly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$ instead of deriving it from $m^\ast$.
 From $A$'s perspective the two games are indistinguishable, unless $A$ queries $G$ with the value of $m^\ast$.
@@
 
-    \emph{Game 3} is identical to game 2, except that on line 7 of algorithm \ref{alg:game-0}, the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $c^\prime$.
+    \emph{Game 3} is identical to game 2, except that the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $c^\prime$.
 From $A$'s perspective the two games are indistinguishable, unless $A$ queries $H$ with $(m^\ast, \cdot)$.
@@
 
-% TODO: What is Poly1305
-
-% TODO: What is GMAC
-
-% TODO: What is CMAC
+Poly1305 and GMAC are both Carter-Wegman style authenticators that compute the tag using finite field arithmetic.
+Generically speaking, Carter-Wegman MAC is parameterized by some finite field $\mathbb{F}$ and the maximal message length $L > 0$.
+Each symmetric key $k = (k_1, k_2) \leftsample \mathbb{F}^2$ is a pair of uniformly ranodm field elements, and the message is parsed into tuples of field elements up to length $L$: $m = (m_1, m_2, \ldots, m_l) \in \mathbb{F}^{\leq L}$.
+The tag $t$ is computed by evaluating a polynomial whose coefficients the message blocks and whose indeterminate is the key:
+
+\begin{equation}\label{eq:carter-wegman-one-time-mac}
+    \mac((k_1, k_2), m) = H_\text{xpoly}(k_1, m) + k_2
+\end{equation}
+
+Where $H_\text{xpoly}$ is given by:
+
+\begin{equation*}
+    H_\text{xpoly}(k_1, m) = k_1^{l+1} + k_1^l \cdot m_1 + k_1^{l-1} \cdot m_2 + \ldots + k_1 \cdot m_l
+\end{equation*}
+
+The authenticator formulated in equation \ref{eq:carter-wegman-one-time-mac} is a one-time MAC. To make the construction many-time secure, a non-repeating nonce $r$ and a PRF is needed:
+
+\begin{equation*}
+    \mac((k_1, k_2), m, r) = H_\text{xpoly}(k_1, m) \oplus \texttt{PRF}(k_2, r)
+\end{equation*}
+
+Specifically, Poly1035 operates in the prime field $\mathbb{F}_q$ where $q = 2^{130} - 5$ whereas GMAC operates in the binary field $\mathbb{F}_{2^{128}}$.
+In OpenSSL's implementation, standalone Poly1305 is a one-time secure MAC, whereas GMAC uses a nonce and AES as the PRF and is thus many-time secure (in OpenSSL GMAC is AES-256-GCM except all data is fed into the ``associated data'' section and thus not encrypted).
+
+The CMAC is based on the CBC-MAC, where the block chosen is chosen to be AES-256.
+To compute a CMAC tag, the message is first broke into 128-bit blocks with appropriate padding.
+Each block is first XOR'd with the previous block's output, then encrypted under AES using the symmetric key.
+The final output is XOR'd with a sub key derived from the symmetric key, before being encrypted for one last time.
+A summary of the computation can be found in figure \ref{fig:aes-256-cmac}
+
+\begin{figure}[h]
+    \centering
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{Sub-key derivation}
+            \begin{algorithmic}[1]
+                \Require 256-bit key $k$
+                \Require $\texttt{const\_Rb = 0x87}$
+                \State $l \leftarrow \texttt{AES-256}(k, 0^{128})$
+                \If{$\texttt{MostSignificantBit}(l) = 0$}
+                    \State $k_1 \leftarrow \texttt{l {<}{<} 1}$
+                \Else 
+                    \State $k_1 \leftarrow \texttt{l {<}{<} 1} \oplus \texttt{const\_Rb}$
+                \EndIf
+                \If{$\texttt{MostSignificantBit}(k_1) = 0$}
+                    \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1}$
+                \Else 
+                    \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1} \oplus \texttt{const\_Rb}$
+                \EndIf
+                \State \Return $k_1, k_2$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{CMAC(k, m)}
+            \begin{algorithmic}[1]
+                \Require 256-bit symmetric key $k$
+                \State $(k_1, k_2) \leftarrow \texttt{deriveSubKey}(k)$
+                \State $n \leftarrow \left\lceil \texttt{bytesLen}(m) / 16\right\rceil$
+                \If{$n = 0$}
+                    \State $n \leftarrow 1$
+                    \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
+                \ElsIf{$\texttt{bytesLen(m)} \mod 16 = 0$}
+                    \State $m_\texttt{last} \leftarrow m_n \oplus k_1$
+                \Else 
+                    \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
+                \EndIf
+
+                \State $x = 0^{128}$
+                \For{$i \in \{1, 2, \ldots, n-1\}$}
+                    \State $y \leftarrow x \oplus m_i$
+                    \State $x \leftarrow \texttt{AES-256}(k, y)$
+                \EndFor
+                \State $y \leftarrow m_\texttt{last} \oplus x$
+                \State $t \leftarrow \texttt{AES-256}(k, y)$
+                \State \Return $t$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{AES-256 CMAC}\label{fig:aes-256-cmac}
+\end{figure}
 

### 2024-09-23 etm-transform/manuscript/main.tex
Draft rewrite of the MAC performance section

@@
 \section{Implementation}\label{sec:implementation}
-Originally known as Kyber \cite{bos2018crystals}\cite{avanzi2019crystals}, ML-KEM is an \texttt{IND-CCA2} secure key encapsulation mechanism standardized in FIPS 203 by NIST.
+ML-KEM is an \texttt{IND-CCA2} secure key encapsulation mechanism standardized by NIST in FIPS 203.
 The \texttt{IND-CCA2} security of ML-KEM is achieved in two steps.
 First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries.
-Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}.\keygen, \texttt{K-PKE}.\encrypt, \texttt{K-PKE}.\decrypt$ as sub-routines.
-Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine suffers from computational inefficiency.
+Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines.
+Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.
+Also worth noting is the modification to CRYSTALS-Kyber (as presented in its third round submission to NIST's PQC) before it is standardized: Kyber constructed its KEM routines using the $U^{\not\bot}$ variant of the Fujisaki-Okamoto transform, where the shared secret is derived from hashing both the plaintext and the ciphertext; on the other hand, ML-KEM used the $U^{\not\bot}_m$ variant, where the shared secret is derived from hasing the plaintext alone.
 
 We implemented the ``encrypt-then-MAC'' KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios.
-The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
+The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment
 
@@
 
-The CMAC is based on the CBC-MAC, where the block chosen is chosen to be AES-256.
+CMAC is based on the CBC-MAC with the block cipher instantiated from AES-256.
 To compute a CMAC tag, the message is first broke into 128-bit blocks with appropriate padding.
@@
 
-% TODO: What is KMAC
-
-% TODO: We benchmarked MAC performance computing tags on various sizes of ML-KEM ciphertexts
+KMAC is defined in NIST SP 800-185 to be based on the family of sponge functions with Keccak permutaiton as the underlying function.
+We chose KMAC-256, which uses Shake256 as the underlying extendable output functions.
+KMAC allows variable-length key and tag, but we chose the 256 bits for key length and 128 bits for tag size for consistency with other authenticators.
+
+To isolate the performance characteristics of each authenticator in our instantiation of $\texttt{ML-KEM}^+$, we measured the CPU cycles needed for each authenticator to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of \texttt{ML-KEM}.
+The measurements are summarized in table \ref{tbl:mac-performance}.
 
@@
     \centering
-    \caption{MAC performance}\label{tbl:mac-performance}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|}
-        \hline
-        \multicolumn{8}{|c|}{ML-KEM-512} \\
-        \hline
-        \multicolumn{4}{|c|}{Ciphertext size (bytes): }
-        & \multicolumn{4}{|c|}{768} \\
-        \hline
-        \multicolumn{2}{|c|}{Poly1305} 
-        & \multicolumn{2}{|c|}{GMAC} 
-        & \multicolumn{2}{|c|}{CMAC} 
-        & \multicolumn{2}{|c|}{KMAC} \\
-        \hline
-        Median & 909 
-        & Median & 3899 
-        & Median & 6291 
-        & Median & 6373\\
-        \hline
-        Average & 2823 
-        & Average & 4859 
-        & Average & 6373 
-        & Average & 7791 \\
-        \hline
-        \hline
-        \multicolumn{8}{|c|}{ML-KEM-768} \\
-        \hline
-        \multicolumn{4}{|c|}{Ciphertext size (bytes): }
-        & \multicolumn{4}{|c|}{1088} \\
-        \hline
-        \multicolumn{2}{|c|}{Poly1305} 
-        & \multicolumn{2}{|c|}{GMAC} 
-        & \multicolumn{2}{|c|}{CMAC} 
-        & \multicolumn{2}{|c|}{KMAC} \\
-        \hline
-        Median & 961 
-        & Median & 3899 
-        & Median & 7305 
-        & Median & 9697 \\
-        \hline
-        Average & 2704 
-        & Average & 4827 
-        & Average & 7588 
-        & Average & 9928 \\
-        \hline
-        \hline
-        \multicolumn{8}{|c|}{ML-KEM-1024} \\
-        \hline
-        \multicolumn{4}{|c|}{Ciphertext size (bytes): }
-        & \multicolumn{4}{|c|}{1568} \\
-        \hline
-        \multicolumn{2}{|c|}{Poly1305} 
-        & \multicolumn{2}{|c|}{GMAC} 
-        & \multicolumn{2}{|c|}{CMAC} 
-        & \multicolumn{2}{|c|}{KMAC} \\
-        \hline
-        Median & 1065 
-        & Median & 4055 
-        & Median & 8735 
-        & Median & 11647 \\
-        \hline
-        Average & 1809 
-        & Average & 5026 
-        & Average & 8772 
-        & Average & 12186 \\
+    \footnotesize
+    \caption{CPU cycles needed to compute tag on various input sizes}\label{tbl:mac-performance}
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \multicolumn{3}{|c|}{\bf Input size: 768 bytes} \\
+        \hline
+        MAC & Median & Average \\
+        \hline
+        Poly1305 & 909 & 2823 \\
+        \hline
+        GMAC & 3899 & 4859 \\
+        \hline
+        CMAC & 6291 & 6373 \\
+        \hline
+        KMAC & 6373 & 7791 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \multicolumn{3}{|c|}{\bf Input size: 1088 bytes} \\
+        \hline
+        MAC & Median & Average \\
+        \hline
+        Poly1305 & 961 & 2704 \\
+        \hline
+        GMAC & 3899 & 4827 \\
+        \hline
+        CMAC & 7305 & 7588 \\
+        \hline
+        KMAC & 9697 & 9928 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}{|c|c|c|}
+        \hline
+        \multicolumn{3}{|c|}{\bf Input size: 1568 bytes} \\
+        \hline
+        MAC & Median & Average \\
+        \hline
+        Poly1305 & 1065 & 1809 \\
+        \hline
+        GMAC & 4055 & 5026 \\
+        \hline
+        CMAC & 8735 & 8772 \\
+        \hline
+        KMAC & 11647 & 12186 \\
         \hline

### 2024-09-23 quantum-information-processing/quantum.py
Rough explanation of Bell's state and teleportation

@@ -1,0 +1,43 @@
+from sympy import Matrix, sqrt, ShapeError
+from sympy.physics.quantum import TensorProduct
+
+PAULI_X = Matrix([
+    [0, 1],
+    [1, 0]
+])
+PAULI_Z = Matrix([
+    [1, 0],
+    [0, -1]
+])
+HADAMARD = Matrix([
+    [1/sqrt(2), 1/sqrt(2)],
+    [1/sqrt(2), -1/sqrt(2)]
+])
+
+def controlled_u(u: Matrix):
+    """Given unitary operator, compute the unitary operator on controlled-U gate
+    -----*-----
+         |
+    ----|U|----
+    """
+    if u.shape != (2, 2):
+        raise ShapeError("U must be unitary oeprator on a single qubit")
+    return Matrix([
+        [Matrix.eye(2), Matrix.zeros(2)],
+        [Matrix.zeros(2), u]
+    ])
+
+def inverted_controlled_u(u: Matrix):
+    """Given unitary operator, compute the inverted unitary operator
+    ---|U|---
+        |
+    ----*----
+    """
+    if u.shape != (2, 2):
+        raise ShapeError("U must be unitary operator on a single qubit")
+    return Matrix([
+        [1, 0, 0, 0],
+        [0, u[0,0], 0, u[0,1]],
+        [0, 0, 1, 0],
+        [0, u[1,0], 0, u[1,1]],
+    ])

### 2024-09-24 etm-transform/manuscript/main.tex
Added Appendix

@@
 Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.
-Also worth noting is the modification to CRYSTALS-Kyber (as presented in its third round submission to NIST's PQC) before it is standardized: Kyber constructed its KEM routines using the $U^{\not\bot}$ variant of the Fujisaki-Okamoto transform, where the shared secret is derived from hashing both the plaintext and the ciphertext; on the other hand, ML-KEM used the $U^{\not\bot}_m$ variant, where the shared secret is derived from hasing the plaintext alone.
 
 We implemented the ``encrypt-then-MAC'' KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios.
-The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment
+The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
+See appendix \ref{sec:ml-kem-and-kyber} for comparison with Kyber's third round submission to NIST's PQC competition.
 
@@
                 \State $t \leftarrow \mac(k, c^\prime)$
-                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
-                \State $c \leftarrow (c^\prime, c)$
+                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert c^\prime)$
+                \State $c \leftarrow (c^\prime, t)$
                 \State \Return $(c, K)$
@@
         \hline
+        % TODO: this title needs to show that parameters are identical across all variants
         \multicolumn{2}{|c|}{\bf ML-KEM-512} \\
@@
         \hline
+        % ML-KEM+-512
         $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
@@
 
+\section{Appendix}\label{sec:appendix}
+\subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
+ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition.
+While their IND-CPA subroutines (see figure \ref{fig:k-pke-routines}) are identical, ML-KEM deviated from Kyber by choosing a different variant of the Fujisaki-Okamoto transformation.
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.KeyGen()}}
+            \begin{algorithmic}[1]
+                \State $A \leftsample R_q^{k \times k}$
+                \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
+                \State $\pk \leftarrow (A, \mathbf{t})$
+                \State $\sk \leftarrow \mathbf{s}$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.Enc(pk, m)}}
+            \begin{algorithmic}[1]
+                \Ensure $\pk = (A, \mathbf{t})$
+                \Ensure $m \in R_2$
+                \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
+                \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
+                \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
+                \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
+                \State \Return $(\mathbf{c}_1, c_2)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.Dec(sk, c)}}
+            \begin{algorithmic}[1]
+                \Ensure $c = (\mathbf{c}_1, c_2)$
+                \Ensure $\sk = \mathbf{s}$
+                \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
+                \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
+                \State \Return $\hat{m}$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
+\end{figure}
+
+CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext.
+On the other hand, because by using \emph{re-encryption} and \emph{de-randomization}, the PKE is already made \emph{rigid}, the CRYSTALS-Kyber team decided to use the $U^{\not\bot}_m$ variant, where the shared secret is derived from the plaintext alone.
+
+\begin{figure}[H]
+    \centering
+    \begin{minipage}[t]{0.45\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{KEM.KeyGen()}}
+            \begin{algorithmic}[1]
+                \State $z \leftsample \{0,1\}^{256}$
+                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
+                \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+        \begin{algorithm}[H]
+            \caption*{\texttt{KEM.Encap(pk)}}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \{0,1\}^{256}$
+                \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
+                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
+                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
+                    \Comment{$U^{\not\bot}$}
+                \State $K \leftarrow \overline{K}$
+                    \Comment{$U^{\not\bot}_m$}
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.5\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{KEM.Decap(sk, c)}}
+            \begin{algorithmic}[1]
+                \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
+                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
+                \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
+                \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
+                    \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
+                        \Comment{$U^{\not\bot}$}
+                    \State $K \leftarrow \overline{K}$
+                        \Comment{$U^{\not\bot}_m$}
+                \Else 
+                    \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
+                \EndIf
+                \State \Return $K$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{
+        Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
+    }\label{fig:kyber-ml-kem-compare}
+\end{figure}
+
+The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold.
+The first reason is performance: using the $U^{\not\bot}_m$ transformation saves the need to hash the ciphertext, and since Kyber/ML-KEM's performance is mainly bottlenecked by the symmetric components, omitting the hash leads to significant runtime savings (up to 17\% in AVX-2 optimized implementations).
+The second reason is the simplified security proof and tighter security bounds of the $U^{\not\bot}_m$ variant compared to the $U^{\not\bot}$ variant.
+We will omit the details of the security proof and refer readers to \cite{hofheinz2017modular}.
+
+In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.
+
 \end{document}

### 2024-09-24 etm-transform/manuscript/main.tex
Use DBLP citation where possible

@@
 % TODO: need more thorough references
-% TODO: need better intro and conclusion (but later)
+% TODO: need better conclusion
 % \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
@@
 \section{Introduction}
-The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks.
-Later works \cite{hofheinz2017modular} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals}, FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
-% TODO: Mention and cite ML-KEM
-
-The current state of the FO transformation enjoys proven tight security bound and quantum resistance \cite{hofheinz2017modular}, but also leaves many open problems.
+The Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks.
+Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{nist2024module}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
+
+The current state of the FO transformation enjoys proven tight security bound and quantum resistance, but also leaves many open problems.
 One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} \cite{bernstein2018towards} for providing ciphertext integrity.
-In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine dominates the runtime cost of the decapsulation routine.
+In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine substantially increases the runtime cost of the decapsulation routine.
 While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.
@@
 The main challenge in applying ``encrypt-then-MAC'' to public-key cryptography is the lack of a pre-shared symmetric key.
-We took inspiration from hybrid public-key encryption schemes (HPKE) such as .We proposed to derive the symmetric key by hashing the plaintext message.
+Inspired by hybrid public-key encryption schemes such as \cite{DBLP:conf/crypto/FujisakiO99} and \cite{DBLP:conf/crypto/CramerS98}, we proposed to derive the symmetric key by hashing the plaintext message.
 In section \ref{sec:main-results}, we prove that under the random oracle model, if the input public-key encryption scheme is one-way secure against plaintext-checking attack and the input message authentication code is one-time existentially unforgeable, then the transformed key encapsulation mechanism is IND-CCA2 secure.
 
-In section \ref{sec:implementation}, we instantiate concrete instances of our constructions by combining Kyber with GMAC and Poly1305.
+In section \ref{sec:implementation}, we instantiate concrete instances of our constructions using ML-KEM.
 Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal runtime overhead in the encapsulation routine and a small increase in ciphertext size.
@@
 \subsection{Public-key encryption scheme}
-A public key encryption scheme $\pke$ is a collection of three routines $(\keygen, \encrypt, \decrypt)$ defined over some message space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
-Where the encryption routine is probabilistic, the source of randomness is denoted by the coin space $\mathcal{R}$.
-
-The encryption routine $\encrypt(\pk, m)$ takes a public key, a plaintext message, and outputs a ciphertext $c \in \mathcal{C}$.
-Where the encryption routine is probabilistic, specifying a pseudorandom seed $r \in \mathcal{R}$ will make the encryption routine behave deterministically.
-The decryption routine $\decrypt(\sk, c)$ takes a secret key, a ciphertext, and outputs the decryption $\hat{m}$ if the ciphertext is valid.
-Some PKE will explicitly reject invalid ciphertext, in which case the decryption routine will output the rejection symbol $\bot$
-
-We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
+\emph{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
+$(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
+The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key.
+The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key.
+Where the encryption routine is randomized, we denote the randomness by $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space.
+The decryption routine is assumed to always be deterministic.
+Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.
+
+\emph{Correctness.} We usually demand that the decryption routine always correctly recovers the plaintext from an honest encryption, meaning that for all possible keypairs $(\pk, \sk)$ and plaintext $m \in \mathcal{M}$.
+
+\begin{equation*}
+    P\left\lbrack \decrypt(\sk, c) = m \mid c \leftsample \encrypt(\pk, m)\right\rbrack = 1
+\end{equation*}
+
+However, many lattice cryptosystems do not achieve perfect correctness. \cite{hofheinz2017modular} defines $\delta$-correctness to mean:
+
+\begin{equation*}
+    E\left[\max_{m \in \mathcal{M}} P\left[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)\right]\right] \leq \delta
+\end{equation*}
+
+Where the expectation is taken over the probability distribution of possible keypairs.
+Decryption failures can leak information about the secret key and need to be delicately handled, though the concrete probability is small enough that we will ignore them in this paper (ML-KEM-512's probability of decryption failure is estimated to be $2^{-80}$)
+
+\emph{Security.} We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
 Specifically, we first define the $\texttt{OW-ATK}$ as they pertain to a public key encryption scheme.
@@
 \subsection{Related works}
-The Fujisaki-Okamoto transformation \cite{fujisaki1999secure}\cite{hofheinz2017modular} is a family of generic transformations that takes as input a PKE with weaker security, such as OW-CPA, and outputs a PKE or KEM with IND-CCA2 security.
+The modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is a family of generic transformations that takes as input a PKE with weaker security, such as OW-CPA, and outputs a PKE or KEM with IND-CCA2 security.
 The key ingredient in achieving ciphertext non-malleability is with \emph{de-randomization} and \emph{re-encryption}, which first transform a OW-CPA PKE into a \emph{rigid} PKE, then transform the rigid PKE into a KEM.
@@
 Converting a one-way secure rigid PKE (which is essentially a trapdoor function) into a IND-CCA2 KEM is well solved problem.
-We refer readers to \cite{boneh2020graduate} for details on such constructions.
+%TODO: (cite: trapdoor permutation can make KEM)
 \end{enumerate}
@@
         % TODO: this title needs to show that parameters are identical across all variants
-        \multicolumn{2}{|c|}{\bf ML-KEM-512} \\
+        \multicolumn{2}{|c|}{\bf 128-bit security} \\
         \hline
         \multicolumn{2}{|c|}{size parameters (bytes)} \\
-        pubkey size & 800 \\
-        seckey size & 1632 \\
-        ciphertext size & 768 \\
+        $\pk$ size & 800 \\
+        $\sk$ size & 1632 \\
+        \texttt{ct} size & 768 \\
         \hline
@@
         \hline
-        \multicolumn{2}{|c|}{\bf ML-KEM-768} \\
+        \multicolumn{2}{|c|}{\bf 192-bit security} \\
         \hline
         \multicolumn{2}{|c|}{size parameters (bytes)} \\
-        pubkey size & 1184 \\
-        seckey size & 2400 \\
-        ciphertext size & 1088 \\
+        $\pk$ size & 1184 \\
+        $\sk$ size & 2400 \\
+        \texttt{ct} size & 1088 \\
         \hline
@@
         \hline
-        \multicolumn{2}{|c|}{\bf ML-KEM-1024} \\
+        \multicolumn{2}{|c|}{\bf 256-bit security} \\
         \hline
         \multicolumn{2}{|c|}{size parameters (bytes)} \\
-        pubkey size & 1568 \\
-        seckey size & 3168 \\
-        ciphertext size & 1568 \\
+        $\pk$ size & 1568 \\
+        $\sk$ size & 3168 \\
+        \texttt{ct} size & 1568 \\
         \hline

### 2024-09-25 etm-transform/manuscript/main.tex
Notes on correctness and stuff

@@
 \subsection{Public-key encryption scheme}
-\emph{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
+\textbf{\textit{Syntax.}} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
 $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
@@
 
-\emph{Correctness.} We usually demand that the decryption routine always correctly recovers the plaintext from an honest encryption, meaning that for all possible keypairs $(\pk, \sk)$ and plaintext $m \in \mathcal{M}$.
-
-\begin{equation*}
-    P\left\lbrack \decrypt(\sk, c) = m \mid c \leftsample \encrypt(\pk, m)\right\rbrack = 1
-\end{equation*}
-
-However, many lattice cryptosystems do not achieve perfect correctness. \cite{hofheinz2017modular} defines $\delta$-correctness to mean:
+\textbf{\textit{Correctness.}} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04} and \cite{hofheinz2017modular}, a $\pke$ is $\delta$-correct if:
 
@@
 
-Where the expectation is taken over the probability distribution of possible keypairs.
-Decryption failures can leak information about the secret key and need to be delicately handled, though the concrete probability is small enough that we will ignore them in this paper (ML-KEM-512's probability of decryption failure is estimated to be $2^{-80}$)
+Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$.
+For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem (see table \ref{tbl:ml-kem-dec-failure-probs}).
+
+\begin{table}[H]
+    \centering
+
+    \begin{tabular}{|c|c|c|c|}
+        \hline
+        & ML-KEM-512 & ML-KEM-768 & ML-KEM-1024 \\
+        \hline
+        $\delta$ & ??? & ??? & ??? \\
+        \hline
+    \end{tabular}
+    \caption{Estimated probability of decryption failure in ML-KEM}\label{tbl:ml-kem-dec-failure-probs}
+\end{table}
+
+On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm.
+For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table.
+We refer readers to \cite{avanzi2019crystals} and \cite{bos2018crystals} for the details.
 

### 2024-09-26 etm-transform/manuscript/main.tex
encrypt-then-MAC works perfectly on ElGamal

@@
 
+\subsection{Instantiation with ElGamal cryptosystem}\label{sec:etm-with-elgamal}
+In this section we instantiate our ``encrypt-then-MAC'' construction with the ElGamal cryptosystem, then prove that the output KEM is IND-CCA2 secure.
+
+The ElGamal cryptosystem \cite{DBLP:journals/tit/Elgamal85} (see figure \ref{fig:elgamal-cryptosystem}) bases its security on the conjectured intractability of the computational Diffie-Hellman problem and the decisional Diffie-Hellman problem.
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.32\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{KeyGen()}}
+            \begin{algorithmic}[1]
+                \State $x \leftsample \mathbb{Z}_q$
+                \State $\sk \leftarrow x$
+                \State $\pk \leftarrow g^x$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\encrypt(\pk, m)$}
+            \begin{algorithmic}[1]
+                \Require $\pk, m \in \mathbb{Z}_q$
+                \State $y \leftsample \mathbb{Z}_q$
+                \State $u \leftarrow g^y$
+                \State $v \leftarrow m \cdot (\pk)^y$
+                \State \Return $(u, v)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\decrypt(\sk, c)$}
+            \begin{algorithmic}[1]
+                \Require $\sk \in \mathbb{Z}_q$
+                \Require $c = (u, v) \in \mathbb{Z}_q^2$
+                \State \Return $(u^{\sk})^{-1} \cdot v$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{The ElGamal cryptosystem}\label{fig:elgamal-cryptosystem}
+\end{figure}
+
+\begin{definition}[Computational Diffie-Hellman (CDH) problem]
+    Let $G = \langle g \rangle$ be a cyclic group with prime order $g$, let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $G, g, g^x, g^y$, compute $g^{xy}$
+\end{definition}
+
+\begin{definition}[Decisional Diffie-Hellman (DDH) problem]
+    Let $G = \langle g \rangle$ be a cyclic group with prime order $g$, let $x, y, z \leftsample \mathbb{Z}_q$ be uniformly random samples, let $h$ be a random variable with probability distribution:
+
+    \begin{equation*}
+        P[h = g^{xy}] = P[h = g^z] = \frac{1}{2}
+    \end{equation*}
+
+    Given $G, g, g^x, g^y, h$, distinguish whether $h = g^{xy}$ or $h = g^{z}$
+\end{definition}
+
+It is well known that if the CDH problem is intractable, then the ElGamal cryptosystem is OW-CPA secure, and if the DDH problem is intractable, then the ElGamal cryptosystem is IND-CPA secure.
+It is easy to show that the DDH problem reduces to the CDH problem.
+However, it is conjectured that the CDH problem remains intractable even if the DDH problem can be solved efficiently.
+
+% TODO: citation needed
+
+\begin{definition}[Interactive computational Diffie-Hellman (ICDH) problem]
+    Let $G = \langle g \rangle$ be a cyclic group with prime order $q$. Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random sample. Let $\mathcal{O}^\text{DDH}$ be a DDH oracle such that:
+
+    \begin{equation*}
+        \mathcal{O}^\text{DDH}: (u, v) \in \mathbb{Z}_q^2 \mapsto \llbrack u^x = v \rrbrack
+    \end{equation*}
+
+    Given $G, g, g^x, g^y$ and access to $\mathcal{O}^\text{DDH}$, compute $g^{xy}$
+\end{definition}
+
+\begin{theorem}\label{thm:elgamal-is-owpca}
+    For every OW-PCA adversary $A$ against the ElGamal cryptosystem, there exists an ICDH adversary $B$ such that:
+
+    \begin{equation*}
+        \texttt{Adv}(A) \leq \texttt{Adv}(B)
+    \end{equation*}
+\end{theorem}
+
+In other words, for cyclic group $G$ where the ICDH assumption holds, the ElGamal cryptosystem is OW-PCA secure.
+From theorem \ref{thm:ow-pca-implies-kem-ind-cca2}, we naturally conclude that combining ``encrypt-then-MAC'' with the ElGamal cryptosystem gives us an IND-CCA2 secure KEM.
+It remains to prove theorem \ref{thm:elgamal-is-owpca}:
+
+\begin{proof}
+    For a sketch of proof, we will construct an ICDH adversary $B$ that uses the OW-PCA adversary $A$ as a sub-routine.
+The main challenge is simulating the plaintext-checking oracle, which can be overcome because $B$ has access to a DDH oracle $\mathcal{O}^\text{DDH}$.
+See figure \ref{fig:icdh-adv} for how $B$ simulates OW-PCA game for $A$:
+
+    \begin{figure}[H]
+        \centering
+
+        \begin{minipage}[t]{0.5\textwidth}
+            \begin{algorithm}[H]
+                \caption*{OW-PCA game for ElGamal cryptosystem}
+                \begin{algorithmic}[1]
+                    \State $x \leftsample \mathbb{Z}_q$
+                    \State $(\pk, \sk) \leftarrow (g^x, x)$
+                    \State $m^\ast \leftsample \mathbb{Z}_q$
+                    \State $y \leftsample \mathbb{Z}_q$
+                    \State $(u, v) \leftarrow (g^y, m \cdot (g^x)^y)$
+                    \State $c^\ast \leftarrow (u, v)$
+                    \State $\hat{m} \leftarrow A^{\pco}(\pk, c^\ast)$
+                    \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \begin{minipage}[t]{0.49\textwidth}
+            \begin{algorithm}[H]
+                \caption*{ICDH adversary $B^{\mathcal{O}^\text{DDH}}(g^x, g^y)$}
+                \begin{algorithmic}[1]
+                    \State $\pk \leftarrow g^x$
+                    \State $u \leftarrow g^y$
+                    \State $v \leftsample \mathbb{Z}_q$
+                    \State $c^\ast = (u, v)$
+                    \State $\hat{m} \leftarrow A^{\mathcal{O}^\pco_1}(\pk, c^\ast)$
+                    \State \Return $\hat{m}^{-1} \cdot v^\ast$
+                \end{algorithmic}
+            \end{algorithm}\vspace{-0.3cm}
+            \begin{algorithm}[H]
+                \caption*{Simulated PCO: $\mathcal{O}^\pco_1(m, c)$}
+                \begin{algorithmic}[1]
+                    \Require $c = (u, v) \in \mathbb{Z}_q^2$
+                    \Require $m \in \mathbb{Z}_q$ \vspace{0.1cm}
+                    \State \Return $\mathcal{O}^\text{DDH}(u, m^{-1}v)$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+
+        \caption{ICDH adversary $B$ can perfectly simulate the OW-PCA game for $A$}\label{fig:icdh-adv}
+    \end{figure}
+
+    First observe that the simulated PCO is exactly the same as a true PCO:
+
+    \begin{equation*}
+        \begin{aligned}
+            & \pco(m, c=(u,v)) = 1 \\
+            & \leftrightarrow m = \decrypt(\sk, c=(u,v)) \\
+            & \leftrightarrow m = u^{-x}v \\
+            & \leftrightarrow u^x = m^{-1}v \\
+            & \leftarrow \mathcal{O}^\texttt{DDH}(u, m^{-1}v) = 1
+        \end{aligned}
+    \end{equation*}
+
+    Second notice that in the OW-PCA game, $m^\ast \leftsample \mathbb{Z}_q$ is a uniformly random element in $\mathbb{Z}_q$, which means that $v \leftarrow m \cdot (g^x)^y$ is a uniformly random element in $\mathbb{Z}_q$.
+When $B$ simulates the OW-PCA game for $A$, it can simply sample a random value for $v$ in the challenge ciphertext, and $A$ will have retained its advantage in recovering the correct decryption without the secret key.
+
+    If $A$ succeeds in recovering the correct decryption while being run by $B$, then $\hat{m} = u^{-x}v$, which means that $\hat{m}^{-1}v = u^x \cdot v^{-1} \cdot v = u^x = (g^y)x = g^{xy}$.
+In this case, $B$ successfully computes $g^{xy}$ and wins the ICDH game.
+Therefore, the advantage of $A$ is at most the advantage of $B$.
+\end{proof}
+
 \section{Implementation}\label{sec:implementation}
@@
 \section{Conclusions and future works}\label{sec:future-works}
-\emph{Comparison with Fujisaki-Okamoto transformation:} We applied the ``encrypt-then-MAC'' transformation to Kyber and saw meaningful performance improvements over using de-randomization and re-encryption.
-Unfortunately the resulting KEM does not achieve the desired full IND-CCA2 security, because Kyber is known to be vulnerable to key-recovery plaintext-checking attack (KR-PCA) \cite{ravi2019generic}\cite{ueno2022curse}.
-We speculate that while Kyber with ``encrypt-then-MAC'' could not achieve the full IND-CCA2 security, it can still be safe for use in ephemeral key exchange, where each secret key is used to decrypt at most one ciphertext (the KR-PCA requires a few hundred decryption queries to recover the secret key).
-
-In section \ref{sec:main-results}, we showed that if the input PKE is OW-PCA secure, then the resulting KEM is IND-CCA2 secure.
-One sufficient condition for OW-PCA security is one-way security plus rigidity.
-If the input PKE is rigid, then $m = \decrypt(\sk, c)$ is equivalent to $c = \encrypt(\pk, m)$, so a plaintext-checking oracle can be simulated without any secret information.
-However, the $U_m^{\not\bot}$ transformation in \cite{hofheinz2017modular} can already transform an OW-CPA secure and rigid PKE into an IND-CCA2 secure KEM with minimal overhead: the encapsulation and decapsulation routines each adds a hash of the plaintext to the encryption and decryption routine.
-In other words, where the input PKE is rigid, ``encrypt-then-MAC'' doesn't offer any performance advantage.
-It remains an open problem whether there exists a PKE that is OW-PCA secure but not rigid.
-If such a PKE exists, then ``encrypt-then-MAC'' would be a preferable strategy for constructing an IND-CCA2 KEM.
+The ``encrypt-then-MAC'' transformation is a generic KEM construction that achieves IND-CCA2 security under the random oracle model if the input PKE is OW-PCA secure.
+Compared to the Fujisaki-Okamoto transformation, our construction replaced \emph{de-randomization} and \emph{re-encryption} with a message authenticator.
+At the cost of some minimal increase in communication size and encapsulation runtime, our construction achieves significant efficiency gains in the decapsulation routine.
+In practical key exchange protocols, our construction saves between 35-45\% in round trip time.
+
+% TODO: write about open problems
 
@@
 \section{Appendix}\label{sec:appendix}
+% TODO: brief introduction to the various authenticators should be in the appendix
+
 \subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}

### 2024-09-30 etm-transform/manuscript/main.tex
Checkpoint

@@
 The Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks.
-Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{nist2024module}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
+Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}
+\cite{bernstein2018towards} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{FIPS203}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
 
 The current state of the FO transformation enjoys proven tight security bound and quantum resistance, but also leaves many open problems.
-One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} \cite{bernstein2018towards} for providing ciphertext integrity.
+One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} for providing ciphertext integrity.
 In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine substantially increases the runtime cost of the decapsulation routine.
 While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.
+
+% TODO: trivially combining PKE and digital signature is not authenticated encryption?
+
+Achiving indistinguishability under adaptive chosen ciphertext attack (IND-CCA2) is a solved problem in symmetric cryptography.
+Given an IND-CPA secure symmetric cipher and an existentially unforgeable message authentication code, combining them in a pattern called ``encrypt-then-MAC'' achieves IND-CCA2 security \cite{DBLP:conf/asiacrypt/BellareN00}\cite{DBLP:conf/eurocrypt/CanettiK01}\cite{DBLP:conf/crypto/Krawczyk01}.
+However, ``encrypt-then-MAC'' cannot be directly applied within the context of public-key cryptography, because the two parties do not have pre-shared symmetric keys.
+Our main contribution is a novel generic key encapsulation mechanism (KEM) construction that applies the ``encrypt-then-MAC'' pattern by deriving the symmetric key from some public-key encryption scheme's plaintext.
 

### 2024-10-01 etm-transform/manuscript/main.tex
Started with rewriting introduction

--- 2024-09-26 etm-transform/manuscript/main.tex
@@
 \section{Introduction}
-The Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks.
-Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{nist2024module}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.
-
-The current state of the FO transformation enjoys proven tight security bound and quantum resistance, but also leaves many open problems.
-One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} \cite{bernstein2018towards} for providing ciphertext integrity.
-In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine substantially increases the runtime cost of the decapsulation routine.
-While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.
-
-We are inspired by how ciphertext integrity is achieved in symmetric cryptography: given a semantically secure symmetric cipher and an existentially unforgeable message authentication code, combining them using in a pattern called ``encrypt-then-MAC'' provides proven authenticated encryption \cite{bellare2000authenticated}.
-``encrypt-then-MAC'' is now the most widely accepted method for doing symmetric encryption as AES-GCM \cite{mcgrew2004galois} and ChaCha20-Poly1305 \cite{nir2018chacha20}.
-
-The main challenge in applying ``encrypt-then-MAC'' to public-key cryptography is the lack of a pre-shared symmetric key.
-Inspired by hybrid public-key encryption schemes such as \cite{DBLP:conf/crypto/FujisakiO99} and \cite{DBLP:conf/crypto/CramerS98}, we proposed to derive the symmetric key by hashing the plaintext message.
-In section \ref{sec:main-results}, we prove that under the random oracle model, if the input public-key encryption scheme is one-way secure against plaintext-checking attack and the input message authentication code is one-time existentially unforgeable, then the transformed key encapsulation mechanism is IND-CCA2 secure.
-
-In section \ref{sec:implementation}, we instantiate concrete instances of our constructions using ML-KEM.
-Our experimental results showed that replacing re-encryption with computing authenticator leads to significant performance improvements in the decapsulation routine while incurring only minimal runtime overhead in the encapsulation routine and a small increase in ciphertext size.
+Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
+The most widely accepted security standard for a KEM is \emph{indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}: no efficient adversary can distinguish a shared secret obtained by running encapsulation from random noise even with access to a decapsulation oracle throughout the attack.
+KEM is usually paired with a data encapsulation mechanism (DEM), such as AES-GCM or ChaCha20-Poly1305 when constructing secure communication protocol.
+
+Provable IND-CCA2 security is difficult to achieve from scratch.
+One popular approach to constructing secure KEM is by starting with a public-key encryption (PKE) scheme with weaker security properties, then adding additional mechanisms for ensuring ciphertext integrity.
+One particularly successful example is the Fujisaki-Okamoto transformation, originally proposed in 1999 by Fujisaki Eiichiro and Okamoto Tatsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}.
+The original Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE that achieves the full IND-CCA2 security in the random oracle model.
+Subsequent works improved on the original construction, culminating in the 2017 paper \cite{hofheinz2017modular} by Hofheinz, Hovelmann, and Kiltz, which provided a modular KEM construction with tight IND-CCA2 security reduction against classical adversary and proven (though non-tight) IND-CCA2 security reduction against quantum adversaries.
+
+The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
+It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{bos2018crystals}\cite{avanzi2019crystals}, Saber \cite{d2018saber}, FrodoKEM \cite{bos2016frodo}, and classic McEliece \cite{} among others.
+When Kyber was standardized by NIST in FIPS 203 ``Module-lattice key-encapsulation mechanism'' (ML-KEM) \cite{FIPS203}, it kept the Fujisaki-Okamoto transformation in its KEM construction.
+However, the Fujisaki-Okamoto transformation is not perfect.
+Among its shortcomings is the use of \emph{de-randomization} and \emph{re-encryption} for ensuring ciphertext integrity: \begin{itemize}
+    \item \emph{Computational inefficiency.}
+    \item \emph{Side-channel vulnerability}
+\end{itemize}
 

### 2024-10-02 etm-transform/manuscript/main.tex
Made progress on introduction

@@
     Ganyu Xu\inst{1}
+    \and Kalikinkar Mandal\inst{2}
     \and Guang Gong\inst{1}
-    \and Kalikinkar Mandal\inst{2}
 }
@@
 Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
-The most widely accepted security standard for a KEM is \emph{indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}: no efficient adversary can distinguish a shared secret obtained by running encapsulation from random noise even with access to a decapsulation oracle throughout the attack.
-KEM is usually paired with a data encapsulation mechanism (DEM), such as AES-GCM or ChaCha20-Poly1305 when constructing secure communication protocol.
-
-Provable IND-CCA2 security is difficult to achieve from scratch.
-One popular approach to constructing secure KEM is by starting with a public-key encryption (PKE) scheme with weaker security properties, then adding additional mechanisms for ensuring ciphertext integrity.
-One particularly successful example is the Fujisaki-Okamoto transformation, originally proposed in 1999 by Fujisaki Eiichiro and Okamoto Tatsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}.
-The original Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE that achieves the full IND-CCA2 security in the random oracle model.
-Subsequent works improved on the original construction, culminating in the 2017 paper \cite{hofheinz2017modular} by Hofheinz, Hovelmann, and Kiltz, which provided a modular KEM construction with tight IND-CCA2 security reduction against classical adversary and proven (though non-tight) IND-CCA2 security reduction against quantum adversaries.
+The combination of KEM and some data encapsulation mechanism (DEM), such as AES-GCM and ChaCha20-Poly1305, is the foundation of many of today's most popular secure communication protocols such as Transport Layer Security (TLS) and Secure Shell (SSH).
+The commonly accepted security standard of a KEM is \emph{Indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}: no efficient adversary can distinguish a shared secret obtained by running the encapsulation routine from random noise, even with access to a decapsulation oracle throughout the attack.
+KEM is related to another important cryptographic primitive called public-key encryption (PKE), and their security standards are similar.
+The desired security standard for PKE, also called IND-CCA2, requires that no efficient adversary can distinguish the encryption of two adversarially chosen messages even with access to decryption oracle.
+
+It is difficult to build a provably IND-CCA2 secure KEM from scratch.
+Instead, secure KEMs are usually built on top of a PKE with weaker security property (e.g.
+being only OW-CPA or IND-CPA secure, not IND-CCA2).
+One such construction is the Fujisaki-Okamoto transformation, proposed in 1999 by Fujisaki Eiichiro and Okamoto Tetsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}.
+The first Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE (HPKE) with proven IND-CCA2 security under the random oracle model.
+Subsequent proposals such as \cite{DBLP:conf/ctrsa/OkamotoP01}, \cite{DBLP:conf/ctrsa/CoronHJPPT02}, and \cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE.
+This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.
 
 The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
-It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{bos2018crystals}\cite{avanzi2019crystals}, Saber \cite{d2018saber}, FrodoKEM \cite{bos2016frodo}, and classic McEliece \cite{} among others.
+It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
 When Kyber was standardized by NIST in FIPS 203 ``Module-lattice key-encapsulation mechanism'' (ML-KEM) \cite{FIPS203}, it kept the Fujisaki-Okamoto transformation in its KEM construction.
 However, the Fujisaki-Okamoto transformation is not perfect.
-Among its shortcomings is the use of \emph{de-randomization} and \emph{re-encryption} for ensuring ciphertext integrity: \begin{itemize}
-    \item \emph{Computational inefficiency.}
-    \item \emph{Side-channel vulnerability}
+Among its shortcomings is the use of \emph{de-randomization} (transform a randomized encryption routine into a deterministic routine by using a pseudorandom coin derived from the input message) and \emph{re-encryption} (the decapsulation routine of the output KEM runs the encryption routine of the input PKE to ensure ciphertext integrity), which causes two problems: \begin{itemize}
+    \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
+    \item \textbf{side-channel vulnerability:} running the input PKE's encryption routine in the output KEM's decapsulation routine introduces risk of side-channel vulnerabilities not found in the input PKE's decryption routine alone.
+In fact, many practical attacks \cite{ueno2022curse}\cite{ravi2019generic} exploit re-encryption to decrypt ciphertext or recover secret keys.
+Countermeasures such as masking have been proposed to address these side channels, but they inevitably carry substantial performance penalty.
 \end{itemize}
+
+\subsection{Our contributions}
+Our main contribution is a novel generic construction that combines a PKE with weaker security property and a one-time secure message authentication code (MAC) into an IND-CCA2 secure KEM in the random oracle model.
+Specifically, we require the input PKE to be one-way secure against plaintext-checking attacks (OW-PCA): no efficient adversary can recover the decryption of a random encryption with access to a plaintext-checking oracle (PCO) that, when queried on a plaintext-ciphertext pair $(m, c)$, answers $m$ is the decryption of $c$.
+While OW-PCA is not a standard security definition, it has appeared useful in past IND-CCA2 KEM constructions \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}, and plays an important role in the modularity of the Fujisaki-Okamoto KEM transformation.
+
+Our KEM construction is inspired by how symmetric cryptography achieves IND-CCA2 security (or equivalently, authenticated encryption \cite{DBLP:conf/crypto/Krawczyk01}) using the ``encrypt-then-MAC'' pattern.
+We applied ``encrypt-then-MAC'' to the construction of the KEM: at encapsulation, a symmetric key is derived from hashing a random PKE plaintext, then used to compute an authenticator against the PKE ciphertext; at decapsulation, the same symmetric key is derived from hashing the decryption of the PKE ciphertext, then used to verify the authenticator.
+Intuitively, for an adversary to be able to produce a valid authenticator, it must know the correct symmetric key and thus the corresponding PKE plaintext.
+In other words, the adversary either produced the ciphertext honestly, or have broken the one-wayness of the input PKE.
+
+\subsubsection{Performance improvements}
+The main advantage of our KEM construction over the Fujisaki-Okamoto transformation is the performance gains: our construction replaces re-encryption with computing a symmetric message authenticator, which is significantly faster, especially for one-time MAC.
+At the cost of computing an authenticator in encapsulation and increasing the ciphertext size by a message authenticator, our construction speeds up decapsulation by 10x.
+
+When instantiated with the underlying PKE routines of ML-KEM and the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
+
+We also implemented and measured the round trip time of key exchange protocols with various modes of authentication.
+When compared to ML-KEM, $\text{ML-KEM}^+$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange, 29\%-35\% reduction in unilaterally authenticated key exchange, and 35\%-48\% reduction in mutually authenticated key exchange.
+
+\subsection{Related works}
+A similar construction was proposed by Abdulla et al \cite{DBLP:conf/ctrsa/AbdallaBR01}, though our construction builds a KEM instead of a HPKE, and our construction is generic over all OW-PCA secure PKE.
 
@@
 The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
-
-\subsection{Related works}
-The modular Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is a family of generic transformations that takes as input a PKE with weaker security, such as OW-CPA, and outputs a PKE or KEM with IND-CCA2 security.
-The key ingredient in achieving ciphertext non-malleability is with \emph{de-randomization} and \emph{re-encryption}, which first transform a OW-CPA PKE into a \emph{rigid} PKE, then transform the rigid PKE into a KEM.
-More specifically: \begin{enumerate}
-    \item \emph{de-randomization} means that a randomized encryption routine $c \leftsample \encrypt(\pk, m)$ is made into a deterministic encryption routine by deriving randomization coin pseudorandomly: $c \leftarrow \encrypt(\pk, m, r = H(m))$ for some hash function $H$
-    \item \emph{re-encryption} means that the transformed decryption routine will run the transformed encryption routine to verify the integrity of the ciphertext.
-Because after \emph{de-randomization}, each plaintext strictly corresponds exacty one ciphertext, tempering with a ciphertext means that even if the ciphertext decrypts back to the same plaintext, the re-encryption will detect that the ciphertext has been tempered with.
-    \item \emph{rigidity} means that the decryption routine is a perfect inverse of the encryption routine: $c = \encrypt(\pk, m) \Leftrightarrow m = \decrypt(\sk, c)$.
-Converting a one-way secure rigid PKE (which is essentially a trapdoor function) into a IND-CCA2 KEM is well solved problem.
-%TODO: (cite: trapdoor permutation can make KEM)
-\end{enumerate}
-
-let $\pke = (\keygen_\pke, \encrypt, \decrypt)$ be defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$.
-Let $G: \mathcal{M} \rightarrow \mathcal{R}$ hash plaintexts into coincs, and let $H: \{0,1\}^\ast \rightarrow \{0,1\}^\ast$ hash byte stream into session keys.
-Depending on whether the constructed KEM uses implicit or explicit rejection, and the security property of the $\pke$, \cite{hofheinz2017modular} described four variations.
-They are summarized in table \ref{tbl:fo-variations} and figure \ref{fig:fo-routines}.
-
-\begin{table}[H]
-    \centering
-    \begin{tabular}{|c|c|c|}
-        \hline
-        name & rejection & PKE security \\
-        \hline
-        $U^\bot$ & explicit & OW-PCVA \\
-        \hline
-        $U^{\not\bot}$ & implicit & OW-PCA \\
-        \hline
-        $U^\bot_m$ & explicit & OW-VA + rigid \\
-        \hline
-        $U^{\not\bot}_m$ & implicit & OW-CPA + rigid \\
-        \hline
-    \end{tabular}
-    \caption{Variants of modular FO transforms}\label{tbl:fo-variations}
-\end{table}
-
-\begin{figure}[H]
-    \centering
-    \begin{minipage}[t]{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\kem\texttt{.}\keygen()$}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
-                \State $z \leftsample \mathcal{M}$
-                \State $\sk \leftarrow (\sk^\prime, z)$
-                    \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
-                \State $\sk \leftarrow \sk^\prime$
-                    \Comment{$U^{\bot}, U^{\bot}_m$}
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-        \begin{algorithm}[H]
-            \caption*{$\kem\texttt{.}\encap(\pk)$}
-            \begin{algorithmic}[1]
-                \State $m \leftsample \mathcal{M}$
-                \State $r \leftarrow G(m)$
-                \State $c \leftarrow \pke\texttt{.}\encrypt(\pk, m, r)$
-                \State $K \leftarrow H(m, c)$
-                    \Comment{$U^\bot, U^{\not\bot}$}
-                \State $K \leftarrow H(m)$
-                    \Comment{$U_m^\bot, U_m^{\not\bot}$}
-                \State \Return $(c, K)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.5\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\kem\texttt{.}\decap(\sk, c)$}
-            \begin{algorithmic}[1]
-                \Ensure $\sk = (\sk^\prime, z)$
-                \Ensure $\sk^\prime$ is valid PKE secret key
-                \Ensure $z$ is valid PKE plaintext
-                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c)$
-                \State $\hat{r} \leftarrow G(\hat{m})$
-                \State $\hat{c} \leftarrow \pke\texttt{.}\encrypt(\pk, \hat{m}, \hat{r})$
-                \If{$\hat{c} = c$}
-                    \State $K \leftarrow H(\hat{m})$
-                        \Comment{$U_m^\bot, U_m^{\not\bot}$}
-                    \State $K \leftarrow H(\hat{m}, c)$
-                        \Comment{$U^\bot, U^{\not\bot}$}
-                \Else
-                    \State $K \leftarrow H(z, c)$
-                        \Comment{$U^{\not\bot}, U^{\not\bot}_m$}
-                    \State $K \leftarrow \bot$
-                        \Comment{$U^{\bot}, U^{\bot}_m$}
-                \EndIf 
-                \State \Return $K$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{Summary of the modular Fujisaki-Okamoto transformation variations}\label{fig:fo-routines}
-\end{figure}
-
-The modular FO transformations enjoy tight security bounds and proven quantum resistance.
-Variations have been deployed to many post-quantum KEMs submitted to NIST's post-quantum cryptography competition.
-Kyber, one of the round 3 finalists, uses the $U^{\not\bot}$ transformation.
-When it was later standardized into FIPS-203, it changed to use the $U^{\not\bot}_m$ transformation for computational efficiencies.
 
@@
 
-\subsection{Instantiation with ElGamal cryptosystem}\label{sec:etm-with-elgamal}
-In this section we instantiate our ``encrypt-then-MAC'' construction with the ElGamal cryptosystem, then prove that the output KEM is IND-CCA2 secure.
-
-The ElGamal cryptosystem \cite{DBLP:journals/tit/Elgamal85} (see figure \ref{fig:elgamal-cryptosystem}) bases its security on the conjectured intractability of the computational Diffie-Hellman problem and the decisional Diffie-Hellman problem.
-
-\begin{figure}[H]
-    \centering
-
-    \begin{minipage}[t]{0.32\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{KeyGen()}}
-            \begin{algorithmic}[1]
-                \State $x \leftsample \mathbb{Z}_q$
-                \State $\sk \leftarrow x$
-                \State $\pk \leftarrow g^x$
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\encrypt(\pk, m)$}
-            \begin{algorithmic}[1]
-                \Require $\pk, m \in \mathbb{Z}_q$
-                \State $y \leftsample \mathbb{Z}_q$
-                \State $u \leftarrow g^y$
-                \State $v \leftarrow m \cdot (\pk)^y$
-                \State \Return $(u, v)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\decrypt(\sk, c)$}
-            \begin{algorithmic}[1]
-                \Require $\sk \in \mathbb{Z}_q$
-                \Require $c = (u, v) \in \mathbb{Z}_q^2$
-                \State \Return $(u^{\sk})^{-1} \cdot v$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-
-    \caption{The ElGamal cryptosystem}\label{fig:elgamal-cryptosystem}
-\end{figure}
-
-\begin{definition}[Computational Diffie-Hellman (CDH) problem]
-    Let $G = \langle g \rangle$ be a cyclic group with prime order $g$, let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $G, g, g^x, g^y$, compute $g^{xy}$
-\end{definition}
-
-\begin{definition}[Decisional Diffie-Hellman (DDH) problem]
-    Let $G = \langle g \rangle$ be a cyclic group with prime order $g$, let $x, y, z \leftsample \mathbb{Z}_q$ be uniformly random samples, let $h$ be a random variable with probability distribution:
-
-    \begin{equation*}
-        P[h = g^{xy}] = P[h = g^z] = \frac{1}{2}
-    \end{equation*}
-
-    Given $G, g, g^x, g^y, h$, distinguish whether $h = g^{xy}$ or $h = g^{z}$
-\end{definition}
-
-It is well known that if the CDH problem is intractable, then the ElGamal cryptosystem is OW-CPA secure, and if the DDH problem is intractable, then the ElGamal cryptosystem is IND-CPA secure.
-It is easy to show that the DDH problem reduces to the CDH problem.
-However, it is conjectured that the CDH problem remains intractable even if the DDH problem can be solved efficiently.
-
-% TODO: citation needed
-
-\begin{definition}[Interactive computational Diffie-Hellman (ICDH) problem]
-    Let $G = \langle g \rangle$ be a cyclic group with prime order $q$. Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random sample. Let $\mathcal{O}^\text{DDH}$ be a DDH oracle such that:
-
-    \begin{equation*}
-        \mathcal{O}^\text{DDH}: (u, v) \in \mathbb{Z}_q^2 \mapsto \llbrack u^x = v \rrbrack
-    \end{equation*}
-
-    Given $G, g, g^x, g^y$ and access to $\mathcal{O}^\text{DDH}$, compute $g^{xy}$
-\end{definition}
-
-\begin{theorem}\label{thm:elgamal-is-owpca}
-    For every OW-PCA adversary $A$ against the ElGamal cryptosystem, there exists an ICDH adversary $B$ such that:
-
-    \begin{equation*}
-        \texttt{Adv}(A) \leq \texttt{Adv}(B)
-    \end{equation*}
-\end{theorem}
-
-In other words, for cyclic group $G$ where the ICDH assumption holds, the ElGamal cryptosystem is OW-PCA secure.
-From theorem \ref{thm:ow-pca-implies-kem-ind-cca2}, we naturally conclude that combining ``encrypt-then-MAC'' with the ElGamal cryptosystem gives us an IND-CCA2 secure KEM.
-It remains to prove theorem \ref{thm:elgamal-is-owpca}:
-
-\begin{proof}
-    For a sketch of proof, we will construct an ICDH adversary $B$ that uses the OW-PCA adversary $A$ as a sub-routine.
-The main challenge is simulating the plaintext-checking oracle, which can be overcome because $B$ has access to a DDH oracle $\mathcal{O}^\text{DDH}$.
-See figure \ref{fig:icdh-adv} for how $B$ simulates OW-PCA game for $A$:
-
-    \begin{figure}[H]
-        \centering
-
-        \begin{minipage}[t]{0.5\textwidth}
-            \begin{algorithm}[H]
-                \caption*{OW-PCA game for ElGamal cryptosystem}
-                \begin{algorithmic}[1]
-                    \State $x \leftsample \mathbb{Z}_q$
-                    \State $(\pk, \sk) \leftarrow (g^x, x)$
-                    \State $m^\ast \leftsample \mathbb{Z}_q$
-                    \State $y \leftsample \mathbb{Z}_q$
-                    \State $(u, v) \leftarrow (g^y, m \cdot (g^x)^y)$
-                    \State $c^\ast \leftarrow (u, v)$
-                    \State $\hat{m} \leftarrow A^{\pco}(\pk, c^\ast)$
-                    \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
-                \end{algorithmic}
-            \end{algorithm}
-        \end{minipage}
-        \begin{minipage}[t]{0.49\textwidth}
-            \begin{algorithm}[H]
-                \caption*{ICDH adversary $B^{\mathcal{O}^\text{DDH}}(g^x, g^y)$}
-                \begin{algorithmic}[1]
-                    \State $\pk \leftarrow g^x$
-                    \State $u \leftarrow g^y$
-                    \State $v \leftsample \mathbb{Z}_q$
-                    \State $c^\ast = (u, v)$
-                    \State $\hat{m} \leftarrow A^{\mathcal{O}^\pco_1}(\pk, c^\ast)$
-                    \State \Return $\hat{m}^{-1} \cdot v^\ast$
-                \end{algorithmic}
-            \end{algorithm}\vspace{-0.3cm}
-            \begin{algorithm}[H]
-                \caption*{Simulated PCO: $\mathcal{O}^\pco_1(m, c)$}
-                \begin{algorithmic}[1]
-                    \Require $c = (u, v) \in \mathbb{Z}_q^2$
-                    \Require $m \in \mathbb{Z}_q$ \vspace{0.1cm}
-                    \State \Return $\mathcal{O}^\text{DDH}(u, m^{-1}v)$
-                \end{algorithmic}
-            \end{algorithm}
-        \end{minipage}
-
-        \caption{ICDH adversary $B$ can perfectly simulate the OW-PCA game for $A$}\label{fig:icdh-adv}
-    \end{figure}
-
-    First observe that the simulated PCO is exactly the same as a true PCO:
-
-    \begin{equation*}
-        \begin{aligned}
-            & \pco(m, c=(u,v)) = 1 \\
-            & \leftrightarrow m = \decrypt(\sk, c=(u,v)) \\
-            & \leftrightarrow m = u^{-x}v \\
-            & \leftrightarrow u^x = m^{-1}v \\
-            & \leftarrow \mathcal{O}^\texttt{DDH}(u, m^{-1}v) = 1
-        \end{aligned}
-    \end{equation*}
-
-    Second notice that in the OW-PCA game, $m^\ast \leftsample \mathbb{Z}_q$ is a uniformly random element in $\mathbb{Z}_q$, which means that $v \leftarrow m \cdot (g^x)^y$ is a uniformly random element in $\mathbb{Z}_q$.
-When $B$ simulates the OW-PCA game for $A$, it can simply sample a random value for $v$ in the challenge ciphertext, and $A$ will have retained its advantage in recovering the correct decryption without the secret key.
-
-    If $A$ succeeds in recovering the correct decryption while being run by $B$, then $\hat{m} = u^{-x}v$, which means that $\hat{m}^{-1}v = u^x \cdot v^{-1} \cdot v = u^x = (g^y)x = g^{xy}$.
-In this case, $B$ successfully computes $g^{xy}$ and wins the ICDH game.
-Therefore, the advantage of $A$ is at most the advantage of $B$.
-\end{proof}
-
 \section{Implementation}\label{sec:implementation}

### 2024-10-03 etm-transform/manuscript/main.tex
Drafted introduction

@@
 \usepackage{multirow}
+\usepackage{tikz}
+\usetikzlibrary{positioning}
 % \usepackage[demo]{graphicx}
@@
 %%%% 7. PAPER CONTENT %%%%
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
@@
 The first Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE (HPKE) with proven IND-CCA2 security under the random oracle model.
-Subsequent proposals such as \cite{DBLP:conf/ctrsa/OkamotoP01}, \cite{DBLP:conf/ctrsa/CoronHJPPT02}, and \cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE.
+Subsequent works \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}\cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE.
 This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.
@@
 However, the Fujisaki-Okamoto transformation is not perfect.
-Among its shortcomings is the use of \emph{de-randomization} (transform a randomized encryption routine into a deterministic routine by using a pseudorandom coin derived from the input message) and \emph{re-encryption} (the decapsulation routine of the output KEM runs the encryption routine of the input PKE to ensure ciphertext integrity), which causes two problems: \begin{itemize}
+As pointed out by the authors of \cite{hofheinz2017modular}, the Fujisaki-Okamoto transformation uses \emph{de-randomization} and \emph{re-encryption} to ensure ciphertext integrity, which brings two problems: \begin{itemize}
     \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
@@
 \subsection{Our contributions}
-Our main contribution is a novel generic construction that combines a PKE with weaker security property and a one-time secure message authentication code (MAC) into an IND-CCA2 secure KEM in the random oracle model.
-Specifically, we require the input PKE to be one-way secure against plaintext-checking attacks (OW-PCA): no efficient adversary can recover the decryption of a random encryption with access to a plaintext-checking oracle (PCO) that, when queried on a plaintext-ciphertext pair $(m, c)$, answers $m$ is the decryption of $c$.
-While OW-PCA is not a standard security definition, it has appeared useful in past IND-CCA2 KEM constructions \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}, and plays an important role in the modularity of the Fujisaki-Okamoto KEM transformation.
-
-Our KEM construction is inspired by how symmetric cryptography achieves IND-CCA2 security (or equivalently, authenticated encryption \cite{DBLP:conf/crypto/Krawczyk01}) using the ``encrypt-then-MAC'' pattern.
-We applied ``encrypt-then-MAC'' to the construction of the KEM: at encapsulation, a symmetric key is derived from hashing a random PKE plaintext, then used to compute an authenticator against the PKE ciphertext; at decapsulation, the same symmetric key is derived from hashing the decryption of the PKE ciphertext, then used to verify the authenticator.
-Intuitively, for an adversary to be able to produce a valid authenticator, it must know the correct symmetric key and thus the corresponding PKE plaintext.
-In other words, the adversary either produced the ciphertext honestly, or have broken the one-wayness of the input PKE.
+Our main contribution is a novel generic KEM construction that combines a PKE with weak security and a message authentication code (MAC).
+We were inspired by the strategy used in symmetric cyrptography for achieving IND-CCA2 security: combining a symmetric cipher with a MAC in a pattern called ``encrypt-then-MAC''.
+In the encapsulation routine of the ``encrypt-then-MAC'' KEM construction, a random PKE plaintext $m$ is sampled and encrypted into some PKE ciphertext $c^\prime$; a symmetric key $k$, derived from hashing the plaintext, is used to compute a message authenticator $t$ on the PKE ciphertext $c^\prime$.
+The KEM ciphertext $c = (c^\prime, t)$ contains both the PKE ciphertext and the message authenticator.
+At decapsulation, the PKE ciphertext $c^\prime$ is decrypted into $\hat{m}$ first, then a symmetric key $\hat{k}$ is derived from hashing the decryption $\hat{m}$.
+The symmetric key is used to verify the authenticator against the PKE ciphertext.
+A illustration of the data flow in the ``encrypt-then-MAC'' KEM construction can be found in figure \ref{fig:etm-diagram}.
+A detailed description of the individual routines can be found in figure \ref{fig:etm-routines}.
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}
+        \tikzstyle{rect} = [draw, rectangle, minimum width=3cm, minimum height=1cm]
+        \tikzstyle{filledrect} = [draw, rectangle, minimum width=3cm, minimum height=1cm, fill=lightgray]
+        \matrix [column sep=7mm, row sep=5mm] {
+            \node (pk) [rect] {$\pk$}; &
+            \node (m) [rect] {$m \leftsample \mathcal{M}$}; &
+            \node (hashg) [rect] {Hash function $G$}; \\
+            \node (pkeenc) [rect] {$\texttt{PKE.Enc}$}; &
+            \node (mackey) [rect] {$k_\mac$}; \\
+            \node (ct) [filledrect] {$c^\prime$}; &
+            \node (mac) [rect] {$\mac$}; &
+            \node (tag) [filledrect] {$t$}; \\
+        };
+        \draw[->, thick] (pk) -- (pkeenc);
+        \draw[->, thick] (m) -- (pkeenc);
+        \draw[->, thick] (m) -- (hashg);
+        \draw[->, thick] (hashg) -- (mackey);
+        \draw[->, thick] (pkeenc) -- (ct);
+        \draw[->, thick] (mackey) -- (mac);
+        \draw[->, thick] (ct) -- (mac);
+        \draw[->, thick] (mac) -- (tag);
+    \end{tikzpicture}
+
+    \caption{Combining PKE with MAC using ``encrypt-then-MAC'' to encapsulate a shared secret. 
+    The returned values are colored grey
+    }\label{fig:etm-diagram}
+\end{figure}
+
+\subsubsection{Security reduction}
+By authenticating the PKE ciphertext using a symmetric key derived from its decryption, the ``encrypt-then-MAC'' construction greatly limits the amount of information an adversary can learn from a chosen-ciphertext attack.
+For an adversary to produce a valid KEM ciphertext, it must produce a valid message authenticator for the PKE ciphertext under the correct symmetric key.
+Under the random oracle model, the symmetric key is indistinguishable from uniformly random unless the adversary also knows the corresponding plaintext completely, thus rendering the output of the decasulation oracle redundant.
+Furthermore, because each encapsulation will pick a random message, within a reasonable lifetime of the KEM keypair, each KEM ciphertext will be authenticated using a distinct symmetric key, so the choice of MAC is only required to be one-time secure, affording us additional performance gains by choosing more efficient MAC instances.
+In section \ref{sec:main-results}, we prove that the IND-CCA2 security of the ``encrypt-then-MAC'' KEM reduces tightly to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC under the random oracle model.
+
+One-wayness under plaintext-checking attack (OW-PCA)\cite{DBLP:conf/ctrsa/OkamotoP01} requires that no efficient adversary can recover some random encryption even with access to a plaintext-checking oracle, who accepts a plaintext-ciphertext pair $(m, c)$ and returns whether the queried ciphertext decrypts to the queried plaintext.
+While it is not a standard security notion, it is useful when reducing the security of the ``encrypt-then-MAC'' KEM to the security of the PKE, and there are many well studied cryptosystems that are known to be OW-PCA: the RSA cryptosystem \cite{DBLP:journals/cacm/RivestSA78} is known to be a trapdoor permutation and thus trivially OW-PCA secure; the ElGamal cryptosystem  \cite{DBLP:journals/tit/Elgamal85} is conjectured to be OW-PCA secure under the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01}.
 
 \subsubsection{Performance improvements}
-The main advantage of our KEM construction over the Fujisaki-Okamoto transformation is the performance gains: our construction replaces re-encryption with computing a symmetric message authenticator, which is significantly faster, especially for one-time MAC.
-At the cost of computing an authenticator in encapsulation and increasing the ciphertext size by a message authenticator, our construction speeds up decapsulation by 10x.
-
+The main advantage of our KEM construction over the Fujisaki-Okamoto transformation is the performance gains: our construction replaces re-encryption with computing a message authenticator, which is significantly faster.
 When instantiated with the underlying PKE routines of ML-KEM and the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
 
+\begin{table}[H]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
+        & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
+        & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
+        \\
+        \hline
+        Encap \newline (ccl/tick) 
+        & 91467 & 93157 \newline (+1.8\%) 
+        & 136405 & 146405 \newline (+7.3\%) 
+        & 199185 & 205763 \newline (+3.3\%) 
+        \\
+        \hline
+        Decap \newline (ccl/tick) 
+        & 121185 & 33733 \newline (-72.2\%) 
+        & 186445 & 43315 \newline (-76.8\%) 
+        & 246245 & 51375 \newline (-79.1\%) 
+        \\
+        \hline
+        CT size \newline (bytes) 
+        & 768 & 784 \newline (+2.1\%) 
+        & 1088 & 1104 \newline (+1.5\%) 
+        & 1568 & 1584 \newline (+1.0\%) 
+        \\
+        \hline
+    \end{tabular}
+
+    \caption{$\text{ML-KEM}^+$ is instantiated with Poly1305}
+\end{table}
+
 We also implemented and measured the round trip time of key exchange protocols with various modes of authentication.
-When compared to ML-KEM, $\text{ML-KEM}^+$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange, 29\%-35\% reduction in unilaterally authenticated key exchange, and 35\%-48\% reduction in mutually authenticated key exchange.
+When compared to ML-KEM, $\text{ML-KEM}^+$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% reduction in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
+        & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
+        & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
+        \\
+        \hline
+        KE RTT \newline ($\us$) 
+        & 92 & 70 \newline (-23.9\%) 
+        & 135 & 99 \newline (-26.7\%) 
+        & 193 & 138 \newline (-28.5\%) 
+        \\
+        \hline
+        UAKE RTT \newline ($\us$) 
+        & 145 & 103 \newline (-29.0\%) 
+        & 215 & 144 \newline (-33.0\%) 
+        & 310 & 202 \newline (-34.8\%) 
+        \\
+        \hline
+        AKE RTT \newline ($\us$) 
+        & 220 & 133 \newline (-39.5\%) 
+        & 294 & 190 \newline (-35.4\%) 
+        & 512 & 266 \newline (-48.0\%) 
+        \\
+        \hline
+    \end{tabular}
+
+    \caption{Key exchange round-trip times}
+\end{table}
 
 \subsection{Related works}
-A similar construction was proposed by Abdulla et al \cite{DBLP:conf/ctrsa/AbdallaBR01}, though our construction builds a KEM instead of a HPKE, and our construction is generic over all OW-PCA secure PKE.
+As mentioned at the beginning of section \ref{sec:introduction}, there is a large volume of previous proposals that combine cryptographic primitives with weak security properties into cryptographic primitives with strong security properties.
+Besides the Fujisaki-Okamoto transformation, another transformation that sees substantial adoption is \emph{Optimal Asymmetric Encryption Padding (OAEP)} \cite{bellare1995optimal}.
+Unfortunately, OAEP can only be applied to one-way trapdoor permutation, a rare occurrence.
+To this day, RSA remains the only viable candidate \cite{DBLP:conf/crypto/FujisakiOPS01}, though RSA-OAEP has been widely adopted after its standardization in PKCS\#1 v2.2 \cite{rfc8017}.
+
+Securing an encryption scheme against chosen-ciphertext attack was a well-studied problem in the context of symmetric cryptography.
+Among the three generic patterns ``encrypt-then-MAC'', ``encrypt-and-MAC'', ``MAC-then-encrypt'', only the first one is formally analyzed and proved to be secure with any combination of symmetric cipher and MAC \cite{DBLP:conf/crypto/Krawczyk01}\cite{DBLP:conf/eurocrypt/CanettiK01}.
+``MAC-then-encrypt'' is only secure with specific choices of symmetric cipher and MAC, and ``encrypt-and-MAC'' is generally not semantically secure since most MAC will leak information about the message it authenticates.
+Today's most well-established symmetric encryption schemes, including AES-GCM and ChaCha20-Poly1305, largely follow the pattern of ``encrypt-then-MAC''.
+
+The idea of using a message authenticator to ensure ciphertext integrity was instantiated with ElGamal cryptosystem by Bellare and Rogaway in 1997 \cite{DBLP:conf/icics/BellareR97}, and its security was later reduced to the Gap Diffie-Hellman assumption by Abdulla, Bellare, and Rogaway in \cite{DBLP:conf/ctrsa/AbdallaBR01} in 2001.
+Our work can be seen as generalizing Bellare and Rogaway's scheme from ElGamal cryptosystem to all PKE with OW-PCA security.
 
@@
 \subsection{Public-key encryption scheme}
-\textbf{\textit{Syntax.}} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
+\textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
 $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
@@
 
-\textbf{\textit{Correctness.}} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04} and \cite{hofheinz2017modular}, a $\pke$ is $\delta$-correct if:
+\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04} and \cite{hofheinz2017modular}, a $\pke$ is $\delta$-correct if:
 
@@
 
-\emph{Security.} We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
+\textbf{Security.} We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
 Specifically, we first define the $\texttt{OW-ATK}$ as they pertain to a public key encryption scheme.

### 2024-10-04 etm-transform/manuscript/main.tex
Minor fixes for the body

@@
 Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$.
-For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem (see table \ref{tbl:ml-kem-dec-failure-probs}).
-
-\begin{table}[H]
-    \centering
-
-    \begin{tabular}{|c|c|c|c|}
-        \hline
-        & ML-KEM-512 & ML-KEM-768 & ML-KEM-1024 \\
-        \hline
-        $\delta$ & ??? & ??? & ??? \\
-        \hline
-    \end{tabular}
-    \caption{Estimated probability of decryption failure in ML-KEM}\label{tbl:ml-kem-dec-failure-probs}
-\end{table}
-
+For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem.
 On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm.
@@
             \begin{algorithmic}[1]
-                \Ensure $\pk$ is some $\pke$ public key
                 \State $m \leftsample \mathcal{M}$
@@
             \begin{algorithmic}[1]
-                \Require $c = c^\prime \Vert t, \sk = sk^\prime \Vert z$
-                \Ensure $c^\prime$ is some $\pke$ ciphertext
-                \Ensure $t$ is some $\mac$ tag
-                \Ensure $\sk^\prime$ is some $\pke$ secret key
-                \Ensure $z$ is some $\pke$ plaintext
                 \State $(c^\prime, t) \leftarrow c$
@@
 
-A detailed description of $\texttt{K-PKE}$'s routines can be found in FIPS 203 (TODO: citation). The ``encrypt-then-MAC'' routines are described in figure \ref{fig:ml-kem-plus-routines}.
+We refer readers to \cite{FIPS203} for the details of the \texttt{K-PKE} routines. The ``encrypt-then-MAC'' KEM routines are described in figure \ref{fig:ml-kem-plus-routines} below.
 
@@
 The final output is XOR'd with a sub key derived from the symmetric key, before being encrypted for one last time.
-A summary of the computation can be found in figure \ref{fig:aes-256-cmac}
-
-\begin{figure}[h]
-    \centering
-    \begin{minipage}[t]{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption*{Sub-key derivation}
-            \begin{algorithmic}[1]
-                \Require 256-bit key $k$
-                \Require $\texttt{const\_Rb = 0x87}$
-                \State $l \leftarrow \texttt{AES-256}(k, 0^{128})$
-                \If{$\texttt{MostSignificantBit}(l) = 0$}
-                    \State $k_1 \leftarrow \texttt{l {<}{<} 1}$
-                \Else 
-                    \State $k_1 \leftarrow \texttt{l {<}{<} 1} \oplus \texttt{const\_Rb}$
-                \EndIf
-                \If{$\texttt{MostSignificantBit}(k_1) = 0$}
-                    \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1}$
-                \Else 
-                    \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1} \oplus \texttt{const\_Rb}$
-                \EndIf
-                \State \Return $k_1, k_2$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption*{CMAC(k, m)}
-            \begin{algorithmic}[1]
-                \Require 256-bit symmetric key $k$
-                \State $(k_1, k_2) \leftarrow \texttt{deriveSubKey}(k)$
-                \State $n \leftarrow \left\lceil \texttt{bytesLen}(m) / 16\right\rceil$
-                \If{$n = 0$}
-                    \State $n \leftarrow 1$
-                    \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
-                \ElsIf{$\texttt{bytesLen(m)} \mod 16 = 0$}
-                    \State $m_\texttt{last} \leftarrow m_n \oplus k_1$
-                \Else 
-                    \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
-                \EndIf
-
-                \State $x = 0^{128}$
-                \For{$i \in \{1, 2, \ldots, n-1\}$}
-                    \State $y \leftarrow x \oplus m_i$
-                    \State $x \leftarrow \texttt{AES-256}(k, y)$
-                \EndFor
-                \State $y \leftarrow m_\texttt{last} \oplus x$
-                \State $t \leftarrow \texttt{AES-256}(k, y)$
-                \State \Return $t$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{AES-256 CMAC}\label{fig:aes-256-cmac}
-\end{figure}
+%A summary of the computation can be found in figure \ref{fig:aes-256-cmac}
+
+% \begin{figure}[h]
+%     \centering
+%     \begin{minipage}[t]{0.49\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{Sub-key derivation}
+%             \begin{algorithmic}[1]
+%                 \Require 256-bit key $k$
+%                 \Require $\texttt{const\_Rb = 0x87}$
+%                 \State $l \leftarrow \texttt{AES-256}(k, 0^{128})$
+%                 \If{$\texttt{MostSignificantBit}(l) = 0$}
+%                     \State $k_1 \leftarrow \texttt{l {<}{<} 1}$
+%                 \Else 
+%                     \State $k_1 \leftarrow \texttt{l {<}{<} 1} \oplus \texttt{const\_Rb}$
+%                 \EndIf
+%                 \If{$\texttt{MostSignificantBit}(k_1) = 0$}
+%                     \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1}$
+%                 \Else 
+%                     \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1} \oplus \texttt{const\_Rb}$
+%                 \EndIf
+%                 \State \Return $k_1, k_2$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}\hfill
+%     \begin{minipage}[t]{0.49\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{CMAC(k, m)}
+%             \begin{algorithmic}[1]
+%                 \Require 256-bit symmetric key $k$
+%                 \State $(k_1, k_2) \leftarrow \texttt{deriveSubKey}(k)$
+%                 \State $n \leftarrow \left\lceil \texttt{bytesLen}(m) / 16\right\rceil$
+%                 \If{$n = 0$}
+%                     \State $n \leftarrow 1$
+%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
+%                 \ElsIf{$\texttt{bytesLen(m)} \mod 16 = 0$}
+%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_1$
+%                 \Else 
+%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
+%                 \EndIf
+
+%                 \State $x = 0^{128}$
+%                 \For{$i \in \{1, 2, \ldots, n-1\}$}
+%                     \State $y \leftarrow x \oplus m_i$
+%                     \State $x \leftarrow \texttt{AES-256}(k, y)$
+%                 \EndFor
+%                 \State $y \leftarrow m_\texttt{last} \oplus x$
+%                 \State $t \leftarrow \texttt{AES-256}(k, y)$
+%                 \State \Return $t$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}
+%     \caption{AES-256 CMAC}\label{fig:aes-256-cmac}
+% \end{figure}
 
@@
 The measurements are summarized in table \ref{tbl:mac-performance}.
+
+From our testing, we found Poly1305 to exhibit the best performance characteristics.
+However, there are additional security considerations that may require the use of other less efficient MAC instances.
+For example, it is possible for an adversary with large computing infrastructure or quantum computers to pre-compute a large lookup table mapping symmetric key to the source plaintext.
+Upon receiving a ciphertext, the adversary can then search through the lookup table for a matching key, which would've revealed the corresponding decryption.
+We partially mitigated such attack by deriving the symmetric key from both the public key and the plaintext, but in case of a long-term keypair, the adversary might still be able to compute a large lookup table AFTER obtaining the long-term public key.
+Further mitigation could include using larger-size keys, which can be accomplished either by using a Carter-Wegman MAC that operates on a larger finite field or using a MAC with a variable key-length such as KMAC.
 
@@
 Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the ``encrypt-then-MAC'' transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
-A comparison between ML-KEM and variations of the ML-KEM-ETM can be found in table \ref{tbl:kem-performance}
+A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in table \ref{tbl:kem-performance}
 
@@
         $\sk$ size & 1632 \\
-        \texttt{ct} size & 768 \\
+        \text{ct} size & 768 \\
         \hline
@@
         \hline
-        \texttt{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
-        \hline
-        \texttt{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
+        \text{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
+        \hline
+        \text{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
         \hline
         % ML-KEM+-512
-        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
+        $\text{ML-KEM}^+$-512 w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
+        \hline
+        $\text{ML-KEM}^+$-512 w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
+        \hline
+        $\text{ML-KEM}^+$-512 w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
+        \hline
+        $\text{ML-KEM}^+$-512 w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
         \hline
@@
         $\sk$ size & 2400 \\
-        \texttt{ct} size & 1088 \\
+        \text{ct} size & 1088 \\
         \hline
@@
         \hline
-        \texttt{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
-        \hline
-        \texttt{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
+        \text{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
+        \hline
+        \text{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
+        \hline
+        $\text{ML-KEM}^+$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
+        \hline
+        $\text{ML-KEM}^+$-768 w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
+        \hline
+        $\text{ML-KEM}^+$-768 w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
+        \hline
+        $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
         \hline
@@
         $\sk$ size & 3168 \\
-        \texttt{ct} size & 1568 \\
+        \text{ct} size & 1568 \\
         \hline
@@
         \hline
-        \texttt{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
-        \hline
-        \texttt{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
+        \text{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
+        \hline
+        \text{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
+        \hline
+        $\text{ML-KEM}^+$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
+        \hline
+        $\text{ML-KEM}^+$-1024 w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
+        \hline
+        $\text{ML-KEM}^+$-1024 w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
+        \hline
+        $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
         \hline

### 2024-10-04 etm-transform/manuscript/main.tex
Open problems

@@
 %%%% 4. TITLE %%%%
-\title{
-Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''}
+\title{Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''}
 %%%% NOTES:
@@
 However, the Fujisaki-Okamoto transformation is not perfect.
-As pointed out by the authors of \cite{hofheinz2017modular}, the Fujisaki-Okamoto transformation uses \emph{de-randomization} and \emph{re-encryption} to ensure ciphertext integrity, which brings two problems: \begin{itemize}
+It uses \emph{de-randomization} and \emph{re-encryption} to achieve \emph{rigidity} \cite{bernstein2018towards} which then ensures ciphertext integrity.
+This brings two problems: \begin{itemize}
     \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
@@
 \subsection{Choosing a message authenticator}
-For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{kelsey2016sha}.
-
-Poly1305 and GMAC are both Carter-Wegman style authenticators that compute the tag using finite field arithmetic.
+For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
+
+Poly1305 and GMAC are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic.
 Generically speaking, Carter-Wegman MAC is parameterized by some finite field $\mathbb{F}$ and the maximal message length $L > 0$.
@@
 
-KMAC is defined in NIST SP 800-185 to be based on the family of sponge functions with Keccak permutaiton as the underlying function.
+KMAC is defined in \cite{SP80053r4} to be based on the family of sponge functions with Keccak permutaiton as the underlying function.
 We chose KMAC-256, which uses Shake256 as the underlying extendable output functions.
@@
 
-%TODO: Two tables, one-time MAC and many-time MAC, separate encap and decap in 2 tables
-%TODO: Look at OpenSSL for GMAC with no AES-256.
-
 \section{Conclusions and future works}\label{sec:future-works}
@@
 
-% TODO: write about open problems
+\textbf{Other suitable post-quantum instantiation.} In section \ref{sec:implementation}, we instantiated the ``encrypt-then-MAC'' KEM transformation with the PKE subroutines of ML-KEM.
+Unfortunately, ML-KEM is vulnerable to plaintext checking attacks that can recover complete secret key if the keypair is reused \cite{DBLP:conf/eurocrypt/BaetuDHTV19}.
+In fact, as Chris Peikert pointed out, because of the search-decision equivalence of LWE problems, almost all LWE-based cryptosystems will be vulerable to plaintext-checking attacks \cite{DBLP:conf/pqcrypto/Peikert14}.
+This naturally raises the question of finding other post-quantum primitives to instantiate ``encrypt-then-MAC'' with, such as other lattice-based cryptosystems not based on LWE, code-based cryptosystems and isogeny-based cryptosystems.
 

### 2024-10-04 etm-transform/manuscript/main.tex
Do not commit references in pdf

@@
 Subsequent works \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}\cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE.
-This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.
+This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.
 
@@
 For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table.
-We refer readers to \cite{avanzi2019crystals} and \cite{bos2018crystals} for the details.
+We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for the details.
 
@@
 A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel.
-\cite{bos2018crystals} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE).
+\cite{DBLP:conf/eurosp/BosDKLLSSSS18} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE).
 We instantiated an implementation for each of the three key exchange protocols using different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with implementations instantiated using ML-KEM.

### 2024-10-07 etm-transform/manuscript/main.tex
Angle the paper differently

@@
 %%%% 7. PAPER CONTENT %%%%
-\section{Introduction}\label{sec:introduction}
-Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
-The combination of KEM and some data encapsulation mechanism (DEM), such as AES-GCM and ChaCha20-Poly1305, is the foundation of many of today's most popular secure communication protocols such as Transport Layer Security (TLS) and Secure Shell (SSH).
+% \section{Introduction}\label{sec:introduction}
+% Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
+The combination of KEM and some data encapsulation mechanism (DEM), such as AES-GCM and ChaCha20-Poly1305, is the foundation of many of today's most widely adopted communication protocols such as Transport Layer Security (TLS) and Secure Shell (SSH).
 The commonly accepted security standard of a KEM is \emph{Indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}: no efficient adversary can distinguish a shared secret obtained by running the encapsulation routine from random noise, even with access to a decapsulation oracle throughout the attack.
@@
 
-It is difficult to build a provably IND-CCA2 secure KEM from scratch.
+% It is difficult to build a provably IND-CCA2 secure KEM from scratch.
 Instead, secure KEMs are usually built on top of a PKE with weaker security property (e.g.
@@
 
-The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
+% The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
 It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
@@
 This brings two problems: \begin{itemize}
-    \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
-    \item \textbf{side-channel vulnerability:} running the input PKE's encryption routine in the output KEM's decapsulation routine introduces risk of side-channel vulnerabilities not found in the input PKE's decryption routine alone.
+%     \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
+%     \item \textbf{side-channel vulnerability:} running the input PKE's encryption routine in the output KEM's decapsulation routine introduces risk of side-channel vulnerabilities not found in the input PKE's decryption routine alone.
 In fact, many practical attacks \cite{ueno2022curse}\cite{ravi2019generic} exploit re-encryption to decrypt ciphertext or recover secret keys.
 Countermeasures such as masking have been proposed to address these side channels, but they inevitably carry substantial performance penalty.
-\end{itemize}
-
-\subsection{Our contributions}
-Our main contribution is a novel generic KEM construction that combines a PKE with weak security and a message authentication code (MAC).
+% \end{itemize}
+
+% \subsection{Our contributions}
+% Our main contribution is a novel generic KEM construction that combines a PKE with weak security and a message authentication code (MAC).
 We were inspired by the strategy used in symmetric cyrptography for achieving IND-CCA2 security: combining a symmetric cipher with a MAC in a pattern called ``encrypt-then-MAC''.
@@
 A detailed description of the individual routines can be found in figure \ref{fig:etm-routines}.
+
+% \begin{figure}
+%     \centering
+%     \begin{tikzpicture}
+%         \tikzstyle{rect} = [draw, rectangle, minimum width=3cm, minimum height=1cm]
+%         \tikzstyle{filledrect} = [draw, rectangle, minimum width=3cm, minimum height=1cm, fill=lightgray]
+%         \matrix [column sep=7mm, row sep=5mm] {
+%             \node (pk) [rect] {$\pk$}; &
+%             \node (m) [rect] {$m \leftsample \mathcal{M}$}; &
+%             \node (hashg) [rect] {Hash function $G$}; \\
+%             \node (pkeenc) [rect] {$\texttt{PKE.Enc}$}; &
+%             \node (mackey) [rect] {$k_\mac$}; \\
+%             \node (ct) [filledrect] {$c^\prime$}; &
+%             \node (mac) [rect] {$\mac$}; &
+%             \node (tag) [filledrect] {$t$}; \\
+%         };
+%         \draw[->, thick] (pk) -- (pkeenc);
+%         \draw[->, thick] (m) -- (pkeenc);
+%         \draw[->, thick] (m) -- (hashg);
+%         \draw[->, thick] (hashg) -- (mackey);
+%         \draw[->, thick] (pkeenc) -- (ct);
+%         \draw[->, thick] (mackey) -- (mac);
+%         \draw[->, thick] (ct) -- (mac);
+%         \draw[->, thick] (mac) -- (tag);
+%     \end{tikzpicture}
+
+%     \caption{Combining PKE with MAC using ``encrypt-then-MAC'' to encapsulate a shared secret. 
+%     The returned values are colored grey
+%     }\label{fig:etm-diagram}
+% \end{figure}
+
+% \subsubsection{Security reduction}
+% By authenticating the PKE ciphertext using a symmetric key derived from its decryption, the ``encrypt-then-MAC'' construction greatly limits the amount of information an adversary can learn from a chosen-ciphertext attack.
+For an adversary to produce a valid KEM ciphertext, it must produce a valid message authenticator for the PKE ciphertext under the correct symmetric key.
+Under the random oracle model, the symmetric key is indistinguishable from uniformly random unless the adversary also knows the corresponding plaintext completely, thus rendering the output of the decasulation oracle redundant.
+Furthermore, because each encapsulation will pick a random message, within a reasonable lifetime of the KEM keypair, each KEM ciphertext will be authenticated using a distinct symmetric key, so the choice of MAC is only required to be one-time secure, affording us additional performance gains by choosing more efficient MAC instances.
+In section \ref{sec:main-results}, we prove that the IND-CCA2 security of the ``encrypt-then-MAC'' KEM reduces tightly to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC under the random oracle model.
+
+% One-wayness under plaintext-checking attack (OW-PCA)\cite{DBLP:conf/ctrsa/OkamotoP01} requires that no efficient adversary can recover some random encryption even with access to a plaintext-checking oracle, who accepts a plaintext-ciphertext pair $(m, c)$ and returns whether the queried ciphertext decrypts to the queried plaintext.
+While it is not a standard security notion, it is useful when reducing the security of the ``encrypt-then-MAC'' KEM to the security of the PKE, and there are many well studied cryptosystems that are known to be OW-PCA: the RSA cryptosystem \cite{DBLP:journals/cacm/RivestSA78} is known to be a trapdoor permutation and thus trivially OW-PCA secure; the ElGamal cryptosystem  \cite{DBLP:journals/tit/Elgamal85} is conjectured to be OW-PCA secure under the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01}.
+
+% \subsubsection{Performance improvements}
+% The main advantage of our KEM construction over the Fujisaki-Okamoto transformation is the performance gains: our construction replaces re-encryption with computing a message authenticator, which is significantly faster.
+When instantiated with the underlying PKE routines of ML-KEM and the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
+
+% \begin{table}[H]
+%     \centering
+%     \footnotesize
+
+%     \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+%         \hline
+%         & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
+%         & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
+%         & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
+%         \\
+%         \hline
+%         Encap \newline (ccl/tick) 
+%         & 91467 & 93157 \newline (+1.8\%) 
+%         & 136405 & 146405 \newline (+7.3\%) 
+%         & 199185 & 205763 \newline (+3.3\%) 
+%         \\
+%         \hline
+%         Decap \newline (ccl/tick) 
+%         & 121185 & 33733 \newline (-72.2\%) 
+%         & 186445 & 43315 \newline (-76.8\%) 
+%         & 246245 & 51375 \newline (-79.1\%) 
+%         \\
+%         \hline
+%         CT size \newline (bytes) 
+%         & 768 & 784 \newline (+2.1\%) 
+%         & 1088 & 1104 \newline (+1.5\%) 
+%         & 1568 & 1584 \newline (+1.0\%) 
+%         \\
+%         \hline
+%     \end{tabular}
+
+%     \caption{$\text{ML-KEM}^+$ is instantiated with Poly1305}
+% \end{table}
+
+% We also implemented and measured the round trip time of key exchange protocols with various modes of authentication.
+When compared to ML-KEM, $\text{ML-KEM}^+$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% reduction in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+
+% \begin{table}[H]
+%     \centering
+%     \footnotesize
+
+%     \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+%         \hline
+%         & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
+%         & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
+%         & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
+%         \\
+%         \hline
+%         KE RTT \newline ($\us$) 
+%         & 92 & 70 \newline (-23.9\%) 
+%         & 135 & 99 \newline (-26.7\%) 
+%         & 193 & 138 \newline (-28.5\%) 
+%         \\
+%         \hline
+%         UAKE RTT \newline ($\us$) 
+%         & 145 & 103 \newline (-29.0\%) 
+%         & 215 & 144 \newline (-33.0\%) 
+%         & 310 & 202 \newline (-34.8\%) 
+%         \\
+%         \hline
+%         AKE RTT \newline ($\us$) 
+%         & 220 & 133 \newline (-39.5\%) 
+%         & 294 & 190 \newline (-35.4\%) 
+%         & 512 & 266 \newline (-48.0\%) 
+%         \\
+%         \hline
+%     \end{tabular}
+
+%     \caption{Key exchange round-trip times}
+% \end{table}
+
+% \subsection{Related works}
+% As mentioned at the beginning of section \ref{sec:introduction}, there is a large volume of previous proposals that combine cryptographic primitives with weak security properties into cryptographic primitives with strong security properties.
+Besides the Fujisaki-Okamoto transformation, another transformation that sees substantial adoption is \emph{Optimal Asymmetric Encryption Padding (OAEP)} \cite{bellare1995optimal}.
+Unfortunately, OAEP can only be applied to one-way trapdoor permutation, a rare occurrence.
+To this day, RSA remains the only viable candidate \cite{DBLP:conf/crypto/FujisakiOPS01}, though RSA-OAEP has been widely adopted after its standardization in PKCS\#1 v2.2 \cite{rfc8017}.
+
+% Securing an encryption scheme against chosen-ciphertext attack was a well-studied problem in the context of symmetric cryptography.
+Among the three generic patterns ``encrypt-then-MAC'', ``encrypt-and-MAC'', ``MAC-then-encrypt'', only the first one is formally analyzed and proved to be secure with any combination of symmetric cipher and MAC \cite{DBLP:conf/crypto/Krawczyk01}\cite{DBLP:conf/eurocrypt/CanettiK01}.
+``MAC-then-encrypt'' is only secure with specific choices of symmetric cipher and MAC, and ``encrypt-and-MAC'' is generally not semantically secure since most MAC will leak information about the message it authenticates.
+Today's most well-established symmetric encryption schemes, including AES-GCM and ChaCha20-Poly1305, largely follow the pattern of ``encrypt-then-MAC''.
+
+% The idea of using a message authenticator to ensure ciphertext integrity was instantiated with ElGamal cryptosystem by Bellare and Rogaway in 1997 \cite{DBLP:conf/icics/BellareR97}, and its security was later reduced to the Gap Diffie-Hellman assumption by Abdulla, Bellare, and Rogaway in \cite{DBLP:conf/ctrsa/AbdallaBR01} in 2001.
+Our work can be seen as generalizing Bellare and Rogaway's scheme from ElGamal cryptosystem to all PKE with OW-PCA security.
+
+\section{Introduction}\label{sec:introduction}
+A key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
+The combination of KEM for key transport and data encapsulation mechanism (DEM), usually instantiated with authenticated encryption with associated data schemes, form the foundation of today's most widely adopted secure communication protocols such as Transpot Layer Security (TLS) and Secure Shell (SSH).
+
+The accepted security standard for a KEM is \emph{indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}.
+IND-CCA2 security requires that no efficient adversary, with access to a decapsulation oracle throughout the attack, can distinguish a pseudorandom shared secret from truly random noise.
+However, building an provably IND-CCA2 secure KEM from scratch is immensely difficult.
+Instead, the most viable approach is to start with a public-key encryption (PKE) scheme with weaker security properties (e.g.
+OW-CPA or IND-CPA), then put on additional checks for ensuring ciphertext integrity.
+
+One such generic transformation was proposed by Abdalla, Rogaway, and Bellare in 2001 \cite{DBLP:conf/ctrsa/AbdallaBR01}.
+In its original form, often referred to as ``Hashed ElGamal'', Abdalla's proposal is a hybrid public-key encryption (HPKE) scheme whose chosen-ciphertext security reduces to the strong Diffie-Hellman assumption (that no efficient adversary can violate the computational Diffie-Hellman assumption even with access to a decisional Diffie-Hellman oracle) under the random oracle model.
+
+\subsection{Our contribution}
+
+In this paper, we adapt the ``Hashed ElGamal'' construction to a generic KEM transformation built on top of a PKE, and reduces the IND-CCA2 security of the KEM to the OW-PCA security of the input PKE in the random oracle model.
+We called our construction the ``encrypt-then-MAC'' KEM transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
+A summary of the data flow in the ``encrypt-then-MAC'' KEM can be found in figure \ref{fig:etm-diagram}.
 
@@
 \subsubsection{Security reduction}
-By authenticating the PKE ciphertext using a symmetric key derived from its decryption, the ``encrypt-then-MAC'' construction greatly limits the amount of information an adversary can learn from a chosen-ciphertext attack.
-For an adversary to produce a valid KEM ciphertext, it must produce a valid message authenticator for the PKE ciphertext under the correct symmetric key.
-Under the random oracle model, the symmetric key is indistinguishable from uniformly random unless the adversary also knows the corresponding plaintext completely, thus rendering the output of the decasulation oracle redundant.
-Furthermore, because each encapsulation will pick a random message, within a reasonable lifetime of the KEM keypair, each KEM ciphertext will be authenticated using a distinct symmetric key, so the choice of MAC is only required to be one-time secure, affording us additional performance gains by choosing more efficient MAC instances.
-In section \ref{sec:main-results}, we prove that the IND-CCA2 security of the ``encrypt-then-MAC'' KEM reduces tightly to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC under the random oracle model.
-
-One-wayness under plaintext-checking attack (OW-PCA)\cite{DBLP:conf/ctrsa/OkamotoP01} requires that no efficient adversary can recover some random encryption even with access to a plaintext-checking oracle, who accepts a plaintext-ciphertext pair $(m, c)$ and returns whether the queried ciphertext decrypts to the queried plaintext.
-While it is not a standard security notion, it is useful when reducing the security of the ``encrypt-then-MAC'' KEM to the security of the PKE, and there are many well studied cryptosystems that are known to be OW-PCA: the RSA cryptosystem \cite{DBLP:journals/cacm/RivestSA78} is known to be a trapdoor permutation and thus trivially OW-PCA secure; the ElGamal cryptosystem  \cite{DBLP:journals/tit/Elgamal85} is conjectured to be OW-PCA secure under the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01}.
+From a high level, the ``encrypt-then-MAC'' KEM encapsulates a shared secret by encrypting a randomly sampled PKE plaintext.
+To ensure the integrity of the ciphertext, a message authenticator is computed on the PKE ciphertext, where the symmetric key is derived by hashing the random PKE plaintext.
+The PKE ciphertext and message authenticator combine into the KEM ciphertext.
+For an adversary to produce a valid KEM ciphertext, it must produce a valid tag.
+If the message authenticator is existentially unforgeable, then the adversary cannot produce a valid tag without knowing the symmetric key, and under the random oracle model, this implies that the adversary must also know the corresonding PKE plaintext, thus rendering the decapsulation oracle's output redundant.
+
+We also extended the discussion of security requirement for the MAC: because each call to the encapsulation routine will use a freshly sampled PKE plaintext, each PKE ciphertext will be signed with a distinct MAC key.
+In other words, the MAC only needs to be one-time existentially unforgeable, which opens up possibilities of using more efficient constructions than HMAC and CBC-MAC as mentioned in the original paper.
+
+In section \ref{sec:main-results}, we formally state the transformation and prove that the ``encrypt-then-MAC'' KEM is IND-CCA2 secure if the input PKE is OW-PCA secure and the input MAC is one-time existentially unforgeable.
 
 \subsubsection{Performance improvements}
-The main advantage of our KEM construction over the Fujisaki-Okamoto transformation is the performance gains: our construction replaces re-encryption with computing a message authenticator, which is significantly faster.
-When instantiated with the underlying PKE routines of ML-KEM and the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
+A major advantage of the ``encrypt-then-MAC'' construction is the low performance overhead added to the underlying PKE scheme.
+We instantiated the ``encrypt-then-MAC'' KEM with the underlying PKE routines of ML-KEM and compared its performance to ML-KEM, which uses the Fujisaki-Okamoto transformation.
+The Fujisaki-Okamoto transformation uses \emph{de-randomization} and \emph{re-encryption} to achieve chosen ciphertext security, which result in substantial performance penalty.
+When combined with the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
 
@@
 \subsection{Related works}
-As mentioned at the beginning of section \ref{sec:introduction}, there is a large volume of previous proposals that combine cryptographic primitives with weak security properties into cryptographic primitives with strong security properties.
-Besides the Fujisaki-Okamoto transformation, another transformation that sees substantial adoption is \emph{Optimal Asymmetric Encryption Padding (OAEP)} \cite{bellare1995optimal}.
-Unfortunately, OAEP can only be applied to one-way trapdoor permutation, a rare occurrence.
-To this day, RSA remains the only viable candidate \cite{DBLP:conf/crypto/FujisakiOPS01}, though RSA-OAEP has been widely adopted after its standardization in PKCS\#1 v2.2 \cite{rfc8017}.
-
-Securing an encryption scheme against chosen-ciphertext attack was a well-studied problem in the context of symmetric cryptography.
-Among the three generic patterns ``encrypt-then-MAC'', ``encrypt-and-MAC'', ``MAC-then-encrypt'', only the first one is formally analyzed and proved to be secure with any combination of symmetric cipher and MAC \cite{DBLP:conf/crypto/Krawczyk01}\cite{DBLP:conf/eurocrypt/CanettiK01}.
-``MAC-then-encrypt'' is only secure with specific choices of symmetric cipher and MAC, and ``encrypt-and-MAC'' is generally not semantically secure since most MAC will leak information about the message it authenticates.
-Today's most well-established symmetric encryption schemes, including AES-GCM and ChaCha20-Poly1305, largely follow the pattern of ``encrypt-then-MAC''.
-
-The idea of using a message authenticator to ensure ciphertext integrity was instantiated with ElGamal cryptosystem by Bellare and Rogaway in 1997 \cite{DBLP:conf/icics/BellareR97}, and its security was later reduced to the Gap Diffie-Hellman assumption by Abdulla, Bellare, and Rogaway in \cite{DBLP:conf/ctrsa/AbdallaBR01} in 2001.
-Our work can be seen as generalizing Bellare and Rogaway's scheme from ElGamal cryptosystem to all PKE with OW-PCA security.
-
-\section{Preliminaries and previous results}
+Optimal Asymmetric Encryption Padding (OAEP) \cite{bellare1995optimal}\cite{bellare1998relations} is a generic PKE construction whose IND-CCA2 security reduces to the one-wayness of the input trapdoor permutation under the random oracle model.
+OAEP enjoys tight security reduction and minimal performance overhead, but its requirement for a trapdoor permutation is difficult to satisfy.
+To this day, RSA remains the only practical candidate to instantiate OAEP with \cite{DBLP:journals/joc/Shoup02}\cite{DBLP:conf/crypto/FujisakiOPS01}, and RSA-OAEP saw widespread adoption in Internet communication protocols after its standardization in PKCS\#1 v2 \cite{rfc8017}.
+
+Originally proposed by Eiichiro Fujisaki and Tatsuaki Okamoto, the Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is a hybid PKE whose security reduces non-tightly to the OW-CPA security of the input PKE.
+Later works by Hofheinz, Hovelmann, and Kiltz \cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} made systematic improvements to the original proposal by providing a modular KEM construction whose security tightly reduces to the IND-CPA security of the input PKE.
+In addition, the authors made extensive security analysis with respect to decryption failure (a non-trivial security flaws in many lattice-based cryptosystems) and provided a non-tight security reduction in the quantum random oracle model (QROM), which made the modular Fujisaki-Okamoto KEM suitable for post-quantum cryptography.
+
+The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
+It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
+When Kyber was standardized by NIST in FIPS 203 ``Module-lattice key-encapsulation mechanism'' (ML-KEM) \cite{FIPS203}, it kept the Fujisaki-Okamoto transformation in its KEM construction.
+However, the Fujisaki-Okamoto transformation is not perfect.
+It uses \emph{de-randomization} and \emph{re-encryption} to achieve \emph{rigidity} \cite{bernstein2018towards} which then ensures ciphertext integrity.
+This brings two problems: \begin{itemize}
+    \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
+    \item \textbf{side-channel vulnerability:} running the input PKE's encryption routine in the output KEM's decapsulation routine introduces risk of side-channel vulnerabilities not found in the input PKE's decryption routine alone.
+In fact, many practical attacks \cite{ueno2022curse}\cite{ravi2019generic} exploit re-encryption to decrypt ciphertext or recover secret keys.
+Countermeasures such as masking have been proposed to address these side channels, but they inevitably carry substantial performance penalty.
+\end{itemize}
+
+% OAEP, REACT, GEM,
+
+\section{Preliminaries}
 

### 2024-10-08 etm-transform/manuscript/main.tex
angle more differently

@@
 \section{Introduction}\label{sec:introduction}
-A key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
-The combination of KEM for key transport and data encapsulation mechanism (DEM), usually instantiated with authenticated encryption with associated data schemes, form the foundation of today's most widely adopted secure communication protocols such as Transpot Layer Security (TLS) and Secure Shell (SSH).
-
-The accepted security standard for a KEM is \emph{indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}.
-IND-CCA2 security requires that no efficient adversary, with access to a decapsulation oracle throughout the attack, can distinguish a pseudorandom shared secret from truly random noise.
-However, building an provably IND-CCA2 secure KEM from scratch is immensely difficult.
-Instead, the most viable approach is to start with a public-key encryption (PKE) scheme with weaker security properties (e.g.
-OW-CPA or IND-CPA), then put on additional checks for ensuring ciphertext integrity.
-
-One such generic transformation was proposed by Abdalla, Rogaway, and Bellare in 2001 \cite{DBLP:conf/ctrsa/AbdallaBR01}.
-In its original form, often referred to as ``Hashed ElGamal'', Abdalla's proposal is a hybrid public-key encryption (HPKE) scheme whose chosen-ciphertext security reduces to the strong Diffie-Hellman assumption (that no efficient adversary can violate the computational Diffie-Hellman assumption even with access to a decisional Diffie-Hellman oracle) under the random oracle model.
+A key encapsulation mechanism (KEM) \cite{DBLP:journals/iacr/Shoup01} is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
+The desired security standard for a KEM is called indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2).
+Intuitively speaking, IND-CCA2 security requires that no efficient adversary can distinguish a pseudorandom shared secret from truly random noise, even with access to a decapsulation oracle throughout the attack.
+However, building a provably IND-CCA2 KEM is tremendously difficult.
+Early attempts without formal proof, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attack \cite{DBLP:conf/crypto/Bleichenbacher98}.
+In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to ensure ciphertext non-malleability \cite{bellare1998relations}.
+Some of the earliest proposals for generic IND-CCA2 secure constructions include OAEP \cite{bellare1995optimal}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.
+
+On the other hand, chosen-ciphertext security is a solve problem in symmetric cryptography.
+It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called ``encrypt-then-MAC'' \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA2 security.
+While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
+Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA2 security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
+The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypt the PKE plaintext, then authenticate the PKE ciphertext using the previously derived MAC key.
+Where the Gap Diffie-Hellman assumption holds and the MAC is existentially unforgeable, no efficient adversary can recover the decryption of an unknown ciphertext even with access to a decryption oracle because it cannot produce a valid tag for such unknown ciphertext.
 
 \subsection{Our contribution}
-
-In this paper, we adapt the ``Hashed ElGamal'' construction to a generic KEM transformation built on top of a PKE, and reduces the IND-CCA2 security of the KEM to the OW-PCA security of the input PKE in the random oracle model.
-We called our construction the ``encrypt-then-MAC'' KEM transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
-A summary of the data flow in the ``encrypt-then-MAC'' KEM can be found in figure \ref{fig:etm-diagram}.
+In this paper, we propose a generic IND-CCA2 secure KEM using techniques adapted from DHIES.
+We call our proposed scheme the ``encrypt-then-MAC'' transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
+A summary of the data flow in the ``encrypt-then-MAC'' transformation can be found in figure \ref{fig:etm-diagram}.
 
@@
 
-\subsubsection{Security reduction}
-From a high level, the ``encrypt-then-MAC'' KEM encapsulates a shared secret by encrypting a randomly sampled PKE plaintext.
-To ensure the integrity of the ciphertext, a message authenticator is computed on the PKE ciphertext, where the symmetric key is derived by hashing the random PKE plaintext.
-The PKE ciphertext and message authenticator combine into the KEM ciphertext.
-For an adversary to produce a valid KEM ciphertext, it must produce a valid tag.
-If the message authenticator is existentially unforgeable, then the adversary cannot produce a valid tag without knowing the symmetric key, and under the random oracle model, this implies that the adversary must also know the corresonding PKE plaintext, thus rendering the decapsulation oracle's output redundant.
-
-We also extended the discussion of security requirement for the MAC: because each call to the encapsulation routine will use a freshly sampled PKE plaintext, each PKE ciphertext will be signed with a distinct MAC key.
-In other words, the MAC only needs to be one-time existentially unforgeable, which opens up possibilities of using more efficient constructions than HMAC and CBC-MAC as mentioned in the original paper.
-
-In section \ref{sec:main-results}, we formally state the transformation and prove that the ``encrypt-then-MAC'' KEM is IND-CCA2 secure if the input PKE is OW-PCA secure and the input MAC is one-time existentially unforgeable.
+Our scheme mainly differs from DHIES in its versatility.
+Whereas the IND-CCA2 security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the ``encrypt-then-MAC'' KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security).
+In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the ``encrypt-then-MAC'' KEM can be instantiated with one-time secure MAC such as Poly1305, which greatly improves the computational efficiency of the scheme.
 
 \subsubsection{Performance improvements}
+In section \ref{sec:implementation}, we compare the computational performance in a variety of scenarios between ``encrypt-then-MAC'' KEM to the modular Fujisaki-Okamoto transformation \cite{DBLP:conf/tcc/HofheinzHK17}.
 A major advantage of the ``encrypt-then-MAC'' construction is the low performance overhead added to the underlying PKE scheme.

### 2024-10-08 etm-transform/manuscript/main.tex
rewrote introduction

@@
 \newcommand{\us}{\mu s}
+\def\mlkemplus{\text{ML-KEM}^+}
 
@@
 Whereas the IND-CCA2 security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the ``encrypt-then-MAC'' KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security).
-In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the ``encrypt-then-MAC'' KEM can be instantiated with one-time secure MAC such as Poly1305, which greatly improves the computational efficiency of the scheme.
+In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the ``encrypt-then-MAC'' KEM can be instantiated with one-time secure MAC such as Poly1305 (Abdalla et al originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
 
 \subsubsection{Performance improvements}
-In section \ref{sec:implementation}, we compare the computational performance in a variety of scenarios between ``encrypt-then-MAC'' KEM to the modular Fujisaki-Okamoto transformation \cite{DBLP:conf/tcc/HofheinzHK17}.
-A major advantage of the ``encrypt-then-MAC'' construction is the low performance overhead added to the underlying PKE scheme.
-We instantiated the ``encrypt-then-MAC'' KEM with the underlying PKE routines of ML-KEM and compared its performance to ML-KEM, which uses the Fujisaki-Okamoto transformation.
-The Fujisaki-Okamoto transformation uses \emph{de-randomization} and \emph{re-encryption} to achieve chosen ciphertext security, which result in substantial performance penalty.
-When combined with the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
+In section \ref{sec:implementation}, we instantiate an instance of ``encrypt-then-MAC'' KEM using the PKE subroutines defined in ML-KEM \cite{FIPS203}.
+For clarity's sake we call this instantiation $\mlkemplus$.
+Compared with the full KEM routine in ML-KEM, which is based on some variant of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates enormous runtime efficiency gain in decapsulation with only minimal performance penalty in encapsulation runtime and ciphertext size.
+On an AMD EPYC 9R14 CPU, $\mlkemplus$ (combined with Poly1305) achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation.
+See table \ref{tbl:cpu-cycles-summary} for a summary of CPU cycles needed for encapsulation and decapsulation in each security level.
 
@@
 
-    \caption{$\text{ML-KEM}^+$ is instantiated with Poly1305}
+    \caption{$\text{ML-KEM}^+$ is instantiated with Poly1305}\label{tbl:cpu-cycles-summary}
 \end{table}
 
-We also implemented and measured the round trip time of key exchange protocols with various modes of authentication.
-When compared to ML-KEM, $\text{ML-KEM}^+$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% reduction in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+Since the decapsulation efficiency comes at the cost of increasing encryption runtime and ciphertext size, we also measured the round trip time of key exchange protocols with various modes of authentication.
+When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+See table \ref{tbl:rtt-summary} for a summary of round trip times.
 
@@
 
-    \caption{Key exchange round-trip times}
+    \caption{Key exchange round-trip times}\label{tbl:rtt-summary}
 \end{table}
@@
 \subsection{Related works}
-Optimal Asymmetric Encryption Padding (OAEP) \cite{bellare1995optimal}\cite{bellare1998relations} is a generic PKE construction whose IND-CCA2 security reduces to the one-wayness of the input trapdoor permutation under the random oracle model.
-OAEP enjoys tight security reduction and minimal performance overhead, but its requirement for a trapdoor permutation is difficult to satisfy.
-To this day, RSA remains the only practical candidate to instantiate OAEP with \cite{DBLP:journals/joc/Shoup02}\cite{DBLP:conf/crypto/FujisakiOPS01}, and RSA-OAEP saw widespread adoption in Internet communication protocols after its standardization in PKCS\#1 v2 \cite{rfc8017}.
-
-Originally proposed by Eiichiro Fujisaki and Tatsuaki Okamoto, the Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is a hybid PKE whose security reduces non-tightly to the OW-CPA security of the input PKE.
-Later works by Hofheinz, Hovelmann, and Kiltz \cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} made systematic improvements to the original proposal by providing a modular KEM construction whose security tightly reduces to the IND-CPA security of the input PKE.
-In addition, the authors made extensive security analysis with respect to decryption failure (a non-trivial security flaws in many lattice-based cryptosystems) and provided a non-tight security reduction in the quantum random oracle model (QROM), which made the modular Fujisaki-Okamoto KEM suitable for post-quantum cryptography.
-
-The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
+\textbf{Optimal Asymmetric Encryption Padding (OAEP)} \cite{DBLP:conf/eurocrypt/BellareR94}\cite{DBLP:conf/crypto/BellareDPR98} is a generic chosen-ciphertext secure PKE.
+Under the random oracle model, the chosen-ciphertext security of the OAEP encryption scheme reduces to the one-wayness of the input trapdoor permutation.
+Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA2 security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al later proved that the OAEP is IND-CCA2 secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
+RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended method of RSA-based encryption scheme.
+Unfortunately, OAEP's requirement for a trapdoor permutation is immensely difficult to satisfy, and no other practical instantiation saw widespread adoption to this day.
+
+The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic chosen-ciphertext secure transformation.
+While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA2 security reduces non-tightly to the OW-CPA security of the input PKE and the IND-CPA security of the symmetric encryption scheme.
+Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build KEM, and proved its security in the quantum random oracle model (QROM).
+
+The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversary.
 It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
-When Kyber was standardized by NIST in FIPS 203 ``Module-lattice key-encapsulation mechanism'' (ML-KEM) \cite{FIPS203}, it kept the Fujisaki-Okamoto transformation in its KEM construction.
+
 However, the Fujisaki-Okamoto transformation is not perfect.
-It uses \emph{de-randomization} and \emph{re-encryption} to achieve \emph{rigidity} \cite{bernstein2018towards} which then ensures ciphertext integrity.
-This brings two problems: \begin{itemize}
-    \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
-    \item \textbf{side-channel vulnerability:} running the input PKE's encryption routine in the output KEM's decapsulation routine introduces risk of side-channel vulnerabilities not found in the input PKE's decryption routine alone.
-In fact, many practical attacks \cite{ueno2022curse}\cite{ravi2019generic} exploit re-encryption to decrypt ciphertext or recover secret keys.
-Countermeasures such as masking have been proposed to address these side channels, but they inevitably carry substantial performance penalty.
+Among its shortfalls are two problems caused by its use of \emph{de-randomization} and \emph{re-encryption} for achieving chosen-ciphertext security.
+\begin{itemize}
+    \item \textbf{Computational inefficiency}.
+The decapsulation routine needs to re-encrypt the decryption to ensure ciphertext has not been tempered with.
+For input PKE whose encryption routine carries substantial computational cost, such as most lattice-based cryptosystems, re-encryption slows down decapsulation significantly.
+    \item \textbf{Side-channel vulnerability}.
+Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext.
+As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key in fewer than 1000 traces.
 \end{itemize}
-
-% OAEP, REACT, GEM,
 
@@
 
-\subsection{Choosing a message authenticator}
+\subsection{Choosing a message authenticator}\label{sec:choosing-a-message-authenticator}
 For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.

### 2024-10-08 etm-transform/manuscript/main.tex
Some words in conclusion

@@
 %
-% TODO: need more thorough references
-% TODO: need better conclusion
 % \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
@@
 Early attempts without formal proof, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attack \cite{DBLP:conf/crypto/Bleichenbacher98}.
-In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to ensure ciphertext non-malleability \cite{bellare1998relations}.
-Some of the earliest proposals for generic IND-CCA2 secure constructions include OAEP \cite{bellare1995optimal}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.
+In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to ensure ciphertext non-malleability \cite{DBLP:conf/asiacrypt/BellareN00}.
+Some of the earliest proposals for generic IND-CCA2 secure constructions include OAEP \cite{DBLP:conf/eurocrypt/BellareR94}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.
 
@@
 
-\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04} and \cite{hofheinz2017modular}, a $\pke$ is $\delta$-correct if:
+\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04}, a $\pke$ is $\delta$-correct if:
 
@@
 First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries.
-Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation \cite{hofheinz2017modular} is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines.
+Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines.
 Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.
@@
 The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
-See appendix \ref{sec:ml-kem-and-kyber} for comparison with Kyber's third round submission to NIST's PQC competition.
 
@@
 
-% TODO: How is ML-KEM different from Kyber round 3?
-
 \begin{table}[h]
@@
         \hline
-        % TODO: this title needs to show that parameters are identical across all variants
         \multicolumn{2}{|c|}{\bf 128-bit security} \\
@@
 \section{Conclusions and future works}\label{sec:future-works}
-The ``encrypt-then-MAC'' transformation is a generic KEM construction that achieves IND-CCA2 security under the random oracle model if the input PKE is OW-PCA secure.
-Compared to the Fujisaki-Okamoto transformation, our construction replaced \emph{de-randomization} and \emph{re-encryption} with a message authenticator.
-At the cost of some minimal increase in communication size and encapsulation runtime, our construction achieves significant efficiency gains in the decapsulation routine.
-In practical key exchange protocols, our construction saves between 35-45\% in round trip time.
-
-\textbf{Other suitable post-quantum instantiation.} In section \ref{sec:implementation}, we instantiated the ``encrypt-then-MAC'' KEM transformation with the PKE subroutines of ML-KEM.
-Unfortunately, ML-KEM is vulnerable to plaintext checking attacks that can recover complete secret key if the keypair is reused \cite{DBLP:conf/eurocrypt/BaetuDHTV19}.
-In fact, as Chris Peikert pointed out, because of the search-decision equivalence of LWE problems, almost all LWE-based cryptosystems will be vulerable to plaintext-checking attacks \cite{DBLP:conf/pqcrypto/Peikert14}.
-This naturally raises the question of finding other post-quantum primitives to instantiate ``encrypt-then-MAC'' with, such as other lattice-based cryptosystems not based on LWE, code-based cryptosystems and isogeny-based cryptosystems.
+The ``encrypt-then-MAC'' transformation is a generic KEM construction whose IND-CCA2 security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.
+Compared to the Fujisaki-Okamoto transformation, the ``encrypt-then-MAC'' replaces the computationally expensive re-encryption with computing a MAC tag.
+At the cost of minimal increase in encapsulation cost and ciphertext size, the ``encrypt-then-MAC'' substantially improves the efficiency of the decapsulation routine.
+Where the input PKE's encryption is slower than decryption, the ``encrypt-then-MAC'' KEM achieves meaningful time savings in practical key exchange protocols.
+
+Unfortunately, ML-KEM is not OW-PCA secure.
+In fact, as Chris Peikert pointed out in \cite{DBLP:conf/pqcrypto/Peikert14}, most lattice-based cryptosystems are not OW-PCA secure due to the search-decision equivalence of lattice problems.
+While we instantiated ``encrypt-then-MAC'' with ML-KEM subroutines for performance comparison, the resulting KEM is not chosen-ciphertext secure, and should not be used in production systems.
+The natural questions is thus to find a suitable cryptosystem to instantiate ``encrypt-then-MAC'' with.
+
+\textbf{RSA and ElGamal.} Combining ``encrypt-then-MAC'' with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}.
+RSA is also known to be OW-PCA secure because it is a trapdoor permutation and thus a rigid PKE.
+However, because of RSA's rigidity, there exists even more efficient KEM transformation (such as RSA-KEM \cite{DBLP:journals/iacr/Shoup01}) that only adds a single hash to the base PKE routines.
+While applying ``encrypt-then-MAC'' to RSA will result in an IND-CCA2 secure KEM, such construction offers no meaningful advantage to RSA-KEM.
+
+\textbf{Code-based cryptography.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems may not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
+Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}.
+On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, though in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying ``encrypt-then-MAC'' may not yield meaningful performance gains.
+
+\textbf{Isogeny-based cryptography.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}.
+While SIKE and SIDH were found to be insecure \cite{DBLP:conf/eurocrypt/CastryckD23}, other isogeny-based cryptosystem such as CSIDH \cite{DBLP:conf/asiacrypt/Castryck0MPR18} remains unaffected by the aforementioned attack and might be suitable candidates.
 
@@
 
-\section{Appendix}\label{sec:appendix}
-% TODO: brief introduction to the various authenticators should be in the appendix
-
-\subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
-ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition.
+% \section{Appendix}\label{sec:appendix}
+% % TODO: brief introduction to the various authenticators should be in the appendix
+
+% \subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
+% ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition.
 While their IND-CPA subroutines (see figure \ref{fig:k-pke-routines}) are identical, ML-KEM deviated from Kyber by choosing a different variant of the Fujisaki-Okamoto transformation.
 
-\begin{figure}[H]
-    \centering
-
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{K-PKE.KeyGen()}}
-            \begin{algorithmic}[1]
-                \State $A \leftsample R_q^{k \times k}$
-                \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
-                \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
-                \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
-                \State $\pk \leftarrow (A, \mathbf{t})$
-                \State $\sk \leftarrow \mathbf{s}$
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{K-PKE.Enc(pk, m)}}
-            \begin{algorithmic}[1]
-                \Ensure $\pk = (A, \mathbf{t})$
-                \Ensure $m \in R_2$
-                \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
-                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
-                \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
-                \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
-                \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
-                \State \Return $(\mathbf{c}_1, c_2)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.3\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{K-PKE.Dec(sk, c)}}
-            \begin{algorithmic}[1]
-                \Ensure $c = (\mathbf{c}_1, c_2)$
-                \Ensure $\sk = \mathbf{s}$
-                \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
-                \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
-                \State \Return $\hat{m}$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
-\end{figure}
-
-CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext.
+% \begin{figure}[H]
+%     \centering
+
+%     \begin{minipage}[t]{0.33\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{\texttt{K-PKE.KeyGen()}}
+%             \begin{algorithmic}[1]
+%                 \State $A \leftsample R_q^{k \times k}$
+%                 \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
+%                 \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
+%                 \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
+%                 \State $\pk \leftarrow (A, \mathbf{t})$
+%                 \State $\sk \leftarrow \mathbf{s}$
+%                 \State \Return $(\pk, \sk)$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}\hfill
+%     \begin{minipage}[t]{0.33\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{\texttt{K-PKE.Enc(pk, m)}}
+%             \begin{algorithmic}[1]
+%                 \Ensure $\pk = (A, \mathbf{t})$
+%                 \Ensure $m \in R_2$
+%                 \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
+%                 \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
+%                 \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
+%                 \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
+%                 \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
+%                 \State \Return $(\mathbf{c}_1, c_2)$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}\hfill
+%     \begin{minipage}[t]{0.3\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{\texttt{K-PKE.Dec(sk, c)}}
+%             \begin{algorithmic}[1]
+%                 \Ensure $c = (\mathbf{c}_1, c_2)$
+%                 \Ensure $\sk = \mathbf{s}$
+%                 \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
+%                 \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
+%                 \State \Return $\hat{m}$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}
+%     \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
+% \end{figure}
+
+% CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext.
 On the other hand, because by using \emph{re-encryption} and \emph{de-randomization}, the PKE is already made \emph{rigid}, the CRYSTALS-Kyber team decided to use the $U^{\not\bot}_m$ variant, where the shared secret is derived from the plaintext alone.
 
-\begin{figure}[H]
-    \centering
-    \begin{minipage}[t]{0.45\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{KEM.KeyGen()}}
-            \begin{algorithmic}[1]
-                \State $z \leftsample \{0,1\}^{256}$
-                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
-                \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-        \begin{algorithm}[H]
-            \caption*{\texttt{KEM.Encap(pk)}}
-            \begin{algorithmic}[1]
-                \State $m \leftsample \{0,1\}^{256}$
-                \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
-                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
-                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
-                    \Comment{$U^{\not\bot}$}
-                \State $K \leftarrow \overline{K}$
-                    \Comment{$U^{\not\bot}_m$}
-                \State \Return $(c, K)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.5\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{KEM.Decap(sk, c)}}
-            \begin{algorithmic}[1]
-                \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
-                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
-                \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
-                \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
-                    \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
-                        \Comment{$U^{\not\bot}$}
-                    \State $K \leftarrow \overline{K}$
-                        \Comment{$U^{\not\bot}_m$}
-                \Else 
-                    \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
-                \EndIf
-                \State \Return $K$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-
-    \caption{
-        Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
-    }\label{fig:kyber-ml-kem-compare}
-\end{figure}
-
-The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold.
+% \begin{figure}[H]
+%     \centering
+%     \begin{minipage}[t]{0.45\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{\texttt{KEM.KeyGen()}}
+%             \begin{algorithmic}[1]
+%                 \State $z \leftsample \{0,1\}^{256}$
+%                 \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
+%                 \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
+%                 \State \Return $(\pk, \sk)$
+%             \end{algorithmic}
+%         \end{algorithm}
+%         \begin{algorithm}[H]
+%             \caption*{\texttt{KEM.Encap(pk)}}
+%             \begin{algorithmic}[1]
+%                 \State $m \leftsample \{0,1\}^{256}$
+%                 \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
+%                 \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
+%                 \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
+%                     \Comment{$U^{\not\bot}$}
+%                 \State $K \leftarrow \overline{K}$
+%                     \Comment{$U^{\not\bot}_m$}
+%                 \State \Return $(c, K)$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}
+%     \begin{minipage}[t]{0.5\textwidth}
+%         \begin{algorithm}[H]
+%             \caption*{\texttt{KEM.Decap(sk, c)}}
+%             \begin{algorithmic}[1]
+%                 \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
+%                 \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
+%                 \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
+%                 \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
+%                     \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
+%                         \Comment{$U^{\not\bot}$}
+%                     \State $K \leftarrow \overline{K}$
+%                         \Comment{$U^{\not\bot}_m$}
+%                 \Else 
+%                     \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
+%                 \EndIf
+%                 \State \Return $K$
+%             \end{algorithmic}
+%         \end{algorithm}
+%     \end{minipage}
+
+%     \caption{
+%         Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
+%     }\label{fig:kyber-ml-kem-compare}
+% \end{figure}
+
+% The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold.
 The first reason is performance: using the $U^{\not\bot}_m$ transformation saves the need to hash the ciphertext, and since Kyber/ML-KEM's performance is mainly bottlenecked by the symmetric components, omitting the hash leads to significant runtime savings (up to 17\% in AVX-2 optimized implementations).
@@
 
-In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.
+% In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.
 

### 2024-10-09 etm-transform/manuscript/main.tex
ElGamal is a special case of encrypt-then-MAC

@@
     Combining all equations above produce the desired security bound.
+\end{proof}
+
+\subsection{ElGamal is OW-PCA secure}
+We show that the DHAES/DHIES hybrid encryption scheme is a special case of the ``encrypt-then-MAC'' transformation. Specifically, we will sketch a proof of the following lemma:
+
+\begin{lemma}\label{lemma:elgamal-is-ow-pca}
+    For every OW-PCA adversary $A$ against the ElGamal cryptosystem, there exists a Gap Diffie-Hellman problem solver $B$ such that:
+
+    \begin{equation*}
+        \text{Adv}_\text{GapDH}(B) = \text{Adv}_\text{OW-PCA}(A)
+    \end{equation*}
+
+    In other words, ElGamal is OW-PCA secure under the Gap Diffie-Hellman assumption.
+\end{lemma}
+
+Each ElGamal cryptosystem \cite{DBLP:journals/tit/Elgamal85} is parameterized by a cyclic group $G = \langle g \rangle$ of prime order $q > 2$. A summary of the routine is shown below:
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.32\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{KeyGen()}}
+            \begin{algorithmic}[1]
+                \State $x \leftsample \mathbb{Z}_q$
+                \State $\sk \leftarrow x$
+                \State $\pk \leftarrow g^x$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{Enc}(\pk = g^x, m \in G)$}
+            \begin{algorithmic}[1]
+                \Require $m \in G$
+                \State $y \leftsample \mathbb{Z}_q$
+                \State $w \leftarrow g^y$
+                \State $v \leftarrow m \cdot (g^x)^y$
+                \State \Return $c = (w, v)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{Dec}(\sk = x, c = (w,v) \in G^2)$}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow (w^x)^{-1}\cdot v$
+                \State \Return $\hat{m}$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{ElGamal cryptosystem}
+\end{figure}
+
+The security of ElGamal cryptosystem reduces to the conjectured intractability of the computational Diffie-Hellman problem and the decisional Diffie-Hellman problem:
+
+\begin{definition}[\textbf{computational Diffie-Hellman problem (CDH)}]
+    Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $(g, g^x, g^y)$, compute $g^{xy}$.
+\end{definition}
+\begin{definition}[\textbf{decisional Diffie-Hellman problem (DDH)}]
+    Let $x, y, z \leftsample \mathbb{Z}_q$ be uniformly random samples.
+Let $h \leftsample \{g^z, g^{xy}\}$ be randomly chosen between $g^z$ and $g^{xy}$.
+Given $(g, g^x, g^y, h)$, determine whether $h$ is $g^{xy}$ or $g^z$
+\end{definition}
+
+It is also conjectured in \cite{DBLP:conf/ctrsa/AbdallaBR01} that for certain choice of cyclic group $G$, the computational Diffie-Hellman problem remains intractable even if the adversary as access to a restricted decisional Diffie-Hellman oracle.
+This assumption is captured in the Gap Diffie-Hellman problem:
+
+\begin{definition}[\textbf{Gap Diffie-Hellman problem}]
+    Let $G = \langle g \rangle$ be a cyclic group of prime order $q > 2$.
+Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples.
+Given $(g, g^x, g^y)$ and a restricted DDH oracle $\mathcal{O}^\text{DDH}: (u, v) \mapsto \llbrack u^x = v \rrbrack$, compute $g^{xy}$.
+\end{definition}
+
+We now present the proof for lemma \ref{lemma:elgamal-is-ow-pca}
+\begin{proof}
+    We will prove by sequence of game. A summary can be found in figure \ref{fig:elgamal-pca-games}
+
+    \begin{figure}
+        \centering
+
+        \begin{minipage}[t]{0.4\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$G_0 - G_2$}
+                \begin{algorithmic}[1]
+                    \State $x \leftsample \mathbb{Z}_q$
+                    \State $m^\ast \leftsample G$
+                    \State $y \leftsample \mathbb{Z}_q, w \leftarrow g^y$
+                    \State $v \leftarrow m^\ast \cdot (g^x)^y$
+                        \Comment{$G_0$ - $G_1$}
+                    \State $v \leftsample G$
+                        \Comment{$G_2$}
+                    \State $c^\ast \leftarrow (w, v)$
+                    \State $\hat{m} \leftsample A^{\pco}(g^x, c^\ast)$
+                        \Comment{$G_0$}
+                    \State $\hat{m} \leftsample A^{\pco_1}(g^x, c^\ast)$
+                        \Comment{$G_1$ - $G_2$}
+                    \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+                        \Comment{$G_0$ - $G_1$}
+                    \State \Return $\llbrack \hat{m} = w^{-x}\cdot v \rrbrack$
+                        \Comment{$G_2$}
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \begin{minipage}[t]{0.4\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$\pco(m, c=(w, v))$}
+                \begin{algorithmic}[1]
+                    \State \Return $\llbrack m = (w^x)^{-1}\cdot v\rrbrack$
+                \end{algorithmic}
+            \end{algorithm}
+            \begin{algorithm}[H]
+                \caption*{$\pco_1(m, c=(w, v))$}
+                \begin{algorithmic}[1]
+                    \State \Return $\llbrack (w^x) = m^{-1} \cdot v \rrbrack$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+
+        \caption{Lemma \ref{lemma:elgamal-is-ow-pca} sequence of games}\label{fig:elgamal-pca-games}
+    \end{figure}
+
+    \emph{Game 0} is the OW-PCA game. Adversary $A$ has access to the plaintext-checking oracle $\pco$ and wins the game if it can correctly recover the challenge plaintext $m^\ast$.
+
+    \emph{Game 1} is identical to game 0, except that the formulation of the $\pco$ is changed.
+When servicing the plaintext-checking query $(m, c = (w, v))$, $\pco_1$ checks whether $w^x$ is equal to $m^{-1} \cdot v$.
+Observe that in the cyclic group $G$, the algebraic expressions in $\pco$ and $\pco_1$ are equivalent, which means that $\pco_1$ behaves identically to $\pco$.
+
+    \emph{Game 2} is identical to game 1 except for two modifications: first, when computing the challenge ciphertext, $v$ is no longer computed from $m^\ast$ but is randomly sampled; second, the win condition changed from $\hat{m} = m^\ast$ to $\hat{m} = w^{-x}\cdot v$.
+Notice that in game 0-1, $v \leftsample m^\ast \cdot (g^x)^y$ follows uniform random distribution on the cyclic group $G$ because $m^\ast$ is uniformly random in the cyclic group, so adversary $A$ retains its advantage in ``recovering the decryption'' when $v$ becomes truly uniformly random in the cyclic group.
+It is easy to verify that the two win conditions are equivalent.
+Up to this point, we have simply moved things around, and game 0 through game 2 are algebraically equivalent:
+
+    \begin{equation*}
+        \adv_0(A) = \adv_1(A) = \adv_2(A)
+    \end{equation*}
+
+    The Gap Diffie-Hellman adversary $B$ can perfectly simulate game 2 for $A$ (see figure \ref{fig:ow-pca-to-gap-dh}): $B$ receives as the Gap Diffie-Hellman problem inputs $g^x$ and $g^y$.
+$g^x$ simulates an ElGamal public key, where as $g^y$ simulates the first component of the challenge ciphertext.
+As in game 2, the second component of the challenge ciphertext can be randomly sampled.
+Finally, the $\pco_1$ from game 2 can be perfectly simulated using the restricted DDH oracle $\mathcal{O}^\text{DDH}$.
+
+    \begin{figure}
+        \centering
+
+        \begin{minipage}[t]{0.45\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$B^{\mathcal{O}^\text{DDH}}(g, g^x, g^y)$}
+                \begin{algorithmic}[1]
+                    \State $w \leftarrow g^y$
+                    \State $v \leftsample G$
+                    \State $c^\ast \leftarrow (w, v)$
+                    \State $\hat{m} \leftsample A^{\pco_2}(g^x, c^\ast)$
+                    \State \Return $\hat{m}^{-1}\cdot v$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+        \begin{minipage}[t]{0.45\textwidth}
+            \begin{algorithm}[H]
+                \caption*{$\mathcal{O}^\text{DDH}(u, v)$}
+                \begin{algorithmic}[1]
+                    \State \Return $\llbrack u^x = v \rrbrack$
+                \end{algorithmic}
+            \end{algorithm}\vspace{-0.3cm}
+            \begin{algorithm}[H]
+                \caption*{$\pco_2(m, c=(w, v))$}
+                \begin{algorithmic}[1]
+                    \State \Return $\mathcal{O}^\text{DDH}(w, m^{-1}\cdot v)$
+                \end{algorithmic}
+            \end{algorithm}
+        \end{minipage}
+
+        \caption{Gap Diffie-Hellman adversary $B$ simulates game 2 for $A$}\label{fig:ow-pca-to-gap-dh}
+    \end{figure}
+
+    If $A$ wins game 2, then its output is $\hat{m} = w^{-x}\cdot v = g^{-xy}\cdot v$, so $m^{-1}\cdot v$ is $g^{xy}$, the correct answer to the Gap Diffie-Hellman problem.
+In other words, $A$ wins game 2 if and only if $B$ solves the Gap Diffie-Hellman problem:
+
+    \begin{equation*}
+        \adv_2(A) = \adv_\text{GapDH}(B)
+    \end{equation*}
 \end{proof}
@@
 In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key.
-This helps the key exchange to achieve weak forward secrecy (citation needed).
+This helps the key exchange to achieve weak forward secrecy \cite{DBLP:conf/eurocrypt/CanettiK01}.
 
 Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
-KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage.
-(citation needed).
+KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage \cite{DBLP:conf/ccs/SchwabeSW20}.
 

### 2024-10-11 etm-transform/manuscript/main.tex
checkpoint

@@
 For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table.
-We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for the details.
-
-\textbf{Security.} We discuss the security of a $\pke$ using the sequence of games described in \cite{shoup2004sequences}.
-Specifically, we first define the $\texttt{OW-ATK}$ as they pertain to a public key encryption scheme.
-In later section we will define the $\texttt{IND-CCA}$ game as it pertains to a key encapsulation mechanism.
-
-\def\algowatkgame{
-    \begin{algorithm}[H]
-        \caption*{The $\texttt{OW-ATK}$ game}
+We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for details of the mitigation techniques, though decryption failure is not a focus of this paper.
+
+\textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}.
+Each game is defined by the goal and the capabilities of the adversary.
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{OW-ATK} Game}
             \begin{algorithmic}[1]
@@
                 \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
-            \State $\hat{m} \leftsample \mathcal{A}^{\mathcal{O}_\texttt{ATK}}(
-                1^\lambda, \pk, c^\ast
-            )$
-            \State \Return $\llbrack m^\ast = \hat{m} \rrbrack$
-        \end{algorithmic}
-    \end{algorithm}
-}
-\def\algpco{
-    \begin{algorithm}[H]
-        \caption*{$\pco(m \in \mathcal{M}, c \in \mathcal{C})$}
-        \begin{algorithmic}[1]
-            \State \Return $\llbrack \decrypt(\sk, c) = m \rrbrack$
-        \end{algorithmic}
-    \end{algorithm}
-}
-\begin{figure}[H]
-    \centering
-    \begin{minipage}[b]{0.49\textwidth}
-        \algowatkgame
+                \State $\hat{m} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
+                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hspace{0.05\textwidth}
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{IND-ATK} Game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+                \State $(m_0, m_1) \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk)$
+                \State $b \leftsample \{0,1\}$
+                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
+                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
+                \State \Return $\llbrack \hat{b} = b \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
     \end{minipage}
-    \begin{minipage}[b]{0.49\textwidth}
-        \algpco
-    \end{minipage}
-    \caption{
-        One-way security game of PKE (left) and plaintext-checking oracle (right)
-    }\label{fig:ow-game-pco}
+    \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
 \end{figure}
 
-In the \texttt{OW-ATK} game (see figure \ref{fig:ow-game-pco}), an adversary's goal is to recover the decryption of a randomly generated ciphertext.
-A challenger randomly samples a keypair and a challenge plaintext $m^\ast$, encrypts the challenge plaintext $c^\ast \leftsample \encrypt(\pk, m^\ast)$, then gives $\pk$ and $c^\ast$ to the adversary $A$.
-The adversary $A$, with access to some oracle $\mathcal{O}_\texttt{ATK}$, outputs a guess decryption $\hat{m}$.
-$A$ wins the game if its guess $\hat{m}$ is equal to the challenge plaintext $m^\ast$.
-The \textit{advantage} $\adv_\texttt{OW-ATK}$ of an adversary in this game is the probability that it wins the game:
-
-\begin{equation*}
-    \adv_\texttt{OW-ATK}(A) = P\left[
-        A(\pk, c^\ast) = m^\ast \vert
-        (\pk, \sk) \leftsample \keygen() ;
-        m^\ast \leftsample \mathcal{M};
-        c^\ast \leftsample \encrypt(\pk, m^\ast)
-    \right]
-\end{equation*}
-
-The capabilities of the oracle $\mathcal{O}_\texttt{ATK}$ depends on the choice of security goal $\texttt{ATK}$.
-Particularly relevant to our result is security against plaintext-checking attack (PCA), for which the adversary has access to a plaintext-checking oracle (PCO) (see figure \ref{fig:ow-game-pco}).
-A PCO takes as input a plaintext-ciphertext pair $(m, c)$ and returns \texttt{True} if $m$ is the decryption of $c$ or \texttt{False} otherwise.
+Depending on the attack model, the adversary may have access to various oracles.
+In public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack.
+If the adversary has access to a plaintext-checking oracle $\pco$ then it can mount plaintext-checking attack (PCA).
+Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA).
+A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary with access to the corresponding oracle(s) can win the corresponding game with non-negligible advantage.
 
 \subsection{Key encapsulation mechanism (KEM)}
-A key encapsulation mechanism is a collection of three routines $(\keygen, \encap, \decap)$ defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
+\textbf{Syntax} A key encapsulation mechanism $\kem(\keygen, \encap, \decap)$ is a collection of three routines defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
 The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$.
@@
 
-The IND-CCA security of a KEM is defined by an adversarial game in which an adversary's goal is to distinguish pseudorandom shared secret (generated by running the $\encap$ routine) and a truly random value.
-
-\begin{figure}[H]
-    \centering
-    \begin{minipage}[b]{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption*{KEM-IND-CCA2 game}
+\textbf{Security} The security of KEMs is similarly discussed in adversarial games (figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game.
+In a KEM's indistinguishability game, an adversary is given the public key and a challenge ciphertext, then asked to distinguish a pseudorandom shared secret $K_0$ associated with the challenge ciphertext from a truly random bit string of equal length.
+
+\begin{figure}[h]
+    \centering
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{IND-ATK} game}
             \begin{algorithmic}[1]
@@
                 \State $b \leftsample \{0, 1\}$
-                \State $\hat{b} \leftsample A^{\mathcal{O}_\decap}(
+                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(
                     1^\lambda, \pk, c^\ast, K_b
@@
     \end{minipage}\hfill
-    \begin{minipage}[b]{0.49\textwidth}
+    \begin{minipage}[t]{0.49\textwidth}
         \begin{algorithm}[H]
@@
     \end{minipage}
-    \caption{IND-CCA2 game for KEM (left) and decapsulation oracle (right)}\label{fig:kem-game}
+    \caption{\texttt{IND-ATK} game for KEM and decapsulation oracle $\mathcal{O}_\decap$}\label{fig:kem-game}
 \end{figure}

### 2024-10-11 etm-transform/manuscript/main.tex
Adding revisions

@@
 %%%% 4. TITLE %%%%
-\title{Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''}
+\title{Faster generic IND-CCA secure KEM using encrypt-then-MAC}
 %%%% NOTES:
@@
     Key encapsulation mechanism, 
+    message authentication code,
     post-quantum cryptography,
@@
 \begin{abstract}
-  The modular Fujisaki-Okamoto (FO) transformation takes public-key encryption with weaker security and constructs a key encapsulation mechanism (KEM) with indistinguishability under adaptive chosen ciphertext attacks.
-While the modular FO transform enjoys tight security bound and quantum resistance, it also suffers from computational inefficiency and vulnerabilities to side-channel attacks due to using de-randomization and re-encryption for providing ciphertext integrity.
-In this work, we propose an alternative KEM construction that achieves ciphertext integrity using a message authentication code (MAC) and instantiate a concrete instance using Kyber.
-Our experimental results showed that where the encryption routine incurs heavy computational cost, replacing re-encryption with MAC provides substantial performance improvements at comparable security level.
+    Indistinguishability under chosen-ciphertext attack (IND-CCA) is the accepted standard of security for key encapsulation mechanisms (KEM).
+However, provably chosen-ciphertext security is difficult to achieve from scratch.
+In this work, we generalize the ``Hashed ElGamal'' encryption scheme proposed by M.
+Abdalla et al into an IND-CCA secure KEM by combining an OW-PCA secure public-key encryption (PKE) scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC.
+Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
+When instantiated with the PKE sub-routines of ML-KEM, the encrypt-then-MAC transformation achieves massive computational efficiency improvements over ML-KEM in a variety of practical scenarios.
 \end{abstract}
-
-%%%% 7. PAPER CONTENT %%%%
-% \section{Introduction}\label{sec:introduction}
-% Key encapsulation mechanism (KEM) is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
-The combination of KEM and some data encapsulation mechanism (DEM), such as AES-GCM and ChaCha20-Poly1305, is the foundation of many of today's most widely adopted communication protocols such as Transport Layer Security (TLS) and Secure Shell (SSH).
-The commonly accepted security standard of a KEM is \emph{Indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2)}: no efficient adversary can distinguish a shared secret obtained by running the encapsulation routine from random noise, even with access to a decapsulation oracle throughout the attack.
-KEM is related to another important cryptographic primitive called public-key encryption (PKE), and their security standards are similar.
-The desired security standard for PKE, also called IND-CCA2, requires that no efficient adversary can distinguish the encryption of two adversarially chosen messages even with access to decryption oracle.
-
-% It is difficult to build a provably IND-CCA2 secure KEM from scratch.
-Instead, secure KEMs are usually built on top of a PKE with weaker security property (e.g.
-being only OW-CPA or IND-CPA secure, not IND-CCA2).
-One such construction is the Fujisaki-Okamoto transformation, proposed in 1999 by Fujisaki Eiichiro and Okamoto Tetsuyaki in their seminal paper \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}.
-The first Fujisaki-Okamoto transformation combines an OW-CPA secure PKE with an IND-CPA secure symmetric cipher into a hybrid PKE (HPKE) with proven IND-CCA2 security under the random oracle model.
-Subsequent works \cite{DBLP:conf/ctrsa/OkamotoP01}\cite{DBLP:conf/ctrsa/CoronHJPPT02}\cite{DBLP:conf/ima/Dent03} improved on the original proposal and adapted it to build KEM instead of HPKE.
-This line of work culminated in a landmark publication in 2017 by Hofhein, Hovelmann, and Kiltz \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}, where the authors provided a versatile variety of modular KEM constructions with tight security reduction in the random oracle model and non-tight security reduction in the quantum random oracle model.
-
-% The modular Fujisaki-Okamoto KEM transformation is remarkably successful.
-It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
-When Kyber was standardized by NIST in FIPS 203 ``Module-lattice key-encapsulation mechanism'' (ML-KEM) \cite{FIPS203}, it kept the Fujisaki-Okamoto transformation in its KEM construction.
-However, the Fujisaki-Okamoto transformation is not perfect.
-It uses \emph{de-randomization} and \emph{re-encryption} to achieve \emph{rigidity} \cite{bernstein2018towards} which then ensures ciphertext integrity.
-This brings two problems: \begin{itemize}
-%     \item \textbf{computational inefficiency:} where the PKE's encryption routine is substantially more expensive than the decryption routine, using re-encryption causes the decapsulation routine in the output KEM to become computationally expensive
-%     \item \textbf{side-channel vulnerability:} running the input PKE's encryption routine in the output KEM's decapsulation routine introduces risk of side-channel vulnerabilities not found in the input PKE's decryption routine alone.
-In fact, many practical attacks \cite{ueno2022curse}\cite{ravi2019generic} exploit re-encryption to decrypt ciphertext or recover secret keys.
-Countermeasures such as masking have been proposed to address these side channels, but they inevitably carry substantial performance penalty.
-% \end{itemize}
-
-% \subsection{Our contributions}
-% Our main contribution is a novel generic KEM construction that combines a PKE with weak security and a message authentication code (MAC).
-We were inspired by the strategy used in symmetric cyrptography for achieving IND-CCA2 security: combining a symmetric cipher with a MAC in a pattern called ``encrypt-then-MAC''.
-In the encapsulation routine of the ``encrypt-then-MAC'' KEM construction, a random PKE plaintext $m$ is sampled and encrypted into some PKE ciphertext $c^\prime$; a symmetric key $k$, derived from hashing the plaintext, is used to compute a message authenticator $t$ on the PKE ciphertext $c^\prime$.
-The KEM ciphertext $c = (c^\prime, t)$ contains both the PKE ciphertext and the message authenticator.
-At decapsulation, the PKE ciphertext $c^\prime$ is decrypted into $\hat{m}$ first, then a symmetric key $\hat{k}$ is derived from hashing the decryption $\hat{m}$.
-The symmetric key is used to verify the authenticator against the PKE ciphertext.
-A illustration of the data flow in the ``encrypt-then-MAC'' KEM construction can be found in figure \ref{fig:etm-diagram}.
-A detailed description of the individual routines can be found in figure \ref{fig:etm-routines}.
-
-% \begin{figure}
-%     \centering
-%     \begin{tikzpicture}
-%         \tikzstyle{rect} = [draw, rectangle, minimum width=3cm, minimum height=1cm]
-%         \tikzstyle{filledrect} = [draw, rectangle, minimum width=3cm, minimum height=1cm, fill=lightgray]
-%         \matrix [column sep=7mm, row sep=5mm] {
-%             \node (pk) [rect] {$\pk$}; &
-%             \node (m) [rect] {$m \leftsample \mathcal{M}$}; &
-%             \node (hashg) [rect] {Hash function $G$}; \\
-%             \node (pkeenc) [rect] {$\texttt{PKE.Enc}$}; &
-%             \node (mackey) [rect] {$k_\mac$}; \\
-%             \node (ct) [filledrect] {$c^\prime$}; &
-%             \node (mac) [rect] {$\mac$}; &
-%             \node (tag) [filledrect] {$t$}; \\
-%         };
-%         \draw[->, thick] (pk) -- (pkeenc);
-%         \draw[->, thick] (m) -- (pkeenc);
-%         \draw[->, thick] (m) -- (hashg);
-%         \draw[->, thick] (hashg) -- (mackey);
-%         \draw[->, thick] (pkeenc) -- (ct);
-%         \draw[->, thick] (mackey) -- (mac);
-%         \draw[->, thick] (ct) -- (mac);
-%         \draw[->, thick] (mac) -- (tag);
-%     \end{tikzpicture}
-
-%     \caption{Combining PKE with MAC using ``encrypt-then-MAC'' to encapsulate a shared secret. 
-%     The returned values are colored grey
-%     }\label{fig:etm-diagram}
-% \end{figure}
-
-% \subsubsection{Security reduction}
-% By authenticating the PKE ciphertext using a symmetric key derived from its decryption, the ``encrypt-then-MAC'' construction greatly limits the amount of information an adversary can learn from a chosen-ciphertext attack.
-For an adversary to produce a valid KEM ciphertext, it must produce a valid message authenticator for the PKE ciphertext under the correct symmetric key.
-Under the random oracle model, the symmetric key is indistinguishable from uniformly random unless the adversary also knows the corresponding plaintext completely, thus rendering the output of the decasulation oracle redundant.
-Furthermore, because each encapsulation will pick a random message, within a reasonable lifetime of the KEM keypair, each KEM ciphertext will be authenticated using a distinct symmetric key, so the choice of MAC is only required to be one-time secure, affording us additional performance gains by choosing more efficient MAC instances.
-In section \ref{sec:main-results}, we prove that the IND-CCA2 security of the ``encrypt-then-MAC'' KEM reduces tightly to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC under the random oracle model.
-
-% One-wayness under plaintext-checking attack (OW-PCA)\cite{DBLP:conf/ctrsa/OkamotoP01} requires that no efficient adversary can recover some random encryption even with access to a plaintext-checking oracle, who accepts a plaintext-ciphertext pair $(m, c)$ and returns whether the queried ciphertext decrypts to the queried plaintext.
-While it is not a standard security notion, it is useful when reducing the security of the ``encrypt-then-MAC'' KEM to the security of the PKE, and there are many well studied cryptosystems that are known to be OW-PCA: the RSA cryptosystem \cite{DBLP:journals/cacm/RivestSA78} is known to be a trapdoor permutation and thus trivially OW-PCA secure; the ElGamal cryptosystem  \cite{DBLP:journals/tit/Elgamal85} is conjectured to be OW-PCA secure under the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01}.
-
-% \subsubsection{Performance improvements}
-% The main advantage of our KEM construction over the Fujisaki-Okamoto transformation is the performance gains: our construction replaces re-encryption with computing a message authenticator, which is significantly faster.
-When instantiated with the underlying PKE routines of ML-KEM and the Poly1305 message authenticator, our construction $\text{ML-KEM}^+$ achieves on average 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles for encapsulation when compared to ML-KEM.
-
-% \begin{table}[H]
-%     \centering
-%     \footnotesize
-
-%     \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-%         \hline
-%         & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
-%         & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
-%         & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
-%         \\
-%         \hline
-%         Encap \newline (ccl/tick) 
-%         & 91467 & 93157 \newline (+1.8\%) 
-%         & 136405 & 146405 \newline (+7.3\%) 
-%         & 199185 & 205763 \newline (+3.3\%) 
-%         \\
-%         \hline
-%         Decap \newline (ccl/tick) 
-%         & 121185 & 33733 \newline (-72.2\%) 
-%         & 186445 & 43315 \newline (-76.8\%) 
-%         & 246245 & 51375 \newline (-79.1\%) 
-%         \\
-%         \hline
-%         CT size \newline (bytes) 
-%         & 768 & 784 \newline (+2.1\%) 
-%         & 1088 & 1104 \newline (+1.5\%) 
-%         & 1568 & 1584 \newline (+1.0\%) 
-%         \\
-%         \hline
-%     \end{tabular}
-
-%     \caption{$\text{ML-KEM}^+$ is instantiated with Poly1305}
-% \end{table}
-
-% We also implemented and measured the round trip time of key exchange protocols with various modes of authentication.
-When compared to ML-KEM, $\text{ML-KEM}^+$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% reduction in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
-
-% \begin{table}[H]
-%     \centering
-%     \footnotesize
-
-%     \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-%         \hline
-%         & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
-%         & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
-%         & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
-%         \\
-%         \hline
-%         KE RTT \newline ($\us$) 
-%         & 92 & 70 \newline (-23.9\%) 
-%         & 135 & 99 \newline (-26.7\%) 
-%         & 193 & 138 \newline (-28.5\%) 
-%         \\
-%         \hline
-%         UAKE RTT \newline ($\us$) 
-%         & 145 & 103 \newline (-29.0\%) 
-%         & 215 & 144 \newline (-33.0\%) 
-%         & 310 & 202 \newline (-34.8\%) 
-%         \\
-%         \hline
-%         AKE RTT \newline ($\us$) 
-%         & 220 & 133 \newline (-39.5\%) 
-%         & 294 & 190 \newline (-35.4\%) 
-%         & 512 & 266 \newline (-48.0\%) 
-%         \\
-%         \hline
-%     \end{tabular}
-
-%     \caption{Key exchange round-trip times}
-% \end{table}
-
-% \subsection{Related works}
-% As mentioned at the beginning of section \ref{sec:introduction}, there is a large volume of previous proposals that combine cryptographic primitives with weak security properties into cryptographic primitives with strong security properties.
-Besides the Fujisaki-Okamoto transformation, another transformation that sees substantial adoption is \emph{Optimal Asymmetric Encryption Padding (OAEP)} \cite{bellare1995optimal}.
-Unfortunately, OAEP can only be applied to one-way trapdoor permutation, a rare occurrence.
-To this day, RSA remains the only viable candidate \cite{DBLP:conf/crypto/FujisakiOPS01}, though RSA-OAEP has been widely adopted after its standardization in PKCS\#1 v2.2 \cite{rfc8017}.
-
-% Securing an encryption scheme against chosen-ciphertext attack was a well-studied problem in the context of symmetric cryptography.
-Among the three generic patterns ``encrypt-then-MAC'', ``encrypt-and-MAC'', ``MAC-then-encrypt'', only the first one is formally analyzed and proved to be secure with any combination of symmetric cipher and MAC \cite{DBLP:conf/crypto/Krawczyk01}\cite{DBLP:conf/eurocrypt/CanettiK01}.
-``MAC-then-encrypt'' is only secure with specific choices of symmetric cipher and MAC, and ``encrypt-and-MAC'' is generally not semantically secure since most MAC will leak information about the message it authenticates.
-Today's most well-established symmetric encryption schemes, including AES-GCM and ChaCha20-Poly1305, largely follow the pattern of ``encrypt-then-MAC''.
-
-% The idea of using a message authenticator to ensure ciphertext integrity was instantiated with ElGamal cryptosystem by Bellare and Rogaway in 1997 \cite{DBLP:conf/icics/BellareR97}, and its security was later reduced to the Gap Diffie-Hellman assumption by Abdulla, Bellare, and Rogaway in \cite{DBLP:conf/ctrsa/AbdallaBR01} in 2001.
-Our work can be seen as generalizing Bellare and Rogaway's scheme from ElGamal cryptosystem to all PKE with OW-PCA security.
 
@@
 A key encapsulation mechanism (KEM) \cite{DBLP:journals/iacr/Shoup01} is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel.
-The desired security standard for a KEM is called indistinguishability under adaptive chosen-ciphertext attack (IND-CCA2).
-Intuitively speaking, IND-CCA2 security requires that no efficient adversary can distinguish a pseudorandom shared secret from truly random noise, even with access to a decapsulation oracle throughout the attack.
-However, building a provably IND-CCA2 KEM is tremendously difficult.
-Early attempts without formal proof, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attack \cite{DBLP:conf/crypto/Bleichenbacher98}.
-In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to ensure ciphertext non-malleability \cite{DBLP:conf/asiacrypt/BellareN00}.
-Some of the earliest proposals for generic IND-CCA2 secure constructions include OAEP \cite{DBLP:conf/eurocrypt/BellareR94}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.
-
-On the other hand, chosen-ciphertext security is a solve problem in symmetric cryptography.
-It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called ``encrypt-then-MAC'' \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA2 security.
+The desired security standard for a KEM is called indistinguishability under chosen-ciphertext attack (IND-CCA).
+Intuitively speaking, IND-CCA security requires that no efficient adversary can distinguish a pseudorandom shared secret from a uniformly random bit string of equal length, even with access to a decapsulation oracle throughout the attack.
+However, building a provably IND-CCA secure KEM is tremendously difficult.
+Early attempts without formal proofs, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attacks \cite{DBLP:conf/crypto/Bleichenbacher98}.
+In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to achieve \emph{ciphertext non-malleability} \cite{DBLP:conf/asiacrypt/BellareN00}.
+Some of the earliest proposals for generic IND-CCA secure constructions include OAEP \cite{DBLP:conf/eurocrypt/BellareR94}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.
+
+On the other hand, chosen-ciphertext security is a solved problem in symmetric cryptography.
+It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA security.
 While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
-Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA2 security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
-The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypt the PKE plaintext, then authenticate the PKE ciphertext using the previously derived MAC key.
-Where the Gap Diffie-Hellman assumption holds and the MAC is existentially unforgeable, no efficient adversary can recover the decryption of an unknown ciphertext even with access to a decryption oracle because it cannot produce a valid tag for such unknown ciphertext.
+Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
+The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypting the PKE plaintext, then authenticating the PKE ciphertext using the previously derived MAC key.
+When the Gap Diffie-Hellman assumption holds and the MAC is existentially unforgeable, no efficient adversary can recover the decryption of an unknown ciphertext even with access to a decryption oracle because it cannot produce a valid tag for such unknown ciphertext.
 
 \subsection{Our contribution}
-In this paper, we propose a generic IND-CCA2 secure KEM using techniques adapted from DHIES.
-We call our proposed scheme the ``encrypt-then-MAC'' transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
-A summary of the data flow in the ``encrypt-then-MAC'' transformation can be found in figure \ref{fig:etm-diagram}.
+In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES.
+We call our proposed scheme the encrypt-then-MAC transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
+Our scheme mainly differs from DHIES in its versatility and input requirement.
+Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security).
+In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (Abdalla et al originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
 
@@
 
-    \caption{Combining PKE with MAC using ``encrypt-then-MAC'' to encapsulate a shared secret. 
+    \caption{Combining PKE with MAC using encrypt-then-MAC to encapsulate a shared secret. 
     The returned values are colored grey
@@
 
-Our scheme mainly differs from DHIES in its versatility.
-Whereas the IND-CCA2 security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the ``encrypt-then-MAC'' KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security).
-In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the ``encrypt-then-MAC'' KEM can be instantiated with one-time secure MAC such as Poly1305 (Abdalla et al originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
-
-\subsubsection{Performance improvements}
-In section \ref{sec:implementation}, we instantiate an instance of ``encrypt-then-MAC'' KEM using the PKE subroutines defined in ML-KEM \cite{FIPS203}.
+\textbf{Construction.} From a high level, the encrypt-then-MAC transformation takes as input a PKE and a MAC, then outputs a KEM.
+The output KEM encapsulates by encrypting a randomly sampled PKE plaintext and hashing the PKE plaintext-ciphertext pair into the shared secret.
+A symmetric key is derived from hashing the PKE plaintext, then used to produce a MAC tag on the PKE ciphertext.
+At decapsulation, the PKE secret key is used to recover the PKE plaintext, which is then hashed into a symmetric key that is used to verify the MAC tag against the input ciphertext.
+A summary of the data flow in the encrypt-then-MAC transformation can be found in Figure \ref{fig:etm-diagram}.
+A detailed description of individual routines can be found in Figure \ref{fig:etm-routines}.
+
+\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against chosen-plaintext attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure.
+The IND-CCA security of the output KEM reduces tightly to the OW-PCA security of the input PKE, and non-tightly to the unforgeability of the MAC, though the non-tightness will likely not be a weakness because the construction makes it easy to plug in MAC of any desired security level.
+
+\textbf{Implementation and performance comparison.} In Section \ref{sec:implementation}, we instantiate an instance of encrypt-then-MAC KEM using the PKE subroutines defined in ML-KEM \cite{FIPS203}.
 For clarity's sake we call this instantiation $\mlkemplus$.
@@
 On an AMD EPYC 9R14 CPU, $\mlkemplus$ (combined with Poly1305) achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation.
-See table \ref{tbl:cpu-cycles-summary} for a summary of CPU cycles needed for encapsulation and decapsulation in each security level.
+See Table \ref{tbl:cpu-cycles-summary} for a summary of performance comparison of individual routines against ML-KEM.
 
@@
 
-    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
-        & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
-        & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
+    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        & $\text{ML-KEM}^+$ \newline 512 
+        & $\text{ML-KEM}^+$ \newline 768 
+        & $\text{ML-KEM}^+$ \newline 1024 
         \\
@@
         Encap \newline (ccl/tick) 
-        & 91467 & 93157 \newline (+1.8\%) 
-        & 136405 & 146405 \newline (+7.3\%) 
-        & 199185 & 205763 \newline (+3.3\%) 
+        & 93157 \newline (+1.8\%) 
+        & 146405 \newline (+7.3\%) 
+        & 205763 \newline (+3.3\%) 
         \\
@@
         Decap \newline (ccl/tick) 
-        & 121185 & 33733 \newline (-72.2\%) 
-        & 186445 & 43315 \newline (-76.8\%) 
-        & 246245 & 51375 \newline (-79.1\%) 
+        & 33733 \newline (-72.2\%) 
+        & 43315 \newline (-76.8\%) 
+        & 51375 \newline (-79.1\%) 
         \\
@@
         CT size \newline (bytes) 
-        & 768 & 784 \newline (+2.1\%) 
-        & 1088 & 1104 \newline (+1.5\%) 
-        & 1568 & 1584 \newline (+1.0\%) 
+        & 784 \newline (+2.1\%) 
+        & 1104 \newline (+1.5\%) 
+        & 1584 \newline (+1.0\%) 
         \\
@@
 
-    \caption{$\text{ML-KEM}^+$ is instantiated with Poly1305}\label{tbl:cpu-cycles-summary}
+    \caption{$\text{ML-KEM}^+$ achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
 \end{table}
@@
 When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
-See table \ref{tbl:rtt-summary} for a summary of round trip times.
+See Table \ref{tbl:rtt-summary} for a summary of round trip times.
 
@@
 
-    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        & ML-KEM \newline 512 & $\text{ML-KEM}^+$ \newline 512 
-        & ML-KEM \newline 768 & $\text{ML-KEM}^+$ \newline 768 
-        & ML-KEM \newline 1024 & $\text{ML-KEM}^+$ \newline 1024 
+    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        &  $\text{ML-KEM}^+$ \newline 512 
+        &  $\text{ML-KEM}^+$ \newline 768 
+        &  $\text{ML-KEM}^+$ \newline 1024 
         \\
@@
         KE RTT \newline ($\us$) 
-        & 92 & 70 \newline (-23.9\%) 
-        & 135 & 99 \newline (-26.7\%) 
-        & 193 & 138 \newline (-28.5\%) 
+        &  70 \newline (-23.9\%) 
+        &  99 \newline (-26.7\%) 
+        &  138 \newline (-28.5\%) 
         \\
@@
         UAKE RTT \newline ($\us$) 
-        & 145 & 103 \newline (-29.0\%) 
-        & 215 & 144 \newline (-33.0\%) 
-        & 310 & 202 \newline (-34.8\%) 
+        &  103 \newline (-29.0\%) 
+        &  144 \newline (-33.0\%) 
+        &  202 \newline (-34.8\%) 
         \\
@@
         AKE RTT \newline ($\us$) 
-        & 220 & 133 \newline (-39.5\%) 
-        & 294 & 190 \newline (-35.4\%) 
-        & 512 & 266 \newline (-48.0\%) 
+        &  133 \newline (-39.5\%) 
+        &  190 \newline (-35.4\%) 
+        &  266 \newline (-48.0\%) 
         \\
@@
 
-    \caption{Key exchange round-trip times}\label{tbl:rtt-summary}
+    \caption{$\mlkemplus$ achieves substantial RTT savings despite increased encapsulation cost and ciphertext size}\label{tbl:rtt-summary}
 \end{table}
@@
 Under the random oracle model, the chosen-ciphertext security of the OAEP encryption scheme reduces to the one-wayness of the input trapdoor permutation.
-Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA2 security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al later proved that the OAEP is IND-CCA2 secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
-RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended method of RSA-based encryption scheme.
+Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al.
+later proved that the OAEP is IND-CCA secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
+RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended of all RSA-based encryption schemes.
 Unfortunately, OAEP's requirement for a trapdoor permutation is immensely difficult to satisfy, and no other practical instantiation saw widespread adoption to this day.
@@
 The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic chosen-ciphertext secure transformation.
-While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA2 security reduces non-tightly to the OW-CPA security of the input PKE and the IND-CPA security of the symmetric encryption scheme.
-Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build KEM, and proved its security in the quantum random oracle model (QROM).
-
-The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversary.
+While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme.
+Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build a KEM, and proved its security in the quantum random oracle model (QROM).
+
+The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries.
 It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
@@
 
-\section{The ``encrypt-then-MAC'' transformation}\label{sec:main-results}
+\section{The encrypt-then-MAC transformation}\label{sec:main-results}
 Let $\mathcal{B}^\ast$ denote the set of finite bit strings.
@@
 Let $H: \mathcal{B}^\ast \rightarrow \mathcal{K}_\kem$ be a hash function that maps bit strings into the set of possible shared secrets.
-The ``encrypt-then-MAC'' transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen_\kem, \encap, \decap)$, whose routines are described in figure \ref{fig:etm-routines}.
+The encrypt-then-MAC transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen_\kem, \encap, \decap)$, whose routines are described in figure \ref{fig:etm-routines}.
 
@@
 
-On the other hand, if the underlying PKE is one-way secure against plaintext-checking attack that makes $q$ plaintext-checking queries, then ``encrypt-then-MAC'' KEM is semantically secure under chosen ciphertext attacks making the same number of decapsulation queries:
+On the other hand, if the underlying PKE is one-way secure against plaintext-checking attack that makes $q$ plaintext-checking queries, then encrypt-then-MAC KEM is semantically secure under chosen ciphertext attacks making the same number of decapsulation queries:
 
 \begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
-    For every \texttt{IND-CCA2} adversary $A$ against $\kem_\etm$ that makes $q$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes at least $q$ plaintext-checking queries against the underlying $\pke$, and an one-time existential forgery adversary $C$ against the underlying $\mac$ such that
+    For every \texttt{IND-CCA} adversary $A$ against $\kem_\etm$ that makes $q$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes at least $q$ plaintext-checking queries against the underlying $\pke$, and an one-time existential forgery adversary $C$ against the underlying $\mac$ such that
 
     \begin{equation*}
-        \texttt{Adv}_\texttt{IND-CCA2}(A) \leq q \cdot \adv_\texttt{OT-MAC}(C) + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
+        \texttt{Adv}_\texttt{IND-CCA}(A) \leq q \cdot \adv_\texttt{OT-MAC}(C) + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
     \end{equation*}
@@
 \begin{lemma}
-    $\kem_\etm$ is IND-CCA2 secure if and only if the input $\pke$ is OW-PCA secure
+    $\kem_\etm$ is IND-CCA secure if and only if the input $\pke$ is OW-PCA secure
 \end{lemma}
@@
 A summary of the the sequence of games can be found in figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
-From a high level we made three incremental modifications to the IND-CCA2 game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
-A OW-PCA adversary can then simulate the modified IND-CCA2 game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.
+From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
+A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.
 
@@
         \begin{algorithm}[H]
-            \caption*{\texttt{IND-CCA2} game for $\kem_\etm$}
+            \caption*{\texttt{IND-CCA} game for $\kem_\etm$}
             \begin{algorithmic}[1]
@@
 \begin{proof}
-    \emph{Game 0} is the standard IND-CCA2 game for KEMs.
+    \emph{Game 0} is the standard IND-CCA game for KEMs.
 The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
@@
         \end{minipage}
-        \caption{OW-PCA adversary $B$ simulates game 3 for IND-CCA2 adversary $A$}\label{fig:ow-pca-simulates-game-3}
+        \caption{OW-PCA adversary $B$ simulates game 3 for IND-CCA adversary $A$}\label{fig:ow-pca-simulates-game-3}
     \end{figure}
@@
 \subsection{ElGamal is OW-PCA secure}
-We show that the DHAES/DHIES hybrid encryption scheme is a special case of the ``encrypt-then-MAC'' transformation. Specifically, we will sketch a proof of the following lemma:
+We show that the DHAES/DHIES hybrid encryption scheme is a special case of the encrypt-then-MAC transformation. Specifically, we will sketch a proof of the following lemma:
 
@@
 \section{Implementation}\label{sec:implementation}
-ML-KEM is an \texttt{IND-CCA2} secure key encapsulation mechanism standardized by NIST in FIPS 203.
-The \texttt{IND-CCA2} security of ML-KEM is achieved in two steps.
+ML-KEM is an \texttt{IND-CCA} secure key encapsulation mechanism standardized by NIST in FIPS 203.
+The \texttt{IND-CCA} security of ML-KEM is achieved in two steps.
 First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries.
@@
 
-We implemented the ``encrypt-then-MAC'' KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios.
-The experimental data showed that while the ``encrypt-then-MAC'' construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
-
-We refer readers to \cite{FIPS203} for the details of the \texttt{K-PKE} routines. The ``encrypt-then-MAC'' KEM routines are described in figure \ref{fig:ml-kem-plus-routines} below.
+We implemented the encrypt-then-MAC KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios.
+The experimental data showed that while the encrypt-then-MAC construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
+
+We refer readers to \cite{FIPS203} for the details of the \texttt{K-PKE} routines. The encrypt-then-MAC KEM routines are described in figure \ref{fig:ml-kem-plus-routines} below.
 
@@
 \subsection{KEM performance}
-Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the ``encrypt-then-MAC'' transformation the following trade-off when given the same input sub-routines: \begin{enumerate}
+Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the encrypt-then-MAC transformation the following trade-off when given the same input sub-routines: \begin{enumerate}
     \item Both encapsulation and decapsulation add a small amount of overhead for needing to hash both the PKE plaintext and the PKE ciphertext when deriving the shared secret, where as the $U^{\not\bot}_m$ transformation only needs to hash the PKE plaintext.
@@
 
-Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the ``encrypt-then-MAC'' transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
+Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the encrypt-then-MAC transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
 A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in table \ref{tbl:kem-performance}
@@
 \cite{DBLP:conf/eurosp/BosDKLLSSSS18} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE).
-We instantiated an implementation for each of the three key exchange protocols using different variations of the ``encrypt-then-MAC'' KEM and compared round trip time with implementations instantiated using ML-KEM.
+We instantiated an implementation for each of the three key exchange protocols using different variations of the encrypt-then-MAC KEM and compared round trip time with implementations instantiated using ML-KEM.
 
@@
 \section{Conclusions and future works}\label{sec:future-works}
-The ``encrypt-then-MAC'' transformation is a generic KEM construction whose IND-CCA2 security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.
-Compared to the Fujisaki-Okamoto transformation, the ``encrypt-then-MAC'' replaces the computationally expensive re-encryption with computing a MAC tag.
-At the cost of minimal increase in encapsulation cost and ciphertext size, the ``encrypt-then-MAC'' substantially improves the efficiency of the decapsulation routine.
-Where the input PKE's encryption is slower than decryption, the ``encrypt-then-MAC'' KEM achieves meaningful time savings in practical key exchange protocols.
+The encrypt-then-MAC transformation is a generic KEM construction whose IND-CCA security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.
+Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC replaces the computationally expensive re-encryption with computing a MAC tag.
+At the cost of minimal increase in encapsulation cost and ciphertext size, the encrypt-then-MAC substantially improves the efficiency of the decapsulation routine.
+Where the input PKE's encryption is slower than decryption, the encrypt-then-MAC KEM achieves meaningful time savings in practical key exchange protocols.
 
@@
 In fact, as Chris Peikert pointed out in \cite{DBLP:conf/pqcrypto/Peikert14}, most lattice-based cryptosystems are not OW-PCA secure due to the search-decision equivalence of lattice problems.
-While we instantiated ``encrypt-then-MAC'' with ML-KEM subroutines for performance comparison, the resulting KEM is not chosen-ciphertext secure, and should not be used in production systems.
-The natural questions is thus to find a suitable cryptosystem to instantiate ``encrypt-then-MAC'' with.
-
-\textbf{RSA and ElGamal.} Combining ``encrypt-then-MAC'' with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}.
+While we instantiated encrypt-then-MAC with ML-KEM subroutines for performance comparison, the resulting KEM is not chosen-ciphertext secure, and should not be used in production systems.
+The natural questions is thus to find a suitable cryptosystem to instantiate encrypt-then-MAC with.
+
+\textbf{RSA and ElGamal.} Combining encrypt-then-MAC with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}.
 RSA is also known to be OW-PCA secure because it is a trapdoor permutation and thus a rigid PKE.
 However, because of RSA's rigidity, there exists even more efficient KEM transformation (such as RSA-KEM \cite{DBLP:journals/iacr/Shoup01}) that only adds a single hash to the base PKE routines.
-While applying ``encrypt-then-MAC'' to RSA will result in an IND-CCA2 secure KEM, such construction offers no meaningful advantage to RSA-KEM.
+While applying encrypt-then-MAC to RSA will result in an IND-CCA secure KEM, such construction offers no meaningful advantage to RSA-KEM.
 
@@
 Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}.
-On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, though in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying ``encrypt-then-MAC'' may not yield meaningful performance gains.
+On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, though in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying encrypt-then-MAC may not yield meaningful performance gains.
 

### 2024-10-11 etm-transform/manuscript/main.tex
Getting ready for CHES submission

@@
 Our scheme mainly differs from DHIES in its versatility and input requirement.
-Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security).
-In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (Abdalla et al originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
+Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies OW-PCA security).
+In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (M.
+Abdalla et al.
+originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
 
@@
 
-\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against chosen-plaintext attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure.
+\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against plaintext-checking attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure.
 The IND-CCA security of the output KEM reduces tightly to the OW-PCA security of the input PKE, and non-tightly to the unforgeability of the MAC, though the non-tightness will likely not be a weakness because the construction makes it easy to plug in MAC of any desired security level.
 
-\textbf{Implementation and performance comparison.} In Section \ref{sec:implementation}, we instantiate an instance of encrypt-then-MAC KEM using the PKE subroutines defined in ML-KEM \cite{FIPS203}.
-For clarity's sake we call this instantiation $\mlkemplus$.
-Compared with the full KEM routine in ML-KEM, which is based on some variant of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates enormous runtime efficiency gain in decapsulation with only minimal performance penalty in encapsulation runtime and ciphertext size.
-On an AMD EPYC 9R14 CPU, $\mlkemplus$ (combined with Poly1305) achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation.
+\textbf{Implementation and performance comparison.} In Section \ref{sec:practical-instantiation-with-mlkem} and \ref{sec:performance-comparison}, we applied the encrypt-then-MAC transformation to the K-PKE subroutines defined in ML-KEM \cite{FIPS203}, which we call $\mlkemplus$.
+Compared with ML-KEM, which is based on the $U^{\not\bot}$ variant (implicit rejection with rigidity) of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates massive performance improvements in decapsulation while only incurring minimal performance penalty in encapsulation runtime and ciphertext size.
+On an AMD EPYC 9R14 CPU, $\mlkemplus$ with Poly1305 achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation.
 See Table \ref{tbl:cpu-cycles-summary} for a summary of performance comparison of individual routines against ML-KEM.
@@
 
-    \caption{$\text{ML-KEM}^+$ achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
+    \caption{$\mlkemplus$ with Poly1305 achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
 \end{table}
 
-Since the decapsulation efficiency comes at the cost of increasing encryption runtime and ciphertext size, we also measured the round trip time of key exchange protocols with various modes of authentication.
+We also measured the round trip time of key exchange protocols with various modes of authentication.
 When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
@@
 
+\begin{figure}
+    \centering
+
+    \begin{minipage}[t]{0.34\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem^{\not\bot}_m.\keygen()$}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
+                \State $z \leftsample \mathcal{M}$
+                \State $\sk \leftarrow (\sk^\prime, \pk, z)$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem^{\not\bot}_m.\encap(\pk)$}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}$
+                \State $r \leftarrow G(m)$
+                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m ,r)$
+                \State $K \leftarrow H(m)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem^{\not\bot}_m.\decap(\sk, c)$}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
+                \State $\hat{r} \leftarrow G(m)$
+                \State $\hat{c} \leftarrow \texttt{PKE.Enc}(\pk, \hat{m}, \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State $K \leftarrow H(\hat{m})$
+                \Else
+                    \State $K \leftarrow H(z, c)$
+                \EndIf
+                \State \Return $K$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{The $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
+\end{figure}
+
 The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries.
 It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
+The $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.
 
 However, the Fujisaki-Okamoto transformation is not perfect.
-Among its shortfalls are two problems caused by its use of \emph{de-randomization} and \emph{re-encryption} for achieving chosen-ciphertext security.
-\begin{itemize}
+Because it uses de-randomization and re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
     \item \textbf{Computational inefficiency}.
@@
 Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext.
-As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key in fewer than 1000 traces.
+As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key.
 \end{itemize}
@@
 \textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
-$(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
+Key generation $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
 The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key.
 The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key.
-Where the encryption routine is randomized, we denote the randomness by $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space.
+Where the encryption routine is randomized, we denote the randomness by a coin $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space.
 The decryption routine is assumed to always be deterministic.
@@
 For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem.
-On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm.
-For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table.
-We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for details of the mitigation techniques, though decryption failure is not a focus of this paper.
 
 \textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}.
-Each game is defined by the goal and the capabilities of the adversary.
+There are two main types of games: in the one-wayness game, the adversary is given a random encryption, then asked to produce the correct decryption; in the indistinguishability game, the adversary is given the encryption of one of two adversary-chosen plaintexts, then asked to decide which of the plaintexts corresopnds with the given encryption.
+Depending on the attack model, the adversary may have access to various oracles.
+Within the context of public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack (CPA).
+If the adversary has access to a plaintext-checking oracle (PCO) \cite{DBLP:conf/ctrsa/OkamotoP01} then it can mount plaintext-checking attack (PCA).
+Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA).
 
@@
 
-    \begin{minipage}[t]{0.4\textwidth}
+    \begin{minipage}[t]{0.3\textwidth}
         \begin{algorithm}[H]
@@
         \end{algorithm}
-    \end{minipage}\hspace{0.05\textwidth}
-    \begin{minipage}[t]{0.4\textwidth}
+    \end{minipage}
+    \begin{minipage}[t]{0.33\textwidth}
         \begin{algorithm}[H]
@@
     \end{minipage}
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\mathcal{O}_\pco(m, c)$}
+            \begin{algorithmic}[1]
+                \State \Return $\llbrack m = \decrypt(\sk, c) \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+        \begin{algorithm}[H]
+            \caption*{$\mathcal{O}_\decrypt(c)$}
+            \begin{algorithmic}[1]
+                \State \Return $\decrypt(\sk, c)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
     \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
@@
 
-Depending on the attack model, the adversary may have access to various oracles.
-In public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack.
-If the adversary has access to a plaintext-checking oracle $\pco$ then it can mount plaintext-checking attack (PCA).
-Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA).
-A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary with access to the corresponding oracle(s) can win the corresponding game with non-negligible advantage.
+The advantage of an adversary in the \texttt{OW-ATK} game is the probability that it ouputs the correct decryption.
+The advantage of an adversary in the \texttt{IND-ATK} game is defined below.
+A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.
+
+\begin{equation*}
+    \texttt{Adv}_\texttt{IND-ATK}(A) = \left\vert P\left[ \hat{b} = b \right] - \frac{1}{2} \right\vert
+\end{equation*}
 
@@
 The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key.
-The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by
+The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by the adversary's ability to correctly distinguish the two cases beyond blind guess:
 
@@
 
+A KEM is \texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.
+
 \subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
-A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
-The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$.
+A message authentication code $\mac(\keygen, \sign, \verify)$ is a collection of routines defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
+The signing routine $\sign(k, m)$ authenticates the message $m$ under the secret key $k$ by producing a tag $t$ (also called digest).
 The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
@@
     \centering
-    \begin{minipage}{0.7\textwidth}
+    \begin{minipage}{0.53\textwidth}
         \begin{algorithm}[H]
             \caption*{\texttt{EUF-CMA} game}
-
             \begin{algorithmic}[1]
@@
 The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
+Some MACs are one-time existentially unforgeable, meaning that each secret key can be used to authenticate only a single message.
+The corresponding security game is modified such that the signing oracle will only answer a single signing query.
 
 \section{The encrypt-then-MAC transformation}\label{sec:main-results}
+In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM.
+In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC.
+In Section \ref{sec:elgamal-is-ow-pca}, we show that the ``Hashed ElGamal'' cryptosystem in \cite{DBLP:conf/ctrsa/AbdallaBR01} is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.
+
 Let $\mathcal{B}^\ast$ denote the set of finite bit strings.
@@
 Let $H: \mathcal{B}^\ast \rightarrow \mathcal{K}_\kem$ be a hash function that maps bit strings into the set of possible shared secrets.
-The encrypt-then-MAC transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen_\kem, \encap, \decap)$, whose routines are described in figure \ref{fig:etm-routines}.
-
-\begin{figure}[H]
+The encrypt-then-MAC transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen, \encap, \decap)$, whose routines are described in Figure \ref{fig:etm-routines}.
+
+\begin{figure}[h]
     \centering
@@
                 \State $z \leftsample \mathcal{M}$
-                \State $\sk \leftarrow \sk^\prime \Vert z$
+                \State $\sk \leftarrow (\sk^\prime, z)$
                 \State \Return $(\pk, \sk)$
@@
                 \State $K \leftarrow H(m, c^\prime)$
-                \State $c \leftarrow c^\prime \Vert t$
+                \State $c \leftarrow (c^\prime, t)$
                 \State \Return $(c, K)$
@@
     \end{minipage}
-    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
+    \caption{The encrypt-then-MAC transformation builds a KEM $\kem_\etm$ using a $\pke(\keygen, \encrypt, \decrypt)$, a $\mac$, and two hash functions $G, H$}\label{fig:etm-routines} 
 \end{figure}
@@
 
+However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed.
+This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see Figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
+
 \begin{figure}[H]
@@
     \end{minipage}
-    \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}\label{fig:pco-using-decap-oracle}
+    \caption{With the encrypt-then-MAC transformation, every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE}\label{fig:pco-using-decap-oracle}
 \end{figure}
 
-However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed.
-This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
-
-On the other hand, if the underlying PKE is one-way secure against plaintext-checking attack that makes $q$ plaintext-checking queries, then encrypt-then-MAC KEM is semantically secure under chosen ciphertext attacks making the same number of decapsulation queries:
+On the other hand, if the underlying PKE is OW-PCA secure and the underlying MAC is one-time existentially unforgeable, then the encrypt-then-MAC KEM is IND-CCA secure:
 
@@
 
-\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
+\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
 We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game.
-A summary of the the sequence of games can be found in figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
+A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
 From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
@@
     We will bound $P[\texttt{QUERY G}]$ and $P[\texttt{QUERY H}]$ by constructing a OW-PCA adversary $B$ against the underlying PKE that uses $A$ as a sub-routine.
-$B$'s behaviors are summarized in figure \ref{fig:ow-pca-simulates-game-3}.
+$B$'s behaviors are summarized in Figure \ref{fig:ow-pca-simulates-game-3}.
 
@@
 
-\subsection{ElGamal is OW-PCA secure}
+\subsection{ElGamal is OW-PCA secure}\label{sec:elgamal-is-ow-pca}
 We show that the DHAES/DHIES hybrid encryption scheme is a special case of the encrypt-then-MAC transformation. Specifically, we will sketch a proof of the following lemma:
@@
 
-Each ElGamal cryptosystem \cite{DBLP:journals/tit/Elgamal85} is parameterized by a cyclic group $G = \langle g \rangle$ of prime order $q > 2$. A summary of the routine is shown below:
+Each ElGamal cryptosystem \cite{DBLP:journals/tit/Elgamal85} is parameterized by a cyclic group $G = \langle g \rangle$ of prime order $q > 2$.
+A summary of the routine is shown in Figure \ref{fig:elgamal-routines}:
 
@@
 
-    \caption{ElGamal cryptosystem}
+    \caption{ElGamal cryptosystem}\label{fig:elgamal-routines}
 \end{figure}
@@
 \begin{proof}
-    We will prove by sequence of game. A summary can be found in figure \ref{fig:elgamal-pca-games}
-
-    \begin{figure}
+    We will prove by sequence of game. A summary can be found in Figure \ref{fig:elgamal-pca-games}
+
+    \begin{figure}[h]
         \centering
@@
 
-    The Gap Diffie-Hellman adversary $B$ can perfectly simulate game 2 for $A$ (see figure \ref{fig:ow-pca-to-gap-dh}): $B$ receives as the Gap Diffie-Hellman problem inputs $g^x$ and $g^y$.
+    The Gap Diffie-Hellman adversary $B$ can perfectly simulate game 2 for $A$ (see Figure \ref{fig:ow-pca-to-gap-dh}): $B$ receives as the Gap Diffie-Hellman problem inputs $g^x$ and $g^y$.
 $g^x$ simulates an ElGamal public key, where as $g^y$ simulates the first component of the challenge ciphertext.
@@
 
-    \begin{figure}
+    \begin{figure}[h]
         \centering
@@
     If $A$ wins game 2, then its output is $\hat{m} = w^{-x}\cdot v = g^{-xy}\cdot v$, so $m^{-1}\cdot v$ is $g^{xy}$, the correct answer to the Gap Diffie-Hellman problem.
-In other words, $A$ wins game 2 if and only if $B$ solves the Gap Diffie-Hellman problem:
+In other words, $B$ solves its Gap Diffie-Hellman problem if and only if $A$ wins the simulated game 2.
 
@@
 
-\section{Implementation}\label{sec:implementation}
-ML-KEM is an \texttt{IND-CCA} secure key encapsulation mechanism standardized by NIST in FIPS 203.
-The \texttt{IND-CCA} security of ML-KEM is achieved in two steps.
-First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries.
-Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines.
-Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.
-
-We implemented the encrypt-then-MAC KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios.
-The experimental data showed that while the encrypt-then-MAC construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.
-
-We refer readers to \cite{FIPS203} for the details of the \texttt{K-PKE} routines. The encrypt-then-MAC KEM routines are described in figure \ref{fig:ml-kem-plus-routines} below.
-
-\begin{figure}[H]
+\section{Practical instantiation with ML-KEM}\label{sec:practical-instantiation-with-mlkem}
+ML-KEM is an IND-CCA secure key encapsulation mechanism standardized by NIST in FIPS 203 \cite{FIPS203}.
+The chosen-ciphertext security of ML-KEM is achieved in two steps.
+First, ML-KEM constructs a PKE whose IND-CPA security reduces to the conjectured intractability of the decisional Module Learning with Error (MLWE) problem.
+Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM.
+A summary of the IND-CPA PKE routines can be found in Figure \ref{fig:k-pke-routines}.
+We refer readers to \cite{DBLP:conf/tcc/HofheinzHK17} and \cite{FIPS203} for more details of the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation.
+
+\begin{figure}[h]
+    \centering
+
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.KeyGen()}}
+            \begin{algorithmic}[1]
+                \State $A \leftsample R_q^{k \times k}$
+                \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
+                \State $\pk \leftarrow (A, \mathbf{t})$
+                \State $\sk \leftarrow \mathbf{s}$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.Enc(pk, m)}}
+            \begin{algorithmic}[1]
+                \Ensure $\pk = (A, \mathbf{t})$
+                \Ensure $m \in R_2$
+                \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
+                \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
+                \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
+                \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
+                \State \Return $(\mathbf{c}_1, c_2)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.Dec(sk, c)}}
+            \begin{algorithmic}[1]
+                \Ensure $c = (\mathbf{c}_1, c_2)$
+                \Ensure $\sk = \mathbf{s}$
+                \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
+                \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
+                \State \Return $\hat{m}$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{K-PKE is an IND-CPA secure PKE based on the conjectured intractability of the Module Learning with Error (MWLE) problem}\label{fig:k-pke-routines}
+\end{figure}
+
+We applied the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM.
+The resulting KEM is called $\mlkemplus$.
+A complete description of the KEM routines can be found in Figure \ref{fig:ml-kem-plus-routines}
+
+\begin{figure}[h]
     \centering
@@
                 \State $m \leftsample \{0,1\}^{256}$
-                \State $(\overline{K}, r, k) \leftarrow \texttt{XOF}(m \Vert H(\pk))$
-                \State $c^\prime \leftarrow \texttt{K-PKE}.\encrypt(\pk, m, r)$
+                \State $(\overline{K}, k) \leftarrow G(m \Vert H(\pk))$
+                \State $c^\prime \leftsample \texttt{K-PKE}.\encrypt(\pk, m)$
                 \State $t \leftarrow \mac(k, c^\prime)$
@@
                 \State $\hat{m} \leftarrow \texttt{K-PKE}.\decrypt(\sk^\prime, c^\prime)$
-                \State $(\overline{K}, \hat{r}, \hat{k}) \leftarrow \texttt{XOF}(\hat{m} \Vert h)$
+                \State $(\overline{K}, \hat{k}) \leftarrow G(\hat{m} \Vert h)$
                 \State $\hat{t} \leftarrow \mac(\hat{k}, c^\prime)$
@@
 
-    \caption{$\texttt{ML-KEM}^+$ routines}\label{fig:ml-kem-plus-routines}
+    \caption{$\texttt{ML-KEM}^+$ applies our encrypt-then-MAC transformation to K-PKE}\label{fig:ml-kem-plus-routines}
 \end{figure}
 
-Our implementation extended from the reference implementation by the PQCrystals team (https://github.com/pq-crystals/kyber).
-All C code is compiled with \texttt{GCC 11.4.1} and \texttt{OpenSSL 3.0.8}.
-All binaries are executed on an AWS c7a.medium instance with an AMD EPYC 9R14 CPU at 3.7 GHz and 1 GB of RAM.
+When applying our encrypt-then-MAC transformation to K-PKE, we modified the routines described in Figure \ref{fig:etm-routines} such that the public key is added into the derivation of the MAC key and the shared secret.
+There are two main reasons behind this modification.
+First, in a realistic key exchange scenario, the public key will be generated by one party while the random PKE plaintext is sampled by the other party.
+Hashing both the public key and the plaintext ensures that both parties in a key exchange have inputs into deriving the shared secret.
+The second reason is to prevent a dictionary attack on the MAC key.
+If the MAC key is derived on plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the corresponding PKE plaintext.
+Upon receiving the ciphertext, the adversary can search through the MAC keys and recover the PKE plaintext.
+Adding the public key into the derivation of the MAC key will dramatically increase the space requirement and render such dictionary attack infeasible in practice.
 
 \subsection{Choosing a message authenticator}\label{sec:choosing-a-message-authenticator}
-For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
+For implementation, we instantiated the MAC with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
 
 Poly1305 and GMAC are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic.
-Generically speaking, Carter-Wegman MAC is parameterized by some finite field $\mathbb{F}$ and the maximal message length $L > 0$.
-Each symmetric key $k = (k_1, k_2) \leftsample \mathbb{F}^2$ is a pair of uniformly ranodm field elements, and the message is parsed into tuples of field elements up to length $L$: $m = (m_1, m_2, \ldots, m_l) \in \mathbb{F}^{\leq L}$.
-The tag $t$ is computed by evaluating a polynomial whose coefficients the message blocks and whose indeterminate is the key:
-
-\begin{equation}\label{eq:carter-wegman-one-time-mac}
-    \mac((k_1, k_2), m) = H_\text{xpoly}(k_1, m) + k_2
-\end{equation}
-
-Where $H_\text{xpoly}$ is given by:
-
-\begin{equation*}
-    H_\text{xpoly}(k_1, m) = k_1^{l+1} + k_1^l \cdot m_1 + k_1^{l-1} \cdot m_2 + \ldots + k_1 \cdot m_l
-\end{equation*}
-
-The authenticator formulated in equation \ref{eq:carter-wegman-one-time-mac} is a one-time MAC. To make the construction many-time secure, a non-repeating nonce $r$ and a PRF is needed:
-
-\begin{equation*}
-    \mac((k_1, k_2), m, r) = H_\text{xpoly}(k_1, m) \oplus \texttt{PRF}(k_2, r)
-\end{equation*}
+Generically speaking, Carter-Wegman MAC operates by breaking the message into message blocks that can then be parsed into finite field elements.
+The tag is computed by evaluating a polynomial whose coefficients are the message blocks and whose interdeterminate is the secret key.
 
@@
 The final output is XOR'd with a sub key derived from the symmetric key, before being encrypted for one last time.
-%A summary of the computation can be found in figure \ref{fig:aes-256-cmac}
-
-% \begin{figure}[h]
-%     \centering
-%     \begin{minipage}[t]{0.49\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{Sub-key derivation}
-%             \begin{algorithmic}[1]
-%                 \Require 256-bit key $k$
-%                 \Require $\texttt{const\_Rb = 0x87}$
-%                 \State $l \leftarrow \texttt{AES-256}(k, 0^{128})$
-%                 \If{$\texttt{MostSignificantBit}(l) = 0$}
-%                     \State $k_1 \leftarrow \texttt{l {<}{<} 1}$
-%                 \Else 
-%                     \State $k_1 \leftarrow \texttt{l {<}{<} 1} \oplus \texttt{const\_Rb}$
-%                 \EndIf
-%                 \If{$\texttt{MostSignificantBit}(k_1) = 0$}
-%                     \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1}$
-%                 \Else 
-%                     \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1} \oplus \texttt{const\_Rb}$
-%                 \EndIf
-%                 \State \Return $k_1, k_2$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}\hfill
-%     \begin{minipage}[t]{0.49\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{CMAC(k, m)}
-%             \begin{algorithmic}[1]
-%                 \Require 256-bit symmetric key $k$
-%                 \State $(k_1, k_2) \leftarrow \texttt{deriveSubKey}(k)$
-%                 \State $n \leftarrow \left\lceil \texttt{bytesLen}(m) / 16\right\rceil$
-%                 \If{$n = 0$}
-%                     \State $n \leftarrow 1$
-%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
-%                 \ElsIf{$\texttt{bytesLen(m)} \mod 16 = 0$}
-%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_1$
-%                 \Else 
-%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
-%                 \EndIf
-
-%                 \State $x = 0^{128}$
-%                 \For{$i \in \{1, 2, \ldots, n-1\}$}
-%                     \State $y \leftarrow x \oplus m_i$
-%                     \State $x \leftarrow \texttt{AES-256}(k, y)$
-%                 \EndFor
-%                 \State $y \leftarrow m_\texttt{last} \oplus x$
-%                 \State $t \leftarrow \texttt{AES-256}(k, y)$
-%                 \State \Return $t$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}
-%     \caption{AES-256 CMAC}\label{fig:aes-256-cmac}
-% \end{figure}
-
-KMAC is defined in \cite{SP80053r4} to be based on the family of sponge functions with Keccak permutaiton as the underlying function.
+
+KMAC is based on the SHA-3 family of sponge functions \cite{FIPS202}.
 We chose KMAC-256, which uses Shake256 as the underlying extendable output functions.
@@
 
-To isolate the performance characteristics of each authenticator in our instantiation of $\texttt{ML-KEM}^+$, we measured the CPU cycles needed for each authenticator to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of \texttt{ML-KEM}.
-The measurements are summarized in table \ref{tbl:mac-performance}.
-
-From our testing, we found Poly1305 to exhibit the best performance characteristics.
-However, there are additional security considerations that may require the use of other less efficient MAC instances.
-For example, it is possible for an adversary with large computing infrastructure or quantum computers to pre-compute a large lookup table mapping symmetric key to the source plaintext.
-Upon receiving a ciphertext, the adversary can then search through the lookup table for a matching key, which would've revealed the corresponding decryption.
-We partially mitigated such attack by deriving the symmetric key from both the public key and the plaintext, but in case of a long-term keypair, the adversary might still be able to compute a large lookup table AFTER obtaining the long-term public key.
-Further mitigation could include using larger-size keys, which can be accomplished either by using a Carter-Wegman MAC that operates on a larger finite field or using a MAC with a variable key-length such as KMAC.
+\section{Performance comparison}\label{sec:performance-comparison}
+We measured the real-world performance of the individual routines of our $\mlkemplus$ construction, as well as the round trip time of performing key exchange over a network in a variety of authentication modes.
+Our implementation extended from the reference implementation by the PQCrystals team \cite{kyberrefimpl}.
+All C code is compiled with \texttt{GCC 11.4.1} and \texttt{OpenSSL 3.0.8}.
+All binaries are executed on an AWS c7a.medium instance (AMD EPYC 9R14 CPU at 3.7 GHz and 1 GB of RAM) in the \texttt{us-west-2} region.
+
+\subsection{MAC performance}
+To isolate the performance characteristics of each authenticator in our instantiation of $\mlkemplus$ we measured the CPU cycles needed for each authenticator to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of ML-KEM.
+The measurements are summarized in Table \ref{tbl:mac-performance}.
 
@@
 
-\subsection{KEM performance}
-Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the encrypt-then-MAC transformation the following trade-off when given the same input sub-routines: \begin{enumerate}
-    \item Both encapsulation and decapsulation add a small amount of overhead for needing to hash both the PKE plaintext and the PKE ciphertext when deriving the shared secret, where as the $U^{\not\bot}_m$ transformation only needs to hash the PKE plaintext.
-    \item The encapsulation routine adds a small amount of run-time overhead for computing the authenticator
-    \item The decapsulation routine enjoys substantial runtime speedup because \emph{re-encryption} is replaced with computing an authenticator
-    \item Ciphertext size increases by the size of an authenticator
-\end{enumerate}
-
+\subsection{Individual routine performance}
+Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the encrypt-then-MAC transformation achieves massive CPU cycle count reduction in decapsulation while incurring only a minimal increase of encapsulation cycle count and ciphertext size.
 Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the encrypt-then-MAC transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
-A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in table \ref{tbl:kem-performance}
-
-\begin{table}[h]
+A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in Table \ref{tbl:kem-performance}.
+
+\emph{Remark.} We also included the performance of Kyber as it is submitted to the third round of the NIST PQC standardization project, which differs from ML-KEM by deriving the shared secret from hashing the public key, the PKE plaintext, and the PKE ciphertext, while ML-KEM derives its shared secret by hashing only the public key and the PKE plaintext.
+This results in ML-KEM having meaningful performance improvement in encapsulation over Kyber, though the performance difference in decapsulation is negligible.
+
+\begin{table}[H]
     \centering
@@
 A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel.
-\cite{DBLP:conf/eurosp/BosDKLLSSSS18} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE).
-We instantiated an implementation for each of the three key exchange protocols using different variations of the encrypt-then-MAC KEM and compared round trip time with implementations instantiated using ML-KEM.
-
-For clarity, we denote the party who sends the first message to be the client and the other party to be the server.
+We used $\mlkemplus$ to implement the key exchange protocols described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18} and compared the round trip time (RTT) of $\mlkemplus$ key exchange with ML-KEM key exchange.
+We denote the party who sends the first message to be the client and the other party to be the server.
 Round trip time (RTT) is defined to be the time interval between the moment before the client starts generating ephemeral keypairs and the moment after the client derives the final session key.
@@
 
-\subsubsection{Unauthenticated key exchange (KE)}\label{sec:unauthenticated-key-exchange}
-In unauthenticated key exchange, a single pair of ephemeral keypair $(\pk_e, \sk_e) \leftsample \keygen()$ is generated by the client.
+\emph{Remark.} Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
+KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage \cite{DBLP:conf/ccs/SchwabeSW20}.
+
+\subsubsection{Unauthenticated key exchange}\label{sec:unauthenticated-key-exchange}
+In unauthenticated key exchange (KE), a single pair of ephemeral keypair $(\pk_e, \sk_e) \leftsample \keygen()$ is generated by the client.
 The client transmits the ephemeral public key $\pk_e$ to the server, who runs the encapsulation routine $(c_e, K_e) \leftsample \encap(\pk_e)$ and transmits the ciphertext $c_e$ back to the client.
 The client finally decapsulates the ciphertext to recover the shared secret $K_e \leftarrow \decap(\sk_e, c_e)$.
-The key exchange routines are summarized in figure \ref{fig:ke-routines}.
-
-Note that in our implementation, a key derivation function (\texttt{KDF}) is applied to the ephemeral shared secret to derive the final session key.
-This step is added to maintain consistency with other authenticated key exchange protocols, where the final session key is derived from multiple shared secrets.
+The key exchange routines are summarized in Figure \ref{fig:ke-routines}.
 The key derivation function is instantiated using Shake256, and the final session key is 256 bits in length.
-
-\begin{figure}[H]
+The RTT comparison is summarized in Table \ref{tbl:ke-rtt}
+
+\begin{figure}[h]
     \centering
@@
 
-The RTT comparison is summarized in table \ref{tbl:ke-rtt}
-
-\begin{table}[H]
-    \centering
-    \small
+\begin{table}[h]
+    \centering
+    \footnotesize
     \caption{KE RTT comparison}\label{tbl:ke-rtt}
@@
         & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
         \cline{4-5}
@@
 
-\subsubsection{Unilaterally authenticated key exchange (UAKE)}\label{sec:uakex}
-In unilaterally authenticated key exchange, the authenticating party proves its identity to the other party by demonstrating possession of a secret key that corresponds to a published long-term public key.
-In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by demonstrating possession of the corresponding long-term secret key $\sk_S$.
-UAKE routines are summarized in figure \ref{fig:uake-routines}.
-
-In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key.
-This helps the key exchange to achieve weak forward secrecy \cite{DBLP:conf/eurocrypt/CanettiK01}.
-
-Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
-KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage \cite{DBLP:conf/ccs/SchwabeSW20}.
-
-\begin{figure}[H]
+\subsubsection{Unilaterally authenticated key exchange}\label{sec:uakex}
+In unilaterally authenticated key exchange (UAKE), the authenticating party proves its identity to the other party by demonstrating possession of a secret key that corresponds to a published long-term public key.
+In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by correctly decrypting the challenge encryption sent by the client.
+UAKE routines are summarized in Figure \ref{fig:uake-routines}.
+
+\begin{figure}[h]
     \centering
@@
 
-\begin{table}[H]
-    \centering
-    \small
+In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key.
+This helps the key exchange to achieve weak forward secrecy \cite{DBLP:conf/eurocrypt/CanettiK01}.
+Performance data is summaried in Table \ref{tbl:uake-rtt}.
+
+\begin{table}[h]
+    \centering
+    \footnotesize
     \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
@@
 The session key is derived by applying \texttt{KDF} onto the concatenation of shared secrets produced under the ephemeral keypair, server's long-term keypair, and client's long-term keypair, in this order.
-
-\begin{figure}[H]
+the key exchange routines are described in Figure \ref{fig:ake-routines} and the performance data summarized in Table \ref{tbl:ake-rtt}.
+
+\begin{figure}[h]
     \centering
@@
 
-\begin{table}[H]
-    \centering
-    \small
+\begin{table}[h]
+    \centering
+    \footnotesize
     \caption{AKE RTT comparison}\label{tbl:ake-rtt}
@@
 
-Unfortunately, ML-KEM is not OW-PCA secure.
+Unfortunately, ML-KEM's K-PKE subroutines are not OW-PCA secure.
 In fact, as Chris Peikert pointed out in \cite{DBLP:conf/pqcrypto/Peikert14}, most lattice-based cryptosystems are not OW-PCA secure due to the search-decision equivalence of lattice problems.
-While we instantiated encrypt-then-MAC with ML-KEM subroutines for performance comparison, the resulting KEM is not chosen-ciphertext secure, and should not be used in production systems.
-The natural questions is thus to find a suitable cryptosystem to instantiate encrypt-then-MAC with.
-
-\textbf{RSA and ElGamal.} Combining encrypt-then-MAC with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}.
+While we applied the encrypt-then-MAC transformation to K-PKE for performance comparison, $\mlkemplus$ is not chosen-ciphertext secure, and should not be used in production systems.
+The natural questions is thus to find a suitable cryptosystem to apply the encrypt-then-MAC transformation to.
+
+\textbf{RSA and ElGamal.} Combining encrypt-then-MAC with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01} which is proven IND-CCA secure.
 RSA is also known to be OW-PCA secure because it is a trapdoor permutation and thus a rigid PKE.
@@
 
-\textbf{Code-based cryptography.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems may not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
+\textbf{Code-based cryptography.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems does not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
 Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}.
-On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, though in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying encrypt-then-MAC may not yield meaningful performance gains.
+On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, although in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying encrypt-then-MAC may not yield meaningful performance gains.
 
 \textbf{Isogeny-based cryptography.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}.
+In fact, SIKE \cite{azarderakhsh2017supersingular} also uses the Fujisaki-Okamoto transformation, and given the heavy computational cost of isogeny computation, replacing re-encryption with MAC will result in substantial performance improvements.
 While SIKE and SIDH were found to be insecure \cite{DBLP:conf/eurocrypt/CastryckD23}, other isogeny-based cryptosystem such as CSIDH \cite{DBLP:conf/asiacrypt/Castryck0MPR18} remains unaffected by the aforementioned attack and might be suitable candidates.
@@
 
-% \section{Appendix}\label{sec:appendix}
-% % TODO: brief introduction to the various authenticators should be in the appendix
-
-% \subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
-% ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition.
-While their IND-CPA subroutines (see figure \ref{fig:k-pke-routines}) are identical, ML-KEM deviated from Kyber by choosing a different variant of the Fujisaki-Okamoto transformation.
-
-% \begin{figure}[H]
-%     \centering
-
-%     \begin{minipage}[t]{0.33\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{\texttt{K-PKE.KeyGen()}}
-%             \begin{algorithmic}[1]
-%                 \State $A \leftsample R_q^{k \times k}$
-%                 \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
-%                 \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
-%                 \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
-%                 \State $\pk \leftarrow (A, \mathbf{t})$
-%                 \State $\sk \leftarrow \mathbf{s}$
-%                 \State \Return $(\pk, \sk)$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}\hfill
-%     \begin{minipage}[t]{0.33\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{\texttt{K-PKE.Enc(pk, m)}}
-%             \begin{algorithmic}[1]
-%                 \Ensure $\pk = (A, \mathbf{t})$
-%                 \Ensure $m \in R_2$
-%                 \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
-%                 \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
-%                 \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
-%                 \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
-%                 \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
-%                 \State \Return $(\mathbf{c}_1, c_2)$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}\hfill
-%     \begin{minipage}[t]{0.3\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{\texttt{K-PKE.Dec(sk, c)}}
-%             \begin{algorithmic}[1]
-%                 \Ensure $c = (\mathbf{c}_1, c_2)$
-%                 \Ensure $\sk = \mathbf{s}$
-%                 \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
-%                 \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
-%                 \State \Return $\hat{m}$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}
-%     \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
-% \end{figure}
-
-% CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext.
-On the other hand, because by using \emph{re-encryption} and \emph{de-randomization}, the PKE is already made \emph{rigid}, the CRYSTALS-Kyber team decided to use the $U^{\not\bot}_m$ variant, where the shared secret is derived from the plaintext alone.
-
-% \begin{figure}[H]
-%     \centering
-%     \begin{minipage}[t]{0.45\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{\texttt{KEM.KeyGen()}}
-%             \begin{algorithmic}[1]
-%                 \State $z \leftsample \{0,1\}^{256}$
-%                 \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
-%                 \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
-%                 \State \Return $(\pk, \sk)$
-%             \end{algorithmic}
-%         \end{algorithm}
-%         \begin{algorithm}[H]
-%             \caption*{\texttt{KEM.Encap(pk)}}
-%             \begin{algorithmic}[1]
-%                 \State $m \leftsample \{0,1\}^{256}$
-%                 \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
-%                 \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
-%                 \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
-%                     \Comment{$U^{\not\bot}$}
-%                 \State $K \leftarrow \overline{K}$
-%                     \Comment{$U^{\not\bot}_m$}
-%                 \State \Return $(c, K)$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}
-%     \begin{minipage}[t]{0.5\textwidth}
-%         \begin{algorithm}[H]
-%             \caption*{\texttt{KEM.Decap(sk, c)}}
-%             \begin{algorithmic}[1]
-%                 \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
-%                 \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
-%                 \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
-%                 \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
-%                     \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
-%                         \Comment{$U^{\not\bot}$}
-%                     \State $K \leftarrow \overline{K}$
-%                         \Comment{$U^{\not\bot}_m$}
-%                 \Else 
-%                     \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
-%                 \EndIf
-%                 \State \Return $K$
-%             \end{algorithmic}
-%         \end{algorithm}
-%     \end{minipage}
-
-%     \caption{
-%         Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
-%     }\label{fig:kyber-ml-kem-compare}
-% \end{figure}
-
-% The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold.
-The first reason is performance: using the $U^{\not\bot}_m$ transformation saves the need to hash the ciphertext, and since Kyber/ML-KEM's performance is mainly bottlenecked by the symmetric components, omitting the hash leads to significant runtime savings (up to 17\% in AVX-2 optimized implementations).
-The second reason is the simplified security proof and tighter security bounds of the $U^{\not\bot}_m$ variant compared to the $U^{\not\bot}$ variant.
-We will omit the details of the security proof and refer readers to \cite{hofheinz2017modular}.
-
-% In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.
-
 \end{document}

### 2024-10-11 etm-transform/manuscript/main.tex
A sweep across figures and tables

@@
 
-\textbf{Security} The security of KEMs is similarly discussed in adversarial games (figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game.
+\textbf{Security} The security of KEMs is similarly discussed in adversarial games (Figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game.
 In a KEM's indistinguishability game, an adversary is given the public key and a challenge ciphertext, then asked to distinguish a pseudorandom shared secret $K_0$ associated with the challenge ciphertext from a truly random bit string of equal length.
@@
 \subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
-We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game.
+We will prove Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game.
 A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
@@
 
-We now present the proof for lemma \ref{lemma:elgamal-is-ow-pca}
+We now present the proof for Lemma \ref{lemma:elgamal-is-ow-pca}
 \begin{proof}
@@
 
-\textbf{Code-based cryptography.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems does not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
+\textbf{Code-based cryptosystems.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems does not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
 Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}.
@@
 
-\textbf{Isogeny-based cryptography.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}.
+\textbf{Isogeny-based cryptosystems.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}.
 In fact, SIKE \cite{azarderakhsh2017supersingular} also uses the Fujisaki-Okamoto transformation, and given the heavy computational cost of isogeny computation, replacing re-encryption with MAC will result in substantial performance improvements.

### 2024-10-12 etm-transform/manuscript/main.tex
Revised manuscirpt in accordance to suggestions

@@
+% TODO: describe ML-KEM in appendix
+%   Added in Figure 14
+
+% TODO: why is key derived from H(preKey || c') vs H(preKey || t)
+% DONE: that was a typo and was fixed. the implementation hashed the tag, not the ciphertext
+
+% TODO: Decapsulation line 9
+% DONE: ML-KEM still uses implicit rejection, see line 7 of Algorithm 18 in FIPS 203 (https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.203.pdf)
+
+% TODO: Explain the removal of the pseudorandom coin
+% DONE: added paragraph in section 3: "Since the encrypt-then-MAC transformation removes re-encryption..."
+
+% TODO: reference for key exchange protocols
+% DONE: Added to the beginning of section 5.3
+
+% TODO: revise abstract and introduction
+% DONE: revised abstract, reorganized introduction, removed RSA/ElGamal in the last section
+
+% TODO: Why only 128-bit tag?
+% DONE: Added a few sentences in section 4.1
+
 %%%% 1. DOCUMENTCLASS %%%%
@@
 \begin{abstract}
-    Indistinguishability under chosen-ciphertext attack (IND-CCA) is the accepted standard of security for key encapsulation mechanisms (KEM).
-However, provably chosen-ciphertext security is difficult to achieve from scratch.
-In this work, we generalize the ``Hashed ElGamal'' encryption scheme proposed by M.
-Abdalla et al into an IND-CCA secure KEM by combining an OW-PCA secure public-key encryption (PKE) scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC.
-Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
-When instantiated with the PKE sub-routines of ML-KEM, the encrypt-then-MAC transformation achieves massive computational efficiency improvements over ML-KEM in a variety of practical scenarios.
+    ML-KEM is a lattice-based IND-CCA secure key encapsulation mechanism (KEM) standardized by NIST in FIPS 203.
+ML-KEM achieves chosen-ciphertext security by applying the Fujisaki-Okamoto transformation to an IND-CPA secure public-key encryption (PKE) scheme.
+The Fujisaki-Okamoto transformation uses de-randomization and re-encryption to ensure ciphertext non-malleability, but because ML-KEM's underlying PKE encryption is significantly slower than the underlying PKE decryption, ML-KEM's decapsulation performance is dominated by re-encryption.
+In this paper, we propose an alternative generic IND-CCA secure KEM transformation that applies the encrypt-then-MAC pattern to an OW-PCA secure PKE and an existentially unforgeable MAC.
+Compared to the Fujisaki-Okamoto transformation, our encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
+We then implemented $\mlkemplus$ by applying the encrypt-then-MAC transformation to the PKE sub-routines of ML-KEM and a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
+At all three security levels (128-bit, 196-bit, and 256-bit security), our implementation demonstrates great performance improvements in a variety of practical scenarios.
+For example, $\mlkemplus$ with Poly1305 reduces decapsulation's CPU cycles by 72\%-79\% and decreases network round trip time in key exchange protocols by 24\%-48\%.
 \end{abstract}
@@
 \subsection{Our contribution}
-In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES.
+\paragraph{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation} In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES.
 We call our proposed scheme the encrypt-then-MAC transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
+We prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
+In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC for further performance improvements.
+
+\paragraph{$\mlkemplus$: an instantiation of encrypt-then-MAC transformation with ML-KEM's PKE sub-routines.} We implemented $\mlkemplus$ by applying our encrypt-then-MAC transformation to ML-KEM's PKE sub-routines.
+Compared with ML-KEM, which is based on the $U^{\not\bot}$ variant (implicit rejection with rigidity) of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates massive performance improvements in decapsulation while only incurring minimal performance penalty in encapsulation runtime and ciphertext size.
+On an AMD EPYC 9R14 CPU, $\mlkemplus$ with Poly1305 achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation (see Table \ref{tbl:cpu-cycles-summary}).
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        & $\text{ML-KEM}^+$ \newline 512 
+        & $\text{ML-KEM}^+$ \newline 768 
+        & $\text{ML-KEM}^+$ \newline 1024 
+        \\
+        \hline
+        Encap \newline (ccl/tick) 
+        & 93157 \newline (+1.8\%) 
+        & 146405 \newline (+7.3\%) 
+        & 205763 \newline (+3.3\%) 
+        \\
+        \hline
+        Decap \newline (ccl/tick) 
+        & 33733 \newline (-72.2\%) 
+        & 43315 \newline (-76.8\%) 
+        & 51375 \newline (-79.1\%) 
+        \\
+        \hline
+        CT size \newline (bytes) 
+        & 784 \newline (+2.1\%) 
+        & 1104 \newline (+1.5\%) 
+        & 1584 \newline (+1.0\%) 
+        \\
+        \hline
+    \end{tabular}
+
+    \caption{$\mlkemplus$ with Poly1305 achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
+\end{table}
+
+\paragraph{Performance improvements in practical key exchange.} We also implemented the three Kyber key exchange protocols described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18} using $\mlkemplus$ and ML-KEM.
+Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+See Table \ref{tbl:rtt-summary} for a summary of round trip times.
+
+\begin{table}[H]
+    \centering
+    \footnotesize
+
+    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        &  $\text{ML-KEM}^+$ \newline 512 
+        &  $\text{ML-KEM}^+$ \newline 768 
+        &  $\text{ML-KEM}^+$ \newline 1024 
+        \\
+        \hline
+        KE RTT \newline ($\us$) 
+        &  70 \newline (-23.9\%) 
+        &  99 \newline (-26.7\%) 
+        &  138 \newline (-28.5\%) 
+        \\
+        \hline
+        UAKE RTT \newline ($\us$) 
+        &  103 \newline (-29.0\%) 
+        &  144 \newline (-33.0\%) 
+        &  202 \newline (-34.8\%) 
+        \\
+        \hline
+        AKE RTT \newline ($\us$) 
+        &  133 \newline (-39.5\%) 
+        &  190 \newline (-35.4\%) 
+        &  266 \newline (-48.0\%) 
+        \\
+        \hline
+    \end{tabular}
+
+    \caption{$\mlkemplus$ achieves substantial RTT savings despite increased encapsulation cost and ciphertext size}\label{tbl:rtt-summary}
+\end{table}
+
+\subsection{Related works}
+\textbf{Optimal Asymmetric Encryption Padding (OAEP)} \cite{DBLP:conf/eurocrypt/BellareR94}\cite{DBLP:conf/crypto/BellareDPR98} is a generic chosen-ciphertext secure PKE.
+Under the random oracle model, the chosen-ciphertext security of the OAEP encryption scheme reduces to the one-wayness of the input trapdoor permutation.
+Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al.
+later proved that the OAEP is IND-CCA secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
+RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended of all RSA-based encryption schemes.
+Unfortunately, OAEP's requirement for a trapdoor permutation is immensely difficult to satisfy, and no other practical instantiation saw widespread adoption to this day.
+
+The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic chosen-ciphertext secure transformation.
+While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme.
+Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build a KEM, and proved its security in the quantum random oracle model (QROM).
+
+\begin{figure}[h]
+    \centering
+
+    \begin{minipage}[t]{0.34\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem^{\not\bot}_m.\keygen()$}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
+                \State $z \leftsample \mathcal{M}$
+                \State $\sk \leftarrow (\sk^\prime, \pk, z)$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem^{\not\bot}_m.\encap(\pk)$}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}$
+                \State $r \leftarrow G(m)$
+                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m ,r)$
+                \State $K \leftarrow H(m)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\kem^{\not\bot}_m.\decap(\sk, c)$}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
+                \State $\hat{r} \leftarrow G(m)$
+                \State $\hat{c} \leftarrow \texttt{PKE.Enc}(\pk, \hat{m}, \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State $K \leftarrow H(\hat{m})$
+                \Else
+                    \State $K \leftarrow H(z, c)$
+                \EndIf
+                \State \Return $K$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{The $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
+\end{figure}
+
+The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries.
+It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
+The $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.
+
+However, the Fujisaki-Okamoto transformation is not perfect.
+Because it uses de-randomization and re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
+    \item \textbf{Computational inefficiency}.
+The decapsulation routine needs to re-encrypt the decryption to ensure ciphertext has not been tempered with.
+For input PKE whose encryption routine carries substantial computational cost, such as most lattice-based cryptosystems, re-encryption slows down decapsulation significantly.
+    \item \textbf{Side-channel vulnerability}.
+Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext.
+As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key.
+\end{itemize}
+
+The rest of the paper is organized as follows.
+In Section \ref{sec:preliminaries}, we define and discuss the preliminary concepts and theorems.
+In Section \ref{sec:main-results}, we present our encrypt-then-MAC transformation in details, prove its security results, and discusses its relationship to the ElGamal cryptosystem.
+In Section \ref{sec:practical-instantiation-with-mlkem}, we present $\mlkemplus$, an instantiation of the encrypt-then-MAC transformation using ML-KEM's PKE sub-routines, then discuss some implementation rationales.
+In Section \ref{sec:performance-comparison}, we benchmark and compare the performance of individual KEM routines and key exchange round trip times between $\mlkemplus$ and ML-KEM.
+
+\section{Preliminaries}\label{sec:preliminaries}
+
+\subsection{Public-key encryption scheme}
+\textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
+Key generation $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
+The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key.
+The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key.
+Where the encryption routine is randomized, we denote the randomness by a coin $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space.
+The decryption routine is assumed to always be deterministic.
+Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.
+
+\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04}, a $\pke$ is $\delta$-correct if:
+
+\begin{equation*}
+    E\left[\max_{m \in \mathcal{M}} P\left[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)\right]\right] \leq \delta
+\end{equation*}
+
+Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$.
+For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem.
+
+\textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}.
+There are two main types of games: in the one-wayness game, the adversary is given a random encryption, then asked to produce the correct decryption; in the indistinguishability game, the adversary is given the encryption of one of two adversary-chosen plaintexts, then asked to decide which of the plaintexts corresopnds with the given encryption.
+Depending on the attack model, the adversary may have access to various oracles.
+Within the context of public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack (CPA).
+If the adversary has access to a plaintext-checking oracle (PCO) \cite{DBLP:conf/ctrsa/OkamotoP01} then it can mount plaintext-checking attack (PCA).
+Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA).
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{OW-ATK} Game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+                \State $m^\ast \leftsample \mathcal{M}$
+                \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
+                \State $\hat{m} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
+                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{IND-ATK} Game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+                \State $(m_0, m_1) \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk)$
+                \State $b \leftsample \{0,1\}$
+                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
+                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
+                \State \Return $\llbrack \hat{b} = b \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\mathcal{O}_\pco(m, c)$}
+            \begin{algorithmic}[1]
+                \State \Return $\llbrack m = \decrypt(\sk, c) \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+        \begin{algorithm}[H]
+            \caption*{$\mathcal{O}_\decrypt(c)$}
+            \begin{algorithmic}[1]
+                \State \Return $\decrypt(\sk, c)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
+\end{figure}
+
+The advantage of an adversary in the \texttt{OW-ATK} game is the probability that it ouputs the correct decryption.
+The advantage of an adversary in the \texttt{IND-ATK} game is defined below.
+A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.
+
+\begin{equation*}
+    \texttt{Adv}_\texttt{IND-ATK}(A) = \left\vert P\left[ \hat{b} = b \right] - \frac{1}{2} \right\vert
+\end{equation*}
+
+\subsection{Key encapsulation mechanism (KEM)}
+\textbf{Syntax} A key encapsulation mechanism $\kem(\keygen, \encap, \decap)$ is a collection of three routines defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
+The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$.
+$\encap(\pk)$ is a probabilistic routine that takes a public key $\pk$ and outputs a pair of values $(c, K)$ where $c \in \mathcal{C}$ is the ciphertext (also called encapsulation) and $K \in \mathcal{K}$ is the shared secret (also called session key).
+$\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $K$ if the ciphertext is valid.
+Some KEM constructions use explicit rejection, where if $c$ is invalid then $\decap$ will return a rejection symbol $\bot$; other KEM constructions use implicit rejection, where if $c$ is invalid then $\decap$ will return a fake session key that depends on the ciphertext and some other secret values.
+
+\textbf{Security} The security of KEMs is similarly discussed in adversarial games (Figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game.
+In a KEM's indistinguishability game, an adversary is given the public key and a challenge ciphertext, then asked to distinguish a pseudorandom shared secret $K_0$ associated with the challenge ciphertext from a truly random bit string of equal length.
+
+\begin{figure}[h]
+    \centering
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{IND-ATK} game}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
+                \State $(c^\ast, K_0) \leftsample \encap(\pk)$
+                \State $K_1 \leftsample \mathcal{K}$
+                \State $b \leftsample \{0, 1\}$
+                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(
+                    1^\lambda, \pk, c^\ast, K_b
+                )$
+                \State \Return $\llbrack \hat{b} = b \rrbrack$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.49\textwidth}
+        \begin{algorithm}[H]
+        \caption*{$\mathcal{O}_\decap(c)$}
+        \begin{algorithmic}[1]
+            \State \Return $\decap(\sk, c)$
+        \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{\texttt{IND-ATK} game for KEM and decapsulation oracle $\mathcal{O}_\decap$}\label{fig:kem-game}
+\end{figure}
+
+The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key.
+The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by the adversary's ability to correctly distinguish the two cases beyond blind guess:
+
+\begin{equation*}
+    \adv_\texttt{IND-CCA}(A) = \norm{
+        P[A^{\mathcal{O}_\decap}(a^\lambda, \pk, c^\ast, K_b) = b] - \frac{1}{2}
+    }
+\end{equation*}
+
+A KEM is \texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.
+
+\subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
+A message authentication code $\mac(\keygen, \sign, \verify)$ is a collection of routines defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
+The signing routine $\sign(k, m)$ authenticates the message $m$ under the secret key $k$ by producing a tag $t$ (also called digest).
+The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
+Many MAC constructions are deterministic.
+For these constructions it is simpler to denote the signing routine by $t \leftarrow \mac(k, m)$ and perform verification using a simple comparison.
+
+The security of a MAC is defined in an adversarial game in which an adversary, with access to some signing oracle $\mathcal{O}_\sign(m)$, tries to forge a new valid message-tag pair that has never been queried before.
+The existential unforgeability under chosen message attack (EUF-CMA) game is shown below:
+
+\begin{figure}[H]
+    \centering
+    \begin{minipage}{0.53\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{EUF-CMA} game}
+            \begin{algorithmic}[1]
+                \State $k^\ast \leftsample \mathcal{K}$
+                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
+                \State \Return $
+                    \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
+                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
+                    \rrbrack
+                $
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{The existential forgery game}\label{fig:euf-cma-game}
+\end{figure}
+
+The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
+Some MACs are one-time existentially unforgeable, meaning that each secret key can be used to authenticate only a single message.
+The corresponding security game is modified such that the signing oracle will only answer a single signing query.
+
+\section{The encrypt-then-MAC transformation}\label{sec:main-results}
+In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM.
 Our scheme mainly differs from DHIES in its versatility and input requirement.
-Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies OW-PCA security).
+Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:elgamal-is-ow-pca} that the Gap Diffie-Hellman assumption implies OW-PCA security).
 In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (M.
@@
 originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
-
-\begin{figure}
+A summary of encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.
+
+\begin{figure}[h]
     \centering
@@
 
-    \caption{Combining PKE with MAC using encrypt-then-MAC to encapsulate a shared secret. 
-    The returned values are colored grey
-    }\label{fig:etm-diagram}
+    \caption{Combining PKE with MAC using encrypt-then-MAC to ensure ciphertext integrity}\label{fig:etm-diagram}
 \end{figure}
 
-\textbf{Construction.} From a high level, the encrypt-then-MAC transformation takes as input a PKE and a MAC, then outputs a KEM.
-The output KEM encapsulates by encrypting a randomly sampled PKE plaintext and hashing the PKE plaintext-ciphertext pair into the shared secret.
-A symmetric key is derived from hashing the PKE plaintext, then used to produce a MAC tag on the PKE ciphertext.
-At decapsulation, the PKE secret key is used to recover the PKE plaintext, which is then hashed into a symmetric key that is used to verify the MAC tag against the input ciphertext.
-A summary of the data flow in the encrypt-then-MAC transformation can be found in Figure \ref{fig:etm-diagram}.
-A detailed description of individual routines can be found in Figure \ref{fig:etm-routines}.
-
-\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against plaintext-checking attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure.
-The IND-CCA security of the output KEM reduces tightly to the OW-PCA security of the input PKE, and non-tightly to the unforgeability of the MAC, though the non-tightness will likely not be a weakness because the construction makes it easy to plug in MAC of any desired security level.
-
-\textbf{Implementation and performance comparison.} In Section \ref{sec:practical-instantiation-with-mlkem} and \ref{sec:performance-comparison}, we applied the encrypt-then-MAC transformation to the K-PKE subroutines defined in ML-KEM \cite{FIPS203}, which we call $\mlkemplus$.
-Compared with ML-KEM, which is based on the $U^{\not\bot}$ variant (implicit rejection with rigidity) of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates massive performance improvements in decapsulation while only incurring minimal performance penalty in encapsulation runtime and ciphertext size.
-On an AMD EPYC 9R14 CPU, $\mlkemplus$ with Poly1305 achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation.
-See Table \ref{tbl:cpu-cycles-summary} for a summary of performance comparison of individual routines against ML-KEM.
-
-\begin{table}[H]
-    \centering
-    \footnotesize
-
-    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        & $\text{ML-KEM}^+$ \newline 512 
-        & $\text{ML-KEM}^+$ \newline 768 
-        & $\text{ML-KEM}^+$ \newline 1024 
-        \\
-        \hline
-        Encap \newline (ccl/tick) 
-        & 93157 \newline (+1.8\%) 
-        & 146405 \newline (+7.3\%) 
-        & 205763 \newline (+3.3\%) 
-        \\
-        \hline
-        Decap \newline (ccl/tick) 
-        & 33733 \newline (-72.2\%) 
-        & 43315 \newline (-76.8\%) 
-        & 51375 \newline (-79.1\%) 
-        \\
-        \hline
-        CT size \newline (bytes) 
-        & 784 \newline (+2.1\%) 
-        & 1104 \newline (+1.5\%) 
-        & 1584 \newline (+1.0\%) 
-        \\
-        \hline
-    \end{tabular}
-
-    \caption{$\mlkemplus$ with Poly1305 achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
-\end{table}
-
-We also measured the round trip time of key exchange protocols with various modes of authentication.
-When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
-See Table \ref{tbl:rtt-summary} for a summary of round trip times.
-
-\begin{table}[H]
-    \centering
-    \footnotesize
-
-    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        &  $\text{ML-KEM}^+$ \newline 512 
-        &  $\text{ML-KEM}^+$ \newline 768 
-        &  $\text{ML-KEM}^+$ \newline 1024 
-        \\
-        \hline
-        KE RTT \newline ($\us$) 
-        &  70 \newline (-23.9\%) 
-        &  99 \newline (-26.7\%) 
-        &  138 \newline (-28.5\%) 
-        \\
-        \hline
-        UAKE RTT \newline ($\us$) 
-        &  103 \newline (-29.0\%) 
-        &  144 \newline (-33.0\%) 
-        &  202 \newline (-34.8\%) 
-        \\
-        \hline
-        AKE RTT \newline ($\us$) 
-        &  133 \newline (-39.5\%) 
-        &  190 \newline (-35.4\%) 
-        &  266 \newline (-48.0\%) 
-        \\
-        \hline
-    \end{tabular}
-
-    \caption{$\mlkemplus$ achieves substantial RTT savings despite increased encapsulation cost and ciphertext size}\label{tbl:rtt-summary}
-\end{table}
-
-\subsection{Related works}
-\textbf{Optimal Asymmetric Encryption Padding (OAEP)} \cite{DBLP:conf/eurocrypt/BellareR94}\cite{DBLP:conf/crypto/BellareDPR98} is a generic chosen-ciphertext secure PKE.
-Under the random oracle model, the chosen-ciphertext security of the OAEP encryption scheme reduces to the one-wayness of the input trapdoor permutation.
-Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al.
-later proved that the OAEP is IND-CCA secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
-RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended of all RSA-based encryption schemes.
-Unfortunately, OAEP's requirement for a trapdoor permutation is immensely difficult to satisfy, and no other practical instantiation saw widespread adoption to this day.
-
-The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic chosen-ciphertext secure transformation.
-While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme.
-Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build a KEM, and proved its security in the quantum random oracle model (QROM).
-
-\begin{figure}
-    \centering
-
-    \begin{minipage}[t]{0.34\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\kem^{\not\bot}_m.\keygen()$}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
-                \State $z \leftsample \mathcal{M}$
-                \State $\sk \leftarrow (\sk^\prime, \pk, z)$
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.3\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\kem^{\not\bot}_m.\encap(\pk)$}
-            \begin{algorithmic}[1]
-                \State $m \leftsample \mathcal{M}$
-                \State $r \leftarrow G(m)$
-                \State $c \leftarrow \texttt{PKE.Enc}(\pk, m ,r)$
-                \State $K \leftarrow H(m)$
-                \State \Return $(c, K)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.3\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\kem^{\not\bot}_m.\decap(\sk, c)$}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
-                \State $\hat{r} \leftarrow G(m)$
-                \State $\hat{c} \leftarrow \texttt{PKE.Enc}(\pk, \hat{m}, \hat{r})$
-                \If{$\hat{c} = c$}
-                    \State $K \leftarrow H(\hat{m})$
-                \Else
-                    \State $K \leftarrow H(z, c)$
-                \EndIf
-                \State \Return $K$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-
-    \caption{The $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
-\end{figure}
-
-The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries.
-It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
-The $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.
-
-However, the Fujisaki-Okamoto transformation is not perfect.
-Because it uses de-randomization and re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
-    \item \textbf{Computational inefficiency}.
-The decapsulation routine needs to re-encrypt the decryption to ensure ciphertext has not been tempered with.
-For input PKE whose encryption routine carries substantial computational cost, such as most lattice-based cryptosystems, re-encryption slows down decapsulation significantly.
-    \item \textbf{Side-channel vulnerability}.
-Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext.
-As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key.
-\end{itemize}
-
-\section{Preliminaries}
-
-\subsection{Public-key encryption scheme}
-\textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$.
-Key generation $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
-The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key.
-The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key.
-Where the encryption routine is randomized, we denote the randomness by a coin $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space.
-The decryption routine is assumed to always be deterministic.
-Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.
-
-\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04}, a $\pke$ is $\delta$-correct if:
-
-\begin{equation*}
-    E\left[\max_{m \in \mathcal{M}} P\left[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)\right]\right] \leq \delta
-\end{equation*}
-
-Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$.
-For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem.
-
-\textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}.
-There are two main types of games: in the one-wayness game, the adversary is given a random encryption, then asked to produce the correct decryption; in the indistinguishability game, the adversary is given the encryption of one of two adversary-chosen plaintexts, then asked to decide which of the plaintexts corresopnds with the given encryption.
-Depending on the attack model, the adversary may have access to various oracles.
-Within the context of public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack (CPA).
-If the adversary has access to a plaintext-checking oracle (PCO) \cite{DBLP:conf/ctrsa/OkamotoP01} then it can mount plaintext-checking attack (PCA).
-Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA).
-
-\begin{figure}[H]
-    \centering
-
-    \begin{minipage}[t]{0.3\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{OW-ATK} Game}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                \State $m^\ast \leftsample \mathcal{M}$
-                \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
-                \State $\hat{m} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
-                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{IND-ATK} Game}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                \State $(m_0, m_1) \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk)$
-                \State $b \leftsample \{0,1\}$
-                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
-                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
-                \State \Return $\llbrack \hat{b} = b \rrbrack$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\mathcal{O}_\pco(m, c)$}
-            \begin{algorithmic}[1]
-                \State \Return $\llbrack m = \decrypt(\sk, c) \rrbrack$
-            \end{algorithmic}
-        \end{algorithm}
-        \begin{algorithm}[H]
-            \caption*{$\mathcal{O}_\decrypt(c)$}
-            \begin{algorithmic}[1]
-                \State \Return $\decrypt(\sk, c)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
-\end{figure}
-
-The advantage of an adversary in the \texttt{OW-ATK} game is the probability that it ouputs the correct decryption.
-The advantage of an adversary in the \texttt{IND-ATK} game is defined below.
-A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.
-
-\begin{equation*}
-    \texttt{Adv}_\texttt{IND-ATK}(A) = \left\vert P\left[ \hat{b} = b \right] - \frac{1}{2} \right\vert
-\end{equation*}
-
-\subsection{Key encapsulation mechanism (KEM)}
-\textbf{Syntax} A key encapsulation mechanism $\kem(\keygen, \encap, \decap)$ is a collection of three routines defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
-The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$.
-$\encap(\pk)$ is a probabilistic routine that takes a public key $\pk$ and outputs a pair of values $(c, K)$ where $c \in \mathcal{C}$ is the ciphertext (also called encapsulation) and $K \in \mathcal{K}$ is the shared secret (also called session key).
-$\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $K$ if the ciphertext is valid.
-Some KEM constructions use explicit rejection, where if $c$ is invalid then $\decap$ will return a rejection symbol $\bot$; other KEM constructions use implicit rejection, where if $c$ is invalid then $\decap$ will return a fake session key that depends on the ciphertext and some other secret values.
-
-\textbf{Security} The security of KEMs is similarly discussed in adversarial games (Figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game.
-In a KEM's indistinguishability game, an adversary is given the public key and a challenge ciphertext, then asked to distinguish a pseudorandom shared secret $K_0$ associated with the challenge ciphertext from a truly random bit string of equal length.
-
-\begin{figure}[h]
-    \centering
-    \begin{minipage}[t]{0.49\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{IND-ATK} game}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
-                \State $(c^\ast, K_0) \leftsample \encap(\pk)$
-                \State $K_1 \leftsample \mathcal{K}$
-                \State $b \leftsample \{0, 1\}$
-                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(
-                    1^\lambda, \pk, c^\ast, K_b
-                )$
-                \State \Return $\llbrack \hat{b} = b \rrbrack$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.49\textwidth}
-        \begin{algorithm}[H]
-        \caption*{$\mathcal{O}_\decap(c)$}
-        \begin{algorithmic}[1]
-            \State \Return $\decap(\sk, c)$
-        \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{\texttt{IND-ATK} game for KEM and decapsulation oracle $\mathcal{O}_\decap$}\label{fig:kem-game}
-\end{figure}
-
-The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key.
-The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by the adversary's ability to correctly distinguish the two cases beyond blind guess:
-
-\begin{equation*}
-    \adv_\texttt{IND-CCA}(A) = \norm{
-        P[A^{\mathcal{O}_\decap}(a^\lambda, \pk, c^\ast, K_b) = b] - \frac{1}{2}
-    }
-\end{equation*}
-
-A KEM is \texttt{IND-ATK} secure if no efficient adversary has non-negligible advantage in the corresponding security game.
-
-\subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
-A message authentication code $\mac(\keygen, \sign, \verify)$ is a collection of routines defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
-The signing routine $\sign(k, m)$ authenticates the message $m$ under the secret key $k$ by producing a tag $t$ (also called digest).
-The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
-Many MAC constructions are deterministic.
-For these constructions it is simpler to denote the signing routine by $t \leftarrow \mac(k, m)$ and perform verification using a simple comparison.
-
-The security of a MAC is defined in an adversarial game in which an adversary, with access to some signing oracle $\mathcal{O}_\sign(m)$, tries to forge a new valid message-tag pair that has never been queried before.
-The existential unforgeability under chosen message attack (EUF-CMA) game is shown below:
-
-\begin{figure}[H]
-    \centering
-    \begin{minipage}{0.53\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{EUF-CMA} game}
-            \begin{algorithmic}[1]
-                \State $k^\ast \leftsample \mathcal{K}$
-                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
-                \State \Return $
-                    \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
-                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
-                    \rrbrack
-                $
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{The existential forgery game}\label{fig:euf-cma-game}
-\end{figure}
-
-The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.
-Some MACs are one-time existentially unforgeable, meaning that each secret key can be used to authenticate only a single message.
-The corresponding security game is modified such that the signing oracle will only answer a single signing query.
-
-\section{The encrypt-then-MAC transformation}\label{sec:main-results}
-In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM.
 In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC.
@@
 In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext $k \leftarrow G(m)$, then used to sign the unauthenticated ciphertext $c^\prime$.
-Because the encryption routine might be randomized, the session key is derived from both the message and the ciphertext.
 Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$ that would be transmitted to the peer.
@@
 
+Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
+If the input PKE is already rigid, then the shared secret may be derived from hashing the PKE plaintext alone.
+However, if the input PKE is not rigid, then the shared secret must be derived from hashing both the PKE plaintext and the PKE ciphertext.
+
+The chosen-ciphertext security of the encrypt-then-MAC scheme can be intuitively argued through an adversary's inability to learn additional information from the decapsulation oracle.
 For an adversary $A$ to produce a valid tag $t$ for some unauthenticated ciphertext $c^\prime$ under the symmetric key $k \leftarrow G(\decrypt(\sk^\prime, c^\prime))$ implies that $A$ must either know the symmetric key $k$ or produce a forgery.
@@
     \centering
-    \begin{minipage}{0.6\textwidth}
+    \begin{minipage}{0.45\textwidth}
     \begin{algorithm}[H]
@@
     \end{minipage}
-    \caption{With the encrypt-then-MAC transformation, every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE}\label{fig:pco-using-decap-oracle}
+    \caption{With the encrypt-then-MAC transformation, every chosen-ciphertext attack against the KEM implies a plaintext-checking attack against the underlying PKE}\label{fig:pco-using-decap-oracle}
 \end{figure}
@@
 The chosen-ciphertext security of ML-KEM is achieved in two steps.
-First, ML-KEM constructs a PKE whose IND-CPA security reduces to the conjectured intractability of the decisional Module Learning with Error (MLWE) problem.
-Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM.
-A summary of the IND-CPA PKE routines can be found in Figure \ref{fig:k-pke-routines}.
-We refer readers to \cite{DBLP:conf/tcc/HofheinzHK17} and \cite{FIPS203} for more details of the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation.
+First, ML-KEM constructs a PKE (called K-PKE, see Figure \ref{fig:k-pke-routines}) whose IND-CPA security reduces to the conjectured intractability of the decisional Module Learning with Error (MLWE) problem.
+Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM (see Figure \ref{fig:ml-kem-routines}).
 
@@
 
-We applied the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM.
-The resulting KEM is called $\mlkemplus$.
-A complete description of the KEM routines can be found in Figure \ref{fig:ml-kem-plus-routines}
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM.KeyGen()}$}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE.KeyGen()}$
+                \State $h \leftarrow \texttt{SHA3-256}(\pk)$
+                \State $z \leftsample \mathcal{M}$
+                \State $\sk \leftarrow (\sk^\prime, \pk, h, z)$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM.Encap}(\pk)$}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}$
+                \State $h \leftarrow \texttt{SHA3-256}(\pk)$
+                \State $(K, r) \leftarrow \texttt{SHA3-512}(m, h)$
+                \State $c \leftarrow \texttt{K-PKE.Enc}(\pk, m ,r)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM.Decap}(\sk, c)$}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \texttt{K-PKE.Dec}(\sk^\prime, c)$
+                \State $(\overline{K}, \hat{r}) \leftarrow \texttt{SHA3-512}(\hat{m}, h)$
+                \State $\hat{c} \leftarrow \texttt{K-PKE.Enc}(\pk, m, \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State $K \leftarrow \overline{K}$
+                \Else 
+                    \State $K \leftarrow \texttt{SHA3-256}(c, z)$
+                \EndIf
+                \State \Return $K$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{ML-KEM uses the $U^{\not\bot}_m$ variant of the modular Fujisaki-Okamoto KEM transformation}\label{fig:ml-kem-routines}
+\end{figure}
+
+We applied the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM. The resulting KEM is called $\mlkemplus$ (See Figure \ref{fig:ml-kem-plus-routines}).
 
@@
             \begin{algorithmic}[1]
-                \Require Public key $\pk$
                 \State $m \leftsample \{0,1\}^{256}$
@@
                 \State $t \leftarrow \mac(k, c^\prime)$
-                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert c^\prime)$
+                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
                 \State $c \leftarrow (c^\prime, t)$
@@
 
-    \caption{$\texttt{ML-KEM}^+$ applies our encrypt-then-MAC transformation to K-PKE}\label{fig:ml-kem-plus-routines}
+    \caption{$\texttt{ML-KEM}^+$ applies our encrypt-then-MAC transformation to K-PKE. $G$ is SHA3-512, $H$ is SHA3-256, $\texttt{KDF}$ is Shake256}\label{fig:ml-kem-plus-routines}
 \end{figure}
@@
 
+We also modified how the shared secret is derived: instead of hashing the entire ciphertext, we hashed the MAC tag, which essentially functions as a hash of the ciphertext and thus serves the same purpose.
+On the other hand, because ML-KEM ciphertext sizes range from 768 bytes to 1568 bytes while the tag size is fixed at 16 bytes, applying the KDF to the tag instead of the entire ciphertext further yields meaningful performance improvements.
+
 \subsection{Choosing a message authenticator}\label{sec:choosing-a-message-authenticator}
 For implementation, we instantiated the MAC with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
+All chosen MACs are parameterized with 256-bit keys and 128-bit tags.
+We chose 256-bit keys for all three security levels (128-bit, 196-bit, and 256-bit security) because choosing a smaller key size does not yield meaningful performance gains (in addition, Poly1305 only supports 256-bit keys).
+On the other hand, a 128-bit tag size is sufficient for all three security levels because unlike brute-force attacks on the MAC key, which can be carried out offline with large computing capacity, attacks on the tag require interaction with an online peer that plays the role of a MAC verification oracle.
+This means that the attack bandwidth is limited by the throughput of the online oracle instead of the attacker's computing power, and in practice it is easy to implement rate limiting in the cryptographic protocol such that the adversary cannot submit large quantities of MAC verification queries.
 
@@
 
-\begin{table}[H]
+\begin{table}[h]
     \centering
@@
 \subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
-A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel.
-We used $\mlkemplus$ to implement the key exchange protocols described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18} and compared the round trip time (RTT) of $\mlkemplus$ key exchange with ML-KEM key exchange.
+The CRYSTALS-Kyber team \cite{DBLP:conf/eurosp/BosDKLLSSSS18} proposed three key exchange protocols: $\texttt{Kyber.KE}$ for a passively secure ephemeral key exchange, $\texttt{Kyber.UAKE}$ for unilaterally authenticated key exchange, and \texttt{Kyber.AKE} for mutually authenticated key exchange.
+First formally proposed in \cite{DBLP:conf/stoc/BellareCK98} and later analyzed in \cite{DBLP:conf/eurocrypt/CanettiK01}, these key exchange protocols all achieve weak forward secrecy, and the authenticated key exchange protocols are resistant to certain categories of active adversaries (e.g.
+Man-in-the-Middle attacks).
+
+We implemented each of the three Kyber key exchange protocols using $\mlkemplus$ and ML-KEM, then measured the round trip time of each handshake.
 We denote the party who sends the first message to be the client and the other party to be the server.
 Round trip time (RTT) is defined to be the time interval between the moment before the client starts generating ephemeral keypairs and the moment after the client derives the final session key.
-All experiements are run on a pair of AWS c7a.medium instances both located in the \texttt{us-west-2} region.
+All experiments are run on a pair of AWS c7a.medium instances both located in the \texttt{us-west-2} region.
 For each experiment, a total of 10,000 rounds of key exchange are performed, with the median and average round trip time (measured in microsecond) recorded.
@@
 \emph{Remark.} Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
-KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage \cite{DBLP:conf/ccs/SchwabeSW20}.
+KEMTLS \cite{DBLP:conf/ccs/SchwabeSW20} was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage.
 
@@
 
-\textbf{RSA and ElGamal.} Combining encrypt-then-MAC with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01} which is proven IND-CCA secure.
-RSA is also known to be OW-PCA secure because it is a trapdoor permutation and thus a rigid PKE.
-However, because of RSA's rigidity, there exists even more efficient KEM transformation (such as RSA-KEM \cite{DBLP:journals/iacr/Shoup01}) that only adds a single hash to the base PKE routines.
-While applying encrypt-then-MAC to RSA will result in an IND-CCA secure KEM, such construction offers no meaningful advantage to RSA-KEM.
-
 \textbf{Code-based cryptosystems.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems does not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.

### 2024-10-13 etm-transform/manuscript/main.tex
More revisions on manuscript

@@
-% TODO: describe ML-KEM in appendix
-%   Added in Figure 14
-
-% TODO: why is key derived from H(preKey || c') vs H(preKey || t)
-% DONE: that was a typo and was fixed. the implementation hashed the tag, not the ciphertext
-
-% TODO: Decapsulation line 9
-% DONE: ML-KEM still uses implicit rejection, see line 7 of Algorithm 18 in FIPS 203 (https://nvlpubs.nist.gov/nistpubs/FIPS/NIST.FIPS.203.pdf)
-
-% TODO: Explain the removal of the pseudorandom coin
-% DONE: added paragraph in section 3: "Since the encrypt-then-MAC transformation removes re-encryption..."
-
-% TODO: reference for key exchange protocols
-% DONE: Added to the beginning of section 5.3
-
-% TODO: revise abstract and introduction
-% DONE: revised abstract, reorganized introduction, removed RSA/ElGamal in the last section
-
-% TODO: Why only 128-bit tag?
-% DONE: Added a few sentences in section 4.1
+% TODO: add the implementation of the protocols in the abstract.
+% DONE: added "We also implemented key exchange protocols and measured..."
+
+% TODO: remove 768/1024 security level data, add remarks about 128-bit tag security.
+% DONE: removed Poly1305 data with ML-KEM-768/1024
+%   GCM's tag is up to 128 bits: https://nvlpubs.nist.gov/nistpubs/legacy/sp/nistspecialpublication800-38d.pdf
+%   CMAC's tag is up to 128 bits because it uses AES as underlying block cipher
 
@@
 We then implemented $\mlkemplus$ by applying the encrypt-then-MAC transformation to the PKE sub-routines of ML-KEM and a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
-At all three security levels (128-bit, 196-bit, and 256-bit security), our implementation demonstrates great performance improvements in a variety of practical scenarios.
-For example, $\mlkemplus$ with Poly1305 reduces decapsulation's CPU cycles by 72\%-79\% and decreases network round trip time in key exchange protocols by 24\%-48\%.
+At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), $\mlkemplus$ achieves massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
+We also implemented key exchange protocols and measured realistic network round trip time (RTT), where $\mlkemplus$ reduces RTT by 23.9\%-39.5\% compared to ML-KEM.
 \end{abstract}
@@
     \centering
-    \footnotesize
-
-    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        & $\text{ML-KEM}^+$ \newline 512 
-        & $\text{ML-KEM}^+$ \newline 768 
-        & $\text{ML-KEM}^+$ \newline 1024 
-        \\
-        \hline
-        Encap \newline (ccl/tick) 
+    % \footnotesize
+    \small
+
+    % TODO: Poly1305 may not be suitable for 192/256 bit security level
+    % \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+    %     \hline
+    %     & $\text{ML-KEM}^+$ \newline 512 
+    %     & $\text{ML-KEM}^+$ \newline 768 
+    %     & $\text{ML-KEM}^+$ \newline 1024 
+    %     \\
+    %     \hline
+    %     Encap \newline (ccl/tick) 
+    %     & 93157 \newline (+1.8\%) 
+    %     & 146405 \newline (+7.3\%) 
+    %     & 205763 \newline (+3.3\%) 
+    %     \\
+    %     \hline
+    %     Decap \newline (ccl/tick) 
+    %     & 33733 \newline (-72.2\%) 
+    %     & 43315 \newline (-76.8\%) 
+    %     & 51375 \newline (-79.1\%) 
+    %     \\
+    %     \hline
+    %     CT size \newline (bytes) 
+    %     & 784 \newline (+2.1\%) 
+    %     & 1104 \newline (+1.5\%) 
+    %     & 1584 \newline (+1.0\%) 
+    %     \\
+    %     \hline
+    % \end{tabular}
+    \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.4cm}|p{1.4cm}|}
+       \hline
+       & Encap \newline (ccl/tick) 
+       & Decap \newline (ccl/tick) 
+       & CT size \newline (bytes) \\
+       \hline
+       $\mlkemplus$-512 \newline (Poly1305)
        & 93157 \newline (+1.8\%)
-        & 146405 \newline (+7.3\%) 
-        & 205763 \newline (+3.3\%) 
-        \\
-        \hline
-        Decap \newline (ccl/tick) 
        & 33733 \newline (-72.2\%)
-        & 43315 \newline (-76.8\%) 
-        & 51375 \newline (-79.1\%) 
-        \\
-        \hline
-        CT size \newline (bytes) 
-        & 784 \newline (+2.1\%) 
-        & 1104 \newline (+1.5\%) 
-        & 1584 \newline (+1.0\%) 
-        \\
+       & 784 \newline (+2.1\%) \\
+       \hline
+       $\mlkemplus$-512 \newline (GMAC)
+       & 97369 \newline (+6.5\%)
+       & 37725 \newline (-68.9\%)
+       & 784 \newline (+2.1\%) \\
+       \hline
+       $\mlkemplus$-512 \newline (CMAC)
+       & 99739 \newline (+9.0\%)
+       & 40117 \newline (-66.9\%)
+       & 784 \newline (+2.1\%) \\
+       \hline
+       $\mlkemplus$-512 \newline (KMAC-256)
+       & 101009 \newline (+10.4\%)
+       & 40741 \newline (-66.4\%)
+       & 784 \newline (+2.1\%) \\
        \hline
@@
 
-    \caption{$\mlkemplus$ with Poly1305 achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
+    \caption{$\mlkemplus$ achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
 \end{table}
@@
     \centering
-    \footnotesize
-
-    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        &  $\text{ML-KEM}^+$ \newline 512 
-        &  $\text{ML-KEM}^+$ \newline 768 
-        &  $\text{ML-KEM}^+$ \newline 1024 
-        \\
-        \hline
-        KE RTT \newline ($\us$) 
+    % \footnotesize
+    \small
+
+    % TODO: Poly1305 might not be suitable for 192/256 bit security
+    % \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+    %     \hline
+    %     &  $\text{ML-KEM}^+$ \newline 512 
+    %     &  $\text{ML-KEM}^+$ \newline 768 
+    %     &  $\text{ML-KEM}^+$ \newline 1024 
+    %     \\
+    %     \hline
+    %     KE RTT \newline ($\us$) 
+    %     &  70 \newline (-23.9\%) 
+    %     &  99 \newline (-26.7\%) 
+    %     &  138 \newline (-28.5\%) 
+    %     \\
+    %     \hline
+    %     UAKE RTT \newline ($\us$) 
+    %     &  103 \newline (-29.0\%) 
+    %     &  144 \newline (-33.0\%) 
+    %     &  202 \newline (-34.8\%) 
+    %     \\
+    %     \hline
+    %     AKE RTT \newline ($\us$) 
+    %     &  133 \newline (-39.5\%) 
+    %     &  190 \newline (-35.4\%) 
+    %     &  266 \newline (-48.0\%) 
+    %     \\
+    %     \hline
+    % \end{tabular}
+    \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.8cm}|p{1.8cm}|}
+       \hline
+       & KE RTT \newline ($\us$) 
+       & UAKE RTT \newline ($\us$) 
+       & AKE RTT \newline ($\us$) \\
+       \hline
+       $\mlkemplus$-512 \newline (Poly1305)
        & 70 \newline (-23.9\%)
-        &  99 \newline (-26.7\%) 
-        &  138 \newline (-28.5\%) 
-        \\
-        \hline
-        UAKE RTT \newline ($\us$) 
        & 103 \newline (-29.0\%)
-        &  144 \newline (-33.0\%) 
-        &  202 \newline (-34.8\%) 
-        \\
-        \hline
-        AKE RTT \newline ($\us$) 
-        &  133 \newline (-39.5\%) 
-        &  190 \newline (-35.4\%) 
-        &  266 \newline (-48.0\%) 
-        \\
+       & 133 \newline (-39.5\%) \\
+       \hline
+       $\mlkemplus$-512 \newline (GMAC)
+       & 73 \newline (-20.7\%)
+       & 106 \newline (-26.9\%)
+       & 139 \newline (-36.8\%) \\
+       \hline
+       $\mlkemplus$-512 \newline (CMAC)
+       & 75 \newline (-18.5\%)
+       & 108 \newline (-25.5\%)
+       & 143 \newline (-35.0\%) \\
+       \hline
+       $\mlkemplus$-512 \newline (KMAC-256)
+       & 76 \newline (-17.4\%)
+       & 109 \newline (-24.8\%)
+       & 145 \newline (-34.1\%) \\
        \hline
@@
 For implementation, we instantiated the MAC with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
-All chosen MACs are parameterized with 256-bit keys and 128-bit tags.
+
+% All chosen MACs are parameterized with 256-bit keys and 128-bit tags.
 We chose 256-bit keys for all three security levels (128-bit, 196-bit, and 256-bit security) because choosing a smaller key size does not yield meaningful performance gains (in addition, Poly1305 only supports 256-bit keys).
@@
 KMAC allows variable-length key and tag, but we chose the 256 bits for key length and 128 bits for tag size for consistency with other authenticators.
+
+\emph{Remark.} Unfortunately Poly1305, GMAC, and CMAC all use tag size of 128 bits with no straightforward ways to increase the tag size.
+We suspect this small tag size to be unsuitable for instantiation with ML-KEM-768 and ML-KEM-1024 due to their higher security target (192-bit and 256-bit security respectively).
+Consequently, we only instantiate with ML-KEM-512, with a target 128-bit security level.
 
@@
     \centering
-    \footnotesize
+    % \footnotesize
+    \small
     \caption{CPU cycles needed to compute tag on various input sizes}\label{tbl:mac-performance}
@@
         \hline
-        MAC & Median & Average \\
+        MAC & Median (ccl/tick) & Average (ccl/tick)\\
         \hline
@@
     \end{tabular}
-    \begin{tabular}{|c|c|c|}
-        \hline
-        \multicolumn{3}{|c|}{\bf Input size: 1088 bytes} \\
-        \hline
-        MAC & Median & Average \\
-        \hline
-        Poly1305 & 961 & 2704 \\
-        \hline
-        GMAC & 3899 & 4827 \\
-        \hline
-        CMAC & 7305 & 7588 \\
-        \hline
-        KMAC & 9697 & 9928 \\
-        \hline
-    \end{tabular}
-    \begin{tabular}{|c|c|c|}
-        \hline
-        \multicolumn{3}{|c|}{\bf Input size: 1568 bytes} \\
-        \hline
-        MAC & Median & Average \\
-        \hline
-        Poly1305 & 1065 & 1809 \\
-        \hline
-        GMAC & 4055 & 5026 \\
-        \hline
-        CMAC & 8735 & 8772 \\
-        \hline
-        KMAC & 11647 & 12186 \\
-        \hline
-    \end{tabular}
+    % TODO: 128-bit tag size is not suitable for higher security level.
+    % \begin{tabular}{|c|c|c|}
+    %     \hline
+    %     \multicolumn{3}{|c|}{\bf Input size: 1088 bytes} \\
+    %     \hline
+    %     MAC & Median & Average \\
+    %     \hline
+    %     Poly1305 & 961 & 2704 \\
+    %     \hline
+    %     GMAC & 3899 & 4827 \\
+    %     \hline
+    %     CMAC & 7305 & 7588 \\
+    %     \hline
+    %     KMAC & 9697 & 9928 \\
+    %     \hline
+    % \end{tabular}
+    % \begin{tabular}{|c|c|c|}
+    %     \hline
+    %     \multicolumn{3}{|c|}{\bf Input size: 1568 bytes} \\
+    %     \hline
+    %     MAC & Median & Average \\
+    %     \hline
+    %     Poly1305 & 1065 & 1809 \\
+    %     \hline
+    %     GMAC & 4055 & 5026 \\
+    %     \hline
+    %     CMAC & 8735 & 8772 \\
+    %     \hline
+    %     KMAC & 11647 & 12186 \\
+    %     \hline
+    % \end{tabular}
 \end{table}
@@
 
-    \begin{tabular}[t]{|cc|}
-        \hline
-        \multicolumn{2}{|c|}{\bf 192-bit security} \\
-        \hline
-        \multicolumn{2}{|c|}{size parameters (bytes)} \\
-        $\pk$ size & 1184 \\
-        $\sk$ size & 2400 \\
-        \text{ct} size & 1088 \\
-        \hline
-        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
-        Median & 129895 \\
-        Average & 130650 \\
-        \hline
-    \end{tabular}
-    \begin{tabular}[t]{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant} 
-        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
-        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
-        \cline{2-5}
-        & Median & Average & Median & Average \\
-        \hline
-        \text{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
-        \hline
-        \text{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
-        \hline
-        $\text{ML-KEM}^+$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
-        \hline
-        $\text{ML-KEM}^+$-768 w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
-        \hline
-        $\text{ML-KEM}^+$-768 w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
-        \hline
-        $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
-        \hline
-    \end{tabular}\vspace{0.3cm}
-
-    \begin{tabular}[t]{|cc|}
-        \hline
-        \multicolumn{2}{|c|}{\bf 256-bit security} \\
-        \hline
-        \multicolumn{2}{|c|}{size parameters (bytes)} \\
-        $\pk$ size & 1568 \\
-        $\sk$ size & 3168 \\
-        \text{ct} size & 1568 \\
-        \hline
-        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
-        Median & 194921 \\
-        Average & 195465 \\
-        \hline
-    \end{tabular}
-    \begin{tabular}[t]{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant} 
-        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
-        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
-        \cline{2-5}
-        & Median & Average & Median & Average \\
-        \hline
-        \text{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
-        \hline
-        \text{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
-        \hline
-        $\text{ML-KEM}^+$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
-        \hline
-        $\text{ML-KEM}^+$-1024 w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
-        \hline
-        $\text{ML-KEM}^+$-1024 w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
-        \hline
-        $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
-        \hline
-    \end{tabular}
+    % TODO: 128-bit tag size is unsuitable for higher security level
+    % \begin{tabular}[t]{|cc|}
+    %     \hline
+    %     \multicolumn{2}{|c|}{\bf 192-bit security} \\
+    %     \hline
+    %     \multicolumn{2}{|c|}{size parameters (bytes)} \\
+    %     $\pk$ size & 1184 \\
+    %     $\sk$ size & 2400 \\
+    %     \text{ct} size & 1088 \\
+    %     \hline
+    %     \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+    %     Median & 129895 \\
+    %     Average & 130650 \\
+    %     \hline
+    % \end{tabular}
+    % \begin{tabular}[t]{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant} 
+    %     & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+    %     & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+    %     \cline{2-5}
+    %     & Median & Average & Median & Average \\
+    %     \hline
+    %     \text{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
+    %     \hline
+    %     \text{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-768 w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-768 w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
+    %     \hline
+    % \end{tabular}\vspace{0.3cm}
+
+    % \begin{tabular}[t]{|cc|}
+    %     \hline
+    %     \multicolumn{2}{|c|}{\bf 256-bit security} \\
+    %     \hline
+    %     \multicolumn{2}{|c|}{size parameters (bytes)} \\
+    %     $\pk$ size & 1568 \\
+    %     $\sk$ size & 3168 \\
+    %     \text{ct} size & 1568 \\
+    %     \hline
+    %     \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+    %     Median & 194921 \\
+    %     Average & 195465 \\
+    %     \hline
+    % \end{tabular}
+    % \begin{tabular}[t]{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant} 
+    %     & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+    %     & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+    %     \cline{2-5}
+    %     & Median & Average & Median & Average \\
+    %     \hline
+    %     \text{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
+    %     \hline
+    %     \text{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-1024 w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-1024 w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
+    %     \hline
+    %     $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
+    %     \hline
+    % \end{tabular}
 \end{table}
@@
         \hline
-        \texttt{ML-KEM-512} & 800 & 768 & 92 & 97 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 800 & 784 & 70 & 72 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ GMAC & 800 & 784 & 73 & 76 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ CMAC & 800 & 784 & 75 & 79 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ KMAC & 800 & 784 & 76 & 78 \\
+        ML-KEM-512 & 800 & 768 & 92 & 97 \\
+        \hline
+        $\mlkemplus$-512 w/ Poly1305 & 800 & 784 & 70 & 72 \\
+        \hline
+        $\mlkemplus$-512 w/ GMAC & 800 & 784 & 73 & 76 \\
+        \hline
+        $\mlkemplus$-512 w/ CMAC & 800 & 784 & 75 & 79 \\
+        \hline
+        $\mlkemplus$-512 w/ KMAC & 800 & 784 & 76 & 78 \\
         \hline
     \end{tabular}\vspace{0.3cm}
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant}
-        & \multirow{2}{*}{Client TX bytes}
-        & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-        \cline{4-5}
-        & & & Median & Average \\
-        \hline
-        \texttt{ML-KEM-768} & 1184 & 1088 & 135 & 140 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 1184 & 1104 & 99 & 104 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ GMAC & 1184 & 1104 & 101 & 105 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ CMAC & 1184 & 1104 & 103 & 109 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ KMAC & 1184 & 1104 & 103 & 107 \\
-        \hline
-    \end{tabular}\vspace{0.3cm}
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant}
-        & \multirow{2}{*}{Client TX bytes}
-        & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-        \cline{4-5}
-        & & & Median & Average \\
-        \hline
-        \texttt{ML-KEM-1024} & 1568 & 1568 & 193 & 199 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 1568 & 1584 & 138 & 141 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 1568 & 1584 & 140 & 145 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 1568 & 1584 & 143 & 148 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 1568 & 1584 & 144 & 149 \\
-        \hline
-    \end{tabular}
+    % TODO: 128-bit tag size is unsuitable for higher security level
+    % \begin{tabular}{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant}
+    %     & \multirow{2}{*}{Client TX bytes}
+    %     & \multirow{2}{*}{Server TX bytes}
+    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+    %     \cline{4-5}
+    %     & & & Median & Average \\
+    %     \hline
+    %     \texttt{ML-KEM-768} & 1184 & 1088 & 135 & 140 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 1184 & 1104 & 99 & 104 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ GMAC & 1184 & 1104 & 101 & 105 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ CMAC & 1184 & 1104 & 103 & 109 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ KMAC & 1184 & 1104 & 103 & 107 \\
+    %     \hline
+    % \end{tabular}\vspace{0.3cm}
+    % \begin{tabular}{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant}
+    %     & \multirow{2}{*}{Client TX bytes}
+    %     & \multirow{2}{*}{Server TX bytes}
+    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+    %     \cline{4-5}
+    %     & & & Median & Average \\
+    %     \hline
+    %     \texttt{ML-KEM-1024} & 1568 & 1568 & 193 & 199 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 1568 & 1584 & 138 & 141 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ GMAC & 1568 & 1584 & 140 & 145 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ CMAC & 1568 & 1584 & 143 & 148 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ KMAC & 1568 & 1584 & 144 & 149 \\
+    %     \hline
+    % \end{tabular}
 \end{table}
@@
         \hline
-        \texttt{ML-KEM-512} & 1568 & 768 & 145 & 151 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 784 & 103 & 106 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 784 & 106 & 110 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 784 & 108 & 112 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 784 & 109 & 113 \\
+        ML-KEM-512 & 1568 & 768 & 145 & 151 \\
+        \hline
+        $\mlkemplus$-512 w/ Poly1305 & 1584 & 784 & 103 & 106 \\
+        \hline
+        $\mlkemplus$-512 w/ GMAC & 1584 & 784 & 106 & 110 \\
+        \hline
+        $\mlkemplus$-512 w/ CMAC & 1584 & 784 & 108 & 112 \\
+        \hline
+        $\mlkemplus$-512 w/ KMAC & 1584 & 784 & 109 & 113 \\
         \hline
     \end{tabular}\vspace{0.3cm}
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant}
-        & \multirow{2}{*}{Client TX bytes}
-        & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-        \cline{4-5}
-        & & & Median & Average \\
-        \hline
-        \texttt{ML-KEM-768} & 2272 & 1088 & 215 & 222 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 1104 & 149 & 156 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 1104 & 153 & 160 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 1104 & 154 & 159 \\
-        \hline
-    \end{tabular}\vspace{0.3cm}
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant}
-        & \multirow{2}{*}{Client TX bytes}
-        & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-        \cline{4-5}
-        & & & Median & Average \\
-        \hline
-        \texttt{ML-KEM-1024} & 3136 & 1568 & 310 & 318 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 1584 & 212 & 228 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 1584 & 212 & 218 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 1584 & 213 & 220 \\
-        \hline
-    \end{tabular}
+
+    % TODO: 128-bit tag size is unsuitable for higher security level
+    % \begin{tabular}{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant}
+    %     & \multirow{2}{*}{Client TX bytes}
+    %     & \multirow{2}{*}{Server TX bytes}
+    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+    %     \cline{4-5}
+    %     & & & Median & Average \\
+    %     \hline
+    %     \texttt{ML-KEM-768} & 2272 & 1088 & 215 & 222 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 1104 & 149 & 156 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 1104 & 153 & 160 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 1104 & 154 & 159 \\
+    %     \hline
+    % \end{tabular}\vspace{0.3cm}
+    % \begin{tabular}{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant}
+    %     & \multirow{2}{*}{Client TX bytes}
+    %     & \multirow{2}{*}{Server TX bytes}
+    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+    %     \cline{4-5}
+    %     & & & Median & Average \\
+    %     \hline
+    %     \texttt{ML-KEM-1024} & 3136 & 1568 & 310 & 318 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 1584 & 212 & 228 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 1584 & 212 & 218 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 1584 & 213 & 220 \\
+    %     \hline
+    % \end{tabular}
 \end{table}
@@
         \hline
-        \texttt{ML-KEM-512} & 1568 & 1536 & 220 & 213 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 1568 & 133 & 138 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 1568 & 139 & 143 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 1568 & 143 & 148 \\
-        \hline
-        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 1568 & 145 & 151 \\
+        ML-KEM-512 & 1568 & 1536 & 220 & 213 \\
+        \hline
+        $\mlkemplus$-512 w/ Poly1305 & 1584 & 1568 & 133 & 138 \\
+        \hline
+        $\mlkemplus$-512 w/ GMAC & 1584 & 1568 & 139 & 143 \\
+        \hline
+        $\mlkemplus$-512 w/ CMAC & 1584 & 1568 & 143 & 148 \\
+        \hline
+        $\mlkemplus$-512 w/ KMAC & 1584 & 1568 & 145 & 151 \\
         \hline
     \end{tabular}\vspace{0.3cm}
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant}
-        & \multirow{2}{*}{Client TX bytes}
-        & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-        \cline{4-5}
-        & & & Median & Average \\
-        \hline
-        \texttt{ML-KEM-768} & 2272 & 2176 & 294 & 301 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 2208 & 197 & 210 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 2208 & 202 & 208 \\
-        \hline
-        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 2208 & 204 & 210 \\
-        \hline
-    \end{tabular}\vspace{0.3cm}
-    \begin{tabular}{|c|c|c|c|c|}
-        \hline
-        \multirow{2}{*}{KEM variant}
-        & \multirow{2}{*}{Client TX bytes}
-        & \multirow{2}{*}{Server TX bytes}
-        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-        \cline{4-5}
-        & & & Median & Average \\
-        \hline
-        \texttt{ML-KEM-1024} & 3136 & 3136 & 512 & 511 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 3168 & 273 & 282 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 3168 & 280 & 287 \\
-        \hline
-        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 3168 & 282 & 288 \\
-        \hline
-    \end{tabular}
+    % TODO: 128-bit tag size is unsuitable for higher security level
+    % \begin{tabular}{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant}
+    %     & \multirow{2}{*}{Client TX bytes}
+    %     & \multirow{2}{*}{Server TX bytes}
+    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+    %     \cline{4-5}
+    %     & & & Median & Average \\
+    %     \hline
+    %     \texttt{ML-KEM-768} & 2272 & 2176 & 294 & 301 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 2208 & 197 & 210 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 2208 & 202 & 208 \\
+    %     \hline
+    %     $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 2208 & 204 & 210 \\
+    %     \hline
+    % \end{tabular}\vspace{0.3cm}
+    % \begin{tabular}{|c|c|c|c|c|}
+    %     \hline
+    %     \multirow{2}{*}{KEM variant}
+    %     & \multirow{2}{*}{Client TX bytes}
+    %     & \multirow{2}{*}{Server TX bytes}
+    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+    %     \cline{4-5}
+    %     & & & Median & Average \\
+    %     \hline
+    %     \texttt{ML-KEM-1024} & 3136 & 3136 & 512 & 511 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 3168 & 273 & 282 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 3168 & 280 & 287 \\
+    %     \hline
+    %     $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 3168 & 282 & 288 \\
+    %     \hline
+    % \end{tabular}
 \end{table}

### 2024-10-13 etm-transform/manuscript/main.tex
removed confusing paragraph in Section 6

@@
-% TODO: add the implementation of the protocols in the abstract.
-% DONE: added "We also implemented key exchange protocols and measured..."
-
-% TODO: remove 768/1024 security level data, add remarks about 128-bit tag security.
-% DONE: removed Poly1305 data with ML-KEM-768/1024
-%   GCM's tag is up to 128 bits: https://nvlpubs.nist.gov/nistpubs/legacy/sp/nistspecialpublication800-38d.pdf
-%   CMAC's tag is up to 128 bits because it uses AES as underlying block cipher
-
 %%%% 1. DOCUMENTCLASS %%%%
@@
 
-\begin{table}[H]
+\paragraph{C implementation and performance comparison} To observe the real-world performance of our encrypt-then-MAC transformation, we implemented $\mlkemplus$ in C and measured the CPU cycle count for encapsulation and decapsulation routines.
+Compared to ML-KEM, $\mlkemplus$ with Poly1305 achieves 72\%-80\% reduction of CPU cycle count for decapsulation while only increasing encapsulation CPU cycle count by 2\%-7\% (See Table \ref{tbl:cpu-cycles-summary}).
+We also implemented the Kyber key exchange protocols (described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18}) using $\mlkemplus$.
+Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+See Table \ref{tbl:rtt-summary} for a summary of round trip times.
+
+\begin{table}[h]
     \centering
@@
 \end{table}
-
-\paragraph{Performance improvements in practical key exchange.} We also implemented the three Kyber key exchange protocols described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18} using $\mlkemplus$ and ML-KEM.
-Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
-See Table \ref{tbl:rtt-summary} for a summary of round trip times.
-
-\begin{table}[H]
+\begin{table}[h]
     \centering
@@
 
-Unfortunately, ML-KEM's K-PKE subroutines are not OW-PCA secure.
-In fact, as Chris Peikert pointed out in \cite{DBLP:conf/pqcrypto/Peikert14}, most lattice-based cryptosystems are not OW-PCA secure due to the search-decision equivalence of lattice problems.
-While we applied the encrypt-then-MAC transformation to K-PKE for performance comparison, $\mlkemplus$ is not chosen-ciphertext secure, and should not be used in production systems.
-The natural questions is thus to find a suitable cryptosystem to apply the encrypt-then-MAC transformation to.
-
-\textbf{Code-based cryptosystems.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems does not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
+In our future work, we would like to look at applying encryption-then-MAC transformation to other post-quantum cryptosystems for improving their practical implementation efficiency.
+
+\paragraph{Code-based cryptosystems.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems does not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security.
 Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}.
-On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, although in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying encrypt-then-MAC may not yield meaningful performance gains.
-
-\textbf{Isogeny-based cryptosystems.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}.
+On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, although in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying encrypt-then-MAC may yield less impressive performance gains.
+
+\paragraph{Isogeny-based cryptosystems.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}.
 In fact, SIKE \cite{azarderakhsh2017supersingular} also uses the Fujisaki-Okamoto transformation, and given the heavy computational cost of isogeny computation, replacing re-encryption with MAC will result in substantial performance improvements.

### 2024-10-13 etm-transform/manuscript/main.tex
Re-wrote contribution

@@
+% TODO: Section 6 confusion
+% DONE: deleted
+
+% TODO: remove indents
+% DONE: we should not do it; other accepted CHES papers kept the default indent, so we should do the same
+
 %%%% 1. DOCUMENTCLASS %%%%
@@
 % - Add "spthm" for LNCS-like theorems
+
+% No indentation
+% \setlength{\parindent}{0px}
 
@@
 
-\subsection{Our contribution}
-\paragraph{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation} In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES.
-We call our proposed scheme the encrypt-then-MAC transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption.
-We prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
-In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC for further performance improvements.
-
-\paragraph{$\mlkemplus$: an instantiation of encrypt-then-MAC transformation with ML-KEM's PKE sub-routines.} We implemented $\mlkemplus$ by applying our encrypt-then-MAC transformation to ML-KEM's PKE sub-routines.
-Compared with ML-KEM, which is based on the $U^{\not\bot}$ variant (implicit rejection with rigidity) of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates massive performance improvements in decapsulation while only incurring minimal performance penalty in encapsulation runtime and ciphertext size.
-On an AMD EPYC 9R14 CPU, $\mlkemplus$ with Poly1305 achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation (see Table \ref{tbl:cpu-cycles-summary}).
-
-\paragraph{C implementation and performance comparison} To observe the real-world performance of our encrypt-then-MAC transformation, we implemented $\mlkemplus$ in C and measured the CPU cycle count for encapsulation and decapsulation routines.
-Compared to ML-KEM, $\mlkemplus$ with Poly1305 achieves 72\%-80\% reduction of CPU cycle count for decapsulation while only increasing encapsulation CPU cycle count by 2\%-7\% (See Table \ref{tbl:cpu-cycles-summary}).
-We also implemented the Kyber key exchange protocols (described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18}) using $\mlkemplus$.
-Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
-See Table \ref{tbl:rtt-summary} for a summary of round trip times.
+\subsection{Contributions}
+Our main contributions are threefold:
+
+\paragraph{1).
+Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
+In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC such as Poly1305 for further performance improvements.
+
+\paragraph{2).
+$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} We describe $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
+Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
+
+\paragraph{3).
+C implementation and performance comparison.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
+Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
 
@@
 \end{table}
+
+We also implemented the Kyber key exchange protocols \cite{DBLP:conf/eurosp/BosDKLLSSSS18} with $\mlkemplus$ and measured the round trip time (RTT) in realistic network settings.
+Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+See Table \ref{tbl:rtt-summary} for a summary of round trip times.
+
 \begin{table}[h]
@@
 
-The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic chosen-ciphertext secure transformation.
+The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic IND-CCA secure transformation.
 While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme.
@@
 However, the Fujisaki-Okamoto transformation is not perfect.
-Because it uses de-randomization and re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
+Because it uses re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
     \item \textbf{Computational inefficiency}.
@@
 In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC.
-In Section \ref{sec:elgamal-is-ow-pca}, we show that the ``Hashed ElGamal'' cryptosystem in \cite{DBLP:conf/ctrsa/AbdallaBR01} is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.
+In Section \ref{sec:elgamal-is-ow-pca}, we show that DHIES \cite{DBLP:conf/ctrsa/AbdallaBR01} is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.
 

### 2024-10-14 etm-transform/manuscript/main.tex
Pulled updates from Overleaf

@@
     Key encapsulation mechanism, 
-    message authentication code,
-    post-quantum cryptography,
-    lattice cryptography,
+    Message authentication code,
+    Post-quantum cryptography,
+    Lattice cryptography,
     Fujisaki-Okamoto transformation
@@
 The Fujisaki-Okamoto transformation uses de-randomization and re-encryption to ensure ciphertext non-malleability, but because ML-KEM's underlying PKE encryption is significantly slower than the underlying PKE decryption, ML-KEM's decapsulation performance is dominated by re-encryption.
-In this paper, we propose an alternative generic IND-CCA secure KEM transformation that applies the encrypt-then-MAC pattern to an OW-PCA secure PKE and an existentially unforgeable MAC.
+In this paper, we propose an alternative generic IND-CCA secure KEM transformation, called ${\sf KEM}_{{\tt EtM}}$ that applies the encrypt-then-MAC mechanism to an OW-PCA secure PKE and an existentially unforgeable MAC.
 Compared to the Fujisaki-Okamoto transformation, our encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
-We then implemented $\mlkemplus$ by applying the encrypt-then-MAC transformation to the PKE sub-routines of ML-KEM and a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
-At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), $\mlkemplus$ achieves massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
-We also implemented key exchange protocols and measured realistic network round trip time (RTT), where $\mlkemplus$ reduces RTT by 23.9\%-39.5\% compared to ML-KEM.
+    We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it $\mlkemplus$. We then implement $\mlkemplus$ with 
+    %by applying the encrypt-then-MAC transformation to the PKE sub-routines of ML-KEM and 
+    a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
+At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), $\mlkemplus$ achieves a massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
+Furthermore, we implement key exchange protocols and measure realistic network round trip times (RTTs), where $\mlkemplus$ reduces RTTs by 23.9\%-39.5\% compared to ML-KEM.
 \end{abstract}
@@
 It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA security.
-While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
-Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
+\textcolor{red}{The encrypt-then-MAC (EtM) mechanism was standard in ISO 19772 \cite{international2009iso-encrypt-then-mac}.} While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
+\textcolor{red}{Some notable techniques for constructing IND-CCA secure public-key cryptosystems in the standard model include the constructions of Naor and Yung \cite{naor1990public}, Cramer and Shoup \cite{cramer1998practical} and Canetti, Halevi and Katz \cite{CCA-IBE-EUROCRYPT2004}.
+The construction of Naor and Yung uses an IND-CPA secure PKE and a non-interactive zero-knowledge (NIZK) proof system.
+Canetti, Halevi and Katz \cite{CCA-IBE-EUROCRYPT2004} proposed a construction to build an IND-CCA secure PKE scheme from an IND-CPA secure IBE scheme and a one-time signature scheme.} Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
 The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypting the PKE plaintext, then authenticating the PKE ciphertext using the previously derived MAC key.

### 2024-10-14 etm-transform/manuscript/main.tex
Made generic transformation consistent with ML-KEM+ description

@@
 originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
-A summary of encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.
+The encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.
 
@@
                 \State $m \leftsample \mathcal{M}$
-                \State $k \leftarrow G(m)$
+                \State $k \leftarrow G(\pk, m)$
                 \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                 \State $t \leftarrow \mac(k, c^\prime)$
-                \State $K \leftarrow H(m, c^\prime)$
+                \State $K \leftarrow H(\pk, m, t)$
                 \State $c \leftarrow (c^\prime, t)$
@@
                 \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
-                \State $\hat{k} \leftarrow G(\hat{m})$
+                \State $\hat{k} \leftarrow G(\pk, \hat{m})$
                 \If{$\mac(\hat{k}, c^\prime) \neq t$}
-                    \State $K \leftarrow H(z, c^\prime)$
+                    \State $K \leftarrow H(z, t)$
                 \Else
-                    \State $K \leftarrow H(\hat{m}, c^\prime)$
+                    \State $K \leftarrow H(\pk, \hat{m}, t)$
                 \EndIf
@@
 
-Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
+Here are some additional design rationales that have practical impact on security and performance but do not affect the security proof: \begin{itemize}
+    \item \textbf{Removing de-randomization.} Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
 If the input PKE is already rigid, then the shared secret may be derived from hashing the PKE plaintext alone.
 However, if the input PKE is not rigid, then the shared secret must be derived from hashing both the PKE plaintext and the PKE ciphertext.
+    \item \textbf{Hashing the public key.} The public key goes into deriving both the MAC key and the shared secret for two main reasons.
+First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties.
+Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret.
+Second, if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
+Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption.
+Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
+    \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext.
+However, we find it unnecessary to hash the entire ciphertext.
+Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements for shared secret derivation.
+On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
+\end{itemize}
 
@@
 
-When applying our encrypt-then-MAC transformation to K-PKE, we modified the routines described in Figure \ref{fig:etm-routines} such that the public key is added into the derivation of the MAC key and the shared secret.
-There are two main reasons behind this modification.
-First, in a realistic key exchange scenario, the public key will be generated by one party while the random PKE plaintext is sampled by the other party.
-Hashing both the public key and the plaintext ensures that both parties in a key exchange have inputs into deriving the shared secret.
-The second reason is to prevent a dictionary attack on the MAC key.
-If the MAC key is derived on plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the corresponding PKE plaintext.
-Upon receiving the ciphertext, the adversary can search through the MAC keys and recover the PKE plaintext.
-Adding the public key into the derivation of the MAC key will dramatically increase the space requirement and render such dictionary attack infeasible in practice.
-
-We also modified how the shared secret is derived: instead of hashing the entire ciphertext, we hashed the MAC tag, which essentially functions as a hash of the ciphertext and thus serves the same purpose.
-On the other hand, because ML-KEM ciphertext sizes range from 768 bytes to 1568 bytes while the tag size is fixed at 16 bytes, applying the KDF to the tag instead of the entire ciphertext further yields meaningful performance improvements.
+Compared to ML-KEM, $\mlkemplus$ makes the following performance trade-offs: \begin{itemize}
+    \item $\mlkemplus$ encapsulation contains two additional steps: computing the MAC tag and hashing the MAC tag into the shared secret.
+    \item $\mlkemplus$ ciphertext size increases by the size of the MAC tag
+    \item $\mlkemplus$ decapsulation replaces re-encryption with computing the MAC tag
+\end{itemize}
 

### 2024-10-14 etm-transform/manuscript/main.tex
Contributions are itemized

@@
 Compared to the Fujisaki-Okamoto transformation, our encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
-    We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it $\mlkemplus$. We then implement $\mlkemplus$ with 
-    %by applying the encrypt-then-MAC transformation to the PKE sub-routines of ML-KEM and 
-    a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
+    We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it $\mlkemplus$.
+We then implement $\mlkemplus$ with a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
 At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), $\mlkemplus$ achieves a massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
@@
 Some of the earliest proposals for generic IND-CCA secure constructions include OAEP \cite{DBLP:conf/eurocrypt/BellareR94}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.
+Some other notable examples for constructing IND-CCA secure public-key cryptosystems in the standard model include the constructions of Naor and Yung \cite{naor1990public}, Cramer and Shoup \cite{cramer1998practical} and Canetti, Halevi and Katz \cite{CCA-IBE-EUROCRYPT2004}.
+The construction of Naor and Yung uses an IND-CPA secure PKE and a non-interactive zero-knowledge (NIZK) proof system.
+Canetti, Halevi and Katz \cite{CCA-IBE-EUROCRYPT2004} proposed a construction to build an IND-CCA secure PKE scheme from an IND-CPA secure IBE scheme and a one-time signature scheme.
 
@@
 It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA security.
-\textcolor{red}{The encrypt-then-MAC (EtM) mechanism was standard in ISO 19772 \cite{international2009iso-encrypt-then-mac}.} While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
-\textcolor{red}{Some notable techniques for constructing IND-CCA secure public-key cryptosystems in the standard model include the constructions of Naor and Yung \cite{naor1990public}, Cramer and Shoup \cite{cramer1998practical} and Canetti, Halevi and Katz \cite{CCA-IBE-EUROCRYPT2004}.
-The construction of Naor and Yung uses an IND-CPA secure PKE and a non-interactive zero-knowledge (NIZK) proof system.
-Canetti, Halevi and Katz \cite{CCA-IBE-EUROCRYPT2004} proposed a construction to build an IND-CCA secure PKE scheme from an IND-CPA secure IBE scheme and a one-time signature scheme.} Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
+The encrypt-then-MAC (EtM) mechanism was standard in ISO 19772 \cite{international2009iso-encrypt-then-mac}.
+While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
+Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model.
 The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypting the PKE plaintext, then authenticating the PKE ciphertext using the previously derived MAC key.
@@
 \subsection{Contributions}
-Our main contributions are threefold:
-
-\paragraph{1).
-Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
+Our main contributions are threefold: \begin{itemize}
+    \item \textbf{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
 In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC such as Poly1305 for further performance improvements.
-
-\paragraph{2).
-$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} We describe $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
+    \item \textbf{$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} We describe $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
 Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
-
-\paragraph{3).
-C implementation and performance comparison.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
+    \item \textbf{C implementation and performance comparison.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
 Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
+We also implemented the Kyber key exchange protocols \cite{DBLP:conf/eurosp/BosDKLLSSSS18} with $\mlkemplus$ and measured the round trip time (RTT) in realistic network settings.
+Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+See Table \ref{tbl:rtt-summary} for a summary of round trip times.
+\end{itemize}
 
@@
 \end{table}
-
-We also implemented the Kyber key exchange protocols \cite{DBLP:conf/eurosp/BosDKLLSSSS18} with $\mlkemplus$ and measured the round trip time (RTT) in realistic network settings.
-Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
-See Table \ref{tbl:rtt-summary} for a summary of round trip times.
 

### 2024-10-15 etm-transform/manuscript/main.tex
Pulled updates from Overleaf

@@
-% TODO: Section 6 confusion
-% DONE: deleted
-
-% TODO: remove indents
-% DONE: we should not do it; other accepted CHES papers kept the default indent, so we should do the same
-
 %%%% 1. DOCUMENTCLASS %%%%
@@
 Our main contributions are threefold: \begin{itemize}
-    \item \textbf{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
+    \item \textbf{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} \textcolor{red}{\bf New IND-CCA secure KEM construction using encrypt-then-MAC.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
 In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC such as Poly1305 for further performance improvements.
-    \item \textbf{$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} We describe $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
+    \item \textbf{$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} \textcolor{red}{\bf $\mlkemplus$: Efficient CCA-secure KEM based on ML-KEM.} We present $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
 Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
-    \item \textbf{C implementation and performance comparison.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
+    \item \textbf{C implementation and performance comparison.}  \textcolor{red}{\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
 Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
@@
                 \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m^\ast)$
-                \State $k^\ast \leftarrow G(m^\ast)$
+                \State $k^\ast \leftarrow G(\pk, m^\ast)$
                     \Comment{Game 0-1}
@@
                 \State $c^\ast \leftarrow c^\prime \Vert t$
-                \State $K_0 \leftarrow H(m^\ast, c^\prime)$
+                \State $K_0 \leftarrow H(\pk, m^\ast, t)$
                     \Comment{Game 0-2}
@@
                 \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
-                \State $\hat{k} \leftarrow G(\hat{m})$
+                \State $\hat{k} \leftarrow G(\pk, \hat{m})$
                 \If{$\mac(\hat{k}, c^\prime) = t$}
-                    \State $K \leftarrow H(\hat{m}, c^\prime)$
+                    \State $K \leftarrow H(\pk, \hat{m}, t)$
                 \Else 
-                    \State $K \leftarrow H(z, c^\prime)$
+                    \State $K \leftarrow H(z, t)$
                 \EndIf 
@@
 \section{Practical instantiation with ML-KEM}\label{sec:practical-instantiation-with-mlkem}
+
 ML-KEM is an IND-CCA secure key encapsulation mechanism standardized by NIST in FIPS 203 \cite{FIPS203}.
@@
 First, ML-KEM constructs a PKE (called K-PKE, see Figure \ref{fig:k-pke-routines}) whose IND-CPA security reduces to the conjectured intractability of the decisional Module Learning with Error (MLWE) problem.
-Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM (see Figure \ref{fig:ml-kem-routines}).
+Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM (\textcolor{blue}{See see Figure \ref{fig:ml-kem-routines} in Appendix}).
 
@@
 
-We applied the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM. The resulting KEM is called $\mlkemplus$ (See Figure \ref{fig:ml-kem-plus-routines}).
+\subsection{$\mlkemplus$} (\textcolor{blue}{In general, you cannot have only one subsection.}
+
+We apply the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM. The resulting KEM is called $\mlkemplus$ (See Figure \ref{fig:ml-kem-plus-routines}).
 
@@
 
-\subsection{Choosing a message authenticator}\label{sec:choosing-a-message-authenticator}
+\subsection{Choosing a message authentication code (MAC)}\label{sec:choosing-a-message-authenticator}
 For implementation, we instantiated the MAC with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
@@
 
+\subsection{Poly1305 and GMAC with 128-bit tags}
 Poly1305 and GMAC are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic.
 Generically speaking, Carter-Wegman MAC operates by breaking the message into message blocks that can then be parsed into finite field elements.
-The tag is computed by evaluating a polynomial whose coefficients are the message blocks and whose interdeterminate is the secret key.
+The tag is computed by evaluating a polynomial whose coefficients are the message blocks and whose interdeterminate is the secret key, which is also called a \textit{hash key}.
 
@@
 
+\textcolor{blue}{\textbf{ I have revised this part. Please consider to get the data, and revise Table 4, i.e., put back the original 192 and 256 cases without Poly1305. }
+}
+
+In order to get the tag sizes of 192-bit and 256-bit respectively. 
+We can extend GMAC, CMAC, and KMAC to get those tags.
+
+\subsection{Constructions of GMAC for 192-bit and 256-bit tags}
+
+For 192-bit and 256-bit tag size for GMAC, the hash key is given by $h=r_0 || r'$ where $r_0=AES_{256}(k, 0)$  and $r'$ is the most significant 64 bits from $r_1=AES_{256}(k, 1)$ for 192-bit tag and $h=r_0||r_1$ for 256-bit tag.
+Now the cipher polynomial, (i.e., it parses the ciphertext  as 192-bit as  blocks or 256-bit as  blocks)  is  evaluated in $\mathbb{F}_{2^{192}}$ and $\mathbb{F}_{2^{256}}$, respectively.
+
+\subsection{CMAC and an extension to get larger tags}
 CMAC is based on the CBC-MAC with the block cipher instantiated from AES-256.
@@
 
+For the other two cases, it runs AES one more time to get another block of 128 bit.
+Using  the same method in the case for getting  192-bit and 256-bit hash keys in GMAC,  we get the tag with  192-bit and 256-bit respectively.
+
+\subsection{KMAC}
 KMAC is based on the SHA-3 family of sponge functions \cite{FIPS202}.
 We chose KMAC-256, which uses Shake256 as the underlying extendable output functions.
-KMAC allows variable-length key and tag, but we chose the 256 bits for key length and 128 bits for tag size for consistency with other authenticators.
-
-\emph{Remark.} Unfortunately Poly1305, GMAC, and CMAC all use tag size of 128 bits with no straightforward ways to increase the tag size.
+KMAC allows variable-length key and tag, but we chose the 256 bits for key length, since the output of Shake256 can output 256 bits, so, we can get tag with size 128-bit, 192-bit and 256-bit, respectively.
+
+\emph{Remark.} Unfortunately Poly1305 uses tag size of 128 bits with no straightforward ways to increase the tag size.
 We suspect this small tag size to be unsuitable for instantiation with ML-KEM-768 and ML-KEM-1024 due to their higher security target (192-bit and 256-bit security respectively).
-Consequently, we only instantiate with ML-KEM-512, with a target 128-bit security level.
+Consequently, we only instantiate  with ML-KEM-512 with Poly1305, with a target 128-bit security level, instead of the other cases.
 

### 2024-10-15 etm-transform/manuscript/main.tex
Cleaned up proof

@@
 Our main contributions are threefold: \begin{itemize}
-    \item \textbf{Encrypt-then-MAC: a generic IND-CCA secure KEM transformation.} \textcolor{red}{\bf New IND-CCA secure KEM construction using encrypt-then-MAC.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
+    \item {\bf New IND-CCA secure KEM construction using encrypt-then-MAC.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
 In addition, we argue that the encrypt-then-MAC transformation can be instantiated with one-time MAC such as Poly1305 for further performance improvements.
-    \item \textbf{$\mlkemplus$: an application of encrypt-then-MAC transformation to ML-KEM.} \textcolor{red}{\bf $\mlkemplus$: Efficient CCA-secure KEM based on ML-KEM.} We present $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
+    \item {\bf $\mlkemplus$: Efficient CCA-secure KEM based on ML-KEM.} We present $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
 Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
-    \item \textbf{C implementation and performance comparison.}  \textcolor{red}{\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
+    \item {\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
 Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
@@
     \centering
-    % \footnotesize
-    \small
+    \footnotesize
+    % \small
 
     % TODO: Poly1305 may not be suitable for 192/256 bit security level
-    % \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-    %     \hline
-    %     & $\text{ML-KEM}^+$ \newline 512 
-    %     & $\text{ML-KEM}^+$ \newline 768 
-    %     & $\text{ML-KEM}^+$ \newline 1024 
-    %     \\
-    %     \hline
-    %     Encap \newline (ccl/tick) 
+    \begin{tabular}{|p{1.2cm}|p{2.1cm}|p{2.6cm}|p{2.7cm}|}
+        \hline
+        & $\mlkemplus$-512 \newline (Poly1305)
+        & $\mlkemplus$-768 \newline (KMAC 192-bit tag)
+        & $\mlkemplus$-1024 \newline (KMAC 256-bit tag)
+        \\
+        \hline
+        Encap \newline (ccl/tick) 
+        & 91155 \newline (???) 
+        & 150929 \newline (???) 
+        & 228357 \newline (???) 
+        \\
+        \hline
+        Decap \newline (ccl/tick) 
+        & ??? \newline (???) 
+        & ??? \newline (???) 
+        & ??? \newline (???) 
+        \\
+        \hline
+        CT size \newline (bytes) 
+        & ??? \newline (???) 
+        & ??? \newline (???) 
+        & ??? \newline (???) 
+        \\
+        \hline
+    \end{tabular}
+
+    % Alternate version
+    % \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.4cm}|p{1.4cm}|}
+    %    \hline
+    %    & Encap \newline (ccl/tick) 
+    %    & Decap \newline (ccl/tick) 
+    %    & CT size \newline (bytes) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (Poly1305)
     %    & 93157 \newline (+1.8\%)
-    %     & 146405 \newline (+7.3\%) 
-    %     & 205763 \newline (+3.3\%) 
-    %     \\
-    %     \hline
-    %     Decap \newline (ccl/tick) 
     %    & 33733 \newline (-72.2\%)
-    %     & 43315 \newline (-76.8\%) 
-    %     & 51375 \newline (-79.1\%) 
-    %     \\
-    %     \hline
-    %     CT size \newline (bytes) 
-    %     & 784 \newline (+2.1\%) 
-    %     & 1104 \newline (+1.5\%) 
-    %     & 1584 \newline (+1.0\%) 
-    %     \\
+    %    & 784 \newline (+2.1\%) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (GMAC)
+    %    & 97369 \newline (+6.5\%)
+    %    & 37725 \newline (-68.9\%)
+    %    & 784 \newline (+2.1\%) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (CMAC)
+    %    & 99739 \newline (+9.0\%)
+    %    & 40117 \newline (-66.9\%)
+    %    & 784 \newline (+2.1\%) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (KMAC-256)
+    %    & 101009 \newline (+10.4\%)
+    %    & 40741 \newline (-66.4\%)
+    %    & 784 \newline (+2.1\%) \\
     %    \hline
     % \end{tabular}
-    \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.4cm}|p{1.4cm}|}
-       \hline
-       & Encap \newline (ccl/tick) 
-       & Decap \newline (ccl/tick) 
-       & CT size \newline (bytes) \\
-       \hline
-       $\mlkemplus$-512 \newline (Poly1305)
-       & 93157 \newline (+1.8\%)
-       & 33733 \newline (-72.2\%)
-       & 784 \newline (+2.1\%) \\
-       \hline
-       $\mlkemplus$-512 \newline (GMAC)
-       & 97369 \newline (+6.5\%)
-       & 37725 \newline (-68.9\%)
-       & 784 \newline (+2.1\%) \\
-       \hline
-       $\mlkemplus$-512 \newline (CMAC)
-       & 99739 \newline (+9.0\%)
-       & 40117 \newline (-66.9\%)
-       & 784 \newline (+2.1\%) \\
-       \hline
-       $\mlkemplus$-512 \newline (KMAC-256)
-       & 101009 \newline (+10.4\%)
-       & 40741 \newline (-66.4\%)
-       & 784 \newline (+2.1\%) \\
-       \hline
-    \end{tabular}
 
@@
     \centering
-    % \footnotesize
-    \small
+    \footnotesize
+    % \small
 
     % TODO: Poly1305 might not be suitable for 192/256 bit security
-    % \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-    %     \hline
-    %     &  $\text{ML-KEM}^+$ \newline 512 
-    %     &  $\text{ML-KEM}^+$ \newline 768 
-    %     &  $\text{ML-KEM}^+$ \newline 1024 
-    %     \\
-    %     \hline
-    %     KE RTT \newline ($\us$) 
+    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
+        \hline
+        &  $\text{ML-KEM}^+$ \newline 512 
+        &  $\text{ML-KEM}^+$ \newline 768 
+        &  $\text{ML-KEM}^+$ \newline 1024 
+        \\
+        \hline
+        KE RTT \newline ($\us$) 
+        &  ??? \newline (???) 
+        &  ??? \newline (???) 
+        &  ??? \newline (???) 
+        \\
+        \hline
+        UAKE RTT \newline ($\us$) 
+        &  ??? \newline (???) 
+        &  ??? \newline (???) 
+        &  ??? \newline (???) 
+        \\
+        \hline
+        AKE RTT \newline ($\us$) 
+        &  ??? \newline (???) 
+        &  ??? \newline (???) 
+        &  ??? \newline (???) 
+        \\
+        \hline
+    \end{tabular}
+
+    % Alternate version
+    % \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.8cm}|p{1.8cm}|}
+    %    \hline
+    %    & KE RTT \newline ($\us$) 
+    %    & UAKE RTT \newline ($\us$) 
+    %    & AKE RTT \newline ($\us$) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (Poly1305)
     %    & 70 \newline (-23.9\%)
-    %     &  99 \newline (-26.7\%) 
-    %     &  138 \newline (-28.5\%) 
-    %     \\
-    %     \hline
-    %     UAKE RTT \newline ($\us$) 
     %    & 103 \newline (-29.0\%)
-    %     &  144 \newline (-33.0\%) 
-    %     &  202 \newline (-34.8\%) 
-    %     \\
-    %     \hline
-    %     AKE RTT \newline ($\us$) 
-    %     &  133 \newline (-39.5\%) 
-    %     &  190 \newline (-35.4\%) 
-    %     &  266 \newline (-48.0\%) 
-    %     \\
+    %    & 133 \newline (-39.5\%) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (GMAC)
+    %    & 73 \newline (-20.7\%)
+    %    & 106 \newline (-26.9\%)
+    %    & 139 \newline (-36.8\%) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (CMAC)
+    %    & 75 \newline (-18.5\%)
+    %    & 108 \newline (-25.5\%)
+    %    & 143 \newline (-35.0\%) \\
+    %    \hline
+    %    $\mlkemplus$-512 \newline (KMAC-256)
+    %    & 76 \newline (-17.4\%)
+    %    & 109 \newline (-24.8\%)
+    %    & 145 \newline (-34.1\%) \\
     %    \hline
     % \end{tabular}
-    \begin{tabular}{|p{2.5cm}|p{1.4cm}|p{1.8cm}|p{1.8cm}|}
-       \hline
-       & KE RTT \newline ($\us$) 
-       & UAKE RTT \newline ($\us$) 
-       & AKE RTT \newline ($\us$) \\
-       \hline
-       $\mlkemplus$-512 \newline (Poly1305)
-       & 70 \newline (-23.9\%)
-       & 103 \newline (-29.0\%)
-       & 133 \newline (-39.5\%) \\
-       \hline
-       $\mlkemplus$-512 \newline (GMAC)
-       & 73 \newline (-20.7\%)
-       & 106 \newline (-26.9\%)
-       & 139 \newline (-36.8\%) \\
-       \hline
-       $\mlkemplus$-512 \newline (CMAC)
-       & 75 \newline (-18.5\%)
-       & 108 \newline (-25.5\%)
-       & 143 \newline (-35.0\%) \\
-       \hline
-       $\mlkemplus$-512 \newline (KMAC-256)
-       & 76 \newline (-17.4\%)
-       & 109 \newline (-24.8\%)
-       & 145 \newline (-34.1\%) \\
-       \hline
-    \end{tabular}
 
@@
 
-    \caption{The $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
+    \caption{The $\kem^{\not\bot}_m$ variant of Fujisaki-Okamoto transformation is used in ML-KEM}\label{fig:u-notbot-m-routines}
 \end{figure}
@@
 Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:elgamal-is-ow-pca} that the Gap Diffie-Hellman assumption implies OW-PCA security).
-In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (M.
+In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 for further performance improvements (M.
 Abdalla et al.
-originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.
+originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}).
 The encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.
@@
 In Section \ref{sec:proof-of-theorem} we reduce the IND-CCA security of the KEM tightly to the OW-PCA security of the underlying PKE, and non-tightly to the unforgeability of the MAC.
-In Section \ref{sec:elgamal-is-ow-pca}, we show that DHIES \cite{DBLP:conf/ctrsa/AbdallaBR01} is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.
+In Section \ref{sec:elgamal-is-ow-pca}, we show that DHIES is a special case of the encrypt-then-MAC transformation by reducing the OW-PCA security of the ElGamal cryptosystem to the Gap Diffie-Hellman assumption.
 
@@
                 \State $m \leftsample \mathcal{M}$
-                \State $k \leftarrow G(\pk, m)$
+                \State $k \leftarrow G(m)$
                 \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                 \State $t \leftarrow \mac(k, c^\prime)$
-                \State $K \leftarrow H(\pk, m, t)$
                 \State $c \leftarrow (c^\prime, t)$
+                \State $K \leftarrow H(m, c)$
                 \State \Return $(c, K)$
@@
                 \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
-                \State $\hat{k} \leftarrow G(\pk, \hat{m})$
-                \If{$\mac(\hat{k}, c^\prime) \neq t$}
-                    \State $K \leftarrow H(z, t)$
+                \State $\hat{k} \leftarrow G(\hat{m})$
+                \If{$\mac(\hat{k}, c^\prime) = t$}
+                    \State $K \leftarrow H(\hat{m}, c)$
                 \Else
-                    \State $K \leftarrow H(\pk, \hat{m}, t)$
+                    % This is directly copied from ML-KEM, DO NOT CHANGE
+                    \State $K \leftarrow H(z, c)$
                 \EndIf
@@
 The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol.
-In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext $k \leftarrow G(m)$, then used to sign the unauthenticated ciphertext $c^\prime$.
+In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext, then used to sign the unauthenticated ciphertext $c^\prime$.
 Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$ that would be transmitted to the peer.
@@
 
-Here are some additional design rationales that have practical impact on security and performance but do not affect the security proof: \begin{itemize}
-    \item \textbf{Removing de-randomization.} Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
+% Here are some additional design rationales that have practical impact on security and performance but do not affect the security proof: \begin{itemize}
+%     \item \textbf{Removing de-randomization.} Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
 If the input PKE is already rigid, then the shared secret may be derived from hashing the PKE plaintext alone.
 However, if the input PKE is not rigid, then the shared secret must be derived from hashing both the PKE plaintext and the PKE ciphertext.
-    \item \textbf{Hashing the public key.} The public key goes into deriving both the MAC key and the shared secret for two main reasons.
+%     \item \textbf{Hashing the public key.} The public key goes into deriving both the MAC key and the shared secret for two main reasons.
 First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties.
@@
 Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
-    \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext.
+%     \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext.
 However, we find it unnecessary to hash the entire ciphertext.
@@
 On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
-\end{itemize}
+% \end{itemize}
 
@@
 However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed.
-This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see Figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
-
-\begin{figure}[H]
-    \centering
-    \begin{minipage}{0.45\textwidth}
-    \begin{algorithm}[H]
-        \caption*{$\pco(m, c)$}
-        \begin{algorithmic}[1]
-            \State $k \leftarrow G(m)$
-            \State $t \leftarrow \mac(k, c)$
-            \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
-        \end{algorithmic}
-    \end{algorithm}
-    \end{minipage}
-    \caption{With the encrypt-then-MAC transformation, every chosen-ciphertext attack against the KEM implies a plaintext-checking attack against the underlying PKE}\label{fig:pco-using-decap-oracle}
-\end{figure}
+This means that a decapsulation oracle can be converted into a plaintext-checking oracle, so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.
 
@@
 A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
-From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
+From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(\pk, m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, t)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
 A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.
@@
                 \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m^\ast)$
-                \State $k^\ast \leftarrow G(\pk, m^\ast)$
+                \State $k^\ast \leftarrow G(m^\ast)$
                     \Comment{Game 0-1}
@@
                 \State $t \leftarrow \mac(k^\ast, c^\prime)$
-                \State $c^\ast \leftarrow c^\prime \Vert t$
-                \State $K_0 \leftarrow H(\pk, m^\ast, t)$
+                \State $c^\ast \leftarrow (c^\prime, t)$
+                \State $K_0 \leftarrow H(m^\ast, c^\ast)$
                     \Comment{Game 0-2}
@@
         \begin{algorithm}[H]
-            \caption*{$\mathcal{O}^G(m)$}
+            \caption*{Hash oracle $\mathcal{O}^G(m)$}
             \begin{algorithmic}[1]
@@
         \begin{algorithm}[H]
-            \caption*{$\mathcal{O}^\decap(c)$}
+            \caption*{Decap oracle $\mathcal{O}^\decap(c)$}
             \begin{algorithmic}[1]
@@
                 \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
-                \State $\hat{k} \leftarrow G(\pk, \hat{m})$
+                \State $\hat{k} \leftarrow G(\hat{m})$
                 \If{$\mac(\hat{k}, c^\prime) = t$}
-                    \State $K \leftarrow H(\pk, \hat{m}, t)$
+                    \State $K \leftarrow H(\hat{m}, c)$
                 \Else 
-                    \State $K \leftarrow H(z, t)$
+                    \State $K \leftarrow H(z, c)$
                 \EndIf 
@@
                 $}
-                    \State $K \leftarrow H(\tilde{m}, c^\prime)$
+                    \State $K \leftarrow H(\tilde{m}, c)$
                 \Else
-                    \State $K \leftarrow H(z, c^\prime)$
+                    \State $K \leftarrow H(z, c)$
                 \EndIf
@@
 Therefore, the true oracle must also accept the queried ciphertext.
-On the other hand, if the true oracle rejects the queried ciphertext (and output the implicit rejection $H(z, c^\prime)$), then the tag is simply invalid under the MAC key $k = G(\decrypt(\sk^\prime, c^\prime))$.
+On the other hand, if the true oracle rejects the queried ciphertext, then the tag is simply invalid under the MAC key $k = G(\decrypt(\sk^\prime, c^\prime))$.
 Therefore, there could not have been a matching query that also validates the tag, and the simulated oracle must also rejects the queried ciphertext.
@@
 
-    \emph{Game 3} is identical to game 2, except that the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $c^\prime$.
+    \emph{Game 3} is identical to game 2, except that the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $t$.
 From $A$'s perspective the two games are indistinguishable, unless $A$ queries $H$ with $(m^\ast, \cdot)$.
@@
                         % \tilde{m} = \decrypt(\sk^\prime, c^\prime)
-                        \pco(c^\prime, \tilde{m}) = 1
+                        \pco(\tilde{m}, c^\prime) = 1
                         \land \mac(\tilde{k}, c^\prime) = t
                     $}
-                        \State $K \leftarrow H(\tilde{m}, c^\prime)$
+                        \State $K \leftarrow H(\tilde{m}, c)$
                     \Else
-                        \State $K \leftarrow H(z, c^\prime)$
+                        \State $K \leftarrow H(z, c)$
                     \EndIf

### 2024-10-15 etm-transform/manuscript/main.tex
major revision; restored 768/1024 data

@@
         \hline
-        & $\mlkemplus$-512 \newline (Poly1305)
-        & $\mlkemplus$-768 \newline (KMAC 192-bit tag)
-        & $\mlkemplus$-1024 \newline (KMAC 256-bit tag)
+        & $\mlkemplus$-512 \newline Poly1305
+        & $\mlkemplus$-768 \newline KMAC 192-bit tag
+        & $\mlkemplus$-1024 \newline KMAC 256-bit tag
         \\
@@
         Encap \newline (ccl/tick) 
-        & 91155 \newline (???) 
-        & 150929 \newline (???) 
-        & 228357 \newline (???) 
+        & 93157 \newline (+1.8\%) 
+        & 155219 \newline (+13.8\%) 
+        & 228357 \newline (+8.8\%) 
         \\
@@
         Decap \newline (ccl/tick) 
-        & ??? \newline (???) 
-        & ??? \newline (???) 
-        & ??? \newline (???) 
+        & 33733 \newline (-72.2\%) 
+        & 52415 \newline (-71.9\%) 
+        & 62269 \newline (-74.7\%) 
         \\
@@
         CT size \newline (bytes) 
-        & ??? \newline (???) 
-        & ??? \newline (???) 
-        & ??? \newline (???) 
+        & 784 \newline (+2.1\%) 
+        & 1112 \newline (+2.2\%) 
+        & 1600 \newline (+2.0\%) 
         \\
@@
     % TODO: Poly1305 might not be suitable for 192/256 bit security
-    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
-        \hline
-        &  $\text{ML-KEM}^+$ \newline 512 
-        &  $\text{ML-KEM}^+$ \newline 768 
-        &  $\text{ML-KEM}^+$ \newline 1024 
+    \begin{tabular}{|p{1.6cm}|p{2.50cm}|p{2.50cm}|p{2.50cm}|}
+        \hline
+        &  $\text{ML-KEM}^+$-512 \newline Poly1305
+        &  $\text{ML-KEM}^+$-768 \newline KMAC 192-bit tag
+        &  $\text{ML-KEM}^+$-1024 \newline KMAC 256-bit tag
         \\
@@
         KE RTT \newline ($\us$) 
-        &  ??? \newline (???) 
-        &  ??? \newline (???) 
-        &  ??? \newline (???) 
+        &  70 \newline (-23.9\%) 
+        &  103 \newline (-23.7\%)
+        &  144 \newline (-25.4\%) 
         \\
@@
         UAKE RTT \newline ($\us$) 
-        &  ??? \newline (???) 
-        &  ??? \newline (???) 
-        &  ??? \newline (???) 
+        &  103 \newline (-29.0\%) 
+        &  154 \newline (-28.4\%) 
+        &  213 \newline (-32.3\%) 
         \\
@@
         AKE RTT \newline ($\us$) 
-        &  ??? \newline (???) 
-        &  ??? \newline (???) 
-        &  ??? \newline (???) 
+        &  133 \newline (-39.5\%) 
+        &  204 \newline (-30.6\%) 
+        &  282 \newline (-44.9\%) 
         \\
@@
 
-The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol.
+% The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol.
 In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext, then used to sign the unauthenticated ciphertext $c^\prime$.
@@
 
-% Here are some additional design rationales that have practical impact on security and performance but do not affect the security proof: \begin{itemize}
-%     \item \textbf{Removing de-randomization.} Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
+Since the encrypt-then-MAC transformation removes re-encryption in decapsulation, there is no longer the need for fixing the pseudorandom coin $r$ in the PKE's encryption routine.
 If the input PKE is already rigid, then the shared secret may be derived from hashing the PKE plaintext alone.
 However, if the input PKE is not rigid, then the shared secret must be derived from hashing both the PKE plaintext and the PKE ciphertext.
-%     \item \textbf{Hashing the public key.} The public key goes into deriving both the MAC key and the shared secret for two main reasons.
-First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties.
-Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret.
-Second, if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
-Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption.
-Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
-%     \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext.
-However, we find it unnecessary to hash the entire ciphertext.
-Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements for shared secret derivation.
-On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
-% \end{itemize}
 
@@
 The chosen-ciphertext security of ML-KEM is achieved in two steps.
-First, ML-KEM constructs a PKE (called K-PKE, see Figure \ref{fig:k-pke-routines}) whose IND-CPA security reduces to the conjectured intractability of the decisional Module Learning with Error (MLWE) problem.
-Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM (\textcolor{blue}{See see Figure \ref{fig:ml-kem-routines} in Appendix}).
-
-\begin{figure}[h]
-    \centering
-
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{K-PKE.KeyGen()}}
-            \begin{algorithmic}[1]
-                \State $A \leftsample R_q^{k \times k}$
-                \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
-                \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
-                \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
-                \State $\pk \leftarrow (A, \mathbf{t})$
-                \State $\sk \leftarrow \mathbf{s}$
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.33\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{K-PKE.Enc(pk, m)}}
-            \begin{algorithmic}[1]
-                \Ensure $\pk = (A, \mathbf{t})$
-                \Ensure $m \in R_2$
-                \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
-                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
-                \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
-                \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
-                \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
-                \State \Return $(\mathbf{c}_1, c_2)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{0.3\textwidth}
-        \begin{algorithm}[H]
-            \caption*{\texttt{K-PKE.Dec(sk, c)}}
-            \begin{algorithmic}[1]
-                \Ensure $c = (\mathbf{c}_1, c_2)$
-                \Ensure $\sk = \mathbf{s}$
-                \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
-                \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
-                \State \Return $\hat{m}$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \caption{K-PKE is an IND-CPA secure PKE based on the conjectured intractability of the Module Learning with Error (MWLE) problem}\label{fig:k-pke-routines}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-
-    \begin{minipage}[t]{0.4\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\texttt{ML-KEM.KeyGen()}$}
-            \begin{algorithmic}[1]
-                \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE.KeyGen()}$
-                \State $h \leftarrow \texttt{SHA3-256}(\pk)$
-                \State $z \leftsample \mathcal{M}$
-                \State $\sk \leftarrow (\sk^\prime, \pk, h, z)$
-                \State \Return $(\pk, \sk)$
-            \end{algorithmic}
-        \end{algorithm}
-        \begin{algorithm}[H]
-            \caption*{$\texttt{ML-KEM.Encap}(\pk)$}
-            \begin{algorithmic}[1]
-                \State $m \leftsample \mathcal{M}$
-                \State $h \leftarrow \texttt{SHA3-256}(\pk)$
-                \State $(K, r) \leftarrow \texttt{SHA3-512}(m, h)$
-                \State $c \leftarrow \texttt{K-PKE.Enc}(\pk, m ,r)$
-                \State \Return $(c, K)$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-    \begin{minipage}[t]{0.4\textwidth}
-        \begin{algorithm}[H]
-            \caption*{$\texttt{ML-KEM.Decap}(\sk, c)$}
-            \begin{algorithmic}[1]
-                \State $\hat{m} \leftarrow \texttt{K-PKE.Dec}(\sk^\prime, c)$
-                \State $(\overline{K}, \hat{r}) \leftarrow \texttt{SHA3-512}(\hat{m}, h)$
-                \State $\hat{c} \leftarrow \texttt{K-PKE.Enc}(\pk, m, \hat{r})$
-                \If{$\hat{c} = c$}
-                    \State $K \leftarrow \overline{K}$
-                \Else 
-                    \State $K \leftarrow \texttt{SHA3-256}(c, z)$
-                \EndIf
-                \State \Return $K$
-            \end{algorithmic}
-        \end{algorithm}
-    \end{minipage}
-
-    \caption{ML-KEM uses the $U^{\not\bot}_m$ variant of the modular Fujisaki-Okamoto KEM transformation}\label{fig:ml-kem-routines}
-\end{figure}
-
-\subsection{$\mlkemplus$} (\textcolor{blue}{In general, you cannot have only one subsection.}
+First, ML-KEM constructs a PKE (called K-PKE, see Figure \ref{fig:k-pke-routines} in Appendix) whose IND-CPA security reduces to the conjectured intractability of the decisional Module Learning with Error (MLWE) problem \cite{DBLP:conf/stoc/Regev05,DBLP:conf/eurocrypt/LyubashevskyPR10,DBLP:journals/jacm/Regev09,DBLP:conf/eurocrypt/LyubashevskyPR10,DBLP:conf/stoc/Peikert09}.
+Then, the $\kem^{\not\bot}_m$ variant (see Figure \ref{fig:u-notbot-m-routines}) of the Fujisaki-Okamoto transformation is applied to convert the IND-CPA secure PKE into an IND-CCA secure KEM (See Figure \ref{fig:ml-kem-routines} in Appendix).
+
+\subsection{ML-KEM+}
 
@@
                 \Else
-                    \State $K \leftarrow \texttt{KDF}(z \Vert t)$
+                    \State $K \leftarrow \texttt{KDF}(z \Vert c)$
                 \EndIf
@@
 \end{figure}
+
+For performance and practical security, $\mlkemplus$ makes two main modifications to the encrypt-then-MAC transformation: \begin{itemize}
+    \item \textbf{Hashing $\pk$ into MAC key and shared secret.} The public key goes into deriving both the MAC key and the shared secret for two main reasons.
+First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties.
+Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret.
+The second reason is to prevent multitarget attacks described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18}: if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
+Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption.
+Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
+    \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext.
+However, we find it unnecessary to hash the entire ciphertext.
+Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements for shared secret derivation.
+On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
+\end{itemize}
 
@@
 For implementation, we instantiated the MAC with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
-
-% All chosen MACs are parameterized with 256-bit keys and 128-bit tags.
-We chose 256-bit keys for all three security levels (128-bit, 196-bit, and 256-bit security) because choosing a smaller key size does not yield meaningful performance gains (in addition, Poly1305 only supports 256-bit keys).
-On the other hand, a 128-bit tag size is sufficient for all three security levels because unlike brute-force attacks on the MAC key, which can be carried out offline with large computing capacity, attacks on the tag require interaction with an online peer that plays the role of a MAC verification oracle.
-This means that the attack bandwidth is limited by the throughput of the online oracle instead of the attacker's computing power, and in practice it is easy to implement rate limiting in the cryptographic protocol such that the adversary cannot submit large quantities of MAC verification queries.
-
-\subsection{Poly1305 and GMAC with 128-bit tags}
-Poly1305 and GMAC are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic.
+All MACs are parameterized with a 256-bit key to ensure that MAC keys are at least as hard to guess as the underlying PKE plaintext.
+On the other hand, all MACs use 128-bit tag except for KMAC, which can have variable tag length.
+We suspect that 128-bit tag might be insufficient for the 192-bit and 256-bit security level, so for instantiation with the higher security levels, we only use KMAC with 192-bit and 256-bit tag length output.
+
+\textbf{Poly1305 and GMAC} are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic.
 Generically speaking, Carter-Wegman MAC operates by breaking the message into message blocks that can then be parsed into finite field elements.
 The tag is computed by evaluating a polynomial whose coefficients are the message blocks and whose interdeterminate is the secret key, which is also called a \textit{hash key}.
-
 Specifically, Poly1035 operates in the prime field $\mathbb{F}_q$ where $q = 2^{130} - 5$ whereas GMAC operates in the binary field $\mathbb{F}_{2^{128}}$.
-In OpenSSL's implementation, standalone Poly1305 is a one-time secure MAC, whereas GMAC uses a nonce and AES as the PRF and is thus many-time secure (in OpenSSL GMAC is AES-256-GCM except all data is fed into the ``associated data'' section and thus not encrypted).
-
-\textcolor{blue}{\textbf{ I have revised this part. Please consider to get the data, and revise Table 4, i.e., put back the original 192 and 256 cases without Poly1305. }
-}
-
-In order to get the tag sizes of 192-bit and 256-bit respectively. 
-We can extend GMAC, CMAC, and KMAC to get those tags.
-
-\subsection{Constructions of GMAC for 192-bit and 256-bit tags}
-
-For 192-bit and 256-bit tag size for GMAC, the hash key is given by $h=r_0 || r'$ where $r_0=AES_{256}(k, 0)$  and $r'$ is the most significant 64 bits from $r_1=AES_{256}(k, 1)$ for 192-bit tag and $h=r_0||r_1$ for 256-bit tag.
-Now the cipher polynomial, (i.e., it parses the ciphertext  as 192-bit as  blocks or 256-bit as  blocks)  is  evaluated in $\mathbb{F}_{2^{192}}$ and $\mathbb{F}_{2^{256}}$, respectively.
-
-\subsection{CMAC and an extension to get larger tags}
-CMAC is based on the CBC-MAC with the block cipher instantiated from AES-256.
+
+\textbf{CMAC} is based on the CBC-MAC with the block cipher instantiated from AES-256.
 To compute a CMAC tag, the message is first broke into 128-bit blocks with appropriate padding.
@@
 
-For the other two cases, it runs AES one more time to get another block of 128 bit.
-Using  the same method in the case for getting  192-bit and 256-bit hash keys in GMAC,  we get the tag with  192-bit and 256-bit respectively.
-
-\subsection{KMAC}
-KMAC is based on the SHA-3 family of sponge functions \cite{FIPS202}.
+\textbf{KMAC} is based on the SHA-3 family of sponge functions \cite{FIPS202}.
 We chose KMAC-256, which uses Shake256 as the underlying extendable output functions.
-KMAC allows variable-length key and tag, but we chose the 256 bits for key length, since the output of Shake256 can output 256 bits, so, we can get tag with size 128-bit, 192-bit and 256-bit, respectively.
-
-\emph{Remark.} Unfortunately Poly1305 uses tag size of 128 bits with no straightforward ways to increase the tag size.
-We suspect this small tag size to be unsuitable for instantiation with ML-KEM-768 and ML-KEM-1024 due to their higher security target (192-bit and 256-bit security respectively).
-Consequently, we only instantiate  with ML-KEM-512 with Poly1305, with a target 128-bit security level, instead of the other cases.
+KMAC is the only MAC to support variable key and tag length, though we fixed the key length at 256 bits.
+On the other hand, we chose the appropriate tag length for the corresponding security level: when instantiated with ML-KEM-512, the tag length is 128 bits; with ML-KEM-768 the tag length is 192 bits; with ML-KEM-1024 the tag length is 256 bits.
 
@@
     \centering
-    % \footnotesize
-    \small
+    \footnotesize
+    % \small
     \caption{CPU cycles needed to compute tag on various input sizes}\label{tbl:mac-performance}
-    \begin{tabular}{|c|c|c|}
-        \hline
-        \multicolumn{3}{|c|}{\bf Input size: 768 bytes} \\
-        \hline
-        MAC & Median (ccl/tick) & Average (ccl/tick)\\
-        \hline
-        Poly1305 & 909 & 2823 \\
-        \hline
-        GMAC & 3899 & 4859 \\
-        \hline
-        CMAC & 6291 & 6373 \\
-        \hline
-        KMAC & 6373 & 7791 \\
+    \begin{tabular}[t]{|p{1.45cm}|p{1.0cm}|p{1.0cm}|}
+        \hline
+        \multicolumn{2}{|c|}{\bf Input size: 768 bytes} \\
+        \hline
+        MAC & Median \newline (ccl)\\
+        \hline
+        Poly1305 & 909 \\
+        \hline
+        GMAC & 3899 \\
+        \hline
+        CMAC & 6291 \\
+        \hline
+        KMAC \newline 128-bit tag & 6373 \\
         \hline
@@
     % TODO: 128-bit tag size is not suitable for higher security level.
-    % \begin{tabular}{|c|c|c|}
-    %     \hline
-    %     \multicolumn{3}{|c|}{\bf Input size: 1088 bytes} \\
-    %     \hline
-    %     MAC & Median & Average \\
+    \begin{tabular}[t]{|p{1.45cm}|p{1.0cm}|p{1.0cm}|}
+        \hline
+        \multicolumn{2}{|c|}{\bf Input size: 1088 bytes} \\
+        \hline
+        MAC & Median \newline (ccl) \\
         % \hline
@@
         % CMAC & 7305 & 7588 \\
-    %     \hline
-    %     KMAC & 9697 & 9928 \\
-    %     \hline
-    % \end{tabular}
-    % \begin{tabular}{|c|c|c|}
-    %     \hline
-    %     \multicolumn{3}{|c|}{\bf Input size: 1568 bytes} \\
-    %     \hline
-    %     MAC & Median & Average \\
+        \hline
+        KMAC \newline 192-bit tag & 9697 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}[t]{|p{1.45cm}|p{1.0cm}|p{1.0cm}|}
+        \hline
+        \multicolumn{2}{|c|}{\bf Input size: 1568 bytes} \\
+        \hline
+        MAC & Median \newline (ccl) \\
         % \hline
@@
         % CMAC & 8735 & 8772 \\
-    %     \hline
-    %     KMAC & 11647 & 12186 \\
-    %     \hline
-    % \end{tabular}
+        \hline
+        KMAC \newline 256-bit tag & 11647 \\
+        \hline
+    \end{tabular}
 \end{table}
@@
     % TODO: 128-bit tag size is unsuitable for higher security level
-    % \begin{tabular}[t]{|cc|}
-    %     \hline
-    %     \multicolumn{2}{|c|}{\bf 192-bit security} \\
-    %     \hline
-    %     \multicolumn{2}{|c|}{size parameters (bytes)} \\
-    %     $\pk$ size & 1184 \\
-    %     $\sk$ size & 2400 \\
-    %     \text{ct} size & 1088 \\
-    %     \hline
-    %     \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
-    %     Median & 129895 \\
-    %     Average & 130650 \\
-    %     \hline
-    % \end{tabular}
-    % \begin{tabular}[t]{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant} 
-    %     & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
-    %     & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
-    %     \cline{2-5}
-    %     & Median & Average & Median & Average \\
-    %     \hline
-    %     \text{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
-    %     \hline
-    %     \text{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
-    %     \hline
+    \begin{tabular}[t]{|cc|}
+        \hline
+        \multicolumn{2}{|c|}{\bf 192-bit security} \\
+        \hline
+        \multicolumn{2}{|c|}{size parameters (bytes)} \\
+        $\pk$ size & 1184 \\
+        $\sk$ size & 2400 \\
+        \text{ct} size & 1088 \\
+        \hline
+        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+        Median & 129895 \\
+        Average & 130650 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}[t]{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant} 
+        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+        \cline{2-5}
+        & Median & Average & Median & Average \\
+        \hline
+        \text{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
+        \hline
+        \text{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
+        \hline
         % $\text{ML-KEM}^+$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
@@
         % \hline
-    %     $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
-    %     \hline
-    % \end{tabular}\vspace{0.3cm}
-
-    % \begin{tabular}[t]{|cc|}
-    %     \hline
-    %     \multicolumn{2}{|c|}{\bf 256-bit security} \\
-    %     \hline
-    %     \multicolumn{2}{|c|}{size parameters (bytes)} \\
-    %     $\pk$ size & 1568 \\
-    %     $\sk$ size & 3168 \\
-    %     \text{ct} size & 1568 \\
-    %     \hline
-    %     \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
-    %     Median & 194921 \\
-    %     Average & 195465 \\
-    %     \hline
-    % \end{tabular}
-    % \begin{tabular}[t]{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant} 
-    %     & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
-    %     & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
-    %     \cline{2-5}
-    %     & Median & Average & Median & Average \\
-    %     \hline
-    %     \text{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
-    %     \hline
-    %     \text{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
-    %     \hline
+        $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+
+    \begin{tabular}[t]{|cc|}
+        \hline
+        \multicolumn{2}{|c|}{\bf 256-bit security} \\
+        \hline
+        \multicolumn{2}{|c|}{size parameters (bytes)} \\
+        $\pk$ size & 1568 \\
+        $\sk$ size & 3168 \\
+        \text{ct} size & 1568 \\
+        \hline
+        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
+        Median & 194921 \\
+        Average & 195465 \\
+        \hline
+    \end{tabular}
+    \begin{tabular}[t]{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant} 
+        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
+        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
+        \cline{2-5}
+        & Median & Average & Median & Average \\
+        \hline
+        \text{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
+        \hline
+        \text{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
+        \hline
         % $\text{ML-KEM}^+$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
@@
         % \hline
-    %     $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
-    %     \hline
-    % \end{tabular}
+        $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
+        \hline
+    \end{tabular}
 \end{table}
@@
 
-\emph{Remark.} Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
+% \emph{Remark.} Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes.
 KEMTLS \cite{DBLP:conf/ccs/SchwabeSW20} was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage.
@@
     % TODO: 128-bit tag size is unsuitable for higher security level
-    % \begin{tabular}{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant}
-    %     & \multirow{2}{*}{Client TX bytes}
-    %     & \multirow{2}{*}{Server TX bytes}
-    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-    %     \cline{4-5}
-    %     & & & Median & Average \\
-    %     \hline
-    %     \texttt{ML-KEM-768} & 1184 & 1088 & 135 & 140 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 1184 & 1104 & 99 & 104 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ GMAC & 1184 & 1104 & 101 & 105 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ CMAC & 1184 & 1104 & 103 & 109 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ KMAC & 1184 & 1104 & 103 & 107 \\
-    %     \hline
-    % \end{tabular}\vspace{0.3cm}
-    % \begin{tabular}{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant}
-    %     & \multirow{2}{*}{Client TX bytes}
-    %     & \multirow{2}{*}{Server TX bytes}
-    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-    %     \cline{4-5}
-    %     & & & Median & Average \\
-    %     \hline
-    %     \texttt{ML-KEM-1024} & 1568 & 1568 & 193 & 199 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 1568 & 1584 & 138 & 141 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ GMAC & 1568 & 1584 & 140 & 145 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ CMAC & 1568 & 1584 & 143 & 148 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ KMAC & 1568 & 1584 & 144 & 149 \\
-    %     \hline
-    % \end{tabular}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        ML-KEM-768 & 1184 & 1088 & 135 & 140 \\
+        \hline
+        % $\mlkemplus$-768 w/ Poly1305 & 1184 & 1120 & 99 & 104 \\
+        % \hline
+        % $\mlkemplus$-768 w/ GMAC & 1184 & 1120 & 101 & 105 \\
+        % \hline
+        % $\mlkemplus$-768 w/ CMAC & 1184 & 1120 & 103 & 109 \\
+        % \hline
+        $\mlkemplus$-768 w/ KMAC & 1184 & 1120 & 103 & 107 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        ML-KEM-1024 & 1568 & 1568 & 193 & 199 \\
+        \hline
+        % $\mlkemplus$-1024 w/ Poly1305 & 1568 & 1600 & 138 & 141 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ GMAC & 1568 & 1600 & 140 & 145 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ CMAC & 1568 & 1600 & 143 & 148 \\
+        % \hline
+        $\mlkemplus$-1024 w/ KMAC & 1568 & 1600 & 144 & 149 \\
+        \hline
+    \end{tabular}
 \end{table}
@@
 UAKE routines are summarized in Figure \ref{fig:uake-routines}.
+Performance data is summaried in Table \ref{tbl:uake-rtt}.
 
@@
 
-In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key.
+% In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key.
 This helps the key exchange to achieve weak forward secrecy \cite{DBLP:conf/eurocrypt/CanettiK01}.
-Performance data is summaried in Table \ref{tbl:uake-rtt}.
 
@@
     % TODO: 128-bit tag size is unsuitable for higher security level
-    % \begin{tabular}{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant}
-    %     & \multirow{2}{*}{Client TX bytes}
-    %     & \multirow{2}{*}{Server TX bytes}
-    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-    %     \cline{4-5}
-    %     & & & Median & Average \\
-    %     \hline
-    %     \texttt{ML-KEM-768} & 2272 & 1088 & 215 & 222 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 1104 & 149 & 156 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 1104 & 153 & 160 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 1104 & 154 & 159 \\
-    %     \hline
-    % \end{tabular}\vspace{0.3cm}
-    % \begin{tabular}{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant}
-    %     & \multirow{2}{*}{Client TX bytes}
-    %     & \multirow{2}{*}{Server TX bytes}
-    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-    %     \cline{4-5}
-    %     & & & Median & Average \\
-    %     \hline
-    %     \texttt{ML-KEM-1024} & 3136 & 1568 & 310 & 318 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 1584 & 212 & 228 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 1584 & 212 & 218 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 1584 & 213 & 220 \\
-    %     \hline
-    % \end{tabular}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        ML-KEM-768 & 2272 & 1088 & 215 & 222 \\
+        \hline
+        % $\mlkemplus$-768 w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
+        % \hline
+        % $\mlkemplus$-768 w/ GMAC & 2288 & 1104 & 149 & 156 \\
+        % \hline
+        % $\mlkemplus$-768 w/ CMAC & 2288 & 1104 & 153 & 160 \\
+        % \hline
+        $\mlkemplus$-768 w/ KMAC & 2288 & 1104 & 154 & 159 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        ML-KEM-1024 & 3136 & 1568 & 310 & 318 \\
+        \hline
+        % $\mlkemplus$-1024 w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ GMAC & 3152 & 1584 & 212 & 228 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ CMAC & 3152 & 1584 & 212 & 218 \\
+        % \hline
+        $\mlkemplus$-1024 w/ KMAC & 3152 & 1584 & 213 & 220 \\
+        \hline
+    \end{tabular}
 \end{table}
@@
     % TODO: 128-bit tag size is unsuitable for higher security level
-    % \begin{tabular}{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant}
-    %     & \multirow{2}{*}{Client TX bytes}
-    %     & \multirow{2}{*}{Server TX bytes}
-    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-    %     \cline{4-5}
-    %     & & & Median & Average \\
-    %     \hline
-    %     \texttt{ML-KEM-768} & 2272 & 2176 & 294 & 301 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 2208 & 197 & 210 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 2208 & 202 & 208 \\
-    %     \hline
-    %     $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 2208 & 204 & 210 \\
-    %     \hline
-    % \end{tabular}\vspace{0.3cm}
-    % \begin{tabular}{|c|c|c|c|c|}
-    %     \hline
-    %     \multirow{2}{*}{KEM variant}
-    %     & \multirow{2}{*}{Client TX bytes}
-    %     & \multirow{2}{*}{Server TX bytes}
-    %     & \multicolumn{2}{|c|}{RTT ($\us$)} \\
-    %     \cline{4-5}
-    %     & & & Median & Average \\
-    %     \hline
-    %     \texttt{ML-KEM-1024} & 3136 & 3136 & 512 & 511 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 3168 & 273 & 282 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 3168 & 280 & 287 \\
-    %     \hline
-    %     $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 3168 & 282 & 288 \\
-    %     \hline
-    % \end{tabular}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        ML-KEM-768 & 2272 & 2176 & 294 & 301 \\
+        \hline
+        % $\mlkemplus$-768 w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
+        % \hline
+        % $\mlkemplus$-768 w/ GMAC & 2288 & 2208 & 197 & 210 \\
+        % \hline
+        % $\mlkemplus$-768 w/ CMAC & 2288 & 2208 & 202 & 208 \\
+        % \hline
+        $\mlkemplus$-768 w/ KMAC & 2288 & 2208 & 204 & 210 \\
+        \hline
+    \end{tabular}\vspace{0.3cm}
+    \begin{tabular}{|c|c|c|c|c|}
+        \hline
+        \multirow{2}{*}{KEM variant}
+        & \multirow{2}{*}{Client TX bytes}
+        & \multirow{2}{*}{Server TX bytes}
+        & \multicolumn{2}{|c|}{RTT ($\us$)} \\
+        \cline{4-5}
+        & & & Median & Average \\
+        \hline
+        ML-KEM-1024 & 3136 & 3136 & 512 & 511 \\
+        \hline
+        % $\mlkemplus$-1024 w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ GMAC & 3152 & 3168 & 273 & 282 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ CMAC & 3152 & 3168 & 280 & 287 \\
+        % \hline
+        $\mlkemplus$-1024 w/ KMAC & 3152 & 3168 & 282 & 288 \\
+        \hline
+    \end{tabular}
 \end{table}
@@
 While SIKE and SIDH were found to be insecure \cite{DBLP:conf/eurocrypt/CastryckD23}, other isogeny-based cryptosystem such as CSIDH \cite{DBLP:conf/asiacrypt/Castryck0MPR18} remains unaffected by the aforementioned attack and might be suitable candidates.
+
+\section{Appendix}\label{sec:appendix}
+Below are high-level summaries of the PKE sub-routines and the KEM routines of ML-KEM:
+
+\begin{figure}[h]
+    \centering
+
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.KeyGen()}}
+            \begin{algorithmic}[1]
+                \State $A \leftsample R_q^{k \times k}$
+                \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
+                \State $\pk \leftarrow (A, \mathbf{t})$
+                \State $\sk \leftarrow \mathbf{s}$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.33\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.Enc(pk, m)}}
+            \begin{algorithmic}[1]
+                \Ensure $\pk = (A, \mathbf{t})$
+                \Ensure $m \in R_2$
+                \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
+                \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
+                \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
+                \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
+                \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
+                \State \Return $(\mathbf{c}_1, c_2)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{0.3\textwidth}
+        \begin{algorithm}[H]
+            \caption*{\texttt{K-PKE.Dec(sk, c)}}
+            \begin{algorithmic}[1]
+                \Ensure $c = (\mathbf{c}_1, c_2)$
+                \Ensure $\sk = \mathbf{s}$
+                \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
+                \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
+                \State \Return $\hat{m}$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{K-PKE is an IND-CPA secure PKE based on the conjectured intractability of the Module Learning with Error (MWLE) problem}\label{fig:k-pke-routines}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM.KeyGen()}$}
+            \begin{algorithmic}[1]
+                \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE.KeyGen()}$
+                \State $h \leftarrow \texttt{SHA3-256}(\pk)$
+                \State $z \leftsample \mathcal{M}$
+                \State $\sk \leftarrow (\sk^\prime, \pk, h, z)$
+                \State \Return $(\pk, \sk)$
+            \end{algorithmic}
+        \end{algorithm}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM.Encap}(\pk)$}
+            \begin{algorithmic}[1]
+                \State $m \leftsample \mathcal{M}$
+                \State $h \leftarrow \texttt{SHA3-256}(\pk)$
+                \State $(K, r) \leftarrow \texttt{SHA3-512}(m, h)$
+                \State $c \leftarrow \texttt{K-PKE.Enc}(\pk, m ,r)$
+                \State \Return $(c, K)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{$\texttt{ML-KEM.Decap}(\sk, c)$}
+            \begin{algorithmic}[1]
+                \State $\hat{m} \leftarrow \texttt{K-PKE.Dec}(\sk^\prime, c)$
+                \State $(\overline{K}, \hat{r}) \leftarrow \texttt{SHA3-512}(\hat{m}, h)$
+                \State $\hat{c} \leftarrow \texttt{K-PKE.Enc}(\pk, m, \hat{r})$
+                \If{$\hat{c} = c$}
+                    \State $K \leftarrow \overline{K}$
+                \Else 
+                    \State $K \leftarrow \texttt{SHA3-256}(c, z)$
+                \EndIf
+                \State \Return $K$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+
+    \caption{ML-KEM uses the $U^{\not\bot}_m$ variant of the modular Fujisaki-Okamoto KEM transformation}\label{fig:ml-kem-routines}
+\end{figure}
 

### 2024-10-15 etm-transform/manuscript/main.tex
Final revision before submission

@@
 % - Add "spthm" for LNCS-like theorems
-
-% No indentation
-% \setlength{\parindent}{0px}
 
@@
 
+\newcommand{\red}[1]{\textcolor{red}{#1}}
+
 %%%% 4. TITLE %%%%
@@
 The Fujisaki-Okamoto transformation uses de-randomization and re-encryption to ensure ciphertext non-malleability, but because ML-KEM's underlying PKE encryption is significantly slower than the underlying PKE decryption, ML-KEM's decapsulation performance is dominated by re-encryption.
-In this paper, we propose an alternative generic IND-CCA secure KEM transformation, called ${\sf KEM}_{{\tt EtM}}$ that applies the encrypt-then-MAC mechanism to an OW-PCA secure PKE and an existentially unforgeable MAC.
+In this paper, we propose an alternative generic IND-CCA secure KEM transformation that achieves chosen-ciphertext security by applying the encrypt-then-MAC mechanism to a OW-PCA secure PKE and an existentially unforgeable MAC.
 Compared to the Fujisaki-Okamoto transformation, our encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
-    We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it $\mlkemplus$.
-We then implement $\mlkemplus$ with a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
-At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), $\mlkemplus$ achieves a massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
-Furthermore, we implement key exchange protocols and measure realistic network round trip times (RTTs), where $\mlkemplus$ reduces RTTs by 23.9\%-39.5\% compared to ML-KEM.
+We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it ML-KEM+.
+We then implement ML-KEM+ with a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
+At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), ML-KEM+ achieves a massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
+Furthermore, we implement key exchange protocols and measure realistic network round trip times (RTTs), where $\mlkemplus$ reduces RTTs by 23.9\%-39.5\%,  compared to ML-KEM.
 \end{abstract}
@@
 
+% TODO: this seems rather unnecessary
+% \red{To our knowledge, there is no work focused on constructing KEMs using the encrypt-then-MAC mechanism.}
+
 \subsection{Contributions}
 Our main contributions are threefold: \begin{itemize}
+    \setlength\itemsep{.1em}
     \item {\bf New IND-CCA secure KEM construction using encrypt-then-MAC.} We propose a generic IND-CCA secure KEM transformation called the encrypt-then-MAC transformation and prove that the IND-CCA security of the encrypt-then-MAC transformation reduces tightly to the OW-PCA security of the underlying PKE and the existential unforgeability of the underlying MAC.
@@
     \item {\bf $\mlkemplus$: Efficient CCA-secure KEM based on ML-KEM.} We present $\mlkemplus$, an IND-CCA secure KEM constructed by applying the encrypt-then-MAC transformation to ML-KEM's underlying PKE sub-routines.
-Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing a MAC tag, which yields substantial performance improvements.
-    \item {\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C and compared its performance with ML-KEM in a variety of scenarios.
+Compared to ML-KEM, $\mlkemplus$ adds a small amount of performance penalty to the encapsulation routine and a small increase in ciphertext size, but replaces the expensive re-encryption step in decapsulation with computing and verifying a MAC tag, which yields substantial performance improvements.
+    \item {\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C by adapting ML-KEM's reference implementation \cite{kyberrefimpl} and compared its performance with ML-KEM in a variety of scenarios.
 Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
@@
 
-    \caption{$\mlkemplus$ achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
+    \caption{$\mlkemplus$ achieves impressive reduction of decapsulation CPU cycles (ccl) while incurring minimal penalty in encapsulation performance and ciphertext size.}\label{tbl:cpu-cycles-summary}
 \end{table}
@@
 
-    \caption{$\mlkemplus$ achieves substantial RTT savings despite increased encapsulation cost and ciphertext size}\label{tbl:rtt-summary}
+    \caption{$\mlkemplus$ achieves substantial RTT savings in key exchange protocols despite increased encapsulation cost and ciphertext size.
+RTT is measured in microseconds $\us$.}\label{tbl:rtt-summary}
 \end{table}
@@
 It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
-The $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.
+The $\kem^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.
 
@@
 Because it uses re-encryption for achieving ciphertext non-malleability, the Fujisaki-Okamoto transformation suffers from the following two problems: \begin{itemize}
+    \setlength\itemsep{.1em}
     \item \textbf{Computational inefficiency}.
@@
 
+% \section{Encrypt-then-MAC: new generic IND-CCA KEM transformation}\label{sec:main-results}
 \section{The encrypt-then-MAC transformation}\label{sec:main-results}
@@
     \begin{tikzpicture}
-        \tikzstyle{rect} = [draw, rectangle, minimum width=3cm, minimum height=1cm]
-        \tikzstyle{filledrect} = [draw, rectangle, minimum width=3cm, minimum height=1cm, fill=lightgray]
+        \tikzstyle{rect} = [draw, rectangle, minimum width=3cm, minimum height=0.5cm]
+        \tikzstyle{filledrect} = [draw, rectangle, minimum width=3cm, minimum height=0.5cm, fill=lightgray]
         \matrix [column sep=7mm, row sep=5mm] {
@@
 \end{theorem}
-
-Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} naturally flows into an equivalence relationship between the security of the KEM and the security of the PKE:
-
-\begin{lemma}
-    $\kem_\etm$ is IND-CCA secure if and only if the input $\pke$ is OW-PCA secure
-\end{lemma}
 
@@
     \end{minipage}
-    \caption{Sequence of games}\label{fig:etm-ind-cc2-sequence-of-games}
+    \caption{Sequence of games in the proof of Theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{fig:etm-ind-cc2-sequence-of-games}
 \end{figure}
-
-\begin{proof}
-    \emph{Game 0} is the standard IND-CCA game for KEMs.
-The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
-The queries made to the hash oracles $\mathcal{O}^G, \mathcal{O}^H$ are recorded to their respective tapes $\mathcal{L}^G, \mathcal{L}^H$.
-
-    \emph{Game 1} is identical to game 0 except that the true decapsulation oracle $\mathcal{O}^\decap$ is replaced with a simulated oracle $\mathcal{O}^\decap_1$.
-Instead of directly decrypting $c^\prime$ as in the decapsulation routine, the simulated oracle searches through the tape $\mathcal{L}^G$ to find a matching query $(\tilde{m}, \tilde{k})$ such that $\tilde{m}$ is the decryption of $c^\prime$.
-The simulated oracle then uses $\tilde{k}$ to validate the tag $t$ against $c^\prime$.
-
-    If the simulated oracle accepts the queried ciphertext as valid, then there is a matching query that also validates the tag, which means that the queried ciphertext is honestly generated.
-Therefore, the true oracle must also accept the queried ciphertext.
-On the other hand, if the true oracle rejects the queried ciphertext, then the tag is simply invalid under the MAC key $k = G(\decrypt(\sk^\prime, c^\prime))$.
-Therefore, there could not have been a matching query that also validates the tag, and the simulated oracle must also rejects the queried ciphertext.
-
-    This means that from the adversary $A$'s perspective, game 1 and game 0 differ only when the true oracle accepts while the simulated oracle rejects, which means that $t$ is a valid tag for $c^\prime$ under $k = G(\decrypt(\sk^\prime, c^\prime))$, but $k$ has never been queried.
-Under the random oracle model, such $k$ is a uniformly random sample of $\mathcal{K}_\mac$ that the adversary does not know, so for $A$ to produce a valid tag is to produce a forgery against the $\mac$ under an unknown and uniformly random key.
-Furthermore, the security game does not include a signing oracle, so this is a zero-time forgery.
-While zero-time forgery is not a standard security definition for a MAC, we can bound it by the advantage of a one-time forgery adversary $C$:
-
-    \begin{equation*}
-        P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]
-        \leq \adv_\texttt{OT-MAC}(C)
-    \end{equation*}
-
-    Across all $q$ decapsulation queries, the probability that at least one query is a forgery is thus at most $q \cdot P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]$.
-By the difference lemma:
-
-    \begin{equation*}
-        \adv_{G_0}(A) - \adv_{G_1}(A) \leq q\cdot  \adv_\texttt{OT-MAC}(C)
-    \end{equation*}
-
-    \emph{Game 2} is identical to game 1, except that the challenger samples a uniformly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$ instead of deriving it from $m^\ast$.
-From $A$'s perspective the two games are indistinguishable, unless $A$ queries $G$ with the value of $m^\ast$.
-Denote the probability that $A$ queries $G$ with $m^\ast$ by $P[\texttt{QUERY G}]$, then:
-
-    \begin{equation*}
-        \adv_{G_1}(A) - \adv_{G_2}(A) \leq P\left[\texttt{QUERY G}\right]
-    \end{equation*}
-
-    \emph{Game 3} is identical to game 2, except that the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $t$.
-From $A$'s perspective the two games are indistinguishable, unless $A$ queries $H$ with $(m^\ast, \cdot)$.
-Denote the probability that $A$ queries $H$ with $(m^\ast, \cdot)$ by $P[\texttt{QUERY H}]$, then:
-
-    \begin{equation*}
-        \adv_{G_2}(A) - \adv_{G_3}(A) \leq P\left[\texttt{QUERY H}\right]
-    \end{equation*}
-
-    Since in game 3, both $K_0$ and $K_1$ are uniformly random and independent of all other variables, no adversary can have any advantage: $\adv_{G_3}(A) = 0$.
-
 \begin{figure}[h]
@@
             \end{algorithmic}
-            \end{algorithm}
+        \end{algorithm}\vspace{-0.5cm}
         \begin{algorithm}[H]
@@
     \end{minipage}
-        \caption{OW-PCA adversary $B$ simulates game 3 for IND-CCA adversary $A$}\label{fig:ow-pca-simulates-game-3}
+    \caption{OW-PCA adversary $B$ simulates game 3 for IND-CCA adversary $A$ in the proof for Theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{fig:ow-pca-simulates-game-3}
 \end{figure}
+
+\begin{proof}
+    \emph{Game 0} is the standard IND-CCA game for KEMs.
+The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
+The queries made to the hash oracles $\mathcal{O}^G, \mathcal{O}^H$ are recorded to their respective tapes $\mathcal{L}^G, \mathcal{L}^H$.
+
+    \emph{Game 1} is identical to game 0 except that the true decapsulation oracle $\mathcal{O}^\decap$ is replaced with a simulated oracle $\mathcal{O}^\decap_1$.
+Instead of directly decrypting $c^\prime$ as in the decapsulation routine, the simulated oracle searches through the tape $\mathcal{L}^G$ to find a matching query $(\tilde{m}, \tilde{k})$ such that $\tilde{m}$ is the decryption of $c^\prime$.
+The simulated oracle then uses $\tilde{k}$ to validate the tag $t$ against $c^\prime$.
+
+    If the simulated oracle accepts the queried ciphertext as valid, then there is a matching query that also validates the tag, which means that the queried ciphertext is honestly generated.
+Therefore, the true oracle must also accept the queried ciphertext.
+On the other hand, if the true oracle rejects the queried ciphertext, then the tag is simply invalid under the MAC key $k = G(\decrypt(\sk^\prime, c^\prime))$.
+Therefore, there could not have been a matching query that also validates the tag, and the simulated oracle must also rejects the queried ciphertext.
+
+    This means that from the adversary $A$'s perspective, game 1 and game 0 differ only when the true oracle accepts while the simulated oracle rejects, which means that $t$ is a valid tag for $c^\prime$ under $k = G(\decrypt(\sk^\prime, c^\prime))$, but $k$ has never been queried.
+Under the random oracle model, such $k$ is a uniformly random sample of $\mathcal{K}_\mac$ that the adversary does not know, so for $A$ to produce a valid tag is to produce a forgery against the $\mac$ under an unknown and uniformly random key.
+Furthermore, the security game does not include a signing oracle, so this is a zero-time forgery.
+While zero-time forgery is not a standard security definition for a MAC, we can bound it by the advantage of a one-time forgery adversary $C$:
+
+    \begin{equation*}
+        P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]
+        \leq \adv_\texttt{OT-MAC}(C)
+    \end{equation*}
+
+    Across all $q$ decapsulation queries, the probability that at least one query is a forgery is thus at most $q \cdot P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]$.
+By the difference lemma:
+
+    \begin{equation*}
+        \adv_{G_0}(A) - \adv_{G_1}(A) \leq q\cdot  \adv_\texttt{OT-MAC}(C)
+    \end{equation*}
+
+    \emph{Game 2} is identical to game 1, except that the challenger samples a uniformly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$ instead of deriving it from $m^\ast$.
+From $A$'s perspective the two games are indistinguishable, unless $A$ queries $G$ with the value of $m^\ast$.
+Denote the probability that $A$ queries $G$ with $m^\ast$ by $P[\texttt{QUERY G}]$, then:
+
+    \begin{equation*}
+        \adv_{G_1}(A) - \adv_{G_2}(A) \leq P\left[\texttt{QUERY G}\right]
+    \end{equation*}
+
+    \emph{Game 3} is identical to game 2, except that the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $t$.
+From $A$'s perspective the two games are indistinguishable, unless $A$ queries $H$ with $(m^\ast, \cdot)$.
+Denote the probability that $A$ queries $H$ with $(m^\ast, \cdot)$ by $P[\texttt{QUERY H}]$, then:
+
+    \begin{equation*}
+        \adv_{G_2}(A) - \adv_{G_3}(A) \leq P\left[\texttt{QUERY H}\right]
+    \end{equation*}
+
+    Since in game 3, both $K_0$ and $K_1$ are uniformly random and independent of all other variables, no adversary can have any advantage: $\adv_{G_3}(A) = 0$.
 
@@
 
-It is also conjectured in \cite{DBLP:conf/ctrsa/AbdallaBR01} that for certain choice of cyclic group $G$, the computational Diffie-Hellman problem remains intractable even if the adversary as access to a restricted decisional Diffie-Hellman oracle.
+It is also conjectured in \cite{DBLP:conf/ctrsa/AbdallaBR01} (and later extensively studied in \cite{DBLP:conf/pkc/OkamotoP01}) that for certain choice of cyclic group $G$, the computational Diffie-Hellman problem remains intractable even if the adversary as access to a restricted decisional Diffie-Hellman oracle.
 This assumption is captured in the Gap Diffie-Hellman problem:
@@
     \emph{Game 2} is identical to game 1 except for two modifications: first, when computing the challenge ciphertext, $v$ is no longer computed from $m^\ast$ but is randomly sampled; second, the win condition changed from $\hat{m} = m^\ast$ to $\hat{m} = w^{-x}\cdot v$.
-Notice that in game 0-1, $v \leftsample m^\ast \cdot (g^x)^y$ follows uniform random distribution on the cyclic group $G$ because $m^\ast$ is uniformly random in the cyclic group, so adversary $A$ retains its advantage in ``recovering the decryption'' when $v$ becomes truly uniformly random in the cyclic group.
-It is easy to verify that the two win conditions are equivalent.
-Up to this point, we have simply moved things around, and game 0 through game 2 are algebraically equivalent:
+It is easy to verify that Game 0 through Game 2 are algebraically equivalent:
 
@@
     If $A$ wins game 2, then its output is $\hat{m} = w^{-x}\cdot v = g^{-xy}\cdot v$, so $m^{-1}\cdot v$ is $g^{xy}$, the correct answer to the Gap Diffie-Hellman problem.
-In other words, $B$ solves its Gap Diffie-Hellman problem if and only if $A$ wins the simulated game 2.
-
-    \begin{equation*}
-        \adv_2(A) = \adv_\text{GapDH}(B)
-    \end{equation*}
+In other words, $B$ solves its Gap Diffie-Hellman problem if and only if $A$ wins the simulated game 2: $\adv_2(A) = \adv_\text{GapDH}(B)$.
 \end{proof}
@@
 
-We apply the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM. The resulting KEM is called $\mlkemplus$ (See Figure \ref{fig:ml-kem-plus-routines}).
+We apply the encrypt-then-MAC transformation to the K-PKE routines of ML-KEM.
+The resulting KEM is called $\mlkemplus$ (See Figure \ref{fig:ml-kem-plus-routines}).
+The key generation routines are identical between $\mlkemplus$ and ML-KEM.
+The $\mlkemplus$  encapsulation and decapsulation algorithms make use of two additional algorithms, namely a {\sf MAC} and a {\sf KDF}.
 
@@
 
-For performance and practical security, $\mlkemplus$ makes two main modifications to the encrypt-then-MAC transformation: \begin{itemize}
-    \item \textbf{Hashing $\pk$ into MAC key and shared secret.} The public key goes into deriving both the MAC key and the shared secret for two main reasons.
-First, in a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties.
+For performance and practical security, $\mlkemplus$ makes two minor modifications to the encrypt-then-MAC transformation: \begin{itemize}
+    \setlength\itemsep{0.1em}
+    \item \textbf{Hashing $\pk$ into MAC key and shared secret.} The public key goes into deriving both the MAC key and the shared secret for the same two reasons why Kyber hashes its public key into the pseudorandom coin and the shared secret \cite{DBLP:conf/eurosp/BosDKLLSSSS18}.
+The first reason is to make the KEM contributory.
+In a realistic key exchange, the public key and the random PKE plaintext are generated by opposite parties.
 Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret.
-The second reason is to prevent multitarget attacks described in \cite{DBLP:conf/eurosp/BosDKLLSSSS18}: if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
+The second reason is to prevent multi-target attacks: if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
 Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption.
@@
 However, we find it unnecessary to hash the entire ciphertext.
-Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements for shared secret derivation.
-On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because asssuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag invalid, so the decapsulation oracle will always return the implicit rejection.
+Instead, since the MAC tag functions as a keyed hash of the ciphertext, we can use the much shorter MAC tag as a substitute ciphertext hash, which yields meaningful performance improvements.
+On the other hand, not hashing the entire ciphertext will not give an adversary additional advantage because assuming that the adversary does not already know the underlying decryption, it cannot tamper with any part of the challenge ciphertext without making the ciphertext-tag pair invalid, so the decapsulation oracle will always return the implicit rejection.
 \end{itemize}
@@
 Compared to ML-KEM, $\mlkemplus$ makes the following performance trade-offs: \begin{itemize}
+    \setlength\itemsep{0.1em}
     \item $\mlkemplus$ encapsulation contains two additional steps: computing the MAC tag and hashing the MAC tag into the shared secret.
     \item $\mlkemplus$ ciphertext size increases by the size of the MAC tag
-    \item $\mlkemplus$ decapsulation replaces re-encryption with computing the MAC tag
+    \item $\mlkemplus$ decapsulation replaces re-execution $\texttt{K-PKE}.\texttt{Enc}$ with computing the MAC tag
 \end{itemize}
@@
 \subsection{Choosing a message authentication code (MAC)}\label{sec:choosing-a-message-authenticator}
+In Figure \ref{fig:ml-kem-plus-routines}, while $H$, $G$ and \textsf{KDF} are instantiated by SHA3-256, SHA3-512 and Shake256, respectively, there could be various choices of standardized and well-analyze MAC algorithms.
 For implementation, we instantiated the MAC with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.
@@
 To isolate the performance characteristics of each authenticator in our instantiation of $\mlkemplus$ we measured the CPU cycles needed for each authenticator to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of ML-KEM.
-The measurements are summarized in Table \ref{tbl:mac-performance}.
+The median CPU cycle counts in 10,000 rounds of testing are summarized in Table \ref{tbl:mac-performance}.
 
@@
     \caption{CPU cycles needed to compute tag on various input sizes}\label{tbl:mac-performance}
-    \begin{tabular}[t]{|p{1.45cm}|p{1.0cm}|p{1.0cm}|}
-        \hline
-        \multicolumn{2}{|c|}{\bf Input size: 768 bytes} \\
-        \hline
-        MAC & Median \newline (ccl)\\
+    \begin{tabular}[t]{|c|c|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{128-bit security}} \\
+        \multicolumn{1}{|c}{\textbf{Tag size}}
+        & \multicolumn{1}{c|}{128 bits} \\
+        \multicolumn{1}{|c}{\textbf{CT size}}
+        & \multicolumn{1}{c|}{768 bytes} \\
+        \hline
+        \textbf{MAC} & \textbf{Median} \\
         \hline
@@
         \hline
-        KMAC \newline 128-bit tag & 6373 \\
+        KMAC & 6373 \\
         \hline
@@
     % TODO: 128-bit tag size is not suitable for higher security level.
-    \begin{tabular}[t]{|p{1.45cm}|p{1.0cm}|p{1.0cm}|}
-        \hline
-        \multicolumn{2}{|c|}{\bf Input size: 1088 bytes} \\
-        \hline
-        MAC & Median \newline (ccl) \\
+    \begin{tabular}[t]{|c|c|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{192-bit security}} \\
+        \multicolumn{1}{|c}{\textbf{Tag size}}
+        & \multicolumn{1}{c|}{192 bits} \\
+        \multicolumn{1}{|c}{\textbf{CT size}}
+        & \multicolumn{1}{c|}{1088 bytes} \\
+        \hline
+        \textbf{MAC} & \textbf{Median} \\
         % \hline
@@
         \hline
-        KMAC \newline 192-bit tag & 9697 \\
+        KMAC & 9697 \\
         \hline
     \end{tabular}
-    \begin{tabular}[t]{|p{1.45cm}|p{1.0cm}|p{1.0cm}|}
-        \hline
-        \multicolumn{2}{|c|}{\bf Input size: 1568 bytes} \\
-        \hline
-        MAC & Median \newline (ccl) \\
+    \begin{tabular}[t]{|c|c|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{256-bit security}} \\
+        \multicolumn{1}{|c}{\textbf{Tag size}}
+        & \multicolumn{1}{c|}{256 bits} \\
+        \multicolumn{1}{|c}{\textbf{CT size}}
+        & \multicolumn{1}{c|}{1568 bytes} \\
+        \hline
+        \textbf{MAC} & \textbf{Median} \\
         % \hline
@@
         \hline
-        KMAC \newline 256-bit tag & 11647 \\
+        KMAC & 11647 \\
         \hline
@@
 \subsection{Individual routine performance}
-Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the encrypt-then-MAC transformation achieves massive CPU cycle count reduction in decapsulation while incurring only a minimal increase of encapsulation cycle count and ciphertext size.
-Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the encrypt-then-MAC transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
+Compared to the $\kem^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the encrypt-then-MAC transformation achieves massive CPU cycle count reduction in decapsulation while incurring only a minimal increase of encapsulation cycle count and ciphertext size.
+Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the encrypt-then-MAC transformation over the $\kem^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
 A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in Table \ref{tbl:kem-performance}.
@@
 The key derivation function is instantiated using Shake256, and the final session key is 256 bits in length.
-The RTT comparison is summarized in Table \ref{tbl:ke-rtt}
+$\mlkemplus$-512 with Poly1305 reduces RTT by 23.9\% compared to ML-KEM-512; $\mlkemplus$-768 with KMAC reduces RTT by 23.7\%; $\mlkemplus$-1024 with KMAC reduces RTT by 25.4\% (See Table \ref{tbl:ke-rtt}).
 
@@
                 \State $K_e \leftarrow \decap(\sk_e, c_e)$
-                \State $K \leftarrow \texttt{KDF}(K)$
+                \State $K \leftarrow \texttt{KDF}(K_e)$
                 \State \Return $K$
@@
 UAKE routines are summarized in Figure \ref{fig:uake-routines}.
-Performance data is summaried in Table \ref{tbl:uake-rtt}.
+$\mlkemplus$-512 with Poly1305 reduces RTT by 29.0\% compared to ML-KEM-512; $\mlkemplus$-768 with KMAC reduces RTT by 28.4\%; $\mlkemplus$-1024 with KMAC reduces RTT by 31.3\% (See Table \ref{tbl:uake-rtt}).
 
@@
 The session key is derived by applying \texttt{KDF} onto the concatenation of shared secrets produced under the ephemeral keypair, server's long-term keypair, and client's long-term keypair, in this order.
-the key exchange routines are described in Figure \ref{fig:ake-routines} and the performance data summarized in Table \ref{tbl:ake-rtt}.
+The key exchange routines are described in Figure \ref{fig:ake-routines}.
+$\mlkemplus$-512 with Poly1305 reduces RTT by 39.5\% compared to ML-KEM-512; $\mlkemplus$-768 with KMAC reduces RTT by 30.6\%; $\mlkemplus$-1024 with KMAC reduces RTT by 44.9\% (See Table \ref{tbl:ake-rtt}).
 
@@
 \section{Conclusions and future works}\label{sec:future-works}
-The encrypt-then-MAC transformation is a generic KEM construction whose IND-CCA security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.
+In this paper we propose a generic KEM transformation based on the encrypt-then-MAC pattern.
+We reduce the IND-CCA security of the encrypt-then-MAC KEM to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.
+Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC transformation replaces the computationally expensive re-encryption with computing a MAC tag.
+At the cost of minimal increase in encapsulation cost and ciphertext size, the encrypt-then-MAC transformation substantially improves the efficiency of the decapsulation routine.
+Where the input PKE's encryption is slower than decryption, the encrypt-then-MAC transformation achieves meaningful time savings in practical key exchange protocols.
+We also propose $\mlkemplus$ as an efficient alternative to ML-KEM.
+Compared to ML-KEM, $\mlkemplus$ achieves great performance improvements in a variety of practical scenarios.
+
+% The encrypt-then-MAC transformation is a generic KEM construction whose IND-CCA security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.
 Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC replaces the computationally expensive re-encryption with computing a MAC tag.
@@
     \end{minipage}
+
     \caption{K-PKE is an IND-CPA secure PKE based on the conjectured intractability of the Module Learning with Error (MWLE) problem}\label{fig:k-pke-routines}
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \centering
@@
             \end{algorithmic}
-        \end{algorithm}
+        \end{algorithm} \vspace{-0.5cm}
         \begin{algorithm}[H]
@@
 
-    \caption{ML-KEM uses the $U^{\not\bot}_m$ variant of the modular Fujisaki-Okamoto KEM transformation}\label{fig:ml-kem-routines}
+    \caption{ML-KEM uses the $\kem^{\not\bot}_m$ variant of the modular Fujisaki-Okamoto KEM transformation}\label{fig:ml-kem-routines}
 \end{figure}

### 2024-10-19 archives/faster-ind-cca-kem-encrypt-then-mac/cca-elgamal/main.tex
"Faster IND-CCA KEM using encrypt-then-MAC" submitted to CHES on Oct 15

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}

% Custom commands
\usepackage{crypto-primitives}
\usepackage{kyber-algos}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    ElGamal cryptosystem
}
\author{
    Ganyu (Bruce) Xu
}
% Leave the date field empty to display the date of compilation
% \date{}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{The ElGamal cryptosystem}
The ElGamal cryptosystem is a public key encryption scheme that mainly operates on the discrete log problem. Each instance of the encryption scheme is parameterized by a cyclic group $G$ with prime order $q$, a generator $g$ of this cyclic group. The routines of the encryption scheme is shown in figure \ref{fig:cpa-elgamal-routines}

\begin{figure}[H]
    \begin{algorithm}[H]
        \caption{$\keygen$}
        \begin{algorithmic}[1]
            \State $x \leftsample \mathbb{Z}_q$
            \State $u \leftarrow g^x$
            \State $\pk \leftarrow u, \sk \leftarrow x$
            \State \Return $(\pk, \sk)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\encrypt(\pk = u, m \in G)$}
        \begin{algorithmic}[1]
            \State $y \leftsample \mathbb{Z}_q$
            \State $v \leftarrow g^y$
            \State $w \leftarrow u^y$
                \Comment{$w = g^{xy}$}
            \State $c \leftarrow (v, m \cdot w)$
            \State \Return $c$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\decrypt(\sk = x, c)$}
        \begin{algorithmic}[1]
            \State $(c_1, c_2) \leftarrow c$
            \State $\hat{w} \leftarrow c_1^x$
            \State $\hat{m} \leftarrow c_2 \cdot \hat{w}^{-1}$
            \State \Return $\hat{m}$
        \end{algorithmic}
    \end{algorithm}
    \caption{ElGamal encryption scheme is IND-CPA secure if DDH holds}\label{fig:cpa-elgamal-routines}
\end{figure}

The IND-CPA security of the ElGamal cryptosystem depends on the hardness of the following two problems:

\begin{definition}[Computational Diffie-Hellman Problem]
    Let $G$ be a cyclic group with prime order $q$ and generator $g$. Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $g, g^x,g^y$, compute $g^{xy}$
\end{definition}

\begin{definition}[Decisional Diffie-Hellman Problem]
    Let $G$ be a cyclic group with prime order $q$ and generator $g$. Let $x, y, z \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $g, g^x, g^y$, distinguish $g^{xy}$ from $g^z$
\end{definition}

\begin{theorem}
    For every IND-CPA adversary $A$ against the ElGamal cryptosystem, there exists an adversary $B$ against the DDH game such that

    \begin{equation*}
        \texttt{Adv}(A) = 2 \cdot \texttt{Adv}(B)
    \end{equation*}
\end{theorem}

% TODO: maybe write the proof?

\section{CCA-secure ElGamal construction}
The ElGamal cryptosystem presented in figure \ref{fig:cpa-elgamal-routines} is not secure against chosen-ciphertext attacks. For a simple attack, consider an adversary who just obtained the challenge encryption $c = (g^y, m \cdot (g^x)^y)$, where $y \leftsample \mathbb{Z}_q^\ast, m \leftsample G$ are sampled by the challenger, and $g^x$ is the public key. The adversary can pick some non-zero $y^\prime \in \mathbb{Z}_q^\ast$ and compute a distinct encryption of $m$ using the challenge ciphertext:

\begin{equation*}
    c^\prime = (g^y \cdot g^{y^\prime}, m \cdot (g^x)^y \cdot (g^x)^{y^\prime})
    = (g^{y + y^\prime}, m \cdot (g^x)^{y + y^\prime})
\end{equation*}

The adversary can then query the decryption oracle on $c^\prime$, and the decryption oracle will return $m$, which can then be returned to win the OW-CCA game.

\cite{boneh2020graduate} presented a hybrid encryption scheme that combined the ElGamal cryptosystem with an IND-CPA symmetric cipher $\mathcal{E} = (\encrypt, \decrypt)$ into a public-key encryption scheme. This CCA-secure ElGamal cryptosystem (which we will denote by HPKE for short) is parameterized by: \begin{enumerate}
    \item A cyclic group $G$ of prime order $q$ with generator $g$
    \item A symmetric cipher $\mathcal{E} = (\encrypt_s, \decrypt_s)$ defined over $(\mathcal{K}, \mathcal{M}, \mathcal{C})$
    \item A hash function $H: G \rightarrow \mathcal{K}$
\end{enumerate}

The routines are listed in figure \ref{fig:cca-elgamal-routines}

\begin{figure}[H]
    \begin{algorithm}[H]
        \caption{$\keygen$}
        \begin{algorithmic}[1]
            \State $x \leftarrow \mathbb{Z}_q$
            \State $u \leftarrow g^x$
            \State $\pk \leftarrow u$
            \State $\sk \leftarrow x$
            \State \Return $(\pk, \sk)$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\encrypt(\pk = u, m \in \mathcal{M})$}
        \begin{algorithmic}[1]
            \State $y \leftsample \mathbb{Z}_q$
            \State $v \leftarrow g^y$
            \State $w \leftarrow u^y$
                \Comment{$w = g^{xy}$}
            \State $k \leftarrow H(w)$
            \State $c^\prime \leftarrow \encrypt_S(k, m)$
            \State $c \leftarrow (v, c^\prime)$
            \State \Return $c$
        \end{algorithmic}
    \end{algorithm}
    \begin{algorithm}[H]
        \caption{$\decrypt(\sk = x, c)$}
        \begin{algorithmic}[1]
            \State $(v, c^\prime) \leftarrow c$
            \State $\hat{w} \leftarrow v^x$
            \State $\hat{k} \leftarrow H(\hat{w})$
            \State $\hat{m} \leftarrow \decrypt_S(\hat{k}, c^\prime)$
            \State \Return $\hat{m}$
        \end{algorithmic}
    \end{algorithm}
    \caption{ElGamal HPKE}\label{fig:cca-elgamal-routines}
\end{figure}

In this construction \emph{the decryption oracle can be used to construct a decisional Diffie-Hellman problem oracle}: \begin{enumerate}
    \item The DDH adversary $A$ receives $g^x, g^y, w$ and needs to decide whether $w$ is $g^{xy}$ or $g^z$
    \item $A$ samples a random message $m \leftarrow \mathcal{M}$ and computes $k \leftarrow H(w)$ and $c \leftarrow \encrypt_s(k, m)$
    \item $A$ queries the decryption oracle on $c$ and receives some ``decryption'' $\hat{m}$
    \item If $\hat{m} = m$, then $w$ is $g^{xy}$, otherwise $w$ is $g^z$. This is because if $w = g^{xy}$, then $k \leftarrow H(w)$ is the correct symmetric key, so the decryption oracle will decrypt correctly. On the other hand, if $w = g^z$, then $k \leftarrow H(w)$ is a uniformly random key, so the decryptino oracle will not decrypt correctly.
\end{enumerate}

Even though the decryption oracle allows an IND-CCA adversary to solve the decisional Diffie-Hellman problem, the hybrid construction remains CCA secure. This is because \emph{to our knowledge today, solving the decisional Diffie-Hellman problem does not give non-negligible advantage to solving the computational Diffie-Hellman problem}. This idea is expressed in a modified assumption:

\begin{definition}[Interactive computational Diffie-Hellman problem]
    Let $G$ be a cyclic group of prime order $q$ with generator $g$. Let $x, y, z \leftsample \mathbb{Z}_q^\ast$ be uniformly random samples. Given $g, g^x, g^y$ and a decisional Diffie-Hellman oracle $\mathcal{O}: (g^x, g^y, w \in \{g^{xy}, g^z\}) \mapsto \llbrack w = g^{xy} \rrbrack$, there is no efficient adversary who can compute $g^{xy}$ with non-negligible advantage.
\end{definition}

Finally we will put everything together into the security theorem for CCA ElGamal.

\begin{theorem}
    Under the random oracle model, for every IND-CCA adversary $A$ against the HPKE, there exists an \emph{interactive computational Diffie-Hellman problem} adversary $B$ and an IND-CPA adversary $C$ against the symmetric encryption scheme such that

    \begin{equation*}
        \texttt{Adv}_\texttt{IND-CCA}(A) \leq \texttt{Adv}_\texttt{ICDH}(B) + \texttt{Adv}_\texttt{IND-CPA}(C)
    \end{equation*}
\end{theorem}

\begin{proof}
    We will prove using a sequence of games. The games are listed in figure \ref{fig:cca-elgamal-sequence}

    \begin{figure}[H]
        \begin{algorithm}[H]
            \caption{Games $G_0 - G_1$}
            \begin{algorithmic}[1]
                \State $x \leftsample \mathbb{Z}_q^\ast$
                    \Comment{$x$ is the secret key}
                \State $u \leftarrow g^x$
                    \Comment{$u$ is the public key}
                \State $(m_0, m_1) \leftarrow A^{\mathcal{O}^\decrypt}(u)$
                \State $y \leftsample \mathbb{Z}_q^\ast$
                \State $v \leftarrow g^y$
                \State $w \leftarrow u^y$
                \State $k \leftarrow H(w)$
                    \Comment{Game 0}
                \State $k \leftsample \mathcal{K}$
                    \Comment{Game 1}
                \State $b \leftsample \{0,1\}$
                \State $c^\prime \leftarrow \encrypt_s(k, m_b)$
                \State $c^\ast \leftarrow (v, c^\prime)$
                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decrypt}(u, c^\ast, (m_0, m_1))$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
        \caption{Sequence of games}\label{fig:cca-elgamal-sequence}
    \end{figure}

    \emph{Game 0} is the standard IND-CCA game: $\texttt{Adv}_0(A) \coloneq \texttt{Adv}_\texttt{IND-CCA}(A)$

    \emph{Game 1} is identical to game 0, except that in the challenge encryption, the symmetric key $k \leftsample \mathcal{K}$ is uniformly random instead of pseudorandomly derived. Under the random oracle model, the two games are statistically indistinguishable from adversary $A$'s perspective unless $A$ queries $H$ on $w = g^{xy}$. Denote this event by $\texttt{QUERY}^\ast$, then by the difference lemma:

    \begin{equation*}
        \texttt{Adv}_0(A) - \texttt{Adv}_1(A) \leq P\left\lbrack \texttt{QUERY}^\ast \right\rbrack
    \end{equation*}

    \emph{Game 1} can be entirely simulated by an IND-CPA adversary $C$ against the symmetric cipher. $C$ can generate the ElGamal keypair on its own, simulate the hash oracle $H$, and service $A$'s decryption queries (using the generated keypair) before $A$ outputs the chosen plaintexts $m_0, m_1$. When $A$ outputs the chosen plaintexts $m_0, m_1$, $C$ outputs them as its own chosen plaintexts and receives the challenge ciphertext $c^\prime$. $C$ then samples random $y \leftsample \mathbb{Z}_q^\ast$, computes $v \leftarrow g^y$, and returns $c^\ast = (v, c^\prime)$ to $A$ as $A$'s challenge encryption. Finally, when $A$ outputs its guess $\hat{b}$, $C$ passes $\hat{b}$ as its own guess. It is easy to see that $C$ wins the IND-CPA game if and only if $A$ wins game 1:

    \begin{equation*}
        \texttt{Adv}_1(A) = \texttt{Adv}_\texttt{IND-CPA}(C)
    \end{equation*}

    We now bound the probability $P\left\lbrack \texttt{QUERY}^\ast \right\rbrack$ by constructing an interactive computational Diffie-Hellman problem adversary $B$ using $A$ as a subroutine. To do that, $B$ needs to service $A$'s hash queries and decryption queries. The simulated hash oracles and decryption oracles are listed in figure \ref{fig:sim-oracles}. Note that $\mathcal{L}^H$ is used to record the queries made to $H$, while $\mathcal{L}^\decrypt$ is used to record symmetric keys used for decryption queries.

    \begin{figure}[H]
        \begin{minipage}{0.5\textwidth}
            \begin{algorithm}[H]
                \caption{$\mathcal{O}^H_1(w)$}
                \begin{algorithmic}[1]
                    \If{$\exists (\tilde{w}, \tilde{k}) \in \mathcal{L}^H:\tilde{w} = w$}
                        \State \Return $\tilde{k}$
                    \ElsIf{$\exists (\tilde{v}, \tilde{k}) \in \mathcal{L}^\decrypt
                        :\mathcal{O}^\texttt{DDH}(g^x, \tilde{v}, w) = 1$}
                        \State $k \leftarrow \tilde{k}$
                    \Else
                        \State $k \leftsample \mathcal{K}$
                    \EndIf
                    \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(w, k)\}$
                    \State \Return $k$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \begin{minipage}{0.49\textwidth}
            \begin{algorithm}[H]
                \caption{$\mathcal{O}^\decrypt_1(v, c^\prime)$}
                \begin{algorithmic}[1]
                    \If{$\exists (\tilde{w}, \tilde{k}) \in \mathcal{L}^H
                        :\mathcal{O}^\texttt{DDH}(g^x, v, \tilde{w}) = 1$}
                        \State \Return $\decrypt_s(\tilde{k}, c^\prime)$
                    \ElsIf{$\exists (\tilde{v}, \tilde{k}) \in \mathcal{L}^\decrypt
                        :\tilde{v} = v$}
                        \State \Return $\decrypt_s(\tilde{k}, c^\prime)$
                    \Else
                        \State $k \leftsample \mathcal{K}$
                        \State $\mathcal{L}^\decrypt \leftarrow \mathcal{L}^\decrypt \cup \{(v, k)\}$
                        \State \Return $\decrypt_s(k, c^\prime)$
                    \EndIf
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \caption{Simulated hash oracle and decryption oracle}\label{fig:sim-oracles}
    \end{figure}

    Assuming that $\mathcal{O}^\texttt{DDH}$ is always correct, the IND-CCA adversary $A$ cannot distinguish the simulated oracles from the true oracles.

    When $A$ produces the chosen plaintexts $m_0, m_1$, $B$ needs to perform the challenge encryption: \begin{enumerate}
        \item $v \leftarrow g^y$, where $g^y$ is given to $B$ as part of ICDH input
        \item $k \leftsample \mathcal{K}$ as per game 1
        \item $b \leftsample \{0,1\}; c^\prime \leftarrow \encrypt_s(k, m_b)$
        \item Return $(v, c^\prime)$ as the challenge ciphertext
    \end{enumerate}

    Afterwards, $B$ continues simulating the oracles for $A$ until $A$ halts. Then, $B$ searches through $\mathcal{L}^H$. If $\texttt{QUERY}^\ast$ happens, then there exists $\tilde{w} \in \mathcal{L}^H$ such that $\mathcal{O}^\texttt{DDH}(g^x, g^y, \tilde{w}) = 1$, and $B$ can return $\tilde{w}$ and win the ICDH game. Therefore:

    \begin{equation*}
        P\left\lbrack \texttt{QUERY}^\ast \right\rbrack \leq \texttt{Adv}_\texttt{ICDH}(B)
    \end{equation*}

    Finally, putting the equations above give us the desired result.
\end{proof}

\section{Does this apply to Kyber-AE?}
\emph{Unfortuantely not}.

In the case of the hybrid ElGamal presented in figure \ref{fig:cca-elgamal-routines}, the decryption oracle is converted into a decisional Diffie-Hellman oracle, but it can be safely assumed that having a decisional Diffie-Hellman oracle does not provide non-negligible help. However, there is no comparable "loosening of security assumption" for Kyber-AE, where the decapsulation oracle can be converted into a plaintext-checking oracle against the underlying PKE, which then recovers the secret key.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}

### 2024-10-19 archives/faster-ind-cca-kem-encrypt-then-mac/manuscript/main.tex
"Faster IND-CCA KEM using encrypt-then-MAC" submitted to CHES on Oct 15

--- 2024-10-15 etm-transform/manuscript/main.tex
@@
 The Fujisaki-Okamoto transformation uses de-randomization and re-encryption to ensure ciphertext non-malleability, but because ML-KEM's underlying PKE encryption is significantly slower than the underlying PKE decryption, ML-KEM's decapsulation performance is dominated by re-encryption.
-In this paper, we propose an alternative generic IND-CCA secure KEM transformation that achieves chosen-ciphertext security by applying the encrypt-then-MAC mechanism to a OW-PCA secure PKE and an existentially unforgeable MAC.
+In this paper, we propose an alternative generic IND-CCA secure KEM transformation that achieves chosen-ciphertext security by applying the encrypt-then-MAC mechanism to an OW-PCA secure PKE and an existentially unforgeable MAC.
 Compared to the Fujisaki-Okamoto transformation, our encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag.
-We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it ML-KEM+.
-We then implement ML-KEM+ with a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
-At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), ML-KEM+ achieves a massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
-Furthermore, we implement key exchange protocols and measure realistic network round trip times (RTTs), where $\mlkemplus$ reduces RTTs by 23.9\%-39.5\%,  compared to ML-KEM.
+We instantiate our proposed KEM with the PKE sub-routines of ML-KEM and call it $\mlkemplus$.
+We then implement $\mlkemplus$ with a wide selection of MACs including Poly1305, GMAC, CMAC, and KMAC.
+At the cost of minimal increase in encapsulation CPU cycles (+1.8\%) and ciphertext size (+2.1\%), $\mlkemplus$ achieves a massive reduction of decapsulation CPU cycles (-72.2\%) compared to ML-KEM.
+Furthermore, we implement key exchange protocols and measure realistic network round trip times (RTT), where $\mlkemplus$ reduces RTT by 23.9\%-39.5\% compared to ML-KEM.
 \end{abstract}
@@
 It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA security.
-The encrypt-then-MAC (EtM) mechanism was standard in ISO 19772 \cite{international2009iso-encrypt-then-mac}.
+The encrypt-then-MAC (EtM) mechanism was standardized in ISO 19772 \cite{international2009iso-encrypt-then-mac}.
 While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits.
@@
 The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypting the PKE plaintext, then authenticating the PKE ciphertext using the previously derived MAC key.
-When the Gap Diffie-Hellman assumption holds and the MAC is existentially unforgeable, no efficient adversary can recover the decryption of an unknown ciphertext even with access to a decryption oracle because it cannot produce a valid tag for such unknown ciphertext.
+If the Gap Diffie-Hellman assumption holds and the MAC is existentially unforgeable, no efficient adversary can recover the decryption of an unknown ciphertext even with access to a decryption oracle because it cannot produce a valid tag for such unknown ciphertext.
 
@@
     \item {\bf Performance evaluation and comparisons.} We implemented $\mlkemplus$ in C by adapting ML-KEM's reference implementation \cite{kyberrefimpl} and compared its performance with ML-KEM in a variety of scenarios.
-Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-79.1\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-7.3\% and ciphertext size by 1.0\%-2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
+Compared to ML-KEM, $\mlkemplus$ achieves 72.2\%-74.7\% reduction of CPU cycle count for decapsulation while only increasing encapsulation's CPU cycle count by 1.8\%-13.8\% and ciphertext size by 2.0\% (see Table \ref{tbl:cpu-cycles-summary}).
 We also implemented the Kyber key exchange protocols \cite{DBLP:conf/eurosp/BosDKLLSSSS18} with $\mlkemplus$ and measured the round trip time (RTT) in realistic network settings.
-Compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE).
+Compared to ML-KEM, $\mlkemplus$ achieves 23.7\%-25.4\% reduction of round trip time in unauthenticated key exchange (KE), 28.4\%-32.3\% in unilaterally authenticated key exchange (UAKE), and 30.6\%-44.9\% reduction in mutually authenticated key exchange (AKE).
 See Table \ref{tbl:rtt-summary} for a summary of round trip times.
@@
         \hline
-        &  $\text{ML-KEM}^+$-512 \newline Poly1305
-        &  $\text{ML-KEM}^+$-768 \newline KMAC 192-bit tag
-        &  $\text{ML-KEM}^+$-1024 \newline KMAC 256-bit tag
+        &  $\mlkemplus$-512 \newline Poly1305
+        &  $\mlkemplus$-768 \newline KMAC 192-bit tag
+        &  $\mlkemplus$-1024 \newline KMAC 256-bit tag
         \\
@@
 Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al.
-later proved that the OAEP is IND-CCA secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
+later proved that OAEP is IND-CCA secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}.
 RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended of all RSA-based encryption schemes.
@@
 
-The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic IND-CCA secure transformation.
-While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme.
-Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build a KEM, and proved its security in the quantum random oracle model (QROM).
+The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99, DBLP:journals/joc/FujisakiO13} is another generic IND-CCA secure transformation.
+In the original publication, Fujisaki and Okamoto proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme.
+Later works \cite{DBLP:conf/ima/Dent03, DBLP:conf/tcc/HofheinzHK17, DBLP:conf/eurocrypt/DworkNR04, DBLP:conf/asiacrypt/HovelmannsHM22, bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build a KEM, and proved its security in the quantum random oracle model (QROM).
 
@@
 
-The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries.
+The Fujisaki-Okamoto transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries.
 It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.
-The $\kem^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM.
+The $\kem^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation (see Figure \ref{fig:u-notbot-m-routines}) is adopted in ML-KEM \cite{FIPS203}.
 
@@
 Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext.
-As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key.
+As demonstrated in \cite{ueno2022curse,ravi2019generic,DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key.
 \end{itemize}
@@
 Key generation $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair.
-The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key.
-The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key.
+The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext $m$ under the input public key $\pk$ and produces a ciphertext $c$.
+The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext $c$ under the input secret key $\sk$ and produces a plaintext $m$.
 Where the encryption routine is randomized, we denote the randomness by a coin $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space.
 The decryption routine is assumed to always be deterministic.
-Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.
 
@@
 \begin{equation*}
-    E\left[\max_{m \in \mathcal{M}} P\left[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)\right]\right] \leq \delta
+    E\left[\max_{m \in \mathcal{M}} P\left[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)\right]\right] \leq \delta.
 \end{equation*}
@@
 \textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}.
-There are two main types of games: in the one-wayness game, the adversary is given a random encryption, then asked to produce the correct decryption; in the indistinguishability game, the adversary is given the encryption of one of two adversary-chosen plaintexts, then asked to decide which of the plaintexts corresopnds with the given encryption.
+There are two main types of games: i) in the one-wayness (OW-ATK) game, the adversary is given a random encryption, then asked to produce the correct decryption; ii) in the indistinguishability (IND-ATK) game, the adversary is given the encryption of one of two adversary-chosen plaintexts, then asked to decide which of the plaintexts corresponds with the given encryption.
 Depending on the attack model, the adversary may have access to various oracles.
@@
 
-The advantage of an adversary in the \texttt{OW-ATK} game is the probability that it ouputs the correct decryption.
+The advantage of an adversary in the \texttt{OW-ATK} game is the probability that it outputs the correct decryption.
 The advantage of an adversary in the \texttt{IND-ATK} game is defined below.
@@
 \begin{equation*}
-    \texttt{Adv}_\texttt{IND-ATK}(A) = \left\vert P\left[ \hat{b} = b \right] - \frac{1}{2} \right\vert
+    \texttt{Adv}_\texttt{IND-ATK}(A) = \left\vert P\left[A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast) = b \right] - \frac{1}{2} \right\vert.
 \end{equation*}
@@
 \subsection{Key encapsulation mechanism (KEM)}
-\textbf{Syntax} A key encapsulation mechanism $\kem(\keygen, \encap, \decap)$ is a collection of three routines defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
+\textbf{Syntax.} A key encapsulation mechanism $\kem(\keygen, \encap, \decap)$ is a collection of three routines defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$.
 The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$.
@@
 
-\textbf{Security} The security of KEMs is similarly discussed in adversarial games (Figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game.
+\textbf{Security.} The security of a KEM is similarly discussed in adversarial games (Figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE indistinguishability game.
 In a KEM's indistinguishability game, an adversary is given the public key and a challenge ciphertext, then asked to distinguish a pseudorandom shared secret $K_0$ associated with the challenge ciphertext from a truly random bit string of equal length.
@@
 The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key.
-The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by the adversary's ability to correctly distinguish the two cases beyond blind guess:
+The advantage of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by the adversary's ability to correctly distinguish the two cases beyond a blind guess:
 
@@
         P[A^{\mathcal{O}_\decap}(a^\lambda, \pk, c^\ast, K_b) = b] - \frac{1}{2}
-    }
+    }.
 \end{equation*}
@@
 \subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
-A message authentication code $\mac(\keygen, \sign, \verify)$ is a collection of routines defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
-The signing routine $\sign(k, m)$ authenticates the message $m$ under the secret key $k$ by producing a tag $t$ (also called digest).
-The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
+\textbf{Syntax.} A message authentication code $\mac(\keygen, \sign, \verify)$ is a collection of routines defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$.
+The signing routine $\sign(k, m)$ authenticates the message $m$ under the secret key $k$ by producing a tag $t$ (also called digest) (we define the process that generates an authentication tag $t$ over message $m$ a \textit{signing routine} in this paper).
+The verification routine $\verify(k, m, t)$ takes the triplet of secret key $k$, message $m$, and tag $t$, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise.
 Many MAC constructions are deterministic.
@@
 
-The security of a MAC is defined in an adversarial game in which an adversary, with access to some signing oracle $\mathcal{O}_\sign(m)$, tries to forge a new valid message-tag pair that has never been queried before.
+\textbf{Security.} The security of a MAC is defined in an adversarial game in which an adversary, with access to a MAC oracle that can answer signing queries $\mac(k, m) \leftarrow \mathcal{O}_\mac(m)$, tries to forge a new valid message-tag pair that has never been queried before.
+The ability to access a MAC oracle is called \emph{chosen-message attack (CMA)}.
+The ability to produce a valid tag on some arbitrary message is called \emph{existential forgery}.
 The existential unforgeability under chosen message attack (EUF-CMA) game is shown below:
@@
     \centering
-    \begin{minipage}{0.53\textwidth}
+    \begin{minipage}[t]{0.53\textwidth}
         \begin{algorithm}[H]
@@
                 \State $k^\ast \leftsample \mathcal{K}$
-                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
+                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\mac}()$
                 \State \Return $
                     \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
-                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
+                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\mac
                     \rrbrack
@@
         \end{algorithm}
-    \end{minipage}
-    \caption{The existential forgery game}\label{fig:euf-cma-game}
+    \end{minipage}\hspace{0.3cm}
+    \begin{minipage}[t]{0.4\textwidth}
+        \begin{algorithm}[H]
+            \caption*{MAC oracle $\mathcal{O}_\mac(m)$}
+            \begin{algorithmic}[1]
+                \State \Return $\mac(k^\ast, m)$
+            \end{algorithmic}
+        \end{algorithm}
+    \end{minipage}
+    \caption{The existential forgery game and the MAC oracle}\label{fig:euf-cma-game}
 \end{figure}
@@
 Some MACs are one-time existentially unforgeable, meaning that each secret key can be used to authenticate only a single message.
-The corresponding security game is modified such that the signing oracle will only answer a single signing query.
+The corresponding security game is modified such that the MAC oracle will only answer a single signing query.
 
@@
 \section{The encrypt-then-MAC transformation}\label{sec:main-results}
-In this section we introduce the encrypt-then-MAC transformation that transforms an OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM.
+\noindent\textbf{Our technique.} We introduce our encrypt-then-MAC transformation that transforms a OW-PCA secure PKE and an one-time existentially unforgeable MAC into an IND-CCA secure KEM.
 Our scheme mainly differs from DHIES in its versatility and input requirement.
-Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in Section \ref{sec:elgamal-is-ow-pca} that the Gap Diffie-Hellman assumption implies OW-PCA security).
-In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 for further performance improvements (M.
-Abdalla et al.
-originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}).
+Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme.
+In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 for further performance improvements (Abdalla, Rogaway, and Bellare originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see Section \ref{sec:choosing-a-message-authenticator}).
 The encapsulation data flow is illustrated in Figure \ref{fig:etm-diagram}.
@@
 
+\subsection{The generic KEM construction}
 Let $\mathcal{B}^\ast$ denote the set of finite bit strings.
 Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$.
-Let $\mac: \mathcal{K}_\mac \times \mathcal{B}^\ast \rightarrow \mathcal{T}$ be a deterministic message authentication code that takes a key $k \in \mathcal{K}_\mac$, some message $m \in \mathcal{B}^\ast$, and outputs a digest $t \in \mathcal{T}$.
+Let $\mac: \mathcal{K}_\mac \times \mathcal{B}^\ast \rightarrow \mathcal{T}$ be a deterministic message authentication code that takes a key $k \in \mathcal{K}_\mac$, some message $m \in \mathcal{B}^\ast$, and outputs a tag $t \in \mathcal{T}$.
 Let $G: \mathcal{M} \rightarrow \mathcal{K}_\mac$ be a hash function that maps from $\pke$'s plaintext space to $\mac$'s key space.
@@
     \end{minipage}
-    \caption{The encrypt-then-MAC transformation builds a KEM $\kem_\etm$ using a $\pke(\keygen, \encrypt, \decrypt)$, a $\mac$, and two hash functions $G, H$}\label{fig:etm-routines} 
+    \caption{The encrypt-then-MAC transformation builds a KEM, denoted by $\kem_\etm$, using a $\pke(\keygen, \encrypt, \decrypt)$, a $\mac$, and two hash functions $G, H$}\label{fig:etm-routines} 
 \end{figure}
@@
 
-The chosen-ciphertext security of the encrypt-then-MAC scheme can be intuitively argued through an adversary's inability to learn additional information from the decapsulation oracle.
+\paragraph{Security analysis.} The chosen-ciphertext security of the encrypt-then-MAC scheme can be intuitively argued through an adversary's inability to learn additional information from the decapsulation oracle.
 For an adversary $A$ to produce a valid tag $t$ for some unauthenticated ciphertext $c^\prime$ under the symmetric key $k \leftarrow G(\decrypt(\sk^\prime, c^\prime))$ implies that $A$ must either know the symmetric key $k$ or produce a forgery.
-Under the random oracle model, $A$ also cannot know $k$ without knowing its preimage $\decrypt(\sk^\prime, c^\prime)$, so $A$ must either have produced $c^\prime$ honestly, or have broken the one-way security of $\pke$.
-This means that the decapsulation oracle will not give out information on decryptions that the adversary does not already know.
+Under the random oracle model, $A$ also cannot know $k$ without knowing its pre-image $\decrypt(\sk^\prime, c^\prime)$, so $A$ must either have produced $c^\prime$ honestly, or have broken the one-way security of $\pke$.
+This means that the decapsulation oracle will not give out information on decryption that the adversary does not already know.
 
@@
     \begin{equation*}
-        \texttt{Adv}_\texttt{IND-CCA}(A) \leq q \cdot \adv_\texttt{OT-MAC}(C) + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
+        \texttt{Adv}_\texttt{IND-CCA}(A) \leq q \cdot \adv_\texttt{OT-MAC}(C) + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B).
     \end{equation*}
@@
 
-\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
+\subsection{Proof of Theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}\label{sec:proof-of-theorem}
 We will prove Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game.
 A summary of the the sequence of games can be found in Figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}.
-From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(\pk, m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, t)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$.
+From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: \begin{enumerate}
+    \setlength\itemsep{0.1em}
+    \item Replace the true decapsulation oracle with a simulated decapsulation oracle
+    \item Replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random $k^\ast \leftsample \mathcal{K}_\mac$
+    \item Replace the pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$
+\end{enumerate}
+
 A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.
@@
 \begin{proof}
-    \emph{Game 0} is the standard IND-CCA game for KEMs.
+    \emph{Game 0} is the standard KEM IND-CCA game.
 The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully.
@@
 Under the random oracle model, such $k$ is a uniformly random sample of $\mathcal{K}_\mac$ that the adversary does not know, so for $A$ to produce a valid tag is to produce a forgery against the $\mac$ under an unknown and uniformly random key.
-Furthermore, the security game does not include a signing oracle, so this is a zero-time forgery.
+Furthermore, the security game does not include a MAC oracle, so this is a zero-time forgery.
 While zero-time forgery is not a standard security definition for a MAC, we can bound it by the advantage of a one-time forgery adversary $C$:
@@
         P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]
-        \leq \adv_\texttt{OT-MAC}(C)
+        \leq \adv_\texttt{OT-MAC}(C).
     \end{equation*}
@@
     \begin{equation*}
-        \adv_{G_0}(A) - \adv_{G_1}(A) \leq q\cdot  \adv_\texttt{OT-MAC}(C)
+        \adv_{G_0}(A) - \adv_{G_1}(A) \leq q\cdot  \adv_\texttt{OT-MAC}(C).
     \end{equation*}
@@
     \begin{equation*}
-        \adv_{G_1}(A) - \adv_{G_2}(A) \leq P\left[\texttt{QUERY G}\right]
+        \adv_{G_1}(A) - \adv_{G_2}(A) \leq P\left[\texttt{QUERY G}\right].
     \end{equation*}
@@
     \begin{equation*}
-        \adv_{G_2}(A) - \adv_{G_3}(A) \leq P\left[\texttt{QUERY H}\right]
+        \adv_{G_2}(A) - \adv_{G_3}(A) \leq P\left[\texttt{QUERY H}\right].
     \end{equation*}
@@
         \begin{aligned}
-            P\left[\texttt{QUERY G}\right] &\leq \adv_\texttt{OW-PCA}(B) \\
-            P\left[\texttt{QUERY H}\right] &\leq \adv_\texttt{OW-PCA}(B) \\
+            P\left[\texttt{QUERY G}\right] &\leq \adv_\texttt{OW-PCA}(B), \\
+            P\left[\texttt{QUERY H}\right] &\leq \adv_\texttt{OW-PCA}(B). \\
         \end{aligned}
@@
     \begin{equation*}
-        \text{Adv}_\text{GapDH}(B) = \text{Adv}_\text{OW-PCA}(A)
+        \text{Adv}_\text{GapDH}(B) = \text{Adv}_\text{OW-PCA}(A).
     \end{equation*}
@@
 
-We now present the proof for Lemma \ref{lemma:elgamal-is-ow-pca}
+We now present the proof for Lemma \ref{lemma:elgamal-is-ow-pca}. 
 \begin{proof}
-    We will prove by sequence of game. A summary can be found in Figure \ref{fig:elgamal-pca-games}
+    We will prove by a sequence of games. A summary can be found in Figure \ref{fig:elgamal-pca-games}
 
@@
 
-        \caption{Lemma \ref{lemma:elgamal-is-ow-pca} sequence of games}\label{fig:elgamal-pca-games}
+        \caption{ The sequence of games in proving Lemma \ref{lemma:elgamal-is-ow-pca}}\label{fig:elgamal-pca-games}
     \end{figure}
@@
     \begin{equation*}
-        \adv_0(A) = \adv_1(A) = \adv_2(A)
+        \adv_0(A) = \adv_1(A) = \adv_2(A). 
     \end{equation*}
@@
 
-\subsection{ML-KEM+}
+% NOTE: section/subsection normally cannot have math expression or the PDF bookmarks will not typeset properly
+%   Use texorpdfstring to set the bookmark name in the second argument
+\subsection{\texorpdfstring{$\mlkemplus$}{ML-KEM+}}
 
@@
 
-    \caption{$\texttt{ML-KEM}^+$ applies our encrypt-then-MAC transformation to K-PKE. $G$ is SHA3-512, $H$ is SHA3-256, $\texttt{KDF}$ is Shake256}\label{fig:ml-kem-plus-routines}
+    \caption{$\mlkemplus$ applies our encrypt-then-MAC transformation to the K-PKE sub-routines of ML-KEM. $G$ is SHA3-512, $H$ is SHA3-256, $\texttt{KDF}$ is Shake256}\label{fig:ml-kem-plus-routines}
 \end{figure}
@@
 Hashing both the public key and the PKE plaintext ensures that both parties have inputs into the shared secret.
-The second reason is to prevent multi-target attacks: if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
+The second reason is multi-user security: if the MAC key is derived from the PKE plaintext alone, then an adversary can pre-compute a large dictionary mapping MAC keys to the pre-image PKE plaintext.
 Upon receiving some ciphertext, an adversary can search through this key-plaintext dictionary and recover the decryption.
-Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.
+Hashing the public key into the MAC key prevents an adversary from computing one large lookup table and use it against multiple keypairs.
+    % \textcolor{blue}{ This is NOT true as I have commented in my previous email, since pk is known to the attacker.
+``Adding the public key into the derivation of the MAC key will dramatically increase the space and time requirement of such dictionary attack and render it infeasible in practice.''}
     \item \textbf{Hashing the MAC tag instead of the entire ciphertext.} Because we allow the PKE to be randomized, the shared secret must have inputs from both the plaintext and the ciphertext.
@@
     \setlength\itemsep{0.1em}
-    \item $\mlkemplus$ encapsulation contains two additional steps: computing the MAC tag and hashing the MAC tag into the shared secret.
-    \item $\mlkemplus$ ciphertext size increases by the size of the MAC tag
-    \item $\mlkemplus$ decapsulation replaces re-execution $\texttt{K-PKE}.\texttt{Enc}$ with computing the MAC tag
+    \item The $\mlkemplus$ encapsulation contains two additional steps: computing the MAC tag and hashing the MAC tag into the shared secret.
+    \item An $\mlkemplus$ ciphertext size increases by the size of the MAC tag.
+    \item The $\mlkemplus$ decapsulation replaces re-execution $\texttt{K-PKE}.\texttt{Enc}$ with computing the MAC tag. 
 \end{itemize}
@@
 All MACs are parameterized with a 256-bit key to ensure that MAC keys are at least as hard to guess as the underlying PKE plaintext.
-On the other hand, all MACs use 128-bit tag except for KMAC, which can have variable tag length.
-We suspect that 128-bit tag might be insufficient for the 192-bit and 256-bit security level, so for instantiation with the higher security levels, we only use KMAC with 192-bit and 256-bit tag length output.
-
-\textbf{Poly1305 and GMAC} are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic.
+On the other hand, all MACs use a 128-bit tag except for KMAC, which can have a variable tag length.
+We suspect that a 128-bit tag might be insufficient for the 192-bit and 256-bit security levels, so for instantiation with the higher security levels, we only use KMAC with 192-bit and 256-bit tag length outputs.
+
+\textbf{Poly1305 and GMAC} are both Carter-Wegman style MACs \cite{wegman1981new} that compute the tag using finite field arithmetic.
 Generically speaking, Carter-Wegman MAC operates by breaking the message into message blocks that can then be parsed into finite field elements.
-The tag is computed by evaluating a polynomial whose coefficients are the message blocks and whose interdeterminate is the secret key, which is also called a \textit{hash key}.
+The tag is computed by evaluating a polynomial whose coefficients are the message blocks and whose indeterminate is the secret key (also called a \textit{hash key}).
 Specifically, Poly1035 operates in the prime field $\mathbb{F}_q$ where $q = 2^{130} - 5$ whereas GMAC operates in the binary field $\mathbb{F}_{2^{128}}$.
@@
 To compute a CMAC tag, the message is first broke into 128-bit blocks with appropriate padding.
-Each block is first XOR'd with the previous block's output, then encrypted under AES using the symmetric key.
-The final output is XOR'd with a sub key derived from the symmetric key, before being encrypted for one last time.
+Each block is first XORed with the previous block's output, then encrypted under AES using the symmetric key.
+The final output is XORed with a sub key derived from the symmetric key, before being encrypted for one last time.
 
@@
 KMAC is the only MAC to support variable key and tag length, though we fixed the key length at 256 bits.
-On the other hand, we chose the appropriate tag length for the corresponding security level: when instantiated with ML-KEM-512, the tag length is 128 bits; with ML-KEM-768 the tag length is 192 bits; with ML-KEM-1024 the tag length is 256 bits.
+On the other hand, we chose the appropriate tag length for the corresponding security level: when instantiated with ML-KEM-512, the tag length is 128 bits; with ML-KEM-768, the tag length is 192 bits; with ML-KEM-1024, the tag length is 256 bits.
 
@@
 We measured the real-world performance of the individual routines of our $\mlkemplus$ construction, as well as the round trip time of performing key exchange over a network in a variety of authentication modes.
-Our implementation extended from the reference implementation by the PQCrystals team \cite{kyberrefimpl}.
+Our implementation extended from the Kyber implementation \cite{kyberrefimpl}.
 All C code is compiled with \texttt{GCC 11.4.1} and \texttt{OpenSSL 3.0.8}.
@@
 \subsection{MAC performance}
-To isolate the performance characteristics of each authenticator in our instantiation of $\mlkemplus$ we measured the CPU cycles needed for each authenticator to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of ML-KEM.
+To isolate the performance characteristics of each MAC, we measured the CPU cycles needed for each MAC to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of ML-KEM.
 The median CPU cycle counts in 10,000 rounds of testing are summarized in Table \ref{tbl:mac-performance}.
@@
     \centering
-    \footnotesize
-    % \small
+    % \footnotesize
+    \small
     \caption{CPU cycles needed to compute tag on various input sizes}\label{tbl:mac-performance}
@@
         \hline
-    \end{tabular}
+    \end{tabular}\hspace{.3cm}
     % TODO: 128-bit tag size is not suitable for higher security level.
@@
         \hline
-    \end{tabular}
+    \end{tabular}\hspace{.3cm}
     \begin{tabular}[t]{|c|c|}
@@
 Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the encrypt-then-MAC transformation over the $\kem^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC.
-A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in Table \ref{tbl:kem-performance}.
+A comparison between ML-KEM and variations of the $\mlkemplus$ can be found in Table \ref{tbl:kem-performance}.
 
@@
         % ML-KEM+-512
-        $\text{ML-KEM}^+$-512 w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
-        \hline
-        $\text{ML-KEM}^+$-512 w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
-        \hline
-        $\text{ML-KEM}^+$-512 w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
-        \hline
-        $\text{ML-KEM}^+$-512 w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
+        $\mlkemplus$-512 w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
+        \hline
+        $\mlkemplus$-512 w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
+        \hline
+        $\mlkemplus$-512 w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
+        \hline
+        $\mlkemplus$-512 w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
         \hline
@@
         \hline
-        % $\text{ML-KEM}^+$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
-        % \hline
-        % $\text{ML-KEM}^+$-768 w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
-        % \hline
-        % $\text{ML-KEM}^+$-768 w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
-        % \hline
-        $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
+        % $\mlkemplus$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
+        % \hline
+        % $\mlkemplus$-768 w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
+        % \hline
+        % $\mlkemplus$-768 w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
+        % \hline
+        $\mlkemplus$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
         \hline
@@
         \hline
-        % $\text{ML-KEM}^+$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
-        % \hline
-        % $\text{ML-KEM}^+$-1024 w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
-        % \hline
-        % $\text{ML-KEM}^+$-1024 w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
-        % \hline
-        $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
+        % $\mlkemplus$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
+        % \hline
+        % $\mlkemplus$-1024 w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
+        % \hline
+        $\mlkemplus$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
         \hline
@@
 All experiments are run on a pair of AWS c7a.medium instances both located in the \texttt{us-west-2} region.
-For each experiment, a total of 10,000 rounds of key exchange are performed, with the median and average round trip time (measured in microsecond) recorded.
+For each experiment, a total of 10,000 rounds of key exchange are performed, with the median and average round trip time (measured in microsecond $\mu s$) recorded.
 
@@
 In unilaterally authenticated key exchange (UAKE), the authenticating party proves its identity to the other party by demonstrating possession of a secret key that corresponds to a published long-term public key.
-In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by correctly decrypting the challenge encryption sent by the client.
-UAKE routines are summarized in Figure \ref{fig:uake-routines}.
+In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by correctly decrypting the challenge encryption sent by the client using its long-term secret key $\sk_S$ (see Figure \ref{fig:uake-routines}).
 $\mlkemplus$-512 with Poly1305 reduces RTT by 29.0\% compared to ML-KEM-512; $\mlkemplus$-768 with KMAC reduces RTT by 28.4\%; $\mlkemplus$-1024 with KMAC reduces RTT by 31.3\% (See Table \ref{tbl:uake-rtt}).
@@
 Mutually authenticated key exchange is largely identical to unilaterally authenticated key exchange, except for that client authentication is required.
-This means that client possesses server's long-term public key and its own long-term secret key, while the server possesses client's long-term public key and its own long-term secret key.
+This means that client possesses server's long-term public key $\pk_S$ and its own long-term secret key $\sk_C$, while the server possesses client's long-term public key $\pk_C$ and its own long-term secret key $\sk_S$.
 The session key is derived by applying \texttt{KDF} onto the concatenation of shared secrets produced under the ephemeral keypair, server's long-term keypair, and client's long-term keypair, in this order.
@@
 \section{Conclusions and future works}\label{sec:future-works}
-In this paper we propose a generic KEM transformation based on the encrypt-then-MAC pattern.
+In this paper, we propose a generic KEM transformation based on the encrypt-then-MAC pattern.
 We reduce the IND-CCA security of the encrypt-then-MAC KEM to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model.

### 2024-11-06 code-based-crypto/talk/main.tex
Maybe I can do a talk

% TODO: outline of the talk:
% - Overview: what is Classic McEliece, summary of Goppa code and the cryptosystem
% - Preliminaries: finite field, polynomial ring, Bernoulli's rule
% - Goppa decoder:
%   - Lagrange interpolation
%   - Best approximant theorem
%   - Interpolate with error as a Reed Solomon decoder
%   - Definition of code space, Goppa squaring theorem
%   - Goppa decoder via Reed-Solomon decoder
% - Canonical parity check matrix, systematic form, syndrome decoding
% - Implementation details
%   - Semi-systematic form
\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}

% Custom commands
\usepackage{crypto-primitives}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{
    A graduate introduction to classic McEliece
}
\author{
    Ganyu (Bruce) Xu
}
% Leave the date field empty to display the date of compilation
\date{November, 2024}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section{Introduction}
Classic McEliece is an IND-CCA2 key encapsulation mechanism (KEM) submitted to NIST's post-quantum cryptography (PQC) standardization project. Its security is based on the conjectured intractability of the \textbf{Syndrome Decoding Problem}.

This document provides an introduction to the mathematics behind classic McEliece with a particular emphasis on the details of binary Goppa code. We assume a graduate level of mathematical maturity, though most of the results can be reasoned about using first or second year undergraduate math.

Classic McEliece is an appealing candidate for PQC because its cryptanalysis has been remarkably stable since its earliest conception in 1978 \cite{mceliece1978public} (whereas lattice-based cryptography saw enormous advance in its cryptanalysis in the last decade alone, forcing many latticed-based schemes to revise their parameters). Classic McEliece also has the following advantages: \begin{itemize}
    \item Encryption is faster
    \item Ciphertext is smaller
    \item There is no decryption failure that can leak information about secret key
\end{itemize}

On the other hand, Classic McEliece has larger public key size and slower decryption routine.

\subsection{A summary of binary Goppa code}
Binary Goppa code is a linear error-correcting code. Each instance is parameteried by: \begin{itemize}
    \item some base field $K = \mathbb{F}_{2^m}$
    \item $n$ distinct field elements $\alpha_1, \alpha_2, \ldots, \alpha_n \in K$
    \item An irreducible polynomial $g \in K[x]$ such that $\deg(g) = t$
\end{itemize}

% TODO: need better summary

\subsection{An overview of Classic McEliece}
Each instance of Classic McEliece is parameterized by three integers: \begin{itemize}
    \item $m$ is the size of the base field $K = \mathbb{F}_{2^m}$
    \item $n$ is the size of code words: $\text{codewords} \subseteq K^n$
    \item $t$ is the error-correcting capacity of the underlying Goppa code
\end{itemize}

\paragraph{Key generation.} Begin by randomly generating the parameters of a Goppa code instance, which include: \begin{itemize}
    \item $n$ distinct field elements $\alpha_1, \alpha_2, \ldots, \alpha_n \in K$
    \item A degree-$t$, square-free polynomial $g \in K[x]$ such that $g(\alpha_i) \neq 0$ for $1 \leq i \leq n$
\end{itemize}

$\alpha_1, \alpha_2, \ldots, \alpha_n$ and $g$ can be used to compute the canonical parity-check matrix $H \in K^{t \times n}$

$$
H_{i,j} = \frac{\alpha_j^{i - 1}}{g(\alpha_j)} \text{ for } 1 \leq i \leq t, 1 \leq j \leq n
$$

$H$ is then transformed into reduced row-echelon form (or systematic form per DJB) $H^\prime = [I_t \mid T]$ for some $T \in K^{t \times (n - t)}$. $T$ is returned as the public key. $\Gamma = (\alpha_1, \alpha_2, \ldots, \alpha_n, g)$ is returned as the secret key.

\paragraph{Encryption.} The message space is the subset of $\mathbb{F}_2^n$ whose Hamming weight is exactly $t$:

$$\mathcal{M} = \{\mathbf{e} \in \mathbb{F}_2^n \mid \mathop{wt}(\mathbf{e}) = t\}$$

To encrypt, compute:

$$\mathbf{y} \leftarrow [I_t \mid T] \mathbf{e}$$

$\mathbf{y} \in K^t$ is returned as the ciphertext.

\paragraph{Decryption.} Given $\mathbf{y} \in K^t$ and $\Gamma = (\alpha_1, \alpha_2, \ldots, \alpha_n, g)$, let $\mathbf{r} \leftarrow (y_1, y_2, \ldots, y_t, 0, 0, \ldots, 0) \in \mathbb{F}_2^n$. then feed $(\Gamma, \mathbf{r})$ into some Goppa decoder, which will directly recover $\mathbf{e} \in \mathbb{F}_2^n$ such that $\mathop{wt}(\mathbf{e}) = t$. Return $\mathbf{e}$ as the decryption.

\section{Implementation details}
% TODO: semi-systematic form
% TODO: how to generate field ordering and irreducible polynomial

\bibliographystyle{alpha}
\bibliography{references}

\end{document}

### 2024-11-08 code-based-crypto/talk/main.tex
more references

@@
 
+\section{Preliminaries}
+% TODO: degree of polynomial
+% TODO: Hamming weight
+% TODO: Bernoulli's rule
+
+\section{Understanding binary Goppa code decoding}
+\subsection{Polynomial interpolation}
+
+\begin{theorem}[Lagrange Interpolation]\label{thm:lagrange-interpolation}
+    Let $K$ be a finite field, $\alpha_1, \alpha_2, \ldots, \alpha_n \in K$ be $n$ distinct field elements, and $r_1, r_2, \ldots, r_n \in K$ be $n$ (possibly non-distinct) elements.
+There exists a unique degree-$(n-1)$ polynomial $f$ such that $f(\alpha_i) = r_i$:
+
+    \begin{equation*}
+        f = \sum_{i=1}^n\left(
+            r_i \prod_{j \neq i}\frac{x - \alpha_j}{\alpha_i - \alpha_j}
+        \right)
+    \end{equation*}
+\end{theorem}
+
+It's easy to check that the interpolation is correct, and that the degree of $f$ is less than $n$.
+If there exists another $g \in K[x]$ with $\deg(g) < n$ that interpolates $(\alpha_i, r_i)_{i=1}^n$, then $f - g$ is a polynomial with $n$ distinct roots.
+However, $\deg(f-g)$ is less than $n$ since $\deg(f), \deg(g) < n$.
+Therefore, it must be that $f-g = 0$, which means that $f = g$, thus proving uniqueness.
+
+\subsection{Shamir's secret sharing}
+Lagrange interpolation can be used to build a secret sharing scheme \cite{shamir1979share}.
+
+% TODO: shamir secret sharing
+
+\subsection{Polynomial approximant}
+% TODO: need intuition?
+
+\begin{theorem}[Best approximant theorem]\label{thm:best-approximant-theorem}
+    Let $K$ be some finite field and $n, t$ be non-negative integers such that $2t < n$.
+Given polynomials $A, B \in K[x]$ such that $\deg(B) < \deg(A)$, then there exists unique pair of polynomials $a, b \in K[x]$ such that $\deg(a) \leq t$, $\deg(b) < \deg(a)$, $\gcd(a, b) = 1$, and $\deg(aB - bA) < n - t$.
+If $c, d \in K[x]$ is such that $\deg(c) \leq t$ and $\deg(cB - dA) < n - t$, then $(c, d) = \lambda (a, b)$ for some $\lambda \in K[x]$.
+\end{theorem}
+
+% TODO: proof
+
+\subsection{Interpolation with error}
+In this section we introduce an algorithm that can recover a degree $(n - 2t)$ polynomial given $n$ points with up to $t$ errors.
+
+Let $K$ be some finite field and $n, t$ be non-negative integers such that $2t < n$.
+Let $f \in K[x]$ be a polynomial with degree $\deg(f) < n - 2t$.
+Let $\alpha_1, \alpha_2, \ldots, \alpha_n$ be $n$ distinct elements.
+Denote $\mathbf{c} = (f(\alpha_1), f(\alpha_2), \ldots, f(\alpha_n)) \in K^n$.
+If $\mathbf{r} \in K^n$ is such that $\mathop{wt}(\mathbf{r} - \mathbf{c}) \leq t$, then the following procedure can recover $f$ using $\alpha_1, \alpha_2, \ldots, \alpha_n$ and $\mathbf{r}$: \begin{enumerate}
+    \item Let $A = \prod_{i=1}^n (x - \alpha_i)$
+    \item Let $B$ interpolate $(\alpha_i, r_i)$
+    \item Compute degree-$t$ approximant $(a, b)$ of $(A, B)$. The error can be corrected if and only if $a \mid A$
+    \item Compute $f = B - bA / a$. Furthermore, $(B - f)(\alpha_i) \neq 0$ if and only if $a(\alpha_i) = 0$.
+\end{enumerate}
+
+\begin{proof}
+    % TODO: proof
+\end{proof}
+
+\subsection{Goppa decoding}
+
+\subsection{Parity check matrix}
+
 \section{Implementation details}

### 2024-11-28 code-based-crypto/talk/main.tex
progress on the proof

@@
 \usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{xcolor}
 
@@
 \section{Introduction}
-Classic McEliece is an IND-CCA2 key encapsulation mechanism (KEM) submitted to NIST's post-quantum cryptography (PQC) standardization project.
+Classic McEliece is an IND-CCA secure key encapsulation mechanism (KEM) submitted to NIST's post-quantum cryptography (PQC) standardization project.
 Its security is based on the conjectured intractability of the \textbf{Syndrome Decoding Problem}.
@@
 
-\subsection{Shamir's secret sharing}
-Lagrange interpolation can be used to build a secret sharing scheme \cite{shamir1979share}.
+\paragraph{Shamir's secret sharing.} In the secret sharing scheme proposed by Shamir in 1979 \cite{shamir1979share}, the secret is a degree-$t-1$ polynomial, and each share is $(\alpha, g(\alpha))$ for some $\alpha\in K$.
+If there are $t$ distinct shares, then the secret polynomial can be recovered; if there are only $t-1$ shares, then there are as many unique polynomials as there are distinct possible points in the field, meaning that if the field $K$ has cryptographically large order, then any adversary with only $t-1$ shares will have negligible probability of recovering the secret.
+This scheme is information-theoretically secure.
 
-% TODO: shamir secret sharing
-
-\subsection{Polynomial approximant}
-% TODO: need intuition?
+\subsection{Low-degree approximant}
+{\color{red} Is low-degree approximant related to approximant in linear algebra? Probably yes, but we need a clear connection}
 
@@
 
-% TODO: proof
+\begin{proof}\label{proof:best-approximant-theorem}
+    Let's first bound the degree of the polynomial $aB - bA$: $\deg(aB) = \deg(a) + \deg(B) < n + t$, and similarly we have $\deg(bA) = \deg(b) + \deg(A) < n + t$.
+Therefore, $\deg(aB - bA) = \max(\deg(bA), \deg(aB)) < n + t$, which means that the maximal degree of $aB - bA$ is $n - t - 1$.
+We can write $aB - bA$ as follows:
 
-\subsection{Interpolation with error}
-In this section we introduce an algorithm that can recover a degree $(n - 2t)$ polynomial given $n$ points with up to $t$ errors.
+    \begin{equation*}
+        aB - bA = \sum_{i=0}^{n+t-1}k_{i} \cdot x^{i}
+    \end{equation*}
 
-Let $K$ be some finite field and $n, t$ be non-negative integers such that $2t < n$.
-Let $f \in K[x]$ be a polynomial with degree $\deg(f) < n - 2t$.
-Let $\alpha_1, \alpha_2, \ldots, \alpha_n$ be $n$ distinct elements.
-Denote $\mathbf{c} = (f(\alpha_1), f(\alpha_2), \ldots, f(\alpha_n)) \in K^n$.
-If $\mathbf{r} \in K^n$ is such that $\mathop{wt}(\mathbf{r} - \mathbf{c}) \leq t$, then the following procedure can recover $f$ using $\alpha_1, \alpha_2, \ldots, \alpha_n$ and $\mathbf{r}$: \begin{enumerate}
-    \item Let $A = \prod_{i=1}^n (x - \alpha_i)$
-    \item Let $B$ interpolate $(\alpha_i, r_i)$
-    \item Compute degree-$t$ approximant $(a, b)$ of $(A, B)$. The error can be corrected if and only if $a \mid A$
-    \item Compute $f = B - bA / a$. Furthermore, $(B - f)(\alpha_i) \neq 0$ if and only if $a(\alpha_i) = 0$.
+    Now consider a function $\sigma$ that maps $(a, b)$ to the higher-order $2t$ coefficients of $aB - bA$.
+Since the map $(a, b) \mapsto aB - bA$ is trivially linear, and the map from $aB - bA$ to its coefficients is trivially linear, it is easy to see that $\sigma: (a, b) \mapsto (k_{n+t-1}, k_{n+t-2}, \ldots, k_{n-t})$ is also a linear map.
+
+    Because $\deg(a) \leq t$ and $\deg(b) < t$, $a$ can have up to $t+1$ non-zero coefficients, and $b$ can have up to $t$ non-zero coefficients, so $(a, b)$ is an $2t+1$-dimensional linear space, while $\mathop{\text{Img}}(\sigma)$ has $2t$ dimensions.
+By the rank-nullity theorem, we know that $\sigma(a, b) = \mathbf{0}$ has non-zero solution.
+With this non-zero solution, the first $2t$ coefficients of $aB - bA$ are all zeros, so the degree of $aB - bA$ must be less than $n - t$.
+
+    % TODO: pathological case where one of a, b, A, B is zero?
+
+    At this point we have arrived at some $(a, b)$ such that $\deg(b) < \deg(a) \leq t$ and $\deg(aB - bA) < n - t$.
+If $\gcd(a, b) \neq 1$, then we can divide $a, b$ by the non-trivial GCD, which will make them co-prime while $\deg(aB - bA)$ must still be less than $n - t$.
+This proves the existence of co-prime degree-$t$ approximants, now we will prove the uniqueness of such a solution.
+
+    Suppose that $c, d \in K[x]$ are polynomials such that $\deg(d) < \deg(c) \leq t$ and $\deg(cB-dA) < n-t$ (but they are not necessarily co-prime), then consider the following equation:
+
+    \begin{equation*}
+        \det\left(
+            \begin{bmatrix}
+                a & c \\
+                aB-bA & cB-dA
+            \end{bmatrix}
+        \right) = a(cB-dA) - c(aB-bA) = (cb - ad)A
+    \end{equation*}
+
+    The degree of the middle term $\deg(a(cB-dA)-c(aB-bA))$ is less than $n$ because $\deg(cB-dA), \deg(aB-bA)$ are both less than $n - t$ and $\deg(a), \deg(c)$ are both at most $t$.
+On the other hand, since we assumed $\deg(A) = n$, $a(cB-dA)-c(aB-bA) = (cb-ad)A$ must imply that $cb-ad=0$.
+In addition, since we assumed $\gcd(a, b) = 1$, it must be that $a \mid c$, from here we naturally get $d = c/a \cdot b$.
+In other words, $(c, d) = c/a \cdot (a, b)$ is a non-zero multiple of $(a, b)$.
+Any other degree-$t$ approximants cannot have lower degree than $(a, b)$, which makes $(a, b)$ the \textit{best approximants}.
+\end{proof}
+
+Note that the proof above also gave an algorithm for finding $(a, b)$: given $A, B$, the coefficients of $aB-bA$ can be explicitly written as a linear system of equation with respect to the coefficients ot $(a, b)$, so finding non-zero $(a, b)$ is equivalent to solving a linear system $M \in K^{(2t+1) \times 2t}$.
+For the remainder of the paper we assume the existence of a well-defined function $\texttt{approximate}: (A, B, t) \mapsto (a, b)$
+
+\subsection{Interpolation with error}\label{sec:interpolate-with-error}
+Let $n, t$ be non-negative integers such that $2t < n$.
+Using low-degree polynomial approximants, we can recover a degree-$(n - 2t - 1)$ polynomial using evaluations at $n$ distinct points with up to $t$ errors.
+Specifically: \begin{enumerate}
+    \item $f \in K[x]$ is a polynomial with degree $\deg(f) < n - 2t$
+    \item $\alpha_1, \alpha_2, \ldots, \alpha_n \in K$ are distinct field elements
+    \item $r_1, r_2, \ldots, r_n \in K$ are such that $r_i \neq f(\alpha_i)$ at no more than $t$ positions
+\end{enumerate}
+
+The algorithm is as follows \begin{enumerate}\label{alg:interpolate-with-error}
+    \item Let $A = \prod_{i=1}^n(x-\alpha_i)$
+    \item Interpolate $B \in K[x]$ such that $B(\alpha_i) = r_i$ for $1 \leq i \leq n$
+    \item Compute degree-$t$ approximant $(a, b)$ of $(A, B)$ using theorem \ref{thm:best-approximant-theorem}. The error can be recovered if an only if $a \mid A$
+    \item Compute $\hat{f} = B - bA/a$, which will be the output
 \end{enumerate}
@@
 \begin{proof}
-    % TODO: proof
+    let $S = \{i \mid r_i = f(\alpha_i) \}$ denote the set of positions where $r_i$ is correct. $A$ can be factored into two polynomials:
+
+    \begin{equation*}
+        A = \left(\prod_{i\in S}(x - \alpha_i)\right)\left(\prod_{i\not\in S}(x - \alpha_i)\right)
+    \end{equation*}
+
+    We denote the first factor by $E = \prod_{i\in S}(x-\alpha_i)$ and the second factor by $c = \prod_{i\not\in S}(x-\alpha_i)$, then $A = Ec$.
+
+    Now observe the polynomial $(B - f)$: since for all $i \in S$ we have $B(\alpha_i) = r_i = f(\alpha_i)$, we know that $x - \alpha_i$ divides $B - f$, and since $\alpha_i$'s are all distinct, $\prod_{i \in S}(x - \alpha_i) = E$ must divide $B - f$.
+In other words:
+
+    \begin{equation*}
+        Ed = B-f
+    \end{equation*}
+
+    for some non-zero polynomial $d \in K[x]$. Observe that:
+
+    \begin{equation*}
+    Ad = (Ec)d = (Ed)c = (B - f)c = Bc - fc
+    \end{equation*}
+
+    which re-arranges to:
+
+    \begin{equation*}
+    fc = cB-dA
+    \end{equation*}
+
+    Let's take a look at the degrees of each term: \begin{itemize}
+        \item $\deg(f) < n - 2t$ by assumption
+        \item $\deg(c) \leq t$ since we assumed there to be no more than $t$ errors
+        \item $\deg(fc) = \deg(f) + \deg(c) < (n - 2t) + t = n - t$
+        \item $\deg(A) = n$, $\deg(B) = n-1 < \deg(A)$
+    \end{itemize}
+
+    In other words, $(c, d)$ are degree-$t$ approximants of $(A, B)$.
+By Theorem \ref{thm:best-approximant-theorem}, we know that $(c, d) = \lambda(a, b)$ for some non-zero $\lambda \in K[x]$ and $(a, b)$ are the best degree-$t$ approximants of $(A, B)$.
+This means that $a \mid c$, and since $c \mid A$, it naturally follows that $a \mid A$.
+
+    Substituting $(c, d)$ with $\lambda(a, b)$ gives us the desired results:
+
+    \begin{equation*}
+        f = B - bA/a
+    \end{equation*}
 \end{proof}
 
+\begin{theorem}
+    $r_i \neq f(\alpha_i)$ if and only if $a(\alpha_i) = 0$
+\end{theorem}
+
+\begin{proof}
+    TODO: prove this theorem
+\end{proof}
+
+This is in fact, a Reed-Solomon decoder.
+In Reed-Solomon code, the set of symbols is the set of degree-$n-2t$ polynomials, encoding involves evaluating the polynomial at $n$ distinct points.
+If there are no more than $t$ errors, then applying algorithm \ref{alg:interpolate-with-error} will allow the decoder to recover the polynomial.
+
 \subsection{Goppa decoding}
+In this section we restrict $K$ to be finite field of characteristic 2 ($K = \mathbb{F}_{2^m}$).
+
+Each binary Goppa code is parameterized by $n$ distinct field elements $\alpha_1, \alpha_2, \ldots, \alpha_n \in K$ (called \textbf{field ordering}) and a degree-$t$ irreducible polynomial $g\in K[c]$ such that $g(\alpha_i)\neq 0$ for all $1 \leq i \leq n$.
+The set of code words is defined as follows:
+
+\begin{equation*}
+    \mathcal{C} = \{
+        \mathbf{c} \in \mathbb{F}_2^n 
+        \mid \sum_{i=1}^n \frac{c_iA}{x - \alpha_i} \equiv 0 \mod g
+    \}
+\end{equation*}
+
+where $A = \prod_{i=1}^n(x - \alpha_i)$.
 

### 2025-01-10 code-based-crypto/understanding-binary-goppa-decoding-in-mceliece-round-4/main.tex
bm summary and double sized parity check matrix

\documentclass[runningheads]{llncs}

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\urlstyle{rm}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{listings}
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}} % require "Require" with "Input"
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{document}

\title{Understanding binary Goppa decoding in Classic McEliece}

% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%\titlerunning{Abbreviated paper title}

\author{
    Ganyu Xu\inst{1}
    % First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
    % Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
    % Third Author\inst{3}\orcidID{2222--3333-4444-5555}
}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\authorrunning{G. Xu et al.}

\institute{
    University of Waterloo, Waterloo, Ontario, Canada
    \email{g66xu@uwaterloo.ca}
}

\maketitle

\begin{abstract}
    Supplementary materials to ``Understanding binary Goppa decoding'' that reflect the actual decoder used in the reference implementations of Classic McEliece in NIST PQC Round 4 submission

    \keywords{
        Error-correcting code
        \and Classic McEliece
        \and Post-quantum cryptography
    }
\end{abstract}

\section{Introduction}

\section{Preliminaries}

\section{Berlekamp-Massey Algorithm}\label{sec:berlekamp-massey-algorithm}
The Berlekamp-Massey algorithm was first proposed by Elwin Berlekamp \cite{DBLP:journals/tit/Berlekamp68} for decoding BCH code and later extended by James Massey \cite{DBLP:journals/tit/Massey69} to synthesize shortest LFSR. In this section, we will review the Berlekamp-Massey algorithm from the perspective of finding the shortest LFSR that generates a given finite sequence, including some key results that proved the optimality of the generated LFSR.

Let $\mathbb{F}$ be some finite field. A linear-feedback shift register (LFSR) of length $L$ is parameterized by the connection coefficients $c_1, c_2, \ldots, c_L \in \mathbb{F}$ and the seed values $s_0, s_1, \ldots, s_{L-1}\in\mathbb{F}$. An LFSR outputs an infinite sequence $s_0, s_1, \ldots$ where the first $L$ digits are the seed values, then for $j \geq L$:

\begin{equation*}
    s_j + \sum_{i=1}^{L}c_is_{j-i} = 0
\end{equation*}

By convention, an empty LFSR with length $0$ outputs an infinite sequence of all $0$'s. We further define the \textbf{connection polynomial} of an LFSR to be $C(x) = 1 + c_1x + c_2x^2 + \ldots + c_Lx^L$. Note that there is no requirement for the connection coefficients to be non-zero, so the degree of the polynomial can be less than the actual length of the corresponding LFSR.

Let $\mathbf{s} = (s_0, s_1, \ldots) \in \mathbb{F}^\infty$ be some infinite sequence. Let $L_N(\mathbf{s})$ denote the length of the shortest LFSR that generates the first $N$ digits of $\mathbf{s}$ (where there is no ambiguity we will simply use $L_N$ for short). First observe that $L_N \leq N$ because one can always find an LFSR that generates the first $N$ digits simply by using these digits as the seed values. Second, observe that $L_N$ is \textbf{monotonically non-decreasing} with respect to $N$: $L_{N+1}\geq L_N$. This makes sense because any LFSR that generates the first $N+1$ digits is also an LFSR that generates the first $N$ digits, so the shortest LFSR that generates the first $N$ digits cannot be longer than the shortest LFSR that generates the first $N + 1$ digits.

A summary of the shortest LFSR synthesis algorithm, as presented in \cite{DBLP:journals/tit/Massey69}, can be found in Figure \ref{fig:shortest-lfsr-synthesis}

\begin{figure}[h]
    \centering
    \begin{algorithm}[H]
        \caption{Shortest LFSR synthesis}\label{alg:shortest-lfsr-synthesis}
        \begin{algorithmic}[1]
            \Require $(s_0, s_1, \ldots, s_{N-1}) \in \mathbb{F}$
            \Ensure The connection polynomial $C(x)$ and length $l$ of a shortest LFSR
            \State $C(x) \leftarrow 1, l \leftarrow 0, n \leftarrow 0, B(x) \leftarrow 1, d_m \leftarrow 1, y \leftarrow 1$
            \While{$n < N$}
                \State $d_n \leftarrow s_n + \sum_{i=1}^{l}c_is_{n-i}$
                \If{$d_n = 0$}
                    \State $y \leftarrow y + 1$
                \ElsIf{$2 l > n$}
                    \State $C(x) \leftarrow C(x) - d_nd_m^{-1}x^yB(x)$
                    \State $y \leftarrow y + 1$
                \Else
                    \State $T(x) \leftarrow C(x)$
                    \State $C(x) \leftarrow C(x) - d_nd_m^{-1}x^yB(x)$
                    \State $l \leftarrow n + 1 - l$
                    \State $B(x) \leftarrow T(x)$
                    \State $d_m \leftarrow d_n$
                    \State $y \leftarrow 1$
                \EndIf
                \State $n \leftarrow n + 1$
            \EndWhile
            \State \Return $C(x), l$
        \end{algorithmic}
    \end{algorithm}
    \caption{The Berlekamp-Massey algorithm iteratively constructs a shortest LFSR that generates some input sequence}\label{fig:shortest-lfsr-synthesis}
\end{figure}

The optimality of Algorithm \ref{alg:shortest-lfsr-synthesis} is proved in Theorem \ref{thm:shortest-lfsr-synthesis} below:

\begin{theorem}[shortest LFSR synthesis]\label{thm:shortest-lfsr-synthesis}
    Let $\mathbf{s}$ be an infinite sequence and $N \geq 0$. If there exists an LFSR with length $L_N(\mathbf{s})$ that generates $s_0, s_1, \ldots, s_{N-1}$ but not $s_{N}$, then $L_{N+1}(\mathbf{s}) = \max(L_N(\mathbf{s}), N + 1 - L_N(\mathbf{s}))$
\end{theorem}

To prove Theorem \ref{thm:shortest-lfsr-synthesis}, we will first prove a Lemma

\begin{lemma}[LFSR length change]\label{lemma:lfsr-length-change}
    If there exists an LFSR with length $L$ that generates $s_0, s_1, \ldots, s_{N-1}$ but not $s_N$, then the length $L^\prime$ of any LFSR that generates $s_0, s_1, \ldots, s_N$ must be such that $L^\prime \geq N + 1 - L$
\end{lemma}

\begin{proof}
    Let $(c_1, c_2, \ldots, c_L)$ be an length-$L$ LFSR that generates $s_0, \ldots, s_{N-1}$ but not $s_N$. Let $(c_1^\prime, c_2^\prime, \ldots, c_{L^\prime}^\prime)$ be some length $L^\prime$ LFSR that generates $s_0, \ldots, s_N$, then by the definition of LFSR we have the following three equations:

    \begin{align}
        s_j &= -\sum_{i=1}^Lc_is_{j-i} & (L \leq j < N) \label{eq:lfsr-l-1}\\
        s_N &\neq -\sum_{i=1}^Lc_is_{N-i} \label{eq:lfsr-l-neq}\\
        s_j &= -\sum_{i=1}^{L^\prime}c_i^\prime s_{j-i} & (L^\prime \leq j \leq N) \label{eq:lfsr-l-prime}
    \end{align}

    If $L^\prime \leq N - L$, then $N-L \leq j \leq N$ falls in the range of equation \ref{eq:lfsr-l-prime}. The RHS of equation \ref{eq:lfsr-l-neq} can be re-written:

    \begin{equation*}\begin{aligned}
        -\sum_{i=1}^Lc_is_{N-i} &= -\sum_{i=1}^Lc_i (-\sum_{k=1}^{L^\prime} c_k^\prime s_{N-i-k}) \\
        &= -\sum_{k=1}^{L^\prime} c_k^\prime (-\sum_{i=1}^Lc_i s_{N-i-k}) \\
        &= -\sum_{k+1}^{L^\prime}c_k^{L^\prime} s_{N-k} \\
        &= s_N
    \end{aligned}\end{equation*}

    This contradicts equation \ref{eq:lfsr-l-neq}, thus it must be that $L^\prime \geq N + 1 - L$.
\end{proof}

Let $n < N$. Let $C^{(n)}(x)$ denote the connection polynomial of the shortest LFSR that generates the first $n$ digits. If $C^{(n)}$ also generates $s_n$, then $C^{(n+1)} = C^{(n)}$ because of monotonicity. If $C^{(n)}$ does not generate $s_n$, then by Lemma \ref{lemma:lfsr-length-change} we have $L_{n+1} \geq n + 1 - L_n$. Furthermore because of monotonicity, $L_{n+1} \geq L_n$, thus we can establish a (recursive) lower bound : $L_{n+1} \geq \max(L_n, n + 1 - L_n)$.

It turns out that this lower bound can be achieved: suppose $C^{(n)}$ generates the first $n$ digits but not the first $n+1$ digits. Let $d_n = s_n + \sum_{i=1}^{L_n}c_i^{(n)}s_{n-i}$, by the earlier assumption we know $d_n \neq 0$. Let $m$ be the number of digits such that $L_m < L_{m+1} = L_n$ ($m$ is the sequence length before the last LFSR length change) and let $d_m = s_m + \sum_{i=1}^{L_m}c_i^{(m)}s_{m - i}$. Because $L_m < L_{m+1}$, $d_m$ must also be non-zero. We claim that:

\begin{equation*}
    C^{(n+1)} = C^{(n)} - d_nd_m^{-1}x^{n-m}C^{(m)}
\end{equation*}

We need to verify two points: one is that $C^{(n+1)}$ indeed generates the first $n+1$ digits, second is that $C^{(n+1)}$ achieves the lower bound. First expand $C^{(n+1)}$:

\begin{equation*}\begin{aligned}
    C^{(n+1)} &= 1 + \sum_{i=1}^{L_n}c^{(n)}_ix^i - d_nd_m^{-1}x^{n-m}\left(1 + \sum_{i=1}^{L_m}c_i^{(m)}x^i\right) \\
    &= 1 + \sum_{i=1}^{L_n}c^{(n)}_ix^i - d_nd_m^{-1}\left(x^{n-m} + \sum_{i=1}^{L_m}c_i^{(m)}x^{n - m + i}\right) \\
\end{aligned}\end{equation*}

Because of monotonicity we know $L_{n+1} \geq L_n, L_m$, so for $j \geq L_{n+1}$:

\begin{equation*}\begin{aligned}
   \text{j-th output} &= -\sum_{i=1}^{L_n}c_i^{(n)}s_{j-i} + d_nd_m^{-1}\left(
        s_{j - (n - m)} + \sum_{i=1}^{L_m}c_i^{(m)}s_{j - (n - m + i)}
    \right) \\
    &= -\sum_{i=1}^{L_n}c_i^{(n)}s_{j-i} + d_nd_m^{-1}\left(
        s_{m - (j - n)} + \sum_{i=1}^{L_m}c_i^{(m)}s_{m - (j - n) + i)}
    \right)
\end{aligned}\end{equation*}

Where $j < n$, $-\sum_{i=1}^{L_n}c_i^{(n)}s_{j-i}$ evaluates to $s_j$, and $m-(j-n) < m$, so $s_{m - (j - n)} + \sum_{i=1}^{L_m}c_i^{(m)}s_{m - (j - n) + i} = 0$. The entire expression evaluates to $s_j$.

On the other hand, if $j = n$, then $-\sum_{i=1}^{L_n}c_i^{(n)}s_{j-i} = s_j - d_n$, and $m-(j-n) = m$, so $s_{m - (j - n)} + \sum_{i=1}^{L_m}c_i^{(m)}s_{m - (j - n) + i} = d_m$. Thus:

$$
\text{j-th output} = (s_n - d_n) + d_nd_m^{-1}d_m = s_n
$$

Hence $C^{(n+1)}$ generates the first $n+1$ digits. For the second point, observe that:

\begin{equation*}\begin{aligned}
    \deg(C^{(n+1)}) &= \deg(C^{(n)} - d_nd_mx^{n-m}C^{(m)}) \\
    &= \max(\deg(C^{(n)}), \deg(x^{n-m}C^{(m)})) \\
    &= \max(L_n, n - m + L_m)
\end{aligned}\end{equation*}

We inductively assume that the lower bound holds for all $j < n$, then $L_n = L_{m+1} = \max(L_m, m+1-L_m)$. Since we assumed $L_m < L_{m+1}$, it must be that $L_n = L_{m+1} = m+1-L_m$. Therefore:

\begin{equation*}\begin{aligned}
    \deg(C^{(n+1)}) &= \max(L_n, n - m + L_m) \\
    &= \max(L_n, n - m + (m + 1 - L_n)) \\
    &= \max(L_n, n + 1 - L_n)
\end{aligned}\end{equation*}

This proves that the lower bound can be achieved, which proves Theorem \ref{thm:shortest-lfsr-synthesis}.

There are two more points to discuss regarding Algorithm \ref{alg:shortest-lfsr-synthesis}. First, in the proof above we assumed that when $C^{(n)}$ fails to generate $s_n$, some $m: L_m < L_{m+1} < L_n$ exists, which is not the case for when $s_n$ is the first non-zero digit of the sequence. In this case, the algorithm simply set the initial condition to be such that $C^{(n+1)}$ is correct, and after this iteration, $m$ and its related parameters $C^{(m)}(x)$ and $d_m$ are guaranteed to be updated, so subsequent iterations will work according to the proof above.

Second, there is a distinction betweeen $2l > n$ and $2l \leq n$ because it decides whether there is going to be a length change, which will decide how $C^{(m)}, d_m, y$ will be updated. When $2l > n$, $l \geq n + 1 - l$, so $l \leftarrow \max(l, n + 1 - l) = l$ does not incur a length change; on the other hand, when $2l \leq n$, $l < n + 1 - l$, so there will be a length change, meaning that the current set of connection polynomial $C^{(n)}$ and difference $d_n$ becomes ``the most recent length change''.

\section{Alternant code decoding}\label{sec:alternant-code-decoding}

\section{Double-sized syndrome}\label{sec:double-sized-syndrome}
Let $\mathbb{F}_{2^m}$ be some binary extension field, let $\alpha_1, \alpha_2, \ldots, \alpha_n \in \mathbb{F}_{2^m}$ be $n$ distinct elements, let $g(x)\in\mathbb{F}_{2^m}[x]$ be a monic irreducible polynomial of degree $t$. The binary Goppa code parameterized by $\Gamma = (\alpha_1, \ldots, \alpha_n, g)$ is defined by:

\begin{equation*}
    \mathcal{C} = \{\mathbf{c}\in\mathbb{F}_2^n \mid \sum_{i=1}^n \frac{c_iA}{x - \alpha_i}\equiv 0 \mod g\}
\end{equation*}

Where $A = \prod_{i=1}^{n}(x - \alpha_i)$ is the vanishing polynomial of $\alpha_1, \ldots, \alpha_n$. Note that the original formulation of Goppa code only required $g\in\mathbb{F}_{2^m}[x]$ to be square-free and co-prime with $A$ (i.e. $g(\alpha_i)\neq 0$ for all $1 \leq i \leq n$), but in Classic McEliece $g$ is sampled to be monic irreducible, so we also require $g$ to be monic ireducible.

Section \ref{sec:alternant-code-decoding} described how Algorithm \ref{alg:shortest-lfsr-synthesis} can be used to produce the error-locator polynomial $a(x)$ given a parity check matrix $H \in \mathbb{F}^{t \times n}$ and a received codeword $\mathbf{r}\in\mathbb{F}^n$, but this algorithm can only correct up to $\lceil\frac{t}{2}\rceil$ errors instead of the full-capacity $t$. In this section, we will review the techniques used in \cite{DBLP:journals/jce/HeyseG13} to allow the application of Berlekamp-Massey algorithm for recovering all $t$ errors.

\begin{theorem}[Double-sized parity check matrix]\label{them:double-sized-parity-check-matrix}
    Given binary Goppa code $\Gamma = (\alpha_1, \ldots, \alpha_n, g)$ over binary extension field $\mathbb{F}_{2^m}$, $\mathbf{c}\in\mathbf{F}_2^n$ is a codeword if and only if for all $0\leq s\leq 2t-1$:

    \begin{equation*}
        \sum_{i=1}^n \frac{\alpha_i^s c_i}{g^2(\alpha_i)} = 0
    \end{equation*}
\end{theorem}

Theorem \ref{them:double-sized-parity-check-matrix} provides an alternative parity check matrix $H^{(2)}$ that has $2t$ rows instead of $t$ rows as in the canonical parity check matrix. For the remainder of this section we will prove Theorem \ref{them:double-sized-parity-check-matrix} using a series of Theorems and Lemmas.

\begin{proof}
    We begin by stating a well-known results:

    \begin{theorem}[Goppa squaring]\label{thm:goppa-squaring}
        Given binary extension field $\mathbb{F}_{2^m}$, $n$ distinct elements $\alpha_1, \ldots, \alpha_n$, and degree-$t$ monic irreducible polynomial $g(x)\in\mathbb{F}_{2^m}[x]$, $\sum_{i=1}^n\frac{c_iA}{x-\alpha_i}\equiv 0 \mod g$ if and only if $\sum_{i=1}^n\frac{c_iA}{x-\alpha_i}\equiv 0 \mod g^2$
    \end{theorem}
    See \cite{DBLP:journals/iacr/Bernstein22} for proof.

    Let $B = \sum_{i=1}^n\frac{c_iA}{g^2(\alpha_i)(x - \alpha_i)}$ and $C = \sum_{i=1}^n\frac{c_iA}{x - \alpha_i}$, we claim:

    \begin{equation}\label{eq:double-sized-lemma-1}
        C \equiv 0 \mod g^2 \Longleftrightarrow \deg(B) < n - 2t
    \end{equation}

    Observe that for each $1\leq j \leq n$, $B(\alpha_j)g^2(\alpha_j) = C(\alpha_j)$, which means that $\alpha_1, \alpha_2, \ldots, \alpha_n$ are roots of the polynomial $Bg^2 - C$. Consequently $Bg^2 - C$ is divisible by $A$.

    If $C \equiv 0 \mod g^2$, then $Bg^2 - C = g^2(B - C/{g^2})$ where $B - C/g^2\in\mathbb{F}_{2^m}[x]$. Because $g$ and $A$ are co-prime, it must be that $A$ divides $B - C/g^2$, but the degree of $B$ and the degree of $C/g^2$ are both less than $n$, so it must be that $B - C/g^2 = 0$. It is easy to show that $\deg(C)\leq n - 1$, so $\deg(B) = \deg(C/g^2) \leq n - 1 - 2t < n - 2t$.

    If $\deg(B) < n - 2t$, then $\deg(Bg^2) < n$, so $\deg(Bg^2 - C) < n$. But since $A$ divides $Bg^2 - C$, it must be that $Bg^2 - C = 0$, which implies that $C = Bg^2$ is divisible by $g^2$.

    Now denote $Q = \sum_{i=1}^{n}\frac{\alpha_i^{2t}c_iA}{g^2(\alpha_i)(x - \alpha_i)}$. It is easy to show the follwing:

    \begin{equation}\label{eq:double-sized-lemma-2}
        \deg(B) < n - 2t \Longleftrightarrow \deg(x^{2t}B - Q) < n
    \end{equation}

    On the RHS of the relation above:

    \begin{equation*}\begin{aligned}
        x^{2t}B - Q &= \sum_{i=1}^{n}\frac{(x^{2t} - \alpha_i^{2t})c_iA}{g^2(\alpha_i)(x - \alpha_i)} \\
        &= \sum_{i=1}^{n}\left(
            \frac{c_iA}{g^2(\alpha_i)}\sum_{s=0}^{2t-1}x^s\alpha_i^{2t-1-s}
        \right) \\
        &= A \sum_{s=0}^{2t-1}\left(\sum_{i=1}^{n}\left(\frac{c_i\alpha_i^{2t-1-s}}{g^2(\alpha_i)}\right)x^s\right)
    \end{aligned}\end{equation*}

    In other words, $x^{2t}B - Q$ is divisible by $A$, so $\deg(x^{2t}B - Q) < n$ if and only if $x^{2t}B - Q = 0$, which happens if and only if $\sum_{i=1}^{n}\frac{c_i\alpha_i^{2t-1-s}}{g^2(\alpha_i)} = 0$ for all $0\leq s \leq 2t-1$:

    \begin{equation}\label{eq:double-sized-lemma-3}
        \deg(x^{2t}B - Q) < n \Longleftrightarrow \sum_{i=1}^{n}\frac{c_i\alpha_i^s}{g^2(\alpha_i)} = 0 \;\;(0 \leq s \leq 2t-1)
    \end{equation}

    Combining Theorem \ref{thm:goppa-squaring} and equations \ref{eq:double-sized-lemma-1}, \ref{eq:double-sized-lemma-2}, \ref{eq:double-sized-lemma-3} proves Theorem \ref{them:double-sized-parity-check-matrix}.
\end{proof}

% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references.bib}
\end{document}

### 2025-02-26 latex-templates/springer-lecture-notes/main.tex
Simple slides with Waterloo Engineering color scheme

\documentclass[runningheads]{llncs}

\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{color}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\urlstyle{rm}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{listings}
% \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algorithm}
\usepackage{algpseudocode}
\begin{document}
\title{Document title}

% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%\titlerunning{Abbreviated paper title}

\author{
    Ganyu Xu\inst{1}
    % First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
    % Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
    % Third Author\inst{3}\orcidID{2222--3333-4444-5555}
}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\authorrunning{G. Xu et al.}

\institute{
    University of Waterloo, Waterloo, Ontario, Canada
    \email{g66xu@uwaterloo.ca}
}

\maketitle

\begin{abstract}
    The abstract should briefly summarize the contents of the paper in
    150--250 words.

    \keywords{
        First keyword  
        \and Second keyword
        \and Another keyword.
    }
\end{abstract}

\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

Some reference to algorithm in Figure \ref{fig:some-algorithm}.

\begin{figure}[H]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
    \begin{algorithm}[H]
        \caption{Some algorithm}
        \begin{algorithmic}[1]
            \State $x \leftarrow 1$
        \end{algorithmic}
    \end{algorithm}
    \end{minipage}
    \caption{Figure containing algorithm}\label{fig:some-algorithm}
\end{figure}

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the tables.}\label{tab1}
    \begin{tabular}{|l|l|l|}
    \hline
    Heading level &  Example & Font size and style\\
    \hline
    Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
    1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
    2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
    3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
    4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
    \hline
    \end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}

For citations of references, we prefer the use of square brackets and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{wegman1981new}. Multiple citations are grouped
\cite{wegman1981new,bernstein2018towards,bellare1998relations}.

\begin{credits}
\subsubsection{\ackname} A bold run-in heading in small font size at the end of the paper is
used for general acknowledgments, for example: This study was funded
by X (grant number Y).

\subsubsection{\discintname}
It is now necessary to declare any competing interests or to specifically
state that the authors have no competing interests. Please place the
statement with a bold run-in heading in small font size beneath the
(optional) acknowledgments\footnote{If EquinOCS, our proceedings submission
system, is used, then the disclaimer can be provided directly in the system.},
for example: The authors have no competing interests to declare that are
relevant to the content of this article. Or: Author A has received research
grants from Company W. Author B has received a speaker honorarium from
Company X and owns stock in Company Y. Author C is a member of committee Z.
\end{credits}

% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{references.bib}
\end{document}
